### 2019-11-19 report/report/report.tex
spin experiment

\documentclass[letterpaper, 12pt]{article}

\input{report/preamble}

\begin{document}

\begin{center}
  {\huge Schuster Lab - Autumn 2019 Report} \\[0.5em]
  {\large Thomas Propson $\vert$ \href{mailto:tcpropson@uchicago.edu}
    {tcpropson@uchicago.edu} \\[0.5em] \today}
\end{center}

\section{Simulation Parameters}
\subsection{Spin System}
Dave told me these system parameters on 08/25/2019. The system hamiltonian is
\begin{align*}
  H &= \omega_{q}\frac{\sigma_{z}}{2}\\
    &+ \epsilon(t) \frac{\sigma_{x}}{2}\\
\end{align*}
$\omega_{q} = 1 * 10^{-2}$GHz, $\epsilon(t) \in \mathbb{R}, |\epsilon(t)| \le 1 * 10^{-1}$GHz.
The sampling rate of the AWG is 1.2 Gs/s with 14 bits. The bandwidth of the pulses should be
less than $5 * 10^{-1}$GHz. We choose the duration of the pulse to be about $150$ns.

\section{Hamiltonian Parameter Robustness}
We want to design pulses that are robust to variaitions
in the hamiltonian parameter $\omega_{q}$. We seek to find pulses that minimize the
norm of the derivative of $\omega_{q}$ w.r.t the derivative of $\epsilon(t)$ w.r.t to the gate fidelity.
We implement the universal gate set $\{R_{x}(\pi), R_{x}(\frac{\pi}{2}), R_{y}(\frac{\pi}{2}), T\}$
\cite{heeres2017implementing}.

\subsection{Spin Experiment 0}
This experiment achieves the $R_{x}(\pi)$ gate
  \[
    \begin{pmatrix}
      1\\
      0\\
    \end{pmatrix}
    \rightarrow
    \begin{pmatrix}
      0\\
      -i\\
    \end{pmatrix}
   \]
   \[
    \begin{pmatrix}
      0\\
      1\\
    \end{pmatrix}
    \rightarrow
    \begin{pmatrix}
      -i\\
      0\\
    \end{pmatrix}
  \]

  \subsection{Spin Experiment 1}
  This experiment achieves the $R_{x}(\frac{\pi}{2})$ gate
  \[
    \begin{pmatrix}
      1\\
      0\\
    \end{pmatrix}
    \rightarrow
    \begin{pmatrix}
      \frac{\sqrt{2}}{2}\\
      -i\frac{\sqrt{2}}{2}\\
    \end{pmatrix}
   \]
   \[
    \begin{pmatrix}
      0\\
      1\\
    \end{pmatrix}
    \rightarrow
    \begin{pmatrix}
      -i\frac{\sqrt{2}}{2}\\
      \frac{\sqrt{2}}{2}\\
    \end{pmatrix}
  \]

  \subsection{Spin Experiment 2}
  This experiment achieves the $R_{y}(\frac{\pi}{2})$ gate
  \[
    \begin{pmatrix}
      1\\
      0\\
    \end{pmatrix}
    \rightarrow
    \begin{pmatrix}
      \frac{\sqrt{2}}{2}\\
      \frac{\sqrt{2}}{2}\\
    \end{pmatrix}
   \]
   \[
    \begin{pmatrix}
      0\\
      1\\
    \end{pmatrix}
    \rightarrow
    \begin{pmatrix}
      -\frac{\sqrt{2}}{2}\\
      \frac{\sqrt{2}}{2}\\
    \end{pmatrix}
  \]

  \subsection{Spin Experiment 3}
  This experiment achieves the $T$ gate
  \[
    \begin{pmatrix}
      1\\
      0\\
    \end{pmatrix}
    \rightarrow
    \begin{pmatrix}
      1\\
      0\\
    \end{pmatrix}
   \]
   \[
    \begin{pmatrix}
      0\\
      1\\
    \end{pmatrix}
    \rightarrow
    \begin{pmatrix}
      0\\
      \frac{\sqrt{2}}{2} + i\frac{\sqrt{2}}{2}\\
    \end{pmatrix}
  \]

\bibliography{report/refs}
\bibliographystyle{plainnat}

\end{document}

### 2019-11-19 experiments/spin/piby2.py

@@ -1,0 +1,85 @@
+"""
+piby2.py - This module searches for pi/2 pulses
+for the spin system.
+"""
+
+import os
+
+import autograd.numpy as anp
+from qoc import grape_schroedinger_discrete
+from qoc.standard import (conjugate_transpose,
+                          generate_save_file_path,
+                          SIGMA_X, SIGMA_Z,
+                          TargetStateInfidelity,
+                          ControlArea,
+                          ControlBandwidth,
+                          LBFGSB,
+                          Adam,)
+
+CORE_COUNT = 8
+os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
+os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
+
+# Define experimental constants.
+CONTROL_SAMPLING_RATE = 1.2 # GS/s
+MAX_AMP_CONTROL_0 = 1e-1 # GHz
+MAX_AMP_BANDWIDTH_CONTROL_0 = 5e-1 # GHz
+OMEGA = 1e-2 # GHz
+
+
+# Define the system.
+SYSTEM_HAMILTONIAN_0 = OMEGA * SIGMA_Z / 2
+CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
+hamiltonian = lambda controls, time: (SYSTEM_HAMILTONIAN_0
+                                      + controls[0] * CONTROL_HAMILTONIAN_0)
+MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
+MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
+
+# Define the optimization.
+ITERATION_COUNT = int(1e4)
+COMPLEX_CONTROLS = False
+CONTROL_COUNT = 1
+EVOLUTION_TIME = 150 # nanoseconds
+CONTROL_STEP_COUNT = int(EVOLUTION_TIME * CONTROL_SAMPLING_RATE)
+OPTIMIZER = Adam()
+
+# Define the problem.
+INITIAL_STATE_0 = anp.array(((1,), (0,)))
+INITIAL_STATE_1 = anp.array(((0,), (1,)))
+INITIAL_STATES = anp.stack((INITIAL_STATE_0, INITIAL_STATE_1,))
+# Rx(pi/2)
[…]

### 2019-11-19 experiments/spin/spin_exp0.py

@@ -1,0 +1,130 @@
+"""
+spin_exp0.py - Spin experiment 0.
+"""
+
+import os
+
+import autograd.numpy as anp
+from filelock import FileLock, Timeout
+import h5py
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+from qoc import grape_schroedinger_discrete, evolve_schroedinger_discrete
+from qoc.standard import (TargetStateInfidelity,
+                          conjugate_transpose,
+                          get_annihilation_operator,
+                          get_creation_operator,
+                          SIGMA_Z, SIGMA_X,
+                          generate_save_file_path,
+                          krons, plot_controls,
+                          plot_state_population,
+                          Adam, LBFGSB,)
+
+CORE_COUNT = 8
+os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
+os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
+
+# Define experimental constants.
+CONTROL_SAMPLING_RATE = 1.2 # GS/s
+MAX_AMP_CONTROL_0 = 1e-1 # GHz
+MAX_AMP_BANDWIDTH_CONTROL_0 = 5e-1 # GHz
+OMEGA_Q = 1e-2 # GHz
+
+# Define the system.
+SYSTEM_HAMILTONIAN_0 = SIGMA_Z / 2
+CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
+HAMILTONIAN_ARGS = anp.array([OMEGA_Q])
+hamiltonian = lambda controls, time, args: (args[0] * SYSTEM_HAMILTONIAN_0
+                                            + controls[0] * CONTROL_HAMILTONIAN_0)
+MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
+MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
+
+# Define the optimization.
+CONTROLS_PATH = "/home/tcpropson/repos/rbqoc/out/spin/spin_exp0/00011_spin_exp0.h5"
+CONTROLS_PATH_LOCK = "{}.lock".format(CONTROLS_PATH)
+try:
+    with FileLock(CONTROLS_PATH_LOCK):
+        with h5py.File(CONTROLS_PATH) as controls_file:
+            index = anp.argmin(controls_file["error"][()])
+            initial_controls = controls_file["controls"][index]
[…]

### 2019-11-19 experiments/spin/spin_exp0_hpsweep.py

@@ -1,0 +1,53 @@
+"""
+spin_exp0_hpsweep.py - Do a sweep of the hamiltonian parameters
+for spin_exp0.
+"""
+
+import os
+
+import numpy as np
+from rbqoc.util import (hamiltonian_args_sweep,
+                        plot_hamiltonian_args_sweep,)
+from spin_exp0 import (EVOL_ARGS, EXPERIMENT_META, EXPERIMENT_NAME, HAMILTONIAN_ARGS)
+
+# Sweep trials
+SWEEP_TRIAL_COUNT = 100
+SWEEP_MULTIPLIERS = np.linspace(0, 3, SWEEP_TRIAL_COUNT)
+HAMILTONIAN_ARGS_SWEEP = np.stack([HAMILTONIAN_ARGS] * SWEEP_TRIAL_COUNT)
+for i, multiplier in enumerate(SWEEP_MULTIPLIERS):
+    HAMILTONIAN_ARGS_SWEEP[i] *= SWEEP_MULTIPLIERS[i]
+
+# I/O
+WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
+SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
+PULSE_FILE_0 = "00000_spin_exp0.h5"
+PULSE_FILE_1 = "00001_spin_exp0.h5"
+PULSE_FILE_2 = "00013_spin_exp0.h5"
+PULSE_FILE_3 = "00010_spin_exp0.h5"
+PULSE_FILES = [PULSE_FILE_0, PULSE_FILE_1, PULSE_FILE_2, PULSE_FILE_3]
+PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
+PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
+PULSE_PATH_2 = os.path.join(SAVE_PATH, PULSE_FILE_2)
+PULSE_PATH_3 = os.path.join(SAVE_PATH, PULSE_FILE_3)
+PULSE_PATHS = [PULSE_PATH_0, PULSE_PATH_1, PULSE_PATH_2, PULSE_PATH_3]
+SWEEP_SUFFIX = "freq_sweep"
+PLOT_SAVE_FILE = "{}_{}.png".format(EXPERIMENT_NAME, SWEEP_SUFFIX)
[…]

### 2019-11-19 experiments/spin/spin_exp1.py

@@ -1,0 +1,103 @@
+"""
+spin_exp1.py - Spin experiment 1.
+"""
+
+import os
+
+
+import autograd.numpy as anp
+from filelock import FileLock, Timeout
+import h5py
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+from qoc import grape_schroedinger_discrete, evolve_schroedinger_discrete
+from qoc.standard import (TargetStateInfidelity,
+                          conjugate_transpose,
+                          get_annihilation_operator,
+                          get_creation_operator,
+                          SIGMA_Z, SIGMA_X,
+                          generate_save_file_path,
+                          krons, plot_controls,
+                          plot_state_population,
+                          Adam, LBFGSB,)
+
+CORE_COUNT = 8
+os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
+os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
+
+# Define experimental constants.
+CONTROL_SAMPLING_RATE = 1.2 # GS/s
+MAX_AMP_CONTROL_0 = 1e-1 # GHz
+MAX_AMP_BANDWIDTH_CONTROL_0 = 5e-1 # GHz
+OMEGA_Q = 1e-2 # GHz
+
+# Define the system.
+SYSTEM_HAMILTONIAN_0 = SIGMA_Z / 2
+CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
+HAMILTONIAN_ARGS = anp.array([OMEGA_Q])
+hamiltonian = lambda controls, time, args: (args[0] * SYSTEM_HAMILTONIAN_0
+                                            + controls[0] * CONTROL_HAMILTONIAN_0)
+MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
+MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
+
+# Define the optimization.
+ITERATION_COUNT = int(1e4)
+COMPLEX_CONTROLS = False
+CONTROL_COUNT = 1
+EVOLUTION_TIME = 200 # nanoseconds
+CONTROL_EVAL_COUNT = SYSTEM_EVAL_COUNT = EVOLUTION_TIME + 1
+OPTIMIZER = Adam(learning_rate=1e-4)
+
+# Define the problem.
+INITIAL_STATE_0 = anp.array(((1,),
[…]

### 2019-11-19 experiments/spin/spin_exp1_hpsweep.py

@@ -1,0 +1,53 @@
+"""
+spin_exp1_hpsweep.py - Do a sweep of the hamiltonian parameters
+for spin_exp1.
+"""
+
+import os
+
+import numpy as np
+from rbqoc.util import (hamiltonian_args_sweep,
+                        plot_hamiltonian_args_sweep,)
+from spin_exp1 import (EVOL_ARGS, EXPERIMENT_META, EXPERIMENT_NAME, HAMILTONIAN_ARGS)
+
+# Sweep trials
+SWEEP_TRIAL_COUNT = 100
+SWEEP_MULTIPLIERS = np.linspace(0, 3, SWEEP_TRIAL_COUNT)
+HAMILTONIAN_ARGS_SWEEP = np.stack([HAMILTONIAN_ARGS] * SWEEP_TRIAL_COUNT)
+for i, multiplier in enumerate(SWEEP_MULTIPLIERS):
+    HAMILTONIAN_ARGS_SWEEP[i] *= SWEEP_MULTIPLIERS[i]
+
+# I/O
+WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
+SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
+PULSE_FILE_0 = "00000_spin_exp1.h5"
+PULSE_FILE_1 = "00001_spin_exp1.h5"
+PULSE_FILE_2 = "00003_spin_exp1.h5"
+PULSE_FILE_3 = "00004_spin_exp1.h5"
+PULSE_FILES = [PULSE_FILE_0, PULSE_FILE_1, PULSE_FILE_2, PULSE_FILE_3]
+PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
+PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
+PULSE_PATH_2 = os.path.join(SAVE_PATH, PULSE_FILE_2)
+PULSE_PATH_3 = os.path.join(SAVE_PATH, PULSE_FILE_3)
+PULSE_PATHS = [PULSE_PATH_0, PULSE_PATH_1, PULSE_PATH_2, PULSE_PATH_3]
+SWEEP_SUFFIX = "freq_sweep"
+PLOT_SAVE_FILE = "{}_{}.png".format(EXPERIMENT_NAME, SWEEP_SUFFIX)
[…]

### 2019-11-19 experiments/spin/spin_exp2.py

@@ -1,0 +1,103 @@
+"""
+spin_exp2.py - Spin experiment 2.
+"""
+
+import os
+
+
+import autograd.numpy as anp
+from filelock import FileLock, Timeout
+import h5py
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+from qoc import grape_schroedinger_discrete, evolve_schroedinger_discrete
+from qoc.standard import (TargetStateInfidelity,
+                          conjugate_transpose,
+                          get_annihilation_operator,
+                          get_creation_operator,
+                          SIGMA_Z, SIGMA_X,
+                          generate_save_file_path,
+                          krons, plot_controls,
+                          plot_state_population,
+                          Adam, LBFGSB,)
+
+CORE_COUNT = 8
+os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
+os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
+
+# Define experimental constants.
+CONTROL_SAMPLING_RATE = 1.2 # GS/s
+MAX_AMP_CONTROL_0 = 1e-1 # GHz
+MAX_AMP_BANDWIDTH_CONTROL_0 = 5e-1 # GHz
+OMEGA_Q = 1e-2 # GHz
+
+# Define the system.
+SYSTEM_HAMILTONIAN_0 = SIGMA_Z / 2
+CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
+HAMILTONIAN_ARGS = anp.array([OMEGA_Q])
+hamiltonian = lambda controls, time, args: (args[0] * SYSTEM_HAMILTONIAN_0
+                                            + controls[0] * CONTROL_HAMILTONIAN_0)
+MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
+MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
+
+# Define the optimization.
+ITERATION_COUNT = int(1e4)
+COMPLEX_CONTROLS = False
+CONTROL_COUNT = 1
+EVOLUTION_TIME = 200 # nanoseconds
+CONTROL_EVAL_COUNT = SYSTEM_EVAL_COUNT = EVOLUTION_TIME + 1
+OPTIMIZER = Adam()
+
+# Define the problem.
+INITIAL_STATE_0 = anp.array(((1,),
[…]

### 2019-11-19 experiments/spin/spin_exp2_hpsweep.py

@@ -1,0 +1,52 @@
+"""
+spin_exp2_hpsweep.py - Do a sweep of the hamiltonian parameters
+for spin_exp2.
+"""
+
+import os
+
+import numpy as np
+from rbqoc.util import (hamiltonian_args_sweep,
+                        plot_hamiltonian_args_sweep,)
+from spin_exp2 import (EVOL_ARGS, EXPERIMENT_META, EXPERIMENT_NAME, HAMILTONIAN_ARGS)
+
+# Sweep trials
+SWEEP_TRIAL_COUNT = 100
+SWEEP_MULTIPLIERS = np.linspace(0, 3, SWEEP_TRIAL_COUNT)
+HAMILTONIAN_ARGS_SWEEP = np.stack([HAMILTONIAN_ARGS] * SWEEP_TRIAL_COUNT)
+for i, multiplier in enumerate(SWEEP_MULTIPLIERS):
+    HAMILTONIAN_ARGS_SWEEP[i] *= SWEEP_MULTIPLIERS[i]
+
+# I/O
+WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
+SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
+PULSE_FILE_0 = "00004_spin_exp2.h5"
+PULSE_FILE_1 = "00005_spin_exp2.h5"
+PULSE_FILE_2 = "00008_spin_exp2.h5"
+PULSE_FILE_3 = "00009_spin_exp2.h5"
+PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
+PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
+PULSE_PATH_2 = os.path.join(SAVE_PATH, PULSE_FILE_2)
+PULSE_PATH_3 = os.path.join(SAVE_PATH, PULSE_FILE_3)
+PULSE_PATHS = [PULSE_PATH_0, PULSE_PATH_1, PULSE_PATH_2, PULSE_PATH_3]
+SWEEP_SUFFIX = "freq_sweep"
+PLOT_SAVE_FILE = "{}_{}.png".format(EXPERIMENT_NAME, SWEEP_SUFFIX)
+PLOT_SAVE_PATH = os.path.join(SAVE_PATH, PLOT_SAVE_FILE)
+
+def main():
+    sweep_path_list = list()
[…]

### 2019-11-19 experiments/spin/spin_exp3.py

@@ -1,0 +1,103 @@
+"""
+spin_exp3.py - Spin experiment 3.
+"""
+
+import os
+
+
+import autograd.numpy as anp
+from filelock import FileLock, Timeout
+import h5py
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+from qoc import grape_schroedinger_discrete, evolve_schroedinger_discrete
+from qoc.standard import (TargetStateInfidelity,
+                          conjugate_transpose,
+                          get_annihilation_operator,
+                          get_creation_operator,
+                          SIGMA_Z, SIGMA_X,
+                          generate_save_file_path,
+                          krons, plot_controls,
+                          plot_state_population,
+                          Adam, LBFGSB,)
+
+CORE_COUNT = 8
+os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
+os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
+
+# Define experimental constants.
+CONTROL_SAMPLING_RATE = 1.2 # GS/s
+MAX_AMP_CONTROL_0 = 1e-1 # GHz
+MAX_AMP_BANDWIDTH_CONTROL_0 = 5e-1 # GHz
+OMEGA_Q = 1e-2 # GHz
+
+# Define the system.
+SYSTEM_HAMILTONIAN_0 = SIGMA_Z / 2
+CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
+HAMILTONIAN_ARGS = anp.array([OMEGA_Q])
+hamiltonian = lambda controls, time, args: (args[0] * SYSTEM_HAMILTONIAN_0
+                                            + controls[0] * CONTROL_HAMILTONIAN_0)
+MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
+MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
+
+# Define the optimization.
+ITERATION_COUNT = int(1e4)
+COMPLEX_CONTROLS = False
+CONTROL_COUNT = 1
+EVOLUTION_TIME = 150 # nanoseconds
+CONTROL_EVAL_COUNT = SYSTEM_EVAL_COUNT = EVOLUTION_TIME + 1
+OPTIMIZER = Adam()
+
+# Define the problem.
+INITIAL_STATE_0 = anp.array(((1,),
[…]

### 2019-11-19 experiments/spin/spin_exp3_hpsweep.py

@@ -1,0 +1,49 @@
+"""
+spin_exp3_hpsweep.py - Do a sweep of the hamiltonian parameters
+for spin_exp3.
+"""
+
+import os
+
+import numpy as np
+from rbqoc.util import (hamiltonian_args_sweep,
+                        plot_hamiltonian_args_sweep,)
+from spin_exp3 import (EVOL_ARGS, EXPERIMENT_META, EXPERIMENT_NAME, HAMILTONIAN_ARGS)
+
+# Sweep trials
+SWEEP_TRIAL_COUNT = 100
+SWEEP_MULTIPLIERS = np.linspace(0, 3, SWEEP_TRIAL_COUNT)
+HAMILTONIAN_ARGS_SWEEP = np.stack([HAMILTONIAN_ARGS] * SWEEP_TRIAL_COUNT)
+for i, multiplier in enumerate(SWEEP_MULTIPLIERS):
+    HAMILTONIAN_ARGS_SWEEP[i] *= SWEEP_MULTIPLIERS[i]
+
+# I/O
+WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
+SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
+PULSE_FILE_0 = "00000_spin_exp3.h5"
+PULSE_FILE_1 = "00001_spin_exp3.h5"
+PULSE_FILES = [PULSE_FILE_0, PULSE_FILE_1]
+PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
+PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
+PULSE_PATHS = [PULSE_PATH_0, PULSE_PATH_1]
+SWEEP_SUFFIX = "freq_sweep"
+PLOT_SAVE_FILE = "{}_{}.png".format(EXPERIMENT_NAME, SWEEP_SUFFIX)
+PLOT_SAVE_PATH = os.path.join(SAVE_PATH, PLOT_SAVE_FILE)
+
+def main():
+    sweep_path_list = list()
+    # Sweep the frequency parameter for each pulse.
+    for i, pulse_path in enumerate(PULSE_PATHS):
+        save_path = hamiltonian_args_sweep(EVOL_ARGS, HAMILTONIAN_ARGS_SWEEP,
+                                           pulse_path, SWEEP_SUFFIX)
+        sweep_path_list.append(save_path)
+
+    # Plot the sweeps together.
[…]

### 2019-11-19 rbqoc/__init__.py

@@ -1,0 +1,3 @@
+"""
+rbqoc - This is the main directory for the package.
+"""

### 2019-11-19 rbqoc/util/__init__.py

@@ -1,0 +1,12 @@
+"""
+util - This directory houses utilities.
+"""
+
+from .hrutil import (hamiltonian_args_sweep,
+                     plot_hamiltonian_args_sweep,
+)
+
+__all__ = [
+    "hamiltonian_args_sweep",
+    "plot_hamiltonian_args_sweep",
+]

### 2019-11-19 rbqoc/util/hrutil.py

@@ -1,0 +1,164 @@
+"""
+hrutil.py - This module exposes utility functions for hamiltonian parameter robustness
+"""
+
+import os
+
+from filelock import FileLock, Timeout
+import h5py
+import matplotlib.pyplot as plt
+import numpy as np
+from qoc import evolve_schroedinger_discrete
+
+COLOR_PALETTE = (
+    "blue", "red", "green", "pink", "purple", "orange",
+    "teal", "grey", "black", "cyan", "magenta", "brown",
+    "azure", "beige", "coral", "crimson",
+)
+
+COLOR_PALETTE_LEN = len(COLOR_PALETTE)
+
+def hamiltonian_args_sweep(evol_args, hamiltonian_args_set, pulse_path, suffix):
+    """
+    Given a set of hamiltonian_args, run evolve_schroedinger_discrete for a control pulse
+    and save the error associated with each evolution to an h5 file.
+
+    Arguments:
+    evol_args :: dict(any) - The arguments to pass to evolve_schroedinger_discrete
+        in addition to the hamiltonian_args and the control pulse
+    hamiltonian_args_set :: ndarray(trial_count x hamiltonian_arg_count)
+        - A list of the hamiltonian arguments that should be swept over.
+    pulse_path :: list(str) - A list of paths to the output file
+        of the pulse that should be evolved.
+    suffix :: str - The suffix to append to the pulse_path that will
+        constitute the new file where data from these evolutions is stored.
+
+    Returns:
+    save_path :: str - The destination of the file where the
+        sweep information was saved
+    """
+    # Obtain the locks and full paths to the pulse and save files.
+    pulse_path_lock = "{}.lock".format(pulse_path)
+    # We assume that `pulse_path` is an h5 file with the suffix ".h5".
+    save_path = "{}_{}.h5".format(pulse_path.split(".h5")[0], suffix)
+    save_path_lock = "{}.lock".format(save_path)
+
+    # Grab the control pulse that achieved the minimum error.
+    try:
+        with FileLock(pulse_path_lock):
+            with h5py.File(pulse_path, "r") as pulse_file:
+                index = np.argmin(pulse_file["error"])
+                controls = pulse_file["controls"][index][()]
+            #ENDWITH
+        #ENDWITH
+    except Timeout:
+        print("Timeout when attempting to access {}".format(pulse_path_lock))
+        exit(1)
+
+    # Create the save file. Or use the existing sweep.
[…]

### 2019-11-19 setup.py

@@ -1,0 +1,25 @@
+"""
+setup.py - This module allows for package installation.
+"""
+
+from distutils.core import setup
+
+NAME = "rbqoc"
+VERSION = "0.1alpha"
+DEPENDENCIES = [
+    "filelock",
+    "h5py",
+    "matplotlib",
+    "numpy",
+]
+DESCRIPTION = "This package has utilities for performing robust qoc experiments."
+AUTHOR = "Thomas Propson"
+AUTHOR_EMAIL = "tcpropson@uchicago.edu"
+
+setup(author=AUTHOR,
+      author_email=AUTHOR_EMAIL,
+      description=DESCRIPTION,
+      install_requires=DEPENDENCIES,
+      name=NAME,
+      version=VERSION,
+)

### 2020-03-24 report/report/report.tex
update

@@
 
+  \subsection{Spin Experiment 4,5,6, pi\_corpse}
+  These experiments were intended to build a CORPSE $R_{x}(\pi)$ pulse
+  following this Ken Brown paper \cite{merrill2014progress}.
+
+  \subsection{Spin Experiment 7}
+  This experiment is the same as experiment 0 with the added constraint of
+  zero integration.
+
+  \section{Amplitude Robustness}
+
+  \subsection{Cavity Experiment 0}
+  0 to 2 in the cavity following \cite{heeres2017implementing}.
+
 \bibliography{report/refs}

### 2020-03-24 experiments/cavity/cavity0.py

@@ -1,0 +1,183 @@
+"""
+cavity0.py - Cavity Experiment 0.
+"""
+
+from argparse import ArgumentParser
+from copy import copy
+import os
+
+import autograd.numpy as anp
+from filelock import FileLock
+import h5py
+import numpy as np
+from qoc import (
+    evolve_schroedinger_discrete,
+    grape_schroedinger_discrete,
+)
+from qoc.standard import (
+    Adam, LBFGSB,
+    TargetStateInfidelity,
+    conjugate_transpose, matmuls, krons,
+    get_annihilation_operator, get_creation_operator,
+    generate_save_file_path,
+)
+
+# Define paths
+META_NAME = "cavity"
+EXPERIMENT_NAME = "cavity0"
+RBQOC_PATH = os.environ["ROBUST_QOC_PATH"]
+SAVE_PATH = os.path.join(RBQOC_PATH, "out", META_NAME, EXPERIMENT_NAME)
+
+# Define experimental constants. All units are in GHz.
+TRANSMON_FREQ = 2 * np.pi * 5.6640
+CAVITY_FREQ = 2 * np.pi * 4.4526 
+CHI = 2 * np.pi * -0.5644535742521878e-3
+ALPHA = 2 * np.pi * -0.13951263895963648
+KAPPA = 2 * np.pi * -3.3e-6
+CHI_PRIME = 2 * np.pi * -0.73e-6 
+MAX_AMP_NORM_CAVITY = np.sqrt(2) * 2 * np.pi * 4e-4
+MAX_AMP_NORM_TRANSMON = np.sqrt(2) * 2 * np.pi * 4e-3
+
+# Define the system
+CAVITY_STATE_COUNT = 3
+CAVITY_ANNIHILATE = get_annihilation_operator(CAVITY_STATE_COUNT)
+CAVITY_CREATE = get_creation_operator(CAVITY_STATE_COUNT)
+CAVITY_NUMBER = np.matmul(CAVITY_CREATE, CAVITY_ANNIHILATE)
+CAVITY_C2_A2 = matmuls(CAVITY_CREATE, CAVITY_CREATE, CAVITY_ANNIHILATE, CAVITY_ANNIHILATE)
+CAVITY_ID = np.eye(CAVITY_STATE_COUNT)
+CAVITY_VACUUM = np.zeros((CAVITY_STATE_COUNT, 1))
+CAVITY_ZERO = np.copy(CAVITY_VACUUM)
[…]

### 2020-03-24 experiments/cavity/cavity0_amp.py

@@ -1,0 +1,51 @@
+"""
+caity0_amp.py - Do a sweep of the control amplitude multiplier
+for cavity experiment 0.
+"""
+
+import os
+
+import numpy as np
+from rbqoc.util import (amp_sweep, plot_amp_sweep)
+from cavity0 import (EVOL_CONFIG, EXPERIMENT_NAME, SAVE_PATH)
+
+# Sweep trials
+SWEEP_TRIAL_COUNT = 100
+AMP_MULTIPLIERS = np.linspace(0.9, 1.1, SWEEP_TRIAL_COUNT)
+
+# I/O
+LABELS = ("QOC", "RBQOC")
+PULSE_FILE_0 = "00003_cavity0.h5"
+PULSE_FILE_1 = "00006_cavity0.h5"
+PULSE_FILES = (
+    PULSE_FILE_0,
+    PULSE_FILE_1,
+)
+PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
+PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
+PULSE_PATHS = (
+    PULSE_PATH_0,
+    PULSE_PATH_1,
+)
+SWEEP_SUFFIX = "amp"
+PLOT_SAVE_FILE = "{}_{}.png".format(EXPERIMENT_NAME, SWEEP_SUFFIX)
+PLOT_SAVE_PATH = os.path.join(SAVE_PATH, PLOT_SAVE_FILE)
+
+def main():
+    sweep_path_list = list()
+    # Sweep the amplitude parameter for each pulse.
+    for i, pulse_path in enumerate(PULSE_PATHS):
+        save_path = amp_sweep(AMP_MULTIPLIERS, EVOL_CONFIG,
+                              pulse_path, SWEEP_SUFFIX)
+        sweep_path_list.append(save_path)
+
+    # Plot the sweeps together.
+    plot_amp_sweep(PLOT_SAVE_PATH, sweep_path_list,
+                   title=EXPERIMENT_NAME, labels=LABELS,
+                   x_label="Amplitude Multiplier",
+                   y_label="Infidelity",
+                   y_max=0.2, x_min=0.9, x_max=1.1)
+
+
+if __name__ == "__main__":
+    main()

### 2020-03-24 experiments/spin/spin8.py

@@ -1,0 +1,183 @@
+"""
+spin8.py - spin experiment 8
+"""
+
+from argparse import ArgumentParser
+from copy import copy
+import os
+
+import autograd.numpy as anp
+from filelock import FileLock
+import h5py
+import numpy as np
+from qoc import (
+    evolve_schroedinger_discrete,
+    grape_schroedinger_discrete,
+)
+from qoc.standard import (
+    Adam, LBFGSB,
+    ControlArea,
+    TargetStateInfidelity,
+    TargetStateInfidelityTime,
+    conjugate_transpose, matmuls, krons,
+    get_annihilation_operator, get_creation_operator,
+    generate_save_file_path,
+    SIGMA_X, SIGMA_Z
+)
+
+# Define paths
+META_NAME = "spin"
+EXPERIMENT_NAME = "spin8"
+SAVE_PATH = os.path.join(os.environ["ROBUST_QOC_PATH"], "out", META_NAME, EXPERIMENT_NAME)
+
+
+# Define experimental constants.
+FLUXONIUM_FREQ = 2 * np.pi * 1e-2 #GHz
+MAX_AMP_NORM_FLUXONIUM = 2 * np.pi * 3e-1 #GHz
+SAMPLE_RATE = 1.2 #samples/ns
+
+
+# Define the system
+FLUXONIUM_STATE_COUNT = 2
+FLUXONIUM_VACUUM = np.zeros((FLUXONIUM_STATE_COUNT, 1))
+FLUXONIUM_G = np.copy(FLUXONIUM_VACUUM)
+FLUXONIUM_G[0][0] = 1.
+FLUXONIUM_E = np.copy(FLUXONIUM_VACUUM)
+FLUXONIUM_E[1][0] = 1.
+HAMILTONIAN_ARGS = np.array([FLUXONIUM_FREQ])
+H_SYSTEM = (
+    SIGMA_Z / 2
+)
+H_CONTROL_0 = (
+    SIGMA_X / 2
+)
+
+hamiltonian = lambda controls, hargs, time: (
+    hargs[0] * H_SYSTEM
+    + controls[0] * H_CONTROL_0
+)
+CONTROL_COUNT = 1
+COMPLEX_CONTROLS = False
+MAX_CONTROL_NORMS = np.array((
+    MAX_AMP_NORM_FLUXONIUM,
+))
+
+
+# Define the problem
[…]

### 2020-03-24 experiments/spin/spin9.py

@@ -1,0 +1,145 @@
+"""
+spin9.py - Spin experiment 9.
+"""
+
+from argparse import ArgumentParser
+import os
+
+import autograd.numpy as anp
+from filelock import FileLock, Timeout
+import h5py
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+from qoc import grape_schroedinger_discrete, evolve_schroedinger_discrete
+from qoc.standard import (TargetStateInfidelity,
+                          conjugate_transpose,
+                          get_annihilation_operator,
+                          get_creation_operator,
+                          SIGMA_Z, SIGMA_X,
+                          generate_save_file_path,
+                          krons, plot_controls,
+                          plot_state_population,
+                          Adam, LBFGSB,)
+
+CORE_COUNT = 8
+os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
+os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
+
+# Define experimental constants.
+CONTROL_SAMPLING_RATE = 1.2 # GS/s
+MAX_AMP_CONTROL_0 = 2 * anp.pi * 1e-1 # GHz
+MAX_AMP_BANDWIDTH_CONTROL_0 = 2 * anp.pi * 5e-1 # GHz
+OMEGA_Q = 2 * anp.pi * 1e-2 # GHz
+
+# Define the system.
+SYSTEM_HAMILTONIAN_0 = SIGMA_Z / 2
+CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
+HAMILTONIAN_ARGS = anp.array([OMEGA_Q])
+hamiltonian = lambda controls, hargs, time: (
+    hargs[0] * SYSTEM_HAMILTONIAN_0
+    + controls[0] * CONTROL_HAMILTONIAN_0
+)
+MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
+MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
+
+# Define the optimization.
+WDIR = os.environ["ROBUST_QOC_PATH"]
+GRAB_CONTROLS = False
+if GRAB_CONTROLS:
+    CONTROLS_PATH = os.path.join(WDIR, "out/spin/spin9/00001_spin9.h5")
+    CONTROLS_PATH_LOCK = "{}.lock".format(CONTROLS_PATH)
+    try:
[…]

### 2020-03-24 experiments/spin/spin9_amp.py

@@ -1,0 +1,57 @@
+"""
+spin9_amp.py - Do a sweep of the control amplitude multiplier
+for spin9.
+"""
+
+import os
+
+import numpy as np
+from rbqoc.util import (amp_sweep, plot_amp_sweep)
+from spin9 import (EVOL_CONFIG, EXPERIMENT_NAME, EXPERIMENT_META)
+
+# Sweep trials
+SWEEP_TRIAL_COUNT = 100
+AMP_MULTIPLIERS = np.linspace(0.98, 1.02, SWEEP_TRIAL_COUNT)
+
+
+# I/O
+WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
+SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
+PULSE_FILE_0 = "00000_spin9.h5"
+PULSE_FILE_1 = "00001_spin9.h5"
+PULSE_FILE_2 = "00005_spin9.h5"
+PULSE_FILES = (
+    PULSE_FILE_0,
+    PULSE_FILE_1,
+    PULSE_FILE_2,
+)
+PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
+PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
+PULSE_PATH_2 = os.path.join(SAVE_PATH, PULSE_FILE_2)
+PULSE_PATHS = (
+    PULSE_PATH_0,
+    PULSE_PATH_1,
+    PULSE_PATH_2,
+)
+SWEEP_SUFFIX = "amp"
+PLOT_SAVE_FILE = "{}_{}.png".format(EXPERIMENT_NAME, SWEEP_SUFFIX)
+PLOT_SAVE_PATH = os.path.join(SAVE_PATH, PLOT_SAVE_FILE)
+
+def main():
+    sweep_path_list = list()
+    # Sweep the amplitude parameter for each pulse.
+    for i, pulse_path in enumerate(PULSE_PATHS):
+        save_path = amp_sweep(AMP_MULTIPLIERS, EVOL_CONFIG,
+                              pulse_path, SWEEP_SUFFIX)
+        sweep_path_list.append(save_path)
+
+    # Plot the sweeps together.
+    plot_amp_sweep(PLOT_SAVE_PATH, sweep_path_list,
+                   title=EXPERIMENT_NAME,
+                   x_label="Amplitude Multiplier",
[…]

### 2020-03-24 experiments/spin/spin9_hp.py

@@ -1,0 +1,63 @@
+"""
+spin9_hp.py - Do a sweep of the hamiltonian parameters
+for spin9.
+"""
+
+import os
+
+import numpy as np
+from rbqoc.util import (hamiltonian_args_sweep,
+                        plot_hamiltonian_args_sweep,)
+from spin_exp0 import (EVOL_ARGS, EXPERIMENT_META, EXPERIMENT_NAME, HAMILTONIAN_ARGS)
+
+# Sweep trials
+SWEEP_TRIAL_COUNT = 100
+SWEEP_MULTIPLIERS = np.linspace(0, 3, SWEEP_TRIAL_COUNT)
+HAMILTONIAN_ARGS_SWEEP = np.stack([HAMILTONIAN_ARGS] * SWEEP_TRIAL_COUNT)
+for i, multiplier in enumerate(SWEEP_MULTIPLIERS):
+    HAMILTONIAN_ARGS_SWEEP[i] *= SWEEP_MULTIPLIERS[i]
+
+# I/O
+WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
+SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
+PULSE_FILE_0 = "00001_spin9.h5"
+PULSE_FILE_1 = "00006_spin_exp0.h5"
+PULSE_FILE_2 = "00013_spin_exp0.h5"
+PULSE_FILE_3 = "00010_spin_exp0.h5"
+PULSE_FILES = [PULSE_FILE_0,
+               PULSE_FILE_1,
+               PULSE_FILE_2,
+               PULSE_FILE_3,
+]
+PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
+PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
+PULSE_PATH_2 = os.path.join(SAVE_PATH, PULSE_FILE_2)
+PULSE_PATH_3 = os.path.join(SAVE_PATH, PULSE_FILE_3)
+PULSE_PATH_4 = os.path.join(SAVE_PATH, PULSE_FILE_4)
+PULSE_PATHS = [
+    PULSE_PATH_0,
+    PULSE_PATH_1,
+    PULSE_PATH_2,
+    PULSE_PATH_3,
+]
+SWEEP_SUFFIX = "freq_sweep"
[…]

### 2020-03-24 experiments/spin/spin_constrained.py

@@ -1,0 +1,136 @@
+"""
+spin_constrained.py
+"""
+
+from argparse import ArgumentParser
+from copy import copy
+
+import autograd.numpy as anp
+import numpy as np
+from qoc import grape_schroedinger_discrete
+from qoc.standard import (TargetStateInfidelity,
+                          conjugate_transpose,
+                          get_annihilation_operator,
+                          get_creation_operator,
+                          SIGMA_Z, SIGMA_X,
+                          generate_save_file_path,)
+
+MAX_AMP_0 = 2 * anp.pi * 3e-1
+OMEGA = 2 * anp.pi * 1e-2
+
+# Define the system.
+HILBERT_SIZE = 2
+H_SYSTEM_0 = OMEGA * SIGMA_Z / 2
+H_CONTROL_0 = SIGMA_X / 2
+hamiltonian = lambda controls, hamiltonian_args, time: (
+    H_SYSTEM_0
+    + controls[0] * H_CONTROL_0
+)
+MAX_CONTROL_NORMS = anp.array([MAX_AMP_0])
+
+# Define the problem.
+INITIAL_STATE_0 = anp.array([[1], [0]])
+TARGET_STATE_0 = anp.array([[0], [1]])
+INITIAL_STATES = anp.stack((INITIAL_STATE_0,),)
+TARGET_STATES = anp.stack((TARGET_STATE_0,),)
+COSTS = [
+    TargetStateInfidelity(TARGET_STATES)
+]
+
+# Define the optimization.
+COMPLEX_CONTROLS = False
+CONTROL_COUNT = 1
+EVOLUTION_TIME = 100 # nanoseconds
+CONTROL_EVAL_COUNT = SYSTEM_EVAL_COUNT = 2 * int(EVOLUTION_TIME) + 1
+ITERATION_COUNT = 1000
+
+def gs(X, row_vecs=True, norm=True):
+    if not row_vecs:
+        X = X.T
+    Y = X[0:1,:].copy()
+    for i in range(1, X.shape[0]):
+        proj = np.diag((X[i,:].dot(Y.T)/np.linalg.norm(Y,axis=1)**2).flat).dot(Y)
+        Y = np.vstack((Y, X[i,:] - proj.sum(0)))
+    if norm:
+        Y = np.diag(1/np.linalg.norm(Y,axis=1)).dot(Y)
+    if row_vecs:
+        return Y
+    else:
+        return Y.T
+
+    
+# Generate orthonormal basis for zero net flux constraint manifold.
[…]

### 2020-03-24 experiments/spin/spin_exp7.py

@@ -1,0 +1,147 @@
+"""
+spin_exp7.py - Spin experiment 7.
+"""
+
+from argparse import ArgumentParser
+import os
+
+import autograd.numpy as anp
+from filelock import FileLock, Timeout
+import h5py
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+from qoc import grape_schroedinger_discrete, evolve_schroedinger_discrete
+from qoc.standard import (TargetStateInfidelity,
+                          conjugate_transpose,
+                          get_annihilation_operator,
+                          get_creation_operator,
+                          SIGMA_Z, SIGMA_X,
+                          generate_save_file_path,
+                          krons, plot_controls,
+                          plot_state_population,
+                          Adam, LBFGSB,)
+
+CORE_COUNT = 8
+os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
+os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
+
+# Define experimental constants.
+CONTROL_SAMPLING_RATE = 1.2 # GS/s
+MAX_AMP_CONTROL_0 = 2 * anp.pi * 1e-1 # GHz
+MAX_AMP_BANDWIDTH_CONTROL_0 = 2 * anp.pi * 5e-1 # GHz
+OMEGA_Q = 2 * anp.pi * 1e-2 # GHz
+
+# Define the system.
+SYSTEM_HAMILTONIAN_0 = SIGMA_Z / 2
+CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
+HAMILTONIAN_ARGS = anp.array([OMEGA_Q])
+hamiltonian = lambda controls, hargs, time: (
+    hargs[0] * SYSTEM_HAMILTONIAN_0
+    + controls[0] * CONTROL_HAMILTONIAN_0
+)
+MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
+MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
+
+# Define the optimization.
+WDIR = os.environ["ROBUST_QOC_PATH"]
+GRAB_CONTROLS = False
+if GRAB_CONTROLS:
+    CONTROLS_PATH = os.path.join(WDIR, "out/spin/spin_exp0/00017_spin_exp0.h5")
+    CONTROLS_PATH_LOCK = "{}.lock".format(CONTROLS_PATH)
+    try:
[…]

### 2020-03-24 rbqoc/util/amputil.py

@@ -103,8 +103,9 @@ def amp_sweep(amp_multipliers, evol_config,
 
 
 def plot_amp_sweep(out_path, sweep_path_list,
-                   dpi=1000,
-                   title=None, x_label=None, y_label=None):
+                   dpi=1000, labels=None, title=None,
+                   x_label=None, x_max=None, x_min=None,
+                   y_label=None, y_max=None, y_min=None):
     """
     Given a list of files generated by `amp_sweep`, plot their error
     against one of the args in amp_multipliers.
@@ -143,13 +144,27 @@ def plot_amp_sweep(out_path, sweep_path_list,
     # Plot each set of amp_multipliers and associated errors.
     for i, errors in enumerate(sweep_errors):
         amp_multipliers = sweep_amp_multipliers[i]
-        plt.plot(amp_multipliers, errors, color=get_color(i))
+        kwargs = dict()
+        if labels is not None:
+            kwargs.update({
+                "label": labels[i]
+            })
+        plt.plot(amp_multipliers, errors, color=get_color(i), **kwargs)
     if title is not None:
         plt.title(title)
     if x_label is not None:
         plt.xlabel(x_label)
     if y_label is not None:
         plt.ylabel(y_label)
+    plt.legend()
+    if x_max is not None:
+        plt.xlim(right=x_max)
+    if x_min is not None:
+        plt.xlim(left=x_min)
+    if y_max is not None:
+        plt.ylim(top=y_max)
+    if y_min is not None:
+        plt.ylim(bottom=y_min)
     plt.savefig(out_path, dpi=dpi)
 
 

### 2020-07-10 paper/paper.tex
started paper dir

\documentclass[
  amsfonts,
  amsmath,
  amssymb,
  aps,
  nobibnotes,
  prl,
  twocolumn,
]{revtex4-2}

\begin{document}

\end{document}

### 2020-07-10 experiments/spin/spin15.jl

@@ -12,11 +12,11 @@ include(joinpath(ENV["ROBUST_QOC_PATH"], "rbqoc.jl"))
 # paths
 EXPERIMENT_META = "spin"
 EXPERIMENT_NAME = "spin15"
-WDIR = ENV["ROBUST_QOC_PATH"]
+WDIR = ENV["RBQOC_PATH"]
 SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 
 # Define the optimization.
-EVOLUTION_TIME = 20.0
+EVOLUTION_TIME = 56.80
 CONTROL_COUNT = 1
 DT_INTEGRATOR = 1
 DT_INIT = DT_PREF
@@ -38,8 +38,14 @@ INITIAL_ASTATE = [
     @SVector zeros(1); # int_gamma
 ]
 ASTATE_SIZE, = size(INITIAL_ASTATE)
-TARGET_STATE_0 = SA[1., 0, -1, 0] / sqrt(2)
-TARGET_STATE_1 = SA[0., 1, 0, 1] / sqrt(2)
+ZPIBY2_0 = SA[1., 0, -1, 0] / sqrt(2)
+ZPIBY2_1 = SA[0., 1, 0, 1] / sqrt(2)
+YPIBY2_0 = SA[1., 1, 0, 0] / sqrt(2)
+YPIBY2_1 = SA[-1., 1, 0, 0] / sqrt(2)
+XPIBY2_0 = SA[1., 0, 0, -1] / sqrt(2)
+XPIBY2_1 = SA[0., 1, -1, 0] / sqrt(2)
+TARGET_STATE_0 = XPIBY2_0
+TARGET_STATE_1 = XPIBY2_1
 TARGET_ASTATE = [
     TARGET_STATE_0;
     TARGET_STATE_1;
@@ -55,14 +61,6 @@ INT_CONTROLS_IDX = STATE_1_IDX[end] + 1:STATE_1_IDX[end] + CONTROL_COUNT
 CONTROLS_IDX = INT_CONTROLS_IDX[end] + 1:INT_CONTROLS_IDX[end] + CONTROL_COUNT
 DCONTROLS_DT_IDX = CONTROLS_IDX[end] + 1:CONTROLS_IDX[end] + CONTROL_COUNT
 INT_GAMMA_IDX = DCONTROLS_DT_IDX[end] + 1:DCONTROLS_DT_IDX[end] + 1
-
-
-# Generate initial controls.
[…]

### 2020-07-10 rbqoc.jl

@@ -3,6 +3,7 @@ rbqoc.jl - common definitions for the rbqoc repo
 """
 
 # imports
+using Dierckx
 using DifferentialEquations
 using HDF5
 using Plots
@@ -86,14 +87,13 @@ function generate_save_file_path(extension, save_file_name, save_path)
 end
 
 
-function grab_controls(controls_file_path; mode="r", save_type=jl)
-    data = h5open(controls_file_path, mode) do save_file
+function grab_controls(save_file_path; save_type=jl)
+    data = h5open(save_file_path, "r") do save_file
         if save_type == jl
             cidx = read(save_file, "controls_idx")
-            controls = read(save_file, "states")[:, cidx]
+            controls = read(save_file, "astates")[:, cidx]
             evolution_time = read(save_file, "evolution_time")
-        end
-        if save_type == samplejl
+        elseif save_type == samplejl
             controls = read(save_file, "controls_sample")
             evolution_time = read(save_file, "evolution_time_sample")
         end
@@ -103,6 +103,18 @@ function grab_controls(controls_file_path; mode="r", save_type=jl)
     return data
 end
 
+
+function read_save(save_file_path)
+    dict = h5open(save_file_path, "r") do save_file
+        dict = Dict()
+        for key in names(save_file)
+            dict[key] = read(save_file, key)
+        end
+        return dict
+    end
+
+    return dict
+end
 
 """
 horner - compute the value of a polynomial using Horner's method
@@ -122,24 +134,28 @@ function horner(coeffs, val)
 end
 
 
-function plot_controls(controls_file_path, save_file_path;
+function plot_controls(save_file_path, plot_file_path;
                        save_type=jl, title=nothing)
     # Grab and prep data.
-    (controls, evolution_time) = grab_controls(controls_file_path; save_type=save_type)
+    (controls, evolution_time) = grab_controls(save_file_path; save_type=save_type)
     controls = controls ./ (2 * pi)
-    (control_eval_count,) = size(controls)
+    (control_eval_count, control_count) = size(controls)
     control_eval_times = Array(1:1:control_eval_count) * DT_PREF
-    file_name = split(basename(controls_file_path), ".h5")[1]
+    file_name = split(basename(save_file_path), ".h5")[1]
[…]

### 2020-07-10 paper/paper.tex
draft up to t1 relaxation section

@@
 
+\input{preamble}
+
 \begin{document}
 
+%% TITLE
+\title{Robust Quantum Optimal Control}
+
+\author{Thomas Propson}
+\email{tcpropson@uchicago.edu}
+\affiliation{
+  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
+}
+
+\date{\today}
+
+\maketitle
+
+%% ABSTRACT
+\begin{abstract}
+  This is a paper about robust quantum optimal control.
+\end{abstract}
+
+%% SECTION 1
+\section{Introduction}
+[Existing work] The leading model of universal quantum computation is gate-based. There are analytic
+techniques to construct gates \cite{huang2020engineering}
+\cite{merrill2014progress} \cite{han2020experimental} \cite{wang2007nonadiabatic}.
+Most methods focus on optimizing a few aspects of the gate, but not all
+experimentally relevant aspects. There is a growing literature on numerical
+techniques to construct gates \cite{leung2017speedup} \cite{abdelhafez2019gradient}.
+%% TODO: Cite CRAB, GOAT, KROTOV, etc.
+These methods formulate the quantum optimal control problem as unconstrained.
+They typically rely on zeroth-order or first-order optimizers. They are not
+sofisticated enough to handle all of the relevant constraints simultaneously.
+Although some analytic techniques exist to design pulses robust
+to decoherence, no numerical techniques have been presented to
+design pulses robust to decoherence (as far as the author is aware).
+
+[This work] We employ the trajectory optimization literature to formulate the quantum
+optimal control problem as a constrained optimization problem.
+We study the quantum optimal control problem on the fluxonium.
+We outline experimentally realistic constraints and map them to
+the trajectory optimization framework. For the device we study
+we achieve a 2x increase in $T_{1}$ times.
+We present two methods for achieving robustness to system parameter deviations,
+and compare to existing dynamic decoupling methods. We find that 
+our methods beat dynamic decoupling and mitigate dephasing by order X.
+
+[Outline]  First we formulate the quantum optimal control problem
+in the trajectory optimization framework. Then, we introduce the dynamics
+of the fluxonium device and outline experimental considerations
+relevant to gate construction. Next we outline a method for
+making the optimization $T_{1}$ aware. Finally, we present
+some methods for engineering robustness to decoherence and
+compare them to existing techniques.
+
+%% SECTION 2
+\section{QOC + AL-iLQR}
+[QOC Problem Statement] I have some initial
+configuration and I want to reach some target configuration,
+either single- or multi-state transfer, subject to the dynamics
+\begin{align}
+  \frac{\partial \ket{\psi}(t)}{\partial t} &= -/frac{i}{\hbar} H(t) \ket{\psi}(t)
+\end{align}
+Common numerical techniques include approximating the analytic unitary propagator solution,
+or employing explicit/implicit Runge-Kutta methods of the form
+\begin{align}
+  \ket{\psi_{k + 1}} &= \ket{\psi_{k}} + \frac{\partial \ket{\psi_{k}}}{\partial t} \cdot dt
+\end{align}
+The interesting part is that your hamiltonian $H(t)$ has a time-dependent
+control parameter $u(t)$ that the experimentalist gets to control, e.g. flux threading
+a superconducting junction.
+\begin{align}
+  H(t) &= H_{0} + \sum_{i = 1}^{m} a_{m}(u_{m}(t)) H_{m}\\
+\end{align}
+Most often $a_{m}(u) = u$ but any arbitrary dependence is allowed.
+The technique of adjusting $\ket{\psi_{N}}$ based on $\textbf{u}(t)$
+is the goal of the quantum optimal control optimizer. This domain is
+know as sensitivity analysis. GRAPE does
+\begin{align}
+  \min_{U:N-1 \times m} &1 - {\lvert \braket{\psi_{N}(U)}{\psi_{t}} \lvert}^{2}\\
+\end{align}
+with
+\begin{align}
+  U^{(k + 1)} &= U^{(k)} - \alpha \textbf{\nabla_{\textbf{U}}} (1 - {\lvert \braket{\psi_{N}(U)}{\psi_{t}} \lvert}^{2})\\
+\end{align}
+In general they add more cost functions than target fidelity.
+%% TOOD: Other update procedures might differ, e.g. Krotov does some sequential updates
+%% and Houck, Google use black box optimizers.
+
+[AL-iLQR Problem Statement] Trajectory optimization gives us gaurantees
+about our updates via Ricatti recursion
+and allows us to put constraints on our cost functions.
+\begin{align}
+  %% This is e.q. 2 from ALTRO paper
+  \min_{U:N-1 \times m} &\mathcal{L}_{N}(X_{N}, \lambda_{N}, \mu_{N}) + \sum_{k = 1}^{N - 1} \mathcal{L}_{k}(X_{k}, U_{k}, \lambda_{k}, \mu_{k})\\
+  \mathcal{L}_{k}(X_{k}, U_{k}, \lambda_{k}, \mu_{k}) &= (\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(X_{k}, U_{k}))^{T} c_{k}(X_{K}, U_{K})
+  &+ (X_{k} - X_{f})^{T} Q (X_{k} - X_{f}) + U^{T} R U\\
+\end{align}
+The important point is that there is an update step (e.q. 17 from ALTRO paper) where we send \lambda -> \infty
+and get all of the nice convergence properties. The weights are adjusted dynamically between iterations until
+all of our constraints are satisfied. The Markovian decision structure of the problem allows
+us to apply differentiable dynamic programming to garauntee that the update for each control $U_{k}$ is
+opimal, as apposed to the greedy updates of first-order optimizers like the naive gradient descent.
+
+%% TODO: (?) Figure comparing update procedures
+
+%% SECTION 3
+\section{Our Problem}
+[Fluxonium + Device Characterization] In the two-level
+approximation we have
+\begin{align}
+  H/h &= \omega_{q} \frac{\sigma_{z}}{2} + A(\Phi_{ext}) \frac{\sigma_{x}}{2}\\
+\end{align}
+This approximation is good up to the avoided crossing at 0.35 $\Phi_{0}$. We get $A$
+by converting via $\bra{g} \hat{\phi} \ket{e}$.
+%% TODO: list device parameters
+
+[Constraints] We want constraints on our pulses. We want pulses start and end at zero
+for concatenation. We want pulses to have zero net flux to mitigate
+hysteresis in flux bias lines.
+%% TODO: See ref. 28 in Helin paper
+We want the amplitude to be constrained
+$\delta \Phi_{\textrm{ext}} \sim 0.06 \Phi_{0}$
+so the two-level approximation stays valid.
+We want our state to obey normalization conditions,
+mitigating numerical error in simulaiton.
+
+We test on the basis gates $X/2, Y/2, Z/2$. Universal
+up to arbitrary Z rotation.
+
+%% SECTION 4
+\section{Robustness to $T_{1}$-type Noise}
+
+%% SECTION 5
+\section{Robustness to $T_{\phi}$-type Noise}
+
+\bibliography{refs}
+
 \end{document}

### 2020-07-17 paper/paper.tex
progress on paper
progress

@@
   amsmath,
+  tbtags,
   amssymb,
@@
 
-\maketitle
-
 %% ABSTRACT
@@
 
+\maketitle
+
 %% SECTION 1
 \section{Introduction}
-[Existing work] The leading model of universal quantum computation is gate-based. There are analytic
-techniques to construct gates \cite{huang2020engineering}
-\cite{merrill2014progress} \cite{han2020experimental} \cite{wang2007nonadiabatic}.
+(Existing work) The leading model of universal
+quantum computation is gate-based. Analytic methods
+have been demonstrated to construct gates.
+Some focus on solving the TDSE analytically
+\cite{}Analytic methods have
+been demonstrated to construct gates
+\cite{zhang2020universal, huang2020engineering,
+  han2020experimental, xu2020nonadiabatic}.
 Most methods focus on optimizing a few aspects of the gate, but not all
-experimentally relevant aspects. There is a growing literature on numerical
-techniques to construct gates \cite{leung2017speedup} \cite{abdelhafez2019gradient}.
+experimentally relevant aspects. There is a growing literature
+on numerical
+techniques to construct gates \cite{leung2017speedup,
+  goerz2019krotov, doria2011optimal, abdelhafez2019gradient,
+  machnes2015gradient}.
 %% TODO: Cite CRAB, GOAT, KROTOV, etc.
-These methods formulate the quantum optimal control problem as unconstrained.
-They typically rely on zeroth-order or first-order optimizers. They are not
-sofisticated enough to handle all of the relevant constraints simultaneously.
+These methods formulate the quantum optimal
+control problem as unconstrained.
+They typically rely on zeroth-order or first-order optimizers.
+They are not
+sofisticated enough to handle all of the relevant
+constraints simultaneously.
 Although some analytic techniques exist to design pulses robust
@@
 
-[This work] We employ the trajectory optimization literature to formulate the quantum
+(This work) We employ the trajectory optimization
+literature to formulate the quantum
 optimal control problem as a constrained optimization problem.
@@
 
-[Outline]  First we formulate the quantum optimal control problem
+(Outline)  First we formulate the quantum optimal control problem
 in the trajectory optimization framework. Then, we introduce the dynamics
@@
 \section{QOC + AL-iLQR}
-[QOC Problem Statement] I have some initial
+(QOC Problem Statement) I have some initial
 configuration and I want to reach some target configuration,
@@
 \begin{align}
-  \frac{\partial \ket{\psi}(t)}{\partial t} &= -/frac{i}{\hbar} H(t) \ket{\psi}(t)
+  \partial_{t} \ket{\psi(t)} &= -\frac{i}{\hbar} H(t) \ket{\psi(t)}
 \end{align}
@@
 \begin{align}
-  \ket{\psi_{k + 1}} &= \ket{\psi_{k}} + \frac{\partial \ket{\psi_{k}}}{\partial t} \cdot dt
+  \ket{\psi^{(k + 1)}} &= \ket{\psi^{(k)}} + \partial_{t} \ket{\psi^{(k)}} \cdot dt
 \end{align}
@@
 \begin{align}
-  H(t) &= H_{0} + \sum_{i = 1}^{m} a_{m}(u_{m}(t)) H_{m}\\
+  H(t) &= H_{0} + \sum_{i = 1}^{m} a_{m}(u_{m}(t)) H_{m}
 \end{align}
@@
 \begin{align}
-  \min_{U:N-1 \times m} &1 - {\lvert \braket{\psi_{N}(U)}{\psi_{t}} \lvert}^{2}\\
+  \min_{U:N-1 \times m} &1 - {\lvert \braket{\psi_{N}(U) \lvert \psi_{t}} \lvert}^{2}
 \end{align}
-with
+subject to the gradient descent update procedure
 \begin{align}
-  U^{(k + 1)} &= U^{(k)} - \alpha \textbf{\nabla_{\textbf{U}}} (1 - {\lvert \braket{\psi_{N}(U)}{\psi_{t}} \lvert}^{2})\\
+  U^{'} &= U - \alpha \nabla_{U} (1 - {\lvert \braket{\psi_{N}(U) \lvert \psi_{t}} \lvert}^{2})
 \end{align}
@@
 
-[AL-iLQR Problem Statement] Trajectory optimization gives us gaurantees
+(AL-iLQR Problem Statement) Trajectory optimization gives us gaurantees
 about our updates via Ricatti recursion
 and allows us to put constraints on our cost functions.
-\begin{align}
 %% This is e.q. 2 from ALTRO paper
-  \min_{U:N-1 \times m} &\mathcal{L}_{N}(X_{N}, \lambda_{N}, \mu_{N}) + \sum_{k = 1}^{N - 1} \mathcal{L}_{k}(X_{k}, U_{k}, \lambda_{k}, \mu_{k})\\
-  \mathcal{L}_{k}(X_{k}, U_{k}, \lambda_{k}, \mu_{k}) &= (\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(X_{k}, U_{k}))^{T} c_{k}(X_{K}, U_{K})
-  &+ (X_{k} - X_{f})^{T} Q (X_{k} - X_{f}) + U^{T} R U\\
-\end{align}
-The important point is that there is an update step (e.q. 17 from ALTRO paper) where we send \lambda -> \infty
+\begin{equation}
+\begin{aligned}[b]
+  \min_{U:N-1 \times m} &\mathcal{L}_{N}(x_{N}, \lambda_{N}, \mu_{N})\\
+  &+ \sum_{k = 1}^{N - 1} \mathcal{L}_{k}(x_{k}, u_{k}, \lambda_{k}, \mu_{k})\\
+\end{aligned}
+\end{equation}
+\begin{equation}
+\begin{aligned}[b]
+  \mathcal{L}_{k}(x_{k}, &u_{k}, \lambda_{k}, \mu_{k}) =\\
+  &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
+  &+ (x_{k} - x_{f})^{T} Q (x_{k} - x_{f}) + u_{k}^{T} R u_{k}\\
+\end{aligned}
+\end{equation}
+The important point is that there is an update step (e.q. 17 from ALTRO paper) where we send $\lambda \rightarrow \infty$
 and get all of the nice convergence properties. The weights are adjusted dynamically between iterations until
@@
 \section{Our Problem}
-[Fluxonium + Device Characterization] In the two-level
+(Fluxonium Device) In the two-level
 approximation we have
 \begin{align}
-  H/h &= \omega_{q} \frac{\sigma_{z}}{2} + A(\Phi_{ext}) \frac{\sigma_{x}}{2}\\
+  H/h &= \omega_{q} \frac{\sigma_{z}}{2} + A(\Phi_{ext}) \frac{\sigma_{x}}{2}
 \end{align}
-This approximation is good up to the avoided crossing at 0.35 $\Phi_{0}$. We get $A$
+This approximation is good up to the avoided crossing at $\Phi_{\textrm{ext}} = 0.35 \Phi_{0}$. We get $A$
 by converting via $\bra{g} \hat{\phi} \ket{e}$.
+We use the gates $X/2, Y/2, Z/2$. $X/2, Y/2$ are universal
+up to an arbitrary Z rotation.
 %% TODO: list device parameters
 
-[Constraints] We want constraints on our pulses. We want pulses start and end at zero
+(Constraints) We want constraints on our pulses. We want pulses start and end at zero
 for concatenation. We want pulses to have zero net flux to mitigate
@@
 mitigating numerical error in simulaiton.
+%% TODO: mention pulse smoothing via optimizing over 2nd derivative
+%% of pulses
+%% TODO: probably want to list equations for the constraints
 
-We test on the basis gates $X/2, Y/2, Z/2$. Universal
-up to arbitrary Z rotation.
+($T_{1}$ and $T_{\phi}$ noise)
+Dissipation to the thermal bath via longitudinal
+relaxation ($T_{1}$) is an irreversible process
+that results in information loss.
+Pure dephasing ($T_{\phi}$) is a reversible process.
+Dielectric loss, etc. contribute
+to $T_{1}$. $1/f$ flux noise, etc. contribute to $T_{\phi}$
+There is a tradeoff between $T_{1}$ and $T_{\phi}$. In the case of white
+noise we have that the sum of the noise weights $W_{1}$ and $W_{\phi}$
+is constant \cite{huang2020engineering}.
+In our case $T_{\phi}$ is saturated at the flux
+frustration point $\Phi_{\textrm{ext}} = 0.5 \Phi_{0}$
+$T_{\phi}$ decreases as we move away from the flux
+frustration point. Conversely, $T_{1}$ is at a minimum
+around the flux frustration point $T_{1}(0.5 \Phi_{0}) = 0.315$ms,
+and increases away from the flux frustration point
+$T_{1}(0.43 \Phi_{0}) = 4.3$ms. Given the nature
+of the decay processes and the tradeoff, we choose
+to optimize the flux bias to have high time-averaged
+$T_{1}$ times and employ some techniques to mitigate
+$T_{\phi}$ dephasing.
+%% TODO: See fig. 3 in Helin paper
 
@@
 \section{Robustness to $T_{1}$-type Noise}
+(Strategy) We seek to minimize the probability
+that the qubit relaxes longitudinally.
+To this end we augment the state vector $x_{k}$ with the
+longitudinal relaxation probability
+\begin{align}
+  P_{1}(t_{k}) = \int_{0}^{t_{k}}
+  \gamma_{1}(\Phi_{\textrm{ext}}(t)) dt
+\end{align}
+where $\gamma_{1} = T_{1}^{-1}$. Setting the target
+longitudinal relaxation probability to 0 results in
+a quadratic cost at each knot point
+of the form ${\lvert P_{1}(t_{k}) \rvert}^{2}$.
+
+The value $T_{1}(\Phi_{\textrm{ext}}(t_{k}))$
+is obtained at each knot point by evaluating
+a spline interpolant fit to
+experimentally obtained data of the form
+$\{(\Phi_{\textrm{ext}}, T_{1})\}$.
+Calculating $T_{1}$ directly from theoretical
+considerations requires many high-dimensional
+eigendecompositions, which
+is computationally expensive. Additionally,
+$T_{1}$ values are known to fluctuate greatly
+with laboratory temperatures \cite{klimov2018fluctuations}.
+Interpolating $T_{1}$ from experimental data increases
+the fridge truth of the simulation.
+
+Furthermore, the probability of longitudinal 
+relaxation is dependent on the gate duration $t_{N}$.
+We allow the optimizer to tune the gate duration by
+augmenting the control vector $u_{k}$ with
+the time step between knot points $\Delta t_{k}$.
+Promoting $\Delta t_{k}$ to a decision variable, rather
+than the number of knot points $N$, preserves the
+Markovian decision structure of the trajectory
+optimization problem. To ensure numerical
+integration accuracy is maintained we add a bound
+constraint at each knot point
+$5\textrm{e-}3 \ \textrm{ns} \le
+\Delta t_{k} \le 2\textrm{e-}2 \ \textrm{ns}$.
+Note that this bound constraint is allowed to be
+broken for intermediate iterations of the optimization,
+so we use the absolute value of the time step
+$\lvert \Delta t_{k} \rvert$ to ensure that it is non-negative.
+
+(Results) We achieve a factor of 1.5 decrease in the probability
+of longitudinal relaxation from the analytic gates we benchmark
+against. We simulate the performance of the gates using the
+lindblad master equation (see Appendix X). We repeatedly apply the basis gates
+and measure the fidelity of the resulting state as a function
+of time as shown in Fig. 1.
+
+%% T4.1
+%% TODO: Update this table with spline T1 optimizations and calculations.
+%% TODO: Maybe best for appendix.
+%% TOOD: Maybe add some commentary about why we see more speedup
+%% for Z/2 and X/2 than Y/2. In particular I think this is because of the
+%% long idle times at the flux frustration point for Z/2 and X/2. It
+%% might be a more fair comparison to show gates that have better T1s
+%% and as good flux noise mitigation.
+\begin{table}
+  \begin{tabular}{c | c | c | c}
+    & Analytic  & QOC &\\
+    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & Speedup\\
+    \hline
+    Z/2 & 4.615 & 2.576 & 1.792\\
+    Y/2 & 2.826 & 1.888 & 1.496\\
+    X/2 & 9.562 & 5.080 & 1.882\\
+  \end{tabular}
+  \caption{Probability of longitudinal relaxation for each gate
+    evaluated at the gate's duration.}
+\end{table}
+
+%% F4.1
+%% TODO: Update images. Use new gates constructed with
+%% spline T1 and use smaller sim times.
+\begin{figure}[ht]
+  \begin{subfigure}{.33\textwidth}
+    \includegraphics[width=\linewidth]{assets/00045_spin16.png}
+    \label{fig:sub-first}
+  \end{subfigure}
+  \begin{subfigure}{.33\textwidth}
+    \includegraphics[width=\linewidth]{assets/00047_spin16.png}
+    \label{fig:sub-second}
+  \end{subfigure}
+  \begin{subfigure}{.33\textwidth}
+    \includegraphics[width=\linewidth]{assets/00049_spin16.png}
+    \label{fig:sub-third}
+  \end{subfigure}
+  \caption{Master equation simulation with $T_{1}$ dissipation for all basis gates.}
+  \label{fig:fig}
+\end{figure}
 
@@
 \section{Robustness to $T_{\phi}$-type Noise}
+(Strategy) System parameter deviations arise due to
+measurement error and noise injection. For example,
+we might measure $\omega_{q}$ inaccurately,
+which leads to a deviation from
+the the drive frequency $\omega_{d}$.
+Further, we may have some temperature dependent gain
+in the lines leading to the device, or 1/$f$ noise
+and $A(\Phi_{\textrm{ext}})$ comes out with some
+$\delta A$ on the flux axis.
+Both of these effects lead to dephasing,
+uncertain dynamics, and thus erroneous computation.
+
+These problems are typically attacked with dynamic decoupling
+sequence where an erroneous rotation is compensated for with
+one that cancels the $\delta$. However, we can do better
+than the first order variants of these pulses.
+
+We propose two methods for engineering robustness
+to system parameter deviations. The first we call the
+derivative method. We use the intuition that
+making the final state insensitive to changes
+in the system parameter is encoded in the derivative
+$\partial_{\omega}^{l} \ket{\psi_{N}}$.
+In particular, we find that going to the second
+derivative is good. The first derivative is already
+small, especially for the $\delta A$ time, because,
+near the end of optimization, the
+optimizer is unable to improve the final state by
+changing the flux bias.
+The dynamics for propgating the derivative of the state
+are found by differentiating the schroedinger equation
+dynamics with respect to the parameter of interest.
+The scheme boils down to propagating sequential
+derivatives of the parameter you want to be
+robust to $\partial_{\omega}^{l}$ and penalizing thier norm
+${\lvert \partial_{\omega}^{l} \lvert}^{2}$.
+
+The other scheme we call the sampling method. This method
+is well-known in the trajectory optimiztion. In this scheme
+we propgate an array of states whose dynamics differ
+in that the parameter of interest is altered slightly
+for each state. In the 3-point sampling method we
+propagate $\psi_{\pm}$, $\psi$ where
+$\omega_{\pm} = \omega \pm \delta \omega$. Here $\delta$ is
+typically taken at standard deviations of the parameter
+of interest.
+
+(Results) We compare the simplest schemes
+for the derivative and sampling methods to
+the CORPSE dynamic decoupling sequence.
+They perform X better by Y metric.
+
+We also find that longer gate durations allow for
+greater robustness.
+
+%% F5.1
+%% TODO: Update this and make sure it is for X/2.
+%% I'm not sure if we should include all gates Z/2, Y/2, X/2
+%% because I don't know what the dynamic decoupling analogy
+%% for Z/2 and Y/2 are. On the other hand we would
+%% want to include all of them to show that the benefit
+%% is not uniform across gates.
+\begin{figure}
+  %% \includegraphics[width=\linewidth]{assets/00005_spin12_hpsweep.png}
+  \caption{fidelity v.s. $\omega$ for the $X/2$ gate via
+    analytic, 3-sample, 2-derivative (maybe add 5-sample and 3-derivative)}
+\end{figure}
+
+%% F5.2
+%% TODO: Update this and make sure it is for X/2.
+%% Add inset to show high-performing variants at large time steps.
+\begin{figure}
+  %% \includegraphics[width=\linewidth]{assets/00006_spin12_hpsweep.png}
+  \caption{fidelity at $\omega \pm \sigma_{\omega}$ v.s. time for the
+    $X/2$ gate via analytic, 3-sample, 2-derivative
+    (maybe add 5-sample and 3-derivative)}
+\end{figure}
+
+%% F5.3
+%% TODO: Make this figure. My guess is we want a 5-sample, with 2 at \omega_{\pm}
+%% and 2 at A_{\pm} as well as 2-derivative of both parameters separately.
+\begin{figure}
+  %% \includegraphics[width=\linewidth]{assets/t2_temp.jpg}
+  \caption{Master equation simulation with $T_{2}$ dissipation
+    comparing the gates robust to $\delta \omega_{q}$ and $\delta A$ against
+    the analytic gate.}
+\end{figure}
+
+%% TODO: Fig: Full ME simulation for best T1 and T\phi pulses
+%% v.s. analytic
+
+%% SECTION 6
+\section{Discussion}
+We have proposed some schemes and they work well.
+
+\appendix
+%% APPENDIX 1
+\section{Ricatti Recursion}
+This will give the reader unfamiliar with trajectory
+optimization intuition for how the trajectory optimization
+update scheme works and why it is better than
+a more naive method.
+
+%% APPENDIX 2
+\section{Measurements}
+We measure $T_{1}$ using the standard experiment
+and $T_{2}$ using te Ramsey experiment. We fit with splines
+and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
+We measure $\omega_{q}$ and $\sigma_{\omega_{q}}$ using X method.
+
+%% APPENDIX 3
+\section{ME Simulation}
+We model dissipation using the lindblad master
+equation and standard collapse operators for
+longitudinal relaxation and dephasing
+of a two-level system.
+
+%% APPENDIX 4
+\section{Derivative Method}
+\begin{equation}
+  \begin{aligned}
+    \partial_{\omega}^{2} \partial_{t} \psi
+    = \frac{-i}{h} [&2 \cdot \frac{\sigma_{z}}{2} \partial_{\omega} \psi\\
+      &+ (\omega \frac{\sigma_{z}}{2} + A \frac{\sigma_{x}}{2}) \partial_{\omega}^{2} \psi]
+  \end{aligned}
+\end{equation}
 

### 2020-07-17 paper/paper.tex
progress on paper

@@
 (Existing work) The leading model of universal
-quantum computation is gate-based. Analytic methods
-have been demonstrated to construct gates.
-Some focus on solving the TDSE analytically
-\cite{}Analytic methods have
-been demonstrated to construct gates
-\cite{zhang2020universal, huang2020engineering,
-  han2020experimental, xu2020nonadiabatic}.
-Most methods focus on optimizing a few aspects of the gate, but not all
-experimentally relevant aspects. There is a growing literature
-on numerical
-techniques to construct gates \cite{leung2017speedup,
-  goerz2019krotov, doria2011optimal, abdelhafez2019gradient,
-  machnes2015gradient}.
-%% TODO: Cite CRAB, GOAT, KROTOV, etc.
-These methods formulate the quantum optimal
-control problem as unconstrained.
-They typically rely on zeroth-order or first-order optimizers.
-They are not
-sofisticated enough to handle all of the relevant
-constraints simultaneously.
-Although some analytic techniques exist to design pulses robust
-to decoherence, no numerical techniques have been presented to
-design pulses robust to decoherence (as far as the author is aware).
+quantum computation is gate-based. The Quantum Optimal Control (QOC)
+literature has saught to answer the question "What is the best way to construct a quantum gate?".
+Analytic methods
+have been demonstrated to construct quantum gates
+\cite{zhang2020universal, huang2020engineering, han2020experimental,
+  xu2020nonadiabatic, carlini2005quantum}.
+Most methods rely on solving the time-dependent
+Schroedinger equation analytically or deriving equations of motion
+from a suitable Lagrangian.
+
+Some analytical methods
+have been presented to address realistic experimental constraints.
+Considering the dynamical and geometric phases of state
+evolution has led to methods for achieving
+robustness to pure dephasing in quantum systems
+\cite{xu2020nonadiabitc, han2020experimental, merrill2014progress}.
+Additionally, suitable choices of bases have been demonstrated that
+allow the experimentalist to infer tradeoffs in longitudinal
+and pure dephasing coherence times \cite{huang2020engineering}.
+To date, however, no analytical method has been presented
+to construct a gate while satisfying all relevant experimental
+constraints. Furthermore, most analytic methods have only
+been demonstrated in the two-level approximation and become
+cumbersome to solve for larger Hilbert space sizes.
+
+Numerical quantum control techniques have developed in parallel
+with the analytic frameworks. Numerical quantum control
+techniques integrate the time-dependent Schroedinger
+equation and use sensitivity analysis to optimize target metrics
+of the quantum gate, such as fidelity
+\cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
+  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
+These techniques typically employ zeroth- or first-order
+optimizers and make no garauntees of optimality.
+Further, they formulate the quantum control
+problem as unconstrained or use
+projective gradient methods \cite{machnes2015gradient}.
+The latter approach relies on constructing
+bases for constraint spaces--which is not always feasible--and
+may hinder convergence.
 
@@
 optimal control problem as a constrained optimization problem.
-We study the quantum optimal control problem on the fluxonium.
+The trajectory optimization framework allows us to handle arbitrary
+constraints on the evolution of the quantum state. Additionally,
+we introduce methods for constructing gates robust to longitudinal
+relaxation and pure dephasing, the dominant barrier to experimentally
+realizing practical quantum computing.
+
+We study the quantum optimal control problem on the fluxonium qubit.
 We outline experimentally realistic constraints and map them to
 the trajectory optimization framework. For the device we study
-we achieve a 2x increase in $T_{1}$ times.
-We present two methods for achieving robustness to system parameter deviations,
-and compare to existing dynamic decoupling methods. We find that 
-our methods beat dynamic decoupling and mitigate dephasing by order X.
+we achieve a 1.5x increase in $T_{1}$ times.
+We present two methods for achieving robustness to pure dephasing,
+and compare them to existing dynamic decoupling methods. We find that 
+our methods mitigate dephasing by order X over dynamic decoupling.
 
@@
 making the optimization $T_{1}$ aware. Finally, we present
-some methods for engineering robustness to decoherence and
+some methods for engineering robustness to pure dephasing and
 compare them to existing techniques.
@@
 \section{QOC + AL-iLQR}
-(QOC Problem Statement) I have some initial
-configuration and I want to reach some target configuration,
-either single- or multi-state transfer, subject to the dynamics
+(QOC Problem Statement) Here we introduce the notation
+we will use throughout the paper,
+review the quantum optimal control problem statement,
+and introduce the trajectory optimization framework.
+Quantum optimal control concerns the evolution of
+a state $\ket{\psi(t)}$ governed by the time-dependent
+Schroedinger equation
 \begin{align}
-  \partial_{t} \ket{\psi(t)} &= -\frac{i}{\hbar} H(t) \ket{\psi(t)}
+  \partial_{t} \ket{\psi(t)} &= -\frac{i}{\hbar} H(\textbf{u}(t), t) \ket{\psi(t)}
 \end{align}
-Common numerical techniques include approximating the analytic unitary propagator solution,
-or employing explicit/implicit Runge-Kutta methods of the form
+The evolution is sometimes cast with $\rho(t) = \ket{\psi(t)}\bra{\psi(t)}$
+and the Lindblad master equation to model the decoherence of the
+state explicitly. The Hamiltonian
+depends arbitrarily on the controls $\textbf{u}(t)$.
+
+Numerical quantum optimal control techniques make
+the problem tractable by discretizing the problem into $N$
+time steps. Typical integration techniques rely on approximating
+the analytical solution to the time-dependent Schroedinger equation
+or using explicit integration methods, e.g. Runge-Kutta, of the form
 \begin{align}
-  \ket{\psi^{(k + 1)}} &= \ket{\psi^{(k)}} + \partial_{t} \ket{\psi^{(k)}} \cdot dt
+  \ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \partial_{t} \ket{\psi_{k}} \cdot \Delta t_{k}
 \end{align}
-The interesting part is that your hamiltonian $H(t)$ has a time-dependent
-control parameter $u(t)$ that the experimentalist gets to control, e.g. flux threading
-a superconducting junction.
-\begin{align}
-  H(t) &= H_{0} + \sum_{i = 1}^{m} a_{m}(u_{m}(t)) H_{m}
-\end{align}
-Most often $a_{m}(u) = u$ but any arbitrary dependence is allowed.
-The technique of adjusting $\ket{\psi_{N}}$ based on $\textbf{u}(t)$
-is the goal of the quantum optimal control optimizer. This domain is
-know as sensitivity analysis. GRAPE does
-\begin{align}
-  \min_{U:N-1 \times m} &1 - {\lvert \braket{\psi_{N}(U) \lvert \psi_{t}} \lvert}^{2}
-\end{align}
-subject to the gradient descent update procedure
-\begin{align}
-  U^{'} &= U - \alpha \nabla_{U} (1 - {\lvert \braket{\psi_{N}(U) \lvert \psi_{t}} \lvert}^{2})
-\end{align}
-In general they add more cost functions than target fidelity.
-%% TOOD: Other update procedures might differ, e.g. Krotov does some sequential updates
-%% and Houck, Google use black box optimizers.
 
-(AL-iLQR Problem Statement) Trajectory optimization gives us gaurantees
+Quantum optimal control seeks the control
+parameters that minimize a functional $J(u, \ket{\psi_{N}(u)})$.
+In the simplest case the functional
+$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u)} \rvert}^{2}$
+which is the infidelity between the inital state evolved
+to the final time step ($\ket{\psi_{N}(u)}$)
+and the target state ($\ket{\psi_{f}}$). In general
+$J$ is a linear combinaion of cost functions on $\ket{\psi_{N}(u)}$, e.g.
+forbidden-state occupation, as well as
+cost functions on $u$, e.g. the norm of the control amplitudes
+\cite{leung2017speedup}.
+
+(AL-iLQR Problem Statement) The trajectory optimization
+literature solves a more general class of non-linear programs that resemble
+the quantum optimal control problem. The quantum optimal control
+problem is a specific case of the linear quadratic regulator (LQR).
+LQR is so called because the dynamics are linear in the state and
+the cost function is quadratic in the state. In the LQR formulation
+the same functional is evaluated at each time step
+\begin{equation}
+  \begin{aligned}
+    J_{\textrm{LQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
+    &+ \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
+  \end{aligned}
+\end{equation}
+where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the current
+and final state and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
+
+The advantage of the LQR formulation
+is that there exists a dynamic programming algorithm to compute the
+optimal update to the controls $u_{k}$ at each time step.
+This algorithm employs the Ricatti recursion (see Appendix).
+
+In order to incorporate constraints into the problem we employ
+the augmented lagrangian method. This method adaptively
+updates the penalty of constraints
+%% LEFT OFF HERE
+
+Trajectory optimization gives us gaurantees
 about our updates via Ricatti recursion
@@
 %% SECTION 3
-\section{Our Problem}
+\section{QOC on the Fluxonium}
 (Fluxonium Device) In the two-level
@@
   \begin{subfigure}{.33\textwidth}
-    \includegraphics[width=\linewidth]{assets/00045_spin16.png}
+    %% \includegraphics[width=\linewidth]{assets/00045_spin16.png}
     \label{fig:sub-first}
@@
   \begin{subfigure}{.33\textwidth}
-    \includegraphics[width=\linewidth]{assets/00047_spin16.png}
+    %% \includegraphics[width=\linewidth]{assets/00047_spin16.png}
     \label{fig:sub-second}
@@
   \begin{subfigure}{.33\textwidth}
-    \includegraphics[width=\linewidth]{assets/00049_spin16.png}
+    %% \includegraphics[width=\linewidth]{assets/00049_spin16.png}
     \label{fig:sub-third}

### 2020-07-20 paper/paper.tex
second pass through section 3

@@
 problem as unconstrained or use
-projective gradient methods \cite{machnes2015gradient}.
+projective gradient methods to enforce constraints
+\cite{machnes2015gradient}.
 The latter approach relies on constructing
@@
 Quantum optimal control concerns the evolution of
-a state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation
+a quantum state $\ket{\psi(t)}$ governed by the time-dependent
+Schroedinger equation (TDSE)
 \begin{align}
-  \partial_{t} \ket{\psi(t)} &= -\frac{i}{\hbar} H(\textbf{u}(t), t) \ket{\psi(t)}
+  \frac{d}{dt} \ket{\psi(t)} &= -\frac{i}{\hbar} H(\textbf{u}(t), t) \ket{\psi(t)}
 \end{align}
-The evolution is sometimes cast with $\rho(t) = \ket{\psi(t)}\bra{\psi(t)}$
-and the Lindblad master equation to model the decoherence of the
-state explicitly. The Hamiltonian
-depends arbitrarily on the controls $\textbf{u}(t)$.
+The evolution is sometimes cast with the evolution
+of a density matrix under the Lindblad master equation to
+model the decoherence of the state explicitly. The Hamiltonian
+depends arbitrarily on the controls ($\textbf{u}(t)$). The controls
+are so called because they are the means the experimentalist has to
+act on the system.
 
@@
 the problem tractable by discretizing the problem into $N$
-time steps. Typical integration techniques rely on approximating
-the analytical solution to the time-dependent Schroedinger equation
-or using explicit integration methods, e.g. Runge-Kutta, of the form
+time steps. Typical integration techniques for the TDSE include
+approximating unitary propagators as well as explicit integration methods,
+such as Runge-Kutta, of the form
+\label{step-approx}
 \begin{align}
-  \ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \partial_{t} \ket{\psi_{k}} \cdot \Delta t_{k}
+  \ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}
 \end{align}
@@
 Quantum optimal control seeks the control
-parameters that minimize a functional $J(u, \ket{\psi_{N}(u)})$.
-In the simplest case the functional
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u)} \rvert}^{2}$
-which is the infidelity between the inital state evolved
-to the final time step ($\ket{\psi_{N}(u)}$)
+parameters that minimize a functional $J(\textbf{u}(t))$.
+In the simplest case the functional is
+$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(\textbf{u}(t))} \rvert}^{2}$
+the infidelity between the inital state evolved
+to the final time step ($\ket{\psi_{N}(\textbf{u}(t))}$)
 and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on $\ket{\psi_{N}(u)}$, e.g.
+$J$ is a linear combinaion of cost functions on the state, e.g.
 forbidden-state occupation, as well as
-cost functions on $u$, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}.
+cost functions on the controls, e.g. the norm of the control amplitudes
+\cite{leung2017speedup}. Typical quantum optimal control
+algorithms employ automatic differentiation
+to compute first order information for the functional ($\nabla_{u} J(u)$).
+They employ a first-order optimizer to minimize $J$ with respect to $u$.
 
@@
 LQR is so called because the dynamics are linear in the state and
-the cost function is quadratic in the state. In the LQR formulation
+the functional is quadratic in the state. In the LQR formulation
 the same functional is evaluated at each time step
 \begin{equation}
-  \begin{aligned}
-    J_{\textrm{LQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-    &+ \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-  \end{aligned}
+  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
+  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
 \end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the current
-and final state and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
+where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
+at time step $k$ and final state, $u_{k}$ are the controls,
+and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
+The state is propagated using a dynamics function similar to \ref{step-approx}
+$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
+In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
+and $f$ encodes the TDSE dynamics.
 
@@
 is that there exists a dynamic programming algorithm to compute the
-optimal update to the controls $u_{k}$ at each time step.
-This algorithm employs the Ricatti recursion (see Appendix).
+optimal update to the controls ($u_{k}$) which minimizes the functional
+($J_{\textrm{iLQR}, k}$) for each time step. This algorithm proceeds by deriving a
+recurrence relation between time steps $k$ and $k + 1$ for the optimal
+feedback law--known as the Ricatti recursion (see Appendix). The
+iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
+and applies the Ricatti recursion to all time steps on multiple
+executions.
 
-In order to incorporate constraints into the problem we employ
-the augmented lagrangian method. This method adaptively
-updates the penalty of constraints
-%% LEFT OFF HERE
-
-Trajectory optimization gives us gaurantees
-about our updates via Ricatti recursion
-and allows us to put constraints on our cost functions.
-%% This is e.q. 2 from ALTRO paper
+In order to incorporate constraints we employ
+the augmented Lagrangian method. Constraints are contributions
+to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
+zero-valued when the constraint is satisfied. The AL-iLQR
+method updates the penalty multiplier between
+iLQR executions. The penalty multiplier associated with the
+constraint functional estimates the Lagrange multiplier of
+the constraint. The functional now takes the form
 \begin{equation}
-\begin{aligned}[b]
-  \min_{U:N-1 \times m} &\mathcal{L}_{N}(x_{N}, \lambda_{N}, \mu_{N})\\
-  &+ \sum_{k = 1}^{N - 1} \mathcal{L}_{k}(x_{k}, u_{k}, \lambda_{k}, \mu_{k})\\
+  \begin{aligned}
+    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
+    &+ J_{\textrm{iLQR}}
   \end{aligned}
 \end{equation}
-\begin{equation}
-\begin{aligned}[b]
-  \mathcal{L}_{k}(x_{k}, &u_{k}, \lambda_{k}, \mu_{k}) =\\
-  &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-  &+ (x_{k} - x_{f})^{T} Q (x_{k} - x_{f}) + u_{k}^{T} R u_{k}\\
-\end{aligned}
-\end{equation}
-The important point is that there is an update step (e.q. 17 from ALTRO paper) where we send $\lambda \rightarrow \infty$
-and get all of the nice convergence properties. The weights are adjusted dynamically between iterations until
-all of our constraints are satisfied. The Markovian decision structure of the problem allows
-us to apply differentiable dynamic programming to garauntee that the update for each control $U_{k}$ is
-opimal, as apposed to the greedy updates of first-order optimizers like the naive gradient descent.
-
-%% TODO: (?) Figure comparing update procedures
+$\lambda_{k}$ is a Lagrange multiplier. $I_{\mu_{k}}$ is a penalty matrix
+with $\mu_{k}$ along the diagonal. $i$ indicates the $i$-th constraint functional.
+$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
+\begin{align}
+  \lambda_{k_{i}}^{'} &= \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*})\\
+  \mu_{k_{i}}^{'} &= \phi \mu_{k_{i}}
+\end{align}
+where $x^{*}, u^{*}$ are the optimal state and controls from the iLQR execution
+and $\phi$ is a hyperparameter. With this updated form of the cost
+functional there still exists a recurrence relation to calculate the optimal control
+updates, see \cite{howell2019altro}.
 
@@
 \section{QOC on the Fluxonium}
-(Fluxonium Device) In the two-level
-approximation we have
+(Fluxonium Device) In the following we study
+the quantum optimal control problem on the fluxonium qubit.
+In the two-level
+approximation the system Hamiltonian takes the form
+\label{hamiltonian}
 \begin{align}
-  H/h &= \omega_{q} \frac{\sigma_{z}}{2} + A(\Phi_{ext}) \frac{\sigma_{x}}{2}
+  H/h &= \omega_{q} \frac{\sigma_{z}}{2} + u(t) \frac{\sigma_{x}}{2}
 \end{align}
-This approximation is good up to the avoided crossing at $\Phi_{\textrm{ext}} = 0.35 \Phi_{0}$. We get $A$
-by converting via $\bra{g} \hat{\phi} \ket{e}$.
-We use the gates $X/2, Y/2, Z/2$. $X/2, Y/2$ are universal
-up to an arbitrary Z rotation.
+$h$ is Planck's constant. The control parameter $u$ is experimentally
+realized by modulating the flux 
+threading the superconducting junction $\Phi_{\textrm{ext}}$.
+The conversion is defined by
+$u = 4 \pi \braket{g \lvert \hat{\phi} \rvert e} E_{L} \delta \Phi_{\textrm{ext}} / \Phi_{0}$
+where $\delta \Phi_{\textrm{ext}} = \Phi_{\textrm{ext}} - 0.5 \Phi_{0}$ is the flux
+offset from the flux frustration point.
+
+In the following we compare the gates found with our methods
+to the $X/2, Y/2, Z/2$ gates reported in \cite{zhang2020universal}
+for the same device. Note that $Y/2$ and arbitrary $Z$ rotations
+are sufficient for universal computation.
 %% TODO: list device parameters
 
-(Constraints) We want constraints on our pulses. We want pulses start and end at zero
-for concatenation. We want pulses to have zero net flux to mitigate
-hysteresis in flux bias lines.
+(Constraints) We now outline the constraints that we require each gate
+to obey.
+We require $u(t = 0) = u(t = t_{N}) = 0$.
+This constraint ensures gates may be concatenated arbitrarily without
+inducing AWG ringing due to high-frequency transitions.
+Furthermore, we require $\int_{0}^{t_{N}} u(t) dt = 0$. This
+constraint ensures the pulse has zero net flux, mitigating
+the hysteresis ubiquitous in flux bias lines.
 %% TODO: See ref. 28 in Helin paper
-We want the amplitude to be constrained
-$\delta \Phi_{\textrm{ext}} \sim 0.06 \Phi_{0}$
-so the two-level approximation stays valid.
-We want our state to obey normalization conditions,
-mitigating numerical error in simulaiton.
-%% TODO: mention pulse smoothing via optimizing over 2nd derivative
-%% of pulses
+We require $-300 \textrm{MHz} \le u(t) \le 300 \textrm{MHz}$
+to ensure the two-level approximation \ref{hamiltonian}
+remains valid. Additionally, we require that each gate achieves
+the desired state transition $\braket{\psi_{f} \lvert \psi_{N}} = 1$.
+In addition to these constraints we penalize the norm
+of the first and second derivatives of the control
+to mitigate AWG ringing.
+
+The equality and inequality constraints on $u$ are handled
+with a bound constraint. $\int_{0}^{t_{k}} u(t) dt$ is added
+to the augmented state vector $x_{k}$ by incorporating the
+integration of $u_{k}$ into the dynamics function $f$.
+Both the zero net flux and target quantum state constraint
+are then handled by ensuring the augmented final state is
+reached $x_{N} - x_{f} = 0$. The optimization is performed
+over the second derivative of the controls $\frac{d^{2}}{dt^{2}} u(t)$
+so that the second derivative, first derivative, zeroth derivative, and
+integral of the controls may be penalized.
 %% TODO: probably want to list equations for the constraints
@@
 ($T_{1}$ and $T_{\phi}$ noise)
+In addition to optimizing the cost functional to achieve a gate
+that obeys experimental constraints and has a high simulated fidelity,
+we also want to make the gate robust to noise that affects the experimental
+fidelity of the gate. Decoherence of the quantum state due to external noise
+is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
+They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
+(see Appendix).
+The main contributions to longitudinal relaxation in our
+device are dielectric loss in the capacitor, resistive loss in the inductor,
+and Purcell loss. The main contributions to pure dephasing in our
+device are $1/f$ flux noise and decay via charge and flux coupling
+to the control lines.
+
 Dissipation to the thermal bath via longitudinal
-relaxation ($T_{1}$) is an irreversible process
+relaxation is an irreversible process
 that results in information loss.
-Pure dephasing ($T_{\phi}$) is a reversible process.
-Dielectric loss, etc. contribute
-to $T_{1}$. $1/f$ flux noise, etc. contribute to $T_{\phi}$
-There is a tradeoff between $T_{1}$ and $T_{\phi}$. In the case of white
+Converesely, pure dephasing is a reversible process.
+There is a tradeoff between the two decoherence processes. In the case of white
 noise we have that the sum of the noise weights $W_{1}$ and $W_{\phi}$
 is constant \cite{huang2020engineering}.
-In our case $T_{\phi}$ is saturated at the flux
-frustration point $\Phi_{\textrm{ext}} = 0.5 \Phi_{0}$
-$T_{\phi}$ decreases as we move away from the flux
+Our device achieves its best pure dephasing
+protection at the flux frustration point
+$T_{2e}(\Phi_{\textrm{ext}} = 0.5 \Phi_{0}) \sim 300 \mu\textrm{s}$
+where the qubit frequency is first-order insensitive to changes in flux.
+It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
 frustration point. Conversely, $T_{1}$ is at a minimum
@@
 of the decay processes and the tradeoff, we choose
-to optimize the flux bias to have high time-averaged
-$T_{1}$ times and employ some techniques to mitigate
-$T_{\phi}$ dephasing.
+to maximize the longitudinal relaxation time ($T_{1}$)
+over the gate duration
+and employ robust control techniques to mitigate
+pure dephasing.
 %% TODO: See fig. 3 in Helin paper
 
+%% LOH
 %% SECTION 4
@@
 %% APPENDIX 2
-\section{Measurements}
+\section{Experiment}
 We measure $T_{1}$ using the standard experiment

### 2020-07-21 paper/paper.tex
progress

@@
   nobibnotes,
-  prl,
+  %% prl,
   twocolumn,
@@
 robustness to pure dephasing in quantum systems
-\cite{xu2020nonadiabitc, han2020experimental, merrill2014progress}.
+\cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
 Additionally, suitable choices of bases have been demonstrated that
@@
 we introduce methods for constructing gates robust to longitudinal
-relaxation and pure dephasing, the dominant barrier to experimentally
+relaxation and pure dephasing, the dominant barriers to experimentally
 realizing practical quantum computing.
@@
 the trajectory optimization framework. For the device we study
-we achieve a 1.5x increase in $T_{1}$ times.
+we achieve a 1.5x decrease in longitudinal relaxation probability.
 We present two methods for achieving robustness to pure dephasing,
@@
 Schroedinger equation (TDSE)
+\label{eq:tdse}
 \begin{align}
-  \frac{d}{dt} \ket{\psi(t)} &= -\frac{i}{\hbar} H(\textbf{u}(t), t) \ket{\psi(t)}
+  \frac{d}{dt} \ket{\psi(t)} &= -\frac{i}{\hbar} H(u(t), t) \ket{\psi(t)}
 \end{align}
@@
 model the decoherence of the state explicitly. The Hamiltonian
-depends arbitrarily on the controls ($\textbf{u}(t)$). The controls
-are so called because they are the means the experimentalist has to
+has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
+The controls are so called because they are the means the experimentalist has to
 act on the system.
@@
 such as Runge-Kutta, of the form
-\label{step-approx}
+\label{eq:step-approx}
 \begin{align}
@@
 Quantum optimal control seeks the control
-parameters that minimize a functional $J(\textbf{u}(t))$.
+parameters that minimize a functional $J(u(t))$.
 In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(\textbf{u}(t))} \rvert}^{2}$
+$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
 the infidelity between the inital state evolved
-to the final time step ($\ket{\psi_{N}(\textbf{u}(t))}$)
+to the final time step ($\ket{\psi_{N}(u(t))}$)
 and the target state ($\ket{\psi_{f}}$). In general
@@
 and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function similar to \ref{step-approx}
+The state is propagated using a dynamics function similar to
+equation \ref{eq:step-approx}
 $x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
@@
 approximation the system Hamiltonian takes the form
-\label{hamiltonian}
+\label{eq:hamiltonian}
 \begin{align}
-  H/h &= \omega_{q} \frac{\sigma_{z}}{2} + u(t) \frac{\sigma_{x}}{2}
+  H/h &= \omega_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
 \end{align}
-$h$ is Planck's constant. The control parameter $u$ is experimentally
+$h$ is Planck's constant. The flux amplitude $a$ is experimentally
 realized by modulating the flux 
-threading the superconducting junction $\Phi_{\textrm{ext}}$.
-The conversion is defined by
-$u = 4 \pi \braket{g \lvert \hat{\phi} \rvert e} E_{L} \delta \Phi_{\textrm{ext}} / \Phi_{0}$
+threading the deivce $\Phi_{\textrm{ext}}$.
+The flux amplitude is obtained from the external flux by
+$a = 4 \pi \braket{g \lvert \hat{\phi} \rvert e} \rvert_{0.5 \Phi_{0}} E_{L}
+\delta \Phi_{\textrm{ext}} / (h \Phi_{0})$
 where $\delta \Phi_{\textrm{ext}} = \Phi_{\textrm{ext}} - 0.5 \Phi_{0}$ is the flux
@@
 to obey.
-We require $u(t = 0) = u(t = t_{N}) = 0$.
+We require $a(t = 0) = a(t = t_{N}) = 0$.
 This constraint ensures gates may be concatenated arbitrarily without
 inducing AWG ringing due to high-frequency transitions.
-Furthermore, we require $\int_{0}^{t_{N}} u(t) dt = 0$. This
+Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
 constraint ensures the pulse has zero net flux, mitigating
@@
 %% TODO: See ref. 28 in Helin paper
-We require $-300 \textrm{MHz} \le u(t) \le 300 \textrm{MHz}$
-to ensure the two-level approximation \ref{hamiltonian}
+We require $-300 \textrm{MHz} \le a(t) \le 300 \textrm{MHz}$
+to ensure the two-level approximation \ref{eq:hamiltonian}
 remains valid. Additionally, we require that each gate achieves
@@
 
-The equality and inequality constraints on $u$ are handled
-with a bound constraint. $\int_{0}^{t_{k}} u(t) dt$ is added
-to the augmented state vector $x_{k}$ by incorporating the
-integration of $u_{k}$ into the dynamics function $f$.
+The optimization is performed
+over the second derivative of the flux modulation amplitude
+($\frac{d^{2}}{dt^{2}} a(t)$) so that the first derivative
+($\frac{d}{dt} a(t)$), proportional ($a(t)$), and integral ($\int a(t)$)
+are tracked in the augmented state vector $x$ and obtained by
+integration in the dynamics function $f$.
 Both the zero net flux and target quantum state constraint
 are then handled by ensuring the augmented final state is
-reached $x_{N} - x_{f} = 0$. The optimization is performed
-over the second derivative of the controls $\frac{d^{2}}{dt^{2}} u(t)$
-so that the second derivative, first derivative, zeroth derivative, and
-integral of the controls may be penalized.
+reached $x_{N} - x_{f} = 0$.
+The equality and inequality constraints on $a(t)$ are handled
+with a bound constraint.
 %% TODO: probably want to list equations for the constraints
@@
 we also want to make the gate robust to noise that affects the experimental
-fidelity of the gate. Decoherence of the quantum state due to external noise
+gate fidelity. Decoherence of the quantum state due to external noise
 is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
@@
 frustration point. Conversely, $T_{1}$ is at a minimum
-around the flux frustration point $T_{1}(0.5 \Phi_{0}) = 0.315$ms,
+at the flux frustration point $T_{1}(0.5 \Phi_{0}) = 0.315$ms,
 and increases away from the flux frustration point
@@
 
-%% LOH
+%% F4.1
+%% TODO: Update images. Use new gates constructed with
+%% spline T1 and use smaller sim times.
+\begin{figure*}[!t]
+  \begin{subfigure}{.3\linewidth}
+    \includegraphics[width=\linewidth]{assets/00045_spin16.png}
+    \label{fig:sub-first}
+  \end{subfigure}
+  \begin{subfigure}{.3\linewidth}
+    \includegraphics[width=\linewidth]{assets/00047_spin16.png}
+    \label{fig:sub-second}
+  \end{subfigure}
+  \begin{subfigure}{.3\linewidth}
+    \includegraphics[width=\linewidth]{assets/00049_spin16.png}
+    \label{fig:sub-third}
+  \end{subfigure}
+  \caption{Master equation simulation with $T_{1}$ dissipation for all basis gates.}
+  \label{fig:fig}
+\end{figure*}
+
 %% SECTION 4
@@
 (Strategy) We seek to minimize the probability
-that the qubit relaxes longitudinally.
-To this end we augment the state vector $x_{k}$ with the
+that the qubit decays as a result of longitudinal
+relaxation. To this end we augment the state vector ($x_{k}$) with the
 longitudinal relaxation probability
-\begin{align}
-  P_{1}(t_{k}) = \int_{0}^{t_{k}}
-  \gamma_{1}(\Phi_{\textrm{ext}}(t)) dt
-\end{align}
+\begin{equation}
+  P_{1}(t_{k}) = \int_{0}^{t_{k}} \gamma_{1}(a(t)) dt
+\end{equation}
 where $\gamma_{1} = T_{1}^{-1}$. Setting the target
@@
 
-The value $T_{1}(\Phi_{\textrm{ext}}(t_{k}))$
+$\gamma_{1}(a_{k})$
 is obtained at each knot point by evaluating
@@
 $\{(\Phi_{\textrm{ext}}, T_{1})\}$.
+$\Phi_{\textrm{ext}}$ is a function of $a_{k}$
+with the inverse of the relation given in section 3.
 Calculating $T_{1}$ directly from theoretical
@@
 against. We simulate the performance of the gates using the
-lindblad master equation (see Appendix X). We repeatedly apply the basis gates
+Lindblad master equation (see Appendix). We repeatedly apply the basis gates
 and measure the fidelity of the resulting state as a function
@@
 %% and as good flux noise mitigation.
-\begin{table}
-  \begin{tabular}{c | c | c | c}
-    & Analytic  & QOC &\\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & Speedup\\
-    \hline
-    Z/2 & 4.615 & 2.576 & 1.792\\
-    Y/2 & 2.826 & 1.888 & 1.496\\
-    X/2 & 9.562 & 5.080 & 1.882\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% F4.1
-%% TODO: Update images. Use new gates constructed with
-%% spline T1 and use smaller sim times.
-\begin{figure}[ht]
-  \begin{subfigure}{.33\textwidth}
-    %% \includegraphics[width=\linewidth]{assets/00045_spin16.png}
-    \label{fig:sub-first}
-  \end{subfigure}
-  \begin{subfigure}{.33\textwidth}
-    %% \includegraphics[width=\linewidth]{assets/00047_spin16.png}
-    \label{fig:sub-second}
-  \end{subfigure}
-  \begin{subfigure}{.33\textwidth}
-    %% \includegraphics[width=\linewidth]{assets/00049_spin16.png}
-    \label{fig:sub-third}
-  \end{subfigure}
-  \caption{Master equation simulation with $T_{1}$ dissipation for all basis gates.}
-  \label{fig:fig}
-\end{figure}
+%% \begin{table}
+%%   \begin{tabular}{c | c | c | c}
+%%     & Analytic  & QOC &\\
+%%     Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & Speedup\\
+%%     \hline
+%%     Z/2 & 4.615 & 2.576 & 1.792\\
+%%     Y/2 & 2.826 & 1.888 & 1.496\\
+%%     X/2 & 9.562 & 5.080 & 1.882\\
+%%   \end{tabular}
+%%   \caption{Probability of longitudinal relaxation for each gate
+%%     evaluated at the gate's duration.}
+%% \end{table}
 
@@
 \section{Robustness to $T_{\phi}$-type Noise}
-(Strategy) System parameter deviations arise due to
-measurement error and noise injection. For example,
-we might measure $\omega_{q}$ inaccurately,
-which leads to a deviation from
-the the drive frequency $\omega_{d}$.
-Further, we may have some temperature dependent gain
-in the lines leading to the device, or 1/$f$ noise
-and $A(\Phi_{\textrm{ext}})$ comes out with some
-$\delta A$ on the flux axis.
-Both of these effects lead to dephasing,
-uncertain dynamics, and thus erroneous computation.
+(Strategy) For the fluxonium, pure dephasing acts
+along the flux axis. Low frequency 1/$f$ noise,
+Johnson-Nyquist current noise present in the flux bias
+lines, and temperature dependent gain fluctuations
+may cause the flux amplitude to deviate from its
+desired value by an amount $\delta a$. Additionally,
+residual calibration errors and fluctuations may cause
+the qubit frequency to deviate from its measured value
+by an amount $\delta \omega_{q}$.
 
-These problems are typically attacked with dynamic decoupling
-sequence where an erroneous rotation is compensated for with
-one that cancels the $\delta$. However, we can do better
-than the first order variants of these pulses.
+System parameter
+deviations are typically addressed with dynamic decoupling
+sequences \cite{merrill2014progress},
+the DRAG scheme \cite{krantz2019quantum}, or
+geometric phase considerations
+\cite{xu2020nonadiabatic} \cite{han2020experimental}.
+Dynamic decoupling sequences compose rotations on the
+Bloch sphere so that erroneous rotations arising due
+to the parameter deviation are cancelled.
+The latter two methods are inflexible. We draw on the
+robust control literature to demonstrate
+numerical techniques for engineering robustness
+to parameter deviations.
 
@@
 to system parameter deviations. The first we call the
-derivative method. We use the intuition that
-making the final state insensitive to changes
+derivative method. We draw on the intuition that
+making the state evolution insensitive to changes
 in the system parameter is encoded in the derivative
-$\partial_{\omega}^{l} \ket{\psi_{N}}$.
-In particular, we find that going to the second
-derivative is good. The first derivative is already
-small, especially for the $\delta A$ time, because,
-near the end of optimization, the
-optimizer is unable to improve the final state by
-changing the flux bias.
-The dynamics for propgating the derivative of the state
-are found by differentiating the schroedinger equation
-dynamics with respect to the parameter of interest.
-The scheme boils down to propagating sequential
-derivatives of the parameter you want to be
-robust to $\partial_{\omega}^{l}$ and penalizing thier norm
-${\lvert \partial_{\omega}^{l} \lvert}^{2}$.
+$\partial_{\omega_{q}}^{l} \ket{\psi}$.
+The $l$ state derviatives are added to the augmented
+state vector. Their dynamics are found by differentiating
+the TDSE dynamics with respect to the parameter of interest
+\ref{eq:tdse}. Setting the target state derivatives
+to zero vectors results in quadratic costs at each
+knot point
+${\lvert \partial_{\omega_{q}}^{l} \ket{\psi_{k}} \rvert}^{2}$.
 
-The other scheme we call the sampling method. This method
-is well-known in the trajectory optimiztion. In this scheme
-we propgate an array of states whose dynamics differ
-in that the parameter of interest is altered slightly
-for each state. In the 3-point sampling method we
-propagate $\psi_{\pm}$, $\psi$ where
-$\omega_{\pm} = \omega \pm \delta \omega$. Here $\delta$ is
-typically taken at standard deviations of the parameter
-of interest.
+The second scheme we analyze is the sampling method,
+which is well studied in the robust control literature
+\cite{manchester2016derivative} \cite{tronarp2016sigma}.
+We augment the state vector with additional
+states that follow deviant dynamics. The parameters of interest
+are altered by a small deviation, typically chosen
+to represent the distribution of the parameter.
+For example, in the 3-point sampling method we
+propagate $\ket{\psi_{\pm, I}}$ where
+$\omega_{q}^{\pm} = \omega_{q} \pm \sigma_{\omega_{q}}$,
+$\omega_{q}^{I} = \omega_{q}$.
+The difference between the final states and the target
+state is penalized, resulting in quadratic costs at
+each knot point
+${\lvert \ket{\psi_{\pm, I}} - \ket{\psi_{f}} \rvert}^{2}$
+
+%% LOH
 
@@
 %% is not uniform across gates.
+%% (maybe add 5-sample and 3-derivative)
 \begin{figure}
@@
   \caption{fidelity v.s. $\omega$ for the $X/2$ gate via
-    analytic, 3-sample, 2-derivative (maybe add 5-sample and 3-derivative)}
+    analytic, 3-sample, 2-derivative}
 \end{figure}
@@
   \caption{fidelity at $\omega \pm \sigma_{\omega}$ v.s. time for the
-    $X/2$ gate via analytic, 3-sample, 2-derivative
-    (maybe add 5-sample and 3-derivative)}
+    $X/2$ gate via analytic, 3-sample, 2-derivative}
 \end{figure}
@@
 %% and 2 at A_{\pm} as well as 2-derivative of both parameters separately.
+%% (maybe add 5-sample and 3-derivative)
 \begin{figure}
@@
   \caption{Master equation simulation with $T_{2}$ dissipation
-    comparing the gates robust to $\delta \omega_{q}$ and $\delta A$ against
+    comparing the $X/2$ gate robust to $\delta \omega_{q}$ and $\delta a$ against
     the analytic gate.}
@@
 %% TODO: Fig: Full ME simulation for best T1 and T\phi pulses
-%% v.s. analytic
+%% v.s. analytic (?)
 
 %% SECTION 6
-\section{Discussion}
+\section{Conclusion}
 We have proposed some schemes and they work well.
@@
 \section{ME Simulation}
-We model dissipation using the lindblad master
+We model dissipation using the Lindblad master
 equation and standard collapse operators for

### 2020-07-22 paper/paper.tex
paper has all sections flushed out

@@
 
-%% SECTION 1
+%% S1
 \section{Introduction}
@@
 
-%% SECTION 2
+%% S2
 \section{QOC + AL-iLQR}
@@
 such as Runge-Kutta, of the form
-\label{eq:step-approx}
-\begin{align}
-  \ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}
-\end{align}
+$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
 
@@
 and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function similar to
-equation \ref{eq:step-approx}
+The state is propagated using a dynamics function
 $x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
@@
 \end{equation}
-$\lambda_{k}$ is a Lagrange multiplier. $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal. $i$ indicates the $i$-th constraint functional.
+where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
+with $\mu_{k}$ along the diagonal.
 $\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
 \begin{align}
-  \lambda_{k_{i}}^{'} &= \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*})\\
-  \mu_{k_{i}}^{'} &= \phi \mu_{k_{i}}
+  \lambda_{k_{i}} &\gets \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*})\\
+  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
 \end{align}
-where $x^{*}, u^{*}$ are the optimal state and controls from the iLQR execution
+where $x^{*}, u^{*}$ are the optimal state and controls from the iLQR execution,
+$i$ indicates the $i$-th constraint functional,
 and $\phi$ is a hyperparameter. With this updated form of the cost
@@
 
-%% SECTION 3
+%% F4.1
+%% TODO: Update images. Use new gates constructed with
+%% spline T1 and use smaller sim times.
+\begin{figure*}[ht]
+  \begin{subfigure}{.3\linewidth}
+    \includegraphics[width=\linewidth]{assets/00045_spin16.png}
+    \label{fig:sub-first}
+  \end{subfigure}
+  \begin{subfigure}{.3\linewidth}
+    \includegraphics[width=\linewidth]{assets/00047_spin16.png}
+    \label{fig:sub-second}
+  \end{subfigure}
+  \begin{subfigure}{.3\linewidth}
+    \includegraphics[width=\linewidth]{assets/00049_spin16.png}
+    \label{fig:sub-third}
+  \end{subfigure}
+  \caption{Master equation simulation with $T_{1}$ dissipation for all basis gates,
+    comparing the performance of the analytic gates and the $T_{1}$ optimized QOC gates.}
+  \label{fig:fig}
+\end{figure*}
+
+%% S3
 \section{QOC on the Fluxonium}
@@
 \end{align}
-$h$ is Planck's constant. The flux amplitude $a$ is experimentally
+where $\omega_{q}$ is the qubit frequency at the flux frustration point,
+$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
+are pauli matrices. The flux amplitude $a$ is experimentally
 realized by modulating the flux 
-threading the deivce $\Phi_{\textrm{ext}}$.
+threading the device ($\Phi_{\textrm{ext}}$).
 The flux amplitude is obtained from the external flux by
@@
 The optimization is performed
-over the second derivative of the flux modulation amplitude
+over the second derivative of the flux amplitude
 ($\frac{d^{2}}{dt^{2}} a(t)$) so that the first derivative
 ($\frac{d}{dt} a(t)$), proportional ($a(t)$), and integral ($\int a(t)$)
-are tracked in the augmented state vector $x$ and obtained by
+are propagated in the state vector and obtained by
 integration in the dynamics function $f$.
@@
 
-%% F4.1
-%% TODO: Update images. Use new gates constructed with
-%% spline T1 and use smaller sim times.
-\begin{figure*}[!t]
-  \begin{subfigure}{.3\linewidth}
-    \includegraphics[width=\linewidth]{assets/00045_spin16.png}
-    \label{fig:sub-first}
-  \end{subfigure}
-  \begin{subfigure}{.3\linewidth}
-    \includegraphics[width=\linewidth]{assets/00047_spin16.png}
-    \label{fig:sub-second}
-  \end{subfigure}
-  \begin{subfigure}{.3\linewidth}
-    \includegraphics[width=\linewidth]{assets/00049_spin16.png}
-    \label{fig:sub-third}
-  \end{subfigure}
-  \caption{Master equation simulation with $T_{1}$ dissipation for all basis gates.}
-  \label{fig:fig}
-\end{figure*}
-
-%% SECTION 4
+%% S4
 \section{Robustness to $T_{1}$-type Noise}
@@
 \begin{equation}
-  P_{1}(t_{k}) = \int_{0}^{t_{k}} \gamma_{1}(a(t)) dt
+  P_{1}(t) = \int_{0}^{t} \gamma_{1}(a(t^{\prime})) dt^{\prime}
 \end{equation}
@@
 %% and as good flux noise mitigation.
-%% \begin{table}
-%%   \begin{tabular}{c | c | c | c}
-%%     & Analytic  & QOC &\\
-%%     Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & Speedup\\
-%%     \hline
-%%     Z/2 & 4.615 & 2.576 & 1.792\\
-%%     Y/2 & 2.826 & 1.888 & 1.496\\
-%%     X/2 & 9.562 & 5.080 & 1.882\\
-%%   \end{tabular}
-%%   \caption{Probability of longitudinal relaxation for each gate
-%%     evaluated at the gate's duration.}
-%% \end{table}
+\begin{table}[h]
+  \begin{tabular}{c | c | c | c}
+    & Analytic  & QOC &\\
+    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & Speedup\\
+    \hline
+    Z/2 & 4.615 & 2.576 & 1.792\\
+    Y/2 & 2.826 & 1.888 & 1.496\\
+    X/2 & 9.562 & 5.080 & 1.882\\
+  \end{tabular}
+  \caption{Probability of longitudinal relaxation for each gate
+    evaluated at the gate's duration.}
+\end{table}
 
-%% SECTION 5
+%% S5
 \section{Robustness to $T_{\phi}$-type Noise}
@@
 We propose two methods for engineering robustness
-to system parameter deviations. The first we call the
+to parameter deviations. The first we call the
 derivative method. We draw on the intuition that
 making the state evolution insensitive to changes
-in the system parameter is encoded in the derivative
+in the parameter is encoded in the derivative
 $\partial_{\omega_{q}}^{l} \ket{\psi}$.
-The $l$ state derviatives are added to the augmented
+The $l$ state derviatives are propgated in the
 state vector. Their dynamics are found by differentiating
 the TDSE dynamics with respect to the parameter of interest
-\ref{eq:tdse}. Setting the target state derivatives
+(see Appendix). Setting the target derivatives of the state
 to zero vectors results in quadratic costs at each
@@
 to represent the distribution of the parameter.
-For example, in the 3-point sampling method we
-propagate $\ket{\psi_{\pm, I}}$ where
+For example, in the 2-point sampling method we
+propagate $\ket{\psi^{\pm, I}(t)}$ where
 $\omega_{q}^{\pm} = \omega_{q} \pm \sigma_{\omega_{q}}$,
@@
 each knot point
-${\lvert \ket{\psi_{\pm, I}} - \ket{\psi_{f}} \rvert}^{2}$
+${\lvert \ket{\psi^{\pm, I}_{k}} - \ket{\psi_{f}} \rvert}^{2}$
 
-%% LOH
+(Results) We compare the robust control
+schemes to a dynamic decoupling pulse
+for a single $R_{x}(\pi/2)$ gate
+subject to qubit frequency detuning
+($\omega_{q} \gets \omega_{q} + \delta \omega_{q}$)
+to demonstrate the applicability of our techniques to mitigate
+system parameter deviations.
+We use a compensating for off-resonance error
+(CORPSE) dynamic decoupling pulse that is optimized
+to mitigate first-order error arising due to the detuning.
+The pulse is composed of
+three rotations about the $\hat{x}$, $-\hat{x}$, and $\hat{x}$ axes
+of the Bloch sphere
+designed to compensate for erroneous rotations
+(see Appendix). We simulate the gate fidelity as
+a function of the qubit frequency detuning for
+variants of the
+derivative method, sample method, and
+CORPSE pulse (see Figure 2a). We also find that
+increasing the gate duration allows the optimizer
+to find a more robust pulse
+(see Figure 3a). Increasing the order of the derivative
+and sample methods requires propagating an additional quantum state
+in the state vector, a linear increase
+in the computational complexity, but finding higher order
+variants of the CORPSE family is prohibitively difficult.
 
-(Results) We compare the simplest schemes
-for the derivative and sampling methods to
-the CORPSE dynamic decoupling sequence.
-They perform X better by Y metric.
-
-We also find that longer gate durations allow for
-greater robustness.
+Additionally, we compare the robust control
+schemes to a dynamic decoupling pulse
+for sequential $R_{x}(\pi/2)$ gates
+subject to $T_{\phi}$ noise to demonstrate the applicability
+of our techniques to mitigate noise of this type.
+We simulate the gate fidelity as a function of time
+for variants of the derivative method,
+sample method, and B2CORPSE pulse (see Figure 3).
+The B2CORPSE pulse is constructed to be robust
+to qubit frequency detuning errors and
+amplitude scaling errors (see Appendix).
+The derivative and sampling methods are employed to
+mitigate error due to qubit frequency detuning
+($\omega_{q} \gets \omega_{q} + \delta \omega_{q}$)
+as well as flux amplitude scaling error
+($a \gets a (1 + \delta a)$).
 
 %% F5.1
-%% TODO: Update this and make sure it is for X/2.
-%% I'm not sure if we should include all gates Z/2, Y/2, X/2
-%% because I don't know what the dynamic decoupling analogy
-%% for Z/2 and Y/2 are. On the other hand we would
-%% want to include all of them to show that the benefit
-%% is not uniform across gates.
-%% (maybe add 5-sample and 3-derivative)
+%% TODO: Make this figure for real. Add inset to second figure
+%% to show high performing variants at high time steps
 \begin{figure}
-  %% \includegraphics[width=\linewidth]{assets/00005_spin12_hpsweep.png}
-  \caption{fidelity v.s. $\omega$ for the $X/2$ gate via
-    analytic, 3-sample, 2-derivative}
+  \begin{subfigure}{\linewidth}
+    \label{fig:5.1:a}
+    \includegraphics[width=\linewidth]{assets/00005_spin12_hpsweep.png}
+  \end{subfigure}
+
+  \begin{subfigure}{\linewidth}
+    \includegraphics[width=\linewidth]{assets/00006_spin12_hpsweep.png}
+    \label{fig:5.1:a}
+  \end{subfigure}
+  \caption{\centering Robustness to qubit frequency detuning for
+    the (2, 4)-point sampling method, (2\textsuperscript{nd}, 3\textsuperscript{rd})
+    order derivative method, and CORPSE pulses for the $R_{x}(\pi/2)$ gate.
+    (a) Gate fidelity as a function of the qubit frequency detuning.
+    (b) Gate fidelity at one standard deviation from the nominal qubit frequency
+    as a function of time.
+  }
 \end{figure}
@@
 %% F5.2
-%% TODO: Update this and make sure it is for X/2.
-%% Add inset to show high-performing variants at large time steps.
+%% TODO: Make this figure.
 \begin{figure}
-  %% \includegraphics[width=\linewidth]{assets/00006_spin12_hpsweep.png}
-  \caption{fidelity at $\omega \pm \sigma_{\omega}$ v.s. time for the
-    $X/2$ gate via analytic, 3-sample, 2-derivative}
+  \includegraphics[width=\linewidth]{assets/t2_temp.jpg}
+  \caption{Robustness to $T_{\phi}$-type noise for the
+    the (2, 4)-point sampling method, (2\textsuperscript{nd}, 3\textsuperscript{rd})
+    order derivative method, and B2CORPSE pulse for the $R_{x}(\pi/2)$ gate.}
 \end{figure}
 
-%% F5.3
-%% TODO: Make this figure. My guess is we want a 5-sample, with 2 at \omega_{\pm}
-%% and 2 at A_{\pm} as well as 2-derivative of both parameters separately.
-%% (maybe add 5-sample and 3-derivative)
-\begin{figure}
-  %% \includegraphics[width=\linewidth]{assets/t2_temp.jpg}
-  \caption{Master equation simulation with $T_{2}$ dissipation
-    comparing the $X/2$ gate robust to $\delta \omega_{q}$ and $\delta a$ against
-    the analytic gate.}
-\end{figure}
-
-%% TODO: Fig: Full ME simulation for best T1 and T\phi pulses
-%% v.s. analytic (?)
-
-%% SECTION 6
+%% S6
 \section{Conclusion}
@@
 \appendix
-%% APPENDIX 1
+%% AA
 \section{Ricatti Recursion}
@@
 
-%% APPENDIX 2
+%% AB
 \section{Experiment}
@@
 
-%% APPENDIX 3
+%% AC
 \section{ME Simulation}
@@
 
-%% APPENDIX 4
+%% AD
 \section{Derivative Method}

### 2020-07-23 paper/paper.tex
first draft

@@
 quantum computation is gate-based. The Quantum Optimal Control (QOC)
-literature has saught to answer the question "What is the best way to construct a quantum gate?".
+literature has saught to construct techniques for engineering waveforms to manipulate
+quantum systems with high fidelity.
 Analytic methods
@@
 relevant to gate construction. Next we outline a method for
-making the optimization $T_{1}$ aware. Finally, we present
+making the optimization longitudinal relaxation aware. Finally, we present
 some methods for engineering robustness to pure dephasing and
@@
 the problem tractable by discretizing the problem into $N$
-time steps. Typical integration techniques for the TDSE include
+knot points (time steps). Typical integration techniques for the TDSE include
 approximating unitary propagators as well as explicit integration methods,
@@
 the infidelity between the inital state evolved
-to the final time step ($\ket{\psi_{N}(u(t))}$)
+to the final knot point ($\ket{\psi_{N}(u(t))}$)
 and the target state ($\ket{\psi_{f}}$). In general
@@
 the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each time step
+the same functional is evaluated at each knot point
 \begin{equation}
@@
 where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at time step $k$ and final state, $u_{k}$ are the controls,
+at knot point $k$ and final state, $u_{k}$ are the controls,
 and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
@@
 In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics.
+and $f$ encodes the TDSE dynamics. In the following
+we refer to $\ket{\psi}$ as the state. We refer to $x$ and $u$ as
+the augmented state and augmented controls, respectively.
 
@@
 is that there exists a dynamic programming algorithm to compute the
-optimal update to the controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each time step. This algorithm proceeds by deriving a
-recurrence relation between time steps $k$ and $k + 1$ for the optimal
+optimal update to the augmented controls ($u_{k}$) which minimizes the functional
+($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
+recurrence relation between knot points $k$ and $k + 1$ for the optimal
 feedback law--known as the Ricatti recursion (see Appendix). The
 iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all time steps on multiple
+and applies the Ricatti recursion to all knot points on multiple
 executions.
@@
 to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero-valued when the constraint is satisfied. The AL-iLQR
-method updates the penalty multiplier between
-iLQR executions. The penalty multiplier associated with the
-constraint functional estimates the Lagrange multiplier of
-the constraint. The functional now takes the form
+zero or negative when the constraint is satisfied. The AL-iLQR
+method associates a penalty multiplier with the functional
+that estimates the constraint's Lagrange multiplier.
+The algorithm updates the penalty multiplier between
+iLQR executions. In this scheme the functional takes the form
 \begin{equation}
@@
 \begin{align}
-  \lambda_{k_{i}} &\gets \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*})\\
+  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
   \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
 \end{align}
-where $x^{*}, u^{*}$ are the optimal state and controls from the iLQR execution,
+where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
 $i$ indicates the $i$-th constraint functional,
@@
 $a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are pauli matrices. The flux amplitude $a$ is experimentally
+are Pauli matrices. The flux amplitude $a$ is experimentally
 realized by modulating the flux 
@@
 \delta \Phi_{\textrm{ext}} / (h \Phi_{0})$
-where $\delta \Phi_{\textrm{ext}} = \Phi_{\textrm{ext}} - 0.5 \Phi_{0}$ is the flux
+where $\hat{\phi}$ is the phase operator, $E_{L}$ is the characteristic inductance energy, $\Phi_{0}$
+is a flux quantum, and
+$\delta \Phi_{\textrm{ext}} = \Phi_{\textrm{ext}} - 0.5 \Phi_{0}$ is the flux
 offset from the flux frustration point.
@@
 In addition to these constraints we penalize the norm
-of the first and second derivatives of the control
+of the first and second derivatives of the flux amplitude
 to mitigate AWG ringing.
 
-The optimization is performed
-over the second derivative of the flux amplitude
-($\frac{d^{2}}{dt^{2}} a(t)$) so that the first derivative
+The optimization is performed over the second derivative of the flux amplitude
+($\frac{d^{2}}{dt^{2}} a(t)$) which is contained in the
+augmented control vector. The first derivative
 ($\frac{d}{dt} a(t)$), proportional ($a(t)$), and integral ($\int a(t)$)
-are propagated in the state vector and obtained by
+flux amplitude terms
+are contained in the augmented state vector. They are obtained from
+the second derivative of the flux amplitude by
 integration in the dynamics function $f$.
@@
 that the qubit decays as a result of longitudinal
-relaxation. To this end we augment the state vector ($x_{k}$) with the
+relaxation. To this end we add the
 longitudinal relaxation probability
+to the augmented state vector
 \begin{equation}
@@
 We allow the optimizer to tune the gate duration by
-augmenting the control vector $u_{k}$ with
-the time step between knot points $\Delta t_{k}$.
+adding the time step between knot points $\Delta t_{k}$
+to the augmented control vector $u_{k}$.
 Promoting $\Delta t_{k}$ to a decision variable, rather
@@
 
+%% F5.1
+%% TODO: Make this figure for real. Add inset to second figure
+%% to show high performing variants at high time steps
+\begin{figure}
+  \begin{subfigure}{\linewidth}
+    \label{fig:5.1:a}
+    \includegraphics[width=\linewidth]{assets/00005_spin12_hpsweep.png}
+  \end{subfigure}
+
+  \begin{subfigure}{\linewidth}
+    \includegraphics[width=\linewidth]{assets/00006_spin12_hpsweep.png}
+    \label{fig:5.1:a}
+  \end{subfigure}
+  \caption{\centering Robustness to qubit frequency detuning for
+    the (2, 4)-point sampling method, (2\textsuperscript{nd}, 3\textsuperscript{rd})
+    order derivative method, and CORPSE pulses for the $R_{x}(\pi/2)$ gate.
+    (a) Gate fidelity as a function of the qubit frequency detuning.
+    (b) Gate fidelity at one standard deviation from the nominal qubit frequency
+    as a function of time.
+  }
+\end{figure}
+
+%% F5.2
+%% TODO: Make this figure.
+\begin{figure}
+  \includegraphics[width=\linewidth]{assets/t2_temp.jpg}
+  \caption{Robustness to $T_{\phi}$-type noise for the
+    the (2, 4)-point sampling method, (2\textsuperscript{nd}, 3\textsuperscript{rd})
+    order derivative method, and B2CORPSE pulse for the $R_{x}(\pi/2)$ gate.}
+\end{figure}
+
 %% S5
@@
 making the state evolution insensitive to changes
-in the parameter is encoded in the derivative
-$\partial_{\omega_{q}}^{l} \ket{\psi}$.
-The $l$ state derviatives are propgated in the
-state vector. Their dynamics are found by differentiating
+in the parameter ($x$) is encoded in the derivative of
+the state with respect to that parameter
+$\partial_{x}^{l} \ket{\psi}$.
+The derviatives of the state are propgated in the
+augmented state vector. Their dynamics are found by differentiating
 the TDSE dynamics with respect to the parameter of interest
@@
 knot point
-${\lvert \partial_{\omega_{q}}^{l} \ket{\psi_{k}} \rvert}^{2}$.
+${\lvert \partial_{x}^{l} \ket{\psi_{k}} \rvert}^{2}$.
 
@@
 \cite{manchester2016derivative} \cite{tronarp2016sigma}.
-We augment the state vector with additional
-states that follow deviant dynamics. The parameters of interest
+We add additional
+states that follow deviant dynamics to the augmented
+state vector. The parameters of interest
 are altered by a small deviation, typically chosen
@@
 to find a more robust pulse
-(see Figure 3a). Increasing the order of the derivative
+(see Figure 2b). Increasing the order of the derivative
 and sample methods requires propagating an additional quantum state
-in the state vector, a linear increase
+in the augmented state vector, a linear increase
 in the computational complexity, but finding higher order
@@
 ($a \gets a (1 + \delta a)$).
-
-%% F5.1
-%% TODO: Make this figure for real. Add inset to second figure
-%% to show high performing variants at high time steps
-\begin{figure}
-  \begin{subfigure}{\linewidth}
-    \label{fig:5.1:a}
-    \includegraphics[width=\linewidth]{assets/00005_spin12_hpsweep.png}
-  \end{subfigure}
-
-  \begin{subfigure}{\linewidth}
-    \includegraphics[width=\linewidth]{assets/00006_spin12_hpsweep.png}
-    \label{fig:5.1:a}
-  \end{subfigure}
-  \caption{\centering Robustness to qubit frequency detuning for
-    the (2, 4)-point sampling method, (2\textsuperscript{nd}, 3\textsuperscript{rd})
-    order derivative method, and CORPSE pulses for the $R_{x}(\pi/2)$ gate.
-    (a) Gate fidelity as a function of the qubit frequency detuning.
-    (b) Gate fidelity at one standard deviation from the nominal qubit frequency
-    as a function of time.
-  }
-\end{figure}
-
-%% F5.2
-%% TODO: Make this figure.
-\begin{figure}
-  \includegraphics[width=\linewidth]{assets/t2_temp.jpg}
-  \caption{Robustness to $T_{\phi}$-type noise for the
-    the (2, 4)-point sampling method, (2\textsuperscript{nd}, 3\textsuperscript{rd})
-    order derivative method, and B2CORPSE pulse for the $R_{x}(\pi/2)$ gate.}
-\end{figure}
 
@@
 %% AC
-\section{ME Simulation}
+\section{Dissipation Simulation}
 We model dissipation using the Lindblad master
-equation and standard collapse operators for
-longitudinal relaxation and dephasing
-of a two-level system.
+equation. 
+\begin{equation}
+  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
+\end{equation}
+where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
+and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
+For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
+and $L_{\uparrow} = \sigma^{+}/2$,
+$L_{\downarrow} = \sigma^{-}/2$
+are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
+$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
+$L_{2} = (I - \sigma_{z})/2$.
 
 %% AD
-\section{Derivative Method}
+\section{Derivative Method Dynamics}
+To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
+we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
+($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
+state with respect to the qubit frequency ($\omega_{q}$) and the flux amplitude ($a$).
+Here we present the dynamics for the second derivative of the state with respect to the
+qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
+of $\omega_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
+brevity.
 \begin{equation}
   \begin{aligned}
-    \partial_{\omega}^{2} \partial_{t} \psi
-    = \frac{-i}{h} [&2 \cdot \frac{\sigma_{z}}{2} \partial_{\omega} \psi\\
-      &+ (\omega \frac{\sigma_{z}}{2} + A \frac{\sigma_{x}}{2}) \partial_{\omega}^{2} \psi]
+      \partial_{\omega_{q}} \partial_{t} \ket{\psi} &= \partial_{\omega_{q}} H \ket{\psi}\\
+      &= (\partial_{\omega_{q}} H) \ket{\psi} + H (\partial_{\omega_{q}} \ket{\psi})\\
+      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{\omega_{q}} \ket{\psi})
   \end{aligned}
 \end{equation}
+\begin{equation}
+  \begin{aligned}
+    \partial_{\omega_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{\omega_{q}} (\partial_{\omega_{q}} H \ket{\psi})\\
+    &= (\partial_{\omega_{q}}^{2} H) \ket{\psi} + 2 (\partial_{\omega_{q}} H)(\partial_{\omega_{q}} \ket{\psi})\\
+    &\quad + H (\partial_{\omega_{q}}^{2} \ket{\psi})\\
+    &= \sigma_{z} (\partial_{\omega_{q}} \ket{\psi}) + H (\partial_{\omega_{q}}^{2} \ket{\psi})
+  \end{aligned}
+\end{equation}
+The augmented state vector carries $\partial_{\omega_{q}}^{l} \ket{\psi}$
+which appears explicitly in its own dynamics. Due to the dependence of $H$ on $\omega_{q}$ and $a$,
+the $l$\textsuperscript{th} state derivative is coupled to the
+$l - 1$\textsuperscript{th} state derivative.
+
+%% AE
+\section{Dynamic Decoupling Sequences}
+Here we present the equations for constructing the CORPSE and
+B2CORPSE sequences.
+
+%% AF
+\section{Complex Tensor Handling}
+We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
+because the software we use does not support complex numbers yet.
 

### 2020-07-23 paper/paper.tex
small update

@@
     \includegraphics[width=\linewidth]{assets/00045_spin16.png}
-    \label{fig:sub-first}
   \end{subfigure}
@@
     \includegraphics[width=\linewidth]{assets/00047_spin16.png}
-    \label{fig:sub-second}
   \end{subfigure}
@@
     \includegraphics[width=\linewidth]{assets/00049_spin16.png}
-    \label{fig:sub-third}
   \end{subfigure}
@@
     comparing the performance of the analytic gates and the $T_{1}$ optimized QOC gates.}
-  \label{fig:fig}
 \end{figure*}
@@
 %% S4
-\section{Robustness to $T_{1}$-type Noise}
+\section{Robustness to Longitudinal Relaxation}
 (Strategy) We seek to minimize the probability
@@
     \includegraphics[width=\linewidth]{assets/00006_spin12_hpsweep.png}
-    \label{fig:5.1:a}
+    \label{fig:5.1:b}
   \end{subfigure}
@@
 %% S5
-\section{Robustness to $T_{\phi}$-type Noise}
+\section{Robustness to Pure Dephasing}
 (Strategy) For the fluxonium, pure dephasing acts

### 2020-07-29 experiments/spin/figures.jl
prog

@@ -1,0 +1,136 @@
+"""
+figures.jl
+"""
+
+using LaTeXStrings
+using Printf
+import Plots
+
+include(joinpath(ENV["RBQOC_PATH"], "rbqoc.jl"))
+
+# Configure paths.
+META_SAVE_PATH = joinpath(ENV["RBQOC_PATH"], "out", "spin")
+EXPERIMENT_NAME = "figures"
+SAVE_PATH = joinpath(META_SAVE_PATH, EXPERIMENT_NAME)
+
+# Configure plotting.
+ENV["GKSwstype"] = "nul"
+Plots.gr()
+
+# constants
+SAVE_FILE_PATH_KEY = 1
+SAVE_TYPE_KEY = 2
+DATA_FILE_PATH_KEY = 3
+F1_GATE_COUNT = Integer(1.5e4)
+
+@enum PulseType begin
+    qoc = 1
+    analytic = 2
+end
+
+PT_TO_STR = Dict(
+    qoc => "QOC",
+    analytic => "Analytic",
+)
+
+# Figure 1
+function plot_fidelity_by_gate_count_single(fig, path; inds=nothing)
+    (fidelities,
+     pulse_type) = h5open(path, "r") do data_file
+         fidelities = read(data_file, "fidelities")
+         pulse_type = PulseType(read(data_file, "pulse_type"))
+         return (fidelities, pulse_type)
+    end
+    gate_count = size(fidelities)[1] - 1
+    gate_count_axis = Array(0:1:gate_count)
+    label = "$(PT_TO_STR[pulse_type])"
+    if isnothing(inds)
+        inds = 1:gate_count + 1
+    end
+    # inds = 1:4:gate_count + 1
+    Plots.plot!(fig, gate_count_axis[inds], fidelities[inds], label=label, legend=:bottomleft)
+end
+
+
+function plot_fidelity_by_gate_count(paths; inds=nothing, title=nothing, ylims=(0, 1), yticks=(0:0.1:1))
+    plot_file_path = generate_save_file_path("png", EXPERIMENT_NAME, SAVE_PATH)
+    fig = Plots.plot(dpi=DPI, ylims=ylims, yticks=yticks, title=title)
+    for path in paths
+        plot_fidelity_by_gate_count_single(fig, path; inds=inds)
+    end
+    Plots.ylabel!("Fidelity")
+    Plots.xlabel!("Gate Count")
+    Plots.savefig(fig, plot_file_path)
[…]

### 2020-07-29 experiments/spin/spin14.py

@@ -49,6 +49,7 @@ AMP_0 = 2 * np.pi * 1.25e-1
 DT = 1e-2
 DT_INV = 1e2
 T_ZPI = 35.714285714285715
+T_TOT_ZPIBY2 = 17.85714285714286
 
 # XPI pulse parameters
 T_XZ_XPI = 4.800613
@@ -120,13 +121,13 @@ def gen_controls_xpi(t, shape=Shape.SQUARE):
         #ENDIF
     #ENDIF
     
-    return [c1]
+    return np.array([c1,])
 
 
 def save_controls_xpi():
     shape = Shape.TRIANGLE
     control_eval_times = np.linspace(0, T_TOT_XPI, N_TOT_XPI)
-    controls = np.array([gen_controls_xpi(t, shape=shape) for t in control_eval_times])
+    controls = np.vstack([gen_controls_xpi(t, shape=shape) for t in control_eval_times])
     evolution_time = controls.shape[0] * DT
     save_file_path = generate_save_file_path(EXPERIMENT_NAME, SAVE_PATH)
     with h5py.File(save_file_path, "a") as save_file:
@@ -142,14 +143,12 @@ def save_controls_xpi():
 T_XZ_YPIBY2 = 2.1656249366575766
 T_Z_YPIBY2 = 15.142330599557274
 T_TOT_YPIBY2 = 2 * T_XZ_YPIBY2 + T_Z_YPIBY2
-N_TOT_YPIBY2 = int(T_TOT_YPIBY2 / DT)
 T0_YPIBY2 = 0.
 T1_YPIBY2 = T_XZ_YPIBY2 / 2
 T2_YPIBY2 = T_XZ_YPIBY2
 T3_YPIBY2 = T2_YPIBY2 + T_Z_YPIBY2
 T4_YPIBY2 = T3_YPIBY2 + T_XZ_YPIBY2 / 2
 T5_YPIBY2 = T3_YPIBY2 + T_XZ_YPIBY2
-
 
 def gen_controls_ypiby2(t, shape=Shape.TRIANGLE):
     if shape == Shape.TRIANGLE:
@@ -175,17 +174,65 @@ def gen_controls_ypiby2(t, shape=Shape.TRIANGLE):
             c1 = -amp
     #ENDIF
 
-    return [c1]
+    return np.array([c1,])
 #ENDDEF
 
 
 def save_controls_ypiby2():
     # generate
-    shape = Shape.TRIANGLE
[…]

### 2020-07-29 experiments/spin/spin15.jl

@@ -16,14 +16,15 @@ WDIR = ENV["RBQOC_PATH"]
 SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 
 # Define the optimization.
-EVOLUTION_TIME = 56.80
+CONSTRAINT_TOLERANCE = 1e-8
 CONTROL_COUNT = 1
-DT_INTEGRATOR = 1
+DT_STATIC = DT_PREF
+DT_STATIC_INV = DT_PREF_INV
 DT_INIT = DT_PREF
 DT_INIT_INV = DT_PREF_INV
 DT_MIN = DT_INIT / 2
 DT_MAX = DT_INIT * 2
-N = Int(EVOLUTION_TIME * DT_INIT_INV) + 1
+EVOLUTION_TIME = 20.0
 
 # Define the problem.
 INITIAL_STATE_0 = SA[1., 0, 0, 0]
@@ -44,8 +45,8 @@ YPIBY2_0 = SA[1., 1, 0, 0] / sqrt(2)
 YPIBY2_1 = SA[-1., 1, 0, 0] / sqrt(2)
 XPIBY2_0 = SA[1., 0, 0, -1] / sqrt(2)
 XPIBY2_1 = SA[0., 1, -1, 0] / sqrt(2)
-TARGET_STATE_0 = XPIBY2_0
-TARGET_STATE_1 = XPIBY2_1
+TARGET_STATE_0 = ZPIBY2_0
+TARGET_STATE_1 = ZPIBY2_1
 TARGET_ASTATE = [
     TARGET_STATE_0;
     TARGET_STATE_1;
@@ -81,12 +82,12 @@ Base.size(model::Model) = (model.n, model.m)
 
 function TrajectoryOptimization.dynamics(model::Model, astate, acontrols, time)
     neg_i_control_hamiltonian = astate[CONTROLS_IDX][1] * NEG_I_H_C1
-            delta_state_0 = (OMEGA_NEG_I_H_S + neg_i_control_hamiltonian) * astate[STATE_0_IDX]
+    delta_state_0 = (OMEGA_NEG_I_H_S + neg_i_control_hamiltonian) * astate[STATE_0_IDX]
     delta_state_1 = (OMEGA_NEG_I_H_S + neg_i_control_hamiltonian) * astate[STATE_1_IDX]
     delta_int_control = astate[CONTROLS_IDX]
     delta_control = astate[DCONTROLS_DT_IDX]
[…]

### 2020-07-29 experiments/spin/t1.jl

@@ -2,7 +2,9 @@
 t1.jl - do some calculations for fluxonium t1
 """
 
+using Dierckx
 using HDF5
+using Interpolations
 using LaTeXStrings
 using LinearAlgebra
 using TrajectoryOptimization
@@ -16,33 +18,15 @@ EXPERIMENT_NAME = "spin15"
 WDIR = ENV["ROBUST_QOC_PATH"]
 SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 T1_SAVE_FILE_PATH = joinpath(SAVE_PATH, "t1.h5")
+T1CMP_PLOT_FILE_PATH = joinpath(SAVE_PATH, "t1splinecmp.png")
 
 # Plotting configuration.
 ENV["GKSwstype"] = "nul"
 Plots.gr()
 DPI = 300
 
-function generate_save_file_path(save_file_name, save_path)
-    # Ensure the path exists.
-    mkpath(save_path)
-
-    # Create a save file name based on the one given; ensure it will
-    # not conflict with others in the directory.
-    max_numeric_prefix = -1
-    for (_, _, files) in walkdir(save_path)
-        for file_name in files
-            if occursin("_$save_file_name.h5", file_name)
-                max_numeric_prefix = max(parse(Int, split(file_name, "_")[1]))
-            end
-        end
-    end
-
-    save_file_name = "_$save_file_name.h5"
-    save_file_name = @sprintf("%05d%s", max_numeric_prefix + 1, save_file_name)
-
-    return joinpath(save_path, save_file_name)
-end
-
+# misc. constants
+SAMPLE_SIZE = Int(1e3)
 
 # Define experimental constants.
 # E / h
@@ -63,6 +47,17 @@ E_CHARGE = 1.60217646e-19
 INVERSE_FLUX_QUANTUM = 4.835977958971655e14
 # FLUX_FRUSTRATION = FLUX_QUANTUM / 2
 FLUX_FRUSTRATION = 1.033917036516689e-15
+# raw T1s reported in units of microseconds
+T1_ARRAY = [
+    1597.923, 1627.93, 301.86, 269.03, 476.33, 1783.19, 2131.76, 2634.50, 
[…]

### 2020-07-29 experiments/spin/verify.py

@@ -3,6 +3,7 @@ verify.py - Quitp verification.
 """
 
 from argparse import ArgumentParser
+from enum import Enum
 from functools import reduce
 import os
 
@@ -88,7 +89,7 @@ NEG_YPIBY2 = np.array([[1., 1.],
 ZPIBY2 = np.array([[1. - 1.j, 0],
                    [0,        1. + 1.j]]) * INV_ROOT_2
 INITIAL_STATE = gen_initial_state(2)
-TARGET_STATE = np.matmul(YPIBY2, INITIAL_STATE)
+TARGET_STATE = np.matmul(XPIBY2, INITIAL_STATE)
 INITIAL_DENSITY = np.matmul(INITIAL_STATE, conjugate_transpose(INITIAL_STATE))
 TARGET_DENSITY = np.matmul(TARGET_STATE, conjugate_transpose(TARGET_STATE))
 
@@ -146,7 +147,6 @@ def main():
     parser.add_argument("--ename", action="store", type=str)
     parser.add_argument("--cname", action="store", type=str)
     parser.add_argument("--cidx", action="store", type=int)
-    parser.add_argument("--cidx", action="store", type=str)
     args = vars(parser.parse_args())
     do_spin = args["spin"]
     experiment_name = args["ename"]

### 2020-07-29 rbqoc.jl
prog

@@ -3,11 +3,13 @@ rbqoc.jl - common definitions for the rbqoc repo
 """
 
 # imports
+using Dates
 using Dierckx
 using DifferentialEquations
 using HDF5
+using Interpolations
+using LinearAlgebra
 using Plots
-using HDF5
 using Polynomials
 using Printf
 using Random
@@ -30,11 +32,13 @@ ALPHA = 0.2
 
 # other constants
 DEQJL_MAXITERS = 1e10
+DEQJL_ADAPTIVE = false
 
 # types
-@enum DissipationType begin
-    dissipation = 1
-    nodissipation = 2
+@enum DynamicsType begin
+    schroed = 1
+    lindbladnodis = 2
+    lindbladdis = 3
 end
 
 
@@ -53,8 +57,9 @@ end
 
 
 DT_TO_STR = Dict(
-    dissipation => "Dissipation",
-    nodissipation => "No Dissipation",
+    schroed => "Schroedinger",
+    lindbladnodis => "Lindblad No Dissipation",
+    lindbladdis => "Lindblad Dissipation",
 )
 
 GT_TO_STR = Dict(
@@ -74,19 +79,23 @@ function generate_save_file_path(extension, save_file_name, save_path)
     max_numeric_prefix = -1
     for (_, _, files) in walkdir(save_path)
         for file_name in files
-            if occursin("_$save_file_name.$(extension)", file_name)
+            if occursin("_$(save_file_name).$(extension)", file_name)
                 max_numeric_prefix = max(parse(Int, split(file_name, "_")[1]))
             end
         end
     end
 
-    save_file_name = "_$save_file_name.$(extension)"
+    save_file_name = "_$(save_file_name).$(extension)"
     save_file_name = @sprintf("%05d%s", max_numeric_prefix + 1, save_file_name)
 
     return joinpath(save_path, save_file_name)
 end
 
 
+"""
+grab_controls - do some extraction of relevant controls
+data for common h5 save formats
+"""
 function grab_controls(save_file_path; save_type=jl)
     data = h5open(save_file_path, "r") do save_file
         if save_type == jl
@@ -96,6 +105,9 @@ function grab_controls(save_file_path; save_type=jl)
         elseif save_type == samplejl
             controls = read(save_file, "controls_sample")
[…]

### 2020-08-12 paper/paper.tex
updated nb folder

@@
 the trajectory optimization framework. For the device we study
-we achieve a 1.5x decrease in longitudinal relaxation probability.
+we achieve a 2.5x decrease in longitudinal relaxation probability.
 We present two methods for achieving robustness to pure dephasing,
@@
 
-(Results) We achieve a factor of 1.5 decrease in the probability
+(Results) We achieve a factor of 2.5 decrease in the probability
 of longitudinal relaxation from the analytic gates we benchmark
@@
 %% T4.1
-%% TODO: Update this table with spline T1 optimizations and calculations.
 %% TODO: Maybe best for appendix.
@@
 %% and as good flux noise mitigation.
-\begin{table}[h]
+\begin{table}[ht]
   \begin{tabular}{c | c | c | c}
@@
     \hline
-    Z/2 & 4.615 & 2.576 & 1.792\\
-    Y/2 & 2.826 & 1.888 & 1.496\\
-    X/2 & 9.562 & 5.080 & 1.882\\
+    Z/2 & 5.745 & 2.144 & 2.680\\
+    Y/2 & 5.253 & 2.011 & 2.612\\
+    X/2 & 16.251 & 6.234 & 2.607\\
   \end{tabular}

### 2020-08-20 src/spin/figures.jl
most sims are functional

@@ -25,29 +25,45 @@ Plots.gr()
     analytic = 2
     derivative = 3
     sample = 4
+    derivative2 = 5
+    sample2 = 6
 end
 
 const PT_STR = Dict(
     qoc => "QOC",
-    analytic => "Analytic",
-    sample => "Sample",
-    derivative => "Derivative"
+    analytic => "Anl.",
+    sample => "S-2",
+    derivative => "D-2",
+    derivative2 => "D-3",
+    sample2 => "S-4"
 )
 
 const PT_MARKER = Dict(
     sample => :circle,
-    derivative => :square,
+    sample2 => :square,
+    derivative2 => :utriangle,
+    derivative => :diamond,
 )
 
 const PT_COLOR = Dict(
     analytic => :lightskyblue,
     qoc => :coral,
-    sample => :green,
-    derivative => :red,
+    sample => :limegreen,
+    sample2 => :darkgreen,
+    derivative => :crimson,
+    derivative2 => :firebrick,
+)
+
+const PT_LINESTYLE = Dict(
+    analytic => :solid,
+    qoc => :solid,
+    sample => :solid,
+    sample2 => :dash,
+    derivative => :solid,
+    derivative2 => :dash,
 )
 
 const GT_LIST = [zpiby2, ypiby2, xpiby2]
-const PT_LIST = [analytic, qoc]
 
 # plotting constants
 const ALPHA_POINT = 0.4
@@ -55,6 +71,9 @@ const MS_DATA = 4
 const MS_POINT = 8
 const FS_AXIS_LABELS = 12
 const FS_AXIS_TICKS = 10
+const FS_ANNOTATE = 10
+const FS_LEGEND = 10
+const FG_COLOR_LEGEND = nothing
 const DPI_FINAL = Integer(1e3)
 
 # common dict keys
@@ -64,6 +83,8 @@ const DATA_FILE_PATH_KEY = 3
 const COLOR_KEY = 4
 const ACORDS_KEY = 5
 const MARKER_KEY = 6
+const LCORDS_KEY = 7
+const DATA2_FILE_PATH_KEY = 8
 
 
 ### ALL ###
@@ -73,7 +94,9 @@ function plot_fidelity_by_gate_count(fidelitiess; inds=nothing, title="", ylims=
                                      xlims=nothing)
     plot_file_path = generate_save_file_path("png", EXPERIMENT_NAME, SAVE_PATH)
[…]

### 2020-08-20 src/spin/spin.jl

@@ -87,6 +87,11 @@ const T1_ARRAY = [
     4364.68, 2587.82, 1661.915, 1794.468, 2173.88, 1188.83, 
     1576.493, 965.183, 560.251, 310.88
 ] * 1e3
+const T1_ARRAY_ERR = [
+    78.034, 70.57, 20.71, 20.93, 12.92, 66.93, 137.06, 319.19, 
+    194.84, 146.87, 170.584, 374.582, 322.533, 125.10,
+    105.987, 58.612, 22.295, 11.29
+] * 1e3
 # raw T1 times are in units of microseconds
 const T1_ARRAY_REDUCED = [
     1597.923, 1627.93, 301.86, 269.03, 476.33, 1783.19, 2131.76, 2634.50, 

### 2020-08-20 src/spin/spin11.jl

@@ -251,8 +251,8 @@ function run_traj(;gate_type=xpiby2, evolution_time=60., solver_type=altro,
     add_constraint!(constraints, control_bnd, 2:N-2)
     add_constraint!(constraints, control_bnd_boundary, N-1:N-1)
     add_constraint!(constraints, target_astate_constraint, N:N);
-    add_constraint!(constraints, normalization_constraint_1, 2:N-1)
-    add_constraint!(constraints, normalization_constraint_2, 2:N-1)
+    # add_constraint!(constraints, normalization_constraint_1, 2:N-1)
+    # add_constraint!(constraints, normalization_constraint_2, 2:N-1)
 
     # Instantiate problem and solve.
     prob = Problem{IT_RDI[integrator_type]}(model, obj, constraints, x0, xf, Z, N, t0, evolution_time)
@@ -317,7 +317,7 @@ function run_traj(;gate_type=xpiby2, evolution_time=60., solver_type=altro,
             write(save_file, "ctol", CONSTRAINT_TOLERANCE)
             write(save_file, "alko", AL_KICKOUT_TOLERANCE)
             write(save_file, "ilqr_dj_tol", ILQR_DJ_TOL)
-            write(save_file, "integrator_type", integrator_type)
+            write(save_file, "integrator_type", Integer(integrator_type))
             write(save_file, "gate_type", Integer(gate_type))
         end
 

### 2020-08-23 paper/paper.tex
paper updates

@@
 \begin{abstract}
-  This is a paper about robust quantum optimal control.
+  %% - 1 + 2 topic importance
+  %% - 1 what I have done
+  %% - few sentences about primary results
+  The ability to engineer high fidelity operations on quantum processors in the presence of
+  decoherence and systematic errors remains the primary challenge to achieving quantum advantage.
+  Quantum optimal control (QOC) techniques have been developed to mitigate deleterious effects
+  by manipulating the time-dependent parameters controlling the system. These techniques often
+  target a single error channel and may be incompatible with experimental constraints. 
+  In this work, we develop techniques to simultaneously mitigate
+  longitudinal relaxation and pure dephasing in a framework that may impose arbitrary
+  experimental constraints.
+  Optimizing a gate set for a fluxonium qubit as an example, we achieve a factor of 4 increase in
+  longitudinal relaxation times and outperform
+  existing dynamic decoupling methods in mitigating pure dephasing, achieving
+  high fidelity gates in the presence of parameter spreads up to $5\%$.
 \end{abstract}
@@
 \section{Introduction}
-(Existing work) The leading model of universal
-quantum computation is gate-based. The Quantum Optimal Control (QOC)
-literature has saught to construct techniques for engineering waveforms to manipulate
-quantum systems with high fidelity.
-Analytic methods
-have been demonstrated to construct quantum gates
-\cite{zhang2020universal, huang2020engineering, han2020experimental,
-  xu2020nonadiabatic, carlini2005quantum}.
-Most methods rely on solving the time-dependent
+%% P1 - general field, lots of citation. cite reviews, original qoc.
+%% P2 - exhaustively find all papers, what people have done to address the issues
+%% P3 - what we do is. maybe add a rant about how
+%% QOC is classical control theory but we've ignored that field
+%% P4 - outline
+(The field) The field of quantum optimal control (QOC) is concerned
+with improving the efficiency and accuracy with which quantum systems
+are manipulated.
+Early QOC techniques were proposed for nuclear magnetic resonance experiments
+\cite{khaneja2005optimal}, and applications now include superconducting
+qubits \cite{heeres2017implementing,
+  leng2019robust, leung2017speedup, xu2020nonadiabatic},
+neutral atoms and ion traps \cite{van2016optimal}, nitrogen-vacancy centers in
+diamonds \cite{rembold2020introduction}, and Bose-Einstien condensates
+\cite{sorensen2018quantum}.
+In quantum control theory, as in classical control theory,
+optimization is performed
+to minimize an objective function. Relevant for the field
+of quantum computation is the task of achieving high fidelity quantum gates.
+In this application, the objective function includes contributions from the
+gate error associated with the state transfer of the system as well as contributions
+due to the decoherence of the system, such as the evolution time of the gate.
+The decision variables of the optimization problem are the time-dependent control
+parameters, particular to the quantum system, that govern its evolution.
+
+(Existing work) Many analytic and numerical techniques have been developed to control
+quantum systems. Analytic methods rely on solving the time-dependent
 Schroedinger equation analytically or deriving equations of motion
-from a suitable Lagrangian.
-
-Some analytical methods
-have been presented to address realistic experimental constraints.
+from a suitable Lagrangian \cite{zhang2020universal, huang2020engineering, han2020experimental,
+  xu2020nonadiabatic, carlini2005quantum}. Analytic methods
+have been presented to address decoherence in quantum systems.
 Considering the dynamical and geometric phases of state
 evolution has led to methods for achieving
-robustness to pure dephasing in quantum systems
+robustness to dynamical errors and mitigating pure dephasing
 \cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
-Additionally, suitable choices of bases have been demonstrated that
-allow the experimentalist to infer tradeoffs in longitudinal
+Additionally, suitable choices of bases have been proposed that
+allow the experimentalist to infer tradeoffs in longitudinal relaxation
 and pure dephasing coherence times \cite{huang2020engineering}.
 To date, however, no analytical method has been presented
-to construct a gate while satisfying all relevant experimental
+to construct a gate while satisfying arbitrary experimental
 constraints. Furthermore, most analytic methods have only
@@
 Numerical quantum control techniques have developed in parallel
-with the analytic frameworks. Numerical quantum control
-techniques integrate the time-dependent Schroedinger
-equation and use sensitivity analysis to optimize target metrics
-of the quantum gate, such as fidelity
-\cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
+with analytic frameworks.
+Most numerical techniques proposed for QOC are indirect shooting
+methods where an initial guess for the control trajectory is forward-simulated,
+e.g. under the time-dependent-Schroedinger-equation,
+and first-order gradients are computed for a functional evaluated
+on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
   abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
-These techniques typically employ zeroth- or first-order
-optimizers and make no garauntees of optimality.
-Further, they formulate the quantum control
+Most techniques formulate the
 problem as unconstrained or use
@@
 bases for constraint spaces--which is not always feasible--and
-may hinder convergence.
+may hinder convergence. Methods for mitigating decoherence using numerical
+techniques have employed master equation dynamics (citation needed),
+Monte Carlo style trajectories \cite{abdelhafez2019gradient},
+and proposals have been made to simulate multiple trajectories
+\cite{reinhold2019controlling, rembold2020introduction}.
 
-(This work) We employ the trajectory optimization
-literature to formulate the quantum
+(This work) In this work, we employ state of the art
+trajectory optimization methods
+to formulate the quantum
 optimal control problem as a constrained optimization problem.
@@
 constraints on the evolution of the quantum state. Additionally,
-we introduce methods for constructing gates robust to longitudinal
-relaxation and pure dephasing, the dominant barriers to experimentally
+we introduce methods for constructing gates that mitigate longitudinal
+relaxation and are robust to experimental errors, and therefore
+pure dephasing, the dominant barriers to experimentally
 realizing practical quantum computing.
 
-We study the quantum optimal control problem on the fluxonium qubit.
-We outline experimentally realistic constraints and map them to
-the trajectory optimization framework. For the device we study
-we achieve a 2.5x decrease in longitudinal relaxation probability.
-We present two methods for achieving robustness to pure dephasing,
+(Outline) As an example, we study the quantum optimal control problem on
+the fluxonium qubit presented in \cite{zhang2020universal}.
+We describe experimentally realistic constraints and map them to
+the trajectory optimization framework. Next, we
+outline a method for making the optimization longitudinal
+relaxation aware. We achieve a 4x decrease in longitudinal relaxation times
+over the baseline gate set.
+We present two methods for achieving robustness to pure dephasing
 and compare them to existing dynamic decoupling methods. We find that
-our methods mitigate dephasing by order X over dynamic decoupling.
-
-(Outline)  First we formulate the quantum optimal control problem
-in the trajectory optimization framework. Then, we introduce the dynamics
-of the fluxonium device and outline experimental considerations
-relevant to gate construction. Next we outline a method for
-making the optimization longitudinal relaxation aware. Finally, we present
-some methods for engineering robustness to pure dephasing and
-compare them to existing techniques.
+the methods we employ outperform dynamic decoupling, producing
+high fidelity gates at system parameter spreads of $5\%$.
 
@@
 %% F4.1
-%% TODO: Update images. Use new gates constructed with
-%% spline T1 and use smaller sim times.
-\begin{figure*}[ht]
-  \begin{subfigure}{.3\linewidth}
-    \includegraphics[width=\linewidth]{assets/00045_spin16.png}
+\begin{figure}[ht]
+  \begin{subfigure}{\linewidth}
+    \includegraphics[width=\linewidth]{assets/f1a.png}
   \end{subfigure}
-  \begin{subfigure}{.3\linewidth}
-    \includegraphics[width=\linewidth]{assets/00047_spin16.png}
+  \begin{subfigure}{\linewidth}
+    \includegraphics[width=\linewidth]{assets/f1b.png}
   \end{subfigure}
-  \begin{subfigure}{.3\linewidth}
-    \includegraphics[width=\linewidth]{assets/00049_spin16.png}
+  \begin{subfigure}{\linewidth}
+    \includegraphics[width=\linewidth]{assets/f1c.png}
   \end{subfigure}
-  \caption{Master equation simulation with $T_{1}$ dissipation for all basis gates,
-    comparing the performance of the analytic gates and the $T_{1}$ optimized QOC gates.}
-\end{figure*}
+  \caption{(a) $T_{1}$ optimized gates (orange) and analytic gates (blue). (b) Master equation simulation with $T_{1}$ dissipation
+    for all gates, over the course of concatenated gate applications. (c) $T_{1}$ interpolation function used in optimization
+    and the corresponding $T_{1}$ times for the average absolute amplitude of each gate.}
+\end{figure}
 
@@
 %% S4
-\section{Robustness to Longitudinal Relaxation}
+\section{Longitudinal Relaxation Awareness}
 (Strategy) We seek to minimize the probability
@@
 %% and as good flux noise mitigation.
+
 \begin{table}[ht]
@@
     \hline
-    Z/2 & 5.745 & 2.144 & 2.680\\
-    Y/2 & 5.253 & 2.011 & 2.612\\
-    X/2 & 16.251 & 6.234 & 2.607\\
+    Z/2 & 5.745 & 1.469 & 3.911\\
+    Y/2 & 5.253 & 1.400 & 3.752\\
+    X/2 & 16.251 & 6.946 & 2.340\\
   \end{tabular}
@@
 %% F5.1
-%% TODO: Make this figure for real. Add inset to second figure
-%% to show high performing variants at high time steps
-\begin{figure}
+\begin{figure}[ht]
   \begin{subfigure}{\linewidth}
     \label{fig:5.1:a}
-    \includegraphics[width=\linewidth]{assets/00005_spin12_hpsweep.png}
+    \includegraphics[width=\linewidth]{assets/f2a.png}
   \end{subfigure}
@@
   \begin{subfigure}{\linewidth}
-    \includegraphics[width=\linewidth]{assets/00006_spin12_hpsweep.png}
     \label{fig:5.1:b}
+    \includegraphics[width=\linewidth]{assets/f2b.png}
   \end{subfigure}
-  \caption{\centering Robustness to qubit frequency detuning for
-    the (2, 4)-point sampling method, (2\textsuperscript{nd}, 3\textsuperscript{rd})
-    order derivative method, and CORPSE pulses for the $R_{x}(\pi/2)$ gate.
-    (a) Gate fidelity as a function of the qubit frequency detuning.
-    (b) Gate fidelity at one standard deviation from the nominal qubit frequency
-    as a function of time.
+
+  \begin{subfigure}{\linewidth}
+    \includegraphics[width=\linewidth]{assets/f2c.png}
+    \label{fig:5.1:c}
+  \end{subfigure}
+
+  \caption{\centering (a) $X/2$ gates constructed from the analytic, (2, 4)-point
+    sampling, and (2, 3)-derivative robustness methods at a gate duration of $56.8 (ns)$.
+    (b) Gate error as a function of qubit frequency detuning from the nominal frequency.
+    (c) Average gate error at one standard deviation from the nominal qubit frequency as
+    a function fo the gate duration.
   }
@@
 %% TODO: Make this figure.
-\begin{figure}
+\begin{figure}[ht]
   \includegraphics[width=\linewidth]{assets/t2_temp.jpg}
@@
 %% S5
-\section{Robustness to Pure Dephasing}
+\section{Robustness to Parameter Deviations}
 (Strategy) For the fluxonium, pure dephasing acts

### 2020-09-01 src/rbqoc.jl
most things are positive

@@ -55,7 +55,7 @@ const IT_RDI = Dict(
 
 
 # methods
-function generate_save_file_path(extension, save_file_name, save_path)
+function generate_file_path(extension, save_file_name, save_path)
     # Ensure the path exists.
     mkpath(save_path)
 

### 2020-09-01 src/spin/figures.jl

@@ -13,7 +13,8 @@ include(joinpath(WDIR, "src", "spin", "spin.jl"))
 # Configure paths.
 const EXPERIMENT_NAME = "figures"
 const SAVE_PATH = joinpath(SPIN_OUT_PATH, EXPERIMENT_NAME)
-const F1B_DATA_PATH = joinpath(SAVE_PATH, "f1b.h5")
+const F1B_DATA_FILE_PATH = joinpath(SAVE_PATH, "f1b.h5")
+const F3C_DATA_FILE_PATH = joinpath(SAVE_PATH, "f3c.h5")
 
 # Configure plotting.
 ENV["GKSwstype"] = "nul"
@@ -273,7 +274,7 @@ function make_figure1c(;save=false)
     Plots.scatter!(amps_data, t1s_data, yerror=t1s_data_err, label=nothing, marker=(:circle, MS_DATA),
                    color=:mediumorchid)
     if save
-        h5open(F1B_DATA_PATH, "w") do save_file
+        h5open(F1B_DATA_FILE_PATH, "w") do save_file
             write(save_file, "amps_fit", amps_fit)
             write(save_file, "t1s_fit", t1s_fit)
             write(save_file, "amps_data", amps_data)
@@ -449,6 +450,23 @@ function make_figure2b()
     println("Plotted Figure2b to $(plot_file_path)")
 end
 
+
+function f2b_sweep(;save_file_path=nothing, save_type=jl)
+    fq_devs = Array(range(-F2B_FQ_DEV, stop=F2B_FQ_DEV, length=F2B_TRIAL_COUNT))
+    fqs = (fq_devs .* FQ) .+ FQ
+    negi_h0s = [NEGI_H0_ISO * fq for fq in fqs]
+    gate_type=xpiby2
+    data_file_path = run_sim_h0sweep_deqjl(
+        gate_type, negi_h0s; save_file_path=save_file_path,
+        dynamics_type=schroed, save_type=save_type, dt=1e-3
+    )
+    h5open(data_file_path, "r+") do data_file
+        write(data_file, "fqs", fqs)
+    end
+    println("datafp: $(data_file_path)")
+end
+
+
[…]

### 2020-09-01 src/spin/figures.py

@@ -20,6 +20,7 @@ META_NAME = "spin"
 EXPERIMENT_NAME = "figures"
 SPIN_OUT_PATH = os.path.join(WDIR, "out", META_NAME)
 SAVE_PATH = os.path.join(WDIR, "out", META_NAME, EXPERIMENT_NAME)
+F3C_DATA_FILE_PATH = os.path.join(SAVE_PATH, "f3c.h5")
 
 # simulation
 DT_PREF = 1e-2
@@ -58,7 +59,7 @@ def grab_controls(save_file_path, save_type=SaveType.jl):
 #ENDDEF
 
 
-def generate_save_file_path(extension, save_file_name, save_path):
+def generate_file_path(extension, save_file_name, save_path):
     # Ensure the path exists.
     os.makedirs(save_path, exist_ok=True)
     
@@ -101,6 +102,8 @@ class PulseType(Enum):
     s4 = 4
     d2 = 5
     d3 = 6
+    s21 = 7
+    s41 = 8
 #ENDDEF
 
 PT_STR = {
@@ -110,6 +113,8 @@ PT_STR = {
     PulseType.s4: "S-4",
     PulseType.d2: "D-2",
     PulseType.d3: "D-3",
+    PulseType.s21: "S-21",
+    PulseType.s41: "S-41",
 }
 
 PT_COLOR = {
@@ -117,8 +122,10 @@ PT_COLOR = {
     PulseType.qoc: "red",
     PulseType.s2: "lime",
     PulseType.s4: "green",
-    PulseType.d2: "crimson",
-    PulseType.d3: "firebrick",
+    PulseType.d2: "red",
+    PulseType.d3: "darkred",
+    PulseType.s21: "black",
+    PulseType.s41: "black",
 }
 
 PT_LS = {
@@ -128,6 +135,8 @@ PT_LS = {
     PulseType.s4: "dashed",
     PulseType.d2: "solid",
     PulseType.d3: "dashed",
+    PulseType.s21: "solid",
+    PulseType.s41: "dashed",
 }
 
 DATAFP_KEY = 1
@@ -139,15 +148,19 @@ AVGT1_KEY = 6
 
 # GENERAL #
 def plot_fidelity_by_gate_count(
-        fidelitiess, inds=None, title="", ylim=(0, 1),
-        yticks=np.arange(0, 1.1, 0.1),
+        fidelitiess, inds=None, title="", ylim=None,
+        yticks=None,
[…]

### 2020-09-01 src/spin/fluxonium.jl

@@ -51,6 +51,18 @@ function fbfq_hamiltonian(fbfq)
 end
 
 
+function fbfq_gphie(fbfq)
+    hamiltonian = fbfq_hamiltonian(fbfq)
+    evals, evecs = eigen(Hermitian(hamiltonian))
+    gphie = evecs[:, 1]' * PHI_OP * evecs[:, 2]
+    return gphie
+end
+
+
+# derivative of the 2-pi reduced amplitude w.r.t the external flux by flux quantum
+const DA_DFBFQ = 4 * pi * EL * fbfq_gphie(0.5)
+
+
 function fbfq_fq(fbfq)
     hamiltonian = fbfq_hamiltonian(fbfq)
     eigvals_ = eigvals(Hermitian(hamiltonian))

### 2020-09-01 src/spin/spin.jl

@@ -48,6 +48,7 @@ end
     lindbladdf = 16
     zpiby2da = 17
     schroedda = 18
+    empty = 19
 end
 
 @enum StateType begin
@@ -77,6 +78,7 @@ const DT_ST = Dict(
     xpiby2da => st_state,
     xpiby2t2 => st_density,
     zpiby2da => st_state,
+    empty => st_state,
 )
 
 const GT_STR = Dict(
@@ -109,17 +111,16 @@ const E_PLASMA = sqrt(8 * EL * EC)
 const PHI_OSC = (8 * EC / EL)^(1//4)
 # qubit frequency at flux frustration point
 const FQ = 1.4e-2 #GHz
-const SIGMAFQ = FQ * 5e-2
-const S1FQ = FQ + SIGMAFQ
-const S2FQ = FQ - SIGMAFQ
-const S3FQ = FQ + 2 * SIGMAFQ
-const S4FQ = FQ - 2 * SIGMAFQ
+const SP1FQ = FQ + FQ * 1e-2
+const SN1FQ = FQ - FQ * 1e-2
+const SP2FQ = FQ + FQ * 2e-2
+const SN2FQ = FQ - FQ * 2e-2
 const MAX_CONTROL_NORM_0 = 5e-1 #GHz
 const FBFQ_A = 0.202407
 const FBFQ_B = 0.5
 const AYPIBY2 = 1.25e-1 #GHz
 const FBFQ_NAMP = 5.21e-6 # flux noise amplitude
-const FBFQ_NDIST = Normal(0., 1.)
+const STD_NORMAL = Normal(0., 1.)
 const GAMMAC = 1 / 3e5 #GHz(T_c = 300 us)
 const SQRTLNIR = 4
 const NAMP_PREFACTOR = FBFQ_NAMP / FBFQ_A
@@ -187,10 +188,10 @@ const SIGMAZ_ISO = SA_F64[1   0   0   0;
 const NEGI_H0_ISO = pi * NEGI * SIGMAZ_ISO
 const NEGI_H1_ISO = pi * NEGI * SIGMAX_ISO
 const FQ_NEGI_H0_ISO = FQ * NEGI_H0_ISO
-const S1FQ_NEGI_H0_ISO = S1FQ * NEGI_H0_ISO
[…]

### 2020-09-01 src/spin/spin11.jl

@@ -20,13 +20,10 @@ const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 
 # Define the optimization.
 const CONTROL_COUNT = 1
-const DORDER = 3
-const DT_STATIC = DT_PREF
-const DT_STATIC_INV = DT_PREF_INV
 const CONSTRAINT_TOLERANCE = 1e-8
 const AL_KICKOUT_TOLERANCE = 1e-7
-const PN_STEPS = 5
-const MAX_PENALTY = 1e10
+const PN_STEPS = 2
+const MAX_PENALTY = 1e11
 const ILQR_DJ_TOL = 1e-4
 
 # Define the problem.
@@ -133,7 +130,8 @@ end
 function run_traj(;gate_type=xpiby2, evolution_time=60., solver_type=altro,
                   postsample=false, initial_save_file_path=nothing,
                   initial_save_type=jl, sqrtbp=false, derivative_order=0,
-                  integrator_type=rk6, max_penalty=MAX_PENALTY, qs=nothing)
+                  integrator_type=rk6, max_penalty=MAX_PENALTY, qs=nothing,
+                  smoke_test=false, dt=5e-3)
     model = Model(derivative_order)
     n = state_dim(model)
     m = control_dim(model)
@@ -192,8 +190,8 @@ function run_traj(;gate_type=xpiby2, evolution_time=60., solver_type=altro,
     ]
 
     if isnothing(initial_save_file_path)
-        dt = DT_STATIC
-        N = Int(floor(evolution_time * DT_STATIC_INV)) + 1
+        dt_inv = 1 / dt
+        N = Int(floor(evolution_time * dt_inv)) + 1
         U0 = [SVector{m}(
             fill(1e-4, CONTROL_COUNT)
         ) for k = 1:N-1]
@@ -209,9 +207,10 @@ function run_traj(;gate_type=xpiby2, evolution_time=60., solver_type=altro,
             end
             return (d2controls, evolution_time)
         end
-        evolution_time = Int(floor(evolution_time * DT_STATIC_INV)) * DT_STATIC
-        dt = DT_STATIC
-        N = Int(floor(evolution_time * DT_STATIC_INV)) + 1
+        dt = DT_PREF
+        dt_inv = DT_PREF_INV
+        evolution_time = Int(floor(evolution_time * dt_inv)) * dt
[…]

### 2020-09-01 src/spin/spin12.jl

@@ -24,7 +24,7 @@ const DT_STATIC_INV = DT_PREF_INV
 const CONSTRAINT_TOLERANCE = 1e-8
 const AL_KICKOUT_TOLERANCE = 1e-7
 const PN_STEPS = 5
-const MAX_PENALTY = 1e10
+const MAX_PENALTY = 1e11
 const ILQR_DJ_TOL = 1e-4
 
 # Define the problem.
@@ -81,8 +81,8 @@ function RobotDynamics.dynamics(model::Model{SO}, astate::StaticVector,
     delta_dcontrol = acontrols[D2CONTROLS_IDX]
 
     if SO == 2
-        negi_s1h = S1FQ_NEGI_H0_ISO + negi_hc
-        negi_s2h = S2FQ_NEGI_H0_ISO + negi_hc
+        negi_s1h = SP1FQ_NEGI_H0_ISO + negi_hc
+        negi_s2h = SN1FQ_NEGI_H0_ISO + negi_hc
         delta_s1state1 = negi_s1h * astate[S1STATE1_IDX]
         delta_s1state2 = negi_s1h * astate[S1STATE2_IDX]
         delta_s2state1 = negi_s2h * astate[S2STATE1_IDX]
@@ -99,10 +99,10 @@ function RobotDynamics.dynamics(model::Model{SO}, astate::StaticVector,
             delta_s2state2;
         ]
     elseif SO == 4
-        negi_s1h = S1FQ_NEGI_H0_ISO + negi_hc
-        negi_s2h = S2FQ_NEGI_H0_ISO + negi_hc
-        negi_s3h = S3FQ_NEGI_H0_ISO + negi_hc
-        negi_s4h = S4FQ_NEGI_H0_ISO + negi_hc
+        negi_s1h = SP1FQ_NEGI_H0_ISO + negi_hc
+        negi_s2h = SN1FQ_NEGI_H0_ISO + negi_hc
+        negi_s3h = SP2FQ_NEGI_H0_ISO + negi_hc
+        negi_s4h = SN2FQ_NEGI_H0_ISO + negi_hc
         delta_s1state1 = negi_s1h * astate[S1STATE1_IDX]
         delta_s1state2 = negi_s1h * astate[S1STATE2_IDX]
         delta_s2state1 = negi_s2h * astate[S2STATE1_IDX]

### 2020-09-01 paper/paper.tex
all 3 figures

@@
   The ability to engineer high fidelity operations on quantum processors in the presence of
-  decoherence and systematic errors remains the primary challenge to achieving quantum advantage.
+  decoherence and systematic errors remains the primary challenge requisite to achieving quantum advantage.
   Quantum optimal control (QOC) techniques have been developed to mitigate deleterious effects
@@
   In this work, we develop techniques to simultaneously mitigate
-  longitudinal relaxation and pure dephasing in a framework that may impose arbitrary
+  longitudinal relaxation and pure dephasing in a framework that can handle arbitrary
   experimental constraints.
@@
   longitudinal relaxation times and outperform
-  existing dynamic decoupling methods in mitigating pure dephasing, achieving
+  existing dynamic decoupling methods in mitigating pure dephasing due to flux noise, achieving
   high fidelity gates in the presence of parameter spreads up to $5\%$.
@@
 outline a method for making the optimization longitudinal
-relaxation aware. We achieve a 4x decrease in longitudinal relaxation times
+relaxation aware. We achieve a 4x increase in longitudinal relaxation times
 over the baseline gate set.
@@
 
-%% F4.1
+%% F1
 \begin{figure}[ht]
-  \begin{subfigure}{\linewidth}
+  \begin{subfigure}{.49\linewidth}
     \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}
-  \begin{subfigure}{\linewidth}
+  \end{subfigure}\hfill%
+  \begin{subfigure}{.49\linewidth}
     \includegraphics[width=\linewidth]{assets/f1b.png}
@@
 
-%% F5.1
+%% F2
 \begin{figure}[ht]
-  \begin{subfigure}{\linewidth}
+  \begin{subfigure}{.49\linewidth}
     \label{fig:5.1:a}
     \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}
-
-  \begin{subfigure}{\linewidth}
+  \end{subfigure}%
+  \begin{subfigure}{.49\linewidth}
     \label{fig:5.1:b}
@@
 
-  \caption{\centering (a) $X/2$ gates constructed from the analytic, (2, 4)-point
-    sampling, and (2, 3)-derivative robustness methods at a gate duration of $56.8 (ns)$.
+  \caption{\centering (a) $X/2$ gates robust to qubit frequency offsets
+    constructed using the analytic, (2, 4)-point
+    sampling, and (2, 3)-derivative methods at a gate duration of $56.8 (ns)$.
     (b) Gate error as a function of qubit frequency detuning from the nominal frequency.
@@
 
-%% F5.2
-%% TODO: Make this figure.
+%% F3
 \begin{figure}[ht]
-  \includegraphics[width=\linewidth]{assets/t2_temp.jpg}
-  \caption{Robustness to $T_{\phi}$-type noise for the
-    the (2, 4)-point sampling method, (2\textsuperscript{nd}, 3\textsuperscript{rd})
-    order derivative method, and B2CORPSE pulse for the $R_{x}(\pi/2)$ gate.}
+  \begin{subfigure}{.5\linewidth}
+    \includegraphics[width=\linewidth]{assets/f3a.png}
+  \end{subfigure}%
+  \begin{subfigure}{.5\linewidth}
+    \includegraphics[width=\linewidth]{assets/f3c.png}
+  \end{subfigure}
+
+  \begin{subfigure}{\linewidth}
+    \includegraphics[width=\linewidth]{assets/f3b.png}
+  \end{subfigure}
+
+  \caption{\centering (a) $X/2$ gates robust to flux offsets constructed using the analytic,
+    (2, 4)-point sampling, and (2, 3)-derivative methods at a gate duration
+    of $56.8 (ns)$. (b) Mean gate error for 10 samples of 1/$f$ flux noise over 5000
+    gate applications. (c) Stochastic 1/$f$ flux noise.}
 \end{figure}

### 2020-09-07 src/rbqoc.jl
updates

@@ -87,14 +87,17 @@ function grab_controls(save_file_path; save_type=jl)
             cidx = read(save_file, "controls_idx")
             controls = read(save_file, "astates")[:, cidx]
             evolution_time = read(save_file, "evolution_time")
+            controls_dt_inv = "dt" in names(save_file) ? read(save_file, "dt")^(-1) : DT_PREF_INV
         elseif save_type == samplejl
             controls = read(save_file, "controls_sample")
             evolution_time = read(save_file, "evolution_time_sample")
+            controls_dt_inv = DT_PREF_INV
         elseif save_type == py
             controls = permutedims(read(save_file, "controls"), (2, 1))
             evolution_time = read(save_file, "evolution_time")
+            controls_dt_inv = DT_PREF_INV
         end
-        return (controls, evolution_time)
+        return (controls, controls_dt_inv, evolution_time)
     end
 
     return data
@@ -168,6 +171,15 @@ end
 
 
 @inline get_vec_iso(vec) = vcat(real(vec), imag(vec))
+
+
+function get_vec_uniso(vec)
+    n = size(vec)[1]
+    nby2 = Integer(n/2)
+    i1 = 1:nby2
+    i2 = (nby2 + 1):n
+    return vec[i1] + 1im * vec[i2]
+end
 
 
 function get_mat_iso(mat)

### 2020-09-07 src/spin/figures.jl

@@ -14,7 +14,9 @@ include(joinpath(WDIR, "src", "spin", "spin.jl"))
 const EXPERIMENT_NAME = "figures"
 const SAVE_PATH = joinpath(SPIN_OUT_PATH, EXPERIMENT_NAME)
 const F1B_DATA_FILE_PATH = joinpath(SAVE_PATH, "f1b.h5")
+const F2C_DATA_FILE_PATH = joinpath(SAVE_PATH, "f2c.h5")
 const F3C_DATA_FILE_PATH = joinpath(SAVE_PATH, "f3c.h5")
+const F3D_DATA_FILE_PATH = joinpath(SAVE_PATH, "f3d.h5")
 
 # Configure plotting.
 ENV["GKSwstype"] = "nul"
@@ -22,46 +24,46 @@ Plots.gr()
 
 # types
 @enum PulseType begin
-    qoc = 1
-    analytic = 2
-    derivative = 3
-    sample = 4
-    derivative2 = 5
-    sample2 = 6
+    analytic = 1
+    qoc = 2
+    s2 = 3
+    s4 = 4
+    d2 = 5
+    d3 = 6
 end
 
 const PT_STR = Dict(
     qoc => "QOC",
     analytic => "Anl.",
-    sample => "S-2",
-    derivative => "D-2",
-    derivative2 => "D-3",
-    sample2 => "S-4"
+    s2 => "S-2",
+    d2 => "D-2",
+    d3 => "D-3",
+    s4 => "S-4"
 )
 
 const PT_MARKER = Dict(
-    sample => :circle,
-    sample2 => :square,
-    derivative2 => :utriangle,
-    derivative => :diamond,
+    s2 => :circle,
+    s4 => :square,
+    d3 => :utriangle,
+    d2 => :diamond,
 )
 
 const PT_COLOR = Dict(
     analytic => :lightskyblue,
     qoc => :coral,
-    sample => :limegreen,
-    sample2 => :darkgreen,
-    derivative => :crimson,
-    derivative2 => :firebrick,
+    s2 => :limegreen,
+    s4 => :darkgreen,
+    d2 => :crimson,
+    d3 => :firebrick,
 )
 
 const PT_LINESTYLE = Dict(
     analytic => :solid,
     qoc => :solid,
-    sample => :solid,
-    sample2 => :dash,
-    derivative => :solid,
-    derivative2 => :dash,
+    s2 => :solid,
+    s4 => :dash,
+    d2 => :solid,
+    d3 => :dash,
 )
 
[…]

### 2020-09-07 src/spin/figures.py

@@ -2,12 +2,14 @@
 figures.py - figures but in python
 """
 
+from argparse import ArgumentParser
 from enum import Enum
 import os
 
 import h5py
 import matplotlib.pyplot as plt
 import numpy as np
+import qutip
 
 
 ## GENERAL ##
@@ -20,7 +22,9 @@ META_NAME = "spin"
 EXPERIMENT_NAME = "figures"
 SPIN_OUT_PATH = os.path.join(WDIR, "out", META_NAME)
 SAVE_PATH = os.path.join(WDIR, "out", META_NAME, EXPERIMENT_NAME)
+F2C_DATA_FILE_PATH = os.path.join(SAVE_PATH, "f2c.h5")
 F3C_DATA_FILE_PATH = os.path.join(SAVE_PATH, "f3c.h5")
+F3D_DATA_FILE_PATH = os.path.join(SAVE_PATH, "f3d.h5")
 
 # simulation
 DT_PREF = 1e-2
@@ -29,6 +33,8 @@ DT_PREF_INV = 1e2
 # plotting
 DPI = 300
 DPI_FINAL = int(1e3)
+TICK_FS = LABEL_FS = LEGEND_FS = TEXT_FS = 10
+PAPER_LW = 3.40457
 
 # TYPES #
 
@@ -102,8 +108,10 @@ class PulseType(Enum):
     s4 = 4
     d2 = 5
     d3 = 6
-    s21 = 7
-    s41 = 8
+    s2b = 7
+    s4b = 8
+    d2b = 9
+    d3b = 10
 #ENDDEF
 
 PT_STR = {
@@ -113,8 +121,10 @@ PT_STR = {
     PulseType.s4: "S-4",
     PulseType.d2: "D-2",
     PulseType.d3: "D-3",
-    PulseType.s21: "S-21",
-    PulseType.s41: "S-41",
+    PulseType.s2b: "S-2",
+    PulseType.s4b: "S-4",
+    PulseType.d2b: "D-2",
+    PulseType.d3b: "D-2",
 }
 
 PT_COLOR = {
@@ -124,19 +134,30 @@ PT_COLOR = {
     PulseType.s4: "green",
     PulseType.d2: "red",
     PulseType.d3: "darkred",
-    PulseType.s21: "black",
-    PulseType.s41: "black",
+    PulseType.s2b: "lime",
+    PulseType.s4b: "green",
+    PulseType.d2b: "red",
+    PulseType.d3b: "darkred",
 }
 
[…]

### 2020-09-07 src/spin/spin.jl

@@ -115,6 +115,8 @@ const SP1FQ = FQ + FQ * 1e-2
 const SN1FQ = FQ - FQ * 1e-2
 const SP2FQ = FQ + FQ * 2e-2
 const SN2FQ = FQ - FQ * 2e-2
+const SP3FQ = FQ + FQ * 3e-2
+const SN3FQ = FQ - FQ * 3e-2
 const MAX_CONTROL_NORM_0 = 5e-1 #GHz
 const FBFQ_A = 0.202407
 const FBFQ_B = 0.5
@@ -173,18 +175,18 @@ const ZPIBY2_GATE_TIME = 17.86
 # ISO indicates the object is defined in the complex to real isomorphism.
 # NEGI is the negative complex unit.
 const NEGI = SA_F64[0   0  1  0 ;
-              0   0  0  1 ;
-              -1  0  0  0 ;
-              0  -1  0  0 ;]
+                    0   0  0  1 ;
+                    -1  0  0  0 ;
+                    0  -1  0  0 ;]
 # SIGMAX, SIGMAZ are the X and Z pauli matrices
 const SIGMAX_ISO = SA_F64[0   1   0   0;
-                    1   0   0   0;
-                    0   0   0   1;
-                    0   0   1   0]
+                          1   0   0   0;
+                          0   0   0   1;
+                          0   0   1   0]
 const SIGMAZ_ISO = SA_F64[1   0   0   0;
-                    0  -1   0   0;
-                    0   0   1   0;
-                    0   0   0  -1]
+                          0  -1   0   0;
+                          0   0   1   0;
+                          0   0   0  -1]
 const NEGI_H0_ISO = pi * NEGI * SIGMAZ_ISO
 const NEGI_H1_ISO = pi * NEGI * SIGMAX_ISO
 const FQ_NEGI_H0_ISO = FQ * NEGI_H0_ISO
@@ -192,6 +194,8 @@ const SP1FQ_NEGI_H0_ISO = SP1FQ * NEGI_H0_ISO
 const SN1FQ_NEGI_H0_ISO = SN1FQ * NEGI_H0_ISO
 const SP2FQ_NEGI_H0_ISO = SP2FQ * NEGI_H0_ISO
 const SN2FQ_NEGI_H0_ISO = SN2FQ * NEGI_H0_ISO
[…]

### 2020-09-07 src/spin/spin11.jl

@@ -19,6 +19,7 @@ const EXPERIMENT_NAME = "spin11"
 const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 
 # Define the optimization.
+const DT_STATIC = 5e-3
 const CONTROL_COUNT = 1
 const CONSTRAINT_TOLERANCE = 1e-8
 const AL_KICKOUT_TOLERANCE = 1e-7
@@ -76,8 +77,20 @@ function RobotDynamics.dynamics(model::Model{DO}, astate::StaticVector,
     delta_intcontrol = astate[CONTROLS_IDX]
     delta_control = astate[DCONTROLS_IDX]
     delta_dcontrol = acontrols[D2CONTROLS_IDX]
-
-    if DO == 2
+    
+    if DO == 1
+        delta_dstate1 = NEGI_H0_ISO * astate[STATE1_IDX] + negi_h * astate[DSTATE1_IDX]
+        delta_dstate2 = NEGI_H0_ISO * astate[STATE2_IDX] + negi_h * astate[DSTATE2_IDX]
+        delta_astate = [
+            delta_state1;
+            delta_state2;
+            delta_intcontrol;
+            delta_control;
+            delta_dcontrol;
+            delta_dstate1;
+            delta_dstate2;
+        ]
+    elseif DO == 2
         delta_dstate1 = NEGI_H0_ISO * astate[STATE1_IDX] + negi_h * astate[DSTATE1_IDX]
         delta_dstate2 = NEGI_H0_ISO * astate[STATE2_IDX] + negi_h * astate[DSTATE2_IDX]
         delta_d2state1 = 2 * NEGI_H0_ISO * astate[DSTATE1_IDX] + negi_h * astate[D2STATE1_IDX]
@@ -127,11 +140,11 @@ function RobotDynamics.dynamics(model::Model{DO}, astate::StaticVector,
 end
 
 
-function run_traj(;gate_type=xpiby2, evolution_time=60., solver_type=altro,
+function run_traj(;gate_type=xpiby2, evolution_time=60., solver_type=alilqr,
                   postsample=false, initial_save_file_path=nothing,
                   initial_save_type=jl, sqrtbp=false, derivative_order=0,
                   integrator_type=rk6, max_penalty=MAX_PENALTY, qs=nothing,
-                  smoke_test=false, dt=5e-3)
+                  smoke_test=false, dt=DT_STATIC)
[…]

### 2020-09-07 src/spin/spin12.jl

@@ -19,11 +19,9 @@ const EXPERIMENT_NAME = "spin12"
 const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 
 # Define the optimization.
-const DT_STATIC = DT_PREF
-const DT_STATIC_INV = DT_PREF_INV
 const CONSTRAINT_TOLERANCE = 1e-8
 const AL_KICKOUT_TOLERANCE = 1e-7
-const PN_STEPS = 5
+const PN_STEPS = 2
 const MAX_PENALTY = 1e11
 const ILQR_DJ_TOL = 1e-4
 
@@ -140,11 +138,12 @@ function RobotDynamics.dynamics(model::Model{SO}, astate::StaticVector,
 end
 
 
-function run_traj(;gate_type=xpiby2, evolution_time=60., solver_type=altro,
+function run_traj(;gate_type=xpiby2, evolution_time=60., solver_type=alilqr,
                   postsample=false, sqrtbp=false, sample_order=0,
-                  integrator_type=rk6, qs=nothing, max_penalty=MAX_PENALTY)
-    dt = DT_STATIC
-    N = Int(floor(evolution_time * DT_STATIC_INV)) + 1
+                  integrator_type=rk6, qs=nothing, max_penalty=MAX_PENALTY,
+                  dt=5e-3, smoke_test=false)
+    dt_inv = dt^(-1)
+    N = Int(floor(evolution_time * dt_inv)) + 1
     model = Model(sample_order)
     n = state_dim(model)
     m = control_dim(model)
@@ -253,10 +252,14 @@ function run_traj(;gate_type=xpiby2, evolution_time=60., solver_type=altro,
     if solver_type == alilqr
         solver = AugmentedLagrangianSolver(prob, opts)
         solver.solver_uncon.opts.square_root = sqrtbp
-        solver.opts.constraint_tolerance = CONSTRAINT_TOLERANCE
-        solver.opts.constraint_tolerance_intermediate = CONSTRAINT_TOLERANCE
+        solver.opts.constraint_tolerance = AL_KICKOUT_TOLERANCE
+        solver.opts.constraint_tolerance_intermediate = AL_KICKOUT_TOLERANCE
         solver.opts.cost_tolerance_intermediate = ILQR_DJ_TOL
         solver.opts.penalty_max = max_penalty
+        if smoke_test
+            solver.opts.iterations = 1
+            solver.solver_uncon.opts.iterations = 1
+        end
     elseif solver_type == altro
[…]

### 2020-09-07 src/spin/spin15.jl

@@ -27,8 +27,8 @@ const DT_MIN = DT_INIT / 2
 const DT_MAX = DT_INIT * 2
 const CONSTRAINT_TOLERANCE = 1e-8
 const AL_KICKOUT_TOLERANCE = 1e-7
-const PN_STEPS = 5
-const MAX_PENALTY = 1e10
+const PN_STEPS = 2
+const MAX_PENALTY = 1e11
 const ILQR_DJ_TOL = 1e-4
 
 # Define the problem.
@@ -63,8 +63,8 @@ const EMPTY_V = []
 struct Model{DA, TO} <: AbstractModel
     Model(DA::Bool=true, TO::Bool=true) = new{DA, TO}()
 end
-RobotDynamics.state_dim(::Model{false, TO}) where TO = ASTATE_SIZE_NODA
-RobotDynamics.state_dim(::Model{true, TO}) where TO = ASTATE_SIZE_NODA + 1
+RobotDynamics.state_dim(::Model{false, TO}) where TO = ASTATE_SIZE_BASE
+RobotDynamics.state_dim(::Model{true, TO}) where TO = ASTATE_SIZE_BASE + 1
 RobotDynamics.control_dim(::Model{DA, false}) where DA = CONTROL_COUNT
 RobotDynamics.control_dim(::Model{DA, true}) where DA = CONTROL_COUNT + 1
 
@@ -105,7 +105,8 @@ function run_traj(;evolution_time=20., gate_type=zpiby2,
                   decay_aware=false,
                   solver_type=altro, sqrtbp=false,
                   integrator_type=rk6, max_penalty=MAX_PENALTY,
-                  qs=[1e0, 1e0, 1e0, 1e-1, 5e1, 1e-1, 1e2])
+                  qs=[1e0, 1e0, 1e0, 1e-1, 5e1, 1e-1, 1e2],
+                  smoke_test=false)
     # Convert to trajectory optimization language.
     model = Model(decay_aware, time_optimal)
     n = state_dim(model)
@@ -117,7 +118,7 @@ function run_traj(;evolution_time=20., gate_type=zpiby2,
         zeros(CONTROL_COUNT); # int_control
         zeros(CONTROL_COUNT); # control
         zeros(CONTROL_COUNT); # dcontrol_dt
-        eval(:($decay_aware ? zeros(1) : EMPTY_V))
+        zeros(eval(:($decay_aware ? 1 : 0)));
     ])
 
[…]

### 2020-09-07 paper/paper.tex

@@
   The ability to engineer high fidelity operations on quantum processors in the presence of
-  decoherence and systematic errors remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have been developed to mitigate deleterious effects
-  by manipulating the time-dependent parameters controlling the system. These techniques often
-  target a single error channel and may be incompatible with experimental constraints. 
-  In this work, we develop techniques to simultaneously mitigate
-  longitudinal relaxation and pure dephasing in a framework that can handle arbitrary
-  experimental constraints.
-  Optimizing a gate set for a fluxonium qubit as an example, we achieve a factor of 4 increase in
-  longitudinal relaxation times and outperform
-  existing dynamic decoupling methods in mitigating pure dephasing due to flux noise, achieving
-  high fidelity gates in the presence of parameter spreads up to $5\%$.
+  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
+  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
+  but they require exquisite calibration to be performant. In this work we employ robust control techniques
+  to achieve high fidelity gates in the presence of parameter spreads up to 5$\%$.
+  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
+  Lindblad master equation dynamics, which drastically increase the computational complexity
+  of the optimization algorithm.
+  In this work we achieve a factor of $4$ increase in longitudinal relaxation times
+  for the gate set of a fluxonium qubit
+  using a computationally efficient metric in optimizaiton.
+
+  %% The ability to engineer high fidelity operations on quantum processors in the presence of
+  %% systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
+  %% Quantum optimal control (QOC) techniques have been developed to mitigate deleterious effects
+  %% by manipulating the time-dependent parameters controlling the system. These techniques often
+  %% target a single error channel and may be incompatible with experimental constraints. 
+  %% In this work, we develop techniques to simultaneously mitigate
+  %% longitudinal relaxation and pure dephasing in a framework that can handle arbitrary
+  %% experimental constraints.
+  %% Optimizing a gate set for a fluxonium qubit as an example, we achieve a factor of 4 increase in
+  %% longitudinal relaxation times and outperform
+  %% existing dynamic decoupling methods in mitigating pure dephasing due to flux noise, achieving
+  %% high fidelity gates in the presence of parameter spreads up to $5\%$.
 \end{abstract}
@@
 \section{Introduction}
+textwidth: \printinunitsof{in}\prntlen{\textwidth}
+linewidth: \printinunitsof{in}\prntlen{\linewidth}
 %% P1 - general field, lots of citation. cite reviews, original qoc.
@@
   leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral atoms and ion traps \cite{van2016optimal}, nitrogen-vacancy centers in
-diamonds \cite{rembold2020introduction}, and Bose-Einstien condensates
+neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
+diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
 \cite{sorensen2018quantum}.
@@
 gate error associated with the state transfer of the system as well as contributions
-due to the decoherence of the system, such as the evolution time of the gate.
+due to experimental constraints, such as the control amplitude.
 The decision variables of the optimization problem are the time-dependent control
@@
 methods where an initial guess for the control trajectory is forward-simulated,
-e.g. under the time-dependent-Schroedinger-equation,
+e.g. under the time-dependent Schroedinger equation,
 and first-order gradients are computed for a functional evaluated
@@
 to formulate the quantum
-optimal control problem as a constrained optimization problem.
-The trajectory optimization framework allows us to handle arbitrary
-constraints on the evolution of the quantum state. Additionally,
+optimal control problem as a constrained optimization problem, allowing
+us to enforce experimental constraints.
+Additionally,
 we introduce methods for constructing gates that mitigate longitudinal
@@
 \begin{figure}[ht]
-  \begin{subfigure}{.49\linewidth}
+  \begin{subfigure}{.6\linewidth}
     \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill%
-  \begin{subfigure}{.49\linewidth}
+  \end{subfigure}%
+    \begin{subfigure}{.4\linewidth}
     \includegraphics[width=\linewidth]{assets/f1b.png}
   \end{subfigure}
+
   \begin{subfigure}{\linewidth}
@@
   \end{subfigure}
-  \caption{(a) $T_{1}$ optimized gates (orange) and analytic gates (blue). (b) Master equation simulation with $T_{1}$ dissipation
-    for all gates, over the course of concatenated gate applications. (c) $T_{1}$ interpolation function used in optimization
-    and the corresponding $T_{1}$ times for the average absolute amplitude of each gate.}
+  \caption{
+    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
+    (b) $T_{1}$ interpolation function used in optimization
+    and the corresponding $T_{1}$ times for the average absolute amplitude of each gate.
+    (c) Master equation simulation with $T_{1}$ dissipation
+    for all gates, over the course of concatenated gate applications.
+  }
 \end{figure}
@@
 
-(Results) We achieve a factor of 2.5 decrease in the probability
+(Results) We achieve a factor of 5 decrease in the probability
 of longitudinal relaxation from the analytic gates we benchmark
@@
 
-%% T4.1
-%% TODO: Maybe best for appendix.
-%% TOOD: Maybe add some commentary about why we see more speedup
-%% for Z/2 and X/2 than Y/2. In particular I think this is because of the
-%% long idle times at the flux frustration point for Z/2 and X/2. It
-%% might be a more fair comparison to show gates that have better T1s
-%% and as good flux noise mitigation.
-
+%% T1
+%% TODO: maybe for appendix.
 \begin{table}[ht]
@@
     & Analytic  & QOC &\\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & Speedup\\
+    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & Ratio\\
     \hline
-    Z/2 & 5.745 & 1.469 & 3.911\\
-    Y/2 & 5.253 & 1.400 & 3.752\\
+    Z/2 & 5.745 & 1.149 & 5.000\\
+    Y/2 & 5.253 & 1.157 & 4.540\\
     X/2 & 16.251 & 6.946 & 2.340\\
@@
 \begin{figure}[ht]
-  \begin{subfigure}{.49\linewidth}
-    \label{fig:5.1:a}
+  \begin{subfigure}{.6\linewidth}
     \includegraphics[width=\linewidth]{assets/f2a.png}
   \end{subfigure}%
-  \begin{subfigure}{.49\linewidth}
-    \label{fig:5.1:b}
+  \begin{subfigure}{.4\linewidth}
     \includegraphics[width=\linewidth]{assets/f2b.png}
@@
     \includegraphics[width=\linewidth]{assets/f2c.png}
-    \label{fig:5.1:c}
   \end{subfigure}
@@
 \begin{figure}[ht]
-  \begin{subfigure}{.5\linewidth}
+  \begin{subfigure}{\linewidth}
     \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}%
-  \begin{subfigure}{.5\linewidth}
-    \includegraphics[width=\linewidth]{assets/f3c.png}
   \end{subfigure}
@@
 
-  \caption{\centering (a) $X/2$ gates robust to flux offsets constructed using the analytic,
+  %% \begin{subfigure}{.5\linewidth}
+  %%   \includegraphics[width=\linewidth]{assets/f3c.png}
+  %% \end{subfigure}
+
+  \caption{\centering
+    (a) $X/2$ gates robust to flux offsets constructed using the analytic,
     (2, 4)-point sampling, and (2, 3)-derivative methods at a gate duration
     of $56.8 (ns)$. (b) Mean gate error for 10 samples of 1/$f$ flux noise over 5000
-    gate applications. (c) Stochastic 1/$f$ flux noise.}
+    gate applications.
+    %% (c) Stochastic 1/$f$ flux noise.
+  }
 \end{figure}

### 2020-09-07 paper/paper.tex
updated figures and abstract

@@
   The ability to engineer high fidelity operations on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
+  calibration errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
   Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
   but they require exquisite calibration to be performant. In this work we employ robust control techniques
-  to achieve high fidelity gates in the presence of parameter spreads up to 5$\%$.
+  to achieve gate errors two orders of magnitude lower than dynamic decoupling for parameter
+  spreads on the order of $1\%$. 
   Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  Lindblad master equation dynamics, which drastically increase the computational complexity
+  Lindblad master equation dynamics, which exponentially increase the computational complexity
   of the optimization algorithm.
-  In this work we achieve a factor of $4$ increase in longitudinal relaxation times
+  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
   for the gate set of a fluxonium qubit
   using a computationally efficient metric in optimizaiton.
-
-  %% The ability to engineer high fidelity operations on quantum processors in the presence of
-  %% systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  %% Quantum optimal control (QOC) techniques have been developed to mitigate deleterious effects
-  %% by manipulating the time-dependent parameters controlling the system. These techniques often
-  %% target a single error channel and may be incompatible with experimental constraints. 
-  %% In this work, we develop techniques to simultaneously mitigate
-  %% longitudinal relaxation and pure dephasing in a framework that can handle arbitrary
-  %% experimental constraints.
-  %% Optimizing a gate set for a fluxonium qubit as an example, we achieve a factor of 4 increase in
-  %% longitudinal relaxation times and outperform
-  %% existing dynamic decoupling methods in mitigating pure dephasing due to flux noise, achieving
-  %% high fidelity gates in the presence of parameter spreads up to $5\%$.
 \end{abstract}
@@
 \begin{figure}[ht]
-  \begin{subfigure}{.6\linewidth}
+  \begin{subfigure}{.55\linewidth}
     \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}%
+  \end{subfigure}\hfill
   \begin{subfigure}{.4\linewidth}
@@
     (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization
-    and the corresponding $T_{1}$ times for the average absolute amplitude of each gate.
+    (b) $T_{1}$ interpolation function used in optimization. Markers
+    denote the time-averaged, absolute amplitude of each gate.
     (c) Master equation simulation with $T_{1}$ dissipation
-    for all gates, over the course of concatenated gate applications.
+    over the course of concatenated gate applications.
   }
@@
 \begin{align}
-  H/h &= \omega_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
+  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
 \end{align}
-where $\omega_{q}$ is the qubit frequency at the flux frustration point,
+where $f_{q}$ is the qubit frequency at the flux frustration point,
 $a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
@@
     & Analytic  & QOC &\\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & Ratio\\
+    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$\\
     \hline
@@
 \begin{figure}[ht]
-  \begin{subfigure}{.6\linewidth}
+  \begin{subfigure}{.55\linewidth}
     \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}%
+  \end{subfigure}\hfill
   \begin{subfigure}{.4\linewidth}
@@
 
-  \caption{\centering (a) $X/2$ gates robust to qubit frequency offsets
-    constructed using the analytic, (2, 4)-point
-    sampling, and (2, 3)-derivative methods at a gate duration of $56.8 (ns)$.
-    (b) Gate error as a function of qubit frequency detuning from the nominal frequency.
-    (c) Average gate error at one standard deviation from the nominal qubit frequency as
-    a function fo the gate duration.
+  \caption{
+    (a) X/2 gates robust to qubit frequency detunings constructed using the
+    analytic, (2, 4)-point sampling, and (2\textsuperscript{nd},
+    3\textsuperscript{rd})-order derivative methods.
+    (b) Gate error as a function of the detuning from the nominal
+    qubit frequency at a fixed gate duration, see (a), and at the best
+    gate duration for each method chosen from a sweep, see (c).
+    (c) Gate error at a one-percent detuning from the nominal qubit frequency as
+    a function of the gate duration set in optimization.
   }
@@
 
-  \caption{\centering
+  \caption{
     (a) $X/2$ gates robust to flux offsets constructed using the analytic,
@@
 the qubit frequency to deviate from its measured value
-by an amount $\delta \omega_{q}$.
+by an amount $\delta f_{q}$.
 
@@
 propagate $\ket{\psi^{\pm, I}(t)}$ where
-$\omega_{q}^{\pm} = \omega_{q} \pm \sigma_{\omega_{q}}$,
-$\omega_{q}^{I} = \omega_{q}$.
+$f_{q}^{\pm} = f_{q} \pm \sigma_{f_{q}}$,
+$f_{q}^{I} = f_{q}$.
 The difference between the final states and the target
@@
 subject to qubit frequency detuning
-($\omega_{q} \gets \omega_{q} + \delta \omega_{q}$)
+($f_{q} \gets f_{q} + \delta f_{q}$)
 to demonstrate the applicability of our techniques to mitigate
@@
 mitigate error due to qubit frequency detuning
-($\omega_{q} \gets \omega_{q} + \delta \omega_{q}$)
+($f_{q} \gets f_{q} + \delta f_{q}$)
 as well as flux amplitude scaling error
@@
 and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $\omega_{q}$ and $\sigma_{\omega_{q}}$ using X method.
+We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
 
@@
 ($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($\omega_{q}$) and the flux amplitude ($a$).
+state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
 Here we present the dynamics for the second derivative of the state with respect to the
 qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $\omega_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
+of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
 brevity.
@@
   \begin{aligned}
-      \partial_{\omega_{q}} \partial_{t} \ket{\psi} &= \partial_{\omega_{q}} H \ket{\psi}\\
-      &= (\partial_{\omega_{q}} H) \ket{\psi} + H (\partial_{\omega_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{\omega_{q}} \ket{\psi})
+      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
+      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
+      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
   \end{aligned}
@@
   \begin{aligned}
-    \partial_{\omega_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{\omega_{q}} (\partial_{\omega_{q}} H \ket{\psi})\\
-    &= (\partial_{\omega_{q}}^{2} H) \ket{\psi} + 2 (\partial_{\omega_{q}} H)(\partial_{\omega_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{\omega_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{\omega_{q}} \ket{\psi}) + H (\partial_{\omega_{q}}^{2} \ket{\psi})
+    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
+    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
+    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
+    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
   \end{aligned}
 \end{equation}
-The augmented state vector carries $\partial_{\omega_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $\omega_{q}$ and $a$,
+The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
+which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
 the $l$\textsuperscript{th} state derivative is coupled to the

### 2020-09-25 paper/main.tex
updates to figs and nbs

--- 2020-09-07 paper/paper.tex
@@
   twocolumn,
+  superscriptaddress,
 ]{revtex4-2}
@@
 %% TITLE
-\title{Robust Quantum Optimal Control}
+\title{Robust Control of a Fluxonium Qubit}
 
@@
 \affiliation{
+  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
+}
+\affiliation{
   Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
+}
+\author{Brian Jackson}
+\author{Zac Manchester}
+\affiliation{
+  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
+}
+\author{David I. Schuster}
+\affiliation{
+  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
+}
+\affiliation{
+  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
+}
+\affiliation{
+  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
 }
@@
   Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant. In this work we employ robust control techniques
-  to achieve gate errors two orders of magnitude lower than dynamic decoupling for parameter
-  spreads on the order of $1\%$. 
+  but they require exquisite calibration to be performant. In this work we employ robust trajectory optimization techniques
+  to achieve high fidelity gates for parameter spreads on the order of $1\%$. 
   Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  Lindblad master equation dynamics, which exponentially increase the computational complexity
+  the Lindblad master equation, which increases the computational complexity
   of the optimization algorithm.
@@
 (The field) The field of quantum optimal control (QOC) is concerned
-with improving the efficiency and accuracy with which quantum systems
-are manipulated.
+with efficiently and accurately manipulating quantum systems.
 Early QOC techniques were proposed for nuclear magnetic resonance experiments
@@
 bases for constraint spaces--which is not always feasible--and
-may hinder convergence. Methods for mitigating decoherence using numerical
+may hinder convergence.
+Methods for mitigating decoherence using numerical
 techniques have employed master equation dynamics (citation needed),
@@
 optimal control problem as a constrained optimization problem, allowing
-us to enforce experimental constraints.
+us to enforce arbitrary experimental constraints.
 Additionally,
@@
 outline a method for making the optimization longitudinal
-relaxation aware. We achieve a 4x increase in longitudinal relaxation times
+relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
 over the baseline gate set.
@@
 and compare them to existing dynamic decoupling methods. We find that
-the methods we employ outperform dynamic decoupling, producing
-high fidelity gates at system parameter spreads of $5\%$.
+the methods we employ are able to produce high fidelity gates when
+subject to parameter spreads of $1\%$.
 
@@
 
-%% F1
-\begin{figure}[ht]
-  \begin{subfigure}{.55\linewidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\linewidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}
-
-  \begin{subfigure}{\linewidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Master equation simulation with $T_{1}$ dissipation
-    over the course of concatenated gate applications.
-  }
-\end{figure}
-
 %% S3
@@
 \end{align}
-where $f_{q}$ is the qubit frequency at the flux frustration point,
+where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
 $a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
@@
 realized by modulating the flux 
-threading the device ($\Phi_{\textrm{ext}}$).
-The flux amplitude is obtained from the external flux by
-$a = 4 \pi \braket{g \lvert \hat{\phi} \rvert e} \rvert_{0.5 \Phi_{0}} E_{L}
-\delta \Phi_{\textrm{ext}} / (h \Phi_{0})$
-where $\hat{\phi}$ is the phase operator, $E_{L}$ is the characteristic inductance energy, $\Phi_{0}$
-is a flux quantum, and
-$\delta \Phi_{\textrm{ext}} = \Phi_{\textrm{ext}} - 0.5 \Phi_{0}$ is the flux
-offset from the flux frustration point.
+threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
+gates for the fluxonium qubit subject to experimental constraints, decoherence, and
+system parameter deviations. We compare the gates we obtain with optimal
+control methods to the analytically constructed gates reported in
+\cite{zhang2020universal} for the same device.
+%% The flux amplitude is obtained from the external flux by
+%% $a = 4 \pi \braket{g \lvert \hat{\phi} \rvert e} \rvert_{0.5 \Phi_{0}} E_{L}
+%% \delta \Phi_{\textrm{ext}} / (h \Phi_{0})$
+%% where $\hat{\phi}$ is the phase operator, $E_{L}$ is the characteristic inductance energy, $\Phi_{0}$
+%% is a flux quantum, and
+%% $\delta \Phi_{\textrm{ext}} = \Phi_{\textrm{ext}} - 0.5 \Phi_{0}$ is the flux
+%% offset from the flux frustration point.
 
-In the following we compare the gates found with our methods
-to the $X/2, Y/2, Z/2$ gates reported in \cite{zhang2020universal}
-for the same device. Note that $Y/2$ and arbitrary $Z$ rotations
-are sufficient for universal computation.
-%% TODO: list device parameters
-
-(Constraints) We now outline the constraints that we require each gate
+(Constraints) We now present the constraints we require each gate
 to obey.
-We require $a(t = 0) = a(t = t_{N}) = 0$.
-This constraint ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions.
+To ensures gates may be concatenated arbitrarily without
+inducing AWG ringing due to high-frequency transitions,
+we require $a(t = 0) = a(t = t_{N}) = 0$.
 Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
@@
 %% TODO: See ref. 28 in Helin paper
-We require $-300 \textrm{MHz} \le a(t) \le 300 \textrm{MHz}$
+We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
 to ensure the two-level approximation \ref{eq:hamiltonian}
 remains valid. Additionally, we require that each gate achieves
-the desired state transition $\braket{\psi_{f} \lvert \psi_{N}} = 1$.
+the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
 In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude
-to mitigate AWG ringing.
+of the first and second derivatives of the flux amplitude to
+ensure its smoothness, and in doing so mitigate AWG ringing.
 
 The optimization is performed over the second derivative of the flux amplitude
-($\frac{d^{2}}{dt^{2}} a(t)$) which is contained in the
+$\frac{d^{2}}{dt^{2}} a$ which is contained in the
 augmented control vector. The first derivative
-($\frac{d}{dt} a(t)$), proportional ($a(t)$), and integral ($\int a(t)$)
+$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
 flux amplitude terms
@@
 the second derivative of the flux amplitude by
-integration in the dynamics function $f$.
+integration in the dynamics function.
 Both the zero net flux and target quantum state constraint
-are then handled by ensuring the augmented final state is
-reached $x_{N} - x_{f} = 0$.
-The equality and inequality constraints on $a(t)$ are handled
+are then handled by ensuring the target augmented state is
+reached $x_{N} = x_{f}$.
+The equality and inequality constraints on $a$ are handled
 with a bound constraint.
@@
 ($T_{1}$ and $T_{\phi}$ noise)
-In addition to optimizing the cost functional to achieve a gate
-that obeys experimental constraints and has a high simulated fidelity,
-we also want to make the gate robust to noise that affects the experimental
+In addition to constructing gates
+that obey experimental constraints and have high simulated fidelities,
+we also desire to make the gates robust to noise that affects the experimental
 gate fidelity. Decoherence of the quantum state due to external noise
@@
 protection at the flux frustration point
-$T_{2e}(\Phi_{\textrm{ext}} = 0.5 \Phi_{0}) \sim 300 \mu\textrm{s}$
-where the qubit frequency is first-order insensitive to changes in flux.
+$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
+where the qubit is first-order insensitive to changes in flux.
 It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point. Conversely, $T_{1}$ is at a minimum
-at the flux frustration point $T_{1}(0.5 \Phi_{0}) = 0.315$ms,
+frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
+at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
 and increases away from the flux frustration point
-$T_{1}(0.43 \Phi_{0}) = 4.3$ms. Given the nature
+$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
 of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time ($T_{1}$)
-over the gate duration
-and employ robust control techniques to mitigate
+to maximize the longitudinal relaxation time directly
+via the flux amplitude and employ robust control techniques to mitigate
 pure dephasing.
-%% TODO: See fig. 3 in Helin paper
+
+%% F1
+\begin{figure*}[ht]
+  \begin{subfigure}{.315\textwidth}
+    \includegraphics[width=\linewidth]{assets/f1a.png}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.23\textwidth}
+    \includegraphics[width=\linewidth]{assets/f1b.png}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.4\textwidth}
+    \includegraphics[width=\linewidth]{assets/f1c.png}
+  \end{subfigure}
+  \caption{
+    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
+    (b) $T_{1}$ interpolation function used in optimization. Markers
+    denote the time-averaged, absolute amplitude of each gate.
+    (c) Lindblad master equation simulation with $T_{1}$ dissipation
+    for successive gate applications. The reported gate error is computed after
+    each gate appliaction.
+  }
+\end{figure*}
 
@@
 \begin{equation}
-  P_{1}(t) = \int_{0}^{t} \gamma_{1}(a(t^{\prime})) dt^{\prime}
+  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
 \end{equation}
-where $\gamma_{1} = T_{1}^{-1}$. Setting the target
+Setting the target
 longitudinal relaxation probability to 0 results in
@@
 
-$\gamma_{1}(a_{k})$
+$T_{1}(a_{k})$
 is obtained at each knot point by evaluating
@@
 experimentally obtained data of the form
-$\{(\Phi_{\textrm{ext}}, T_{1})\}$.
-$\Phi_{\textrm{ext}}$ is a function of $a_{k}$
-with the inverse of the relation given in section 3.
+$\{(a, T_{1})\}$.
 Calculating $T_{1}$ directly from theoretical
-considerations requires many high-dimensional
+considerations requires many high-rank
 eigendecompositions, which
@@
 
-Furthermore, the probability of longitudinal 
-relaxation is dependent on the gate duration $t_{N}$.
+For a given $T_{1}$ time, a shorter gate duration would tend to
+favor a lower longitudinal relaxation probability.
 We allow the optimizer to tune the gate duration by
-adding the time step between knot points $\Delta t_{k}$
-to the augmented control vector $u_{k}$.
+making the time step between each knot point $\Delta t_{k}$
+a decision variable.
 Promoting $\Delta t_{k}$ to a decision variable, rather
@@
 constraint at each knot point
-$5\textrm{e-}3 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}2 \ \textrm{ns}$.
+$2.5\textrm{e-}3 \ \textrm{ns} \le
+\Delta t_{k} \le 1\textrm{e-}2 \ \textrm{ns}$.
 Note that this bound constraint is allowed to be
 broken for intermediate iterations of the optimization,
-so we use the absolute value of the time step
-$\lvert \Delta t_{k} \rvert$ to ensure that it is non-negative.
+so we add the square root of the time step $\sqrt{\Delta t_{k}}$
+to the augmented control vector and use the squared root
+of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
 
-(Results) We achieve a factor of 5 decrease in the probability
-of longitudinal relaxation from the analytic gates we benchmark
-against. We simulate the performance of the gates using the
-Lindblad master equation (see Appendix). We repeatedly apply the basis gates
-and measure the fidelity of the resulting state as a function
-of time as shown in Fig. 1.
+(Results) The $Y/2$ and $Z/2$ gates converge on similar solutions, a periodic
+waveform with amplitude $\sim 0.2 \textrm{GHz}$ (see Figure 1). Both extend their gate times
+beyond their analytic counterpart, trading longer gate times for access
+to higher amplitudes and therefore higher $T_{1}$ times. Both gates reduce
+their probability of longitudinal relaxation per gate by a factor of $\sim 5$ over
+their analytic counterparts. Furthermore, they perform simlarly in
+the concatenated gate application comparision, maintaining a total gate error under $10^{-2}$
+over $10^{3}$ gate applications, for an average gate error on the order of $10^{-5}$.
+We define gate error as the infidelity of the final state and the target state,
+averaged over multiple pseudo-randomly
+generated initial states. The $X/2$ gate
 
@@
     Y/2 & 5.253 & 1.157 & 4.540\\
-    X/2 & 16.251 & 6.946 & 2.340\\
+    X/2 & 16.251 & 4.488 & 3.621\\
   \end{tabular}
@@
 %% F2
-\begin{figure}[ht]
-  \begin{subfigure}{.55\linewidth}
+\begin{figure*}[ht]
+  \begin{subfigure}{.315\textwidth}
     \includegraphics[width=\linewidth]{assets/f2a.png}
   \end{subfigure}\hfill
-  \begin{subfigure}{.4\linewidth}
+  \begin{subfigure}{.4\textwidth}
+    \includegraphics[width=\linewidth]{assets/f2c.png}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.23\textwidth}
     \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \begin{subfigure}{\linewidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
   \end{subfigure}
@@
     (a) X/2 gates robust to qubit frequency detunings constructed using the
-    analytic, (2, 4)-point sampling, and (2\textsuperscript{nd},
-    3\textsuperscript{rd})-order derivative methods.
-    (b) Gate error as a function of the detuning from the nominal
-    qubit frequency at a fixed gate duration, see (a), and at the best
-    gate duration for each method chosen from a sweep, see (c).
-    (c) Gate error at a one-percent detuning from the nominal qubit frequency as
-    a function of the gate duration set in optimization.
+    analytic, 2-point sampling, 8-point unscented sampling, and (2\textsuperscript{nd},
+    3\textsuperscript{rd})-order derivative methods at a gate duration of $56.8$ns.
+    (b) Single gate error as
+    a function of the gate duration at a one-percent
+    detuning from the nominal qubit frequency for all methods.
+    (c) Single gate error as a function of the detuning from the nominal
+    qubit frequency at gate durations of $56.8$ns and $120$ns for the analytic
+    and 2\textsuperscript{nd}-order derivative methods.
   }
-\end{figure}
+\end{figure*}
 
 %% F3
-\begin{figure}[ht]
+\begin{figure}[h]
   \begin{subfigure}{\linewidth}
@@
 
-  %% \begin{subfigure}{.5\linewidth}
-  %%   \includegraphics[width=\linewidth]{assets/f3c.png}
-  %% \end{subfigure}
-
   \caption{
     (a) $X/2$ gates robust to flux offsets constructed using the analytic,
-    (2, 4)-point sampling, and (2, 3)-derivative methods at a gate duration
-    of $56.8 (ns)$. (b) Mean gate error for 10 samples of 1/$f$ flux noise over 5000
-    gate applications.
-    %% (c) Stochastic 1/$f$ flux noise.
+    (2, 4)-point sampling, and (2\textsuperscript{nd}, 3\textsuperscript{rd})-order
+    derivative methods at a gate duration
+    of $56.8$ns. (b) Simulation of stochastic 1/$f$ flux noise for
+    successive gate applications. The reported gate error is measured
+    after each gate application.
   }
@@
 \section{Robustness to Parameter Deviations}
-(Strategy) For the fluxonium, pure dephasing acts
-along the flux axis. Low frequency 1/$f$ noise,
+(Strategy) Magentic flux noise,
 Johnson-Nyquist current noise present in the flux bias
@@
 to the parameter deviation are cancelled.
-The latter two methods are inflexible. We draw on the
+%% The latter two methods are inflexible.
+We draw on the
 robust control literature to demonstrate
@@
 We propose two methods for engineering robustness
-to parameter deviations. The first we call the
+to parameter deviations. The first is the
 derivative method. We draw on the intuition that
-making the state evolution insensitive to changes
-in the parameter ($x$) is encoded in the derivative of
+the sensitivity of the state evolution to the parameter
+$\lambda$ is encoded in the derivative of
 the state with respect to that parameter
-$\partial_{x}^{l} \ket{\psi}$.
-The derviatives of the state are propgated in the
+$\partial_{\lambda}^{l} \ket{\psi}$. The derivative
+scheme aims to minimize the norm of this derivative.
+The derviatives of the state with respect to the parameter
+of interest are propgated in the
 augmented state vector. Their dynamics are found by differentiating
@@
 (see Appendix). Setting the target derivatives of the state
-to zero vectors results in quadratic costs at each
+to the zero vector results in quadratic costs at each
 knot point
-${\lvert \partial_{x}^{l} \ket{\psi_{k}} \rvert}^{2}$.
+${\lvert \partial_{\lambda}^{l} \ket{\psi} \rvert}^{2}$.
+In the $l$\textsuperscript{th}-order derivative method, the norm of
+all derivatives of order less than or equal to $l$ are penalized. For example,
+in the 2\textsuperscript{nd}-order derivative method both the
+1\textsuperscript{st}-order state derivative
+(${\lvert \partial_{\lambda} \ket{\psi_{k}} \rvert}^{2}$)
+and the 2\textsuperscript{nd}-order state derivative
+(${\lvert \partial^{2}_{\lambda} \ket{\psi_{k}} \rvert}^{2}$)
+contribute quadratic costs at each knot point.
 
@@
 We add additional
-states that follow deviant dynamics to the augmented
-state vector. The parameters of interest
-are altered by a small deviation, typically chosen
-to represent the distribution of the parameter.
+states to the augmented state vector that follow deviant dynamics, and
+penalize the state's difference from the target state.
+The deviant dynamics are given by the system
+Hamiltonian with the parameter of interest replaced by
+a deviant value $\lambda \gets \lambda + \delta \lambda$
+The parameter's deviation is typically chosen
+to represent the statistical distribution of the parameter.
 For example, in the 2-point sampling method we
-propagate $\ket{\psi^{\pm, I}(t)}$ where
-$f_{q}^{\pm} = f_{q} \pm \sigma_{f_{q}}$,
-$f_{q}^{I} = f_{q}$.
+propagate the additional states $\ket{\psi^{\pm}}$ where
+$\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$, with
+$\sigma_{\lambda}$ being the standard deviation of $\lambda$.
 The difference between the final states and the target
@@
 each knot point
-${\lvert \ket{\psi^{\pm, I}_{k}} - \ket{\psi_{f}} \rvert}^{2}$
+${\lvert \ket{\psi^{\pm}_{k}} - \ket{\psi_{f}} \rvert}^{2}$.
+For the 4-point sampling method we sample at two standard deviations
+above and below the nominal parameter value in addition to one standard
+deviation above and below.
 
-(Results) We compare the robust control
-schemes to a dynamic decoupling pulse
-for a single $R_{x}(\pi/2)$ gate
-subject to qubit frequency detuning
-($f_{q} \gets f_{q} + \delta f_{q}$)
-to demonstrate the applicability of our techniques to mitigate
-system parameter deviations.
-We use a compensating for off-resonance error
-(CORPSE) dynamic decoupling pulse that is optimized
-to mitigate first-order error arising due to the detuning.
-The pulse is composed of
-three rotations about the $\hat{x}$, $-\hat{x}$, and $\hat{x}$ axes
-of the Bloch sphere
-designed to compensate for erroneous rotations
-(see Appendix). We simulate the gate fidelity as
-a function of the qubit frequency detuning for
-variants of the
-derivative method, sample method, and
-CORPSE pulse (see Figure 2a). We also find that
-increasing the gate duration allows the optimizer
-to find a more robust pulse
-(see Figure 2b). Increasing the order of the derivative
-and sample methods requires propagating an additional quantum state
-in the augmented state vector, a linear increase
-in the computational complexity, but finding higher order
-variants of the CORPSE family is prohibitively difficult.
+(Comparison) To demonstrate the applicability of these techniques
+to mitigate system parameter deviations,
+we consider the task of achieving a single $X/2$
+gate subject to a constant qubit frequency detuning
+($f_{q} \gets f_{q} + \delta f_{q}$).
+%% We use a compensating for off-resonance error
+%% (CORPSE) dynamic decoupling pulse that is optimized
+%% to mitigate first-order error arising due to the detuning.
+%% The pulse is composed of
+%% three rotations about the $\hat{x}$, $-\hat{x}$, and $\hat{x}$ axes
+%% of the Bloch sphere
+%% designed to compensate for erroneous rotations
+%% (see Appendix).
+We take $\sigma_{f_{q}} = 1\% f_{q}$ to be one standard devation, and equip
+the sampling robustness methods accordingly.
+For each method we compute the gate error for one simulated gate application
+subject to the stated deviant dynamics.
 
-Additionally, we compare the robust control
-schemes to a dynamic decoupling pulse
-for sequential $R_{x}(\pi/2)$ gates
-subject to $T_{\phi}$ noise to demonstrate the applicability
-of our techniques to mitigate noise of this type.
-We simulate the gate fidelity as a function of time
-for variants of the derivative method,
-sample method, and B2CORPSE pulse (see Figure 3).
-The B2CORPSE pulse is constructed to be robust
-to qubit frequency detuning errors and
-amplitude scaling errors (see Appendix).
-The derivative and sampling methods are employed to
-mitigate error due to qubit frequency detuning
-($f_{q} \gets f_{q} + \delta f_{q}$)
-as well as flux amplitude scaling error
-($a \gets a (1 + \delta a)$).
+(Results) Both the 2-point and 4-point
+sampling methods converge on a similar solution with a low amplitude and
+a single amplitude parity switch (see Figure 2a). At the gate time of the analytic pulse
+($56.8$ ns) and a one percent qubit frequency detuning they achieve
+gate errors $\sim 1.1 \cdot 10^{-4}$ commensurate with the analytic
+pulse $\sim 1.2 \cdot 10^{-4}$. The 2-point and 4-point sampling methods
+achieve similar metrics in simulation (see Table 2).
+
+Both the 2\textsuperscript{nd}-order and 3\textsuperscript{rd}-order derivative methods
+employ multiple amplitude parity switches and idle periods at the maximum
+amplitude $0.5$ GHz. At the gate time of the analytic pulse and a one percent qubit
+frequency detuning, the 2\textsuperscript{nd}-order method achieves a gate error
+$\sim 0.8 \cdot 10^{-4}$ and the 3\textsuperscript{rd}-order method achieves a gate error
+$\sim 1.1 \cdot 10^{-4}$ (see Figure 2b). Inspection of the simulation metrics reveals
+that the 3\textsuperscript{rd}-order method is able to achieve a lower norm of the
+3\textsuperscript{rd}-order state derivative at the cost of higher 1\textsuperscript{st}-order
+and 2\textsuperscript{nd}-order derivatives than the 2\textsuperscript{nd}-order method (see Table 3).
+
+%% T2
+\begin{table}
+  \begin{tabular}{c | c | c | c}
+    Method & ${\lvert \partial_{f_{q}} \ket{\psi} \rvert}^{2}$ ($10^{3}$)
+    & ${\lvert \partial^{2}_{f_{q}} \ket{\psi} \rvert}^{2}$ ($10^{7}$)
+    & ${\lvert \partial^{3}_{f_{q}} \ket{\psi} \rvert}^{2}$ ($10^{12}$)\\
+    \hline
+    D-2 & 8.706 & 8.331 & 3.692\\
+    D-3 & 9.643 & 9.095 & 3.345\\
+  \end{tabular}
+  \caption{Norm of basis state derivatives undergoing a $X/2$ gate for the derivative methods at $t_{N} = 56.8$ns}
+\end{table}
+
+Further, we observe that the derivative methods are able
+to take advantage of an increased gate duration to achieve greater robustness,
+while the sampling methods are unable to. The tradeoff between gate error and
+gate duration appears roughly linear, where trading 3 times the gate duration
+results in a factor of 4 reduction in gate error for the 2\textsuperscript{nd}-order
+derivative method (using the 50ns and 150ns points, include this remark?). Additionally,
+no optimization converged for gate durations less than 50ns., and robustness commensurate
+with the analytic pulse at a shorter gate duration was not able to be achieved.
+
+%% TODO: maybe add new section
+(Comparison) An additional task of interest is to mitigate
+stochastic parameter deviations. For the fluxonium qubit, magnetic flux noise
+acts to modify the flux amplitude from its intended value by an amount $\delta a$.
+It is well studied that the spectral density of $\delta a$ follows a
+1/$f$ distribution, consisting primarily of low frequency noise. We compare
+the methods discussed for static parameter deviations to the task of
+realizing a $X/2$ gate subject to 1/$f$ flux noise. We simulate successive
+applications of the gate constructed by each method and measure the gate error
+after each application (see Figure 3b). The system is subject to flux noise
+$\delta a (t)$ that is generated from a standard normal distribution, normalized
+by 1/$f^{1/2}$ in frequency space, and modulated by the experimentally obtained
+flux noise amplitude $A_{\Phi} = 5.21 \mu \Phi_{0}$.
+
+(Results) ...
 
@@
 We have proposed some schemes and they work well.
+
+%% ACK
+\begin{acknowledgments}
+  The authors would like to thank Helin Zhang for experimental assistance,
+  and Stephen Lyon (maybe) and Daniel Weiss for useful discussions.
+\end{acknowledgments}
 

### 2020-09-25 src/spin/figures.jl
changes b4 fig2 gets zpiby2

@@ -219,8 +219,8 @@ end
 ### FIGURE 2 ###
 const F2_DATA = Dict(
     analytic => joinpath(SPIN_OUT_PATH, "spin14/00004_spin14.h5"),
-    s2 => joinpath(SPIN_OUT_PATH, "spin12/00496_spin12.h5"),
-    s4 => joinpath(SPIN_OUT_PATH, "spin12/00498_spin12.h5"),
+    s2 => joinpath(SPIN_OUT_PATH, "spin12/00627_spin12.h5"),
+    sut8 => joinpath(SPIN_OUT_PATH, "spin23/00094_spin23.h5"),
     d2 => joinpath(SPIN_OUT_PATH, "spin11/00432_spin11.h5"),
     d3 => joinpath(SPIN_OUT_PATH, "spin11/00429_spin11.h5"),
     s2b => joinpath(SPIN_OUT_PATH, "spin12/00508_spin12.h5"),
@@ -230,7 +230,7 @@ const F2_DATA = Dict(
 )
 
 
-const F2A_PT_LIST = [analytic, s2, s4, d2, d3]
+const F2A_PT_LIST = [analytic, s2, sut8, d2, d3]
 function gen_2a()
     pulse_types_integer = [Integer(pulse_type) for pulse_type in F2A_PT_LIST]
     pulse_type_count = size(F2A_PT_LIST)[1]
@@ -250,7 +250,7 @@ end
 
 const F2B_TRIAL_COUNT = Integer(1e2)
 const F2B_FQ_DEV = 3e-2
-const F2B_PT_LIST = [analytic, s2, s4, d2, d3, s2b, s4b, d2b, d3b]
+const F2B_PT_LIST = [analytic, s2, sut8, d2, d3, s2b, s4b, d2b, d3b]
 const F2B_AVG_COUNT = 10
 function gen_2b(;use_previous=true)
     @assert iseven(F2B_TRIAL_COUNT)
@@ -299,20 +299,20 @@ function gen_2b(;use_previous=true)
                 if (!isnothing(save_file_path_old) && save_file_path == save_file_path_old
                     && k <= avg_count_old)
                     gate_errors[i, j, k] = gate_errors_old[i, j, k]
-                    continue
-                end
-                if pulse_type == analytic
-                    res = run_sim_deqjl(
-                        1, gate_type; dynamics_type=xpiby2nodis,
-                        dt=1e-3, save=false, negi_h0=negi_h0, seed=k
-                    )
                 else
[…]

### 2020-09-25 src/spin/spin12.jl

@@ -22,12 +22,12 @@ const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 # problem
 const CONTROL_COUNT = 1
 const STATE_COUNT = 2
+const SAMPLE_COUNT = 2
 const ASTATE_SIZE_BASE = STATE_COUNT * HDIM_ISO + 3 * CONTROL_COUNT
+const ASTATE_SIZE = ASTATE_SIZE_BASE + SAMPLE_COUNT * HDIM_ISO
+const ACONTROL_SIZE = CONTROL_COUNT
 const INITIAL_STATE1 = [1., 0, 0, 0]
 const INITIAL_STATE2 = [0., 1, 0, 0]
-const SIGMA = 1e-2
-const S1FQ_NEGI_H0_ISO = (FQ + FQ * SIGMA) * NEGI_H0_ISO
-const S2FQ_NEGI_H0_ISO = (FQ - FQ * SIGMA) * NEGI_H0_ISO
 # state indices
 const STATE1_IDX = 1:HDIM_ISO
 const STATE2_IDX = STATE1_IDX[end] + 1:STATE1_IDX[end] + HDIM_ISO
@@ -40,145 +40,240 @@ const S2STATE1_IDX = S1STATE1_IDX[end] + 1:S1STATE1_IDX[end] + HDIM_ISO
 const D2CONTROLS_IDX = 1:CONTROL_COUNT
 
 # model
-struct Model{SC} <: AbstractModel
+struct Model <: AbstractModel
+    h0_samples::Vector{SMatrix{HDIM_ISO, HDIM_ISO}}
 end
-RD.state_dim(::Model{SC}) where SC = (
-    ASTATE_SIZE_BASE + SC * HDIM_ISO
+@inline RD.state_dim(::Model) = ASTATE_SIZE
+@inline RD.control_dim(::Model) = ACONTROL_SIZE
+
+
+# This cost puts a gate error cost on
+# the sample states and a LQR cost on the other terms.
+# The hessian w.r.t the state and controls is constant.
+struct Cost{N,M,T} <: TO.CostFunction
+    Q::Diagonal{T, SVector{N,T}}
+    R::Diagonal{T, SVector{M,T}}
+    q::SVector{N, T}
+    c::T
+    hess_astate::Symmetric{T, SMatrix{N,N,T}}
+    target_state1::SVector{HDIM_ISO, T}
+    q_ss::T
[…]

### 2020-09-30 paper/main.tex
before switch to z/2 in f2

@@
   for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimizaiton.
+  using a computationally efficient metric in optimization.
 \end{abstract}
@@
 %% S5
-\section{Robustness to Parameter Deviations}
-(Strategy) Magentic flux noise,
-Johnson-Nyquist current noise present in the flux bias
-lines, and temperature dependent gain fluctuations
-may cause the flux amplitude to deviate from its
-desired value by an amount $\delta a$. Additionally,
-residual calibration errors and fluctuations may cause
+\section{Robustness to Static Parameter Deviations}
+(Strategy) Uncertainty in the Hamiltonian parameters lead
+to erroneous 
+The uncertainty Hamiltonian parameters are known to
+some finite experimental accuracy. Further,
+experimental fluctuations may cause parameters to deviate
+from their
 the qubit frequency to deviate from its measured value
-by an amount $\delta f_{q}$.
+by an amount $\delta f_{q}$. There are other situations that
+arise where this is useful, such as static amplitude
+offsets or static gain offsets.
 
@@
 %% (see Appendix).
-We take $\sigma_{f_{q}} = 1\% f_{q}$ to be one standard devation, and equip
+We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
 the sampling robustness methods accordingly.
@@
 
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c | c}
-    Method & ${\lvert \partial_{f_{q}} \ket{\psi} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \partial^{2}_{f_{q}} \ket{\psi} \rvert}^{2}$ ($10^{7}$)
-    & ${\lvert \partial^{3}_{f_{q}} \ket{\psi} \rvert}^{2}$ ($10^{12}$)\\
-    \hline
-    D-2 & 8.706 & 8.331 & 3.692\\
-    D-3 & 9.643 & 9.095 & 3.345\\
-  \end{tabular}
-  \caption{Norm of basis state derivatives undergoing a $X/2$ gate for the derivative methods at $t_{N} = 56.8$ns}
-\end{table}
-
 Further, we observe that the derivative methods are able
@@
 
-%% TODO: maybe add new section
+%% T2
+\begin{table}
+  \begin{tabular}{c | c | c}
+    Method & ${\lvert \partial_{f_{q}} \ket{\psi_{N}} \rvert}^{2}$ ($10^{3}$)
+    & ${\lvert \partial^{2}_{f_{q}} \ket{\psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
+    \hline
+    D-1 & 2.509 & 13.292\\
+    D-2 & 3.711 & 1.452\\
+  \end{tabular}
+  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
+    for $X/2$ gates optimized using the derivative methods. The norms are computed
+    at the end of the gate's duration $t_{N} = 110$ns.}
+\end{table}
+
+\section{Robustness to Stochastic Parameter Deviations} %% S6
 (Comparison) An additional task of interest is to mitigate
@@
 
-%% S6
+%% S7
 \section{Conclusion}
@@
   The authors would like to thank Helin Zhang for experimental assistance,
-  and Stephen Lyon (maybe) and Daniel Weiss for useful discussions.
+  and Daniel Weiss for useful discussions.
 \end{acknowledgments}

### 2020-09-30 src/spin/figures.jl
progg

@@ -34,6 +34,7 @@ const F3D_DATA_FILE_PATH = joinpath(SAVE_PATH, "f3d.h5")
     d1b = 14
     sut8b = 15
     d1bb = 16
+    d1bbb = 17
 end
 
 const GT_LIST = [zpiby2, ypiby2, xpiby2]
@@ -220,41 +221,53 @@ end
 
 
 ### FIGURE 2 ###
-const F2_DATA = Dict(
-    analytic => joinpath(SPIN_OUT_PATH, "spin14/00004_spin14.h5"),
-    s2 => joinpath(SPIN_OUT_PATH, "spin12/00627_spin12.h5"),
-    sut8 => joinpath(SPIN_OUT_PATH, "spin23/00094_spin23.h5"),
-    d1 => joinpath(SPIN_OUT_PATH, "spin11/00504_spin11.h5"),
-    d2 => joinpath(SPIN_OUT_PATH, "spin11/00528_spin11.h5"),
-    d1b => joinpath(SPIN_OUT_PATH, "spin11/00514_spin11.h5"),
-    d1bb => joinpath(SPIN_OUT_PATH, "spin11/00524_spin11.h5"),
+const F2A_DATA_ZPIBY2 = Dict(
+    analytic => joinpath(SPIN_OUT_PATH, "spin14/00000_spin14.h5"),
+    s2 => joinpath(SPIN_OUT_PATH, "spin12/00731_spin12.h5"), #"spin12/00692_spin12.h5"),
+    sut8 => "", #joinpath(SPIN_OUT_PATH, "spin23/00094_spin23.h5"),
+    d1 => joinpath(SPIN_OUT_PATH, "spin11/00539_spin11.h5"), #"spin11/00468_spin11.h5"),
+    d2 => joinpath(SPIN_OUT_PATH, "spin11/00547_spin11.h5"), #"spin11/00487_spin11.h5"),
 )
-
 const F2A_PT_LIST = [analytic, s2, sut8, d1, d2]
-function gen_2a()
+function gen_2a(;gate_type=zpiby2)
+    if gate_type == zpiby2
+        data = F2A_DATA_ZPIBY2
+    end
     pulse_types_integer = [Integer(pulse_type) for pulse_type in F2A_PT_LIST]
     pulse_type_count = size(F2A_PT_LIST)[1]
     save_file_paths = Array{String, 1}(undef, pulse_type_count)
[…]

### 2020-09-30 src/spin/spin.jl

@@ -91,6 +91,28 @@ const DT_ST = Dict(
     xpicorpse => st_state,
 )
 
+const DT_AN = Dict(
+    schroed => false,
+    schroedda => false,
+    lindbladnodis => false,
+    lindbladt1 => false,
+    lindbladt2 => false,
+    ypiby2nodis => true,
+    ypiby2t1 => true,
+    xpiby2nodis => true,
+    xpiby2t1 => true,
+    zpiby2nodis => true,
+    zpiby2t1 => true,
+    xpiby2da => true,
+    xpiby2t2 => true,
+    zpiby2da => true,
+    empty => true,
+    ypiby2da => true,
+    xpiby2corpserwa => true,
+    xpiby2corpse => true,
+    xpicorpse => true,
+)
+
 const GT_STR = Dict(
     zpiby2 => "Z/2",
     ypiby2 => "Y/2",
@@ -424,6 +446,17 @@ const GAMMA_ZPIBY2 = amp_t1_spline(0)^(-1)
 )
 
 
+function integrate_prop_zpiby2nodis!(gate_count::Int, states::Array{T, 2},
+                                     state::SVector, params::SimParams) where {T}
+    states[1, :] = state
+    for i = 2:gate_count + 1
+        prop = exp(TTOT_ZPIBY2 * params.negi_h0)
+        state = prop * state
+        states[i, :] = state
+    end
+end
+
+
 @inline dynamics_zpiby2t1_deqjl(density::StaticMatrix, params::SimParams, time::Float64) = (
     FQ_NEGI_H0_ISO * density - density * FQ_NEGI_H0_ISO
     + GAMMA_ZPIBY2 * (G_E * density * E_G + NEG_E_E_BY2 * density + density * NEG_E_E_BY2
@@ -512,6 +545,15 @@ function dynamics_xpiby2nodis_deqjl(state::StaticVector, params::SimParams, time
     return(
         negi_h * state
     )
+end
+
+
+function integrate_prop_xpiby2nodis!(gate_count::Int, states::Array{T, 2},
+                                     state::SVector, params::SimParams) where {T}
+    states[1, :] = state
+    for i = 2:gate_count + 1
+        continue
+    end
 end
 
 
@@ -766,6 +808,8 @@ const DT_DYN = Dict(
 
 # integartor lookup
 const DT_INT = Dict(
[…]

### 2020-09-30 src/spin/spin12.jl

@@ -151,7 +151,7 @@ end
 
 
 # main
-function run_traj(;gate_type=zpiby2, evolution_time=17.5, solver_type=altro,
+function run_traj(;gate_type=zpiby2, evolution_time=18., solver_type=altro,
                   sqrtbp=false, integrator_type=rk3, qs=[1e0, 1e0, 1e0, 1e-1, 1e0, 1e-1],
                   dt_inv=Int64(1e1), smoke_test=false, constraint_tol=1e-8, al_tol=1e-4,
                   pn_steps=2, max_penalty=1e11, verbose=true, save=true, max_iterations=Int64(2e5),

### 2020-09-30 src/spin/spin18.jl

@@ -1,6 +1,9 @@
 """
-spin18.jl - sampling robustness for control amplitude
+spin18.jl - sampling robustness
 """
+
+WDIR = joinpath(@__DIR__, "../../")
+include(joinpath(WDIR, "src", "spin", "spin.jl"))
 
 using Altro
 using HDF5
@@ -8,206 +11,219 @@ using LinearAlgebra
 using RobotDynamics
 using StaticArrays
 using TrajectoryOptimization
+const RD = RobotDynamics
 const TO = TrajectoryOptimization
-
-WDIR = get(ENV, "ROBUST_QOC_PATH", "../../")
-include(joinpath(WDIR, "src", "spin", "spin.jl"))
 
 # paths
 const EXPERIMENT_META = "spin"
 const EXPERIMENT_NAME = "spin18"
 const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 
-# Define the optimization.
-const DT_STATIC = DT_PREF
-const DT_STATIC_INV = DT_PREF_INV
-const CONSTRAINT_TOLERANCE = 1e-8
-const AL_KICKOUT_TOLERANCE = 1e-7
-const PN_STEPS = 5
-const MAX_PENALTY = 1e11
-const ILQR_DJ_TOL = 1e-4
-
-# Define the problem.
+# problem
 const CONTROL_COUNT = 1
 const STATE_COUNT = 2
-const ASTATE_SIZE_BASE = STATE_COUNT * STATE_SIZE_ISO + 3 * CONTROL_COUNT
+const SAMPLE_COUNT = 2
+const ASTATE_SIZE_BASE = STATE_COUNT * HDIM_ISO + 3 * CONTROL_COUNT
+const ASTATE_SIZE = ASTATE_SIZE_BASE + SAMPLE_COUNT * HDIM_ISO
+const ACONTROL_SIZE = CONTROL_COUNT
 const INITIAL_STATE1 = [1., 0, 0, 0]
 const INITIAL_STATE2 = [0., 1, 0, 0]
 # state indices
-const STATE1_IDX = 1:STATE_SIZE_ISO
-const STATE2_IDX = STATE1_IDX[end] + 1:STATE1_IDX[end] + STATE_SIZE_ISO
+const STATE1_IDX = 1:HDIM_ISO
[…]

### 2020-09-30 src/spin/spin25.jl

@@ -1,5 +1,5 @@
 """
-spin23.jl - unscented transform robustness for the δfq problem
+spin25.jl - unscented sampling robustness for the δa problem
 """
 
 WDIR = joinpath(@__DIR__, "../../")
@@ -20,7 +20,7 @@ const TO = TrajectoryOptimization
 
 # paths
 const EXPERIMENT_META = "spin"
-const EXPERIMENT_NAME = "spin23"
+const EXPERIMENT_NAME = "spin25"
 const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 
 # problem
@@ -51,24 +51,42 @@ const ACONTROL_SIZE = CONTROL_COUNT
 const D2CONTROLS_IDX = 1:CONTROL_COUNT
 
 # model
-module Data
-using RobotDynamics
-using StaticArrays
-const RD = RobotDynamics
-const HDIM_ISO = 4
 mutable struct Model <: RD.AbstractModel
-    cov::MMatrix{HDIM_ISO + 1, HDIM_ISO + 1}
-    fq_cov::Float64
+    s1_samples::Array{SVector{HDIM_ISO}, 1}
+    fq_samples::MVector{SAMPLE_COUNT}
+    fq_dist::Distributions.Sampleable
     alpha::Float64
 end
-end
-Model = Data.Model
 @inline RD.state_dim(model::Model) = ASTATE_SIZE
 @inline RD.control_dim(model::Model) = ACONTROL_SIZE
 
 
 function unscented_transform!(model::Model, z::AbstractKnotPoint{T,N,M}) where {T,N,M}
-
+    s11 = SVector{HDIM_ISO}(z.z[S1STATE1_IDX])
+    s21 = SVector{HDIM_ISO}(z.z[S2STATE1_IDX])
+    s31 = SVector{HDIM_ISO}(z.z[S3STATE1_IDX])
+    s41 = SVector{HDIM_ISO}(z.z[S4STATE1_IDX])
+    s51 = SVector{HDIM_ISO}(z.z[S5STATE1_IDX])
+    s61 = SVector{HDIM_ISO}(z.z[S6STATE1_IDX])
+    s71 = SVector{HDIM_ISO}(z.z[S7STATE1_IDX])
+    s81 = SVector{HDIM_ISO}(z.z[S8STATE1_IDX])
+    
+    s1m = SAMPLE_COUNT_INV .* (
[…]

### 2020-09-30 paper/main.tex

@@
   \caption{
-    (a) X/2 gates robust to qubit frequency detunings constructed using the
-    analytic, 2-point sampling, 8-point unscented sampling, and (2\textsuperscript{nd},
-    3\textsuperscript{rd})-order derivative methods at a gate duration of $56.8$ns.
+    (a) Z/2 gates constructed using the
+    analytic, 2-point sampling, 8-point unscented sampling, and the 1\textsuperscript{st}
+    and 2\textsuperscript{nd} order derivative methods. The gates
+    shown for the robust methods are the solutions at twice the analytic
+    gate time.
     (b) Single gate error as
@@
     (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency at gate durations of $56.8$ns and $120$ns for the analytic
-    and 2\textsuperscript{nd}-order derivative methods.
+    qubit frequency at multiples of the analytic gate duration for the analytic
+    and 1\textsuperscript{st} order derivative methods.
   }

### 2020-10-12 paper/main.tex
updates

@@
   The ability to engineer high fidelity operations on quantum processors in the presence of
-  calibration errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
+  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
   Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant. In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$. 
+  but they require exquisite calibration to be performant.
+  In this work we employ robust trajectory optimization techniques
+  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
+  takes advantage of deviant parameter derivative information while maintaining
+  computational efficiency by transforming
+  higher-order differential equations into a set of coupled first-order equations.
   Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
@@
 \begin{align}
-  \frac{d}{dt} \ket{\psi(t)} &= -\frac{i}{\hbar} H(u(t), t) \ket{\psi(t)}
+  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
 \end{align}
@@
 and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state. We refer to $x$ and $u$ as
-the augmented state and augmented controls, respectively.
+we refer to $\ket{\psi}$ as the state, $x$ as the augmented
+state, and $u$ as the augmented controls.
 
@@
 Setting the target
-longitudinal relaxation probability to 0 results in
+probability to zero results in
 a quadratic cost at each knot point
-of the form ${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-
-$T_{1}(a_{k})$
-is obtained at each knot point by evaluating
-a spline interpolant fit to
-experimentally obtained data of the form
+${\lvert P_{1}(t_{k}) \rvert}^{2}$.
+$T_{1}(a_{k})$ is obtained at each knot point by evaluating
+a spline fit to
+experimental data of the form
 $\{(a, T_{1})\}$.
@@
 
-For a given $T_{1}$ time, a shorter gate duration would tend to
+For a constant $T_{1}$ time, a shorter gate duration would
 favor a lower longitudinal relaxation probability.
@@
 constraint at each knot point
-$2.5\textrm{e-}3 \ \textrm{ns} \le
-\Delta t_{k} \le 1\textrm{e-}2 \ \textrm{ns}$.
-Note that this bound constraint is allowed to be
-broken for intermediate iterations of the optimization,
+$5\textrm{e-}2 \ \textrm{ns} \le
+\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
+This bound constraint may be
+violated for intermediate iterations of the optimization,
 so we add the square root of the time step $\sqrt{\Delta t_{k}}$
@@
 
-(Results) The $Y/2$ and $Z/2$ gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$ (see Figure 1). Both extend their gate times
+The optimizations for the $Y/2$ and $Z/2$ gates converge on similar solutions, a periodic
+waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
+Both solutions extend their gate times
 beyond their analytic counterpart, trading longer gate times for access
@@
 their probability of longitudinal relaxation per gate by a factor of $\sim 5$ over
-their analytic counterparts. Furthermore, they perform simlarly in
+their analytic counterparts, a similar scaling for their single gate error improvements
+$\sim 3 \cdot 10^{-5} \rightarrow \sim 7 \cdot 10^{-6}$.
+We define gate error as the infidelity of the final state and the target state
+averaged over multiple pseudo-randomly generated initial states.
+Furthermore, they perform simlarly in
 the concatenated gate application comparision, maintaining a total gate error under $10^{-2}$
-over $10^{3}$ gate applications, for an average gate error on the order of $10^{-5}$.
-We define gate error as the infidelity of the final state and the target state,
-averaged over multiple pseudo-randomly
-generated initial states. The $X/2$ gate
-
-%% T1
-%% TODO: maybe for appendix.
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c}
-    & Analytic  & QOC &\\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745 & 1.149 & 5.000\\
-    Y/2 & 5.253 & 1.157 & 4.540\\
-    X/2 & 16.251 & 4.488 & 3.621\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
+over $10^{3}$ gate applications. 
+The $T_{1}$ optimized $X/2$ gate achieves a single gate error...
 
@@
     (a) Z/2 gates constructed using the
-    analytic, 2-point sampling, 8-point unscented sampling, and the 1\textsuperscript{st}
+    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}
     and 2\textsuperscript{nd} order derivative methods. The gates
@@
 \section{Robustness to Static Parameter Deviations}
-(Strategy) Uncertainty in the Hamiltonian parameters lead
-to erroneous 
-The uncertainty Hamiltonian parameters are known to
-some finite experimental accuracy. Further,
-experimental fluctuations may cause parameters to deviate
-from their
-the qubit frequency to deviate from its measured value
-by an amount $\delta f_{q}$. There are other situations that
-arise where this is useful, such as static amplitude
-offsets or static gain offsets.
+Gate errors may be incurred if Hamiltonian parameters deviate
+from their nominal values in the experiment. For example, the qubit frequency
+may be measurable up to a known precision $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$
+and drift in laboratory temperatures may increase
+amplitude gain in the electronics chain $a \gets a (1 + \epsilon)$.
+Multiple analytic methods have been presented to construct gates
+that mitigate the error arising from parameter deviations
+including composite pulse sequences \cite{merrill2014progress},
+the DRAG scheme \cite{krantz2019quantum}, and
+geometric phase considerations
+\cite{xu2020nonadiabatic, han2020experimental}.
+Composite pulse sequences are derived by computing
+the erroroneous rotation arising from the static parameter deviation
+with a finite order magnus expansion.
+Pulses are then suitably composed to eliminate the
+error. In principle the error may be eliminated
+to arbitrary order with sufficiently many pulses.
+It is difficult to choose an appropriate composite pulse
+for comparison on the problem we study here. We
+propose comparing numerical methods to existing analytic
+techniques for future work.
 
-System parameter
-deviations are typically addressed with dynamic decoupling
-sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, or
-geometric phase considerations
-\cite{xu2020nonadiabatic} \cite{han2020experimental}.
-Dynamic decoupling sequences compose rotations on the
-Bloch sphere so that erroneous rotations arising due
-to the parameter deviation are cancelled.
-%% The latter two methods are inflexible.
-We draw on the
-robust control literature to demonstrate
-numerical techniques for engineering robustness
-to parameter deviations.
+We consider two numerical strategies for achieving robustness
+to parameter deviations. The first we call the
+derivative method. We draw on the intuition that
+the sensitivity of the state evolution to a parameter
+$\lambda$ is encoded in the $l$\textsuperscript{th}-order
+derivative of the state with respect to that parameter
+$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
+derivative method minimizes the norm of the first $m$
+state derivatives with a quadratic cost at each knot point
+${\lvert \braket{\partial^{l}_{\lambda} \psi_{k} | \partial^{l}_{\lambda}
+    \psi_{k}} \rvert}^{2}$, $l \in \{1, \dots, m\}$.
 
-We propose two methods for engineering robustness
-to parameter deviations. The first is the
-derivative method. We draw on the intuition that
-the sensitivity of the state evolution to the parameter
-$\lambda$ is encoded in the derivative of
-the state with respect to that parameter
-$\partial_{\lambda}^{l} \ket{\psi}$. The derivative
-scheme aims to minimize the norm of this derivative.
-The derviatives of the state with respect to the parameter
-of interest are propgated in the
-augmented state vector. Their dynamics are found by differentiating
-the TDSE dynamics with respect to the parameter of interest
-(see Appendix). Setting the target derivatives of the state
-to the zero vector results in quadratic costs at each
-knot point
-${\lvert \partial_{\lambda}^{l} \ket{\psi} \rvert}^{2}$.
-In the $l$\textsuperscript{th}-order derivative method, the norm of
-all derivatives of order less than or equal to $l$ are penalized. For example,
-in the 2\textsuperscript{nd}-order derivative method both the
-1\textsuperscript{st}-order state derivative
-(${\lvert \partial_{\lambda} \ket{\psi_{k}} \rvert}^{2}$)
-and the 2\textsuperscript{nd}-order state derivative
-(${\lvert \partial^{2}_{\lambda} \ket{\psi_{k}} \rvert}^{2}$)
-contribute quadratic costs at each knot point.
+The state derivatives could be obtained with backwards mode differentiation.
+Naive automatic differentiation would compute
+the state derivative at all $1, \dots, k - 1$ knot points
+to obtain the state derivative at knot point $k$.
+For $d$ sensitivity parameters and $N$ knot points
+this requires $O(d N^2)$ matrix multiplications. Instead,
+we forward propagate the state derivatives in the
+augmented state vector under coupled equations of motion, resulting
+in $O(d N)$ matrix multiplications.
+For example, the equations of motion for the 1\textsuperscript{st}-order derivative
+method are
+\begin{align}
+  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
+  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
+  H \ket{\partial_{\lambda} \psi} +
+  (\partial_{\lambda} H) \ket{\psi}
+\end{align}
+Matrix exponential schemes that account for the non-linear
+term may be used to efficiently integrate the coupled dynamics
+\cite{berland2005solving, einkemmer2017performance}.
 
 The second scheme we analyze is the sampling method,
-which is well studied in the robust control literature
-\cite{manchester2016derivative} \cite{tronarp2016sigma}.
-We add additional
-states to the augmented state vector that follow deviant dynamics, and
-penalize the state's difference from the target state.
-The deviant dynamics are given by the system
-Hamiltonian with the parameter of interest replaced by
-a deviant value $\lambda \gets \lambda + \delta \lambda$
-The parameter's deviation is typically chosen
-to represent the statistical distribution of the parameter.
-For example, in the 2-point sampling method we
-propagate the additional states $\ket{\psi^{\pm}}$ where
-$\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$, with
-$\sigma_{\lambda}$ being the standard deviation of $\lambda$.
-The difference between the final states and the target
-state is penalized, resulting in quadratic costs at
-each knot point
-${\lvert \ket{\psi^{\pm}_{k}} - \ket{\psi_{f}} \rvert}^{2}$.
-For the 4-point sampling method we sample at two standard deviations
-above and below the nominal parameter value in addition to one standard
-deviation above and below.
+which has been proposed previously for
+QOC \cite{reinhold2019controlling}.
+We propagate additional states in the augmented state vector that follow deviant dynamics
+and penalize their differences from the target state.
+The states evolve under the deviant
+Hamiltonian where the parameter of interest is replaced by
+a deviant value $\lambda \gets \lambda + \delta \lambda$.
+We propagate the additional states $\ket{\psi^{\pm}}$ with
+deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
+The differences between the final states and the target
+state are penalized, resulting in quadratic costs at
+each knot point ${\lvert \braket{\tilde{\psi}^{\pm}_{k} | \tilde{\psi}^{\pm}_{k}}
+  \rvert}^{2}, \ket{\tilde{\psi}^{\pm}_{k}} = \ket{\psi^{\pm}_{k}} - \ket{\psi_{f}}$.
+We sample two copies ($\ket{\psi^{\pm}}$) of four initial states with this method,
+resulting in eight additional states added to the augmented state vector.
+The four initial states are chosen so
+that their outer products span the operators on the
+Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
+  (\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}. We observe
+that sampling only the basis states $\{\ket{0}$, $\ket{1}\}$
+is insufficient because, as in the case of a $Z$ rotation, their
+trajectories are insensitive to some parameter deviations and therefore
+they are undesirable samples for those operations.
 
-(Comparison) To demonstrate the applicability of these techniques
+Additionaly, we consider the unscented sampling method,
+which is well-studied in the robust control iterature
+\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
+The strategy is to
+propagate an ensemble of states (sigma points) which represent a distribution
+over every element of $\ket{\psi}$. The distribution
+models the uncertainty in each state element arising from the
+parameter deviation. The ensemble consists of $2 (n + d)$ states
+where $n$ is the size of the state vector and $d$ is the number of
+deviant parameters. Each state in the ensemble is propagated to the next
+knot point using separate deviant dynamics. Then, the mean and covariance of
+the propagated ensemble is calculated. New states are sampled from the distribution
+given by the calculated statistics for propagation to the next knot point.
+This resampling procedure, the unscented transformation, accurately propagates
+first and second moments of the distribution and ensures the points
+lie on the $\sqrt{n}$\textsuperscript{th}
+covariance contour at each knot point. A detailed update procedure is given
+in the appendix. The initial state of all $2 (n + d)$ samples is
+$(\ket{0} + i\ket{1} / \sqrt{2})$ with added noise. Repeating this procedure for the other
+three operator basis states would provide the optimizer with more information,
+but come at a computational cost which far exceeds that of the other methods,
+resulting in $O(n^2)$ additional states in the augmented state vector.
+
+To demonstrate the applicability of these techniques
 to mitigate system parameter deviations,
-we consider the task of achieving a single $X/2$
+we consider the task of achieving a single $Z/2$
 gate subject to a constant qubit frequency detuning
 ($f_{q} \gets f_{q} + \delta f_{q}$).
-%% We use a compensating for off-resonance error
-%% (CORPSE) dynamic decoupling pulse that is optimized
-%% to mitigate first-order error arising due to the detuning.
-%% The pulse is composed of
-%% three rotations about the $\hat{x}$, $-\hat{x}$, and $\hat{x}$ axes
-%% of the Bloch sphere
-%% designed to compensate for erroneous rotations
-%% (see Appendix).
 We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling robustness methods accordingly.
-For each method we compute the gate error for one simulated gate application
-subject to the stated deviant dynamics.
+the sampling methods accordingly. For each method we compute the gate error for
+one simulated gate application subject to the deviant dynamics given by the
+stated qubit frequency detuning.
 
-(Results) Both the 2-point and 4-point
-sampling methods converge on a similar solution with a low amplitude and
-a single amplitude parity switch (see Figure 2a). At the gate time of the analytic pulse
-($56.8$ ns) and a one percent qubit frequency detuning they achieve
-gate errors $\sim 1.1 \cdot 10^{-4}$ commensurate with the analytic
-pulse $\sim 1.2 \cdot 10^{-4}$. The 2-point and 4-point sampling methods
-achieve similar metrics in simulation (see Table 2).
+We compare the robust methods to an analytically derived $Z/2$ pulse which corresponds to
+idling at the flux frustration point $a = 0$. The analytic pulse
+is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
+is simple to derive. It is linearly sensitive to qubit frequency detunings,
+the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
+qubit frequency detuning, it achieves a gate error of $\sim 4.5 \cdot 10^{-5}$, sufficient
+for quantum error correction.
 
-Both the 2\textsuperscript{nd}-order and 3\textsuperscript{rd}-order derivative methods
-employ multiple amplitude parity switches and idle periods at the maximum
-amplitude $0.5$ GHz. At the gate time of the analytic pulse and a one percent qubit
-frequency detuning, the 2\textsuperscript{nd}-order method achieves a gate error
-$\sim 0.8 \cdot 10^{-4}$ and the 3\textsuperscript{rd}-order method achieves a gate error
-$\sim 1.1 \cdot 10^{-4}$ (see Figure 2b). Inspection of the simulation metrics reveals
-that the 3\textsuperscript{rd}-order method is able to achieve a lower norm of the
-3\textsuperscript{rd}-order state derivative at the cost of higher 1\textsuperscript{st}-order
-and 2\textsuperscript{nd}-order derivatives than the 2\textsuperscript{nd}-order method (see Table 3).
+Each numerical method is able to find solutions at
+all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
+The ability to perform $Z$ rotations in arbitrary times
+is critical for operating multi-qubit experiments in the lab frame.
+These numerical methods give an effective mechanism for synchronizing
+qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
+The sampling and unscented sampling methods
+converge on qualitatively similar solutions which combine idling periods
+with fast ramps to the maximum amplitude. The gate error at a one-percent
+detuning from the nominal qubit frequency achieved
+by the sampling method does not improve substantially over the
+range of gate durations. The unscented sampling method
+achieves linear decreases in its gate error with longer gate durations
+until half the larmor period $\sim 1 / 2 f_{q}$ after which it achieves a consistent
+gate error $\sim 3.5 \cdot 10^{-5}$.
 
-Further, we observe that the derivative methods are able
-to take advantage of an increased gate duration to achieve greater robustness,
-while the sampling methods are unable to. The tradeoff between gate error and
-gate duration appears roughly linear, where trading 3 times the gate duration
-results in a factor of 4 reduction in gate error for the 2\textsuperscript{nd}-order
-derivative method (using the 50ns and 150ns points, include this remark?). Additionally,
-no optimization converged for gate durations less than 50ns., and robustness commensurate
-with the analytic pulse at a shorter gate duration was not able to be achieved.
+The derivative methods also converge on qualitatively similar solutions that
+use fast traingle pulses at the boundaries and balance time
+on either side of the flux-frustration point symmetrically at low amplitudes.
+They achieve super-linear scaling in their gate error as
+a function of the gate duration which approaches zero at four multiples
+of $t_{Z/2}$. We observe that on a range of problems at small parameter detunings,
+the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
+derivative method. We believe this is due to the low contribution of second-order
+terms to the gate error in this deviation regime, see Table 2.
 
 %% T2
+%% TODO: Appendix
 \begin{table}
   \begin{tabular}{c | c | c}
-    Method & ${\lvert \partial_{f_{q}} \ket{\psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \partial^{2}_{f_{q}} \ket{\psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
+    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
+    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
     \hline
-    D-1 & 2.509 & 13.292\\
-    D-2 & 3.711 & 1.452\\
+    D-1 & 2.051 & 17.935\\
+    D-2 & 2.511 & 7.628\\
   \end{tabular}
   \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $X/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 110$ns.}
+    for $Z/2$ gates optimized using the derivative methods. The norms are computed
+    at the end of the gate's duration $t_{N} = 38$ns.}
 \end{table}
 
-\section{Robustness to Stochastic Parameter Deviations} %% S6
-(Comparison) An additional task of interest is to mitigate
-stochastic parameter deviations. For the fluxonium qubit, magnetic flux noise
-acts to modify the flux amplitude from its intended value by an amount $\delta a$.
+%% S6
+\section{Robustness to Stochastic Parameter Deviations}
+An additional source of error can be modeled by stochastic
+Hamiltonian parameter deviations. For the fluxonium qubit, magnetic flux noise
+modifies the flux amplitude from its nominal value by an amount $\delta a$.
+
 It is well studied that the spectral density of $\delta a$ follows a
@@
 
+%% T1
+\begin{table}[ht]
+  \begin{tabular}{c | c | c | c | c | c | c}
+         & Analytic & QOC & & Analytic & QOC & \\
+    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
+    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
+    \hline
+    Z/2 & 5.745  & 1.149 & 5.000 & 3.568  & 0.727 & 4.905\\
+    Y/2 & 5.253  & 1.157 & 4.540 & 3.107  & 0.763 & 4.072\\
+    X/2 & 16.251 & 4.488 & 3.621 & 10.542 & 3.563 & 3.449\\
+  \end{tabular}
+  \caption{Probability of longitudinal relaxation for each gate
+    evaluated at the gate's duration.}
+\end{table}
+
 %% AD
@@
 %% AE
-\section{Dynamic Decoupling Sequences}
-Here we present the equations for constructing the CORPSE and
-B2CORPSE sequences.
-
-%% AF
 \section{Complex Tensor Handling}

### 2020-10-12 paper/main.tex
main.tex

@@
 
-($T_{1}$ and $T_{\phi}$ noise)
 In addition to constructing gates
@@
 There is a tradeoff between the two decoherence processes. In the case of white
-noise we have that the sum of the noise weights $W_{1}$ and $W_{\phi}$
+noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
 is constant \cite{huang2020engineering}.
@@
     for successive gate applications. The reported gate error is computed after
-    each gate appliaction.
+    each gate appliaction. The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
+    in the figure.
   }
@@
 Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation.
+the fridge truth of the simulation. Furthermore, this scheme for mitigating
+longitudinal relaxation requires one
+additional real number be added to the augmented state vector, avoiding
+the $O(n^{2})$ increase in computational and memory complexities
+associated with propagating a density matrix under the Lindblad
+master equation. Here $n$ is the size of the state vector.
 
@@
 to higher amplitudes and therefore higher $T_{1}$ times. Both gates reduce
-their probability of longitudinal relaxation per gate by a factor of $\sim 5$ over
-their analytic counterparts, a similar scaling for their single gate error improvements
-$\sim 3 \cdot 10^{-5} \rightarrow \sim 7 \cdot 10^{-6}$.
-We define gate error as the infidelity of the final state and the target state
+their single gate probability of longitudinal relaxation by a factor of $5$ over
+their analytic counterparts, scaling similarly to their single gate error reduction
+$3 \cdot 10^{-5} \rightarrow 7 \cdot 10^{-6}$, see Appendix.
+The gate error reported in this text is the infidelity of the evolved state and the target state
 averaged over multiple pseudo-randomly generated initial states.
 Furthermore, they perform simlarly in
-the concatenated gate application comparision, maintaining a total gate error under $10^{-2}$
-over $10^{3}$ gate applications. 
+the concatenated gate application comparision, maintaining total gate errors under $10^{-2}$
+over $10^{3}$ gate applications. Their single gate errors are sufficient for
+quantum error correction, requisite for fault-tolerant quantum computing,
+and their low gate errors for extend computations are critical for noisy, intermediate-scale
+quantum (NISQ) applications.
 The $T_{1}$ optimized $X/2$ gate achieves a single gate error...
@@
 
-We consider two numerical strategies for achieving robustness
+We consider three numerical strategies for achieving robustness
 to parameter deviations. The first we call the
@@
 state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi_{k} | \partial^{l}_{\lambda}
-    \psi_{k}} \rvert}^{2}$, $l \in \{1, \dots, m\}$.
+${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
+  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
 
@@
 \end{align}
-Matrix exponential schemes that account for the non-linear
+Exponential integrators that account for the non-linear
 term may be used to efficiently integrate the coupled dynamics
@@
 deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The differences between the final states and the target
-state are penalized, resulting in quadratic costs at
-each knot point ${\lvert \braket{\tilde{\psi}^{\pm}_{k} | \tilde{\psi}^{\pm}_{k}}
-  \rvert}^{2}, \ket{\tilde{\psi}^{\pm}_{k}} = \ket{\psi^{\pm}_{k}} - \ket{\psi_{f}}$.
-We sample two copies ($\ket{\psi^{\pm}}$) of four initial states with this method,
+The infidelities of the samples and the target
+state are penalized at each knot point
+$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
+We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
 resulting in eight additional states added to the augmented state vector.
@@
 is insufficient because, as in the case of a $Z$ rotation, their
-trajectories are insensitive to some parameter deviations and therefore
-they are undesirable samples for those operations.
+trajectories are insensitive to some parameter deviations, making
+them undesirable samples for those operations.
 
@@
 but come at a computational cost which far exceeds that of the other methods,
-resulting in $O(n^2)$ additional states in the augmented state vector.
+resulting in $O(n^2)$ additional states in the augmented state vector. The
+costs at each knot point are the infidelities
+between the samples and the target state, as in the
+basic sampling method.
 
@@
 the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
-qubit frequency detuning, it achieves a gate error of $\sim 4.5 \cdot 10^{-5}$, sufficient
-for quantum error correction.
+qubit frequency detuning, it achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
+sufficient for quantum error correction.
 
@@
 They achieve super-linear scaling in their gate error as
-a function of the gate duration which approaches zero at four multiples
-of $t_{Z/2}$. We observe that on a range of problems at small parameter detunings,
+a function of the gate duration which approaches zero at the larmor
+period $1 / f_{q}$. We observe that on a range of problems at small parameter detunings,
 the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order

### 2020-10-13 src/rbqoc.jl
numerical dissipation in viso_iso

@@ -208,3 +208,18 @@ function get_mat_iso(mat)
     return vcat(hcat(mat_r, -mat_i),
                 hcat(mat_i,  mat_r))
 end
+
+
+function get_mat_uniso(mat)
+    n = size(mat, 1)
+    nby2 = Integer(n/2)
+    i1 = 1:nby2
+    i2 = (nby2 + 1):n
+    return mat[i1, i1] + mat[i2, i1]
+end
+
+
+@inline get_vec_viso(mat) = reshape(mat, length(mat))
+
+
+@inline get_vec_unviso(vec) = reshape(vec, (Int(sqrt(length(vec))), Int(sqrt(length(vec)))))

### 2020-10-13 src/spin/figures.jl

@@ -54,16 +54,20 @@ const INVAL = 99999
 
 const F1_DATA = Dict(
     zpiby2 => Dict(
-        qoc => joinpath(SPIN_OUT_PATH, "spin15/00209_spin15.h5"),
+        qoc => "",# qoc => joinpath(SPIN_OUT_PATH, "spin15/00209_spin15.h5"),
         analytic => joinpath(SPIN_OUT_PATH, "spin14/00000_spin14.h5"),
     ),
     ypiby2 => Dict(
-        qoc => joinpath(SPIN_OUT_PATH, "spin15/00205_spin15.h5"),
-        analytic => joinpath(SPIN_OUT_PATH, "spin14/00003_spin14.h5"),
+        qoc => "",
+        analytic => "",
+        # qoc => joinpath(SPIN_OUT_PATH, "spin15/00205_spin15.h5"),
+        # analytic => joinpath(SPIN_OUT_PATH, "spin14/00003_spin14.h5"),
     ),
     xpiby2 => Dict(
-        qoc => joinpath(SPIN_OUT_PATH, "spin15/00239_spin15.h5"),
-        analytic => joinpath(SPIN_OUT_PATH, "spin14/00004_spin14.h5"),
+        qoc => "",
+        analytic => "",
+        # qoc => joinpath(SPIN_OUT_PATH, "spin15/00262_spin15.h5"),
+        # analytic => joinpath(SPIN_OUT_PATH, "spin14/00004_spin14.h5"),
     ),
 )
 const F1_PT_LIST = [analytic, qoc]
@@ -131,9 +135,8 @@ function gen_1b()
 end
 
 
-const F1C_GATE_COUNT = Integer(1.6e3)
+const F1C_GATE_COUNT = Integer(1e3)
 const F1C_AVG_COUNT = 10
-const F1C_DT = 1e-3
 function gen_1c(;use_previous=true)
     gate_types_integer = [Integer(gt) for gt in GT_LIST]
     gate_type_count = size(GT_LIST)[1]
@@ -174,10 +177,13 @@ function gen_1c(;use_previous=true)
             print("pt[$(j)]: $(pulse_type) ")
             save_file_path = F1_DATA[gate_type][pulse_type]
             save_file_paths[i, j] = save_file_path
-            save_file_path_sim = pulse_type == analytic ? nothing : save_file_path
             save_file_path_old = ((isnothing(save_file_paths_old) || i > gate_type_count_old
[…]

### 2020-10-13 src/spin/spin.jl

@@ -121,6 +121,8 @@ const GT_STR = Dict(
 
 const HDIM = 2
 const HDIM_ISO = HDIM * 2
+const HDIM_VISO = HDIM^2
+const HDIM_VISO_ISO = HDIM_VISO * 2
 struct SimParams
     controls :: Array{Float64, 2}
     control_knot_count :: Int64
@@ -161,6 +163,14 @@ const GAMMAC = 1 / 3e5 #GHz(T_c = 300 us)
 const SQRTLNIR = 4
 const NAMP_PREFACTOR = FBFQ_NAMP / FBFQ_A
 const GAMMAF_PREFACTOR = FBFQ_NAMP * SQRTLNIR * 2 * pi
+const PFIR_B1 = 0.049922035
+const PFIR_B2 = -0.095993537
+const PFIR_B3 = 0.050612699
+const PFIR_B4 = -0.004408786
+const PFIR_A1 = 1
+const PFIR_A2 = -2.494956002
+const PFIR_A3 = 2.017265875
+const PFIR_A4 = -0.522189400
 # coefficients are listed in descending order
 # raw coefficients are in units of seconds
 const FBFQ_T1_COEFFS = [
@@ -205,6 +215,7 @@ const FBFQ_T1_REDUCED_SPLINE_ITP = extrapolate(interpolate(
 
 # Define the system.
 # ISO indicates the object is defined in the complex to real isomorphism.
+# VISO indicates the object is defined in the vector and complex to real isomorphism.
 # NEGI is the negative complex unit.
 const NEGI = SA_F64[0   0  1  0 ;
                     0   0  0  1 ;
@@ -226,6 +237,20 @@ const NEGI_H1_ISO = pi * NEGI * SIGMAX_ISO
 const NEGI_H1_ISO_BIG = pi * NEGI * SMatrix{HDIM_ISO, HDIM_ISO, BigFloat}(SIGMAX_ISO)
 const FQ_NEGI_H0_ISO = FQ * NEGI_H0_ISO
 const AYPIBY2_NEGI_H1_ISO = AYPIBY2 * NEGI_H1_ISO
+const FQ_NEGI_H0 = FQ * -1im * pi * SIGMAZ
+const FQ_NEGI_H0_VISO = SMatrix{HDIM_VISO, HDIM_VISO}(
+    kron(I(HDIM), FQ_NEGI_H0)
[…]

### 2020-10-13 src/spin/spin13.jl

@@ -2,281 +2,228 @@
 spin13.jl - vanilla
 """
 
+WDIR = get(ENV, "ROBUST_QOC_PATH", "../../")
+include(joinpath(WDIR, "src", "spin", "spin.jl"))
+
+using Altro
 using HDF5
 using LinearAlgebra
-using TrajectoryOptimization
-import Plots
-using Printf
+using RobotDynamics
 using StaticArrays
+using TrajectoryOptimization
+const RD = RobotDynamics
+const TO = TrajectoryOptimization
 
-# Construct paths.
-EXPERIMENT_META = "spin"
-EXPERIMENT_NAME = "spin13"
-WDIR = ENV["ROBUST_QOC_PATH"]
-SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
+# paths
+const EXPERIMENT_META = "spin"
+const EXPERIMENT_NAME = "spin13"
+const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 
-# Plotting configuration.
-ENV["GKSwstype"] = "nul"
-Plots.gr()
-DPI = 300
+# problem
+const CONTROL_COUNT = 1
+const STATE_COUNT = 2
+const ASTATE_SIZE = STATE_COUNT * HDIM_ISO + 3 * CONTROL_COUNT
+const ACONTROL_SIZE = CONTROL_COUNT
+const INITIAL_STATE1 = [1., 0, 0, 0]
+const INITIAL_STATE2 = [0., 1, 0, 0]
+# state indices
+const STATE1_IDX = 1:HDIM_ISO
+const STATE2_IDX = STATE1_IDX[end] + 1:STATE1_IDX[end] + HDIM_ISO
+const INTCONTROLS_IDX = STATE2_IDX[end] + 1:STATE2_IDX[end] + CONTROL_COUNT
+const CONTROLS_IDX = INTCONTROLS_IDX[end] + 1:INTCONTROLS_IDX[end] + CONTROL_COUNT
+const DCONTROLS_IDX = CONTROLS_IDX[end] + 1:CONTROLS_IDX[end] + CONTROL_COUNT
+# control indices
+const D2CONTROLS_IDX = 1:CONTROL_COUNT
 
-function generate_save_file_path(save_file_name, save_path)
-    # Ensure the path exists.
-    mkpath(save_path)
-
-    # Create a save file name based on the one given; ensure it will
[…]

### 2020-10-13 src/spin/spin15.jl

@@ -2,6 +2,9 @@
 spin15.jl - T1 optimized pulses
 """
 
+WDIR = get(ENV, "ROBUST_QOC_PATH", "../../")
+include(joinpath(WDIR, "src", "spin", "spin.jl"))
+
 using Altro
 using HDF5
 using LinearAlgebra
@@ -9,9 +12,6 @@ using RobotDynamics
 using StaticArrays
 using TrajectoryOptimization
 const TO = TrajectoryOptimization
-
-WDIR = get(ENV, "ROBUST_QOC_PATH", "../../")
-include(joinpath(WDIR, "src", "spin", "spin.jl"))
 
 # paths
 const EXPERIMENT_META = "spin"
@@ -21,12 +21,12 @@ const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 # problem
 const CONTROL_COUNT = 1
 const STATE_COUNT = 2
-const ASTATE_SIZE_BASE = STATE_COUNT * STATE_SIZE_ISO + 3 * CONTROL_COUNT
+const ASTATE_SIZE_BASE = STATE_COUNT * HDIM_ISO + 3 * CONTROL_COUNT
 const INITIAL_STATE1 = [1., 0, 0, 0]
 const INITIAL_STATE2 = [0., 1, 0, 0]
 # state indices
-const STATE1_IDX = 1:STATE_SIZE_ISO
-const STATE2_IDX = STATE1_IDX[end] + 1:STATE1_IDX[end] + STATE_SIZE_ISO
+const STATE1_IDX = 1:HDIM_ISO
+const STATE2_IDX = STATE1_IDX[end] + 1:STATE1_IDX[end] + HDIM_ISO
 const INTCONTROLS_IDX = STATE2_IDX[end] + 1:STATE2_IDX[end] + CONTROL_COUNT
 const CONTROLS_IDX = INTCONTROLS_IDX[end] + 1:INTCONTROLS_IDX[end] + CONTROL_COUNT
 const DCONTROLS_IDX = CONTROLS_IDX[end] + 1:CONTROLS_IDX[end] + CONTROL_COUNT
@@ -80,7 +80,8 @@ function run_traj(;evolution_time=20., gate_type=zpiby2, time_optimal=false,
                   decay_aware=false, solver_type=altro, sqrtbp=false,
                   integrator_type=rk3, qs=ones(7), smoke_test=false,
[…]

### 2020-10-13 src/spin/spin18.jl

@@ -22,12 +22,14 @@ const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 # problem
 const CONTROL_COUNT = 1
 const STATE_COUNT = 2
-const SAMPLE_COUNT = 2
+const SAMPLE_COUNT = 8
 const ASTATE_SIZE_BASE = STATE_COUNT * HDIM_ISO + 3 * CONTROL_COUNT
 const ASTATE_SIZE = ASTATE_SIZE_BASE + SAMPLE_COUNT * HDIM_ISO
 const ACONTROL_SIZE = CONTROL_COUNT
 const INITIAL_STATE1 = [1., 0, 0, 0]
 const INITIAL_STATE2 = [0., 1, 0, 0]
+const INITIAL_STATE3 = [1., 0, 0, 1] ./ sqrt(2)
+const INITIAL_STATE4 = [1., -1, 0, 0] ./ sqrt(2)
 # state indices
 const STATE1_IDX = 1:HDIM_ISO
 const STATE2_IDX = STATE1_IDX[end] + 1:STATE1_IDX[end] + HDIM_ISO
@@ -168,8 +170,8 @@ function RD.discrete_dynamics(::Type{RD.RK3}, model::Model, astate::StaticVector
     controls = astate[CONTROLS_IDX[1]] + dt * astate[DCONTROLS_IDX[1]]
     dcontrols = astate[DCONTROLS_IDX[1]] + dt * acontrols[D2CONTROLS_IDX[1]]
 
-    hp_prop = exp(FQ_NEGI_H0_ISO + (camp + model.namp) * NEGI_H1_ISO) * dt)
-    hn_prop = exp(FQ_NEGI_H0_ISO + (camp - model.namp) * NEGI_H1_ISO) * dt)
+    hp_prop = exp((FQ_NEGI_H0_ISO + (camp + model.namp) * NEGI_H1_ISO) * dt)
+    hn_prop = exp((FQ_NEGI_H0_ISO + (camp - model.namp) * NEGI_H1_ISO) * dt)
     s1 = hp_prop * astate[S1_IDX]
     s2 = hp_prop * astate[S2_IDX]
     s3 = hp_prop * astate[S3_IDX]

### 2020-10-13 src/spin/spin23.jl

@@ -22,7 +22,7 @@ const TO = TrajectoryOptimization
 # paths
 const EXPERIMENT_META = "spin"
 const EXPERIMENT_NAME = "spin23"
-const SAVE_PATH = abspath(joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME))
+const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
 
 # problem
 const CONTROL_COUNT = 1
@@ -49,6 +49,7 @@ const S8_IDX = S7_IDX[end] + 1:S7_IDX[end] + HDIM_ISO
 const S9_IDX = S8_IDX[end] + 1:S8_IDX[end] + HDIM_ISO
 const S10_IDX = S9_IDX[end] + 1:S9_IDX[end] + HDIM_ISO
 const SAMPLE_COUNT = 10
+const SAMPLE_COUNT_INV = 1//10
 const ASTATE_SIZE = ASTATE_SIZE_BASE + SAMPLE_COUNT * HDIM_ISO
 const ACONTROL_SIZE = CONTROL_COUNT
 # control indices
@@ -68,7 +69,6 @@ end
 Model = Data.Model
 @inline RD.state_dim(model::Model) = ASTATE_SIZE
 @inline RD.control_dim(model::Model) = ACONTROL_SIZE
-
 
 # dynamics
 function RD.discrete_dynamics(::Type{RK3}, model::Model, astate::SVector{ASTATE_SIZE},
@@ -94,8 +94,9 @@ function RD.discrete_dynamics(::Type{RK3}, model::Model, astate::SVector{ASTATE_
     s9 = SVector{HDIM_ISO}(astate[S9_IDX])
     s10 = SVector{HDIM_ISO}(astate[S10_IDX])
     # compute state mean
-    sm = 1//10 .* (
-        s1 + s2 + s3 + s4 + s5 + s6 + s7 + s8 + s9 + s10
+    sm = SAMPLE_COUNT_INV .* (
+        s1 + s2 + s3 + s4 + s5
+        + s6 + s7 + s8 + s9 + s10
     )
     # compute state covariance
     d1 = s1 - sm
@@ -109,8 +110,9 @@ function RD.discrete_dynamics(::Type{RK3}, model::Model, astate::SVector{ASTATE_
     d9 = s9 - sm
     d10 = s10 - sm
[…]

### 2020-10-13 src/spin/spin25.jl

@@ -1,5 +1,5 @@
 """
-spin25.jl - unscented sampling robustness for the δa problem
+spin25.jl - unscented transform robustness for the δa problem
 """
 
 WDIR = joinpath(@__DIR__, "../../")
@@ -8,6 +8,7 @@ include(joinpath(WDIR, "src", "spin", "spin.jl"))
 using Altro
 using Distributions
 using HDF5
+using Hyperopt
 using ForwardDiff
 using LinearAlgebra
 using Random
@@ -29,137 +30,185 @@ const STATE_COUNT = 2
 const ASTATE_SIZE_BASE = STATE_COUNT * HDIM_ISO + 3 * CONTROL_COUNT
 const INITIAL_STATE1 = [1., 0, 0, 0]
 const INITIAL_STATE2 = [0., 1, 0, 0]
+const INITIAL_STATE3 = [1., 0, 0, 1] ./ sqrt(2)
+const INITIAL_STATE4 = [1., -1, 0, 0] ./ sqrt(2)
 # state indices
 const STATE1_IDX = 1:HDIM_ISO
 const STATE2_IDX = STATE1_IDX[end] + 1:STATE1_IDX[end] + HDIM_ISO
 const INTCONTROLS_IDX = STATE2_IDX[end] + 1:STATE2_IDX[end] + CONTROL_COUNT
 const CONTROLS_IDX = INTCONTROLS_IDX[end] + 1:INTCONTROLS_IDX[end] + CONTROL_COUNT
 const DCONTROLS_IDX = CONTROLS_IDX[end] + 1:CONTROLS_IDX[end] + CONTROL_COUNT
-const S1STATE1_IDX = DCONTROLS_IDX[end] + 1:DCONTROLS_IDX[end] + HDIM_ISO
-const S2STATE1_IDX = S1STATE1_IDX[end] + 1:S1STATE1_IDX[end] + HDIM_ISO
-const S3STATE1_IDX = S2STATE1_IDX[end] + 1:S2STATE1_IDX[end] + HDIM_ISO
-const S4STATE1_IDX = S3STATE1_IDX[end] + 1:S3STATE1_IDX[end] + HDIM_ISO
-const S5STATE1_IDX = S4STATE1_IDX[end] + 1:S4STATE1_IDX[end] + HDIM_ISO
-const S6STATE1_IDX = S5STATE1_IDX[end] + 1:S5STATE1_IDX[end] + HDIM_ISO
[…]

### 2020-10-13 src/spin/figures.jl
working f1c

@@ -54,20 +54,16 @@ const INVAL = 99999
 
 const F1_DATA = Dict(
     zpiby2 => Dict(
-        qoc => "",# qoc => joinpath(SPIN_OUT_PATH, "spin15/00209_spin15.h5"),
+        qoc => joinpath(SPIN_OUT_PATH, "spin15/00209_spin15.h5"),
         analytic => joinpath(SPIN_OUT_PATH, "spin14/00000_spin14.h5"),
     ),
     ypiby2 => Dict(
-        qoc => "",
-        analytic => "",
-        # qoc => joinpath(SPIN_OUT_PATH, "spin15/00205_spin15.h5"),
-        # analytic => joinpath(SPIN_OUT_PATH, "spin14/00003_spin14.h5"),
+        qoc => joinpath(SPIN_OUT_PATH, "spin15/00205_spin15.h5"),
+        analytic => joinpath(SPIN_OUT_PATH, "spin14/00003_spin14.h5"),
     ),
     xpiby2 => Dict(
-        qoc => "",
-        analytic => "",
-        # qoc => joinpath(SPIN_OUT_PATH, "spin15/00262_spin15.h5"),
-        # analytic => joinpath(SPIN_OUT_PATH, "spin14/00004_spin14.h5"),
+        qoc => joinpath(SPIN_OUT_PATH, "spin15/00262_spin15.h5"),
+        analytic => joinpath(SPIN_OUT_PATH, "spin14/00004_spin14.h5"),
     ),
 )
 const F1_PT_LIST = [analytic, qoc]
@@ -135,8 +131,8 @@ function gen_1b()
 end
 
 
-const F1C_GATE_COUNT = Integer(1e3)
-const F1C_AVG_COUNT = 10
+const F1C_GATE_COUNT = Integer(2e3)
+const F1C_AVG_COUNT = Integer(1e3)
 function gen_1c(;use_previous=true)
     gate_types_integer = [Integer(gt) for gt in GT_LIST]
     gate_type_count = size(GT_LIST)[1]
@@ -201,9 +197,9 @@ function gen_1c(;use_previous=true)
                     && k <= avg_count_old)
                     gate_errors[i, j, k, :] = gate_errors_old[i, j, k, :]
                 else
-                    res = run_sim_deqjl(
+                    res = run_sim_prop(
                         F1C_GATE_COUNT, gate_type; dynamics_type=dynamics_type,
-                        save_file_path=save_file_path, save=false, seed=k, dt_inv=Int(1e3)
[…]

### 2020-10-13 src/spin/figures.py

@@ -409,16 +409,19 @@ def make_figure1c():
 
     fig = plt.figure(figsize=(PAPER_TW * 0.4, PAPER_TW * 0.315))
     ax = plt.gca()
-    inax = fig.add_axes((0.71, 0.4, 0.25, 0.25))
+    inax = fig.add_axes((0.7, 0.69, 0.25, 0.25))
     for (i, gate_type) in enumerate(gate_types):
         for (j, pulse_type) in enumerate(pulse_types):
             color = PT_COLOR[pulse_type]
             linestyle = F1C_GT_LS[gate_type]
             gate_errors_ = gate_errors[i, j, :]
+            gate_errors_trans_ = [1e-8 if np.allclose(gate_errors_[k], 0) else gate_errors_[k]
+                                  for k in range(gate_errors_.shape[0])]
+            log_gate_errors_ = log_transform(gate_errors_trans_)
             alpha = F1C_ALPHA[i, j]
             ax.plot(gate_count_axis[inds], gate_errors_[inds],
                      color=color, linestyle=linestyle, linewidth=LW, alpha=alpha)
-            inax.plot(gate_count_axis[inds], log_transform(gate_errors_[inds]),
+            inax.plot(gate_count_axis[inds], log_gate_errors_[inds],
                       color=color, linestyle=linestyle, linewidth=LW, alpha=alpha)
         #ENDFOR
     #ENDFOR
@@ -427,28 +430,33 @@ def make_figure1c():
     ax.set_ylabel("Gate Error", fontsize=LABEL_FS)
     ax.set_xlabel("Gate Count", fontsize=LABEL_FS)
     ax.set_xlim(0, gate_count)
-    ax.set_xticks([0, 500, 1000, 1500])
-    ax.set_xticklabels(["0", "500", "1000", "1500"])
-    # ax.set_ylim(0, 1e-3)
+    xticks_ = [0, 500, 1000, 1500, 2000]
+    ax.set_xticks(xticks_)
+    ax.set_xticklabels([str(xtick) for xtick in xticks_])
+    ax.set_ylim(0, 5e-2)
     ax.tick_params(direction="in", labelsize=TICK_FS)
     ax.plot([], [], label="Z/2", linestyle=F1C_GT_LS[GateType.zpiby2], color="black")
[…]

### 2020-10-13 src/spin/spin.jl

@@ -127,7 +127,7 @@ struct SimParams
     controls :: Array{Float64, 2}
     control_knot_count :: Int64
     controls_dt_inv :: Int64
-    negi_h0 :: StaticMatrix{HDIM_ISO, HDIM_ISO}
+    negi_h0 :: SMatrix
     noise_offsets :: Array{Float64, 1}
     noise_dt_inv :: Int64
     sim_dt_inv :: Int64
@@ -237,20 +237,14 @@ const NEGI_H1_ISO = pi * NEGI * SIGMAX_ISO
 const NEGI_H1_ISO_BIG = pi * NEGI * SMatrix{HDIM_ISO, HDIM_ISO, BigFloat}(SIGMAX_ISO)
 const FQ_NEGI_H0_ISO = FQ * NEGI_H0_ISO
 const AYPIBY2_NEGI_H1_ISO = AYPIBY2 * NEGI_H1_ISO
-const FQ_NEGI_H0 = FQ * -1im * pi * SIGMAZ
-const FQ_NEGI_H0_VISO = SMatrix{HDIM_VISO, HDIM_VISO}(
-    kron(I(HDIM), FQ_NEGI_H0)
-    - kron(transpose(FQ_NEGI_H0), I(HDIM))
-)
-const FQ_NEGI_H0_VISO_ISO = SMatrix{HDIM_VISO_ISO, HDIM_VISO_ISO}(
-    get_mat_iso(kron(I(HDIM), FQ_NEGI_H0)
-                - kron(transpose(FQ_NEGI_H0), I(HDIM)))
-)
-const NEGI_H1 = -1im * pi * SIGMAX
-const NEGI_H1_VISO_ISO = SMatrix{HDIM_VISO_ISO, HDIM_VISO_ISO}(
-    get_mat_iso(kron(I(HDIM), NEGI_H1)
-                - kron(transpose(NEGI_H1), I(HDIM)))
-)
+const FQ_NEGI_H0_ = FQ * -1im * pi * SIGMAZ
+const FQ_NEGI_H0 = SMatrix{HDIM, HDIM}(FQ_NEGI_H0_)
+const FQ_NEGI_H0_VISO_ = kron(I(HDIM), FQ_NEGI_H0) - kron(transpose(FQ_NEGI_H0), I(HDIM))
[…]

### 2020-10-13 paper/main.tex
updated f1 figures

@@
 quantum (NISQ) applications.
-The $T_{1}$ optimized $X/2$ gate achieves a single gate error...
+
+The $T_{1}$ optimized $X/2$ gate achieves a factor of 3.5 reduction in
+its single gate probability of longitudinal relaxation over its analytic
+counterpart, commensurate with its
+reduction in single gate error $1 \cdot 10^{-4} \rightarrow 3.5 \cdot 10^{-5}$.
+The analytic $X/2$ gate is composed of the analytic $Z/2$ and $Y/2$ gates $-Y/2 \cdot Z/2 \cdot Y/2$.
+In principle the $T_{1}$ optimized $X/2$ gate should be able to perform at least as well
+as the analogous composite of the $T_{1}$ optimized $Z/2$ and $Y/2$ gates, achieving a factor
+of 5 reduction in its single gate probability and single gate errors.
+For all of the gates optimized here, the longitudinal relaxation probability is an effective
+metric in predicting performance in simulated relaxation and in decreasing the erro.
 
@@
     \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 3.568  & 0.727 & 4.905\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 3.107  & 0.763 & 4.072\\
-    X/2 & 16.251 & 4.488 & 3.621 & 10.542 & 3.563 & 3.449\\
+    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
+    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
+    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
   \end{tabular}

### 2020-10-26 src/spin/spin.jl
updates to spin11

@@ -429,9 +429,9 @@ end
 function integrate_prop_schroed!(gate_count::Int, states::Array{T, 2}, state::SVector, params::SimParams) where {T}
     states[1, :] = state
     dt = params.controls_dt_inv^(-1)
-    prop = I(HDIM)
+    prop = I(HDIM_ISO)
     for j = 1:params.control_knot_count
-        hamiltonian = params.negi_h0 + params.controls[j, 1] * NEGI_H1
+        hamiltonian = params.negi_h0 + params.controls[j, 1] * NEGI_H1_ISO
         prop_ = exp(hamiltonian * dt)
         prop = prop_ * prop
     end

### 2020-10-26 src/spin/spin11.jl

@@ -26,15 +26,25 @@ const ASTATE_SIZE_BASE = STATE_COUNT * HDIM_ISO + 3 * CONTROL_COUNT
 const ACONTROL_SIZE = CONTROL_COUNT
 const INITIAL_STATE1 = [1., 0, 0, 0]
 const INITIAL_STATE2 = [0., 1, 0, 0]
+const INITIAL_STATE3 = [1., 0, 0, 1] ./ sqrt(2)
+const INITIAL_STATE4 = [1., -1, 0, 0] ./ sqrt(2)
+const SAMPLE_COUNT = 4
 # state indices
 const STATE1_IDX = 1:HDIM_ISO
 const STATE2_IDX = STATE1_IDX[end] + 1:STATE1_IDX[end] + HDIM_ISO
 const INTCONTROLS_IDX = STATE2_IDX[end] + 1:STATE2_IDX[end] + CONTROL_COUNT
 const CONTROLS_IDX = INTCONTROLS_IDX[end] + 1:INTCONTROLS_IDX[end] + CONTROL_COUNT
 const DCONTROLS_IDX = CONTROLS_IDX[end] + 1:CONTROLS_IDX[end] + CONTROL_COUNT
-const DSTATE1_IDX = DCONTROLS_IDX[end] + 1:DCONTROLS_IDX[end] + HDIM_ISO
-const D2STATE1_IDX = DSTATE1_IDX[end] + 1:DSTATE1_IDX[end] + HDIM_ISO
-const D3STATE1_IDX = D2STATE1_IDX[end] + 1:D2STATE1_IDX[end] + HDIM_ISO
+const STATE3_IDX = DCONTROLS_IDX[end] + 1:DCONTROLS_IDX[end] + HDIM_ISO
+const STATE4_IDX = STATE3_IDX[end] + 1:STATE3_IDX[end] + HDIM_ISO
+const DSTATE1_IDX = STATE4_IDX[end] + 1:STATE4_IDX[end] + HDIM_ISO
+const DSTATE2_IDX = DSTATE1_IDX[end] + 1:DSTATE1_IDX[end] + HDIM_ISO
+const DSTATE3_IDX = DSTATE2_IDX[end] + 1:DSTATE2_IDX[end] + HDIM_ISO
+const DSTATE4_IDX = DSTATE3_IDX[end] + 1:DSTATE3_IDX[end] + HDIM_ISO
[…]

### 2020-10-26 paper/main.tex

--- 2020-10-12 paper/main.tex
@@
   for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimization.
+  using a computationally efficient metric in optimization and time-optimal control.
 \end{abstract}
@@
 \section{Introduction}
-textwidth: \printinunitsof{in}\prntlen{\textwidth}
-linewidth: \printinunitsof{in}\prntlen{\linewidth}
+%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
+%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
 %% P1 - general field, lots of citation. cite reviews, original qoc.
@@
 %% P4 - outline
-(The field) The field of quantum optimal control (QOC) is concerned
+The field of quantum optimal control (QOC) is concerned
 with efficiently and accurately manipulating quantum systems.
@@
 
-(Existing work) Many analytic and numerical techniques have been developed to control
+Many analytic and numerical techniques have been developed to control
 quantum systems. Analytic methods rely on solving the time-dependent
@@
 
-(This work) In this work, we employ state of the art
+In this work, we employ state of the art
 trajectory optimization methods
@@
 
-(Outline) As an example, we study the quantum optimal control problem on
+As an example, we study the quantum optimal control problem on
 the fluxonium qubit presented in \cite{zhang2020universal}.
+%%TODO: Fluxonium qubit is exciting and computationally feasible,
+%%but our methods extend to arbitrary platforms.
 We describe experimentally realistic constraints and map them to
@@
 \section{QOC on the Fluxonium}
-(Fluxonium Device) In the following we study
+In the following we study
 the quantum optimal control problem on the fluxonium qubit.
-In the two-level
+The fluxonium qubit is a promising building block for superconducting
+circuits, and the experimental constraints we encode in the optimization reflect
+those of a realistic device. Furthermore, the accurate
+two-level approximation of the system Hamiltonian makes it
+efficient to perform quantum optimal control on
+a classical computer. In the two-level
 approximation the system Hamiltonian takes the form
@@
 gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-system parameter deviations. We compare the gates we obtain with optimal
-control methods to the analytically constructed gates reported in
+systematic errors. We compare the gates we obtain with numerical
+methods to the analytically constructed gates reported in
 \cite{zhang2020universal} for the same device.
-%% The flux amplitude is obtained from the external flux by
-%% $a = 4 \pi \braket{g \lvert \hat{\phi} \rvert e} \rvert_{0.5 \Phi_{0}} E_{L}
-%% \delta \Phi_{\textrm{ext}} / (h \Phi_{0})$
-%% where $\hat{\phi}$ is the phase operator, $E_{L}$ is the characteristic inductance energy, $\Phi_{0}$
-%% is a flux quantum, and
-%% $\delta \Phi_{\textrm{ext}} = \Phi_{\textrm{ext}} - 0.5 \Phi_{0}$ is the flux
-%% offset from the flux frustration point.
 
-(Constraints) We now present the constraints we require each gate
-to obey.
+In addition to imposing optimization constraints that
+reflect physical limitations of the apparatus, we impose
+constraints that improve the experimental realization of the control pulse.
 To ensures gates may be concatenated arbitrarily without
@@
 
-In addition to constructing gates
-that obey experimental constraints and have high simulated fidelities,
-we also desire to make the gates robust to noise that affects the experimental
-gate fidelity. Decoherence of the quantum state due to external noise
+In addition to obeying experimental constraints
+and achieving low simulated gate errors, we desire
+to make our gates perform well in the presence of decoherence.
+Decoherence of the quantum state due to external noise
 is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
@@
 \section{Longitudinal Relaxation Awareness}
-(Strategy) We seek to minimize the probability
-that the qubit decays as a result of longitudinal
-relaxation. To this end we add the
-longitudinal relaxation probability
-to the augmented state vector
+The longitudinal relaxation time $T_{1}$ varies with
+control parameters in a range of superconducting circuit platforms,
+and it is advantageous to maximize it by tuning the controls.
+In the fluxonium, the longitudinal relaxation time varies with
+the amplitude $a$. To completely simulate
+the effect of longitudinal relaxation on the gate error requires
+propagating density matrices of size $n \times n$ under Master equation
+dynamics rather than state vectors of size $n$ under the TDSE dynamics.
+We avoid this increase in computational complexity by
+penalizing the probability that the qubit decays as a result of
+longitudinal relaxation. This modification only
+increases the size of the augmented state vector by a single real number
+and allows us to effectively mitigate the error arising from longitudinal
+relaxation. Additionally, for a constant $T_{1}$ time, a shorter gate duration
+would favor a lower longitudinal relaxation probability. We allow
+the optimizer to tune the gate duration in order to minimize the
+longitudinal relaxation probability. Our scheme for time-optimal
+control is applicable to any time-optimal problem, not only
+the one we study here.
+
+The longitudinal relaxation probability is given by
 \begin{equation}
@@
 \end{equation}
-Setting the target
-probability to zero results in
-a quadratic cost at each knot point
+$P_{1}$ is penalized using a quadratic cost at each knot point
 ${\lvert P_{1}(t_{k}) \rvert}^{2}$.
 $T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to
-experimental data of the form
-$\{(a, T_{1})\}$.
+a spline fit to experimental data of the form $\{(a, T_{1})\}$.
 Calculating $T_{1}$ directly from theoretical
@@
 Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation. Furthermore, this scheme for mitigating
-longitudinal relaxation requires one
-additional real number be added to the augmented state vector, avoiding
-the $O(n^{2})$ increase in computational and memory complexities
-associated with propagating a density matrix under the Lindblad
-master equation. Here $n$ is the size of the state vector.
+the fridge truth of the simulation.
 
-For a constant $T_{1}$ time, a shorter gate duration would
-favor a lower longitudinal relaxation probability.
 We allow the optimizer to tune the gate duration by
 making the time step between each knot point $\Delta t_{k}$
-a decision variable.
-Promoting $\Delta t_{k}$ to a decision variable, rather
+a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
 than the number of knot points $N$, preserves the
@@
 
-The optimizations for the $Y/2$ and $Z/2$ gates converge on similar solutions, a periodic
+The $T_{1}$ optimized gates converge on similar solutions, a periodic
 waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-Both solutions extend their gate times
-beyond their analytic counterpart, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. Both gates reduce
+They extend their gate times
+beyond their analytic counterparts, trading longer gate times for access
+to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
 their single gate probability of longitudinal relaxation by a factor of $5$ over
-their analytic counterparts, scaling similarly to their single gate error reduction
-$3 \cdot 10^{-5} \rightarrow 7 \cdot 10^{-6}$, see Appendix.
-The gate error reported in this text is the infidelity of the evolved state and the target state
-averaged over multiple pseudo-randomly generated initial states.
-Furthermore, they perform simlarly in
-the concatenated gate application comparision, maintaining total gate errors under $10^{-2}$
-over $10^{3}$ gate applications. Their single gate errors are sufficient for
-quantum error correction, requisite for fault-tolerant quantum computing,
-and their low gate errors for extend computations are critical for noisy, intermediate-scale
-quantum (NISQ) applications.
-The $T_{1}$ optimized $X/2$ gate achieves a single gate error...
+their analytic counterparts and achieve a commensurate reduction in their single gate error,
+see Appendix. The gate error reported in this text is the infidelity
+of the evolved state and the target state averaged over 1000 pseudo-randomly
+generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
+the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
+over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
+achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
+Both the analytic and optimized gates attain single gate errors sufficient for
+quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
+The low gate errors achieved by the optimized gates for extend computations
+are critical for noisy, intermediate-scale quantum (NISQ) applications.
+These improvements are significant for the realistic constraints we have put on the
+device, and do not represent a fundamental limit to the optimization methods we have
+employed.
 
@@
   \caption{
-    (a) Z/2 gates constructed using the
-    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}
-    and 2\textsuperscript{nd} order derivative methods. The gates
-    shown for the robust methods are the solutions at twice the analytic
+    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
+    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
+    and 2\textsuperscript{nd}-order derivative methods. The gates
+    shown for the numerical methods are the solutions at twice the analytic
     gate time.
@@
     a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods.
+    detuning from the nominal qubit frequency for all methods. Missing
+    points represent solutions with a gate error above $5 \cdot 10^{-5}$.
     (c) Single gate error as a function of the detuning from the nominal
     qubit frequency at multiples of the analytic gate duration for the analytic
-    and 1\textsuperscript{st} order derivative methods.
+    and 1\textsuperscript{st}-order derivative methods.
   }
@@
 
+%% S5
+\section{Robustness to Static Parameter Deviations}
+We have formulated the quantum optimal control
+problem as an open loop optimization problem, i.e.
+feedback from the experiment is not incorporated in optimization.
+However, the device typically deviates from the model we use in optimization,
+the Hamiltonian, leading to poor experimental performance. To mitigate
+these errors we employ robust control techniques
+to make the state evolution insensitive to Hamiltonian
+parameter deviations. As an example
+we study mitigating errors arising from the finite measurement
+precision of the qubit frequency $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$.
+We consider three flavors of robust control
+techniques. The first is the sampling method, which has
+been preposed previously in the context of QOC
+\cite{rembold2020introduction, reinhold2019controlling}. In the
+sampling method, multiple trajectories are evolved under distinct deviant dynamics
+to capture the effect of parameter deviations. We also
+study the unscented sampling method which uses the unscented
+transformation to accurately propagate a distribution
+representing the evolving state as a result of a parameter deviation.
+The unscented transformation was designed for nonlinear Kalman
+filtering and is frequently utilized in robust control
+\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
+Finally, we propose the derivative method. This scheme
+uses derivative information about how the state
+evolution is affected by the deviant parameter to
+mitigate trajectory deviations.
+
+In the sampling method we propagate additional sample states
+in the augmented state vector that follow deviant dynamics
+and penalize their differences from the target state.
+The states evolve under the deviant
+Hamiltonian where the parameter of interest is replaced by
+a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
+deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
+The infidelities of the samples and the target
+state are penalized at each knot point
+$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
+We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
+resulting in eight additional states added to the augmented state vector.
+The four initial states are chosen so
+that their outer products span the operators on the
+Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
+(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
+This results in $O(n^{2})$ states added to the augmented
+state vector where $n$ is both the dimension of the Hilbert space
+and the size of the state vector.
+
 %% F3
-\begin{figure}[h]
-  \begin{subfigure}{\linewidth}
+\begin{figure*}[ht]
+  \begin{subfigure}{.4\textwidth}
     \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}
-
-  \begin{subfigure}{\linewidth}
+  \end{subfigure}\hspace{0.025\textwidth}
+  \begin{subfigure}{.4\textwidth}
     \includegraphics[width=\linewidth]{assets/f3b.png}
@@
   \caption{
-    (a) $X/2$ gates robust to flux offsets constructed using the analytic,
-    (2, 4)-point sampling, and (2\textsuperscript{nd}, 3\textsuperscript{rd})-order
-    derivative methods at a gate duration
-    of $56.8$ns. (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is measured
+    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
+    8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
+    and 2\textsuperscript{nd}-order derivative methods. The gates shown
+    are the solutions at the analytic gate time.
+    (b) Simulation of stochastic 1/$f$ flux noise for
+    successive gate applications. The reported gate error is computed
     after each gate application.
   }
-\end{figure}
+\end{figure*}
 
-%% S5
-\section{Robustness to Static Parameter Deviations}
-Gate errors may be incurred if Hamiltonian parameters deviate
-from their nominal values in the experiment. For example, the qubit frequency
-may be measurable up to a known precision $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$
-and drift in laboratory temperatures may increase
-amplitude gain in the electronics chain $a \gets a (1 + \epsilon)$.
-Multiple analytic methods have been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
+The strategy of the unscented sampling method is to
+propagate an ensemble of sample states (sigma points) which represent a distribution
+over every element of $\ket{\psi}$. The distribution
+models the uncertainty in each state element arising from the
+parameter deviation. The ensemble consists of $2 (n + d)$ states
+where $d$ is the number of deviant parameters. Each state in the
+ensemble is propagated to the next
+knot point using separate deviant dynamics. Then, the mean and covariance of
+the propagated ensemble is calculated. New states are sampled from the distribution
+given by the calculated statistics for propagation to the next knot point.
+This resampling procedure, the unscented transformation, accurately propagates
+first and second moments of the distribution and ensures the points
+lie on the $\sqrt{n}$\textsuperscript{th}
+covariance contour at each knot point. A detailed update procedure is given
+in the appendix. Each initial state from the operator basis
+is represented with a distribution of $2 (n + d)$ samples. This results
+in $O(n^{3})$ states added to the augmented state vector.
+The costs at each knot point are the infidelities between the samples
+and the target state, as in the sampling method.
 
-We consider three numerical strategies for achieving robustness
-to parameter deviations. The first we call the
-derivative method. We draw on the intuition that
+The derivative method draws on the intuition that
 the sensitivity of the state evolution to a parameter
@@
   \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-
 The state derivatives could be obtained with backwards mode differentiation.
@@
 we forward propagate the state derivatives in the
-augmented state vector under coupled equations of motion, resulting
-in $O(d N)$ matrix multiplications.
-For example, the equations of motion for the 1\textsuperscript{st}-order derivative
-method are
+augmented state vector under coupled dynamics, resulting
+in $O(d N)$ matrix multiplications. For the piece-wise constant control
+scheme we employ here, $N$ is the number of control parameters. Typically
+$d \ll N$ so this is a good tradeoff. As an example, the dynamics
+for the 1\textsuperscript{st}-order derivative method are
 \begin{align}
@@
 
-The second scheme we analyze is the sampling method,
-which has been proposed previously for
-QOC \cite{reinhold2019controlling}.
-We propagate additional states in the augmented state vector that follow deviant dynamics
-and penalize their differences from the target state.
-The states evolve under the deviant
-Hamiltonian where the parameter of interest is replaced by
-a deviant value $\lambda \gets \lambda + \delta \lambda$.
-We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The infidelities of the samples and the target
-state are penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
-resulting in eight additional states added to the augmented state vector.
-The four initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-  (\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}. We observe
-that sampling only the basis states $\{\ket{0}$, $\ket{1}\}$
-is insufficient because, as in the case of a $Z$ rotation, their
-trajectories are insensitive to some parameter deviations, making
-them undesirable samples for those operations.
-
-Additionaly, we consider the unscented sampling method,
-which is well-studied in the robust control iterature
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-The strategy is to
-propagate an ensemble of states (sigma points) which represent a distribution
-over every element of $\ket{\psi}$. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states
-where $n$ is the size of the state vector and $d$ is the number of
-deviant parameters. Each state in the ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. The initial state of all $2 (n + d)$ samples is
-$(\ket{0} + i\ket{1} / \sqrt{2})$ with added noise. Repeating this procedure for the other
-three operator basis states would provide the optimizer with more information,
-but come at a computational cost which far exceeds that of the other methods,
-resulting in $O(n^2)$ additional states in the augmented state vector. The
-costs at each knot point are the infidelities
-between the samples and the target state, as in the
-basic sampling method.
-
 To demonstrate the applicability of these techniques
@@
 gate subject to a constant qubit frequency detuning
-($f_{q} \gets f_{q} + \delta f_{q}$).
+$f_{q} \gets f_{q} + \delta f_{q}$.
 We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
@@
 one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning.
-
-We compare the robust methods to an analytically derived $Z/2$ pulse which corresponds to
-idling at the flux frustration point $a = 0$. The analytic pulse
+stated qubit frequency detuning $\delta f_{q}$. We compare the numerical methods
+to an analytically derived $Z/2$ gate which corresponds to
+idling at the flux frustration point $a = 0$. The analytic gate
 is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
@@
 is critical for operating multi-qubit experiments in the lab frame.
-These numerical methods give an effective mechanism for synchronizing
+These numerical methods offer an effective scheme for synchronizing
 qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
@@
 achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $\sim 1 / 2 f_{q}$ after which it achieves a consistent
+until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
 gate error $\sim 3.5 \cdot 10^{-5}$.
@@
 a function of the gate duration which approaches zero at the larmor
-period $1 / f_{q}$. We observe that on a range of problems at small parameter detunings,
+period $1 / f_{q}$. We observe that on a range of problems with small parameter detunings,
 the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
@@
 
-%% T2
-%% TODO: Appendix
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
+%% TODO: how do I tie this in?
+Multiple analytic methods have also been presented to construct gates
+that mitigate the error arising from parameter deviations
+including composite pulse sequences \cite{merrill2014progress},
+the DRAG scheme \cite{krantz2019quantum}, and
+geometric phase considerations
+\cite{xu2020nonadiabatic, han2020experimental}.
+Composite pulse sequences are derived by computing
+the erroneous rotation arising from the static parameter deviation
+with a finite order magnus expansion.
+Pulses are then suitably composed to eliminate the
+error. In principle the error may be eliminated
+to arbitrary order with sufficiently many pulses.
+It is difficult to choose an appropriate composite pulse
+for comparison on the problem we study here. We
+propose comparing numerical methods to existing analytic
+techniques for future work.
 
@@
 \section{Robustness to Stochastic Parameter Deviations}
-An additional source of error can be modeled by stochastic
-Hamiltonian parameter deviations. For the fluxonium qubit, magnetic flux noise
-modifies the flux amplitude from its nominal value by an amount $\delta a$.
+An additional source of experimental error arises from stochastic
+Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
+superconducting circuit elements, magnetic flux noise is a significant
+source of coherent errors. Magnetic flux noise
+modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
+It is well studied that the spectral density of $\delta a$ follows a
+1/$f$ distribution for a range of devices, consisting primarily of low frequency
+noise (citations needed). Analytic methods to combat 1/$f$ noise
+take advantage of the low frequency characteristic and
+treat the noise as quasi-static, performing generalizations of the spin-echo technique
+to compensate for erroneous drift. This is the strategy employed by the analytic gate
+considered here.
 
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution, consisting primarily of low frequency noise. We compare
-the methods discussed for static parameter deviations to the task of
-realizing a $X/2$ gate subject to 1/$f$ flux noise. We simulate successive
-applications of the gate constructed by each method and measure the gate error
-after each application (see Figure 3b). The system is subject to flux noise
-$\delta a (t)$ that is generated from a standard normal distribution, normalized
-by 1/$f^{1/2}$ in frequency space, and modulated by the experimentally obtained
-flux noise amplitude $A_{\Phi} = 5.21 \mu \Phi_{0}$.
+We compare the numerical methods discussed in the previous section
+on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
+The noise is generated by
+filtering white noise sampled from a standard normal distribution using a finite
+impulse response filter. It is then scaled by the 
+flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
+\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
+The unscented sampling method is modified so that its sampled deviations
+follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
+in the augmented state vector. In principle the basic sampling method could be modified
+similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
+methods require no modification from the static case.
 
-(Results) ...
+We simulate successive applications of the gate constructed by each method and compute the gate error
+after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
+sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
+numerical methods perform similarly in the concatenated gate application comparision, achieving a two
+order of magnitude reduction in gate error over the analytic gate at $200$
+gate applications $\sim 11 \mu\textrm{s}$.
+Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
+an effective method for mitigating them.
 
@@
 \begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance,
+  The authors would like to thank Helin Zhang for experimental assistance
   and Daniel Weiss for useful discussions.
@@
 %% AC
-\section{Dissipation Simulation}
+\section{Dissipation}
 We model dissipation using the Lindblad master
@@
 %% AD
-\section{Derivative Method Dynamics}
+\section{Derivative Method}
 To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
@@
 
+%% T2
+\begin{table}
+  \begin{tabular}{c | c | c}
+    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
+    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
+    \hline
+    D-1 & 2.051 & 17.935\\
+    D-2 & 2.511 & 7.628\\
+  \end{tabular}
+  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
+    for $Z/2$ gates optimized using the derivative methods. The norms are computed
+    at the end of the gate's duration $t_{N} = 38$ns.}
+\end{table}
+
 %% AE

### 2020-10-30 paper/main.tex
intro is separated by contributions

@@
   In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
-  takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by transforming
-  higher-order differential equations into a set of coupled first-order equations.
+  to achieve high fidelity gates for parameter spreads on the order of $1\%$.
+  We propose a method that takes advantage of deviant parameter derivative information while maintaining
+  computational efficiency by utilizing mixed-mode differentation.
   Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
   the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm.
-  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
-  for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimization and time-optimal control.
+  of the optimization algorithm. We propose a computationally efficient metric
+  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
+  We apply these techniques to a fluxonium qubit with realisitc experimental constraints
+  and noise values. We are able to reduce gate errors by orders of magnitudes in the presence
+  of deviant parameters, achieving super-linear scaling for gate robustness as a function of the gate duration.
+  We use the same techniques to mitigate coherent errors due to time-varying magnetic flux noise
+  by two orders of magnitude. Additionally, we mitigate the gate error arising from longitudinal
+  relaxation by a factor of $5$. Further, we find that the numerical techniques we employ
+  are able to perform phase gates in arbitrary times, which is critical for operating multi-qubit
+  experiments in the lab frame.
 \end{abstract}
@@
 \cite{khaneja2005optimal}, and applications now include superconducting
-qubits \cite{heeres2017implementing,
+circuits \cite{heeres2017implementing,
   leng2019robust, leung2017speedup, xu2020nonadiabatic},
@@
 diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}.
-In quantum control theory, as in classical control theory,
-optimization is performed
-to minimize an objective function. Relevant for the field
-of quantum computation is the task of achieving high fidelity quantum gates.
-In this application, the objective function includes contributions from the
-gate error associated with the state transfer of the system as well as contributions
-due to experimental constraints, such as the control amplitude.
+\cite{sorensen2018quantum}. QOC techniques are used to
+control a system such that a set of optimization objectives are maximally fulfilled.
+For quantum computation, relevant objectives include achieving high fidelity
+gates and adhering to experimental constraints.
 The decision variables of the optimization problem are the time-dependent control
 parameters, particular to the quantum system, that govern its evolution.
+It is desirable to make the optimization aware of experimental errors
+such as noise, system model errors, and control errors, to
+improve experimental performance on the optimization objectives.
+The field of robust control, deriving from classical control theory, has developed
+techniques to account for these errors in optimization (citations appreciated).
+In this work we employ robust control techniques to mitigate
+releastic decoherence and systematic errors that arise when controlling
+a superconducting fluxonium qubit.
 
-Many analytic and numerical techniques have been developed to control
-quantum systems. Analytic methods rely on solving the time-dependent
-Schroedinger equation analytically or deriving equations of motion
-from a suitable Lagrangian \cite{zhang2020universal, huang2020engineering, han2020experimental,
-  xu2020nonadiabatic, carlini2005quantum}. Analytic methods
-have been presented to address decoherence in quantum systems.
-Considering the dynamical and geometric phases of state
-evolution has led to methods for achieving
+Numerical techniques for QOC have been remarkably successful in
+state preparation, designing high fidelity gates, and determining
+optimized pulse sequences (citations needed). Many popular numerical techniques
+are iterative, indirect shooting
+methods \cite{leung2017speedup, goer2019krotov, abdelhafez2019gradient,
+  machnes2015gradient}.
+An initial guess for the control parameters is used to forward-simulate
+the quantum state trajectory under the time-dependent Schroedinger equation. Then, gradients
+with respect to the controls are computed for an objective evaluated on
+the evolved quantum state. The gradients are used to update
+the control parameters and the procedure is repeated iteratively.
+A convenient way to enforce constraints in this framework is
+to employ control filtering or project gradients onto the constraint manifold.
+In this work we perform optimization using ALTRO, which combines
+an iterative indirect shooting method with the augmented Lagrangian
+method \cite{howell2019altro}. The augmented Lagrangian method
+allows us to enforce arbitrary, simultaneous constraints and achieve
+fast convergence without being restricted to the constraint
+manifold.
+
+QOC techniques have been effective in mitigating the effects
+of systematic errors. Typically, the model used in optimization
+differs from the experimental system due to parameter drift, the inability
+to realize control parameters exactly, and noise.
+These systematic errors cause the trajectory of the quantum system
+in the experiment to deviate from that predicted by the optimization.
+A set of controls is said to be robust if the experimental state trajectory
+it produces is insensitive to systematic errors.
+Considering the dynamical and geometric phases of quantum state
+evolution has led to analytic methods for achieving
 robustness to dynamical errors and mitigating pure dephasing
-\cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
-Additionally, suitable choices of bases have been proposed that
-allow the experimentalist to infer tradeoffs in longitudinal relaxation
-and pure dephasing coherence times \cite{huang2020engineering}.
-To date, however, no analytical method has been presented
-to construct a gate while satisfying arbitrary experimental
-constraints. Furthermore, most analytic methods have only
-been demonstrated in the two-level approximation and become
-cumbersome to solve for larger Hilbert space sizes.
+\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
+We mitigate systematic errors numerically by employing
+open loop optimization. We draw on robust trajectory optimization
+techniques to encode systematic errors in optimization objectives.
+We consider three
+techniques. The first is the sampling method, which has been applied
+previously in the context of QOC
+\cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
+The second is the unscented sampling
+method, which derives from the unscented transformation used
+for non-linear Kalman filtering
+\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
+We propose a third method, which we call the derivative method.
+This method takes advantage of derivative information of systematic
+errors with respect to the state evolution. We make this method efficient
+by employing mixed-mode differentiation.
 
-Numerical quantum control techniques have developed in parallel
-with analytic frameworks.
-Most numerical techniques proposed for QOC are indirect shooting
-methods where an initial guess for the control trajectory is forward-simulated,
-e.g. under the time-dependent Schroedinger equation,
-and first-order gradients are computed for a functional evaluated
-on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
-  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
-Most techniques formulate the
-problem as unconstrained or use
-projective gradient methods to enforce constraints
-\cite{machnes2015gradient}.
-The latter approach relies on constructing
-bases for constraint spaces--which is not always feasible--and
-may hinder convergence.
-Methods for mitigating decoherence using numerical
-techniques have employed master equation dynamics (citation needed),
-Monte Carlo style trajectories \cite{abdelhafez2019gradient},
-and proposals have been made to simulate multiple trajectories
-\cite{reinhold2019controlling, rembold2020introduction}.
+Further, many analytic and numerical techniques have been developed to mitigate decoherence.
+Floquet techniques have been experimentally demonstrated to simultaneously mitigate
+decoherence due to longitudinal relaxation and pure dephasing
+\cite{huang2020engineering, mundada2020floquet}.
+Numerical techniques have mitigated decoherence by evolving the
+system using master equations \cite{rembold2020introduction} and employing
+Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
+In this work we take advantage of the known dependence of the
+quantum system's controls on longitudinal relaxation. We encode
+the decoherence in an efficient optimization objective that does
+not pay the increased computational cost of integrating a master equation.
 
-In this work, we employ state of the art
-trajectory optimization methods
-to formulate the quantum
-optimal control problem as a constrained optimization problem, allowing
-us to enforce arbitrary experimental constraints.
-Additionally,
-we introduce methods for constructing gates that mitigate longitudinal
-relaxation and are robust to experimental errors, and therefore
-pure dephasing, the dominant barriers to experimentally
-realizing practical quantum computing.
-
-As an example, we study the quantum optimal control problem on
-the fluxonium qubit presented in \cite{zhang2020universal}.
-%%TODO: Fluxonium qubit is exciting and computationally feasible,
-%%but our methods extend to arbitrary platforms.
-We describe experimentally realistic constraints and map them to
-the trajectory optimization framework. Next, we
-outline a method for making the optimization longitudinal
-relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set.
-We present two methods for achieving robustness to pure dephasing
-and compare them to existing dynamic decoupling methods. We find that
-the methods we employ are able to produce high fidelity gates when
-subject to parameter spreads of $1\%$.
+We compare our techniques using the heavy fluxonium qubit presented in
+in \cite{zhang2020universal}. This quantum system is an exciting
+platform for quantum computation due to its long longitudinal relaxation
+and pure dephasing times at the flux frustation point. This system is also
+efficient to simulate due to its accurate two-level approximation.
+We emphasize that the numerical techniques we employ extend to arbitrary quantum systems,
+not only the one we study here. We describe experimentally realistic constraints for this system
+and map them to the ALTRO framework. Next, we
+outline a method for making the optimization aware of longitudinal
+relaxation. We achieve a factor of 5 increase in longitudinal relaxation times
+over the baseline gate set. Then we present three methods for achieving
+robustness to systematic errors. We find that we are able to
+decrease the gate error arising from systematic errors super-linearly
+in the gate duration, achieving orders of magntidue reduction in gate errors.
+We also find that the numerical techniques we employ are able to produce
+fixed phase gates in arbitrary times, a method that could be used
+to control multi-qubit systems in the lab frame. Finally,
+we employ the robust control techniques to mitigate time-varying
+magnetic flux noise. We achieve a two order of magnitude improvement
+in gate errors for extended computations, an improvement critical
+for noisy, intermediate-scale quantum (NISQ) applications.
 
@@
 
+%% TODO: talk about quadratic cost vs. infidelity
+
 %% F1
@@
     (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is computed after
-    each gate appliaction. The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
+    for successive gate applications. The reported gate error is the cumulative
+    gate error after each gate application.
+    The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
     in the figure.
@@
 The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms,
-and it is advantageous to maximize it by tuning the controls.
-In the fluxonium, the longitudinal relaxation time varies with
-the amplitude $a$. To completely simulate
-the effect of longitudinal relaxation on the gate error requires
-propagating density matrices of size $n \times n$ under Master equation
-dynamics rather than state vectors of size $n$ under the TDSE dynamics.
+control parameters in a range of superconducting circuit platforms.
+It is advantageous to tune the controls to extend the longitudinal
+relaxation time.
+To calculate the gate error due to longitudinal relaxation
+requires propagating density matrices of size $n \times n$ under Master equation
+dynamics, rather than state vectors of size $n$ under the TDSE dynamics.
 We avoid this increase in computational complexity by
-penalizing the probability that the qubit decays as a result of
-longitudinal relaxation. This modification only
-increases the size of the augmented state vector by a single real number
-and allows us to effectively mitigate the error arising from longitudinal
-relaxation. Additionally, for a constant $T_{1}$ time, a shorter gate duration
+penalizing the integrated rate of longitudinal relaxation,
+i.e. the probability of longitudinal relaxation.
+Using this probability as proxy for the gate error incurred
+is reasonable because losses due to longitudinal relaxation
+increase monotonically in time.
+This technique can be extended to
+error channels which share the monotonically increasing property.
+Additionally, for a constant $T_{1}$ time, a shorter gate duration
 would favor a lower longitudinal relaxation probability. We allow
@@
 a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-Calculating $T_{1}$ directly from theoretical
-considerations requires many high-rank
-eigendecompositions, which
-is computationally expensive. Additionally,
-$T_{1}$ values are known to fluctuate greatly
+It is also possible to use a spline fit to theoretically obtained data.
+However, $T_{1}$ values are known to fluctuate greatly
 with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation.
+Interpolating $T_{1}$ from experimental data
+increases the fridge truth of the simulation.
 
@@
 
-The $T_{1}$ optimized gates converge on similar solutions, a periodic
+We compare the numerical method we have developed to the analytic gates
+on the task of achieving low gate errors in the presence of longitudinal relaxation
+for the $Z/2$, $Y/2$, and $X/2$ gates.
+The numerically optimized gates converge on similar solutions, a periodic
 waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
@@
 to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate probability of longitudinal relaxation by a factor of $5$ over
-their analytic counterparts and achieve a commensurate reduction in their single gate error,
-see Appendix. The gate error reported in this text is the infidelity
+their single gate errors by a factor of $5$ over
+their analytic counterparts which is commensurate to their
+probability of longitudinal relaxation reductions, see Appendix.
+The gate error reported in this text is the infidelity
 of the evolved state and the target state averaged over 1000 pseudo-randomly
@@
 are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have put on the
-device, and do not represent a fundamental limit to the optimization methods we have
+These improvements are significant for the realistic constraints we have imposed
+on the gates, and do not represent a fundamental limit to the optimization methods we have
 employed.
@@
     (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
+    analytic, sampling, unscented sampling, and the 1\textsuperscript{st}-
     and 2\textsuperscript{nd}-order derivative methods. The gates
@@
     detuning from the nominal qubit frequency for all methods. Missing
-    points represent solutions with a gate error above $5 \cdot 10^{-5}$.
+    data points represent solutions with a gate error above $5 \cdot 10^{-5}$.
     (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency at multiples of the analytic gate duration for the analytic
-    and 1\textsuperscript{st}-order derivative methods.
+    qubit frequency. The solutions for the analytic and
+    1\textsuperscript{st}-order derivative methods are shown at multiples
+    of the analytic gate time. The performance of the two methods is
+    indistinguishable at the analytic gate time $18$ns.
   }
@@
 feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the model we use in optimization,
-the Hamiltonian, leading to poor experimental performance. To mitigate
+However, the device typically deviates from the Hamiltonian we use in optimization,
+leading to poor experimental performance. To mitigate
 these errors we employ robust control techniques
@@
 parameter deviations. As an example
-we study mitigating errors arising from the finite measurement
-precision of the qubit frequency $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$.
-We consider three flavors of robust control
-techniques. The first is the sampling method, which has
+we mitigate errors arising from the drift and finite measurement
+precision of the qubit frequency $\tilde{f_{q}} = f_{q} \pm \sigma_{f_{q}}$.
+We consider three robust control techniques.
+The first is the sampling method, which has
 been preposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling}. In the
-sampling method, multiple trajectories are evolved under distinct deviant dynamics
+\cite{rembold2020introduction, reinhold2019controlling, carvalho2020error}. In the
+sampling method, multiple states are evolved under distinct deviant dynamics
 to capture the effect of parameter deviations. We also
@@
 transformation to accurately propagate a distribution
-representing the evolving state as a result of a parameter deviation.
+representing the uncertainty in an evolving state
+due to a parameter deviation.
 The unscented transformation was designed for nonlinear Kalman
@@
 \cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. This scheme
-uses derivative information about how the state
-evolution is affected by the deviant parameter to
-mitigate trajectory deviations.
+Finally, we propose the derivative method. Derivative
+information encoding the sensitivity of the
+state trajectory with respect to the deviant parameter
+is used to penalize state trajectory deviations.
 
-In the sampling method we propagate additional sample states
-in the augmented state vector that follow deviant dynamics
-and penalize their differences from the target state.
-The states evolve under the deviant
-Hamiltonian where the parameter of interest is replaced by
+In the sampling method, sample states evolve under a
+Hamiltonian where a parameter is replaced by
 a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The infidelities of the samples and the target
-state are penalized at each knot point
+deviant values $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
+The gate error of each sample state is 
+penalized at each knot point
 $1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
-resulting in eight additional states added to the augmented state vector.
-The four initial states are chosen so
+Each initial state is used to initialize two
+sample states $\ket{\psi^{\pm}}$.
+The initial states are chosen so
 that their outer products span the operators on the
@@
 (\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-This results in $O(n^{2})$ states added to the augmented
-state vector where $n$ is both the dimension of the Hilbert space
-and the size of the state vector.
+For this method the number of states in the augmented state vector
+scales as $O(dn^{2})$ because each of $n^{2}$ initial
+states is represented by $2d$ samples. $n$ is both the dimension of the Hilbert space
+and the size of the state vector. $d$ is the number of deviant parameters.
 
@@
     (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
+    sampling, unscented sampling, and the 1\textsuperscript{st}-
     and 2\textsuperscript{nd}-order derivative methods. The gates shown
@@
     (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is computed
-    after each gate application.
+    successive gate applications. The reported gate error is the cumulative
+    gate error after each gate application.
   }
@@
 propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of $\ket{\psi}$. The distribution
+over every element of the initial state. The distribution
 models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states
-where $d$ is the number of deviant parameters. Each state in the
+parameter deviation. The ensemble consists of $2 (n + d)$ states.
+Each state in the
 ensemble is propagated to the next
@@
 in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. This results
-in $O(n^{3})$ states added to the augmented state vector.
-The costs at each knot point are the infidelities between the samples
-and the target state, as in the sampling method.
+is represented with a distribution of $2 (n + d)$ samples. Hence,
+the number of states in the augmented state vector scales as $O(n^{3} + dn^{2})$.
+The gate error of each sample state is penalized as in the sampling method.
 
@@
 to obtain the state derivative at knot point $k$.
-For $d$ sensitivity parameters and $N$ knot points
-this requires $O(d N^2)$ matrix multiplications. Instead,
-we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting
-in $O(d N)$ matrix multiplications. For the piece-wise constant control
-scheme we employ here, $N$ is the number of control parameters. Typically
-$d \ll N$ so this is a good tradeoff. As an example, the dynamics
+For a single state derivative and $N$ knot points this
+requires $O(N^2)$ matrix multiplications.
+Instead, we forward propagate the state derivatives in the
+augmented state vector under coupled dynamics, resulting in
+$O(N)$ matrix multiplications. For example, the dynamics
 for the 1\textsuperscript{st}-order derivative method are
@@
 \cite{berland2005solving, einkemmer2017performance}.
+For this method $m$ state derivatives are associated with each
+initial state from the operator basis.
+So, the number of states in the augmented state vector scales as
+$O(dmn^{2})$.
 
@@
 one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$. We compare the numerical methods
-to an analytically derived $Z/2$ gate which corresponds to
+stated qubit frequency detuning $\delta f_{q}$.
+
+We compare the numerical methods
+to an analytically derived $Z/2$ gate. This gate corresponds to
 idling at the flux frustration point $a = 0$. The analytic gate
 is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. It is linearly sensitive to qubit frequency detunings,
-the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
-qubit frequency detuning, it achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-sufficient for quantum error correction.
+is simple to derive. Its erroneous rotation angle $2 \pi t_{Z/2} \delta f_{q}$ is linearly sensitive to
+the qubit frequency detuning, resulting in a gate error that is quadratically sensitive
+to the qubit frequency detutning.
+At a one-percent
+qubit frequency detuning the analytic gate achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
+which is sufficient for quantum error correction, see Figure 2b.
+Although the analytic $Z/2$ gate performs well, it works
+only at the gate time $t_{Z/2}$. The ability to perform $Z$ rotations in arbitrary times is critical
+for operating multi-qubit experiments in the lab frame.
+Each numerical method is able to find solutions at
+all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times,
+see Figure 2b. These numerical methods offer an effective scheme for synchronizing
+qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
 
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
-The ability to perform $Z$ rotations in arbitrary times
-is critical for operating multi-qubit experiments in the lab frame.
-These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
+%% TOOD: intro sentence
+%% TODO: comment on unscented when it is redone
 The sampling and unscented sampling methods
@@
 gate error $\sim 3.5 \cdot 10^{-5}$.
-
-The derivative methods also converge on qualitatively similar solutions that
+The derivative methods converge on qualitatively similar solutions that
 use fast traingle pulses at the boundaries and balance time
 on either side of the flux-frustration point symmetrically at low amplitudes.
-They achieve super-linear scaling in their gate error as
-a function of the gate duration which approaches zero at the larmor
-period $1 / f_{q}$. We observe that on a range of problems with small parameter detunings,
-the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
-derivative method. We believe this is due to the low contribution of second-order
+Both methods achieve a super-linear scaling in their gate error as
+a function of the gate duration. The gate error for the 1\textsuperscript{st}-order
+derivative method approaches zero at the larmor period $1 / f_{q}$.
+We believe the 1\textsuperscript{st}-order method outperforms the \textsuperscript{nd}-order
+method due to the low contribution of second-order
 terms to the gate error in this deviation regime, see Table 2.
 
+%% TODO: comment on the fact that we can reduce the number of sample states
 %% TODO: how do I tie this in?

### 2020-11-03 paper/main.tex
updates to paper

@@
   %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
+  The ability to engineer high fidelity gates on quantum processors in the presence of
   systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$.
+  Quantum optimal control (QOC) techniques have proven effective in experimentally
+  realizing high fidelity gates, but they require exquisite calibration to be performant.
+  We apply robust trajectory optimization techniques to suppress gate errors arising from system
+  model deviations and noise.
   We propose a method that takes advantage of deviant parameter derivative information while maintaining
   computational efficiency by utilizing mixed-mode differentation.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm. We propose a computationally efficient metric
+  Additionally, completely modeling decoherence effects due to longitudinal relaxation requires
+  integrating the Lindblad master equation, which is computationally expensive.
+  We propose a computationally efficient metric
   and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
-  We apply these techniques to a fluxonium qubit with realisitc experimental constraints
-  and noise values. We are able to reduce gate errors by orders of magnitudes in the presence
-  of deviant parameters, achieving super-linear scaling for gate robustness as a function of the gate duration.
-  We use the same techniques to mitigate coherent errors due to time-varying magnetic flux noise
-  by two orders of magnitude. Additionally, we mitigate the gate error arising from longitudinal
-  relaxation by a factor of $5$. Further, we find that the numerical techniques we employ
-  are able to perform phase gates in arbitrary times, which is critical for operating multi-qubit
-  experiments in the lab frame.
+  We demonstrate these techniques numerically on a fluxonium qubit with realisitc experimental parameters and constraints,
+  achieving orders of magnitude gate error reductions from our baseline gate set.
+  %% TODO: maybe add arbitrary phase bit
 \end{abstract}
@@
 %% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
+Quantum optimal control (QOC) techniques are a class of optimization
+algorithms for accurately and efficiently manipulating quantum systems.
+Early techniques were proposed for nuclear magnetic resonance experiments
 \cite{khaneja2005optimal}, and applications now include superconducting
 circuits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
+  huang2020engineering, leng2019robust, leung2017speedup, xu2020nonadiabatic},
 neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
 diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}. QOC techniques are used to
-control a system such that a set of optimization objectives are maximally fulfilled.
-For quantum computation, relevant objectives include achieving high fidelity
+\cite{sorensen2018quantum}. QOC techniques aim to control the system
+such that a set of objectives are optimally satisfied.
+For quantum computation relevant objectives include achieving high fidelity
 gates and adhering to experimental constraints.
@@
 parameters, particular to the quantum system, that govern its evolution.
-It is desirable to make the optimization aware of experimental errors
-such as noise, system model errors, and control errors, to
-improve experimental performance on the optimization objectives.
-The field of robust control, deriving from classical control theory, has developed
-techniques to account for these errors in optimization (citations appreciated).
+Experimental errors may cause the system evolution to deviate from that predicted in
+optimization, hampering performance.
+The field of classical control theory has developed robust control techniques
+to encode experimental errors in optimization objectives, improving
+performance on the experiment (citations appreciated).
 In this work we employ robust control techniques to mitigate
@@
 
-Numerical techniques for QOC have been remarkably successful in
-state preparation, designing high fidelity gates, and determining
-optimized pulse sequences (citations needed). Many popular numerical techniques
-are iterative, indirect shooting
-methods \cite{leung2017speedup, goer2019krotov, abdelhafez2019gradient,
-  machnes2015gradient}.
-An initial guess for the control parameters is used to forward-simulate
-the quantum state trajectory under the time-dependent Schroedinger equation. Then, gradients
-with respect to the controls are computed for an objective evaluated on
-the evolved quantum state. The gradients are used to update
-the control parameters and the procedure is repeated iteratively.
-A convenient way to enforce constraints in this framework is
-to employ control filtering or project gradients onto the constraint manifold.
-In this work we perform optimization using ALTRO, which combines
-an iterative indirect shooting method with the augmented Lagrangian
-method \cite{howell2019altro}. The augmented Lagrangian method
-allows us to enforce arbitrary, simultaneous constraints and achieve
-fast convergence without being restricted to the constraint
-manifold.
+Analytic and numerical techniques for QOC have been remarkably successful in
+designing high fidelity gates (citations needed). In addition to the high
+simulated fidelities which can be obtained with these techniques, it is desirable for the
+resultant quantum state trajectory to be insensitive to experimental errors
+such as parameter drift, noise, and finite control resolution. Controls that produce
+quantum state trajectories that are insensitive to these effects are said to be robust.
+In line with this objective, recent work has sought to mitigate systematic errors
+and decoherence present in the experiment.
+Considering the dynamical and geometric phases of quantum state
+trajectories has led to analytic methods for achieving
+robustness to systematic errors and mitigating pure dephasing
+\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
+Floquet techniques have been experimentally demonstrated to simultaneously mitigate
+decoherence due to longitudinal relaxation and pure dephasing
+\cite{huang2020engineering, mundada2020floquet}.
+Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
+by modeling master equations \cite{rembold2020introduction} and employing
+Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
+Additionally, efforts have been made to incorporate experimental feedback
+into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
+and pulse design are under active development \cite{wittler2020integrated}.
 
-QOC techniques have been effective in mitigating the effects
-of systematic errors. Typically, the model used in optimization
-differs from the experimental system due to parameter drift, the inability
-to realize control parameters exactly, and noise.
-These systematic errors cause the trajectory of the quantum system
-in the experiment to deviate from that predicted by the optimization.
-A set of controls is said to be robust if the experimental state trajectory
-it produces is insensitive to systematic errors.
-Considering the dynamical and geometric phases of quantum state
-evolution has led to analytic methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
-We mitigate systematic errors numerically by employing
-open loop optimization. We draw on robust trajectory optimization
-techniques to encode systematic errors in optimization objectives.
-We consider three
-techniques. The first is the sampling method, which has been applied
+In this work we employ state of the art trajectory optimization techniques
+to mitigate systematic errors and decoherence. We frame the QOC problem as an
+open loop optimization problem, optimizing the controls to make the
+quantum state trajectory robust to systematic errors and decoherence.
+We consider three robust trajectory optimization techniques.
+The first is the sampling method, which has been applied
 previously in the context of QOC
@@
 \cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-We propose a third method, which we call the derivative method.
-This method takes advantage of derivative information of systematic
-errors with respect to the state evolution. We make this method efficient
-by employing mixed-mode differentiation.
-
-Further, many analytic and numerical techniques have been developed to mitigate decoherence.
-Floquet techniques have been experimentally demonstrated to simultaneously mitigate
-decoherence due to longitudinal relaxation and pure dephasing
-\cite{huang2020engineering, mundada2020floquet}.
-Numerical techniques have mitigated decoherence by evolving the
-system using master equations \cite{rembold2020introduction} and employing
-Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
-In this work we take advantage of the known dependence of the
-quantum system's controls on longitudinal relaxation. We encode
+We propose a third method, which uses derivative information
+of parameter deviations with respect to the quantum state evolution.
+We make this method efficient by employing mixed-mode differentiation.
+Additionally, we take advantage of the known dependence of the
+system's controls on longitudinal relaxation. We encode
 the decoherence in an efficient optimization objective that does
 not pay the increased computational cost of integrating a master equation.
+We perform these optimization using the ALTRO framework \cite{howell2019altro}.
+Many popular numerical techniques
+for QOC use gradient-based control update procedures, which are readily amenable to control filtering
+or constraint manifold projection techniques to enforce constraints \cite{leung2017speedup,
+  goer2019krotov, abdelhafez2019gradient, machnes2015gradient}.
+The ALTRO method we employ combines
+an iterative indirect shooting method with the augmented Lagrangian
+method. The augmented Lagrangian method
+allows us to enforce arbitrary, simultaneous constraints and achieve
+fast convergence without being restricted to the constraint
+manifold.
 
-We compare our techniques using the heavy fluxonium qubit presented in
-in \cite{zhang2020universal}. This quantum system is an exciting
+We compare our techniques using the heavy fluxonium qubit and assocaited analytic gates
+presented in \cite{zhang2020universal}. This quantum system is an exciting
 platform for quantum computation due to its long longitudinal relaxation
@@
 over the baseline gate set. Then we present three methods for achieving
-robustness to systematic errors. We find that we are able to
-decrease the gate error arising from systematic errors super-linearly
-in the gate duration, achieving orders of magntidue reduction in gate errors.
+robustness to static parameter deviations. We find that we are able to
+decrease the gate error arising from parameter deviations super-linearly
+in the gate duration, achieving orders of magntidue reductions in single gate errors.
 We also find that the numerical techniques we employ are able to produce
@@
 ensure its smoothness, and in doing so mitigate AWG ringing.
-
 The optimization is performed over the second derivative of the flux amplitude
@@
 to the control lines.
-
 Dissipation to the thermal bath via longitudinal
@@
 to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
+by tuning the flux amplitude and employ robust control techniques to mitigate
 pure dephasing.
@@
 The first is the sampling method, which has
-been preposed previously in the context of QOC
+been proposed previously in the context of QOC
 \cite{rembold2020introduction, reinhold2019controlling, carvalho2020error}. In the
@@
 \section{Conclusion}
-We have proposed some schemes and they work well.
+In conclusion, we have demonstrated techniques for achieving robustness to systematic
+errors and mitigating decoherence on a quantum system using state-of-the-art trajectory
+optimization. We have proposed a scheme for mitigating longitudinal relaxation with time-optimal
+control and an efficient optimization metric that comes at a constant computational cost as
+opposed to integrating a master equation which scales quadratically with
+the dimension of the Hilbert space.
+We have proposed the derivative method for robust control which achieves
+super-linear gate error reductions in the gate duration for the problem we studied here.
+We have shown that robust control techniques can be used to mitigate decoherence due
+to 1/$f$ flux noise, a dominant source of coherent errors for flux controlled qubits.
+The numerical techniques we have studied can be used to perform phase gates in arbitrary times,
+which will be critical for synchronizing multi-qubit systems. Our
+techniques significantly improve the state-of-the-art for controlling quantum systems subject
+systematic errors and decoherence, and will be used in efforts to lower gate errors
+for fault-tolerant quantum computing and NISQ applications.
 

### 2020-11-03 paper/0abstract.tex
separated sections by file

--- 2020-10-26 paper/main.tex
@@
-\documentclass[
-  amsfonts,
-  amsmath,
-  tbtags,
-  amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
-  twocolumn,
-  superscriptaddress,
-]{revtex4-2}
-
-\input{preamble}
-
-\begin{document}
-
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
 %% ABSTRACT
@@
   %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
+  The ability to engineer high fidelity gates on quantum processors in the presence of
   systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
-  takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by transforming
-  higher-order differential equations into a set of coupled first-order equations.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm.
-  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
-  for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimization and time-optimal control.
+  Quantum optimal control (QOC) techniques have proven effective in experimentally
+  realizing high fidelity gates, but they require exquisite calibration to be performant.
+  We apply robust trajectory optimization techniques to suppress gate errors arising from system
+  model deviations and noise.
+  We propose a method that takes advantage of deviant parameter derivative information while maintaining
+  computational efficiency by utilizing mixed-mode differentation.
+  Additionally, completely modeling decoherence effects due to longitudinal relaxation requires
+  integrating the Lindblad master equation, which is computationally expensive.
+  We propose a computationally efficient metric
+  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
+  We demonstrate these techniques numerically on a fluxonium qubit with realisitc
+  experimental parameters and constraints,
+  achieving orders of magnitude gate error reductions from our baseline gate set.
+  %% TODO: maybe add arbitrary phase bit
 \end{abstract}
-
-\maketitle
-
-%% S1
-\section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-qubits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}.
-In quantum control theory, as in classical control theory,
-optimization is performed
-to minimize an objective function. Relevant for the field
-of quantum computation is the task of achieving high fidelity quantum gates.
-In this application, the objective function includes contributions from the
-gate error associated with the state transfer of the system as well as contributions
-due to experimental constraints, such as the control amplitude.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-
-Many analytic and numerical techniques have been developed to control
-quantum systems. Analytic methods rely on solving the time-dependent
-Schroedinger equation analytically or deriving equations of motion
-from a suitable Lagrangian \cite{zhang2020universal, huang2020engineering, han2020experimental,
-  xu2020nonadiabatic, carlini2005quantum}. Analytic methods
-have been presented to address decoherence in quantum systems.
-Considering the dynamical and geometric phases of state
-evolution has led to methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
-Additionally, suitable choices of bases have been proposed that
-allow the experimentalist to infer tradeoffs in longitudinal relaxation
-and pure dephasing coherence times \cite{huang2020engineering}.
-To date, however, no analytical method has been presented
-to construct a gate while satisfying arbitrary experimental
-constraints. Furthermore, most analytic methods have only
-been demonstrated in the two-level approximation and become
-cumbersome to solve for larger Hilbert space sizes.
-
-Numerical quantum control techniques have developed in parallel
-with analytic frameworks.
-Most numerical techniques proposed for QOC are indirect shooting
-methods where an initial guess for the control trajectory is forward-simulated,
-e.g. under the time-dependent Schroedinger equation,
-and first-order gradients are computed for a functional evaluated
-on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
-  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
-Most techniques formulate the
-problem as unconstrained or use
-projective gradient methods to enforce constraints
-\cite{machnes2015gradient}.
-The latter approach relies on constructing
-bases for constraint spaces--which is not always feasible--and
-may hinder convergence.
-Methods for mitigating decoherence using numerical
-techniques have employed master equation dynamics (citation needed),
-Monte Carlo style trajectories \cite{abdelhafez2019gradient},
-and proposals have been made to simulate multiple trajectories
-\cite{reinhold2019controlling, rembold2020introduction}.
-
-In this work, we employ state of the art
-trajectory optimization methods
-to formulate the quantum
-optimal control problem as a constrained optimization problem, allowing
-us to enforce arbitrary experimental constraints.
-Additionally,
-we introduce methods for constructing gates that mitigate longitudinal
-relaxation and are robust to experimental errors, and therefore
-pure dephasing, the dominant barriers to experimentally
-realizing practical quantum computing.
-
-As an example, we study the quantum optimal control problem on
-the fluxonium qubit presented in \cite{zhang2020universal}.
-%%TODO: Fluxonium qubit is exciting and computationally feasible,
-%%but our methods extend to arbitrary platforms.
-We describe experimentally realistic constraints and map them to
-the trajectory optimization framework. Next, we
-outline a method for making the optimization longitudinal
-relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set.
-We present two methods for achieving robustness to pure dephasing
-and compare them to existing dynamic decoupling methods. We find that
-the methods we employ are able to produce high fidelity gates when
-subject to parameter spreads of $1\%$.
-
-%% S2
-\section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
-\label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
-In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
-algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
-
-(AL-iLQR Problem Statement) The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
-\begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-\end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
-
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-  \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
-  \end{aligned}
-\end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
-\end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
-
-%% S3
-\section{QOC on the Fluxonium}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
-\end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
-
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
-The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
-
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
-
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is computed after
-    each gate appliaction. The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
-  }
-\end{figure*}
-
-%% S4
-\section{Longitudinal Relaxation Awareness}
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms,
-and it is advantageous to maximize it by tuning the controls.
-In the fluxonium, the longitudinal relaxation time varies with
-the amplitude $a$. To completely simulate
-the effect of longitudinal relaxation on the gate error requires
-propagating density matrices of size $n \times n$ under Master equation
-dynamics rather than state vectors of size $n$ under the TDSE dynamics.
-We avoid this increase in computational complexity by
-penalizing the probability that the qubit decays as a result of
-longitudinal relaxation. This modification only
-increases the size of the augmented state vector by a single real number
-and allows us to effectively mitigate the error arising from longitudinal
-relaxation. Additionally, for a constant $T_{1}$ time, a shorter gate duration
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
-the one we study here.
-
-The longitudinal relaxation probability is given by
-\begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
-\end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-Calculating $T_{1}$ directly from theoretical
-considerations requires many high-rank
-eigendecompositions, which
-is computationally expensive. Additionally,
-$T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation.
-
-We allow the optimizer to tune the gate duration by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
-integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
-
-The $T_{1}$ optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate probability of longitudinal relaxation by a factor of $5$ over
-their analytic counterparts and achieve a commensurate reduction in their single gate error,
-see Appendix. The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have put on the
-device, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
-    points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency at multiples of the analytic gate duration for the analytic
-    and 1\textsuperscript{st}-order derivative methods.
-  }
-\end{figure*}
-
-%% S5
-\section{Robustness to Static Parameter Deviations}
-We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the model we use in optimization,
-the Hamiltonian, leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
-we study mitigating errors arising from the finite measurement
-precision of the qubit frequency $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$.
-We consider three flavors of robust control
-techniques. The first is the sampling method, which has
-been preposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling}. In the
-sampling method, multiple trajectories are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
-transformation to accurately propagate a distribution
-representing the evolving state as a result of a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. This scheme
-uses derivative information about how the state
-evolution is affected by the deviant parameter to
-mitigate trajectory deviations.
-
-In the sampling method we propagate additional sample states
-in the augmented state vector that follow deviant dynamics
-and penalize their differences from the target state.
-The states evolve under the deviant
-Hamiltonian where the parameter of interest is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The infidelities of the samples and the target
-state are penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
-resulting in eight additional states added to the augmented state vector.
-The four initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-This results in $O(n^{2})$ states added to the augmented
-state vector where $n$ is both the dimension of the Hilbert space
-and the size of the state vector.
-
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}\hspace{0.025\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is computed
-    after each gate application.
-  }
-\end{figure*}
-
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of $\ket{\psi}$. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states
-where $d$ is the number of deviant parameters. Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. This results
-in $O(n^{3})$ states added to the augmented state vector.
-The costs at each knot point are the infidelities between the samples
-and the target state, as in the sampling method.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
-derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For $d$ sensitivity parameters and $N$ knot points
-this requires $O(d N^2)$ matrix multiplications. Instead,
-we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting
-in $O(d N)$ matrix multiplications. For the piece-wise constant control
-scheme we employ here, $N$ is the number of control parameters. Typically
-$d \ll N$ so this is a good tradeoff. As an example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
-
-To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$. We compare the numerical methods
-to an analytically derived $Z/2$ gate which corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. It is linearly sensitive to qubit frequency detunings,
-the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
-qubit frequency detuning, it achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-sufficient for quantum error correction.
-
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
-The ability to perform $Z$ rotations in arbitrary times
-is critical for operating multi-qubit experiments in the lab frame.
-These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-
-The derivative methods also converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-They achieve super-linear scaling in their gate error as
-a function of the gate duration which approaches zero at the larmor
-period $1 / f_{q}$. We observe that on a range of problems with small parameter detunings,
-the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
-derivative method. We believe this is due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
-
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
-
-%% S6
-\section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
-take advantage of the low frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
-
-We compare the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
-flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-
-We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
-gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
-
-%% S7
-\section{Conclusion}
-We have proposed some schemes and they work well.
-
-%% ACK
-\begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
-\end{acknowledgments}
-
-\appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
-\section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
-\begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
-For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
-and $L_{\uparrow} = \sigma^{+}/2$,
-$L_{\downarrow} = \sigma^{-}/2$
-are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
-$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
-$L_{2} = (I - \sigma_{z})/2$.
-
-%% T1
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & Analytic & QOC & & Analytic & QOC & \\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
-    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% AD
-\section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
-
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
-
-\bibliography{refs}
-
-\end{document}

### 2020-11-03 paper/1introduction.tex
separated sections by file

--- 2020-11-03 paper/main.tex
@@
-\documentclass[
-  amsfonts,
-  amsmath,
-  tbtags,
-  amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
-  twocolumn,
-  superscriptaddress,
-]{revtex4-2}
-
-\input{preamble}
-
-\begin{document}
-
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
-%% ABSTRACT
-\begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
-  The ability to engineer high fidelity gates on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in experimentally
-  realizing high fidelity gates, but they require exquisite calibration to be performant.
-  We apply robust trajectory optimization techniques to suppress gate errors arising from system
-  model deviations and noise.
-  We propose a method that takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by utilizing mixed-mode differentation.
-  Additionally, completely modeling decoherence effects due to longitudinal relaxation requires
-  integrating the Lindblad master equation, which is computationally expensive.
-  We propose a computationally efficient metric
-  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
-  We demonstrate these techniques numerically on a fluxonium qubit with realisitc experimental parameters and constraints,
-  achieving orders of magnitude gate error reductions from our baseline gate set.
-  %% TODO: maybe add arbitrary phase bit
-\end{abstract}
-
-\maketitle
-
 %% S1
@@
 to encode experimental errors in optimization objectives, improving
-performance on the experiment (citations appreciated).
+experimental performance (citations appreciated).
 In this work we employ robust control techniques to mitigate
@@
 We propose a third method, which uses derivative information
-of parameter deviations with respect to the quantum state evolution.
+of parameter deviations with respect to the quantum state trajectory.
 We make this method efficient by employing mixed-mode differentiation.
@@
 not pay the increased computational cost of integrating a master equation.
-We perform these optimization using the ALTRO framework \cite{howell2019altro}.
+We perform these optimizations using the ALTRO method \cite{howell2019altro}.
 Many popular numerical techniques
@@
 
-We compare our techniques using the heavy fluxonium qubit and assocaited analytic gates
-presented in \cite{zhang2020universal}. This quantum system is an exciting
+We compare the performance of these numerical techniques
+on the heavy fluxonium qubit and against the assocaited analytic techniques
+presented in \cite{zhang2020universal}. The fluxonium is an exciting
 platform for quantum computation due to its long longitudinal relaxation
@@
 outline a method for making the optimization aware of longitudinal
-relaxation. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set. Then we present three methods for achieving
+relaxation. We achieve a factor of 5 decrease in single gate errors due to
+longitudinal relaxation over the analytic gate set. Next we outline three methods for achieving
 robustness to static parameter deviations. We find that we are able to
-decrease the gate error arising from parameter deviations super-linearly
+decrease the gate error arising from static parameter deviations super-linearly
 in the gate duration, achieving orders of magntidue reductions in single gate errors.
@@
 for noisy, intermediate-scale quantum (NISQ) applications.
-
-%% S2
-\section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
-\label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
-In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
-algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
-
-(AL-iLQR Problem Statement) The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
-\begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-\end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
-
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-  \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
-  \end{aligned}
-\end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
-\end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
-
-%% S3
-\section{QOC on the Fluxonium}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
-\end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
-
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
-The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
-
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-by tuning the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
-
-%% TODO: talk about quadratic cost vs. infidelity
-
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is the cumulative
-    gate error after each gate application.
-    The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
-  }
-\end{figure*}
-
-%% S4
-\section{Longitudinal Relaxation Awareness}
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms.
-It is advantageous to tune the controls to extend the longitudinal
-relaxation time.
-To calculate the gate error due to longitudinal relaxation
-requires propagating density matrices of size $n \times n$ under Master equation
-dynamics, rather than state vectors of size $n$ under the TDSE dynamics.
-We avoid this increase in computational complexity by
-penalizing the integrated rate of longitudinal relaxation,
-i.e. the probability of longitudinal relaxation.
-Using this probability as proxy for the gate error incurred
-is reasonable because losses due to longitudinal relaxation
-increase monotonically in time.
-This technique can be extended to
-error channels which share the monotonically increasing property.
-Additionally, for a constant $T_{1}$ time, a shorter gate duration
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
-the one we study here.
-
-The longitudinal relaxation probability is given by
-\begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
-\end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-It is also possible to use a spline fit to theoretically obtained data.
-However, $T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data
-increases the fridge truth of the simulation.
-
-We allow the optimizer to tune the gate duration by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
-integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
-
-We compare the numerical method we have developed to the analytic gates
-on the task of achieving low gate errors in the presence of longitudinal relaxation
-for the $Z/2$, $Y/2$, and $X/2$ gates.
-The numerically optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate errors by a factor of $5$ over
-their analytic counterparts which is commensurate to their
-probability of longitudinal relaxation reductions, see Appendix.
-The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have imposed
-on the gates, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, sampling, unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
-    data points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency. The solutions for the analytic and
-    1\textsuperscript{st}-order derivative methods are shown at multiples
-    of the analytic gate time. The performance of the two methods is
-    indistinguishable at the analytic gate time $18$ns.
-  }
-\end{figure*}
-
-%% S5
-\section{Robustness to Static Parameter Deviations}
-We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the Hamiltonian we use in optimization,
-leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
-we mitigate errors arising from the drift and finite measurement
-precision of the qubit frequency $\tilde{f_{q}} = f_{q} \pm \sigma_{f_{q}}$.
-We consider three robust control techniques.
-The first is the sampling method, which has
-been proposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling, carvalho2020error}. In the
-sampling method, multiple states are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
-transformation to accurately propagate a distribution
-representing the uncertainty in an evolving state
-due to a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. Derivative
-information encoding the sensitivity of the
-state trajectory with respect to the deviant parameter
-is used to penalize state trajectory deviations.
-
-In the sampling method, sample states evolve under a
-Hamiltonian where a parameter is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant values $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The gate error of each sample state is 
-penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-Each initial state is used to initialize two
-sample states $\ket{\psi^{\pm}}$.
-The initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-For this method the number of states in the augmented state vector
-scales as $O(dn^{2})$ because each of $n^{2}$ initial
-states is represented by $2d$ samples. $n$ is both the dimension of the Hilbert space
-and the size of the state vector. $d$ is the number of deviant parameters.
-
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}\hspace{0.025\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    sampling, unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is the cumulative
-    gate error after each gate application.
-  }
-\end{figure*}
-
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of the initial state. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states.
-Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. Hence,
-the number of states in the augmented state vector scales as $O(n^{3} + dn^{2})$.
-The gate error of each sample state is penalized as in the sampling method.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
-derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For a single state derivative and $N$ knot points this
-requires $O(N^2)$ matrix multiplications.
-Instead, we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting in
-$O(N)$ matrix multiplications. For example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
-For this method $m$ state derivatives are associated with each
-initial state from the operator basis.
-So, the number of states in the augmented state vector scales as
-$O(dmn^{2})$.
-
-To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$.
-
-We compare the numerical methods
-to an analytically derived $Z/2$ gate. This gate corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. Its erroneous rotation angle $2 \pi t_{Z/2} \delta f_{q}$ is linearly sensitive to
-the qubit frequency detuning, resulting in a gate error that is quadratically sensitive
-to the qubit frequency detutning.
-At a one-percent
-qubit frequency detuning the analytic gate achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-which is sufficient for quantum error correction, see Figure 2b.
-Although the analytic $Z/2$ gate performs well, it works
-only at the gate time $t_{Z/2}$. The ability to perform $Z$ rotations in arbitrary times is critical
-for operating multi-qubit experiments in the lab frame.
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times,
-see Figure 2b. These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-
-%% TOOD: intro sentence
-%% TODO: comment on unscented when it is redone
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-The derivative methods converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-Both methods achieve a super-linear scaling in their gate error as
-a function of the gate duration. The gate error for the 1\textsuperscript{st}-order
-derivative method approaches zero at the larmor period $1 / f_{q}$.
-We believe the 1\textsuperscript{st}-order method outperforms the \textsuperscript{nd}-order
-method due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
-
-%% TODO: comment on the fact that we can reduce the number of sample states
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
-
-%% S6
-\section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
-take advantage of the low frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
-
-We compare the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
-flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-
-We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
-gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
-
-%% S7
-\section{Conclusion}
-In conclusion, we have demonstrated techniques for achieving robustness to systematic
-errors and mitigating decoherence on a quantum system using state-of-the-art trajectory
-optimization. We have proposed a scheme for mitigating longitudinal relaxation with time-optimal
-control and an efficient optimization metric that comes at a constant computational cost as
-opposed to integrating a master equation which scales quadratically with
-the dimension of the Hilbert space.
-We have proposed the derivative method for robust control which achieves
-super-linear gate error reductions in the gate duration for the problem we studied here.
-We have shown that robust control techniques can be used to mitigate decoherence due
-to 1/$f$ flux noise, a dominant source of coherent errors for flux controlled qubits.
-The numerical techniques we have studied can be used to perform phase gates in arbitrary times,
-which will be critical for synchronizing multi-qubit systems. Our
-techniques significantly improve the state-of-the-art for controlling quantum systems subject
-systematic errors and decoherence, and will be used in efforts to lower gate errors
-for fault-tolerant quantum computing and NISQ applications.
-
-%% ACK
-\begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
-\end{acknowledgments}
-
-\appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
-\section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
-\begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
-For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
-and $L_{\uparrow} = \sigma^{+}/2$,
-$L_{\downarrow} = \sigma^{-}/2$
-are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
-$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
-$L_{2} = (I - \sigma_{z})/2$.
-
-%% T1
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & Analytic & QOC & & Analytic & QOC & \\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
-    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% AD
-\section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
-
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
-
-\bibliography{refs}
-
-\end{document}

### 2020-11-03 paper/2background.tex
separated sections by file

--- 2020-10-26 paper/main.tex
@@
-\documentclass[
-  amsfonts,
-  amsmath,
-  tbtags,
-  amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
-  twocolumn,
-  superscriptaddress,
-]{revtex4-2}
-
-\input{preamble}
-
-\begin{document}
-
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
-%% ABSTRACT
-\begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
-  takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by transforming
-  higher-order differential equations into a set of coupled first-order equations.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm.
-  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
-  for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimization and time-optimal control.
-\end{abstract}
-
-\maketitle
-
-%% S1
-\section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-qubits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}.
-In quantum control theory, as in classical control theory,
-optimization is performed
-to minimize an objective function. Relevant for the field
-of quantum computation is the task of achieving high fidelity quantum gates.
-In this application, the objective function includes contributions from the
-gate error associated with the state transfer of the system as well as contributions
-due to experimental constraints, such as the control amplitude.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-
-Many analytic and numerical techniques have been developed to control
-quantum systems. Analytic methods rely on solving the time-dependent
-Schroedinger equation analytically or deriving equations of motion
-from a suitable Lagrangian \cite{zhang2020universal, huang2020engineering, han2020experimental,
-  xu2020nonadiabatic, carlini2005quantum}. Analytic methods
-have been presented to address decoherence in quantum systems.
-Considering the dynamical and geometric phases of state
-evolution has led to methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
-Additionally, suitable choices of bases have been proposed that
-allow the experimentalist to infer tradeoffs in longitudinal relaxation
-and pure dephasing coherence times \cite{huang2020engineering}.
-To date, however, no analytical method has been presented
-to construct a gate while satisfying arbitrary experimental
-constraints. Furthermore, most analytic methods have only
-been demonstrated in the two-level approximation and become
-cumbersome to solve for larger Hilbert space sizes.
-
-Numerical quantum control techniques have developed in parallel
-with analytic frameworks.
-Most numerical techniques proposed for QOC are indirect shooting
-methods where an initial guess for the control trajectory is forward-simulated,
-e.g. under the time-dependent Schroedinger equation,
-and first-order gradients are computed for a functional evaluated
-on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
-  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
-Most techniques formulate the
-problem as unconstrained or use
-projective gradient methods to enforce constraints
-\cite{machnes2015gradient}.
-The latter approach relies on constructing
-bases for constraint spaces--which is not always feasible--and
-may hinder convergence.
-Methods for mitigating decoherence using numerical
-techniques have employed master equation dynamics (citation needed),
-Monte Carlo style trajectories \cite{abdelhafez2019gradient},
-and proposals have been made to simulate multiple trajectories
-\cite{reinhold2019controlling, rembold2020introduction}.
-
-In this work, we employ state of the art
-trajectory optimization methods
-to formulate the quantum
-optimal control problem as a constrained optimization problem, allowing
-us to enforce arbitrary experimental constraints.
-Additionally,
-we introduce methods for constructing gates that mitigate longitudinal
-relaxation and are robust to experimental errors, and therefore
-pure dephasing, the dominant barriers to experimentally
-realizing practical quantum computing.
-
-As an example, we study the quantum optimal control problem on
-the fluxonium qubit presented in \cite{zhang2020universal}.
-%%TODO: Fluxonium qubit is exciting and computationally feasible,
-%%but our methods extend to arbitrary platforms.
-We describe experimentally realistic constraints and map them to
-the trajectory optimization framework. Next, we
-outline a method for making the optimization longitudinal
-relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set.
-We present two methods for achieving robustness to pure dephasing
-and compare them to existing dynamic decoupling methods. We find that
-the methods we employ are able to produce high fidelity gates when
-subject to parameter spreads of $1\%$.
-
 %% S2
@@
 The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
+act on the system. Numerical quantum optimal control techniques make
 the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
+knot points (time steps). Typical explicit integration techniques for the TDSE include
+exponential integrators and integrators of the Runge-Kutta type
 
@@
 updates, see \cite{howell2019altro}.
-
-%% S3
-\section{QOC on the Fluxonium}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
-\end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
-
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
-The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
-
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
-
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is computed after
-    each gate appliaction. The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
-  }
-\end{figure*}
-
-%% S4
-\section{Longitudinal Relaxation Awareness}
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms,
-and it is advantageous to maximize it by tuning the controls.
-In the fluxonium, the longitudinal relaxation time varies with
-the amplitude $a$. To completely simulate
-the effect of longitudinal relaxation on the gate error requires
-propagating density matrices of size $n \times n$ under Master equation
-dynamics rather than state vectors of size $n$ under the TDSE dynamics.
-We avoid this increase in computational complexity by
-penalizing the probability that the qubit decays as a result of
-longitudinal relaxation. This modification only
-increases the size of the augmented state vector by a single real number
-and allows us to effectively mitigate the error arising from longitudinal
-relaxation. Additionally, for a constant $T_{1}$ time, a shorter gate duration
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
-the one we study here.
-
-The longitudinal relaxation probability is given by
-\begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
-\end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-Calculating $T_{1}$ directly from theoretical
-considerations requires many high-rank
-eigendecompositions, which
-is computationally expensive. Additionally,
-$T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation.
-
-We allow the optimizer to tune the gate duration by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
-integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
-
-The $T_{1}$ optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate probability of longitudinal relaxation by a factor of $5$ over
-their analytic counterparts and achieve a commensurate reduction in their single gate error,
-see Appendix. The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have put on the
-device, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
-    points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency at multiples of the analytic gate duration for the analytic
-    and 1\textsuperscript{st}-order derivative methods.
-  }
-\end{figure*}
-
-%% S5
-\section{Robustness to Static Parameter Deviations}
-We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the model we use in optimization,
-the Hamiltonian, leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
-we study mitigating errors arising from the finite measurement
-precision of the qubit frequency $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$.
-We consider three flavors of robust control
-techniques. The first is the sampling method, which has
-been preposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling}. In the
-sampling method, multiple trajectories are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
-transformation to accurately propagate a distribution
-representing the evolving state as a result of a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. This scheme
-uses derivative information about how the state
-evolution is affected by the deviant parameter to
-mitigate trajectory deviations.
-
-In the sampling method we propagate additional sample states
-in the augmented state vector that follow deviant dynamics
-and penalize their differences from the target state.
-The states evolve under the deviant
-Hamiltonian where the parameter of interest is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The infidelities of the samples and the target
-state are penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
-resulting in eight additional states added to the augmented state vector.
-The four initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-This results in $O(n^{2})$ states added to the augmented
-state vector where $n$ is both the dimension of the Hilbert space
-and the size of the state vector.
-
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}\hspace{0.025\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is computed
-    after each gate application.
-  }
-\end{figure*}
-
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of $\ket{\psi}$. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states
-where $d$ is the number of deviant parameters. Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. This results
-in $O(n^{3})$ states added to the augmented state vector.
-The costs at each knot point are the infidelities between the samples
-and the target state, as in the sampling method.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
-derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For $d$ sensitivity parameters and $N$ knot points
-this requires $O(d N^2)$ matrix multiplications. Instead,
-we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting
-in $O(d N)$ matrix multiplications. For the piece-wise constant control
-scheme we employ here, $N$ is the number of control parameters. Typically
-$d \ll N$ so this is a good tradeoff. As an example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
-
-To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$. We compare the numerical methods
-to an analytically derived $Z/2$ gate which corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. It is linearly sensitive to qubit frequency detunings,
-the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
-qubit frequency detuning, it achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-sufficient for quantum error correction.
-
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
-The ability to perform $Z$ rotations in arbitrary times
-is critical for operating multi-qubit experiments in the lab frame.
-These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-
-The derivative methods also converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-They achieve super-linear scaling in their gate error as
-a function of the gate duration which approaches zero at the larmor
-period $1 / f_{q}$. We observe that on a range of problems with small parameter detunings,
-the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
-derivative method. We believe this is due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
-
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
-
-%% S6
-\section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
-take advantage of the low frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
-
-We compare the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
-flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-
-We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
-gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
-
-%% S7
-\section{Conclusion}
-We have proposed some schemes and they work well.
-
-%% ACK
-\begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
-\end{acknowledgments}
-
-\appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
-\section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
-\begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
-For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
-and $L_{\uparrow} = \sigma^{+}/2$,
-$L_{\downarrow} = \sigma^{-}/2$
-are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
-$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
-$L_{2} = (I - \sigma_{z})/2$.
-
-%% T1
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & Analytic & QOC & & Analytic & QOC & \\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
-    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% AD
-\section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
-
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
-
-\bibliography{refs}
-
-\end{document}

### 2020-11-03 paper/3fluxonium.tex
separated sections by file

--- 2020-10-26 paper/main.tex
@@
-\documentclass[
-  amsfonts,
-  amsmath,
-  tbtags,
-  amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
-  twocolumn,
-  superscriptaddress,
-]{revtex4-2}
-
-\input{preamble}
-
-\begin{document}
-
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
-%% ABSTRACT
-\begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
-  takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by transforming
-  higher-order differential equations into a set of coupled first-order equations.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm.
-  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
-  for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimization and time-optimal control.
-\end{abstract}
-
-\maketitle
-
-%% S1
-\section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-qubits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}.
-In quantum control theory, as in classical control theory,
-optimization is performed
-to minimize an objective function. Relevant for the field
-of quantum computation is the task of achieving high fidelity quantum gates.
-In this application, the objective function includes contributions from the
-gate error associated with the state transfer of the system as well as contributions
-due to experimental constraints, such as the control amplitude.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-
-Many analytic and numerical techniques have been developed to control
-quantum systems. Analytic methods rely on solving the time-dependent
-Schroedinger equation analytically or deriving equations of motion
-from a suitable Lagrangian \cite{zhang2020universal, huang2020engineering, han2020experimental,
-  xu2020nonadiabatic, carlini2005quantum}. Analytic methods
-have been presented to address decoherence in quantum systems.
-Considering the dynamical and geometric phases of state
-evolution has led to methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
-Additionally, suitable choices of bases have been proposed that
-allow the experimentalist to infer tradeoffs in longitudinal relaxation
-and pure dephasing coherence times \cite{huang2020engineering}.
-To date, however, no analytical method has been presented
-to construct a gate while satisfying arbitrary experimental
-constraints. Furthermore, most analytic methods have only
-been demonstrated in the two-level approximation and become
-cumbersome to solve for larger Hilbert space sizes.
-
-Numerical quantum control techniques have developed in parallel
-with analytic frameworks.
-Most numerical techniques proposed for QOC are indirect shooting
-methods where an initial guess for the control trajectory is forward-simulated,
-e.g. under the time-dependent Schroedinger equation,
-and first-order gradients are computed for a functional evaluated
-on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
-  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
-Most techniques formulate the
-problem as unconstrained or use
-projective gradient methods to enforce constraints
-\cite{machnes2015gradient}.
-The latter approach relies on constructing
-bases for constraint spaces--which is not always feasible--and
-may hinder convergence.
-Methods for mitigating decoherence using numerical
-techniques have employed master equation dynamics (citation needed),
-Monte Carlo style trajectories \cite{abdelhafez2019gradient},
-and proposals have been made to simulate multiple trajectories
-\cite{reinhold2019controlling, rembold2020introduction}.
-
-In this work, we employ state of the art
-trajectory optimization methods
-to formulate the quantum
-optimal control problem as a constrained optimization problem, allowing
-us to enforce arbitrary experimental constraints.
-Additionally,
-we introduce methods for constructing gates that mitigate longitudinal
-relaxation and are robust to experimental errors, and therefore
-pure dephasing, the dominant barriers to experimentally
-realizing practical quantum computing.
-
-As an example, we study the quantum optimal control problem on
-the fluxonium qubit presented in \cite{zhang2020universal}.
-%%TODO: Fluxonium qubit is exciting and computationally feasible,
-%%but our methods extend to arbitrary platforms.
-We describe experimentally realistic constraints and map them to
-the trajectory optimization framework. Next, we
-outline a method for making the optimization longitudinal
-relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set.
-We present two methods for achieving robustness to pure dephasing
-and compare them to existing dynamic decoupling methods. We find that
-the methods we employ are able to produce high fidelity gates when
-subject to parameter spreads of $1\%$.
-
-%% S2
-\section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
-\label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
-In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
-algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
-
-(AL-iLQR Problem Statement) The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
-\begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-\end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
-
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-  \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
-  \end{aligned}
-\end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
-\end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
-
 %% S3
@@
 to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
+by tuning the flux amplitude and employ robust control techniques to mitigate
 pure dephasing.
 
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is computed after
-    each gate appliaction. The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
-  }
-\end{figure*}
-
-%% S4
-\section{Longitudinal Relaxation Awareness}
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms,
-and it is advantageous to maximize it by tuning the controls.
-In the fluxonium, the longitudinal relaxation time varies with
-the amplitude $a$. To completely simulate
-the effect of longitudinal relaxation on the gate error requires
-propagating density matrices of size $n \times n$ under Master equation
-dynamics rather than state vectors of size $n$ under the TDSE dynamics.
-We avoid this increase in computational complexity by
-penalizing the probability that the qubit decays as a result of
-longitudinal relaxation. This modification only
-increases the size of the augmented state vector by a single real number
-and allows us to effectively mitigate the error arising from longitudinal
-relaxation. Additionally, for a constant $T_{1}$ time, a shorter gate duration
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
-the one we study here.
-
-The longitudinal relaxation probability is given by
-\begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
-\end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-Calculating $T_{1}$ directly from theoretical
-considerations requires many high-rank
-eigendecompositions, which
-is computationally expensive. Additionally,
-$T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation.
-
-We allow the optimizer to tune the gate duration by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
-integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
-
-The $T_{1}$ optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate probability of longitudinal relaxation by a factor of $5$ over
-their analytic counterparts and achieve a commensurate reduction in their single gate error,
-see Appendix. The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have put on the
-device, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
-    points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency at multiples of the analytic gate duration for the analytic
-    and 1\textsuperscript{st}-order derivative methods.
-  }
-\end{figure*}
-
-%% S5
-\section{Robustness to Static Parameter Deviations}
-We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the model we use in optimization,
-the Hamiltonian, leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
-we study mitigating errors arising from the finite measurement
-precision of the qubit frequency $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$.
-We consider three flavors of robust control
-techniques. The first is the sampling method, which has
-been preposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling}. In the
-sampling method, multiple trajectories are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
-transformation to accurately propagate a distribution
-representing the evolving state as a result of a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. This scheme
-uses derivative information about how the state
-evolution is affected by the deviant parameter to
-mitigate trajectory deviations.
-
-In the sampling method we propagate additional sample states
-in the augmented state vector that follow deviant dynamics
-and penalize their differences from the target state.
-The states evolve under the deviant
-Hamiltonian where the parameter of interest is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The infidelities of the samples and the target
-state are penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
-resulting in eight additional states added to the augmented state vector.
-The four initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-This results in $O(n^{2})$ states added to the augmented
-state vector where $n$ is both the dimension of the Hilbert space
-and the size of the state vector.
-
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}\hspace{0.025\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is computed
-    after each gate application.
-  }
-\end{figure*}
-
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of $\ket{\psi}$. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states
-where $d$ is the number of deviant parameters. Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. This results
-in $O(n^{3})$ states added to the augmented state vector.
-The costs at each knot point are the infidelities between the samples
-and the target state, as in the sampling method.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
-derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For $d$ sensitivity parameters and $N$ knot points
-this requires $O(d N^2)$ matrix multiplications. Instead,
-we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting
-in $O(d N)$ matrix multiplications. For the piece-wise constant control
-scheme we employ here, $N$ is the number of control parameters. Typically
-$d \ll N$ so this is a good tradeoff. As an example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
-
-To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$. We compare the numerical methods
-to an analytically derived $Z/2$ gate which corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. It is linearly sensitive to qubit frequency detunings,
-the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
-qubit frequency detuning, it achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-sufficient for quantum error correction.
-
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
-The ability to perform $Z$ rotations in arbitrary times
-is critical for operating multi-qubit experiments in the lab frame.
-These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-
-The derivative methods also converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-They achieve super-linear scaling in their gate error as
-a function of the gate duration which approaches zero at the larmor
-period $1 / f_{q}$. We observe that on a range of problems with small parameter detunings,
-the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
-derivative method. We believe this is due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
-
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
-
-%% S6
-\section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
-take advantage of the low frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
-
-We compare the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
-flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-
-We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
-gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
-
-%% S7
-\section{Conclusion}
-We have proposed some schemes and they work well.
-
-%% ACK
-\begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
-\end{acknowledgments}
-
-\appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
-\section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
-\begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
-For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
-and $L_{\uparrow} = \sigma^{+}/2$,
-$L_{\downarrow} = \sigma^{-}/2$
-are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
-$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
-$L_{2} = (I - \sigma_{z})/2$.
-
-%% T1
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & Analytic & QOC & & Analytic & QOC & \\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
-    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% AD
-\section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
-
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
-
-\bibliography{refs}
-
-\end{document}
+%% TODO: talk about quadratic cost vs. infidelity

### 2020-11-03 paper/4longitude.tex
separated sections by file

--- 2020-10-30 paper/main.tex
@@
-\documentclass[
-  amsfonts,
-  amsmath,
-  tbtags,
-  amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
-  twocolumn,
-  superscriptaddress,
-]{revtex4-2}
-
-\input{preamble}
-
-\begin{document}
-
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
-%% ABSTRACT
-\begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$.
-  We propose a method that takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by utilizing mixed-mode differentation.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm. We propose a computationally efficient metric
-  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
-  We apply these techniques to a fluxonium qubit with realisitc experimental constraints
-  and noise values. We are able to reduce gate errors by orders of magnitudes in the presence
-  of deviant parameters, achieving super-linear scaling for gate robustness as a function of the gate duration.
-  We use the same techniques to mitigate coherent errors due to time-varying magnetic flux noise
-  by two orders of magnitude. Additionally, we mitigate the gate error arising from longitudinal
-  relaxation by a factor of $5$. Further, we find that the numerical techniques we employ
-  are able to perform phase gates in arbitrary times, which is critical for operating multi-qubit
-  experiments in the lab frame.
-\end{abstract}
-
-\maketitle
-
-%% S1
-\section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-circuits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}. QOC techniques are used to
-control a system such that a set of optimization objectives are maximally fulfilled.
-For quantum computation, relevant objectives include achieving high fidelity
-gates and adhering to experimental constraints.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-It is desirable to make the optimization aware of experimental errors
-such as noise, system model errors, and control errors, to
-improve experimental performance on the optimization objectives.
-The field of robust control, deriving from classical control theory, has developed
-techniques to account for these errors in optimization (citations appreciated).
-In this work we employ robust control techniques to mitigate
-releastic decoherence and systematic errors that arise when controlling
-a superconducting fluxonium qubit.
-
-Numerical techniques for QOC have been remarkably successful in
-state preparation, designing high fidelity gates, and determining
-optimized pulse sequences (citations needed). Many popular numerical techniques
-are iterative, indirect shooting
-methods \cite{leung2017speedup, goer2019krotov, abdelhafez2019gradient,
-  machnes2015gradient}.
-An initial guess for the control parameters is used to forward-simulate
-the quantum state trajectory under the time-dependent Schroedinger equation. Then, gradients
-with respect to the controls are computed for an objective evaluated on
-the evolved quantum state. The gradients are used to update
-the control parameters and the procedure is repeated iteratively.
-A convenient way to enforce constraints in this framework is
-to employ control filtering or project gradients onto the constraint manifold.
-In this work we perform optimization using ALTRO, which combines
-an iterative indirect shooting method with the augmented Lagrangian
-method \cite{howell2019altro}. The augmented Lagrangian method
-allows us to enforce arbitrary, simultaneous constraints and achieve
-fast convergence without being restricted to the constraint
-manifold.
-
-QOC techniques have been effective in mitigating the effects
-of systematic errors. Typically, the model used in optimization
-differs from the experimental system due to parameter drift, the inability
-to realize control parameters exactly, and noise.
-These systematic errors cause the trajectory of the quantum system
-in the experiment to deviate from that predicted by the optimization.
-A set of controls is said to be robust if the experimental state trajectory
-it produces is insensitive to systematic errors.
-Considering the dynamical and geometric phases of quantum state
-evolution has led to analytic methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
-We mitigate systematic errors numerically by employing
-open loop optimization. We draw on robust trajectory optimization
-techniques to encode systematic errors in optimization objectives.
-We consider three
-techniques. The first is the sampling method, which has been applied
-previously in the context of QOC
-\cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
-The second is the unscented sampling
-method, which derives from the unscented transformation used
-for non-linear Kalman filtering
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-We propose a third method, which we call the derivative method.
-This method takes advantage of derivative information of systematic
-errors with respect to the state evolution. We make this method efficient
-by employing mixed-mode differentiation.
-
-Further, many analytic and numerical techniques have been developed to mitigate decoherence.
-Floquet techniques have been experimentally demonstrated to simultaneously mitigate
-decoherence due to longitudinal relaxation and pure dephasing
-\cite{huang2020engineering, mundada2020floquet}.
-Numerical techniques have mitigated decoherence by evolving the
-system using master equations \cite{rembold2020introduction} and employing
-Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
-In this work we take advantage of the known dependence of the
-quantum system's controls on longitudinal relaxation. We encode
-the decoherence in an efficient optimization objective that does
-not pay the increased computational cost of integrating a master equation.
-
-We compare our techniques using the heavy fluxonium qubit presented in
-in \cite{zhang2020universal}. This quantum system is an exciting
-platform for quantum computation due to its long longitudinal relaxation
-and pure dephasing times at the flux frustation point. This system is also
-efficient to simulate due to its accurate two-level approximation.
-We emphasize that the numerical techniques we employ extend to arbitrary quantum systems,
-not only the one we study here. We describe experimentally realistic constraints for this system
-and map them to the ALTRO framework. Next, we
-outline a method for making the optimization aware of longitudinal
-relaxation. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set. Then we present three methods for achieving
-robustness to systematic errors. We find that we are able to
-decrease the gate error arising from systematic errors super-linearly
-in the gate duration, achieving orders of magntidue reduction in gate errors.
-We also find that the numerical techniques we employ are able to produce
-fixed phase gates in arbitrary times, a method that could be used
-to control multi-qubit systems in the lab frame. Finally,
-we employ the robust control techniques to mitigate time-varying
-magnetic flux noise. We achieve a two order of magnitude improvement
-in gate errors for extended computations, an improvement critical
-for noisy, intermediate-scale quantum (NISQ) applications.
-
-%% S2
-\section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
-\label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
-In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
-algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
-
-(AL-iLQR Problem Statement) The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
-\begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-\end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
-
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-  \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
-  \end{aligned}
-\end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
-\end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
-
-%% S3
-\section{QOC on the Fluxonium}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
-\end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
-
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
-The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
-
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
-
-%% TODO: talk about quadratic cost vs. infidelity
-
 %% F1
@@
   \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
+    (a) Numerically optimized gates (red) and analytically optimized gates (blue).
     (b) $T_{1}$ interpolation function used in optimization. Markers
@@
     gate error after each gate application.
-    The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
+    The numerically optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
     in the figure.
@@
 relaxation time.
-To calculate the gate error due to longitudinal relaxation
+Computing the gate error due to longitudinal relaxation
 requires propagating density matrices of size $n \times n$ under Master equation
@@
 
-We compare the numerical method we have developed to the analytic gates
-on the task of achieving low gate errors in the presence of longitudinal relaxation
-for the $Z/2$, $Y/2$, and $X/2$ gates.
-The numerically optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate errors by a factor of $5$ over
-their analytic counterparts which is commensurate to their
-probability of longitudinal relaxation reductions, see Appendix.
-The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have imposed
-on the gates, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
 %% F2
@@
 
-%% S5
-\section{Robustness to Static Parameter Deviations}
-We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the Hamiltonian we use in optimization,
-leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
-we mitigate errors arising from the drift and finite measurement
-precision of the qubit frequency $\tilde{f_{q}} = f_{q} \pm \sigma_{f_{q}}$.
-We consider three robust control techniques.
-The first is the sampling method, which has
-been preposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling, carvalho2020error}. In the
-sampling method, multiple states are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
-transformation to accurately propagate a distribution
-representing the uncertainty in an evolving state
-due to a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. Derivative
-information encoding the sensitivity of the
-state trajectory with respect to the deviant parameter
-is used to penalize state trajectory deviations.
-
-In the sampling method, sample states evolve under a
-Hamiltonian where a parameter is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant values $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The gate error of each sample state is 
-penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-Each initial state is used to initialize two
-sample states $\ket{\psi^{\pm}}$.
-The initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-For this method the number of states in the augmented state vector
-scales as $O(dn^{2})$ because each of $n^{2}$ initial
-states is represented by $2d$ samples. $n$ is both the dimension of the Hilbert space
-and the size of the state vector. $d$ is the number of deviant parameters.
-
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}\hspace{0.025\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    sampling, unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is the cumulative
-    gate error after each gate application.
-  }
-\end{figure*}
-
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of the initial state. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states.
-Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. Hence,
-the number of states in the augmented state vector scales as $O(n^{3} + dn^{2})$.
-The gate error of each sample state is penalized as in the sampling method.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
-derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For a single state derivative and $N$ knot points this
-requires $O(N^2)$ matrix multiplications.
-Instead, we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting in
-$O(N)$ matrix multiplications. For example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
-For this method $m$ state derivatives are associated with each
-initial state from the operator basis.
-So, the number of states in the augmented state vector scales as
-$O(dmn^{2})$.
-
-To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$.
-
-We compare the numerical methods
-to an analytically derived $Z/2$ gate. This gate corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. Its erroneous rotation angle $2 \pi t_{Z/2} \delta f_{q}$ is linearly sensitive to
-the qubit frequency detuning, resulting in a gate error that is quadratically sensitive
-to the qubit frequency detutning.
-At a one-percent
-qubit frequency detuning the analytic gate achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-which is sufficient for quantum error correction, see Figure 2b.
-Although the analytic $Z/2$ gate performs well, it works
-only at the gate time $t_{Z/2}$. The ability to perform $Z$ rotations in arbitrary times is critical
-for operating multi-qubit experiments in the lab frame.
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times,
-see Figure 2b. These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-
-%% TOOD: intro sentence
-%% TODO: comment on unscented when it is redone
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-The derivative methods converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-Both methods achieve a super-linear scaling in their gate error as
-a function of the gate duration. The gate error for the 1\textsuperscript{st}-order
-derivative method approaches zero at the larmor period $1 / f_{q}$.
-We believe the 1\textsuperscript{st}-order method outperforms the \textsuperscript{nd}-order
-method due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
-
-%% TODO: comment on the fact that we can reduce the number of sample states
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
-
-%% S6
-\section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
-take advantage of the low frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
-
-We compare the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
-flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-
-We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
-gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
-
-%% S7
-\section{Conclusion}
-We have proposed some schemes and they work well.
-
-%% ACK
-\begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
-\end{acknowledgments}
-
-\appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
-\section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
-\begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
-For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
-and $L_{\uparrow} = \sigma^{+}/2$,
-$L_{\downarrow} = \sigma^{-}/2$
-are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
-$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
-$L_{2} = (I - \sigma_{z})/2$.
-
-%% T1
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & Analytic & QOC & & Analytic & QOC & \\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
-    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% AD
-\section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
-
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
-
-\bibliography{refs}
-
-\end{document}
+We compare the numerical method we have developed to the analytic gates
+on the task of achieving low gate errors in the presence of longitudinal relaxation
+for the $Z/2$, $Y/2$, and $X/2$ gates.
+The numerically optimized gates converge on similar solutions, a periodic
+waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
+They extend their gate times
+beyond their analytic counterparts, trading longer gate times for access
+to higher amplitudes and therefore higher $T_{1}$ times. All numerically optimized gates reduce
+their single gate errors by a factor of $5$ over
+their analytic counterparts which is commensurate to their
+probability of longitudinal relaxation reductions, see Appendix.
+The gate error reported in this text is the infidelity
+of the evolved state and the target state averaged over 1000 pseudo-randomly
+generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
+the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
+over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
+achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
+Both the analytic and optimized gates attain single gate errors sufficient for
+quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
+The low gate errors achieved by the optimized gates for extend computations
+are critical for noisy, intermediate-scale quantum (NISQ) applications.
+These improvements are significant for the realistic constraints we have imposed
+on the gates, and do not represent a fundamental limit to the optimization methods we have
+employed.

### 2020-11-03 paper/5static.tex
separated sections by file

--- 2020-10-30 paper/main.tex
@@
-\documentclass[
-  amsfonts,
-  amsmath,
-  tbtags,
-  amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
-  twocolumn,
-  superscriptaddress,
-]{revtex4-2}
-
-\input{preamble}
-
-\begin{document}
-
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
-%% ABSTRACT
-\begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$.
-  We propose a method that takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by utilizing mixed-mode differentation.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm. We propose a computationally efficient metric
-  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
-  We apply these techniques to a fluxonium qubit with realisitc experimental constraints
-  and noise values. We are able to reduce gate errors by orders of magnitudes in the presence
-  of deviant parameters, achieving super-linear scaling for gate robustness as a function of the gate duration.
-  We use the same techniques to mitigate coherent errors due to time-varying magnetic flux noise
-  by two orders of magnitude. Additionally, we mitigate the gate error arising from longitudinal
-  relaxation by a factor of $5$. Further, we find that the numerical techniques we employ
-  are able to perform phase gates in arbitrary times, which is critical for operating multi-qubit
-  experiments in the lab frame.
-\end{abstract}
-
-\maketitle
-
-%% S1
-\section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-circuits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}. QOC techniques are used to
-control a system such that a set of optimization objectives are maximally fulfilled.
-For quantum computation, relevant objectives include achieving high fidelity
-gates and adhering to experimental constraints.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-It is desirable to make the optimization aware of experimental errors
-such as noise, system model errors, and control errors, to
-improve experimental performance on the optimization objectives.
-The field of robust control, deriving from classical control theory, has developed
-techniques to account for these errors in optimization (citations appreciated).
-In this work we employ robust control techniques to mitigate
-releastic decoherence and systematic errors that arise when controlling
-a superconducting fluxonium qubit.
-
-Numerical techniques for QOC have been remarkably successful in
-state preparation, designing high fidelity gates, and determining
-optimized pulse sequences (citations needed). Many popular numerical techniques
-are iterative, indirect shooting
-methods \cite{leung2017speedup, goer2019krotov, abdelhafez2019gradient,
-  machnes2015gradient}.
-An initial guess for the control parameters is used to forward-simulate
-the quantum state trajectory under the time-dependent Schroedinger equation. Then, gradients
-with respect to the controls are computed for an objective evaluated on
-the evolved quantum state. The gradients are used to update
-the control parameters and the procedure is repeated iteratively.
-A convenient way to enforce constraints in this framework is
-to employ control filtering or project gradients onto the constraint manifold.
-In this work we perform optimization using ALTRO, which combines
-an iterative indirect shooting method with the augmented Lagrangian
-method \cite{howell2019altro}. The augmented Lagrangian method
-allows us to enforce arbitrary, simultaneous constraints and achieve
-fast convergence without being restricted to the constraint
-manifold.
-
-QOC techniques have been effective in mitigating the effects
-of systematic errors. Typically, the model used in optimization
-differs from the experimental system due to parameter drift, the inability
-to realize control parameters exactly, and noise.
-These systematic errors cause the trajectory of the quantum system
-in the experiment to deviate from that predicted by the optimization.
-A set of controls is said to be robust if the experimental state trajectory
-it produces is insensitive to systematic errors.
-Considering the dynamical and geometric phases of quantum state
-evolution has led to analytic methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
-We mitigate systematic errors numerically by employing
-open loop optimization. We draw on robust trajectory optimization
-techniques to encode systematic errors in optimization objectives.
-We consider three
-techniques. The first is the sampling method, which has been applied
-previously in the context of QOC
-\cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
-The second is the unscented sampling
-method, which derives from the unscented transformation used
-for non-linear Kalman filtering
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-We propose a third method, which we call the derivative method.
-This method takes advantage of derivative information of systematic
-errors with respect to the state evolution. We make this method efficient
-by employing mixed-mode differentiation.
-
-Further, many analytic and numerical techniques have been developed to mitigate decoherence.
-Floquet techniques have been experimentally demonstrated to simultaneously mitigate
-decoherence due to longitudinal relaxation and pure dephasing
-\cite{huang2020engineering, mundada2020floquet}.
-Numerical techniques have mitigated decoherence by evolving the
-system using master equations \cite{rembold2020introduction} and employing
-Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
-In this work we take advantage of the known dependence of the
-quantum system's controls on longitudinal relaxation. We encode
-the decoherence in an efficient optimization objective that does
-not pay the increased computational cost of integrating a master equation.
-
-We compare our techniques using the heavy fluxonium qubit presented in
-in \cite{zhang2020universal}. This quantum system is an exciting
-platform for quantum computation due to its long longitudinal relaxation
-and pure dephasing times at the flux frustation point. This system is also
-efficient to simulate due to its accurate two-level approximation.
-We emphasize that the numerical techniques we employ extend to arbitrary quantum systems,
-not only the one we study here. We describe experimentally realistic constraints for this system
-and map them to the ALTRO framework. Next, we
-outline a method for making the optimization aware of longitudinal
-relaxation. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set. Then we present three methods for achieving
-robustness to systematic errors. We find that we are able to
-decrease the gate error arising from systematic errors super-linearly
-in the gate duration, achieving orders of magntidue reduction in gate errors.
-We also find that the numerical techniques we employ are able to produce
-fixed phase gates in arbitrary times, a method that could be used
-to control multi-qubit systems in the lab frame. Finally,
-we employ the robust control techniques to mitigate time-varying
-magnetic flux noise. We achieve a two order of magnitude improvement
-in gate errors for extended computations, an improvement critical
-for noisy, intermediate-scale quantum (NISQ) applications.
-
-%% S2
-\section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
-\label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
-In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
-algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
-
-(AL-iLQR Problem Statement) The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
-\begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-\end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
-
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-  \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
-  \end{aligned}
-\end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
-\end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
-
-%% S3
-\section{QOC on the Fluxonium}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
-\end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
-
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
-The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
-
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
-
-%% TODO: talk about quadratic cost vs. infidelity
-
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is the cumulative
-    gate error after each gate application.
-    The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
-  }
-\end{figure*}
-
-%% S4
-\section{Longitudinal Relaxation Awareness}
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms.
-It is advantageous to tune the controls to extend the longitudinal
-relaxation time.
-To calculate the gate error due to longitudinal relaxation
-requires propagating density matrices of size $n \times n$ under Master equation
-dynamics, rather than state vectors of size $n$ under the TDSE dynamics.
-We avoid this increase in computational complexity by
-penalizing the integrated rate of longitudinal relaxation,
-i.e. the probability of longitudinal relaxation.
-Using this probability as proxy for the gate error incurred
-is reasonable because losses due to longitudinal relaxation
-increase monotonically in time.
-This technique can be extended to
-error channels which share the monotonically increasing property.
-Additionally, for a constant $T_{1}$ time, a shorter gate duration
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
-the one we study here.
-
-The longitudinal relaxation probability is given by
-\begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
-\end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-It is also possible to use a spline fit to theoretically obtained data.
-However, $T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data
-increases the fridge truth of the simulation.
-
-We allow the optimizer to tune the gate duration by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
-integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
-
-We compare the numerical method we have developed to the analytic gates
-on the task of achieving low gate errors in the presence of longitudinal relaxation
-for the $Z/2$, $Y/2$, and $X/2$ gates.
-The numerically optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate errors by a factor of $5$ over
-their analytic counterparts which is commensurate to their
-probability of longitudinal relaxation reductions, see Appendix.
-The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have imposed
-on the gates, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, sampling, unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
-    data points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency. The solutions for the analytic and
-    1\textsuperscript{st}-order derivative methods are shown at multiples
-    of the analytic gate time. The performance of the two methods is
-    indistinguishable at the analytic gate time $18$ns.
-  }
-\end{figure*}
-
 %% S5
@@
 techniques for future work.
-
-%% S6
-\section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
-take advantage of the low frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
-
-We compare the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
-flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-
-We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
-gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
-
-%% S7
-\section{Conclusion}
-We have proposed some schemes and they work well.
-
-%% ACK
-\begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
-\end{acknowledgments}
-
-\appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
-\section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
-\begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
-For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
-and $L_{\uparrow} = \sigma^{+}/2$,
-$L_{\downarrow} = \sigma^{-}/2$
-are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
-$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
-$L_{2} = (I - \sigma_{z})/2$.
-
-%% T1
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & Analytic & QOC & & Analytic & QOC & \\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
-    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% AD
-\section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
-
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
-
-\bibliography{refs}
-
-\end{document}

### 2020-11-03 paper/6stochastic.tex
separated sections by file

--- 2020-10-26 paper/main.tex
@@
-\documentclass[
-  amsfonts,
-  amsmath,
-  tbtags,
-  amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
-  twocolumn,
-  superscriptaddress,
-]{revtex4-2}
-
-\input{preamble}
-
-\begin{document}
-
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
-%% ABSTRACT
-\begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
-  takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by transforming
-  higher-order differential equations into a set of coupled first-order equations.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm.
-  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
-  for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimization and time-optimal control.
-\end{abstract}
-
-\maketitle
-
-%% S1
-\section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-qubits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}.
-In quantum control theory, as in classical control theory,
-optimization is performed
-to minimize an objective function. Relevant for the field
-of quantum computation is the task of achieving high fidelity quantum gates.
-In this application, the objective function includes contributions from the
-gate error associated with the state transfer of the system as well as contributions
-due to experimental constraints, such as the control amplitude.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-
-Many analytic and numerical techniques have been developed to control
-quantum systems. Analytic methods rely on solving the time-dependent
-Schroedinger equation analytically or deriving equations of motion
-from a suitable Lagrangian \cite{zhang2020universal, huang2020engineering, han2020experimental,
-  xu2020nonadiabatic, carlini2005quantum}. Analytic methods
-have been presented to address decoherence in quantum systems.
-Considering the dynamical and geometric phases of state
-evolution has led to methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
-Additionally, suitable choices of bases have been proposed that
-allow the experimentalist to infer tradeoffs in longitudinal relaxation
-and pure dephasing coherence times \cite{huang2020engineering}.
-To date, however, no analytical method has been presented
-to construct a gate while satisfying arbitrary experimental
-constraints. Furthermore, most analytic methods have only
-been demonstrated in the two-level approximation and become
-cumbersome to solve for larger Hilbert space sizes.
-
-Numerical quantum control techniques have developed in parallel
-with analytic frameworks.
-Most numerical techniques proposed for QOC are indirect shooting
-methods where an initial guess for the control trajectory is forward-simulated,
-e.g. under the time-dependent Schroedinger equation,
-and first-order gradients are computed for a functional evaluated
-on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
-  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
-Most techniques formulate the
-problem as unconstrained or use
-projective gradient methods to enforce constraints
-\cite{machnes2015gradient}.
-The latter approach relies on constructing
-bases for constraint spaces--which is not always feasible--and
-may hinder convergence.
-Methods for mitigating decoherence using numerical
-techniques have employed master equation dynamics (citation needed),
-Monte Carlo style trajectories \cite{abdelhafez2019gradient},
-and proposals have been made to simulate multiple trajectories
-\cite{reinhold2019controlling, rembold2020introduction}.
-
-In this work, we employ state of the art
-trajectory optimization methods
-to formulate the quantum
-optimal control problem as a constrained optimization problem, allowing
-us to enforce arbitrary experimental constraints.
-Additionally,
-we introduce methods for constructing gates that mitigate longitudinal
-relaxation and are robust to experimental errors, and therefore
-pure dephasing, the dominant barriers to experimentally
-realizing practical quantum computing.
-
-As an example, we study the quantum optimal control problem on
-the fluxonium qubit presented in \cite{zhang2020universal}.
-%%TODO: Fluxonium qubit is exciting and computationally feasible,
-%%but our methods extend to arbitrary platforms.
-We describe experimentally realistic constraints and map them to
-the trajectory optimization framework. Next, we
-outline a method for making the optimization longitudinal
-relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set.
-We present two methods for achieving robustness to pure dephasing
-and compare them to existing dynamic decoupling methods. We find that
-the methods we employ are able to produce high fidelity gates when
-subject to parameter spreads of $1\%$.
-
-%% S2
-\section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
-\label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
-In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
-algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
-
-(AL-iLQR Problem Statement) The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
-\begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-\end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
-
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-  \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
-  \end{aligned}
-\end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
-\end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
-
-%% S3
-\section{QOC on the Fluxonium}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
-\end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
-
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
-The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
-
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
-
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is computed after
-    each gate appliaction. The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
-  }
-\end{figure*}
-
-%% S4
-\section{Longitudinal Relaxation Awareness}
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms,
-and it is advantageous to maximize it by tuning the controls.
-In the fluxonium, the longitudinal relaxation time varies with
-the amplitude $a$. To completely simulate
-the effect of longitudinal relaxation on the gate error requires
-propagating density matrices of size $n \times n$ under Master equation
-dynamics rather than state vectors of size $n$ under the TDSE dynamics.
-We avoid this increase in computational complexity by
-penalizing the probability that the qubit decays as a result of
-longitudinal relaxation. This modification only
-increases the size of the augmented state vector by a single real number
-and allows us to effectively mitigate the error arising from longitudinal
-relaxation. Additionally, for a constant $T_{1}$ time, a shorter gate duration
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
-the one we study here.
-
-The longitudinal relaxation probability is given by
-\begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
-\end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-Calculating $T_{1}$ directly from theoretical
-considerations requires many high-rank
-eigendecompositions, which
-is computationally expensive. Additionally,
-$T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation.
-
-We allow the optimizer to tune the gate duration by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
-integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
-
-The $T_{1}$ optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate probability of longitudinal relaxation by a factor of $5$ over
-their analytic counterparts and achieve a commensurate reduction in their single gate error,
-see Appendix. The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have put on the
-device, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
-    points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency at multiples of the analytic gate duration for the analytic
-    and 1\textsuperscript{st}-order derivative methods.
-  }
-\end{figure*}
-
-%% S5
-\section{Robustness to Static Parameter Deviations}
-We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the model we use in optimization,
-the Hamiltonian, leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
-we study mitigating errors arising from the finite measurement
-precision of the qubit frequency $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$.
-We consider three flavors of robust control
-techniques. The first is the sampling method, which has
-been preposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling}. In the
-sampling method, multiple trajectories are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
-transformation to accurately propagate a distribution
-representing the evolving state as a result of a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. This scheme
-uses derivative information about how the state
-evolution is affected by the deviant parameter to
-mitigate trajectory deviations.
-
-In the sampling method we propagate additional sample states
-in the augmented state vector that follow deviant dynamics
-and penalize their differences from the target state.
-The states evolve under the deviant
-Hamiltonian where the parameter of interest is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The infidelities of the samples and the target
-state are penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
-resulting in eight additional states added to the augmented state vector.
-The four initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-This results in $O(n^{2})$ states added to the augmented
-state vector where $n$ is both the dimension of the Hilbert space
-and the size of the state vector.
-
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}\hspace{0.025\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is computed
-    after each gate application.
-  }
-\end{figure*}
-
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of $\ket{\psi}$. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states
-where $d$ is the number of deviant parameters. Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. This results
-in $O(n^{3})$ states added to the augmented state vector.
-The costs at each knot point are the infidelities between the samples
-and the target state, as in the sampling method.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
-derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For $d$ sensitivity parameters and $N$ knot points
-this requires $O(d N^2)$ matrix multiplications. Instead,
-we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting
-in $O(d N)$ matrix multiplications. For the piece-wise constant control
-scheme we employ here, $N$ is the number of control parameters. Typically
-$d \ll N$ so this is a good tradeoff. As an example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
-
-To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$. We compare the numerical methods
-to an analytically derived $Z/2$ gate which corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. It is linearly sensitive to qubit frequency detunings,
-the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
-qubit frequency detuning, it achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-sufficient for quantum error correction.
-
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
-The ability to perform $Z$ rotations in arbitrary times
-is critical for operating multi-qubit experiments in the lab frame.
-These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-
-The derivative methods also converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-They achieve super-linear scaling in their gate error as
-a function of the gate duration which approaches zero at the larmor
-period $1 / f_{q}$. We observe that on a range of problems with small parameter detunings,
-the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
-derivative method. We believe this is due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
-
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
-
 %% S6
@@
 an effective method for mitigating them.
-
-%% S7
-\section{Conclusion}
-We have proposed some schemes and they work well.
-
-%% ACK
-\begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
-\end{acknowledgments}
-
-\appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
-\section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
-\begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
-For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
-and $L_{\uparrow} = \sigma^{+}/2$,
-$L_{\downarrow} = \sigma^{-}/2$
-are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
-$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
-$L_{2} = (I - \sigma_{z})/2$.
-
-%% T1
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & Analytic & QOC & & Analytic & QOC & \\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
-    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% AD
-\section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
-
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
-
-\bibliography{refs}
-
-\end{document}

### 2020-11-03 paper/7conclusion.tex
separated sections by file

--- 2020-10-26 paper/main.tex
@@
-\documentclass[
-  amsfonts,
-  amsmath,
-  tbtags,
-  amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
-  twocolumn,
-  superscriptaddress,
-]{revtex4-2}
-
-\input{preamble}
-
-\begin{document}
-
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
-%% ABSTRACT
-\begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
-  takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by transforming
-  higher-order differential equations into a set of coupled first-order equations.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm.
-  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
-  for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimization and time-optimal control.
-\end{abstract}
-
-\maketitle
-
-%% S1
-\section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-qubits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}.
-In quantum control theory, as in classical control theory,
-optimization is performed
-to minimize an objective function. Relevant for the field
-of quantum computation is the task of achieving high fidelity quantum gates.
-In this application, the objective function includes contributions from the
-gate error associated with the state transfer of the system as well as contributions
-due to experimental constraints, such as the control amplitude.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-
-Many analytic and numerical techniques have been developed to control
-quantum systems. Analytic methods rely on solving the time-dependent
-Schroedinger equation analytically or deriving equations of motion
-from a suitable Lagrangian \cite{zhang2020universal, huang2020engineering, han2020experimental,
-  xu2020nonadiabatic, carlini2005quantum}. Analytic methods
-have been presented to address decoherence in quantum systems.
-Considering the dynamical and geometric phases of state
-evolution has led to methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
-Additionally, suitable choices of bases have been proposed that
-allow the experimentalist to infer tradeoffs in longitudinal relaxation
-and pure dephasing coherence times \cite{huang2020engineering}.
-To date, however, no analytical method has been presented
-to construct a gate while satisfying arbitrary experimental
-constraints. Furthermore, most analytic methods have only
-been demonstrated in the two-level approximation and become
-cumbersome to solve for larger Hilbert space sizes.
-
-Numerical quantum control techniques have developed in parallel
-with analytic frameworks.
-Most numerical techniques proposed for QOC are indirect shooting
-methods where an initial guess for the control trajectory is forward-simulated,
-e.g. under the time-dependent Schroedinger equation,
-and first-order gradients are computed for a functional evaluated
-on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
-  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
-Most techniques formulate the
-problem as unconstrained or use
-projective gradient methods to enforce constraints
-\cite{machnes2015gradient}.
-The latter approach relies on constructing
-bases for constraint spaces--which is not always feasible--and
-may hinder convergence.
-Methods for mitigating decoherence using numerical
-techniques have employed master equation dynamics (citation needed),
-Monte Carlo style trajectories \cite{abdelhafez2019gradient},
-and proposals have been made to simulate multiple trajectories
-\cite{reinhold2019controlling, rembold2020introduction}.
-
-In this work, we employ state of the art
-trajectory optimization methods
-to formulate the quantum
-optimal control problem as a constrained optimization problem, allowing
-us to enforce arbitrary experimental constraints.
-Additionally,
-we introduce methods for constructing gates that mitigate longitudinal
-relaxation and are robust to experimental errors, and therefore
-pure dephasing, the dominant barriers to experimentally
-realizing practical quantum computing.
-
-As an example, we study the quantum optimal control problem on
-the fluxonium qubit presented in \cite{zhang2020universal}.
-%%TODO: Fluxonium qubit is exciting and computationally feasible,
-%%but our methods extend to arbitrary platforms.
-We describe experimentally realistic constraints and map them to
-the trajectory optimization framework. Next, we
-outline a method for making the optimization longitudinal
-relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set.
-We present two methods for achieving robustness to pure dephasing
-and compare them to existing dynamic decoupling methods. We find that
-the methods we employ are able to produce high fidelity gates when
-subject to parameter spreads of $1\%$.
-
-%% S2
-\section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
-\label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
-In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
-algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
-
-(AL-iLQR Problem Statement) The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
-\begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-\end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
-
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-  \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
-  \end{aligned}
-\end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
-\end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
-
-%% S3
-\section{QOC on the Fluxonium}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
-\end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
-
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
-The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
-
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
-
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is computed after
-    each gate appliaction. The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
-  }
-\end{figure*}
-
-%% S4
-\section{Longitudinal Relaxation Awareness}
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms,
-and it is advantageous to maximize it by tuning the controls.
-In the fluxonium, the longitudinal relaxation time varies with
-the amplitude $a$. To completely simulate
-the effect of longitudinal relaxation on the gate error requires
-propagating density matrices of size $n \times n$ under Master equation
-dynamics rather than state vectors of size $n$ under the TDSE dynamics.
-We avoid this increase in computational complexity by
-penalizing the probability that the qubit decays as a result of
-longitudinal relaxation. This modification only
-increases the size of the augmented state vector by a single real number
-and allows us to effectively mitigate the error arising from longitudinal
-relaxation. Additionally, for a constant $T_{1}$ time, a shorter gate duration
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
-the one we study here.
-
-The longitudinal relaxation probability is given by
-\begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
-\end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-Calculating $T_{1}$ directly from theoretical
-considerations requires many high-rank
-eigendecompositions, which
-is computationally expensive. Additionally,
-$T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation.
-
-We allow the optimizer to tune the gate duration by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
-integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
-
-The $T_{1}$ optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate probability of longitudinal relaxation by a factor of $5$ over
-their analytic counterparts and achieve a commensurate reduction in their single gate error,
-see Appendix. The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have put on the
-device, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
-    points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency at multiples of the analytic gate duration for the analytic
-    and 1\textsuperscript{st}-order derivative methods.
-  }
-\end{figure*}
-
-%% S5
-\section{Robustness to Static Parameter Deviations}
-We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the model we use in optimization,
-the Hamiltonian, leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
-we study mitigating errors arising from the finite measurement
-precision of the qubit frequency $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$.
-We consider three flavors of robust control
-techniques. The first is the sampling method, which has
-been preposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling}. In the
-sampling method, multiple trajectories are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
-transformation to accurately propagate a distribution
-representing the evolving state as a result of a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. This scheme
-uses derivative information about how the state
-evolution is affected by the deviant parameter to
-mitigate trajectory deviations.
-
-In the sampling method we propagate additional sample states
-in the augmented state vector that follow deviant dynamics
-and penalize their differences from the target state.
-The states evolve under the deviant
-Hamiltonian where the parameter of interest is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The infidelities of the samples and the target
-state are penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
-resulting in eight additional states added to the augmented state vector.
-The four initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-This results in $O(n^{2})$ states added to the augmented
-state vector where $n$ is both the dimension of the Hilbert space
-and the size of the state vector.
-
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}\hspace{0.025\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is computed
-    after each gate application.
-  }
-\end{figure*}
-
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of $\ket{\psi}$. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states
-where $d$ is the number of deviant parameters. Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. This results
-in $O(n^{3})$ states added to the augmented state vector.
-The costs at each knot point are the infidelities between the samples
-and the target state, as in the sampling method.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
-derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For $d$ sensitivity parameters and $N$ knot points
-this requires $O(d N^2)$ matrix multiplications. Instead,
-we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting
-in $O(d N)$ matrix multiplications. For the piece-wise constant control
-scheme we employ here, $N$ is the number of control parameters. Typically
-$d \ll N$ so this is a good tradeoff. As an example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
-
-To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$. We compare the numerical methods
-to an analytically derived $Z/2$ gate which corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. It is linearly sensitive to qubit frequency detunings,
-the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
-qubit frequency detuning, it achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-sufficient for quantum error correction.
-
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
-The ability to perform $Z$ rotations in arbitrary times
-is critical for operating multi-qubit experiments in the lab frame.
-These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-
-The derivative methods also converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-They achieve super-linear scaling in their gate error as
-a function of the gate duration which approaches zero at the larmor
-period $1 / f_{q}$. We observe that on a range of problems with small parameter detunings,
-the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
-derivative method. We believe this is due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
-
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
-
-%% S6
-\section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
-take advantage of the low frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
-
-We compare the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
-flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-
-We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
-gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
-
 %% S7
 \section{Conclusion}
-We have proposed some schemes and they work well.
-
-%% ACK
-\begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
-\end{acknowledgments}
-
-\appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
-\section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
-\begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
-For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
-and $L_{\uparrow} = \sigma^{+}/2$,
-$L_{\downarrow} = \sigma^{-}/2$
-are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
-$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
-$L_{2} = (I - \sigma_{z})/2$.
-
-%% T1
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & Analytic & QOC & & Analytic & QOC & \\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
-    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% AD
-\section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
-
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
-
-\bibliography{refs}
-
-\end{document}
+In conclusion, we have demonstrated techniques for achieving robustness to systematic
+errors and mitigating decoherence on a quantum system using state-of-the-art trajectory
+optimization. We have proposed a scheme for mitigating longitudinal relaxation with time-optimal
+control and an efficient optimization metric that comes at a constant computational cost as
+opposed to integrating a master equation which scales quadratically with
+the dimension of the Hilbert space.
+We have proposed the derivative method for robust control which achieves
+super-linear gate error reductions in the gate duration for the problem we studied here.
+We have shown that robust control techniques can be used to mitigate decoherence due
+to 1/$f$ flux noise, a dominant source of coherent errors for flux controlled qubits.
+The numerical techniques we have studied here can be used to perform phase gates in arbitrary times,
+which will be critical for synchronizing multi-qubit systems. Our
+techniques significantly improve the state-of-the-art for controlling quantum systems subject
+to systematic errors and decoherence, and will be used in efforts to lower gate errors
+for fault-tolerant quantum computing and NISQ applications.

### 2020-11-03 paper/8ack.tex
separated sections by file

--- 2020-10-26 paper/main.tex
@@
-\documentclass[
-  amsfonts,
-  amsmath,
-  tbtags,
-  amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
-  twocolumn,
-  superscriptaddress,
-]{revtex4-2}
-
-\input{preamble}
-
-\begin{document}
-
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
-%% ABSTRACT
-\begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
-  takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by transforming
-  higher-order differential equations into a set of coupled first-order equations.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm.
-  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
-  for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimization and time-optimal control.
-\end{abstract}
-
-\maketitle
-
-%% S1
-\section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-qubits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}.
-In quantum control theory, as in classical control theory,
-optimization is performed
-to minimize an objective function. Relevant for the field
-of quantum computation is the task of achieving high fidelity quantum gates.
-In this application, the objective function includes contributions from the
-gate error associated with the state transfer of the system as well as contributions
-due to experimental constraints, such as the control amplitude.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-
-Many analytic and numerical techniques have been developed to control
-quantum systems. Analytic methods rely on solving the time-dependent
-Schroedinger equation analytically or deriving equations of motion
-from a suitable Lagrangian \cite{zhang2020universal, huang2020engineering, han2020experimental,
-  xu2020nonadiabatic, carlini2005quantum}. Analytic methods
-have been presented to address decoherence in quantum systems.
-Considering the dynamical and geometric phases of state
-evolution has led to methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
-Additionally, suitable choices of bases have been proposed that
-allow the experimentalist to infer tradeoffs in longitudinal relaxation
-and pure dephasing coherence times \cite{huang2020engineering}.
-To date, however, no analytical method has been presented
-to construct a gate while satisfying arbitrary experimental
-constraints. Furthermore, most analytic methods have only
-been demonstrated in the two-level approximation and become
-cumbersome to solve for larger Hilbert space sizes.
-
-Numerical quantum control techniques have developed in parallel
-with analytic frameworks.
-Most numerical techniques proposed for QOC are indirect shooting
-methods where an initial guess for the control trajectory is forward-simulated,
-e.g. under the time-dependent Schroedinger equation,
-and first-order gradients are computed for a functional evaluated
-on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
-  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
-Most techniques formulate the
-problem as unconstrained or use
-projective gradient methods to enforce constraints
-\cite{machnes2015gradient}.
-The latter approach relies on constructing
-bases for constraint spaces--which is not always feasible--and
-may hinder convergence.
-Methods for mitigating decoherence using numerical
-techniques have employed master equation dynamics (citation needed),
-Monte Carlo style trajectories \cite{abdelhafez2019gradient},
-and proposals have been made to simulate multiple trajectories
-\cite{reinhold2019controlling, rembold2020introduction}.
-
-In this work, we employ state of the art
-trajectory optimization methods
-to formulate the quantum
-optimal control problem as a constrained optimization problem, allowing
-us to enforce arbitrary experimental constraints.
-Additionally,
-we introduce methods for constructing gates that mitigate longitudinal
-relaxation and are robust to experimental errors, and therefore
-pure dephasing, the dominant barriers to experimentally
-realizing practical quantum computing.
-
-As an example, we study the quantum optimal control problem on
-the fluxonium qubit presented in \cite{zhang2020universal}.
-%%TODO: Fluxonium qubit is exciting and computationally feasible,
-%%but our methods extend to arbitrary platforms.
-We describe experimentally realistic constraints and map them to
-the trajectory optimization framework. Next, we
-outline a method for making the optimization longitudinal
-relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set.
-We present two methods for achieving robustness to pure dephasing
-and compare them to existing dynamic decoupling methods. We find that
-the methods we employ are able to produce high fidelity gates when
-subject to parameter spreads of $1\%$.
-
-%% S2
-\section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
-\label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
-In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
-algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
-
-(AL-iLQR Problem Statement) The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
-\begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-\end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
-
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-  \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
-  \end{aligned}
-\end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
-\end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
-
-%% S3
-\section{QOC on the Fluxonium}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
-\end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
-
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
-The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
-
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
-
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is computed after
-    each gate appliaction. The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
-  }
-\end{figure*}
-
-%% S4
-\section{Longitudinal Relaxation Awareness}
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms,
-and it is advantageous to maximize it by tuning the controls.
-In the fluxonium, the longitudinal relaxation time varies with
-the amplitude $a$. To completely simulate
-the effect of longitudinal relaxation on the gate error requires
-propagating density matrices of size $n \times n$ under Master equation
-dynamics rather than state vectors of size $n$ under the TDSE dynamics.
-We avoid this increase in computational complexity by
-penalizing the probability that the qubit decays as a result of
-longitudinal relaxation. This modification only
-increases the size of the augmented state vector by a single real number
-and allows us to effectively mitigate the error arising from longitudinal
-relaxation. Additionally, for a constant $T_{1}$ time, a shorter gate duration
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
-the one we study here.
-
-The longitudinal relaxation probability is given by
-\begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
-\end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-Calculating $T_{1}$ directly from theoretical
-considerations requires many high-rank
-eigendecompositions, which
-is computationally expensive. Additionally,
-$T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation.
-
-We allow the optimizer to tune the gate duration by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
-integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
-
-The $T_{1}$ optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate probability of longitudinal relaxation by a factor of $5$ over
-their analytic counterparts and achieve a commensurate reduction in their single gate error,
-see Appendix. The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have put on the
-device, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
-    points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency at multiples of the analytic gate duration for the analytic
-    and 1\textsuperscript{st}-order derivative methods.
-  }
-\end{figure*}
-
-%% S5
-\section{Robustness to Static Parameter Deviations}
-We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the model we use in optimization,
-the Hamiltonian, leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
-we study mitigating errors arising from the finite measurement
-precision of the qubit frequency $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$.
-We consider three flavors of robust control
-techniques. The first is the sampling method, which has
-been preposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling}. In the
-sampling method, multiple trajectories are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
-transformation to accurately propagate a distribution
-representing the evolving state as a result of a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. This scheme
-uses derivative information about how the state
-evolution is affected by the deviant parameter to
-mitigate trajectory deviations.
-
-In the sampling method we propagate additional sample states
-in the augmented state vector that follow deviant dynamics
-and penalize their differences from the target state.
-The states evolve under the deviant
-Hamiltonian where the parameter of interest is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The infidelities of the samples and the target
-state are penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
-resulting in eight additional states added to the augmented state vector.
-The four initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-This results in $O(n^{2})$ states added to the augmented
-state vector where $n$ is both the dimension of the Hilbert space
-and the size of the state vector.
-
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}\hspace{0.025\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is computed
-    after each gate application.
-  }
-\end{figure*}
-
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of $\ket{\psi}$. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states
-where $d$ is the number of deviant parameters. Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. This results
-in $O(n^{3})$ states added to the augmented state vector.
-The costs at each knot point are the infidelities between the samples
-and the target state, as in the sampling method.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
-derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For $d$ sensitivity parameters and $N$ knot points
-this requires $O(d N^2)$ matrix multiplications. Instead,
-we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting
-in $O(d N)$ matrix multiplications. For the piece-wise constant control
-scheme we employ here, $N$ is the number of control parameters. Typically
-$d \ll N$ so this is a good tradeoff. As an example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
-
-To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$. We compare the numerical methods
-to an analytically derived $Z/2$ gate which corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. It is linearly sensitive to qubit frequency detunings,
-the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
-qubit frequency detuning, it achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-sufficient for quantum error correction.
-
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
-The ability to perform $Z$ rotations in arbitrary times
-is critical for operating multi-qubit experiments in the lab frame.
-These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-
-The derivative methods also converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-They achieve super-linear scaling in their gate error as
-a function of the gate duration which approaches zero at the larmor
-period $1 / f_{q}$. We observe that on a range of problems with small parameter detunings,
-the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
-derivative method. We believe this is due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
-
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
-
-%% S6
-\section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
-take advantage of the low frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
-
-We compare the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
-flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-
-We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
-gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
-
-%% S7
-\section{Conclusion}
-We have proposed some schemes and they work well.
-
 %% ACK
@@
 \end{acknowledgments}
-
-\appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
-\section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
-\begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
-For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
-and $L_{\uparrow} = \sigma^{+}/2$,
-$L_{\downarrow} = \sigma^{-}/2$
-are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
-$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
-$L_{2} = (I - \sigma_{z})/2$.
-
-%% T1
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & Analytic & QOC & & Analytic & QOC & \\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
-    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% AD
-\section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
-
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
-
-\bibliography{refs}
-
-\end{document}

### 2020-11-03 paper/a2dissipation.tex
separated sections by file

--- 2020-10-26 paper/main.tex
@@
-\documentclass[
-  amsfonts,
-  amsmath,
-  tbtags,
-  amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
-  twocolumn,
-  superscriptaddress,
-]{revtex4-2}
-
-\input{preamble}
-
-\begin{document}
-
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
-%% ABSTRACT
-\begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
-  takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by transforming
-  higher-order differential equations into a set of coupled first-order equations.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm.
-  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
-  for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimization and time-optimal control.
-\end{abstract}
-
-\maketitle
-
-%% S1
-\section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-qubits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}.
-In quantum control theory, as in classical control theory,
-optimization is performed
-to minimize an objective function. Relevant for the field
-of quantum computation is the task of achieving high fidelity quantum gates.
-In this application, the objective function includes contributions from the
-gate error associated with the state transfer of the system as well as contributions
-due to experimental constraints, such as the control amplitude.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-
-Many analytic and numerical techniques have been developed to control
-quantum systems. Analytic methods rely on solving the time-dependent
-Schroedinger equation analytically or deriving equations of motion
-from a suitable Lagrangian \cite{zhang2020universal, huang2020engineering, han2020experimental,
-  xu2020nonadiabatic, carlini2005quantum}. Analytic methods
-have been presented to address decoherence in quantum systems.
-Considering the dynamical and geometric phases of state
-evolution has led to methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
-Additionally, suitable choices of bases have been proposed that
-allow the experimentalist to infer tradeoffs in longitudinal relaxation
-and pure dephasing coherence times \cite{huang2020engineering}.
-To date, however, no analytical method has been presented
-to construct a gate while satisfying arbitrary experimental
-constraints. Furthermore, most analytic methods have only
-been demonstrated in the two-level approximation and become
-cumbersome to solve for larger Hilbert space sizes.
-
-Numerical quantum control techniques have developed in parallel
-with analytic frameworks.
-Most numerical techniques proposed for QOC are indirect shooting
-methods where an initial guess for the control trajectory is forward-simulated,
-e.g. under the time-dependent Schroedinger equation,
-and first-order gradients are computed for a functional evaluated
-on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
-  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
-Most techniques formulate the
-problem as unconstrained or use
-projective gradient methods to enforce constraints
-\cite{machnes2015gradient}.
-The latter approach relies on constructing
-bases for constraint spaces--which is not always feasible--and
-may hinder convergence.
-Methods for mitigating decoherence using numerical
-techniques have employed master equation dynamics (citation needed),
-Monte Carlo style trajectories \cite{abdelhafez2019gradient},
-and proposals have been made to simulate multiple trajectories
-\cite{reinhold2019controlling, rembold2020introduction}.
-
-In this work, we employ state of the art
-trajectory optimization methods
-to formulate the quantum
-optimal control problem as a constrained optimization problem, allowing
-us to enforce arbitrary experimental constraints.
-Additionally,
-we introduce methods for constructing gates that mitigate longitudinal
-relaxation and are robust to experimental errors, and therefore
-pure dephasing, the dominant barriers to experimentally
-realizing practical quantum computing.
-
-As an example, we study the quantum optimal control problem on
-the fluxonium qubit presented in \cite{zhang2020universal}.
-%%TODO: Fluxonium qubit is exciting and computationally feasible,
-%%but our methods extend to arbitrary platforms.
-We describe experimentally realistic constraints and map them to
-the trajectory optimization framework. Next, we
-outline a method for making the optimization longitudinal
-relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set.
-We present two methods for achieving robustness to pure dephasing
-and compare them to existing dynamic decoupling methods. We find that
-the methods we employ are able to produce high fidelity gates when
-subject to parameter spreads of $1\%$.
-
-%% S2
-\section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
-\label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
-In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
-algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
-
-(AL-iLQR Problem Statement) The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
-\begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-\end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
-
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-  \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
-  \end{aligned}
-\end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
-\end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
-
-%% S3
-\section{QOC on the Fluxonium}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
-\end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
-
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
-The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
-
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
-
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is computed after
-    each gate appliaction. The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
-  }
-\end{figure*}
-
-%% S4
-\section{Longitudinal Relaxation Awareness}
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms,
-and it is advantageous to maximize it by tuning the controls.
-In the fluxonium, the longitudinal relaxation time varies with
-the amplitude $a$. To completely simulate
-the effect of longitudinal relaxation on the gate error requires
-propagating density matrices of size $n \times n$ under Master equation
-dynamics rather than state vectors of size $n$ under the TDSE dynamics.
-We avoid this increase in computational complexity by
-penalizing the probability that the qubit decays as a result of
-longitudinal relaxation. This modification only
-increases the size of the augmented state vector by a single real number
-and allows us to effectively mitigate the error arising from longitudinal
-relaxation. Additionally, for a constant $T_{1}$ time, a shorter gate duration
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
-the one we study here.
-
-The longitudinal relaxation probability is given by
-\begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
-\end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-Calculating $T_{1}$ directly from theoretical
-considerations requires many high-rank
-eigendecompositions, which
-is computationally expensive. Additionally,
-$T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation.
-
-We allow the optimizer to tune the gate duration by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
-integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
-
-The $T_{1}$ optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate probability of longitudinal relaxation by a factor of $5$ over
-their analytic counterparts and achieve a commensurate reduction in their single gate error,
-see Appendix. The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have put on the
-device, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
-    points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency at multiples of the analytic gate duration for the analytic
-    and 1\textsuperscript{st}-order derivative methods.
-  }
-\end{figure*}
-
-%% S5
-\section{Robustness to Static Parameter Deviations}
-We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the model we use in optimization,
-the Hamiltonian, leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
-we study mitigating errors arising from the finite measurement
-precision of the qubit frequency $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$.
-We consider three flavors of robust control
-techniques. The first is the sampling method, which has
-been preposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling}. In the
-sampling method, multiple trajectories are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
-transformation to accurately propagate a distribution
-representing the evolving state as a result of a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. This scheme
-uses derivative information about how the state
-evolution is affected by the deviant parameter to
-mitigate trajectory deviations.
-
-In the sampling method we propagate additional sample states
-in the augmented state vector that follow deviant dynamics
-and penalize their differences from the target state.
-The states evolve under the deviant
-Hamiltonian where the parameter of interest is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The infidelities of the samples and the target
-state are penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
-resulting in eight additional states added to the augmented state vector.
-The four initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-This results in $O(n^{2})$ states added to the augmented
-state vector where $n$ is both the dimension of the Hilbert space
-and the size of the state vector.
-
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}\hspace{0.025\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is computed
-    after each gate application.
-  }
-\end{figure*}
-
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of $\ket{\psi}$. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states
-where $d$ is the number of deviant parameters. Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. This results
-in $O(n^{3})$ states added to the augmented state vector.
-The costs at each knot point are the infidelities between the samples
-and the target state, as in the sampling method.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
-derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For $d$ sensitivity parameters and $N$ knot points
-this requires $O(d N^2)$ matrix multiplications. Instead,
-we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting
-in $O(d N)$ matrix multiplications. For the piece-wise constant control
-scheme we employ here, $N$ is the number of control parameters. Typically
-$d \ll N$ so this is a good tradeoff. As an example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
-
-To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$. We compare the numerical methods
-to an analytically derived $Z/2$ gate which corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. It is linearly sensitive to qubit frequency detunings,
-the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
-qubit frequency detuning, it achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-sufficient for quantum error correction.
-
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
-The ability to perform $Z$ rotations in arbitrary times
-is critical for operating multi-qubit experiments in the lab frame.
-These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-
-The derivative methods also converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-They achieve super-linear scaling in their gate error as
-a function of the gate duration which approaches zero at the larmor
-period $1 / f_{q}$. We observe that on a range of problems with small parameter detunings,
-the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
-derivative method. We believe this is due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
-
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
-
-%% S6
-\section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
-take advantage of the low frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
-
-We compare the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
-flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-
-We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
-gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
-
-%% S7
-\section{Conclusion}
-We have proposed some schemes and they work well.
-
-%% ACK
-\begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
-\end{acknowledgments}
-
-\appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
+%% A1
 \section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
+To completely model the effect of longitudinal
+relaxation on the quantum system we employ
+the Lindblad master equation. This equation takes the form
 \begin{equation}
@@
 \end{table}
-
-%% AD
-\section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
-
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
-
-\bibliography{refs}
-
-\end{document}

### 2020-11-03 paper/a3derivative.tex
separated sections by file

--- 2020-10-26 paper/main.tex
@@
-\documentclass[
-  amsfonts,
-  amsmath,
-  tbtags,
-  amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
-  twocolumn,
-  superscriptaddress,
-]{revtex4-2}
-
-\input{preamble}
-
-\begin{document}
-
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
-%% ABSTRACT
-\begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
-  takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by transforming
-  higher-order differential equations into a set of coupled first-order equations.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm.
-  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
-  for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimization and time-optimal control.
-\end{abstract}
-
-\maketitle
-
-%% S1
-\section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-qubits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}.
-In quantum control theory, as in classical control theory,
-optimization is performed
-to minimize an objective function. Relevant for the field
-of quantum computation is the task of achieving high fidelity quantum gates.
-In this application, the objective function includes contributions from the
-gate error associated with the state transfer of the system as well as contributions
-due to experimental constraints, such as the control amplitude.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-
-Many analytic and numerical techniques have been developed to control
-quantum systems. Analytic methods rely on solving the time-dependent
-Schroedinger equation analytically or deriving equations of motion
-from a suitable Lagrangian \cite{zhang2020universal, huang2020engineering, han2020experimental,
-  xu2020nonadiabatic, carlini2005quantum}. Analytic methods
-have been presented to address decoherence in quantum systems.
-Considering the dynamical and geometric phases of state
-evolution has led to methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
-Additionally, suitable choices of bases have been proposed that
-allow the experimentalist to infer tradeoffs in longitudinal relaxation
-and pure dephasing coherence times \cite{huang2020engineering}.
-To date, however, no analytical method has been presented
-to construct a gate while satisfying arbitrary experimental
-constraints. Furthermore, most analytic methods have only
-been demonstrated in the two-level approximation and become
-cumbersome to solve for larger Hilbert space sizes.
-
-Numerical quantum control techniques have developed in parallel
-with analytic frameworks.
-Most numerical techniques proposed for QOC are indirect shooting
-methods where an initial guess for the control trajectory is forward-simulated,
-e.g. under the time-dependent Schroedinger equation,
-and first-order gradients are computed for a functional evaluated
-on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
-  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
-Most techniques formulate the
-problem as unconstrained or use
-projective gradient methods to enforce constraints
-\cite{machnes2015gradient}.
-The latter approach relies on constructing
-bases for constraint spaces--which is not always feasible--and
-may hinder convergence.
-Methods for mitigating decoherence using numerical
-techniques have employed master equation dynamics (citation needed),
-Monte Carlo style trajectories \cite{abdelhafez2019gradient},
-and proposals have been made to simulate multiple trajectories
-\cite{reinhold2019controlling, rembold2020introduction}.
-
-In this work, we employ state of the art
-trajectory optimization methods
-to formulate the quantum
-optimal control problem as a constrained optimization problem, allowing
-us to enforce arbitrary experimental constraints.
-Additionally,
-we introduce methods for constructing gates that mitigate longitudinal
-relaxation and are robust to experimental errors, and therefore
-pure dephasing, the dominant barriers to experimentally
-realizing practical quantum computing.
-
-As an example, we study the quantum optimal control problem on
-the fluxonium qubit presented in \cite{zhang2020universal}.
-%%TODO: Fluxonium qubit is exciting and computationally feasible,
-%%but our methods extend to arbitrary platforms.
-We describe experimentally realistic constraints and map them to
-the trajectory optimization framework. Next, we
-outline a method for making the optimization longitudinal
-relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set.
-We present two methods for achieving robustness to pure dephasing
-and compare them to existing dynamic decoupling methods. We find that
-the methods we employ are able to produce high fidelity gates when
-subject to parameter spreads of $1\%$.
-
-%% S2
-\section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
-\label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
-In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
-algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
-
-(AL-iLQR Problem Statement) The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
-\begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-\end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
-
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-  \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
-  \end{aligned}
-\end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
-\end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
-
-%% S3
-\section{QOC on the Fluxonium}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
-\end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
-
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
-The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
-
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
-
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is computed after
-    each gate appliaction. The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
-  }
-\end{figure*}
-
-%% S4
-\section{Longitudinal Relaxation Awareness}
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms,
-and it is advantageous to maximize it by tuning the controls.
-In the fluxonium, the longitudinal relaxation time varies with
-the amplitude $a$. To completely simulate
-the effect of longitudinal relaxation on the gate error requires
-propagating density matrices of size $n \times n$ under Master equation
-dynamics rather than state vectors of size $n$ under the TDSE dynamics.
-We avoid this increase in computational complexity by
-penalizing the probability that the qubit decays as a result of
-longitudinal relaxation. This modification only
-increases the size of the augmented state vector by a single real number
-and allows us to effectively mitigate the error arising from longitudinal
-relaxation. Additionally, for a constant $T_{1}$ time, a shorter gate duration
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
-the one we study here.
-
-The longitudinal relaxation probability is given by
-\begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
-\end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-Calculating $T_{1}$ directly from theoretical
-considerations requires many high-rank
-eigendecompositions, which
-is computationally expensive. Additionally,
-$T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation.
-
-We allow the optimizer to tune the gate duration by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
-integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
-
-The $T_{1}$ optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate probability of longitudinal relaxation by a factor of $5$ over
-their analytic counterparts and achieve a commensurate reduction in their single gate error,
-see Appendix. The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have put on the
-device, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
-    points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency at multiples of the analytic gate duration for the analytic
-    and 1\textsuperscript{st}-order derivative methods.
-  }
-\end{figure*}
-
-%% S5
-\section{Robustness to Static Parameter Deviations}
-We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the model we use in optimization,
-the Hamiltonian, leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
-we study mitigating errors arising from the finite measurement
-precision of the qubit frequency $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$.
-We consider three flavors of robust control
-techniques. The first is the sampling method, which has
-been preposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling}. In the
-sampling method, multiple trajectories are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
-transformation to accurately propagate a distribution
-representing the evolving state as a result of a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. This scheme
-uses derivative information about how the state
-evolution is affected by the deviant parameter to
-mitigate trajectory deviations.
-
-In the sampling method we propagate additional sample states
-in the augmented state vector that follow deviant dynamics
-and penalize their differences from the target state.
-The states evolve under the deviant
-Hamiltonian where the parameter of interest is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The infidelities of the samples and the target
-state are penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
-resulting in eight additional states added to the augmented state vector.
-The four initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-This results in $O(n^{2})$ states added to the augmented
-state vector where $n$ is both the dimension of the Hilbert space
-and the size of the state vector.
-
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}\hspace{0.025\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is computed
-    after each gate application.
-  }
-\end{figure*}
-
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of $\ket{\psi}$. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states
-where $d$ is the number of deviant parameters. Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. This results
-in $O(n^{3})$ states added to the augmented state vector.
-The costs at each knot point are the infidelities between the samples
-and the target state, as in the sampling method.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
-derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For $d$ sensitivity parameters and $N$ knot points
-this requires $O(d N^2)$ matrix multiplications. Instead,
-we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting
-in $O(d N)$ matrix multiplications. For the piece-wise constant control
-scheme we employ here, $N$ is the number of control parameters. Typically
-$d \ll N$ so this is a good tradeoff. As an example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
-
-To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$. We compare the numerical methods
-to an analytically derived $Z/2$ gate which corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. It is linearly sensitive to qubit frequency detunings,
-the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
-qubit frequency detuning, it achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-sufficient for quantum error correction.
-
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
-The ability to perform $Z$ rotations in arbitrary times
-is critical for operating multi-qubit experiments in the lab frame.
-These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-
-The derivative methods also converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-They achieve super-linear scaling in their gate error as
-a function of the gate duration which approaches zero at the larmor
-period $1 / f_{q}$. We observe that on a range of problems with small parameter detunings,
-the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
-derivative method. We believe this is due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
-
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
-
-%% S6
-\section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
-take advantage of the low frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
-
-We compare the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
-flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-
-We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
-gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
-
-%% S7
-\section{Conclusion}
-We have proposed some schemes and they work well.
-
-%% ACK
-\begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
-\end{acknowledgments}
-
-\appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
-\section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
-\begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
-For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
-and $L_{\uparrow} = \sigma^{+}/2$,
-$L_{\downarrow} = \sigma^{-}/2$
-are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
-$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
-$L_{2} = (I - \sigma_{z})/2$.
-
-%% T1
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & Analytic & QOC & & Analytic & QOC & \\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
-    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% AD
 \section{Derivative Method}
@@
 \end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
-
-\bibliography{refs}
-
-\end{document}

### 2020-11-03 paper/main.tex
separated sections by file

--- 2020-10-26 paper/main.tex
@@
 
-%% ABSTRACT
-\begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
-  The ability to engineer high fidelity operations on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
-  but they require exquisite calibration to be performant.
-  In this work we employ robust trajectory optimization techniques
-  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
-  takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by transforming
-  higher-order differential equations into a set of coupled first-order equations.
-  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
-  the Lindblad master equation, which increases the computational complexity
-  of the optimization algorithm.
-  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
-  for the gate set of a fluxonium qubit
-  using a computationally efficient metric in optimization and time-optimal control.
-\end{abstract}
+\input{0abstract.tex}
 
@@
 
-%% S1
-\section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
-The field of quantum optimal control (QOC) is concerned
-with efficiently and accurately manipulating quantum systems.
-Early QOC techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-qubits \cite{heeres2017implementing,
-  leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}.
-In quantum control theory, as in classical control theory,
-optimization is performed
-to minimize an objective function. Relevant for the field
-of quantum computation is the task of achieving high fidelity quantum gates.
-In this application, the objective function includes contributions from the
-gate error associated with the state transfer of the system as well as contributions
-due to experimental constraints, such as the control amplitude.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-
-Many analytic and numerical techniques have been developed to control
-quantum systems. Analytic methods rely on solving the time-dependent
-Schroedinger equation analytically or deriving equations of motion
-from a suitable Lagrangian \cite{zhang2020universal, huang2020engineering, han2020experimental,
-  xu2020nonadiabatic, carlini2005quantum}. Analytic methods
-have been presented to address decoherence in quantum systems.
-Considering the dynamical and geometric phases of state
-evolution has led to methods for achieving
-robustness to dynamical errors and mitigating pure dephasing
-\cite{xu2020nonadiabatic, han2020experimental, merrill2014progress}.
-Additionally, suitable choices of bases have been proposed that
-allow the experimentalist to infer tradeoffs in longitudinal relaxation
-and pure dephasing coherence times \cite{huang2020engineering}.
-To date, however, no analytical method has been presented
-to construct a gate while satisfying arbitrary experimental
-constraints. Furthermore, most analytic methods have only
-been demonstrated in the two-level approximation and become
-cumbersome to solve for larger Hilbert space sizes.
-
-Numerical quantum control techniques have developed in parallel
-with analytic frameworks.
-Most numerical techniques proposed for QOC are indirect shooting
-methods where an initial guess for the control trajectory is forward-simulated,
-e.g. under the time-dependent Schroedinger equation,
-and first-order gradients are computed for a functional evaluated
-on the state of the evolved system \cite{leung2017speedup,  goerz2019krotov, doria2011optimal,
-  abdelhafez2019gradient, machnes2015gradient, leng2019robust}.
-Most techniques formulate the
-problem as unconstrained or use
-projective gradient methods to enforce constraints
-\cite{machnes2015gradient}.
-The latter approach relies on constructing
-bases for constraint spaces--which is not always feasible--and
-may hinder convergence.
-Methods for mitigating decoherence using numerical
-techniques have employed master equation dynamics (citation needed),
-Monte Carlo style trajectories \cite{abdelhafez2019gradient},
-and proposals have been made to simulate multiple trajectories
-\cite{reinhold2019controlling, rembold2020introduction}.
-
-In this work, we employ state of the art
-trajectory optimization methods
-to formulate the quantum
-optimal control problem as a constrained optimization problem, allowing
-us to enforce arbitrary experimental constraints.
-Additionally,
-we introduce methods for constructing gates that mitigate longitudinal
-relaxation and are robust to experimental errors, and therefore
-pure dephasing, the dominant barriers to experimentally
-realizing practical quantum computing.
-
-As an example, we study the quantum optimal control problem on
-the fluxonium qubit presented in \cite{zhang2020universal}.
-%%TODO: Fluxonium qubit is exciting and computationally feasible,
-%%but our methods extend to arbitrary platforms.
-We describe experimentally realistic constraints and map them to
-the trajectory optimization framework. Next, we
-outline a method for making the optimization longitudinal
-relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
-over the baseline gate set.
-We present two methods for achieving robustness to pure dephasing
-and compare them to existing dynamic decoupling methods. We find that
-the methods we employ are able to produce high fidelity gates when
-subject to parameter spreads of $1\%$.
-
-%% S2
-\section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
-\label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
-act on the system.
-
-Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical integration techniques for the TDSE include
-approximating unitary propagators as well as explicit integration methods,
-such as Runge-Kutta, of the form
-$\ket{\psi_{k + 1}} \approx \ket{\psi_{k}} + \frac{d}{dt} \ket{\psi_{k}} \cdot \Delta t_{k}$
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
-In the simplest case the functional is
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
-algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
-
-(AL-iLQR Problem Statement) The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
-\begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
-\end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
-
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-  \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
-  \end{aligned}
-\end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
-\end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
-
-%% S3
-\section{QOC on the Fluxonium}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
-\end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
-
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
-The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
-
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-via the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
-
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-  \end{subfigure}
-  \caption{
-    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is computed after
-    each gate appliaction. The $T_{1}$ optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
-  }
-\end{figure*}
-
-%% S4
-\section{Longitudinal Relaxation Awareness}
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms,
-and it is advantageous to maximize it by tuning the controls.
-In the fluxonium, the longitudinal relaxation time varies with
-the amplitude $a$. To completely simulate
-the effect of longitudinal relaxation on the gate error requires
-propagating density matrices of size $n \times n$ under Master equation
-dynamics rather than state vectors of size $n$ under the TDSE dynamics.
-We avoid this increase in computational complexity by
-penalizing the probability that the qubit decays as a result of
-longitudinal relaxation. This modification only
-increases the size of the augmented state vector by a single real number
-and allows us to effectively mitigate the error arising from longitudinal
-relaxation. Additionally, for a constant $T_{1}$ time, a shorter gate duration
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
-the one we study here.
-
-The longitudinal relaxation probability is given by
-\begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
-\end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
-Calculating $T_{1}$ directly from theoretical
-considerations requires many high-rank
-eigendecompositions, which
-is computationally expensive. Additionally,
-$T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data increases
-the fridge truth of the simulation.
-
-We allow the optimizer to tune the gate duration by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
-integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
-
-The $T_{1}$ optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All optimized gates reduce
-their single gate probability of longitudinal relaxation by a factor of $5$ over
-their analytic counterparts and achieve a commensurate reduction in their single gate error,
-see Appendix. The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have put on the
-device, and do not represent a fundamental limit to the optimization methods we have
-employed.
-
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
-    points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency at multiples of the analytic gate duration for the analytic
-    and 1\textsuperscript{st}-order derivative methods.
-  }
-\end{figure*}
-
-%% S5
-\section{Robustness to Static Parameter Deviations}
-We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
-However, the device typically deviates from the model we use in optimization,
-the Hamiltonian, leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
-we study mitigating errors arising from the finite measurement
-precision of the qubit frequency $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$.
-We consider three flavors of robust control
-techniques. The first is the sampling method, which has
-been preposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling}. In the
-sampling method, multiple trajectories are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
-transformation to accurately propagate a distribution
-representing the evolving state as a result of a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. This scheme
-uses derivative information about how the state
-evolution is affected by the deviant parameter to
-mitigate trajectory deviations.
-
-In the sampling method we propagate additional sample states
-in the augmented state vector that follow deviant dynamics
-and penalize their differences from the target state.
-The states evolve under the deviant
-Hamiltonian where the parameter of interest is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The infidelities of the samples and the target
-state are penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-We sample two trajectories ($\ket{\psi^{\pm}}$) for each of four initial states,
-resulting in eight additional states added to the augmented state vector.
-The four initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-This results in $O(n^{2})$ states added to the augmented
-state vector where $n$ is both the dimension of the Hilbert space
-and the size of the state vector.
-
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-  \end{subfigure}\hspace{0.025\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-  \end{subfigure}
-
-  \caption{
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
-    8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is computed
-    after each gate application.
-  }
-\end{figure*}
-
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of $\ket{\psi}$. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states
-where $d$ is the number of deviant parameters. Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. This results
-in $O(n^{3})$ states added to the augmented state vector.
-The costs at each knot point are the infidelities between the samples
-and the target state, as in the sampling method.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
-derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For $d$ sensitivity parameters and $N$ knot points
-this requires $O(d N^2)$ matrix multiplications. Instead,
-we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting
-in $O(d N)$ matrix multiplications. For the piece-wise constant control
-scheme we employ here, $N$ is the number of control parameters. Typically
-$d \ll N$ so this is a good tradeoff. As an example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
-
-To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$. We compare the numerical methods
-to an analytically derived $Z/2$ gate which corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. It is linearly sensitive to qubit frequency detunings,
-the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
-qubit frequency detuning, it achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-sufficient for quantum error correction.
-
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
-The ability to perform $Z$ rotations in arbitrary times
-is critical for operating multi-qubit experiments in the lab frame.
-These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-
-The derivative methods also converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-They achieve super-linear scaling in their gate error as
-a function of the gate duration which approaches zero at the larmor
-period $1 / f_{q}$. We observe that on a range of problems with small parameter detunings,
-the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
-derivative method. We believe this is due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
-
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
-It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
-
-%% S6
-\section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
-take advantage of the low frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
-
-We compare the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
-flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-
-We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
-gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
-
-%% S7
-\section{Conclusion}
-We have proposed some schemes and they work well.
-
-%% ACK
-\begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
-\end{acknowledgments}
+\input{1introduction.tex}
+\input{2background.tex}
+\input{3fluxonium.tex}
+\input{4longitude.tex}
+\input{5static.tex}
+\input{6stochastic.tex}
+\input{7conclusion.tex}
+\input{8ack.tex}
 
 \appendix
-%% AA
-\section{Ricatti Recursion}
-This will give the reader unfamiliar with trajectory
-optimization intuition for how the trajectory optimization
-update scheme works and why it is better than
-a more naive method.
-
-%% AB
-\section{Experiment}
-We measure $T_{1}$ using the standard experiment
-and $T_{2}$ using te Ramsey experiment. We fit with splines
-and the data looks like fig. 3 in Helin's paper \cite{zhang2020universal}.
-We measure $f_{q}$ and $\sigma_{f_{q}}$ using X method.
-
-%% AC
-\section{Dissipation}
-We model dissipation using the Lindblad master
-equation. 
-\begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho] + \sum_{i = 1}^{N^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger} - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $N = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
-For longitudinal relaxation $\gamma_{1} = T_{1}^{-1} = T_{1, \uparrow}^{-1} + T_{1, \downarrow}^{-1}$
-and $L_{\uparrow} = \sigma^{+}/2$,
-$L_{\downarrow} = \sigma^{-}/2$
-are the ladder operators $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$. For pure dephasing
-$\gamma_{2} = T_{2}^{-1} = (2 T_{1})^{-1} + T_{\phi}^{-1}$ and
-$L_{2} = (I - \sigma_{z})/2$.
-
-%% T1
-\begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & Analytic & QOC & & Analytic & QOC & \\
-    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
-    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
-    \hline
-    Z/2 & 5.745  & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253  & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
-  \end{tabular}
-  \caption{Probability of longitudinal relaxation for each gate
-    evaluated at the gate's duration.}
-\end{table}
-
-%% AD
-\section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
-
-%% T2
-\begin{table}
-  \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
-    \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
-  \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
-\end{table}
-
-%% AE
-\section{Complex Tensor Handling}
-We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-because the software we use does not support complex numbers yet.
+\input{a1experiment.tex}
+\input{a2dissipation.tex}
+\input{a3derivative.tex}
+\input{a4ctensor.tex}
 

### 2020-11-04 paper/1introduction.tex
updated many, added pdf

@@
-%% S1
 \section{Introduction}
-%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
-%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
-%% P1 - general field, lots of citation. cite reviews, original qoc.
-%% P2 - exhaustively find all papers, what people have done to address the issues
-%% P3 - what we do is. maybe add a rant about how
-%% QOC is classical control theory but we've ignored that field
-%% P4 - outline
 Quantum optimal control (QOC) techniques are a class of optimization
@@
 diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}. QOC techniques aim to control the system
-such that a set of objectives are optimally satisfied.
-For quantum computation relevant objectives include achieving high fidelity
-gates and adhering to experimental constraints.
+\cite{sorensen2018quantum}. For quantum computation
+QOC techniques are employed to achieve high fidelity gates
+while adhering to experimental constraints.
 The decision variables of the optimization problem are the time-dependent control
@@
 In this work we employ robust control techniques to mitigate
-releastic decoherence and systematic errors that arise when controlling
+realistic system parameter deviations and decoherence that arise when controlling
 a superconducting fluxonium qubit.
@@
 resultant quantum state trajectory to be insensitive to experimental errors
-such as parameter drift, noise, and finite control resolution. Controls that produce
-quantum state trajectories that are insensitive to these effects are said to be robust.
-In line with this objective, recent work has sought to mitigate systematic errors
-and decoherence present in the experiment.
+such as parameter drift, noise, and finite control resolution. Quantum state
+trajectories that are insensitive to these effects are said to be robust.
+In line with this objective, recent work has sought to mitigate decoherence
+and errors due to parameter deviations.
 Considering the dynamical and geometric phases of quantum state
-trajectories has led to analytic methods for achieving
-robustness to systematic errors and mitigating pure dephasing
+trajectories has led to analytic methods for mitigating
+errors due to parameter deviations and pure dephasing
 \cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
@@
 
-In this work we employ state of the art trajectory optimization techniques
-to mitigate systematic errors and decoherence. We frame the QOC problem as an
-open loop optimization problem, optimizing the controls to make the
-quantum state trajectory robust to systematic errors and decoherence.
-We consider three robust trajectory optimization techniques.
+In this work we develop robust trajectory optimization techniques to mitigate
+errors arising from longitudinal relaxation, system parameter deviations, and 1/$f$ flux noise
+on the fluxonium qubit. We perform these optimization using the ALTRO
+method \cite{howell2019altro} which combines the iterative, indirect shooting
+flavor of methods such as GOAT \cite{machnes2015gradient}, GRAPE
+\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}
+with the augmented Lagrangian method. The augmented Lagrangian method
+allows us to enforce arbitrary, simultaneous constraints and achieve
+fast convergence without being restricted to the constraint
+manifold. To mitigate errors due to longitudinal
+relaxation, we perform time-optimal control and
+encode the dependence of longitudinal relaxation
+on the controls in an efficient optimization objective that does
+not pay the increased computational cost of integrating a master equation.
+Additionally, we consider three robust
+robust trajectory opitmization techniques to make the
+quantum state trajectory robust to system parameter deviations.
 The first is the sampling method, which has been applied
@@
 We make this method efficient by employing mixed-mode differentiation.
-Additionally, we take advantage of the known dependence of the
-system's controls on longitudinal relaxation. We encode
-the decoherence in an efficient optimization objective that does
-not pay the increased computational cost of integrating a master equation.
-We perform these optimizations using the ALTRO method \cite{howell2019altro}.
-Many popular numerical techniques
-for QOC use gradient-based control update procedures, which are readily amenable to control filtering
-or constraint manifold projection techniques to enforce constraints \cite{leung2017speedup,
-  goer2019krotov, abdelhafez2019gradient, machnes2015gradient}.
-The ALTRO method we employ combines
-an iterative indirect shooting method with the augmented Lagrangian
-method. The augmented Lagrangian method
-allows us to enforce arbitrary, simultaneous constraints and achieve
-fast convergence without being restricted to the constraint
-manifold.
 
@@
 on the heavy fluxonium qubit and against the assocaited analytic techniques
-presented in \cite{zhang2020universal}. The fluxonium is an exciting
-platform for quantum computation due to its long longitudinal relaxation
-and pure dephasing times at the flux frustation point. This system is also
-efficient to simulate due to its accurate two-level approximation.
-We emphasize that the numerical techniques we employ extend to arbitrary quantum systems,
-not only the one we study here. We describe experimentally realistic constraints for this system
-and map them to the ALTRO framework. Next, we
+presented in \cite{zhang2020universal}. This paper is outlined as follows.
+First, we introduce the ALTRO method in the context of QOC.
+We describe experimentally
+realistic constraints for the fluxonium and
+map them to the ALTRO method. Then, we
 outline a method for making the optimization aware of longitudinal
-relaxation. We achieve a factor of 5 decrease in single gate errors due to
-longitudinal relaxation over the analytic gate set. Next we outline three methods for achieving
-robustness to static parameter deviations. We find that we are able to
-decrease the gate error arising from static parameter deviations super-linearly
-in the gate duration, achieving orders of magntidue reductions in single gate errors.
-We also find that the numerical techniques we employ are able to produce
-fixed phase gates in arbitrary times, a method that could be used
-to control multi-qubit systems in the lab frame. Finally,
-we employ the robust control techniques to mitigate time-varying
-magnetic flux noise. We achieve a two order of magnitude improvement
-in gate errors for extended computations, an improvement critical
-for noisy, intermediate-scale quantum (NISQ) applications.
+relaxation. Next, we outline three methods for achieving
+robustness to static parameter deviations. Finally,
+we employ the robust control techniques to mitigate magnetic flux noise.

### 2020-11-04 paper/a1dissipation.tex
updated many, added pdf

\section{Longitudinal Relaxation}
The probability of longitudinal relaxation metric and gate error due to longitudinal
relaxation metric are compared in Table I for the experiment shown in Figure 1c.
The relative performance between the analytic and numerical techniques we study
is similar across the two metrics.

\begin{table}[h!]
  \begin{tabular}{c | c | c}
    Gate & $P_{1\textrm{A}} / P_{1\textrm{N}}$ & GE$_{\textrm{A}}$ / GE$_{\textrm{N}}$\\
    \hline
    Z/2 & 5.745/1.149 = 5.000 & 1.776/0.371 = 4.787\\
    Y/2 & 5.253/1.157 = 4.540 & 1.539/0.370 = 4.159\\
    X/2 & 16.251/2.660 = 6.109 & 5.347/0.863 = 6.196\\
  \end{tabular}
  \caption{
    Single gate probability of longitudinal relaxation
    ratios and single gate error due to longitudinal relaxation
    ratios for the analytic and numerical methods. Both probabilities
    and gate errors are shown in units of $10^{-5}$.
  }
\end{table}

To compute the gate error due to longitudinal relaxation,
we require the final state of the quantum system subject
to longitudinal relaxation. To obtain the final state we employ
the Lindblad master equation. This equation takes the form
\begin{equation}
  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho]
  + \sum_{i = 1}^{n^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger}
  - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
\end{equation}
where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $n = \textrm{dim}(\mathcal{H})$,
and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
For longitudinal relaxation $\gamma_{\uparrow} = T_{1, \uparrow}^{-1}$,
$\gamma_{\downarrow} = T_{1, \downarrow}^{-1}$,
$L_{\uparrow} = \sigma^{+}/2$, and
$L_{\downarrow} = \sigma^{-}/2$ where $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$.
Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained from the spline shown in Figure 1b.
The $T_{1}$ values in this spline are obtained experimentally by driving the qubit at the desired flux amplitude
and monitoring the resultant decay. For more details consult \cite{zhang2020universal}.

Exponential integrators can be employed to integrate the Lindblad master equation
using the Vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018}
\begin{equation}
  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}} \textrm{vec}({\rho})
\end{equation}
\begin{equation}
  \begin{aligned}
    \hat{\mathcal{L}} &= -i(I \otimes H - H^{T} \otimes I)\\
    &+ \sum_{i = 1}^{n^{2} - 1} \gamma_{i}
    (L_{i}^{*} \otimes L_{i} - \frac{1}{2} (I \otimes L_{i}^{\dagger}L_{i}
    - L_{i}^{T}L_{i}^{*} \otimes I))
  \end{aligned}
\end{equation}
where $\rho = \sum_{i, j} \alpha_{i, j} \ket{i}\bra{j}$
and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{i, j} \ket{i} \otimes \ket{j}$.
We use zero-order hold on the controls so the integration is exact
$\textrm{vec}(\rho_{k + 1}) = \exp(\hat{\mathcal{L}} \Delta t_{k}) \textrm{vec}(\rho_{k})$.
This isomorphism transforms $(n^{2} \times n^{2}) \times (n^{2} \times n^{2})$
matrix-matrix multiplications to $(n^{4} \times n^{4}) \times n^{4}$ matrix-vector
multiplications. For small $n$ and zero-order held controls, we find that it is
faster to use an exponential integrator on the vectorized equation than to perform
Runge-Kutta on the unvectorized equation. The latter requires decreasing the integration
time step to maintain accuracy, resulting in more knot points.

### 2020-11-04 paper/a2unscented.tex
updated many, added pdf

\section{Unscented Transformation}
In this section we outline the full unscented sampling procedure.
First we remark that ALTRO does not natively support complex numbers, so our optimization
is performed under the isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$.
The unscented sampling method concerns a state $\psi \in \mathbb{R}^{2n}$ with dynamics
$\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$ with deviant parameters $\lambda \in \mathbb{R}^{d}$.
The nominal initial state and parameters are given $\psi_{0}$, $\lambda_{0}$ with associated
positive-definite covariance matrices $P \in \mathbb{R}_{++}^{n \times n}$,
$L \in \mathbb{R}_{++}^{d \times d}$. We set $\lambda_{0} = 0$ for convenience and modify
any offset in the dynamics function. Physically, $P$ can represent the state preparation error,
but is typically chosen to be non-zero even if there is a small state preparation error.
Using these initial values we generate
\begin{equation}
  \begin{bmatrix} \Psi_{0} \\ \Lambda_{0} \end{bmatrix} = \mathcal{N}(
  \begin{bmatrix} \psi_{0} \\ 0\end{bmatrix}, \begin{bmatrix} P & 0\\ 0 & L\end{bmatrix})
\end{equation}
Then the initial states are propagated in the dynamics function
\begin{equation}
  \tilde{\Psi_{1}} = f(\Psi_{0}, \Lambda_{0})
\end{equation}
We then compute the mean and covariance of the sigma points.

The desired initial state is given $x_{0}$ with a known covariance $P_{0} \in \mathbb{R}^{n \times n}$.
These two 
which is used to generate the initial samples 
In the case of QOC $x = \ket{\psi} \in $

We will consider a single quantum state $\ket{\psi}$ and associated
deviant parameters $\vec{\omega}$. $\ket{\psi_{0}}$
with associated samples $\Psi_{0} = \{\ket{\psi_{0}^{i}}\}$, $i \in \{1, \dots, 2(n + d)\}$.
The initial samples are generated using covariance matrices for the

### 2020-11-04 paper/a3derivative.tex
updated many, added pdf

@@
 \section{Derivative Method}
-To obtain the dynamics for the derivative of the state $\partial_{x}^{l} \ket{\psi(t)}$
-we differentiate the TDSE dynamics \ref{eq:tdse} with respect to the parameter of interest
-($x$). Using the fluxonium hamiltonian \ref{eq:hamiltonian} we obtain the derivative of the
-state with respect to the qubit frequency ($f_{q}$) and the flux amplitude ($a$).
-Here we present the dynamics for the second derivative of the state with respect to the
-qubit frequency. The case is analogous for the flux amplitude. Both $H$ and $\ket{\psi}$ are functions
-of $f_{q}$, $a$, and $t$, but we omit the explicit dependence in notation for
-brevity.
-\begin{equation}
-  \begin{aligned}
-      \partial_{f_{q}} \partial_{t} \ket{\psi} &= \partial_{f_{q}} H \ket{\psi}\\
-      &= (\partial_{f_{q}} H) \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})\\
-      &= \frac{\sigma_{z}}{2} \ket{\psi} + H (\partial_{f_{q}} \ket{\psi})
-  \end{aligned}
-\end{equation}
-\begin{equation}
-  \begin{aligned}
-    \partial_{f_{q}}^{2} \partial_{t} \ket{\psi} &= \partial_{f_{q}} (\partial_{f_{q}} H \ket{\psi})\\
-    &= (\partial_{f_{q}}^{2} H) \ket{\psi} + 2 (\partial_{f_{q}} H)(\partial_{f_{q}} \ket{\psi})\\
-    &\quad + H (\partial_{f_{q}}^{2} \ket{\psi})\\
-    &= \sigma_{z} (\partial_{f_{q}} \ket{\psi}) + H (\partial_{f_{q}}^{2} \ket{\psi})
-  \end{aligned}
-\end{equation}
-The augmented state vector carries $\partial_{f_{q}}^{l} \ket{\psi}$
-which appears explicitly in its own dynamics. Due to the dependence of $H$ on $f_{q}$ and $a$,
-the $l$\textsuperscript{th} state derivative is coupled to the
-$l - 1$\textsuperscript{th} state derivative.
+We believe the 1\textsuperscript{st}-order derivative method tends
+to outperform the 2\textsuperscript{nd}-order
+derivative method in the small, static detuning regime, see Figure 2b,
+because second-order contributions to the gate error are small.
+The norm of the state derivatives for the 1\textsuperscript{st}- and 2\textsuperscript{nd}-order
+derivative methods are provided in Table 2 for the solutions of the optimization outlined
+in section V at the gate duration $t_{N} = 38$ns.
 
-%% T2
-\begin{table}
+%%TODO: the values in this table are not up to date until this comment is removed
+\begin{table}[h]
   \begin{tabular}{c | c | c}
-    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
+    Method & ${\lvert \braket{\partial_{f_{q}} \psi_{N} | \partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
+    & ${\lvert \braket{\partial^{2}_{f_{q}} \psi_{N} | \partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
     \hline
@@
 \end{table}
+
+The dynamics for the derivative methods can be integrated efficiently using exponential integrators.
+General exponential integrators break the dynamics into a linear term and a non-linear term. For example, consider
+integrating the dynamics of the first state derivative $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
+H \ket{\partial_{\lambda} \psi} + (\partial_{\lambda} H) \ket{\psi}$ in units of $i\hbar = 1$.
+The linear term is $L = H$ and the non-linear term is $N = (\partial_{\lambda} H) \ket{\psi}$.
+With zero-order hold on the controls the exact propagation is
+\begin{equation}
+  \begin{aligned}
+    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(L_{k} \Delta t_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
+    &+ \int_{0}^{\Delta t_{k}} \exp(L_{k}(\Delta t_{k} - t^{'})) N(t_{k} + t^{'}) dt^{'}\\
+  \end{aligned}
+\end{equation}
+General exponential integrators proceed by breaking the integral into a discrete sum, similar to the procedure
+for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler method \cite{berland2005solving}
+\begin{equation}
+  \begin{aligned}
+    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(L_{k} \Delta t_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
+    &+ \exp(L_{k} \Delta t_{k}) N_{k}\\
+  \end{aligned}
+\end{equation}
+This method provides a good tradeoff between accuracy and efficiency, requiring one unique matrix
+exponential computation per stage. Exact integration accuracy is not of the utmost importance because the
+state derivatives guide the robust optimization, and do not correspond to experimental parameters
+which must be realized with high accuracy.

### 2020-11-06 paper/1introduction.tex
misc updates

@@
 In this work we develop robust trajectory optimization techniques to mitigate
-errors arising from longitudinal relaxation, system parameter deviations, and 1/$f$ flux noise
-on the fluxonium qubit. We perform these optimization using the ALTRO
+errors arising from longitudinal relaxation, parameter deviations, and 1/$f$ flux noise
+on the fluxonium qubit presented in \cite{zhang2020universal}. We perform these optimization using the ALTRO
 method \cite{howell2019altro} which combines the iterative, indirect shooting
@@
 robust trajectory opitmization techniques to make the
-quantum state trajectory robust to system parameter deviations.
+quantum state trajectory robust to parameter deviations.
 The first is the sampling method, which has been applied
@@
 
-We compare the performance of these numerical techniques
-on the heavy fluxonium qubit and against the assocaited analytic techniques
-presented in \cite{zhang2020universal}. This paper is outlined as follows.
+This paper is organized as follows.
 First, we introduce the ALTRO method in the context of QOC.
-We describe experimentally
-realistic constraints for the fluxonium and
+We describe realistic constraints for the fluxonium and
 map them to the ALTRO method. Then, we
@@
 robustness to static parameter deviations. Finally,
-we employ the robust control techniques to mitigate magnetic flux noise.
+we employ the robust control techniques to mitigate 1/$f$ flux noise.

### 2020-11-06 paper/4longitude.tex
misc updates

@@
 Computing the gate error due to longitudinal relaxation
-requires propagating density matrices of size $n \times n$ under Master equation
+requires propagating density matrices of size $n \times n$ under master equation
 dynamics, rather than state vectors of size $n$ under the TDSE dynamics.
@@
 beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All numerically optimized gates reduce
+to higher amplitudes and therefore higher $T_{1}$ times. All numerical gates reduce
 their single gate errors by a factor of $5$ over
@@
 of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The optimized $Z/2$ and $Y/2$ gates perform simlarly in
+generated initial states. The numerical $Z/2$ and $Y/2$ gates perform simlarly in
 the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The optimized $X/2$ gate
+over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The numerical $X/2$ gate
 achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-Both the analytic and optimized gates attain single gate errors sufficient for
+Both the analytic and numerical gates attain single gate errors sufficient for
 quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the optimized gates for extend computations
+The low gate errors achieved by the numerical gates for extend computations
 are critical for noisy, intermediate-scale quantum (NISQ) applications.

### 2020-11-06 paper/7conclusion.tex
misc updates

@@
-%% S7
 \section{Conclusion}
@@
 The numerical techniques we have studied here can be used to perform phase gates in arbitrary times,
-which will be critical for synchronizing multi-qubit systems. Our
-techniques significantly improve the state-of-the-art for controlling quantum systems subject
-to systematic errors and decoherence, and will be used in efforts to lower gate errors
-for fault-tolerant quantum computing and NISQ applications.
+which will be critical for synchronizing multi-qubit systems. Additionally,
+interleaving the error models with existing
+characterization methods will improve their effectiveness in experiments.
+These techniques will be employed to achieve the low gate errors
+required for fault-tolerant quantum computing applications.

### 2020-11-06 paper/a1dissipation.tex
misc updates

@@
 \end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $n = \textrm{dim}(\mathcal{H})$,
+Here $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $n = \textrm{dim}(\mathcal{H})$,
 and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
@@
 \end{equation}
-where $\rho = \sum_{i, j} \alpha_{i, j} \ket{i}\bra{j}$
+Here $\rho = \sum_{i, j} \alpha_{i, j} \ket{i}\bra{j}$
 and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{i, j} \ket{i} \otimes \ket{j}$.
 We use zero-order hold on the controls so the integration is exact
-$\textrm{vec}(\rho_{k + 1}) = \exp(\hat{\mathcal{L}} \Delta t_{k}) \textrm{vec}(\rho_{k})$.
+$\textrm{vec}(\rho_{k + 1}) = \exp(\Delta t_{k} \hat{\mathcal{L}}_{k}) \textrm{vec}(\rho_{k})$.
 This isomorphism transforms $(n^{2} \times n^{2}) \times (n^{2} \times n^{2})$

### 2020-11-06 paper/a2unscented.tex
misc updates

@@
 In this section we outline the full unscented sampling procedure.
-First we remark that ALTRO does not natively support complex numbers, so our optimization
-is performed under the isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$.
-The unscented sampling method concerns a state $\psi \in \mathbb{R}^{2n}$ with dynamics
-$\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$ with deviant parameters $\lambda \in \mathbb{R}^{d}$.
-The nominal initial state and parameters are given $\psi_{0}$, $\lambda_{0}$ with associated
-positive-definite covariance matrices $P \in \mathbb{R}_{++}^{n \times n}$,
-$L \in \mathbb{R}_{++}^{d \times d}$. We set $\lambda_{0} = 0$ for convenience and modify
-any offset in the dynamics function. Physically, $P$ can represent the state preparation error,
-but is typically chosen to be non-zero even if there is a small state preparation error.
-Using these initial values we generate
+First, we remark that ALTRO does not natively support complex numbers, so our optimization
+is performed under an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$.
 \begin{equation}
-  \begin{bmatrix} \Psi_{0} \\ \Lambda_{0} \end{bmatrix} = \mathcal{N}(
-  \begin{bmatrix} \psi_{0} \\ 0\end{bmatrix}, \begin{bmatrix} P & 0\\ 0 & L\end{bmatrix})
+  H \psi \cong \begin{pmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{pmatrix}
+  \begin{pmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{pmatrix}
 \end{equation}
-Then the initial states are propagated in the dynamics function
+
+The unscented sampling method concerns a quantum state $\psi \in \mathbb{R}^{2n}$ with
+deviant parameters $\lambda \in \mathbb{R}^{d}$ and dynamics $\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
+The nominal initial quantum state is given by $\psi_{0}$ with an associated
+positive-definite covariance matrix $P_{0} \in \mathbb{R}_{++}^{n \times n}$
+which describes the uncertainty in the initial state.
+$P_{0}$ is typically chosen to be non-zero even if the state preparation error is negligible.
+The deviant parameter
+is assumed to have zero mean and its distribution is given by the covariance matrix
+$L_{k} \in \mathbb{R}_{++}^{d \times d}$ at knot point $k$. The zero mean assumption
+is convenient for deriving the update procedure. A non-zero mean can be encoded
+in the dynamics $f$.
+
+The initial $4n + 2d$ sigma points and initial $4n + 2d$ deviant parameters are sampled
+from the initial distributions
 \begin{equation}
-  \tilde{\Psi_{1}} = f(\Psi_{0}, \Lambda_{0})
+  \begin{bmatrix} \Psi_{0}^{i} \\ \Lambda_{0}^{i} \end{bmatrix} =
+  \begin{bmatrix} \bar{\Psi}_{0} \\ 0\end{bmatrix}
+    \pm \beta \sqrt{\begin{bmatrix} P_{0} & 0\\ 0 & L_{0}\end{bmatrix}}^{\; i}
 \end{equation}
-We then compute the mean and covariance of the sigma points.
-
-The desired initial state is given $x_{0}$ with a known covariance $P_{0} \in \mathbb{R}^{n \times n}$.
-These two 
-which is used to generate the initial samples 
-In the case of QOC $x = \ket{\psi} \in $
-
-We will consider a single quantum state $\ket{\psi}$ and associated
-deviant parameters $\vec{\omega}$. $\ket{\psi_{0}}$
-with associated samples $\Psi_{0} = \{\ket{\psi_{0}^{i}}\}$, $i \in \{1, \dots, 2(n + d)\}$.
-The initial samples are generated using covariance matrices for the
+We have written $\bar{\Psi}_{0} = \psi_{0}$. $\beta$
+is a hyperparameter which controls the spacing of the covariance contour.
+The $\pm$ is understood to take $+$ for $ i \in \{1, \dots, 2n + d\}$ and $-$ for
+$i \in \{2n + d + 1, \dots, 4n + 2d\}$. We use the Cholesky factorization
+to compute the square root of the
+joint covariance matrix, though other methods
+such as the principal square root may be employed.
+The superscript indicates the $i$\textsuperscript{th} column of the
+lower triangular Cholesky factor.
+Then, the sigma points are normalized
+\begin{equation}
+  \Psi_{0}^{i} \gets \frac{\Psi_{0}^{i}}{\sqrt{\Psi_{0}^{i^{T}} \Psi_{0}^{i}}}
+\end{equation}
+The sigma points are propagated to the next knot point
+\begin{equation}
+  \Psi_{1} = f(\Psi_{0}, \Lambda_{0})
+\end{equation}
+The mean and covariance of the sigma points is computed
+\begin{align}
+  \bar{\Psi}_{1} &= \frac{1}{4n + 2d} \sum_{i = 1}^{4n + 2d} \Psi_{1}^{i}\\
+  P_{1} &= \frac{1}{\beta^{2}(4n + 2d)} \sum_{i = 1}^{4n + 2d}
+  (\Psi^{i}_{1} - \bar{\Psi}_{1})^{T}(\Psi^{i}_{1} - \bar{\Psi}_{1})
+\end{align}
+The sigma points are then resampled and propagated to the next knot point using eq B2, B3, and B4.

### 2020-11-12 paper/0abstract.tex
updated many, added pdf
misc updates

@@
+%% TITLE
+\title{Robust Control of a Fluxonium Qubit}
+
+\author{Thomas Propson}
+\email{tcpropson@uchicago.edu}
+\affiliation{
+  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
+}
+\affiliation{
+  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
+}
+\author{Brian Jackson}
+\author{Zac Manchester}
+\affiliation{
+  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
+}
+\author{David I. Schuster}
+\affiliation{
+  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
+}
+\affiliation{
+  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
+}
+\affiliation{
+  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
+}
+
+\date{\today}
+
 %% ABSTRACT
@@
   systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control (QOC) techniques have proven effective in experimentally
+  Quantum optimal control techniques have proven effective in experimentally
   realizing high fidelity gates, but they require exquisite calibration to be performant.
-  We apply robust trajectory optimization techniques to suppress gate errors arising from system
-  model deviations and noise.
+  We apply robust trajectory optimization techniques to supress gate errors arising from system
+  parameter deviations and noise.
   We propose a method that takes advantage of deviant parameter derivative information while maintaining
@@
   achieving orders of magnitude gate error reductions from our baseline gate set.
-  %% TODO: maybe add arbitrary phase bit
 \end{abstract}

### 2020-11-12 paper/2background.tex
updated many, added pdf
misc updates

@@
-%% S2
 \section{QOC + AL-iLQR}
@@
 we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
+state, and $u$ as the augmented controls. In practice the final state
+of the quantum system matching the target state
+is the most important metric, which is reflected in the weight
+bias $Q_{N} \sim N \cdot Q_{1}$ and $Q_{k} = Q_{1}$.
 
@@
 recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
+feedback law--known as the Ricatti recursion. The
 iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
@@
 updates, see \cite{howell2019altro}.
+
+%% TODO: include comment about control filtering and constraint manifold
+%% projection of GOAT, Krotov, GRAPE, etc.

### 2020-11-12 paper/4longitude.tex
misc updates

@@
 \begin{figure*}[ht]
+  \label{fig:longitude}
   \begin{subfigure}{.315\textwidth}
@@
 \begin{figure*}[ht]
+  \label{fig:static}
   \begin{subfigure}{.315\textwidth}
@@
 The numerically optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
+waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitude}.
 They extend their gate times
@@
 their analytic counterparts which is commensurate to their
-probability of longitudinal relaxation reductions, see Appendix.
+probability of longitudinal relaxation reductions, see Appendix \ref{appendix:longitude}.
 The gate error reported in this text is the infidelity

### 2020-11-12 paper/5static.tex
misc updates

@@
 \begin{figure*}[ht]
+  \label{fig:stochastic}
   \begin{subfigure}{.4\textwidth}
@@
 covariance contour at each knot point. A detailed update procedure is given
-in the appendix. Each initial state from the operator basis
+in Appendix \ref{appendix:unscented}. Each initial state from the operator basis
 is represented with a distribution of $2 (n + d)$ samples. Hence,
@@
 
+%% TODO: comment on the fact that we can reduce the number of sample states
 The derivative method draws on the intuition that
@@
 Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics
-\cite{berland2005solving, einkemmer2017performance}.
+term may be used to efficiently integrate the coupled dynamics,
+see Appendix \ref{appendix:derivative}.
 For this method $m$ state derivatives are associated with each
@@
 
+%% TODO: fix figure and table referencing
 We compare the numerical methods
-to an analytically derived $Z/2$ gate. This gate corresponds to
+to an analytically derived $Z/2$ gate, see Figure \ref{fig:static}a. This gate corresponds to
 idling at the flux frustration point $a = 0$. The analytic gate
@@
 qubit frequency detuning the analytic gate achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-which is sufficient for quantum error correction, see Figure 2b.
+which is sufficient for quantum error correction, see Figure \ref{fig:static}b.
 Although the analytic $Z/2$ gate performs well, it works
@@
 Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times,
-see Figure 2b. These numerical methods offer an effective scheme for synchronizing
+all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
+These numerical methods offer an effective scheme for synchronizing
 qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
@@
 a function of the gate duration. The gate error for the 1\textsuperscript{st}-order
-derivative method approaches zero at the larmor period $1 / f_{q}$.
+derivative method approaches zero at the larmor period $1 / f_{q}$, see Figure \ref{fig:static}c.
 We believe the 1\textsuperscript{st}-order method outperforms the \textsuperscript{nd}-order
 method due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Table 2.
+terms to the gate error in this deviation regime, see Appendix \ref{appendix:derivative}.
 
-%% TODO: comment on the fact that we can reduce the number of sample states
 %% TODO: how do I tie this in?

### 2020-11-12 paper/6stochastic.tex
misc updates

@@
 filtering white noise sampled from a standard normal distribution using a finite
-impulse response filter. It is then scaled by the 
+impulse response filter \cite{saspweb2011}
+\footnote{\url{https://ccrma.stanford.edu/~jos/sasp/Example_Synthesis_1_F_Noise.html}}.
+It is then scaled by the 
 flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
@@
 We simulate successive applications of the gate constructed by each method and compute the gate error
-after each application, see Figure 3. Both the analytic and numerical methods achieve single gate errors
+after each application, see Figure \ref{fig:stochastic}. Both the analytic and numerical methods achieve single gate errors
 sufficient for quantum error correction. Despite converging on qualitatively different solutions, the

### 2020-11-12 paper/a1longitude.tex
misc updates

--- 2020-11-06 paper/a1dissipation.tex
@@
 \section{Longitudinal Relaxation}
-The probability of longitudinal relaxation metric and gate error due to longitudinal
+\label{appendix:longitude}
+We comment on the longitudinal relaxation metrics and then give
+our procedure for integrating the Lindblad master equation.
+The longitudal relaxation probability metric and gate error due to longitudinal
 relaxation metric are compared in Table I for the experiment shown in Figure 1c.
-The relative performance between the analytic and numerical techniques we study
+The relative performance of the analytic and numerical techniques we study
 is similar across the two metrics.
@@
 \begin{table}[h!]
-  \begin{tabular}{c | c | c}
-    Gate & $P_{1\textrm{A}} / P_{1\textrm{N}}$ & GE$_{\textrm{A}}$ / GE$_{\textrm{N}}$\\
+  \begin{tabular}{c | c | c | c | c | c | c}
+         & $P_{1\textrm{A}}$ & $P_{1\textrm{N}}$ & & GE$_{\textrm{A}}$ & GE$_{\textrm{N}}$ &\\
+    Gate & $(10^{-5})$ & $(10^{-5})$ & $P_{1\textrm{A}}/ P_{1\textrm{N}}$ & $(10^{-5})$ &
+    $(10^{-5})$ & GE$_{\textrm{A}} / \textrm{GE}_{\textrm{N}}$\\
     \hline
-    Z/2 & 5.745/1.149 = 5.000 & 1.776/0.371 = 4.787\\
-    Y/2 & 5.253/1.157 = 4.540 & 1.539/0.370 = 4.159\\
-    X/2 & 16.251/2.660 = 6.109 & 5.347/0.863 = 6.196\\
+    Z/2 & 5.745 & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
+    Y/2 & 5.253 & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
+    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
   \end{tabular}
   \caption{
-    Single gate probability of longitudinal relaxation
-    ratios and single gate error due to longitudinal relaxation
-    ratios for the analytic and numerical methods. Both probabilities
-    and gate errors are shown in units of $10^{-5}$.
+    Single gate longitudinal relaxation probability ratios
+    and single gate error due to longitudinal relaxation ratios.
+    Values are reported for the analytic (A) and numerical (N) techniques
+    presented in the main text.
   }

### 2020-11-12 paper/a2unscented.tex
misc updates

@@
 \section{Unscented Transformation}
+\label{appendix:unscented}
 In this section we outline the full unscented sampling procedure.
 First, we remark that ALTRO does not natively support complex numbers, so our optimization
-is performed under an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$.
+is performed under an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
+\cite{leung2017speedup}
 \begin{equation}
@@
 from the initial distributions
-\begin{equation}
+\begin{equation}\label{eq:uupdate}
   \begin{bmatrix} \Psi_{0}^{i} \\ \Lambda_{0}^{i} \end{bmatrix} =
@@
 is a hyperparameter which controls the spacing of the covariance contour.
-The $\pm$ is understood to take $+$ for $ i \in \{1, \dots, 2n + d\}$ and $-$ for
+The $(\pm)$ is understood to take $(+)$ for $ i \in \{1, \dots, 2n + d\}$ and $(-)$ for
 $i \in \{2n + d + 1, \dots, 4n + 2d\}$. We use the Cholesky factorization
@@
 such as the principal square root may be employed.
-The superscript indicates the $i$\textsuperscript{th} column of the
-lower triangular Cholesky factor.
+The superscript on the matrix square root indicates the $i$\textsuperscript{th}
+column (mod $2n + d$) of the lower triangular Cholesky factor.
 Then, the sigma points are normalized
-\begin{equation}
+\begin{equation}\label{eq:unormalize}
   \Psi_{0}^{i} \gets \frac{\Psi_{0}^{i}}{\sqrt{\Psi_{0}^{i^{T}} \Psi_{0}^{i}}}
@@
 The sigma points are propagated to the next knot point
-\begin{equation}
+\begin{equation}\label{eq:upropagate}
   \Psi_{1} = f(\Psi_{0}, \Lambda_{0})
 \end{equation}
-The mean and covariance of the sigma points is computed
+The mean and covariance of the sigma points are computed
 \begin{align}
@@
 \end{align}
-The sigma points are then resampled and propagated to the next knot point using eq B2, B3, and B4.
+The sigma points are then resampled and propagated to the next knot point using
+\eqref{eq:uupdate}, \eqref{eq:unormalize}, and \eqref{eq:upropagate}.

### 2020-11-12 paper/a3derivative.tex
misc updates

@@
 \section{Derivative Method}
+\label{appendix:derivative}
+We comment on the relative performance of the derivative methods and then
+outline how to efficiently integrate their dynamics.
 We believe the 1\textsuperscript{st}-order derivative method tends
@@
 The norm of the state derivatives for the 1\textsuperscript{st}- and 2\textsuperscript{nd}-order
-derivative methods are provided in Table 2 for the solutions of the optimization outlined
+derivative methods are provided in Table \ref{tab:dnorm} for the solutions of the optimization outlined
 in section V at the gate duration $t_{N} = 38$ns.
@@
 \begin{table}[h]
+  \label{tab:dnorm}
   \begin{tabular}{c | c | c}
@@
 
-The dynamics for the derivative methods can be integrated efficiently using exponential integrators.
+The dynamics for the derivative methods can be integrated efficiently using exponential integrators
+\cite{berland2005solving, einkemmer2017performance}.
 General exponential integrators break the dynamics into a linear term and a non-linear term. For example, consider
@@
 \begin{equation}
+  \label{eq:dgeneralexp}
   \begin{aligned}
-    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(L_{k} \Delta t_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
-    &+ \int_{0}^{\Delta t_{k}} \exp(L_{k}(\Delta t_{k} - t^{'})) N(t_{k} + t^{'}) dt^{'}\\
+    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(\Delta t_{k} L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
+    &+ \int_{0}^{\Delta t_{k}} \exp((\Delta t_{k} - t^{'})L_{k}) N(t_{k} + t^{'}) dt^{'}\\
   \end{aligned}
 \end{equation}
-General exponential integrators proceed by breaking the integral into a discrete sum, similar to the procedure
-for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler method \cite{berland2005solving}
+General exponential integrators proceed by breaking the integral in \eqref{eq:dgeneralexp}
+into a discrete sum, similar to the procedure
+for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler
+method \cite{berland2005solving}
 \begin{equation}
   \begin{aligned}
-    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(L_{k} \Delta t_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
-    &+ \exp(L_{k} \Delta t_{k}) N_{k}\\
+    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(\Delta t_{k}L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
+    &+ \exp(\Delta t_{k}L_{k}) N_{k}\\
   \end{aligned}
@@
 This method provides a good tradeoff between accuracy and efficiency, requiring one unique matrix
-exponential computation per stage. Exact integration accuracy is not of the utmost importance because the
-state derivatives guide the robust optimization, and do not correspond to experimental parameters
+exponential computation per stage. Integration accuracy is not of the utmost importance because the
+state derivatives guide the optimization, and do not correspond to experimental parameters
 which must be realized with high accuracy.

### 2020-11-12 paper/main.tex
updated many, added pdf
misc updates

@@
 
-%% TITLE
-\title{Robust Control of a Fluxonium Qubit}
-
-\author{Thomas Propson}
-\email{tcpropson@uchicago.edu}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\author{Brian Jackson}
-\author{Zac Manchester}
-\affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
-}
-\author{David I. Schuster}
-\affiliation{
-  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
-}
-\affiliation{
-  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
-}
-
-\date{\today}
-
 \input{0abstract.tex}
-
 \maketitle
@@
 \appendix
-\input{a1experiment.tex}
-\input{a2dissipation.tex}
+\input{a1longitude.tex}
+\input{a2unscented.tex}
 \input{a3derivative.tex}
-\input{a4ctensor.tex}
 

### 2020-11-13 paper/5static.tex
paper updates

@@
-%% S5
-\section{Robustness to Static Parameter Deviations}
+\section{Robustness to Static Parameter Deviations \label{sec:static}}
 We have formulated the quantum optimal control
@@
 \begin{figure*}[ht]
-  \label{fig:stochastic}
   \begin{subfigure}{.4\textwidth}
@@
   \end{subfigure}
-
   \caption{
@@
   }
+  \label{fig:stochastic}
 \end{figure*}

### 2020-11-13 paper/a3derivative.tex
paper updates

@@
 \label{appendix:derivative}
-We comment on the relative performance of the derivative methods and then
+We comment on the optimization metrics of the derivative methods and then
 outline how to efficiently integrate their dynamics.
-We believe the 1\textsuperscript{st}-order derivative method tends
+The 1\textsuperscript{st}-order derivative method tends
 to outperform the 2\textsuperscript{nd}-order
-derivative method in the small, static detuning regime, see Figure 2b,
-because second-order contributions to the gate error are small.
-The norm of the state derivatives for the 1\textsuperscript{st}- and 2\textsuperscript{nd}-order
-derivative methods are provided in Table \ref{tab:dnorm} for the solutions of the optimization outlined
-in section V at the gate duration $t_{N} = 38$ns.
+derivative method in the small, static detuning regime, see Figure \ref{fig:static}b.
+The norms of the state derivatives for the 1\textsuperscript{st}- and 2\textsuperscript{nd}-order
+methods are provided in Table \ref{tab:dnorm} for the solutions at $t_{N} = 60$ns.
+The 2\textsuperscript{nd}-order method is able to decrease the 2\textsuperscript{nd}-order
+state derivate norm relative to the 1\textsuperscript{st}-order method at the expense of increasing
+its 1\textsuperscript{st}-order derivative norm. In the problem
+we have studied it is likely the 2\textsuperscript{nd}-order derivative norm has a smaller
+contribution to the gate error than the 1\textsuperscript{st}-order derivative norm. A
+careful analysis could be completed for future problems to predict the efficacy of the
+derivative method at each order.
 
-%%TODO: the values in this table are not up to date until this comment is removed
 \begin{table}[h]
-  \label{tab:dnorm}
   \begin{tabular}{c | c | c}
     Method & ${\lvert \braket{\partial_{f_{q}} \psi_{N} | \partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \braket{\partial^{2}_{f_{q}} \psi_{N} | \partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
+    & ${\lvert \braket{\partial^{2}_{f_{q}} \psi_{N} | \partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{6}$)\\
     \hline
-    D-1 & 2.051 & 17.935\\
-    D-2 & 2.511 & 7.628\\
+    D-1 & 0.436 & 57.817\\
+    D-2 & 1.702 & 9.030\\
   \end{tabular}
-  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
+  \caption{Norm of state derivates with respect to the qubit frequency
     for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 38$ns.}
+    at the end of the gate's duration $t_{N} = 60$ns and are averaged over each initial state.}
+  \label{tab:dnorm}
 \end{table}

### 2020-11-14 src/spin/figures.jl
spin23 functional

@@ -224,7 +224,7 @@ const F2A_DATA_ZPIBY2 = Dict(
     analytic => joinpath(SPIN_OUT_PATH, "spin14/00000_spin14.h5"),
     s2 => joinpath(SPIN_OUT_PATH, "spin12/00808_spin12.h5"),
     sut8 => joinpath(SPIN_OUT_PATH, "spin23/00293_spin23.h5"),
-    d1 => joinpath(SPIN_OUT_PATH, "spin11/00683_spin11.h5"),
+    d1 => joinpath(SPIN_OUT_PATH, "spin11/00691_spin11.h5"),
     d2 => joinpath(SPIN_OUT_PATH, "spin11/00685_spin11.h5"),
 )
 const F2A_PT_LIST = [analytic, s2, sut8, d1, d2]
@@ -253,7 +253,7 @@ end
 const F2B_DATA_ZPIBY2 = Dict(
     analytic => joinpath(SPIN_OUT_PATH, "spin14/00000_spin14.h5"),
     d1 => joinpath(SPIN_OUT_PATH, "spin11/00612_spin11.h5"),
-    d1b => joinpath(SPIN_OUT_PATH, "spin11/00683_spin11.h5"),
+    d1b => joinpath(SPIN_OUT_PATH, "spin11/00691_spin11.h5"),
     d1bb => joinpath(SPIN_OUT_PATH, "spin11/00632_spin11.h5"),
     d1bbb => joinpath(SPIN_OUT_PATH, "spin11/00585_spin11.h5"),
 )
@@ -371,7 +371,7 @@ const F2C_DATA_ZPIBY2 = Dict(
         301, 304, 302, 305, 307, 306, 308, 309,
     ]],
     d1 => [joinpath(SPIN_OUT_PATH, "spin11/$(lpad(index, 5, '0'))_spin11.h5") for index in [
-        612, 614, 617, 616, 618, 619, 620, 621, 622, 683,
+        612, 614, 617, 616, 618, 619, 620, 621, 622, 691,
         628, 624, 625, 627, 629, 630, 609, 631, 632, 633,
         634, 635, 636, 637, 638, 639, 610, 585,
     ]],
@@ -478,8 +478,8 @@ const F3_DATA = Dict(
     analytic => joinpath(SPIN_OUT_PATH, "spin14/00004_spin14.h5"),
[…]

### 2020-11-14 src/spin/figures.py

@@ -93,19 +93,19 @@ class PulseType(Enum):
 #ENDDEF
 
 PT_STR = {
-    PulseType.analytic: "Anl.",
+    PulseType.analytic: "A",
     PulseType.qoc: "QOC",
-    PulseType.s2: "S-8",
+    PulseType.s2: "S",
     PulseType.s4: "S-4",
-    PulseType.d1: "D-1",
-    PulseType.d2: "D-2",
+    PulseType.d1: "D1",
+    PulseType.d2: "D2",
     PulseType.d3: "D-3",
     PulseType.s2b: "S-2",
     PulseType.s4b: "S-4",
     PulseType.d2b: "D-2",
     PulseType.d3b: "D-2",
     PulseType.corpse: "C-2",
-    PulseType.sut8: "SU-10",
+    PulseType.sut8: "SU",
 }
 
 PT_COLOR = {
@@ -427,7 +427,7 @@ def make_figure1c():
     #ENDFOR
 
     # configure main 
-    ax.set_ylabel("Gate Error", fontsize=LABEL_FS)
+    ax.set_ylabel("Total Gate Error", fontsize=LABEL_FS)
     ax.set_xlabel("Gate Count", fontsize=LABEL_FS)
     ax.set_xlim(0, gate_count)
     xticks_ = [0, 500, 1000, 1500, 2000]
@@ -496,11 +496,11 @@ def make_figure2a():
         axs[i].tick_params(direction="in", labelsize=TICK_FS)
         axs[i].set_xticks([])
     #ENDFOR
-    fig.text(0.27, 0.91, "Anl.", fontsize=TEXT_FS)
-    fig.text(0.27, 0.75, "S-8", fontsize=TEXT_FS)
-    fig.text(0.27, 0.59, "SU-10", fontsize=TEXT_FS)
-    fig.text(0.27, 0.425, "D-1", fontsize=TEXT_FS)
-    fig.text(0.27, 0.265, "D-2", fontsize=TEXT_FS)
+    fig.text(0.27, 0.91, PT_STR[PulseType.analytic], fontsize=TEXT_FS)
+    fig.text(0.27, 0.75, PT_STR[PulseType.s2], fontsize=TEXT_FS)
+    fig.text(0.27, 0.59, PT_STR[PulseType.sut8], fontsize=TEXT_FS)
[…]

### 2020-11-14 src/spin/spin17.jl

@@ -153,7 +153,6 @@ function run_traj(;gate_type=xpiby2, evolution_time=56.8, solver_type=altro,
     x_min = fill(-Inf, n)
     x_min[CONTROLS_IDX] .= -MAX_CONTROL_NORM_0
     x_min = SVector{n}(x_min)
-    x_max = fill(Inf, n)
     
     # control amplitude constraint at boundary
     x_max_boundary = fill(Inf, n)

### 2020-11-14 paper/3fluxonium.tex
misc updates
removed old dirs

@@
-%% S3
 \section{QOC on the Fluxonium}
@@
 constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines.
-%% TODO: See ref. 28 in Helin paper
+the hysteresis ubiquitous in flux bias lines,
+\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
 We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
@@
 is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
-(see Appendix).
+They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively.
 The main contributions to longitudinal relaxation in our

### 2020-11-14 paper/5static.tex
removed old dirs

@@
+%% TODO: standardize use of "paramter" vs "system parameter" language
 \section{Robustness to Static Parameter Deviations \label{sec:static}}
@@
 However, the device typically deviates from the Hamiltonian we use in optimization,
-leading to poor experimental performance. To mitigate
-these errors we employ robust control techniques
-to make the state evolution insensitive to Hamiltonian
-parameter deviations. As an example
+leading to poor experimental performance. We combat errors
+of this form using robust control techniques,
+making the state evolution insensitive
+to Hamiltonian parameter deviations. As an example
 we mitigate errors arising from the drift and finite measurement
@@
 to capture the effect of parameter deviations. We also
-study the unscented sampling method which uses the unscented
+study the unscented sampling method, which uses the unscented
 transformation to accurately propagate a distribution
@@
 (\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-For this method the number of states in the augmented state vector
+This is a standard perscription, though knowledge
+of the problem can typically be used to reduced the number
+of required samples. For this method the number of states in the augmented state vector
 scales as $O(dn^{2})$ because each of $n^{2}$ initial
@@
 models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $2 (n + d)$ states.
+parameter deviation. The ensemble consists of $4n + 2d$ states.
 Each state in the
@@
 This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{n}$\textsuperscript{th}
+first and second moments of the distribution and ensurnes the points
+lie on the $\sqrt{2n}$\textsuperscript{th}
 covariance contour at each knot point. A detailed update procedure is given
 in Appendix \ref{appendix:unscented}. Each initial state from the operator basis
-is represented with a distribution of $2 (n + d)$ samples. Hence,
+is represented with a distribution of $4n + 2d$ samples. Hence,
 the number of states in the augmented state vector scales as $O(n^{3} + dn^{2})$.
@@
 
-%% TODO: comment on the fact that we can reduce the number of sample states
 The derivative method draws on the intuition that
@@
 
-%% TODO: fix figure and table referencing
 We compare the numerical methods
@@
 derivative method approaches zero at the larmor period $1 / f_{q}$, see Figure \ref{fig:static}c.
-We believe the 1\textsuperscript{st}-order method outperforms the \textsuperscript{nd}-order
+We believe the 1\textsuperscript{st}-order method outperforms the 2\textsuperscript{nd}-order
 method due to the low contribution of second-order

### 2020-11-14 paper/a1longitude.tex
removed old dirs

@@
 our procedure for integrating the Lindblad master equation.
-The longitudal relaxation probability metric and gate error due to longitudinal
-relaxation metric are compared in Table I for the experiment shown in Figure 1c.
-The relative performance of the analytic and numerical techniques we study
+The longitudal relaxation probability and the gate error due to longitudinal
+relaxation metrics are compared in Table \ref{tab:longitude} for the experiment shown in Figure \ref{fig:longitude}c.
+The relative performance of the analytic and numerical techniques
 is similar across the two metrics.
@@
   }
+  \label{tab:longitude}
 \end{table}
@@
 $L_{\downarrow} = \sigma^{-}/2$ where $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$.
-Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained from the spline shown in Figure 1b.
+Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained from the spline shown in Figure \ref{fig:longitude}b.
 The $T_{1}$ values in this spline are obtained experimentally by driving the qubit at the desired flux amplitude

### 2020-11-14 rbqoc/util/hrutil.py

@@ -1,191 +1,0 @@
-"""
-hrutil.py - This module exposes utility functions for hamiltonian parameter robustness
-"""
-
-from copy import copy
-import os
-
-from filelock import FileLock, Timeout
-import h5py
-import matplotlib.pyplot as plt
-import numpy as np
-from qoc import evolve_schroedinger_discrete
-
-COLOR_PALETTE = (
-    "blue", "red", "green", "pink", "purple", "orange",
-    "teal", "grey", "black", "cyan", "magenta", "brown",
-    "azure", "beige", "coral", "crimson",
-)
-
-COLOR_PALETTE_LEN = len(COLOR_PALETTE)
-
-def hamiltonian_args_sweep(evol_args, hamiltonian_args_set, pulse_path, suffix):
-    """
-    Given a set of hamiltonian_args, run evolve_schroedinger_discrete for a control pulse
-    and save the error associated with each evolution to an h5 file.
-
-    Arguments:
-    evol_args :: dict(any) - The arguments to pass to evolve_schroedinger_discrete
-        in addition to the hamiltonian_args and the control pulse
-    hamiltonian_args_set :: ndarray(trial_count x hamiltonian_arg_count)
-        - A list of the hamiltonian arguments that should be swept over.
-    pulse_path :: list(str) - A list of paths to the output file
-        of the pulse that should be evolved.
-    suffix :: str - The suffix to append to the pulse_path that will
-        constitute the new file where data from these evolutions is stored.
-
-    Returns:
-    save_path :: str - The destination of the file where the
-        sweep information was saved
-    """
-    # Obtain the locks and full paths to the pulse and save files.
-    pulse_path_lock = "{}.lock".format(pulse_path)
-    # We assume that `pulse_path` is an h5 file with the suffix ".h5".
-    save_path = "{}_{}.h5".format(pulse_path.split(".h5")[0], suffix)
-    save_path_lock = "{}.lock".format(save_path)
-
-    # Grab the control pulse that achieved the minimum error.
-    try:
-        with FileLock(pulse_path_lock):
-            with h5py.File(pulse_path, "r") as pulse_file:
-                if "error" in pulse_file.keys():
-                    if "constraint" in pulse_file.keys():
-                        indices = np.nonzero(pulse_file["constraint"][()])[0]
-                        errors = pulse_file["error"][()][indices]
-                        index = indices[np.argmin(errors)]
-                    else:
-                        index = np.argmin(pulse_file["error"][()])
[…]

### 2020-11-14 src/cavity/cavity0.py

@@ -1,183 +1,0 @@
-"""
-cavity0.py - Cavity Experiment 0.
-"""
-
-from argparse import ArgumentParser
-from copy import copy
-import os
-
-import autograd.numpy as anp
-from filelock import FileLock
-import h5py
-import numpy as np
-from qoc import (
-    evolve_schroedinger_discrete,
-    grape_schroedinger_discrete,
-)
-from qoc.standard import (
-    Adam, LBFGSB,
-    TargetStateInfidelity,
-    conjugate_transpose, matmuls, krons,
-    get_annihilation_operator, get_creation_operator,
-    generate_save_file_path,
-)
-
-# Define paths
-META_NAME = "cavity"
-EXPERIMENT_NAME = "cavity0"
-RBQOC_PATH = os.environ["ROBUST_QOC_PATH"]
-SAVE_PATH = os.path.join(RBQOC_PATH, "out", META_NAME, EXPERIMENT_NAME)
-
-# Define experimental constants. All units are in GHz.
-TRANSMON_FREQ = 2 * np.pi * 5.6640
-CAVITY_FREQ = 2 * np.pi * 4.4526 
-CHI = 2 * np.pi * -0.5644535742521878e-3
-ALPHA = 2 * np.pi * -0.13951263895963648
-KAPPA = 2 * np.pi * -3.3e-6
-CHI_PRIME = 2 * np.pi * -0.73e-6 
-MAX_AMP_NORM_CAVITY = np.sqrt(2) * 2 * np.pi * 4e-4
-MAX_AMP_NORM_TRANSMON = np.sqrt(2) * 2 * np.pi * 4e-3
-
-# Define the system
-CAVITY_STATE_COUNT = 3
-CAVITY_ANNIHILATE = get_annihilation_operator(CAVITY_STATE_COUNT)
-CAVITY_CREATE = get_creation_operator(CAVITY_STATE_COUNT)
-CAVITY_NUMBER = np.matmul(CAVITY_CREATE, CAVITY_ANNIHILATE)
-CAVITY_C2_A2 = matmuls(CAVITY_CREATE, CAVITY_CREATE, CAVITY_ANNIHILATE, CAVITY_ANNIHILATE)
-CAVITY_ID = np.eye(CAVITY_STATE_COUNT)
-CAVITY_VACUUM = np.zeros((CAVITY_STATE_COUNT, 1))
-CAVITY_ZERO = np.copy(CAVITY_VACUUM)
[…]

### 2020-11-14 src/spin/piby2.py
removed some redundant things

@@ -1,85 +1,0 @@
-"""
-piby2.py - This module searches for pi/2 pulses
-for the spin system.
-"""
-
-import os
-
-import autograd.numpy as anp
-from qoc import grape_schroedinger_discrete
-from qoc.standard import (conjugate_transpose,
-                          generate_save_file_path,
-                          SIGMA_X, SIGMA_Z,
-                          TargetStateInfidelity,
-                          ControlArea,
-                          ControlBandwidth,
-                          LBFGSB,
-                          Adam,)
-
-CORE_COUNT = 8
-os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
-os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
-
-# Define experimental constants.
-CONTROL_SAMPLING_RATE = 1.2 # GS/s
-MAX_AMP_CONTROL_0 = 1e-1 # GHz
-MAX_AMP_BANDWIDTH_CONTROL_0 = 5e-1 # GHz
-OMEGA = 1e-2 # GHz
-
-
-# Define the system.
-SYSTEM_HAMILTONIAN_0 = OMEGA * SIGMA_Z / 2
-CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
-hamiltonian = lambda controls, time: (SYSTEM_HAMILTONIAN_0
-                                      + controls[0] * CONTROL_HAMILTONIAN_0)
-MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
-MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
-
-# Define the optimization.
-ITERATION_COUNT = int(1e4)
-COMPLEX_CONTROLS = False
-CONTROL_COUNT = 1
-EVOLUTION_TIME = 150 # nanoseconds
-CONTROL_STEP_COUNT = int(EVOLUTION_TIME * CONTROL_SAMPLING_RATE)
-OPTIMIZER = Adam()
-
-# Define the problem.
-INITIAL_STATE_0 = anp.array(((1,), (0,)))
-INITIAL_STATE_1 = anp.array(((0,), (1,)))
-INITIAL_STATES = anp.stack((INITIAL_STATE_0, INITIAL_STATE_1,))
-# Rx(pi/2)
[…]

### 2020-11-14 src/spin/spin12_hpsweep.py

@@ -1,321 +1,0 @@
-"""
-spin12_hpsweep.py - hp sweep for spin11 and spin12
-"""
-
-from argparse import ArgumentParser
-from enum import Enum
-import os
-
-import h5py
-import numpy as np
-import matplotlib.pyplot as plt
-from qoc.standard import (
-    conjugate_transpose,
-    generate_save_file_path,
-)
-from qutip import (
-    mesolve, Qobj,
-)
-
-class PulseType(Enum):
-    TRAJ = 0
-    QOC = 1
-#ENDDEF
-
-# Directory.
-WDIR = os.environ.get("ROBUST_QOC_PATH", ".")
-OUT_PATH = os.path.join(WDIR, "out")
-SAVE_PATH = os.path.join(OUT_PATH, "spin", "spin12")
-EXPERIMENT_NAME = "spin12_hpsweep"
-EXPERIMENT_META = "spin"
-
-# Data constants.
-PULSE_DATA = [
-    ("spin12", "00018_spin12.h5", 13, PulseType.TRAJ, "sample"),
-    ("spin12", "00027_spin12.h5", 21, PulseType.TRAJ, "sample 2"),
-]
-
-# Sweep building.
-SWEEP_COUNT = int(2e2)
-SWEEP_MULTIPLIERS = np.linspace(-0.1, 0.1, SWEEP_COUNT)
-
-
-# other constants
-DPI = int(5e2)
-ALPHA = 0.5
-MS = 15
-LINE_WIDTH = 1
-LINE_ALPHA = 0.8
-LINE_COLOR = "black"
-COLORS = [
-    "blue", "red", "green", "orange", "purple"
-]
-
-
-# Computational constants.
-SIGMA_X = np.array([[0, 1],
-                    [1, 0]])
-SIGMA_Z = np.array([[1, 0],
-                    [0, -1]])
-H_S = SIGMA_Z / 2
-H_C1 = SIGMA_X / 2
-OMEGA_RAW = 2 * np.pi * 1.4e-2
-OMEGA_STD = 5e-2
-DOMEGA = OMEGA_RAW * OMEGA_STD
-OMEGA_PLUS = OMEGA_RAW + DOMEGA
-OMEGA_MINUS = OMEGA_RAW - DOMEGA
-
-
-def fidelity(v1, v2):
-    ip = np.matmul(conjugate_transpose(v1), v2)[0, 0]
-    return np.real(ip) ** 2 + np.imag(ip) ** 2
-
-
[…]

### 2020-11-14 src/spin/spin15_bench.jl

@@ -1,789 +1,0 @@
-"""
-spin15_bench.jl - lindblad evolution for spin15.jl in Julia
-"""
-
-using Dates
-using HDF5
-import LaTeXStrings
-using LinearAlgebra
-using MPI
-import Plots
-using Printf
-using Random
-using StaticArrays
-using Statistics
-using TrajectoryOptimization
-
-
-EXPERIMENT_META = "spin"
-EXPERIMENT_NAME = "spin15_bench"
-WDIR = ENV["ROBUST_QOC_PATH"]
-META_PATH = joinpath(WDIR, "out", EXPERIMENT_META)
-SAVE_PATH = joinpath(META_PATH, EXPERIMENT_NAME)
-ZPIBY2_COMPARISON_FILE_PATH = joinpath(SAVE_PATH, "zpiby2_comparsion_spin15_bench.png")
-
-# plotting configuration
-ENV["GKSwstype"] = "nul"
-Plots.gr()
-
-# Define experimental constants.
-# qubit frequency at flux frustration point
-OMEGA = 2 * pi * 1.4e-2 #GHz
-FBFQ_A = 0.202407
-FBFQ_B = 0.5
-# coefficients are listed in descending order
-# raw coefficients are in units of seconds
-FBFQ_T1_COEFFS = [
-    3276.06057; -7905.24414; 8285.24137; -4939.22432;
-    1821.23488; -415.520981; 53.9684414; -3.04500484
-] * 1e9
-
-# Define the system.
-NEG_I = SA_F64[0  0 1 0;
-               0  0 0 1;
-               -1 0 0 0;
-               0 -1 0 0;]
-SIGMA_X = SA_F64[0   1   0   0;
-                 1   0   0   0;
-                 0   0   0   1;
-                 0   0   1   0;]
-SIGMA_Z = SA_F64[1   0   0   0;
-                 0  -1   0   0;
-                 0   0   1   0;
-                 0   0   0  -1;]
-H_S = SIGMA_Z / 2
-OMEGA_NEG_I_H_S = OMEGA * NEG_I * H_S
-H_C1 = SIGMA_X / 2
-NEG_I_2_PI_H_C1 = 2 * pi * NEG_I * H_C1
-# L_{0} = |g> <e|
-# L_{0}^{\dagger} = |e> <g|
-# L_{0}^{\dagger} L_{0} = |e> <e|
-# L_{1} = L_{0}^{\dagger} = |e> <g|
-# L_{1}^{\dagger} = L_{0} = |g> <e|
[…]

### 2020-11-14 src/spin/spin15_bench.py

@@ -1,396 +1,0 @@
-"""
-spin15_bench.py - benchmark the spin15.jl t1 pulses
-"""
-
-from argparse import ArgumentParser
-from copy import copy
-from enum import Enum
-import os
-
-import h5py
-import matplotlib.pyplot as plt
-import numpy as np
-from qoc.standard import (
-    conjugate_transpose, generate_save_file_path,
-    matmuls
-)
-from qutip import (
-    mesolve, Qobj, basis
-)
-
-
-# paths
-EXPERIMENT_META = "spin"
-EXPERIMENT_NAME = "spin15_bench"
-WDIR = os.environ.get("ROBUST_QOC_PATH", ".")
-META_PATH = os.path.join(WDIR, "out", EXPERIMENT_META)
-SAVE_PATH = os.path.join(META_PATH, EXPERIMENT_NAME)
-VE_FILE_NAME = "ve_spin15_bench"
-
-# comp. onstants
-OMEGA = 2 * np.pi * 1.4e-2
-INV_ROOT_2 = 2 ** (-1/2)
-DT = 1e-2
-DT_INV = 1e2
-CONTROL_COUNT = 1
-CONTROLS_ZERO = np.zeros((1, CONTROL_COUNT))
-FBFQ_A = 0.202407
-FBFQ_B = 0.5
-# Sorted from highest order to lowest order.
-# These coeffecients are in units of s.
-FBFQ_T1_COEFFS = np.array([
-    3276.06057, -7905.24414, 8285.24137, -4939.22432,
-    1821.23488, -415.520981, 53.9684414, -3.04500484
-]) * 1e9
-
-INITIAL_STATE = np.array([[1.], [0.]])
-INITIAL_DENSITY = np.matmul(INITIAL_STATE, conjugate_transpose(INITIAL_STATE))
-SIGMA_X = np.array([[0, 1],
-                    [1, 0]])
-SIGMA_Z = np.array([[1, 0],
-                    [0, -1]])
-H_S = SIGMA_Z / 2
-H_C1 = SIGMA_X / 2
-XPIBY2 = np.array([[1.,  -1j],
-                   [-1.j, 1.]]) * INV_ROOT_2
-YPIBY2 = np.array([[1., -1.],
-                   [1.,  1.]]) * INV_ROOT_2
-ZPIBY2 = np.array([[1. - 1.j, 0],
-                   [0,        1. + 1.j]]) * INV_ROOT_2
-C_G_TO_E = np.array([[0., 1.],
[…]

### 2020-11-14 src/spin/spin8.py

@@ -1,183 +1,0 @@
-"""
-spin8.py - spin experiment 8
-"""
-
-from argparse import ArgumentParser
-from copy import copy
-import os
-
-import autograd.numpy as anp
-from filelock import FileLock
-import h5py
-import numpy as np
-from qoc import (
-    evolve_schroedinger_discrete,
-    grape_schroedinger_discrete,
-)
-from qoc.standard import (
-    Adam, LBFGSB,
-    ControlArea,
-    TargetStateInfidelity,
-    TargetStateInfidelityTime,
-    conjugate_transpose, matmuls, krons,
-    get_annihilation_operator, get_creation_operator,
-    generate_save_file_path,
-    SIGMA_X, SIGMA_Z
-)
-
-# Define paths
-META_NAME = "spin"
-EXPERIMENT_NAME = "spin8"
-SAVE_PATH = os.path.join(os.environ["ROBUST_QOC_PATH"], "out", META_NAME, EXPERIMENT_NAME)
-
-
-# Define experimental constants.
-FLUXONIUM_FREQ = 2 * np.pi * 1e-2 #GHz
-MAX_AMP_NORM_FLUXONIUM = 2 * np.pi * 3e-1 #GHz
-SAMPLE_RATE = 1.2 #samples/ns
-
-
-# Define the system
-FLUXONIUM_STATE_COUNT = 2
-FLUXONIUM_VACUUM = np.zeros((FLUXONIUM_STATE_COUNT, 1))
-FLUXONIUM_G = np.copy(FLUXONIUM_VACUUM)
-FLUXONIUM_G[0][0] = 1.
-FLUXONIUM_E = np.copy(FLUXONIUM_VACUUM)
-FLUXONIUM_E[1][0] = 1.
-HAMILTONIAN_ARGS = np.array([FLUXONIUM_FREQ])
-H_SYSTEM = (
-    SIGMA_Z / 2
-)
-H_CONTROL_0 = (
-    SIGMA_X / 2
-)
-
-hamiltonian = lambda controls, hargs, time: (
-    hargs[0] * H_SYSTEM
-    + controls[0] * H_CONTROL_0
-)
-CONTROL_COUNT = 1
-COMPLEX_CONTROLS = False
-MAX_CONTROL_NORMS = np.array((
-    MAX_AMP_NORM_FLUXONIUM,
-))
-
-
-# Define the problem
[…]

### 2020-11-14 src/spin/spin9_amp.py

@@ -1,57 +1,0 @@
-"""
-spin9_amp.py - Do a sweep of the control amplitude multiplier
-for spin9.
-"""
-
-import os
-
-import numpy as np
-from rbqoc.util import (amp_sweep, plot_amp_sweep)
-from spin9 import (EVOL_CONFIG, EXPERIMENT_NAME, EXPERIMENT_META)
-
-# Sweep trials
-SWEEP_TRIAL_COUNT = 100
-AMP_MULTIPLIERS = np.linspace(0.98, 1.02, SWEEP_TRIAL_COUNT)
-
-
-# I/O
-WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
-SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
-PULSE_FILE_0 = "00000_spin9.h5"
-PULSE_FILE_1 = "00001_spin9.h5"
-PULSE_FILE_2 = "00005_spin9.h5"
-PULSE_FILES = (
-    PULSE_FILE_0,
-    PULSE_FILE_1,
-    PULSE_FILE_2,
-)
-PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
-PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
-PULSE_PATH_2 = os.path.join(SAVE_PATH, PULSE_FILE_2)
-PULSE_PATHS = (
-    PULSE_PATH_0,
-    PULSE_PATH_1,
-    PULSE_PATH_2,
-)
-SWEEP_SUFFIX = "amp"
-PLOT_SAVE_FILE = "{}_{}.png".format(EXPERIMENT_NAME, SWEEP_SUFFIX)
-PLOT_SAVE_PATH = os.path.join(SAVE_PATH, PLOT_SAVE_FILE)
-
-def main():
-    sweep_path_list = list()
-    # Sweep the amplitude parameter for each pulse.
-    for i, pulse_path in enumerate(PULSE_PATHS):
-        save_path = amp_sweep(AMP_MULTIPLIERS, EVOL_CONFIG,
-                              pulse_path, SWEEP_SUFFIX)
-        sweep_path_list.append(save_path)
-
-    # Plot the sweeps together.
-    plot_amp_sweep(PLOT_SAVE_PATH, sweep_path_list,
-                   title=EXPERIMENT_NAME,
-                   x_label="Amplitude Multiplier",
[…]

### 2020-11-14 src/spin/spin_exp0_hpsweep.py

@@ -1,63 +1,0 @@
-"""
-spin_exp0_hpsweep.py - Do a sweep of the hamiltonian parameters
-for spin_exp0.
-"""
-
-import os
-
-import numpy as np
-from rbqoc.util import (hamiltonian_args_sweep,
-                        plot_hamiltonian_args_sweep,)
-from spin_exp0 import (EVOL_ARGS, EXPERIMENT_META, EXPERIMENT_NAME, HAMILTONIAN_ARGS)
-
-# Sweep trials
-SWEEP_TRIAL_COUNT = 100
-SWEEP_MULTIPLIERS = np.linspace(0, 3, SWEEP_TRIAL_COUNT)
-HAMILTONIAN_ARGS_SWEEP = np.stack([HAMILTONIAN_ARGS] * SWEEP_TRIAL_COUNT)
-for i, multiplier in enumerate(SWEEP_MULTIPLIERS):
-    HAMILTONIAN_ARGS_SWEEP[i] *= SWEEP_MULTIPLIERS[i]
-
-# I/O
-WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
-SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
-PULSE_FILE_0 = "00000_spin_exp0.h5"
-PULSE_FILE_1 = "00001_spin_exp0.h5"
-PULSE_FILE_2 = "00013_spin_exp0.h5"
-PULSE_FILE_3 = "00010_spin_exp0.h5"
-PULSE_FILES = [PULSE_FILE_0,
-               PULSE_FILE_1,
-               PULSE_FILE_2,
-               PULSE_FILE_3,
-]
-PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
-PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
-PULSE_PATH_2 = os.path.join(SAVE_PATH, PULSE_FILE_2)
-PULSE_PATH_3 = os.path.join(SAVE_PATH, PULSE_FILE_3)
-PULSE_PATH_4 = os.path.join(SAVE_PATH, PULSE_FILE_4)
-PULSE_PATHS = [
-    PULSE_PATH_0,
-    PULSE_PATH_1,
-    PULSE_PATH_2,
-    PULSE_PATH_3,
-]
[…]

### 2020-11-14 src/spin/spin_exp1.py

@@ -1,136 +1,0 @@
-"""
-spin_exp1.py - Spin experiment 1.
-"""
-
-import os
-
-import autograd.numpy as anp
-from filelock import FileLock, Timeout
-import h5py
-import matplotlib.pyplot as plt
-import matplotlib.patches as mpatches
-from qoc import grape_schroedinger_discrete, evolve_schroedinger_discrete
-from qoc.standard import (TargetStateInfidelity,
-                          conjugate_transpose,
-                          get_annihilation_operator,
-                          get_creation_operator,
-                          SIGMA_Z, SIGMA_X,
-                          generate_save_file_path,
-                          krons, plot_controls,
-                          plot_state_population,
-                          Adam, LBFGSB,)
-
-CORE_COUNT = 8
-os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
-os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
-
-# Define experimental constants.
-CONTROL_SAMPLING_RATE = 1.2 # GS/s
-MAX_AMP_CONTROL_0 = 2 * anp.pi * 1e-1 # GHz
-MAX_AMP_BANDWIDTH_CONTROL_0 = 2 * anp.pi * 5e-1 # GHz
-OMEGA_Q = 2 * anp.pi * 1e-2 # GHz
-
-# Define the system.
-SYSTEM_HAMILTONIAN_0 = SIGMA_Z / 2
-CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
-HAMILTONIAN_ARGS = anp.array([OMEGA_Q])
-hamiltonian = lambda controls, hargs, time: (hargs[0] * SYSTEM_HAMILTONIAN_0
-                                             + controls[0] * CONTROL_HAMILTONIAN_0)
-MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
-MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
-
-# Define the optimization.
-WDIR = os.environ["ROBUST_QOC_PATH"]
-CONTROLS_PATH = os.path.join(WDIR, "out/spin/spin_exp1/00000_spin_exp1.h5")
-CONTROLS_PATH_LOCK = "{}.lock".format(CONTROLS_PATH)
-# try:
[…]

### 2020-11-14 src/spin/spin_exp1_ampsweep.py

@@ -1,52 +1,0 @@
-"""
-spin_exp1_ampsweep.py - Do a sweep of the control amplitude multiplier
-for spin_exp1.
-"""
-
-import os
-
-import numpy as np
-from rbqoc.util import (amp_sweep, plot_amp_sweep)
-from spin_exp1 import (EVOL_CONFIG, EXPERIMENT_NAME, EXPERIMENT_META)
-
-# Sweep trials
-SWEEP_TRIAL_COUNT = 100
-AMP_MULTIPLIERS = np.linspace(0.9, 1.1, SWEEP_TRIAL_COUNT)
-
-
-# I/O
-WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
-SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
-PULSE_FILE_0 = "00000_spin_exp1.h5"
-PULSE_FILE_1 = "00007_spin_exp1.h5"
-PULSE_FILES = (
-    PULSE_FILE_0,
-    PULSE_FILE_1,
-)
-PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
-PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
-PULSE_PATHS = (
-    PULSE_PATH_0,
-    PULSE_PATH_1,
-)
-SWEEP_SUFFIX = "amp_sweep"
-PLOT_SAVE_FILE = "{}_{}.png".format(EXPERIMENT_NAME, SWEEP_SUFFIX)
-PLOT_SAVE_PATH = os.path.join(SAVE_PATH, PLOT_SAVE_FILE)
-
-def main():
-    sweep_path_list = list()
-    # Sweep the amplitude parameter for each pulse.
-    for i, pulse_path in enumerate(PULSE_PATHS):
-        save_path = amp_sweep(AMP_MULTIPLIERS, EVOL_CONFIG,
-                              pulse_path, SWEEP_SUFFIX)
-        sweep_path_list.append(save_path)
-
-    # Plot the sweeps together.
-    plot_amp_sweep(PLOT_SAVE_PATH, sweep_path_list,
-                                title=EXPERIMENT_NAME,
-                                x_label="Amplitude Multiplier",
-                                y_label="Infidelity")
-
-
-if __name__ == "__main__":
-    main()

### 2020-11-14 src/spin/spin_exp2_ampsweep.py

@@ -1,52 +1,0 @@
-"""
-spin_exp2_ampsweep.py - Do a sweep of the control amplitude multiplier
-for spin_exp2.
-"""
-
-import os
-
-import numpy as np
-from rbqoc.util import (amp_sweep, plot_amp_sweep)
-from spin_exp2 import (EVOL_CONFIG, EXPERIMENT_NAME, EXPERIMENT_META)
-
-# Sweep trials
-SWEEP_TRIAL_COUNT = 100
-AMP_MULTIPLIERS = np.linspace(0.9, 1.1, SWEEP_TRIAL_COUNT)
-
-
-# I/O
-WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
-SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
-PULSE_FILE_0 = "00000_spin_exp2.h5"
-PULSE_FILE_1 = "00007_spin_exp2.h5"
-PULSE_FILES = (
-    PULSE_FILE_0,
-    PULSE_FILE_1,
-)
-PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
-PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
-PULSE_PATHS = (
-    PULSE_PATH_0,
-    PULSE_PATH_1,
-)
-SWEEP_SUFFIX = "amp_sweep"
-PLOT_SAVE_FILE = "{}_{}.png".format(EXPERIMENT_NAME, SWEEP_SUFFIX)
-PLOT_SAVE_PATH = os.path.join(SAVE_PATH, PLOT_SAVE_FILE)
-
-def main():
-    sweep_path_list = list()
-    # Sweep the amplitude parameter for each pulse.
-    for i, pulse_path in enumerate(PULSE_PATHS):
-        save_path = amp_sweep(AMP_MULTIPLIERS, EVOL_CONFIG,
-                              pulse_path, SWEEP_SUFFIX)
-        sweep_path_list.append(save_path)
-
-    # Plot the sweeps together.
-    plot_amp_sweep(PLOT_SAVE_PATH, sweep_path_list,
-                                title=EXPERIMENT_NAME,
-                                x_label="Amplitude Multiplier",
-                                y_label="Infidelity")
-
-
-if __name__ == "__main__":
-    main()

### 2020-11-14 src/spin/spin_exp3_hpsweep.py

@@ -1,49 +1,0 @@
-"""
-spin_exp3_hpsweep.py - Do a sweep of the hamiltonian parameters
-for spin_exp3.
-"""
-
-import os
-
-import numpy as np
-from rbqoc.util import (hamiltonian_args_sweep,
-                        plot_hamiltonian_args_sweep,)
-from spin_exp3 import (EVOL_ARGS, EXPERIMENT_META, EXPERIMENT_NAME, HAMILTONIAN_ARGS)
-
-# Sweep trials
-SWEEP_TRIAL_COUNT = 100
-SWEEP_MULTIPLIERS = np.linspace(0, 3, SWEEP_TRIAL_COUNT)
-HAMILTONIAN_ARGS_SWEEP = np.stack([HAMILTONIAN_ARGS] * SWEEP_TRIAL_COUNT)
-for i, multiplier in enumerate(SWEEP_MULTIPLIERS):
-    HAMILTONIAN_ARGS_SWEEP[i] *= SWEEP_MULTIPLIERS[i]
-
-# I/O
-WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
-SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
-PULSE_FILE_0 = "00000_spin_exp3.h5"
-PULSE_FILE_1 = "00001_spin_exp3.h5"
-PULSE_FILES = [PULSE_FILE_0, PULSE_FILE_1]
-PULSE_PATH_0 = os.path.join(SAVE_PATH, PULSE_FILE_0)
-PULSE_PATH_1 = os.path.join(SAVE_PATH, PULSE_FILE_1)
-PULSE_PATHS = [PULSE_PATH_0, PULSE_PATH_1]
-SWEEP_SUFFIX = "freq_sweep"
-PLOT_SAVE_FILE = "{}_{}.png".format(EXPERIMENT_NAME, SWEEP_SUFFIX)
-PLOT_SAVE_PATH = os.path.join(SAVE_PATH, PLOT_SAVE_FILE)
-
-def main():
-    sweep_path_list = list()
-    # Sweep the frequency parameter for each pulse.
-    for i, pulse_path in enumerate(PULSE_PATHS):
-        save_path = hamiltonian_args_sweep(EVOL_ARGS, HAMILTONIAN_ARGS_SWEEP,
-                                           pulse_path, SWEEP_SUFFIX)
-        sweep_path_list.append(save_path)
-
-    # Plot the sweeps together.
[…]

### 2020-11-14 src/spin/spin_exp5.py

@@ -1,165 +1,0 @@
-"""
-spin_exp5.py
-"""
-
-from argparse import ArgumentParser
-import os
-
-import autograd.numpy as anp
-from filelock import FileLock, Timeout
-import h5py
-import matplotlib.pyplot as plt
-import matplotlib.patches as mpatches
-from qoc import grape_schroedinger_discrete, evolve_schroedinger_discrete
-from qoc.standard import (
-    TargetStateInfidelity, ControlVariation, ControlNorm,
-    conjugate_transpose,
-    get_annihilation_operator,
-    get_creation_operator,
-    SIGMA_Z, SIGMA_X,
-    generate_save_file_path,
-    matrix_to_column_vector_list,
-    krons, plot_controls,
-    plot_state_population,
-    Adam, LBFGSB,
-)
-
-CORE_COUNT = 8
-os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
-os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
-
-# Define experimental constants.
-CONTROL_SAMPLING_RATE = 1.2 # GS/s
-MAX_AMP_CONTROL_0 = 2 * anp.pi * 1e-1 # GHz
-MAX_AMP_BANDWIDTH_CONTROL_0 = 2 * anp.pi * 5e-1 # GHz
-OMEGA_Q = 2 * anp.pi * 1e-2 # GHz
-
-# Define the system.
-SYSTEM_HAMILTONIAN_0 = SIGMA_Z / 2
-CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
-HAMILTONIAN_ARGS = anp.array([OMEGA_Q])
-hamiltonian = lambda controls, time: (
-    OMEGA_Q * SYSTEM_HAMILTONIAN_0
-    + controls[0] * CONTROL_HAMILTONIAN_0)
-MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
-MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
-
-# Define the optimization.
-WDIR = os.environ["ROBUST_QOC_PATH"]
-# CONTROLS_PATH = os.path.join(WDIR, "out/spin/spin_exp5/00001_spin_exp5.h5")
-# CONTROLS_PATH_LOCK = "{}.lock".format(CONTROLS_PATH)
[…]

### 2020-11-14 src/spin/spin_exp6.py

@@ -1,166 +1,0 @@
-"""
-spin_exp6.py
-"""
-
-from argparse import ArgumentParser
-import os
-
-import autograd.numpy as anp
-from filelock import FileLock, Timeout
-import h5py
-import matplotlib.pyplot as plt
-import matplotlib.patches as mpatches
-from qoc import grape_schroedinger_discrete, evolve_schroedinger_discrete
-from qoc.standard import (
-    TargetStateInfidelity, ControlVariation, ControlNorm,
-    conjugate_transpose,
-    get_annihilation_operator,
-    get_creation_operator,
-    SIGMA_Z, SIGMA_X,
-    column_vector_list_to_matrix,
-    generate_save_file_path,
-    matrix_to_column_vector_list,
-    krons, plot_controls,
-    plot_state_population,
-    Adam, LBFGSB,
-)
-
-CORE_COUNT = 8
-os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
-os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
-
-# Define experimental constants.
-CONTROL_SAMPLING_RATE = 1.2 # GS/s
-MAX_AMP_CONTROL_0 = 2 * anp.pi * 1e-1 # GHz
-MAX_AMP_BANDWIDTH_CONTROL_0 = 2 * anp.pi * 5e-1 # GHz
-OMEGA_Q = 2 * anp.pi * 1e-2 # GHz
-
-# Define the system.
-SYSTEM_HAMILTONIAN_0 = SIGMA_Z / 2
-CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
-HAMILTONIAN_ARGS = anp.array([OMEGA_Q])
-hamiltonian = lambda controls, time: (
-    OMEGA_Q * SYSTEM_HAMILTONIAN_0
-    + controls[0] * CONTROL_HAMILTONIAN_0)
-MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
-MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
-
-# Define the optimization.
-WDIR = os.environ["ROBUST_QOC_PATH"]
-# CONTROLS_PATH = os.path.join(WDIR, "out/spin/spin_exp6/00000_spin_exp6.h5")
[…]

### 2020-11-14 src/spin/spin_exp7.py

@@ -1,147 +1,0 @@
-"""
-spin_exp7.py - Spin experiment 7.
-"""
-
-from argparse import ArgumentParser
-import os
-
-import autograd.numpy as anp
-from filelock import FileLock, Timeout
-import h5py
-import matplotlib.pyplot as plt
-import matplotlib.patches as mpatches
-from qoc import grape_schroedinger_discrete, evolve_schroedinger_discrete
-from qoc.standard import (TargetStateInfidelity,
-                          conjugate_transpose,
-                          get_annihilation_operator,
-                          get_creation_operator,
-                          SIGMA_Z, SIGMA_X,
-                          generate_save_file_path,
-                          krons, plot_controls,
-                          plot_state_population,
-                          Adam, LBFGSB,)
-
-CORE_COUNT = 8
-os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
-os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
-
-# Define experimental constants.
-CONTROL_SAMPLING_RATE = 1.2 # GS/s
-MAX_AMP_CONTROL_0 = 2 * anp.pi * 1e-1 # GHz
-MAX_AMP_BANDWIDTH_CONTROL_0 = 2 * anp.pi * 5e-1 # GHz
-OMEGA_Q = 2 * anp.pi * 1e-2 # GHz
-
-# Define the system.
-SYSTEM_HAMILTONIAN_0 = SIGMA_Z / 2
-CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
-HAMILTONIAN_ARGS = anp.array([OMEGA_Q])
-hamiltonian = lambda controls, hargs, time: (
-    hargs[0] * SYSTEM_HAMILTONIAN_0
-    + controls[0] * CONTROL_HAMILTONIAN_0
-)
-MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
-MAX_CONTROL_BANDWIDTHS = anp.array((MAX_AMP_BANDWIDTH_CONTROL_0,))
-
-# Define the optimization.
-WDIR = os.environ["ROBUST_QOC_PATH"]
-GRAB_CONTROLS = False
-if GRAB_CONTROLS:
-    CONTROLS_PATH = os.path.join(WDIR, "out/spin/spin_exp0/00017_spin_exp0.h5")
-    CONTROLS_PATH_LOCK = "{}.lock".format(CONTROLS_PATH)
-    try:
[…]

### 2020-11-14 src/spin/spin_pi_corpse.py

@@ -1,204 +1,0 @@
-"""
-spin_pi_corpse.py
-"""
-
-from argparse import ArgumentParser
-import os
-
-import autograd.numpy as anp
-from filelock import FileLock, Timeout
-import h5py
-import matplotlib.pyplot as plt
-import matplotlib.patches as mpatches
-import numpy as np
-from qoc import grape_schroedinger_discrete, evolve_schroedinger_discrete
-from qoc.standard import (TargetStateInfidelity,
-                          conjugate_transpose,
-                          get_annihilation_operator,
-                          get_creation_operator,
-                          SIGMA_Z, SIGMA_X, RX,
-                          generate_save_file_path,
-                          matrix_to_column_vector_list,
-                          column_vector_list_to_matrix,
-                          matmuls,
-                          krons, plot_controls,
-                          plot_state_population,
-                          Adam, LBFGSB,)
-from qutip import mesolve, Qobj
-
-CORE_COUNT = 8
-os.environ["MKL_NUM_THREADS"] = "{}".format(CORE_COUNT)
-os.environ["OPENBLAS_NUM_THREADS"] = "{}".format(CORE_COUNT)
-
-# Define experimental constants.
-CONTROL_SAMPLING_RATE = 1.2 # GS/s
-MAX_AMP_CONTROL_0 = 2 * anp.pi * 1e-1 # GHz
-MAX_AMP_BANDWIDTH_CONTROL_0 = 2 * anp.pi * 5e-1 # GHz
-OMEGA_Q = 2 * anp.pi * 1e-2 # GHz
-
-# Define the system.
-SYSTEM_HAMILTONIAN_0 = SIGMA_Z / 2
-CONTROL_HAMILTONIAN_0 = SIGMA_X / 2
-HAMILTONIAN_ARGS = anp.array([OMEGA_Q])
-hamiltonian = lambda controls, hargs, time: (
-    hargs[0] * SYSTEM_HAMILTONIAN_0
-    + controls[0] * CONTROL_HAMILTONIAN_0)
-MAX_CONTROL_NORMS = anp.array((MAX_AMP_CONTROL_0,))
[…]

### 2020-11-14 src/spin/spin_pi_corpse_hpsweep.py

@@ -1,74 +1,0 @@
-"""
-spin_pi_corpse_hpsweep.py - Do a sweep of the hamiltonian parameters
-for spin_pi_corpse.
-"""
-
-import os
-
-import numpy as np
-from rbqoc.util import (hamiltonian_args_sweep,
-                        plot_hamiltonian_args_sweep,)
-from spin_pi_corpse import (EVOL_CONFIG, EXPERIMENT_META, EXPERIMENT_NAME, HAMILTONIAN_ARGS)
-
-EVOL_CONFIG.pop("controls", None)
-EVOL_CONFIG.pop("hamiltonian_args", None)
-
-# Sweep trials
-SWEEP_TRIAL_COUNT = 100
-HAMILTONIAN_ARGS_SWEEP = np.linspace(2e-2, 1e-1, SWEEP_TRIAL_COUNT)
-HAMILTONIAN_ARGS_SWEEP = np.expand_dims(HAMILTONIAN_ARGS_SWEEP, axis=1)
-# SWEEP_MULTIPLIERS = np.linspace(0, 3, SWEEP_TRIAL_COUNT)
-# HAMILTONIAN_ARGS_SWEEP = np.stack([HAMILTONIAN_ARGS] * SWEEP_TRIAL_COUNT)
-# for i, multiplier in enumerate(SWEEP_MULTIPLIERS):
-#     HAMILTONIAN_ARGS_SWEEP[i] *= SWEEP_MULTIPLIERS[i]
-
-
-# I/O
-WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
-SAVE_PATH_0 = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
-SAVE_PATH_1 = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, "spin_exp0")
-PULSE_FILE_0 = "00000_spin_pi_corpse.h5"
-PULSE_FILE_1 = "00015_spin_exp0.h5"
-PULSE_FILE_2 = "00017_spin_exp0.h5"
-LABELS = [
-    "CORPSE",
-    "QOC",
-    "RBQOC",
-]
-PULSE_FILES = [PULSE_FILE_0,
-               PULSE_FILE_1,
-               PULSE_FILE_2,
-]
-PULSE_PATH_0 = os.path.join(SAVE_PATH_0, PULSE_FILE_0)
[…]

### 2020-11-14 src/spin/spincnstr_hpsweep.py

@@ -1,130 +1,0 @@
-"""
-spincnstr_hpsweep.py - Do a sweep of the hamiltonian parameters
-for spincnstr.
-"""
-
-import os
-
-import h5py
-from filelock import FileLock, Timeout
-import matplotlib.pyplot as plt
-import numpy as np
-from rbqoc.util import (hamiltonian_args_sweep,
-                        plot_hamiltonian_args_sweep,)
-from spincnstr import (EVOL_CONFIG, EXPERIMENT_META, EXPERIMENT_NAME, HAMILTONIAN_ARGS)
-
-# Plot
-DPI = 1000
-
-# Sweep trials
-SWEEP_TRIAL_COUNT = 100
-SWEEP_MULTIPLIERS = np.linspace(0.9, 1.1, SWEEP_TRIAL_COUNT)
-HAMILTONIAN_ARGS_SWEEP = np.stack([HAMILTONIAN_ARGS] * SWEEP_TRIAL_COUNT)
-for i, multiplier in enumerate(SWEEP_MULTIPLIERS):
-    HAMILTONIAN_ARGS_SWEEP[i] *= SWEEP_MULTIPLIERS[i]
-
-# I/O
-WDIR_PATH = os.environ["ROBUST_QOC_PATH"]
-SAVE_PATH = os.path.join(WDIR_PATH, "out", EXPERIMENT_META, EXPERIMENT_NAME)
-SAVE_FILE_PATH = os.path.join(SAVE_PATH, "{}_hpsweep_plot.png".format(EXPERIMENT_NAME))
-PULSE_FILES = [
-    # "00033_spincnstr.h5",
-    "00034_spincnstr.h5",
-    "00035_spincnstr.h5",
-    "00036_spincnstr.h5",
-    "00037_spincnstr.h5",
-    # "00038_spincnstr.h5",
-    # "00039_spincnstr.h5",
-    "00040_spincnstr.h5",
-    "00041_spincnstr.h5",
-]
-TIMES = [
-    # 150,
-    100,
-    80,
-    60,
-    120,
-    # 70,
-    # 90,
-    110,
-    130,
-]
-PULSE_PATHS = list()
-for pulse_file in PULSE_FILES:
-    PULSE_PATHS.append(os.path.join(SAVE_PATH, pulse_file))
-SWEEP_SUFFIX = "freq_sweep"
[…]

### 2020-11-16 paper/6stochastic.tex

@@
-%% S6
 \section{Robustness to Stochastic Parameter Deviations}
-An additional source of experimental error arises from stochastic
+An additional source of experimental error arises from stochastic, time-varying
 Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
@@
 1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat 1/$f$ noise
+noise (citations needed). Analytic methods to combat flux noise
 take advantage of the low frequency characteristic and
@@
 on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The noise is generated by
-filtering white noise sampled from a standard normal distribution using a finite
+The flux noise is generated by
+filtering white noise sampled from a standard normal distribution with a finite
 impulse response filter \cite{saspweb2011}
@@
 
-We simulate successive applications of the gate constructed by each method and compute the gate error
+We simulate successive applications of the gate constructed by each method and compute the cumulative gate error
 after each application, see Figure \ref{fig:stochastic}. Both the analytic and numerical methods achieve single gate errors
 sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision, achieving a two
-order of magnitude reduction in gate error over the analytic gate at $200$
+numerical methods perform similarly in the concatenated gate application comparision. They achieve a two
+order of magnitude reduction in the total gate error over the analytic method after $200$
 gate applications $\sim 11 \mu\textrm{s}$.
-Coherent errors are a signficiant source of error in NISQ applications and these numerical techniques give
-an effective method for mitigating them.
+1/$f$ flux noise is a signficiant source of coherent errors in NISQ applications and
+these numerical techniques offer effective avenues to mitigate them.

### 2020-11-16 paper/2background.tex
full build

--- 2020-11-03 paper/2background.tex
@@
-%% S2
 \section{QOC + AL-iLQR}
-(QOC Problem Statement) Here we introduce the notation
+In this section we introduce the notation
 we will use throughout the paper,
@@
 a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schroedinger equation (TDSE)
+Schr{\"o}dinger equation (TDSE)
+\begin{equation}
+  i \hbar \frac{d}{dt} \ket{\psi} = H(u(t), t) \ket{\psi}
   \label{eq:tdse}
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
-\end{align}
+\end{equation}
 The evolution is sometimes cast with the evolution
@@
 knot points (time steps). Typical explicit integration techniques for the TDSE include
-exponential integrators and integrators of the Runge-Kutta type
+exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}
+and Runge-Kutta integrators \cite{jorgensen2011numerical}.
 
@@
 In the simplest case the functional is
+the infidelity between the inital state evolved
+to the final knot point and the target state
 $J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-the infidelity between the inital state evolved
-to the final knot point ($\ket{\psi_{N}(u(t))}$)
-and the target state ($\ket{\psi_{f}}$). In general
-$J$ is a linear combinaion of cost functions on the state, e.g.
+In general $J$ is a linear combinaion of cost functions on the state, e.g.
 forbidden-state occupation, as well as
@@
 algorithms employ automatic differentiation
-to compute first order information for the functional ($\nabla_{u} J(u)$).
-They employ a first-order optimizer to minimize $J$ with respect to $u$.
+to compute first order information for the functional $\nabla_{u} J(u)$.
+This information is used in a first-order optimizer to minimize $J$ with respect to $u$.
+This scheme lends itself to projected gradient methods
+which restrict the optimization to the constraint manifold
+\cite{clarkson2010coresets, hauswirth2016projected}.
 
-(AL-iLQR Problem Statement) The trajectory optimization
+The trajectory optimization
 literature solves a more general class of non-linear programs that resemble
-the quantum optimal control problem. The quantum optimal control
+the QOC problem. The QOC
 problem is a specific case of the linear quadratic regulator (LQR).
 LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. In the LQR formulation
-the same functional is evaluated at each knot point
+the functional is quadratic in the state. A general LQR cost
+function takes the form
 \begin{equation}
-  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
-  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
+  J_{\textrm{iLQR}} = \ell_{N}({x}_{N})
+  + \sum_{k = 0}^{N - 1} \ell_{k}({x}_{k}, u_{k})
 \end{equation}
-where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
-at knot point $k$ and final state, $u_{k}$ are the controls,
-and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
-The state is propagated using a dynamics function
-$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics. In the following
-we refer to $\ket{\psi}$ as the state, $x$ as the augmented
-state, and $u$ as the augmented controls.
-
-The advantage of the LQR formulation
-is that there exists a dynamic programming algorithm to compute the
-optimal update to the augmented controls ($u_{k}$) which minimizes the functional
-($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
+The subscript $k$ denotes the knot point, $x$ is the augmented state vector,
+and $u$ is the augmented control vector. The augmented state is propagated
+to the next knot point by the dynamics function $x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
+In the case of QOC $\ket{\psi_{k}} \subseteq x_{k}$
+and $f$ encodes the TDSE dynamics \eqref{eq:tdse}. The
+cost function $\ell_{k}$ includes contributions from the infidelity
+$1 - {\lvert \braket{\psi_{f} \lvert \psi_{k}} \rvert}^{2}$, final state deviations
+$\tilde{x}^{T}_{k} Q_{k} \tilde{x}_{k}$ where $\tilde{x}_{k} = x_{f} - x_{k}$,
+and control regularization $u^{T}_{k} R_{k} u_{k}$.
+Placing cost functions at each knot point smoothens the optimization
+landscape. In QOC realizing the final state
+is the most important objective so the cost function at the final knot point
+carries a higher realtive weight $\ell_{N} \sim N \cdot \ell_{k}$.
+The LQR formulation restricts the cost functions to operate on the augmented
+state and controls at each knot point. The advantage of the LQR formulation
+is that there exists a differentiable dynamic programming algorithm to compute the
+optimal update to the augmented controls $u_{k}$ which minimizes the cost
+function $\ell_{k}$ at each knot point. This algorithm proceeds by deriving a
 recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (see Appendix). The
+feedback law--known as the Ricatti recursion (citations needed). The
 iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
@@
 zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the functional
+method associates a penalty multiplier with the constraint
 that estimates the constraint's Lagrange multiplier.
@@
 \begin{aligned}
-    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-    &+ J_{\textrm{iLQR}}
+    J_{\textrm{AL-iLQR}} &= J_{\textrm{iLQR}}\\
+    &+ (\lambda_{N} + \frac{1}{2}I_{\mu_{N}} c_{N}(x_{N}))^{T} c_{N}(x_{N})\\
+    &+ \sum^{N - 1}_{k = 0} (\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
 \end{aligned}
 \end{equation}
-where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
+Here $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
 with $\mu_{k}$ along the diagonal.
@@
 \begin{align}
-  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
+  \lambda^{i}_{k} &\gets \max(0, \lambda^{i}_{k} + \mu^{i}_{k} c^{i}_{k}(x_{k}^{*}, u_{k}^{*}))\\
+  \mu^{i}_{k} &\gets \phi \mu^{i}_{k}
 \end{align}
-where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
+Here $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
 $i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. With this updated form of the cost
-functional there still exists a recurrence relation to calculate the optimal control
-updates, see \cite{howell2019altro}.
+and $\phi$ is a hyperparameter. An optimal control update, similar to the Ricatti
+recursion, can be formed to calculate the optimal control updates for this functional,
+see \cite{howell2019altro}.

### 2020-11-16 paper/3fluxonium.tex
full build

@@
 approximation the system Hamiltonian takes the form
-\label{eq:hamiltonian}
 \begin{align}
   H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
+  \label{eq:hamiltonian}
 \end{align}
@@
 We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \ref{eq:hamiltonian}
+to ensure the two-level approximation \eqref{eq:hamiltonian}
 remains valid. Additionally, we require that each gate achieves
@@
 are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$.
+reached $x_{N} = x_{f}$, the constraint functional takes the form
+$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
 The equality and inequality constraints on $a$ are handled
-with a bound constraint.
-%% TODO: probably want to list equations for the constraints
+with a bound constraint which takes the form
+$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
+and $0$ otherwise.
+In addition the norm of the quantum state
+is constrained to ensure discretization error is not
+exploited $c_{k} = {\lvert \braket{\psi_{k} \lvert \psi_{k}} \rvert}^{2} - 1$.
 
@@
 Converesely, pure dephasing is a reversible process.
-There is a tradeoff between the two decoherence processes. In the case of white
-noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
-is constant \cite{huang2020engineering}.
 Our device achieves its best pure dephasing
@@
 pure dephasing.
-
-%% TODO: talk about quadratic cost vs. infidelity

### 2020-11-16 paper/5static.tex
full build

@@
-%% TODO: standardize use of "paramter" vs "system parameter" language
 \section{Robustness to Static Parameter Deviations \label{sec:static}}
@@
 filtering and is frequently utilized in robust control
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
+\cite{howell2020direct, julier2004unscented, lee2013sigma, manchester2016derivative}.
 Finally, we propose the derivative method. Derivative
@@
 (\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-This is a standard perscription, though knowledge
-of the problem can typically be used to reduced the number
-of required samples. For this method the number of states in the augmented state vector
+For this method the number of states in the augmented state vector
 scales as $O(dn^{2})$ because each of $n^{2}$ initial
@@
     (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The reported gate error is the cumulative
-    gate error after each gate application.
+    successive gate applications. The cumulative
+    gate error is computed after each gate application.
   }
@@
 So, the number of states in the augmented state vector scales as
-$O(dmn^{2})$.
+$O(dmn^{2})$. Additionally, we note that the number of initial
+states and samples we have employed for each of the three
+methods follows standard perscriptions. However, knowledge
+of the problem can typically be used to reduced the number
+of required samples.
 
 To demonstrate the applicability of these techniques
-to mitigate system parameter deviations,
+to mitigate parameter deviations,
 we consider the task of achieving a single $Z/2$
@@
 
-%% TODO: how do I tie this in?
-Multiple analytic methods have also been presented to construct gates
-that mitigate the error arising from parameter deviations
+Additionally, there are multiple analytic methods 
+to mitigate errors due to parameter deviations
 including composite pulse sequences \cite{merrill2014progress},
@@
 Composite pulse sequences are derived by computing
-the erroneous rotation arising from the static parameter deviation
+the erroneous rotation arising from the parameter deviation
 with a finite order magnus expansion.
@@
 error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses.
+to arbitrary order with sufficiently many pulses \cite{merill2014progress},
+which mirrors the negative correlation between
+gate error and gate duration we observe with the derivative method.
 It is difficult to choose an appropriate composite pulse
-for comparison on the problem we study here. We
-propose comparing numerical methods to existing analytic
-techniques for future work.
+for the problem studied here,
+so we propose comparisons between composite pulses and numerical techniques
+for future work.

### 2020-11-16 paper/6stochastic.tex
full build

@@
 
-We compare the numerical methods discussed in the previous section
+We compare the analytic gate and those produced by
+the numerical methods discussed in the previous section
 on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
@@
 
-We simulate successive applications of the gate constructed by each method and compute the cumulative gate error
-after each application, see Figure \ref{fig:stochastic}. Both the analytic and numerical methods achieve single gate errors
+We simulate successive applications of the gate constructed by each method
+and compute the cumulative gate error
+after each application, see Figure \ref{fig:stochastic}. Both the analytic
+and numerical methods achieve single gate errors
 sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
 numerical methods perform similarly in the concatenated gate application comparision. They achieve a two
-order of magnitude reduction in the total gate error over the analytic method after $200$
+order of magnitude cumulative gate error reduction over the analytic method after $200$
 gate applications $\sim 11 \mu\textrm{s}$.

### 2020-11-16 paper/a3derivative.tex
removed old dirs
full build

@@
     for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 60$ns and are averaged over each initial state.}
+    at the end of the gate's duration $t_{N} = 60$ns and are averaged over the four state derivatives.}
   \label{tab:dnorm}
@@
 
-The dynamics for the derivative methods can be integrated efficiently using exponential integrators
-\cite{berland2005solving, einkemmer2017performance}.
+The dynamics for the derivative methods can be integrated efficiently using exponential integrators.
 General exponential integrators break the dynamics into a linear term and a non-linear term. For example, consider

### 2020-11-18 paper/1introduction.tex
full build
brian edits

@@
 Analytic and numerical techniques for QOC have been remarkably successful in
-designing high fidelity gates (citations needed). In addition to the high
-simulated fidelities which can be obtained with these techniques, it is desirable for the
-resultant quantum state trajectory to be insensitive to experimental errors
-such as parameter drift, noise, and finite control resolution. Quantum state
-trajectories that are insensitive to these effects are said to be robust.
-In line with this objective, recent work has sought to mitigate decoherence
-and errors due to parameter deviations.
+designing high fidelity gates (citations needed). 
+While these methods are able to predict system behavior with high accuracy, they tend 
+to be sensitive to experimental errors such as parameter drift, noise, and finite control 
+resolution. 
+% In addition to the high
+% simulated fidelities which can be obtained with these techniques, it is desirable for the
+% resultant quantum state trajectory to be insensitive to experimental errors
+% such as parameter drift, noise, and finite control resolution. Quantum state
+% trajectories that are insensitive to these effects are said to be robust.
+% In line with this objective, 
+In order to address these issues,
+recent work has sought to mitigate decoherence
+and errors due to parameter deviations
 Considering the dynamical and geometric phases of quantum state
@@
 
-In this work we develop robust trajectory optimization techniques to mitigate
+In this work we mitigate
 errors arising from longitudinal relaxation, parameter deviations, and 1/$f$ flux noise
-on the fluxonium qubit presented in \cite{zhang2020universal}. We perform these optimization using the ALTRO
-method \cite{howell2019altro} which combines the iterative, indirect shooting
-flavor of methods such as GOAT \cite{machnes2015gradient}, GRAPE
-\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}
-with the augmented Lagrangian method. The augmented Lagrangian method
-allows us to enforce arbitrary, simultaneous constraints and achieve
-fast convergence without being restricted to the constraint
-manifold. To mitigate errors due to longitudinal
+on the fluxonium qubit \cite{zhang2020universal} by solving a robust trajectory 
+optimization problem. 
+Leveraging recent advances in trajectory optimization within the field of robotics, we
+solve this optimization problem using the ALTRO solver, which uses iterative LQR 
+(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
+methods such as GOAT \cite{machnes2015gradient}, GRAPE
+\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
+an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
+each time step \cite{howell2019altro}. 
+% The augmented Lagrangian method
+% allows us to enforce arbitrary, simultaneous constraints and achieve
+% fast convergence without being restricted to the constraint
+% manifold. 
+To mitigate errors due to longitudinal
 relaxation, we perform time-optimal control and
@@
 not pay the increased computational cost of integrating a master equation.
-Additionally, we consider three robust
-robust trajectory opitmization techniques to make the
-quantum state trajectory robust to parameter deviations.
-The first is the sampling method, which has been applied
-previously in the context of QOC
-\cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
-The second is the unscented sampling
-method, which derives from the unscented transformation used
-for non-linear Kalman filtering
-\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
-We propose a third method, which uses derivative information
-of parameter deviations with respect to the quantum state trajectory.
-We make this method efficient by employing mixed-mode differentiation.
+Additionally, to make the quantum state trajectory robust to parameter deviations,
+we consider three different approaches to making the optimized trajectory less 
+sensitive to model uncertainty:
+% we consider three robust
+% robust trajectory opitmization techniques to make the
+% quantum state trajectory robust to parameter deviations.
+\begin{enumerate}
+  \item A sampling method, similar to the work of \cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
+  \item An unscented sampling method, adapted from the unscented tranform used in the 
+  state estimation community \cite{howell2020direct, julier2004unscented, lee2013sigma, manchester2016derivative}.
+  \item A derivative-penalization method, which uses efficient mixed-mode differentiation
+  to compute derivative information of parameter deviations with respect to the 
+  quantum state trajectory.
+\end{enumerate}
+% The first is the sampling method, which has been applied
+% previously in the context of QOC
+%%TODO: the citations for unscented sampling
+%% and sampling are the same as in section V
+% \cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
+% The second is the unscented sampling
+% method, which derives from the unscented transformation used
+% for non-linear Kalman filtering
+% \cite{howell2020direct, julier2004unscented, lee2013sigma, manchester2016derivative}.
+% We propose a third method, which uses derivative information
+% of parameter deviations with respect to the quantum state trajectory.
+% We make this method efficient by employing mixed-mode differentiation.
 
-This paper is organized as follows.
+This paper is organized as follows. \todo{use section labels?}
 First, we introduce the ALTRO method in the context of QOC.

### 2020-11-18 paper/2background.tex
brian edits

@@
 
-The trajectory optimization
-literature solves a more general class of non-linear programs that resemble
-the QOC problem. The QOC
-problem is a specific case of the linear quadratic regulator (LQR).
-LQR is so called because the dynamics are linear in the state and
-the functional is quadratic in the state. A general LQR cost
-function takes the form
-\begin{equation}
-  J_{\textrm{iLQR}} = \ell_{N}({x}_{N})
-  + \sum_{k = 0}^{N - 1} \ell_{k}({x}_{k}, u_{k})
-\end{equation}
-The subscript $k$ denotes the knot point, $x$ is the augmented state vector,
-and $u$ is the augmented control vector. The augmented state is propagated
-to the next knot point by the dynamics function $x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-In the case of QOC $\ket{\psi_{k}} \subseteq x_{k}$
-and $f$ encodes the TDSE dynamics \eqref{eq:tdse}. The
-cost function $\ell_{k}$ includes contributions from the infidelity
-$1 - {\lvert \braket{\psi_{f} \lvert \psi_{k}} \rvert}^{2}$, final state deviations
-$\tilde{x}^{T}_{k} Q_{k} \tilde{x}_{k}$ where $\tilde{x}_{k} = x_{f} - x_{k}$,
-and control regularization $u^{T}_{k} R_{k} u_{k}$.
-Placing cost functions at each knot point smoothens the optimization
-landscape. In QOC realizing the final state
-is the most important objective so the cost function at the final knot point
-carries a higher realtive weight $\ell_{N} \sim N \cdot \ell_{k}$.
-The LQR formulation restricts the cost functions to operate on the augmented
-state and controls at each knot point. The advantage of the LQR formulation
-is that there exists a differentiable dynamic programming algorithm to compute the
-optimal update to the augmented controls $u_{k}$ which minimizes the cost
-function $\ell_{k}$ at each knot point. This algorithm proceeds by deriving a
-recurrence relation between knot points $k$ and $k + 1$ for the optimal
-feedback law--known as the Ricatti recursion (citations needed). The
-iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-and applies the Ricatti recursion to all knot points on multiple
-executions.
+Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
+and solved using any of the state-of-the-art solvers developed by the robotics community.
+Trajectory optimization problems are typically of the following form: 
+\begin{mini}[2]
+    {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell(x_k,u_k) }{}{}
+    \addConstraint{x_{k+1} = f(x_k,u_k)}
+    \addConstraint{g_k(x_k,u_k)}{\leq 0}
+    \addConstraint{h_k(x_k,u_k)}{=0}
+    \label{opt:discrete_trajopt},
+\end{mini}
+where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^n$ and
+$u_k \in \R^m$ are the state and input control variables, $f(x_k,u_k)$ is the discrete
+dynamics function, and $g_k(x_k,u_k)$ and $h_k(x_k,u_k)$ are the inequality
+and equality constraints, potentially including initial and final conditions,
+at time step $k$.
 
-In order to incorporate constraints we employ
-the augmented Lagrangian method. Constraints are contributions
-to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-zero or negative when the constraint is satisfied. The AL-iLQR
-method associates a penalty multiplier with the constraint
-that estimates the constraint's Lagrange multiplier.
-The algorithm updates the penalty multiplier between
-iLQR executions. In this scheme the functional takes the form
-\begin{equation}
-\begin{aligned}
-    J_{\textrm{AL-iLQR}} &= J_{\textrm{iLQR}}\\
-    &+ (\lambda_{N} + \frac{1}{2}I_{\mu_{N}} c_{N}(x_{N}))^{T} c_{N}(x_{N})\\
-    &+ \sum^{N - 1}_{k = 0} (\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-\end{aligned}
-\end{equation}
-Here $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-with $\mu_{k}$ along the diagonal.
-$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-\begin{align}
-  \lambda^{i}_{k} &\gets \max(0, \lambda^{i}_{k} + \mu^{i}_{k} c^{i}_{k}(x_{k}^{*}, u_{k}^{*}))\\
-  \mu^{i}_{k} &\gets \phi \mu^{i}_{k}
-\end{align}
-Here $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-$i$ indicates the $i$-th constraint functional,
-and $\phi$ is a hyperparameter. An optimal control update, similar to the Ricatti
-recursion, can be formed to calculate the optimal control updates for this functional,
-see \cite{howell2019altro}.
+Many techniques have been proposed for solving \eqref{opt:discrete_trajopt}. Standard 
+methods include direction collocation \cite{Hargraves87} and differential-dynamic programming
+(DDP) \cite{Mayne1966a}. Recent state-of-the-art solvers, such as ALTRO \cite{howell2019altro},
+have combined principles from both of these approaches.
+
+ALTRO uses iterative LQR (iLQR) \cite{Li2004a} as the internal solver of an augmented 
+Lagrangian method (ALM). iLQR solves an unconstrained trajectory optimization problem 
+using a backward Riccati recursion to derive a closed-loop linear feedback law about the current 
+trajectory. By simulating the system forward with the feedback law, the trajectory is 
+brought closer to the (locally) optimal trajectory. DDP-based solvers such as iLQR
+are popular since they are very computationally efficient, are always dynamically feasible,
+and provide a closed-loop control policy about the optimal trajectory for free; however,
+standard implementations have no ability to deal with nonlinear equality and inequality 
+constraints. ALM handles constraints by successively solving unconstrained minimization 
+problems of the form:
+\begin{mini}[2]
+    {z}{f(z) + \lambda^T c(z) + \half c(z)^T I_\mu c(z)}{}{}
+    \label{opt:alm}
+\end{mini}
+where $f(z)$ is the objective function, $c(z) : \R^n \mapsto \R^p$ is the constraint 
+function, $\lambda \in \R^p$ are the Lagrange multipliers, and $I_\mu$ is a diagonal matrix
+of penalty weights, whose magnitude depend on whether the constraint is active or inactive.
+For ALTRO, $f(z)$ is the objective function of \eqref{opt:discrete_trajopt}, $c(z)$ is the 
+concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
+across all time steps. After minimizing \eqref{opt:alm} using iLQR, the penalty terms and 
+Lagrange multipliers are updated, and the process repeats until convergence.
+
+ALM converges superlinearly but tends to exhibit slow constraint convergence near the optimal 
+solution due to poor numerical conditioning. To address this shortcoming, ALTRO provides a 
+solution-polishing phase that takes 1-2 Newton steps on the active constraint set to provide 
+machine-precision constraint satisfaction. For more information on the details of the ALTRO
+solver, see \cite{howell2019altro,Jackson2020altroc}.
+
+% For the QOC problem, the state $x$ consists of \todo{fill this out} while the control input
+% $u$ includes the flux drive amplitude. The cost functions $\ell_N$
+% The trajectory optimization
+% literature, driven primarily by the robotics community,
+% solves a more general class of non-linear programs that resemble
+% the QOC problem. The QOC
+% problem is a specific case of the linear quadratic regulator (LQR).
+% LQR is so called because the dynamics are linear in the state and
+% the functional is quadratic in the state. A general LQR cost
+% function takes the form
+% \begin{equation}
+%   J_{\textrm{iLQR}} = \ell_{N}({x}_{N})
+%   + \sum_{k = 0}^{N - 1} \ell_{k}({x}_{k}, u_{k})
+% \end{equation}
+% The subscript $k$ denotes the knot point, $x$ is the augmented state vector,
+% and $u$ is the augmented control vector. The augmented state is propagated
+% to the next knot point by the dynamics function $x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
+% In the case of QOC $\ket{\psi_{k}} \subseteq x_{k}$
+% and $f$ encodes the TDSE dynamics \eqref{eq:tdse}. The
+% cost function $\ell_{k}$ includes contributions from the infidelity
+% $1 - {\lvert \braket{\psi_{f} \lvert \psi_{k}} \rvert}^{2}$, final state deviations
+% $\tilde{x}^{T}_{k} Q_{k} \tilde{x}_{k}$ where $\tilde{x}_{k} = x_{f} - x_{k}$,
+% and control regularization $u^{T}_{k} R_{k} u_{k}$.
+% Placing cost functions at each knot point smoothens the optimization
+% landscape. In QOC realizing the final state
+% is the most important objective so the cost function at the final knot point
+% carries a higher realtive weight $\ell_{N} \sim N \cdot \ell_{k}$.
+% The LQR formulation restricts the cost functions to operate on the augmented
+% state and controls at each knot point. The advantage of the LQR formulation
+% is that there exists a differentiable dynamic programming algorithm to compute the
+% optimal update to the augmented controls $u_{k}$ which minimizes the cost
+% function $\ell_{k}$ at each knot point. This algorithm proceeds by deriving a
+% recurrence relation between knot points $k$ and $k + 1$ for the optimal
+% feedback law--known as the Ricatti recursion (citations needed). The
+% iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
+% and applies the Ricatti recursion to all knot points on multiple
+% executions.
+
+% In order to incorporate constraints we employ
+% the augmented Lagrangian method. Constraints are contributions
+% to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
+% zero or negative when the constraint is satisfied. The AL-iLQR
+% method associates a penalty multiplier with the constraint
+% that estimates the constraint's Lagrange multiplier.
+% The algorithm updates the penalty multiplier between
+% iLQR executions. In this scheme the functional takes the form
+% \begin{equation}
+% \begin{aligned}
+%     J_{\textrm{AL-iLQR}} &= J_{\textrm{iLQR}}\\
+%     &+ (\lambda_{N} + \frac{1}{2}I_{\mu_{N}} c_{N}(x_{N}))^{T} c_{N}(x_{N})\\
+%     &+ \sum^{N - 1}_{k = 0} (\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
+% \end{aligned}
+% \end{equation}
+% Here $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
+% with $\mu_{k}$ along the diagonal.
+% $\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
+% \begin{align}
+%   \lambda^{i}_{k} &\gets \max(0, \lambda^{i}_{k} + \mu^{i}_{k} c^{i}_{k}(x_{k}^{*}, u_{k}^{*}))\\
+%   \mu^{i}_{k} &\gets \phi \mu^{i}_{k}
+% \end{align}
+% Here $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
+% $i$ indicates the $i$-th constraint functional,
+% and $\phi$ is a hyperparameter. An optimal control update, similar to the Ricatti
+% recursion, can be formed to calculate the optimal control updates for this functional,
+% see \cite{howell2019altro}.

### 2020-11-18 paper/3fluxonium.tex
brian edits

@@
 
+\todo{I highly suggest writing out the optimization problem explicitly here. I've started it 
+for you below. You can reference the individual lines as you explain your problem setup.
+Also explicitly spell out what your state and control vectors are.}
+{\color{red}
+\begin{mini!}[2]
+    {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k} + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{}
+    \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
+    \addConstraint{x_1 }{= x_\text{init}} \label{eq:init_con}
+    \addConstraint{x_N}{= x_f} \label{eq:term_con}
+\end{mini!}
+}
+
 In addition to imposing optimization constraints that
@@
 The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
+$\pdv*[2]{a}{t}$ which is contained in the
 augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
+$\pdv*{a}{t}$, proportional $a$, and integral $\int a$
 flux amplitude terms
@@
 are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$, the constraint functional takes the form
+reached $x_{N} = x_{f}$, the constraint functional 
+\todo{not sure what the convention is in your community, but for us a functional is a 
+function of functions (or trajectories). This would just be a function. Also, shouldn't 
+this just be $x_N = x_f$ and $\abs{x_k} \le b$? Why are you squaring the functions?}
+ takes the form
 $c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
@@
 exploited $c_{k} = {\lvert \braket{\psi_{k} \lvert \psi_{k}} \rvert}^{2} - 1$.
+\todo{is this last constraint supposed to be an inequality constraint?}
 
@@
 pure dephasing.
+
+\todo{maybe reference the next Sections as you mention these points?}

### 2020-11-18 paper/4longitude.tex
paper updates
removed old dirs
full build
brian edits

@@
-%% F1
 \begin{figure*}[ht]
-  \label{fig:longitude}
+
   \begin{subfigure}{.315\textwidth}
@@
     (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The reported gate error is the cumulative
-    gate error after each gate application.
+    for successive gate applications. The cumulative
+    gate error is computed after each gate application.
     The numerically optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
@@
   }
+  \label{fig:longitude}
 \end{figure*}
 
-%% S4
 \section{Longitudinal Relaxation Awareness}
@@
 the one we study here.
+
+\todo{again, I'd recommend writing out either the entire optimization problem or at least
+point to the parts of the optimization problem defined in the previous section you had to 
+change. Just make it a little more concrete.}
 
@@
 \begin{figure*}[ht]
-  \label{fig:static}
   \begin{subfigure}{.315\textwidth}
@@
   }
+  \label{fig:static}
 \end{figure*}
@@
 quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the numerical gates for extend computations
+The low gate errors achieved by the numerical gates \todo{typo / awkward sentence} 
+for extend computations
 are critical for noisy, intermediate-scale quantum (NISQ) applications.

### 2020-11-18 paper/5static.tex
brian edits

@@
 we mitigate errors arising from the drift and finite measurement
-precision of the qubit frequency $\tilde{f_{q}} = f_{q} \pm \sigma_{f_{q}}$.
+precision of the qubit frequency $\tilde{f_{q}} = f_{q} \pm \sigma_{f_{q}}$ \todo{maybe 
+reference \eqref{eq:hamiltonian} here?}.
 We consider three robust control techniques.
@@
 
+\todo{I think you can make this more concrete and simpler to understand by just defining
+the new augmented state vector and explaining you stack multiple copies of the state vector
+together, each evolving with a different static parameter. In this case there's only one 
+parameter being changed, so I'd just state it explicitly.}
 In the sampling method, sample states evolve under a
@@
 For this method the number of states in the augmented state vector
-scales as $O(dn^{2})$ because each of $n^{2}$ initial
+scales as $O(dn^{2})$ \todo{is this right?}
+because each of $n^{2}$ initial
 states is represented by $2d$ samples. $n$ is both the dimension of the Hilbert space
@@
 This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensurnes the points
+first and second moments of the distribution and ensures the points
 lie on the $\sqrt{2n}$\textsuperscript{th}
@@
 The gate error of each sample state is penalized as in the sampling method.
+\todo{I think this section needs some more detailed explanation.}
 
@@
 \end{align}
+\todo{inconsistent use of $\hbar$ vs $h$.}
 Exponential integrators that account for the non-linear
@@
 of the problem can typically be used to reduced the number
-of required samples.
+of required samples. \todo{I think this also needs some more explanation. It's not 
+clear exactly what you're doing. You're penalizing the derivatives of the dynamics with
+respect to the uncertain parameter in the objective, so why do you need to augment the 
+state vector? Or are you adding the derivatives to the state vector so that you have 
+standard quadratic penalty terms in the objective? Again, I think this confusion would 
+be easily cleared up by explicitly writing down your new dynamics, objective function, 
+and any additional constraints you impose on the system. Since this is really the meat of 
+the paper, it's worth spending the time to make sure this is clear and easy to understand,
+especially for those who aren't very familiar with trajectory optimization or optimization 
+in general. Also, I think you should explicitly state the size of the problem you're
+actually solving. Might be worth putting in a table, along with some timing results, to 
+give an idea of how well these methods scale.}
 
@@
 We compare the numerical methods
-to an analytically derived $Z/2$ gate, see Figure \ref{fig:static}a. This gate corresponds to
+to an analytically derived $Z/2$ gate, see Figure \ref{fig:static}a 
+\todo{{\LaTeX} suggetion: don't hard-code the sublabels. You can directly reference them.}. 
+This gate corresponds to
 idling at the flux frustration point $a = 0$. The analytic gate

### 2020-11-18 paper/7conclusion.tex
removed old dirs
full build
brian edits

@@
 The numerical techniques we have studied here can be used to perform phase gates in arbitrary times,
-which will be critical for synchronizing multi-qubit systems. Additionally,
-interleaving the error models with existing
+which will be critical for synchronizing multi-qubit systems 
+\todo{it might be worth mentioning the scaling properties of your method
+here, since you suggest scaling it up to systems with more states. How may
+the different methods you suggest perform at scale? Maybe there will be an
+advantage to using the sampling-based methods as the problems grow? Also you
+don't mention the sampling methods in the conclusion? The derivative methods only
+outperformed the sampling methods on the simpler of your 2 test cases,
+so I'd be hesitant to imply that the derivative method is the better choice}. 
+Additionally, interleaving the error models with existing
 characterization methods will improve their effectiveness in experiments.
 These techniques will be employed to achieve the low gate errors
-required for fault-tolerant quantum computing applications.
+required for fault-tolerant quantum computing applications. Our
+implementation of the techniques described in this work is available
+at \url{https://github.com/SchusterLab/rbqoc}.

### 2020-11-18 paper/a2unscented.tex
full build
brian edits

@@
 First, we remark that ALTRO does not natively support complex numbers, so our optimization
-is performed under an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-\cite{leung2017speedup}
+is performed under an isomorphism 
+$\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
+\cite{leung2017speedup}. 
+\todo{this should probably be mentioned explicitly in the main body of the paper. Also 
+make sure you're being consistent on your definition of $n$. The definition here doesn't seem 
+to match your definition other places, especially when you start augmenting the dynamics.}
 \begin{equation}
@@
   \bar{\Psi}_{1} &= \frac{1}{4n + 2d} \sum_{i = 1}^{4n + 2d} \Psi_{1}^{i}\\
-  P_{1} &= \frac{1}{\beta^{2}(4n + 2d)} \sum_{i = 1}^{4n + 2d}
+  P_{1} &= \frac{1}{2 \beta^{2}} \sum_{i = 1}^{4n + 2d}
   (\Psi^{i}_{1} - \bar{\Psi}_{1})^{T}(\Psi^{i}_{1} - \bar{\Psi}_{1})
@@
 The sigma points are then resampled and propagated to the next knot point using
-\eqref{eq:uupdate}, \eqref{eq:unormalize}, and \eqref{eq:upropagate}.
+\eqref{eq:uupdate}, \eqref{eq:unormalize}, and \eqref{eq:upropagate}. Our
+choice of sigma points follows the prescription in equation 11 of \cite{julier2004unscented}.
+Other prescriptions may be chosen to reduce the number of required sigma points \cite{julier2002reduced}.

### 2020-11-20 paper/1introduction.tex
minor edits

@@
 to encode experimental errors in optimization objectives, improving
-experimental performance (citations appreciated).
+experimental performance \cite{Zhou97,Morimoto00,Manchester18}.
 In this work we employ robust control techniques to mitigate
@@
 Analytic and numerical techniques for QOC have been remarkably successful in
-designing high fidelity gates (citations needed). 
+designing high fidelity gates \todo{citations needed}. 
 While these methods are able to predict system behavior with high accuracy, they tend 

### 2020-11-20 paper/2background.tex
minor edits

@@
 algorithms employ automatic differentiation
-to compute first order information for the functional $\nabla_{u} J(u)$.
-This information is used in a first-order optimizer to minimize $J$ with respect to $u$.
-This scheme lends itself to projected gradient methods
-which restrict the optimization to the constraint manifold
+to compute gradients of the functional $\nabla_{u} J(u)$,
+which can easily be used to implement first-order optimization methods
 \cite{clarkson2010coresets, hauswirth2016projected}.
@@
 Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
-and solved using any of the state-of-the-art solvers developed by the robotics community.
+and solved using a variety of specialized solvers developed by the robotics community \cite{Schulman13,Tedrake16,Hereid2017FROST,howell2019altro}.
 Trajectory optimization problems are typically of the following form: 
@@
 are popular since they are very computationally efficient, are always dynamically feasible,
-and provide a closed-loop control policy about the optimal trajectory for free; however,
+and provide a locally stabilizing closed-loop control policy about the optimal trajectory. However,
 standard implementations have no ability to deal with nonlinear equality and inequality 
@@
 where $f(z)$ is the objective function, $c(z) : \R^n \mapsto \R^p$ is the constraint 
-function, $\lambda \in \R^p$ are the Lagrange multipliers, and $I_\mu$ is a diagonal matrix
-of penalty weights, whose magnitude depend on whether the constraint is active or inactive.
+function, $\lambda \in \R^p$ is a Lagrange multiplier estimate, and $I_\mu$ is a diagonal matrix
+of penalty weights, $\mu$, whose magnitudes depend on whether the corresponding constraint is active or inactive.
 For ALTRO, $f(z)$ is the objective function of \eqref{opt:discrete_trajopt}, $c(z)$ is the 
 concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
-across all time steps. After minimizing \eqref{opt:alm} using iLQR, the penalty terms and 
-Lagrange multipliers are updated, and the process repeats until convergence.
+across all time steps. After minimizing \eqref{opt:alm} using iLQR, the Lagrange multiplier estimates are updated according to,
+\begin{equation}
+	\lambda \gets \lambda + \mu c(z) ,
+\end{equation}
+the penalty terms are updated, and the process repeats until convergence.
 

### 2020-11-20 paper/3fluxonium.tex
minor edits

@@
 constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
+To ensure gates may be concatenated arbitrarily without
 inducing AWG ringing due to high-frequency transitions,

### 2020-11-24 paper/1introduction.tex
progress on todos

--- 2020-11-18 paper/1introduction.tex
@@
 In order to address these issues,
-recent work has sought to mitigate decoherence
-and errors due to parameter deviations
+recent work has sought to mitigate
+errors due to parameter deviations.
 Considering the dynamical and geometric phases of quantum state
@@
 Floquet techniques have been experimentally demonstrated to simultaneously mitigate
-decoherence due to longitudinal relaxation and pure dephasing
+longitudinal relaxation and pure dephasing decoherence
 \cite{huang2020engineering, mundada2020floquet}.
@@
 \begin{enumerate}
-  \item A sampling method, similar to the work of \cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
+\item A sampling method, similar to the work of \cite{carvalho2020error, reinhold2019controlling,
+  rembold2020introduction}.
   \item An unscented sampling method, adapted from the unscented tranform used in the 
-  state estimation community \cite{howell2020direct, julier2004unscented, lee2013sigma, manchester2016derivative}.
+    state estimation community \cite{howell2020direct, julier2004unscented,
+      lee2013sigma, manchester2016derivative}.
   \item A derivative-penalization method, which uses efficient mixed-mode differentiation
@@
 
-This paper is organized as follows. \todo{use section labels?}
-First, we introduce the ALTRO method in the context of QOC.
+This paper is organized as follows.
+First, we introduce the ALTRO method in the context of QOC \ref{sec:background}.
 We describe realistic constraints for the fluxonium and
-map them to the ALTRO method. Then, we
+map them to the ALTRO method \ref{sec:fluxonium}. Then, we
 outline a method for making the optimization aware of longitudinal
-relaxation. Next, we outline three methods for achieving
-robustness to static parameter deviations. Finally,
-we employ the robust control techniques to mitigate 1/$f$ flux noise.
+relaxation \ref{sec:longitude}. Next, we outline three methods for achieving
+robustness to static parameter deviations \ref{sec:static}. Finally,
+we employ the robust control techniques to mitigate 1/$f$ flux noise
+\ref{sec:stochastic}.

### 2020-11-24 paper/2background.tex
progress on todos

--- 2020-11-18 paper/2background.tex
@@
-\section{QOC + AL-iLQR}
+\section{QOC + AL-iLQR \label{sec:background}}
 In this section we introduce the notation
@@
 \end{equation}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
+%% The evolution is sometimes cast with the evolution
+%% of a density matrix under the Lindblad master equation to
+%% model the decoherence of the state explicitly.
+The Hamiltonian
 has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
@@
 to the final knot point and the target state
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
+$J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u(t))} \rvert}^{2}$.
 In general $J$ is a linear combinaion of cost functions on the state, e.g.

### 2020-11-24 paper/3fluxonium.tex
progress on todos

--- 2020-11-16 paper/3fluxonium.tex
@@
-\section{QOC on the Fluxonium}
+\section{QOC on the Fluxonium \label{sec:fluxonium}}
 In the following we study
@@
 
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
-constraints that improve the experimental realization of the control pulse.
-To ensures gates may be concatenated arbitrarily without
+The optimization problem takes the form:
+\begin{mini!}[2]
+    {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k} + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{}
+    \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
+    \addConstraint{x_1}{= x_\text{init}} \label{eq:init_con}
+    \addConstraint{x_N}{= x_f} \label{eq:term_con}
+\end{mini!}
+where the augmented state and controls are given by:
+\begin{equation}
+  x = \begin{bmatrix} \psi^{(1)} \\ \psi^{(2)} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
+  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
+\end{equation}
+The initial and final conditions are determined by the
+desired unitary transformation, which requires two quantum states
+to specify $\psi^{(1)}$, $\psi^{(2)}$.
+The ALTRO implementation we use does not currently support complex numbers so
+we compute in the isomorphism $\mathcal{H}(\mathbb{R}^{2n}) \cong \mathcal{H}(\mathbb{C}^{n})$
+given in \cite{leung2017speedup}:
+\begin{equation}
+  H \psi \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
+  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
+\end{equation}
+The dynamics function $f(x_{k}, u_{k})$ integrates the TDSE dynamics for the quantum states
+\eqref{eq:tdse}, \eqref{eq:hamiltonian}
+and integrates the moments of the flux drive amplitude. Exposing lower order moments of the flux
+drive amplitude allows us to penalize their norms, smoothing the flux drive amplitude
+and mitigating AWG ringing due to high frequency transitions.
+The matrices $Q_{k}$ and $R_{k}$ define the penalty metric. We choose $Q$
+to be diagonal because it is computationally efficient. This corresponds
+to penalizing phase differences between the quantum states, although
+other phase-insensitive metrics such as infidelity may be employed.
+The final state transfer is the most important objective which is
+encoded in the realtive metric weight $Q_{N} \sim N \cdot Q_{k}$.
+%%, $k \in \{1, ..., N - 1\}$.
+
+Additionally, we impose constraints that reflect limitations
+of the apparatus and ones that improve the experimental realization
+of the control pulse. To ensure gates may be concatenated arbitrarily without
 inducing AWG ringing due to high-frequency transitions,
@@
 remains valid. Additionally, we require that each gate achieves
-the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
-In addition to these constraints we penalize the norm
-of the first and second derivatives of the flux amplitude to
-ensure its smoothness, and in doing so mitigate AWG ringing.
-The optimization is performed over the second derivative of the flux amplitude
-$\frac{d^{2}}{dt^{2}} a$ which is contained in the
-augmented control vector. The first derivative
-$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
-flux amplitude terms
-are contained in the augmented state vector. They are obtained from
-the second derivative of the flux amplitude by
-integration in the dynamics function.
+the desired state transition $\psi_{N} = \psi_{f}$.
 Both the zero net flux and target quantum state constraint
 are then handled by ensuring the target augmented state is
-reached $x_{N} = x_{f}$, the constraint functional takes the form
+reached $x_{N} = x_{f}$, the constraint function
+takes the form
 $c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
@@
 and $0$ otherwise.
-In addition the norm of the quantum state
-is constrained to ensure discretization error is not
-exploited $c_{k} = {\lvert \braket{\psi_{k} \lvert \psi_{k}} \rvert}^{2} - 1$.
+In addition the norm of the quantum states
+are constrained to $1$ to ensure discretization error is not
+exploited $c_{k} = x_{k}^{T}x_{k} - 1$.
 
@@
 to maximize the longitudinal relaxation time directly
-by tuning the flux amplitude and employ robust control techniques to mitigate
-pure dephasing.
+by tuning the flux amplitude \ref{sec:longitude} and employ robust control techniques to mitigate
+pure dephasing \ref{sec:static}, \ref{sec:stochastic}.

### 2020-11-24 paper/4longitude.tex
progress on todos

@@
 
-\section{Longitudinal Relaxation Awareness}
+\section{Longitudinal Relaxation Awareness \label{sec:longitude}}
 The longitudinal relaxation time $T_{1}$ varies with

### 2020-11-24 paper/6stochastic.tex
brian edits
progress on todos

@@
-\section{Robustness to Stochastic Parameter Deviations}
+\section{Robustness to Stochastic Parameter Deviations \label{sec:stochastic}}
 An additional source of experimental error arises from stochastic, time-varying
@@
 similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
+methods require no modification from the static case. \todo{can you provide a little more 
+detail on how this noise factors into the dynamics? Are you just adding some white 
+noise to $a(t)$? Are you also considering uncertainty in $f_q$ at the same time?}
 

### 2020-11-24 paper/a2unscented.tex
progress on todos

@@
 In this section we outline the full unscented sampling procedure.
-First, we remark that ALTRO does not natively support complex numbers, so our optimization
-is performed under an isomorphism 
-$\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-\cite{leung2017speedup}. 
-\todo{this should probably be mentioned explicitly in the main body of the paper. Also 
-make sure you're being consistent on your definition of $n$. The definition here doesn't seem 
-to match your definition other places, especially when you start augmenting the dynamics.}
-\begin{equation}
-  H \psi \cong \begin{pmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{pmatrix}
-  \begin{pmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{pmatrix}
-\end{equation}
-
 The unscented sampling method concerns a quantum state $\psi \in \mathbb{R}^{2n}$ with

### 2020-11-25 paper/0abstract.tex
brian edits

@@
 \author{Brian Jackson}
-\author{Zac Manchester}
+\author{Zachary Manchester}
 \affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
+  Robotics Institute, Carnegie Mellon University, Pittsburgh, Pennsylvania 15213, USA
+  % Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
 }
@@
   %% - few sentences about primary results
-  The ability to engineer high fidelity gates on quantum processors in the presence of
+  The ability to engineer high-fidelity gates on quantum processors in the presence of
   systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
   Quantum optimal control techniques have proven effective in experimentally
-  realizing high fidelity gates, but they require exquisite calibration to be performant.
-  We apply robust trajectory optimization techniques to supress gate errors arising from system
+  realizing high-fidelity gates, but they require exquisite calibration to be performant.
+  We apply robust trajectory optimization techniques to suppress gate errors arising from system
   parameter deviations and noise.
   We propose a method that takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by utilizing mixed-mode differentation.
+  computational efficiency by utilizing mixed-mode differentiation.
   Additionally, completely modeling decoherence effects due to longitudinal relaxation requires
@@
   and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
-  We demonstrate these techniques numerically on a fluxonium qubit with realisitc
+  We demonstrate these techniques numerically on a fluxonium qubit with realistic
   experimental parameters and constraints,

### 2020-11-25 paper/1introduction.tex

@@
 diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}. For quantum computation
+\cite{sorensen2018quantum}.
+For quantum computation
 QOC techniques are employed to achieve high fidelity gates
 while adhering to experimental constraints.
-The decision variables of the optimization problem are the time-dependent control
-parameters, particular to the quantum system, that govern its evolution.
-Experimental errors may cause the system evolution to deviate from that predicted in
+Experimental errors may cause the system's trajectory to deviate from that predicted in
 optimization, hampering performance.
-The field of classical control theory has developed robust control techniques
-to encode experimental errors in optimization objectives, improving
-experimental performance (citations appreciated).
+Robust control techniques 
+address this issue by encoding
+experimental errors in optimization objectives, yielding performance
+garauntees in the presence of model uncertainty \cite{Zhou97,Morimoto00,Manchester18}.
 In this work we employ robust control techniques to mitigate
@@
 resolution. 
-% In addition to the high
-% simulated fidelities which can be obtained with these techniques, it is desirable for the
-% resultant quantum state trajectory to be insensitive to experimental errors
-% such as parameter drift, noise, and finite control resolution. Quantum state
-% trajectories that are insensitive to these effects are said to be robust.
-% In line with this objective, 
 In order to address these issues,
@@
 
-In this work we mitigate
-errors arising from longitudinal relaxation, parameter deviations, and 1/$f$ flux noise
-on the fluxonium qubit \cite{zhang2020universal} by solving a robust trajectory 
-optimization problem. 
-Leveraging recent advances in trajectory optimization within the field of robotics, we
-solve this optimization problem using the ALTRO solver, which uses iterative LQR 
-(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
-methods such as GOAT \cite{machnes2015gradient}, GRAPE
-\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
-an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
-each time step \cite{howell2019altro}. 
-% The augmented Lagrangian method
-% allows us to enforce arbitrary, simultaneous constraints and achieve
-% fast convergence without being restricted to the constraint
-% manifold. 
-To mitigate errors due to longitudinal
-relaxation, we perform time-optimal control and
-encode the dependence of longitudinal relaxation
-on the controls in an efficient optimization objective that does
-not pay the increased computational cost of integrating a master equation.
-Additionally, to make the quantum state trajectory robust to parameter deviations,
-we consider three different approaches to making the optimized trajectory less 
-sensitive to model uncertainty:
-% we consider three robust
-% robust trajectory opitmization techniques to make the
-% quantum state trajectory robust to parameter deviations.
+In this work, we study three robust control techniques that
+make the optimized trajectory less sensitive
+to static and time-dependent parameter deviations:
 \begin{enumerate}
@@
 \end{enumerate}
-% The first is the sampling method, which has been applied
-% previously in the context of QOC
-%%TODO: the citations for unscented sampling
-%% and sampling are the same as in section V
-% \cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
-% The second is the unscented sampling
-% method, which derives from the unscented transformation used
-% for non-linear Kalman filtering
-% \cite{howell2020direct, julier2004unscented, lee2013sigma, manchester2016derivative}.
-% We propose a third method, which uses derivative information
-% of parameter deviations with respect to the quantum state trajectory.
-% We make this method efficient by employing mixed-mode differentiation.
+We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
+We also show that QOC can solve important problems for fluxonium-based qubits,
+in particular taking advantage of the $T_{1}$-dependence of the controls
+to mitigate longitudinal relaxation and
+performing phase gates in arbitrary times.
+To mitigate errors due to longitudinal
+relaxation, we perform time-optimal control and
+utilize an efficient optimization objective that does
+not pay the increased computational cost of integrating a master equation.
+Leveraging recent advances in trajectory optimization within the field of robotics, we
+solve these optimization problems using the ALTRO solver, which uses iterative LQR 
+(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
+methods such as GOAT \cite{machnes2015gradient}, GRAPE
+\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
+an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
+each time step \cite{howell2019altro}.
 
 This paper is organized as follows.
-First, we introduce the ALTRO method in the context of QOC \ref{sec:background}.
+First, we introduce the ALTRO method in the context of QOC
+in Section \ref{sec:background}.
 We describe realistic constraints for the fluxonium and
-map them to the ALTRO method \ref{sec:fluxonium}. Then, we
+map them to the ALTRO method in Section \ref{sec:fluxonium}. Then, we
 outline a method for making the optimization aware of longitudinal
-relaxation \ref{sec:longitude}. Next, we outline three methods for achieving
-robustness to static parameter deviations \ref{sec:static}. Finally,
+relaxation in Section \ref{sec:longitude}. Next, we outline three methods for achieving
+robustness to static parameter deviations in Section \ref{sec:static}. Finally,
 we employ the robust control techniques to mitigate 1/$f$ flux noise
-\ref{sec:stochastic}.
+in Section \ref{sec:stochastic}.

### 2020-11-25 paper/2background.tex

--- 2020-11-20 paper/2background.tex
@@
 a quantum state $\ket{\psi(t)}$ governed by the time-dependent
-Schr{\"o}dinger equation (TDSE)
+Schr{\"o}dinger equation (TDSE),
 \begin{equation}
@@
 \end{equation}
-The evolution is sometimes cast with the evolution
-of a density matrix under the Lindblad master equation to
-model the decoherence of the state explicitly. The Hamiltonian
+The Hamiltonian
 has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
@@
 to the final knot point and the target state
-$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
-In general $J$ is a linear combinaion of cost functions on the state, e.g.
-forbidden-state occupation, as well as
-cost functions on the controls, e.g. the norm of the control amplitudes
-\cite{leung2017speedup}. Typical quantum optimal control
+$J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u(t))} \rvert}^{2}$.
+In general $J$ is a linear combinaion of cost functions on the state
+as well as cost functions on the controls \cite{leung2017speedup}.
+Typical quantum optimal control
 algorithms employ automatic differentiation
@@
 Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
-and solved using a variety of specialized solvers developed by the robotics community \cite{Schulman13,Tedrake16,Hereid2017FROST,howell2019altro}.
+and solved using a variety of specialized solvers developed by the robotics community
+\cite{Schulman13,Tedrake16,Hereid2017FROST,howell2019altro}.
 Trajectory optimization problems are typically of the following form: 
@@
 concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
-across all time steps. After minimizing \eqref{opt:alm} using iLQR, the Lagrange multiplier estimates are updated according to,
+across all time steps. After minimizing \eqref{opt:alm} using iLQR,
+the Lagrange multiplier estimates are updated according to,
 \begin{equation}
@@
 machine-precision constraint satisfaction. For more information on the details of the ALTRO
-solver, see \cite{howell2019altro,Jackson2020altroc}.
+solver, see \cite{howell2019altro, Jackson2020altroc}.
 
-% For the QOC problem, the state $x$ consists of \todo{fill this out} while the control input
-% $u$ includes the flux drive amplitude. The cost functions $\ell_N$
-% The trajectory optimization
-% literature, driven primarily by the robotics community,
-% solves a more general class of non-linear programs that resemble
-% the QOC problem. The QOC
-% problem is a specific case of the linear quadratic regulator (LQR).
-% LQR is so called because the dynamics are linear in the state and
-% the functional is quadratic in the state. A general LQR cost
-% function takes the form
-% \begin{equation}
-%   J_{\textrm{iLQR}} = \ell_{N}({x}_{N})
-%   + \sum_{k = 0}^{N - 1} \ell_{k}({x}_{k}, u_{k})
-% \end{equation}
-% The subscript $k$ denotes the knot point, $x$ is the augmented state vector,
-% and $u$ is the augmented control vector. The augmented state is propagated
-% to the next knot point by the dynamics function $x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
-% In the case of QOC $\ket{\psi_{k}} \subseteq x_{k}$
-% and $f$ encodes the TDSE dynamics \eqref{eq:tdse}. The
-% cost function $\ell_{k}$ includes contributions from the infidelity
-% $1 - {\lvert \braket{\psi_{f} \lvert \psi_{k}} \rvert}^{2}$, final state deviations
-% $\tilde{x}^{T}_{k} Q_{k} \tilde{x}_{k}$ where $\tilde{x}_{k} = x_{f} - x_{k}$,
-% and control regularization $u^{T}_{k} R_{k} u_{k}$.
-% Placing cost functions at each knot point smoothens the optimization
-% landscape. In QOC realizing the final state
-% is the most important objective so the cost function at the final knot point
-% carries a higher realtive weight $\ell_{N} \sim N \cdot \ell_{k}$.
-% The LQR formulation restricts the cost functions to operate on the augmented
-% state and controls at each knot point. The advantage of the LQR formulation
-% is that there exists a differentiable dynamic programming algorithm to compute the
-% optimal update to the augmented controls $u_{k}$ which minimizes the cost
-% function $\ell_{k}$ at each knot point. This algorithm proceeds by deriving a
-% recurrence relation between knot points $k$ and $k + 1$ for the optimal
-% feedback law--known as the Ricatti recursion (citations needed). The
-% iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
-% and applies the Ricatti recursion to all knot points on multiple
-% executions.
-
-% In order to incorporate constraints we employ
-% the augmented Lagrangian method. Constraints are contributions
-% to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
-% zero or negative when the constraint is satisfied. The AL-iLQR
-% method associates a penalty multiplier with the constraint
-% that estimates the constraint's Lagrange multiplier.
-% The algorithm updates the penalty multiplier between
-% iLQR executions. In this scheme the functional takes the form
-% \begin{equation}
-% \begin{aligned}
-%     J_{\textrm{AL-iLQR}} &= J_{\textrm{iLQR}}\\
-%     &+ (\lambda_{N} + \frac{1}{2}I_{\mu_{N}} c_{N}(x_{N}))^{T} c_{N}(x_{N})\\
-%     &+ \sum^{N - 1}_{k = 0} (\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
-% \end{aligned}
-% \end{equation}
-% Here $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
-% with $\mu_{k}$ along the diagonal.
-% $\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
-% \begin{align}
-%   \lambda^{i}_{k} &\gets \max(0, \lambda^{i}_{k} + \mu^{i}_{k} c^{i}_{k}(x_{k}^{*}, u_{k}^{*}))\\
-%   \mu^{i}_{k} &\gets \phi \mu^{i}_{k}
-% \end{align}
-% Here $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
-% $i$ indicates the $i$-th constraint functional,
-% and $\phi$ is a hyperparameter. An optimal control update, similar to the Ricatti
-% recursion, can be formed to calculate the optimal control updates for this functional,
-% see \cite{howell2019altro}.
+The ALTRO method has fast tail-convergence and
+allows us to impose arbitrary constraints without restricting
+the optimization to the constraint manifold.
+This contrasts typical first-order QOC methods that employ the same
+update procedure for the duration of optimization
+and employ projected gradient methods
+which restrict the optimization to the constraint manifold.
+To see how to use the ALTRO method for QOC, we will present
+a concrete example in the following section.

### 2020-11-25 paper/3fluxonium.tex

@@
 a classical computer. In the two-level
-approximation the system Hamiltonian takes the form
+approximation the system Hamiltonian takes the form:
 \begin{align}
@@
 \begin{mini!}[2]
-    {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k} + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{}
+  {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
+    + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{} \label{eq:costfun}
     \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
-    \addConstraint{x_1}{= x_\text{init}} \label{eq:init_con}
-    \addConstraint{x_N}{= x_f} \label{eq:term_con}
+    \addConstraint{(x_1, x_N \; \textrm{given})}
+    \addConstraint{g_k(x_k,u_k)}{\leq 0}
+    \addConstraint{h_k(x_k,u_k)}{=0}
 \end{mini!}
@@
   u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
+  \label{eq:astatecontrols}
 \end{equation}
@@
 \end{equation}
-The dynamics function $f(x_{k}, u_{k})$ integrates the TDSE dynamics for the quantum states
+The discrete dynamics function $f(x_{k}, u_{k})$ integrates the TDSE dynamics for the quantum states
 \eqref{eq:tdse}, \eqref{eq:hamiltonian}
@@
 encoded in the realtive metric weight $Q_{N} \sim N \cdot Q_{k}$.
-%%, $k \in \{1, ..., N - 1\}$.
 
-Additionally, we impose constraints that reflect limitations
-of the apparatus and ones that improve the experimental realization
-of the control pulse. To ensure gates may be concatenated arbitrarily without
+In addition to imposing optimization constraints that
+reflect physical limitations of the apparatus, we impose
+constraints that improve the experimental realization of the control pulse.
+To ensure gates may be concatenated arbitrarily without
 inducing AWG ringing due to high-frequency transitions,
@@
 
-In addition to obeying experimental constraints
-and achieving low simulated gate errors, we desire
-to make our gates perform well in the presence of decoherence.
-Decoherence of the quantum state due to external noise
-is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
-They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively.
-The main contributions to longitudinal relaxation in our
-device are dielectric loss in the capacitor, resistive loss in the inductor,
-and Purcell loss. The main contributions to pure dephasing in our
-device are $1/f$ flux noise and decay via charge and flux coupling
-to the control lines.
-Dissipation to the thermal bath via longitudinal
-relaxation is an irreversible process
-that results in information loss.
-Converesely, pure dephasing is a reversible process.
-Our device achieves its best pure dephasing
-protection at the flux frustration point
-$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-where the qubit is first-order insensitive to changes in flux.
-It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-and increases away from the flux frustration point
-$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-of the decay processes and the tradeoff, we choose
-to maximize the longitudinal relaxation time directly
-by tuning the flux amplitude \ref{sec:longitude} and employ robust control techniques to mitigate
-pure dephasing \ref{sec:static}, \ref{sec:stochastic}.
+%% In addition to obeying experimental constraints
+%% and achieving low simulated gate errors, we desire
+%% to make our gates perform well in the presence of decoherence.
+%% Decoherence of the quantum state due to external noise
+%% is modeled by two phenomena: longitudinal relaxation and pure dephasing.
+%% They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively.
+%% Dissipation to the thermal bath via longitudinal
+%% relaxation is an irreversible process
+%% that results in information loss.
+%% Converesely, pure dephasing is a reversible process.
+%% We choose to maximize the longitudinal relaxation time directly
+%% by tuning the flux amplitude, as in Section \ref{sec:longitude}, and
+%% employ robust control techniques to mitigate
+%% pure dephasing in Section \ref{sec:stochastic}.
+
+%% The main contributions to longitudinal relaxation in our
+%% device are dielectric loss in the capacitor, resistive loss in the inductor,
+%% and Purcell loss. The main contributions to pure dephasing in our
+%% device are $1/f$ flux noise and decay via charge and flux coupling
+%% to the control lines.
+%% Our device achieves its best pure dephasing
+%% protection at the flux frustration point
+%% $T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
+%% where the qubit is first-order insensitive to changes in flux.
+%% It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
+%% frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
+%% at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
+%% and increases away from the flux frustration point
+%% $T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
+%% of the decay processes and the tradeoff,

### 2020-11-25 paper/4longitude.tex

--- 2020-11-18 paper/4longitude.tex
@@
     \includegraphics[width=\linewidth]{assets/f1a.png}
+    \caption{\label{fig:longitudea}}
   \end{subfigure}\hfill
@@
     \includegraphics[width=\linewidth]{assets/f1b.png}
+    \caption{\label{fig:longitudeb}}
   \end{subfigure}\hfill
@@
     \includegraphics[width=\linewidth]{assets/f1c.png}
+    \caption{\label{fig:longitudec}}
   \end{subfigure}
@@
 
-\section{Longitudinal Relaxation Awareness}
+\section{Longitudinal Relaxation Awareness \label{sec:longitude}}
+Dissipation to the thermal bath via longitudinal relaxation
+is an irreversible process that results in information loss.
+This process is modeled using its $1/e$ decay time $T_{1}$.
 The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms.
+control parameters in a range of superconducting circuit platforms,
+including the fluxonium qubit.
 It is advantageous to tune the controls to extend the longitudinal
@@
 requires propagating density matrices of size $n \times n$ under master equation
-dynamics, rather than state vectors of size $n$ under the TDSE dynamics.
+dynamics, rather than state vectors of size $n$ under the TDSE dynamics \eqref{eq:tdse}.
 We avoid this increase in computational complexity by
@@
 
-\todo{again, I'd recommend writing out either the entire optimization problem or at least
-point to the parts of the optimization problem defined in the previous section you had to 
-change. Just make it a little more concrete.}
-
-The longitudinal relaxation probability is given by
+The longitudinal relaxation probability is given by,
 \begin{equation}
@@
 \end{equation}
-$P_{1}$ is penalized using a quadratic cost at each knot point
-${\lvert P_{1}(t_{k}) \rvert}^{2}$.
+This value is appended to the augmented state vector \eqref{eq:astatecontrols}
+and penalized according to \eqref{eq:costfun}.
 $T_{1}(a_{k})$ is obtained at each knot point by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$.
+a spline fit to experimental data of the form $\{(a, T_{1})\}$,
+see Figure \ref{fig:longitudeb}.
 It is also possible to use a spline fit to theoretically obtained data.
@@
 Markovian decision structure of the trajectory
-optimization problem. To ensure numerical
+optimization problem. The square root of the time step $\sqrt{\Delta t_{k}}$
+is appended to the augmented control vector \eqref{eq:astatecontrols}
+and the squared root
+of the time step $\lvert \Delta t_{k} \rvert$ is used
+for integration in the discrete dynamics function \eqref{eq:dyn_con}.
+To ensure numerical
 integration accuracy is maintained we add a bound
-constraint at each knot point
-$5\textrm{e-}2 \ \textrm{ns} \le
-\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
-This bound constraint may be
-violated for intermediate iterations of the optimization,
-so we add the square root of the time step $\sqrt{\Delta t_{k}}$
-to the augmented control vector and use the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ in optimization.
+constraint on the time step at each knot point.
 
@@
     \includegraphics[width=\linewidth]{assets/f2a.png}
+    \label{fig:statica}
   \end{subfigure}\hfill
@@
     \includegraphics[width=\linewidth]{assets/f2c.png}
+    \label{fig:staticb}
   \end{subfigure}\hfill
@@
     \includegraphics[width=\linewidth]{assets/f2b.png}
+    \label{fig:staticc}
   \end{subfigure}

### 2020-11-25 paper/0abstract.tex
added lots of references to introduction

@@
   Robotics Institute, Carnegie Mellon University, Pittsburgh, Pennsylvania 15213, USA
-  % Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
 }
@@
 \begin{abstract}
-  %% - 1 + 2 topic importance
-  %% - 1 what I have done
-  %% - few sentences about primary results
   The ability to engineer high-fidelity gates on quantum processors in the presence of

### 2020-11-25 paper/1introduction.tex
added lots of references to introduction

@@
 algorithms for accurately and efficiently manipulating quantum systems.
-Early techniques were proposed for nuclear magnetic resonance experiments
+\todo{more nmr and sqc cite} Early techniques were proposed for nuclear magnetic resonance experiments
 \cite{khaneja2005optimal}, and applications now include superconducting
@@
   huang2020engineering, leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
-\cite{sorensen2018quantum}.
-For quantum computation
-QOC techniques are employed to achieve high fidelity gates
+neutral and ionized atoms \cite{
+  brouzos2015quantum, de2008optimal, goerz2011quantum, guo2019high, jensen2019time,
+  larrouy2020fast, omran2019generation, rosi2013fast, sorensen2019qengine,
+  treutlein2006microwave, van2016optimal}, nitrogen-vacancy centers in
+diamond \cite{chou2015optimal, dolde2014high, geng2016experimental,
+  nobauer2015smooth, poggiali2018optimal, rembold2020introduction, tian2019optimal},
+and Bose-Einstien condensates \cite{amri2019optimal, sorensen2018quantum}.
+For quantum computation,
+optimal control techniques are employed to achieve high fidelity gates
 while adhering to experimental constraints.
-Experimental errors may cause the system's trajectory to deviate from that predicted in
-optimization, hampering performance.
-Robust control techniques 
-address this issue by encoding
-experimental errors in optimization objectives, yielding performance
-garauntees in the presence of model uncertainty \cite{Zhou97,Morimoto00,Manchester18}.
-In this work we employ robust control techniques to mitigate
-realistic system parameter deviations and decoherence that arise when controlling
+Experimental errors may cause the system to deviate
+from the model used in optimization, leading
+to poor experimental performance.
+Robust control techniques improve upon
+standard optimal control techniques by encoding
+model uncertainties
+in optimization objectives, yielding performance
+guarantees over a range of parameters \cite{Zhou97,Morimoto00,Manchester18}.
+We adapt robust control techniques from the robotics community to mitigate
+realistic model uncertainties for
 a superconducting fluxonium qubit.
 
-Analytic and numerical techniques for QOC have been remarkably successful in
-designing high fidelity gates \todo{citations needed}. 
-While these methods are able to predict system behavior with high accuracy, they tend 
-to be sensitive to experimental errors such as parameter drift, noise, and finite control 
-resolution. 
-In order to address these issues,
-recent work has sought to mitigate
-errors due to parameter deviations.
-Considering the dynamical and geometric phases of quantum state
-trajectories has led to analytic methods for mitigating
-errors due to parameter deviations and pure dephasing
-\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
-Floquet techniques have been experimentally demonstrated to simultaneously mitigate
-longitudinal relaxation and pure dephasing decoherence
+Analytic and numerical techniques for QOC have had tremendous success in
+engineering high fidelity quantum gates \cite{
+  chakram2020multimode, chou2015optimal, dolde2014high, egger2013optimized, egger2014adaptive,
+  gokhale2019partial, grace2007optimal, heeres2017implementing, huang2014optimal,
+  kelly2014optimal, leng2019robust, liebermann2016optimal,
+  nebendahl2009optimal, rebentrost2009optimal, rebentrost2009optimal2,
+  spiteri2018quantum, sporl2007optimal}. 
+While traditional QOC techniques can predict system behavior
+with high accuracy, they are sensitive to experimental
+errors such as parameter drift, noise, finite control 
+resolution, and decoherence. Analytic techniques to mitigate parameter deviation errors
+include dynamic and geometric phase considerations
+\cite{han2020experimental, xu2020nonadiabatic} and
+composite pulses \cite{merrill2014progress}.
+To mitigate decoherence, Floquet techniques have been employed
 \cite{huang2020engineering, mundada2020floquet}.
-Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
-by modeling master equations \cite{rembold2020introduction} and employing
+Numerical techniques to mitigate parameter deviation errors
+include closed-loop methods \cite{egger2014adaptive, feng2018gradient, huang2020engineering,
+  wittler2020integrated} and open-loop methods \cite{
+  allen2019robust, carvalho2020error, reinhold2019controlling,
+  rembold2020introduction, kosut2013robust}.
+Numerical techniques to mitigate decoherence include
+modeling master equations \cite{rembold2020introduction} and employing
 Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
-Additionally, efforts have been made to incorporate experimental feedback
-into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
-and pulse design are under active development \cite{wittler2020integrated}.
 
 In this work, we study three robust control techniques that
-make the optimized trajectory less sensitive
+make the system's state trajectory less sensitive
 to static and time-dependent parameter deviations:
 \begin{enumerate}
-\item A sampling method, similar to the work of \cite{carvalho2020error, reinhold2019controlling,
-  rembold2020introduction}.
-  \item An unscented sampling method, adapted from the unscented tranform used in the 
+  \item A sampling method, similar to the work of \cite{allen2019robust,
+    carvalho2020error, reinhold2019controlling, rembold2020introduction}.
+  \item An unscented sampling method adapted from the unscented transform used in the 
     state estimation community \cite{howell2020direct, julier2004unscented,
@@
 We also show that QOC can solve important problems for fluxonium-based qubits,
-in particular taking advantage of the $T_{1}$-dependence of the controls
-to mitigate longitudinal relaxation and
+in particular, taking advantage of the $T_{1}$-dependence of the controls
+to mitigate longitudinal relaxation type decoherence and
 performing phase gates in arbitrary times.

### 2020-11-28 paper/1introduction.tex
solid progress through sec4

@@
 algorithms for accurately and efficiently manipulating quantum systems.
-\todo{more nmr and sqc cite} Early techniques were proposed for nuclear magnetic resonance experiments
-\cite{khaneja2005optimal}, and applications now include superconducting
-circuits \cite{heeres2017implementing,
-  huang2020engineering, leng2019robust, leung2017speedup, xu2020nonadiabatic},
-neutral and ionized atoms \cite{
-  brouzos2015quantum, de2008optimal, goerz2011quantum, guo2019high, jensen2019time,
+Early techniques were proposed for nuclear magnetic resonance experiments
+\cite{vandersypen2005nmr, kehlet2004improving, khaneja2005optimal,
+  maximov2008optimal, nielsen2010optimal, skinner2003application, tosner2009optimal},
+and applications now include superconducting circuits \cite{abdelhafez2020universal,
+  chakram2020multimode, fisher2010optimal, gokhale2019partial,
+  huang2020engineering, leng2019robust, leung2017speedup, li2020fast,
+  xu2020nonadiabatic},
+neutral atoms and ions \cite{brouzos2015quantum,
+  de2008optimal, goerz2011quantum, guo2019high, jensen2019time,
   larrouy2020fast, omran2019generation, rosi2013fast, sorensen2019qengine,
-  treutlein2006microwave, van2016optimal}, nitrogen-vacancy centers in
-diamond \cite{chou2015optimal, dolde2014high, geng2016experimental,
+  treutlein2006microwave, van2016optimal},
+nitrogen-vacancy centers in diamond \cite{chou2015optimal,
+  dolde2014high, geng2016experimental,
   nobauer2015smooth, poggiali2018optimal, rembold2020introduction, tian2019optimal},
@@
 
-Analytic and numerical techniques for QOC have had tremendous success in
+QOC techniques have had tremendous success in
 engineering high fidelity quantum gates \cite{
-  chakram2020multimode, chou2015optimal, dolde2014high, egger2013optimized, egger2014adaptive,
-  gokhale2019partial, grace2007optimal, heeres2017implementing, huang2014optimal,
+  chou2015optimal, dolde2014high, egger2013optimized, egger2014adaptive,
+  grace2007optimal, heeres2017implementing, huang2014optimal,
   kelly2014optimal, leng2019robust, liebermann2016optimal,
@@
   spiteri2018quantum, sporl2007optimal}. 
-While traditional QOC techniques can predict system behavior
+While standard QOC techniques can predict system behavior
 with high accuracy, they are sensitive to experimental
 errors such as parameter drift, noise, finite control 
-resolution, and decoherence. Analytic techniques to mitigate parameter deviation errors
+resolution, and decoherence. Multiple techniques have been developed to address
+these shortcomings. Analytic techniques to mitigate parameter deviation errors
 include dynamic and geometric phase considerations
 \cite{han2020experimental, xu2020nonadiabatic} and
-composite pulses \cite{merrill2014progress}.
+composite pulses \cite{cummins2000use, cummins2003tackling,
+  kupce1995stretched, merrill2014progress}.
 To mitigate decoherence, Floquet techniques have been employed

### 2020-11-28 paper/2background.tex
solid progress through sec4

@@
-\section{QOC + AL-iLQR \label{sec:background}}
-In this section we introduce the notation
-we will use throughout the paper,
-review the quantum optimal control problem statement,
-and introduce the trajectory optimization framework.
+\section{QOC + ALTRO \label{sec:background}}
+In this section we
+review the quantum optimal control problem statement
+and introduce the ALTRO solver.
 Quantum optimal control concerns the evolution of
@@
 \begin{equation}
-  i \hbar \frac{d}{dt} \ket{\psi} = H(u(t), t) \ket{\psi}
+  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(u(t), t) \ket{\psi(t)}
   \label{eq:tdse}
 \end{equation}
-The Hamiltonian
-has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
+The Hamiltonian has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
 The controls are so called because they are the means the experimentalist has to
-act on the system. Numerical quantum optimal control techniques make
-the problem tractable by discretizing the problem into $N$
-knot points (time steps). Typical explicit integration techniques for the TDSE include
-exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}
-and Runge-Kutta integrators \cite{jorgensen2011numerical}.
-
-Quantum optimal control seeks the control
-parameters that minimize a functional $J(u(t))$.
+act on the system. To make the problem numerically tractable,
+the controls and quantum state are discretized into $N$ knot points (time steps).
+The optimization problem is to find the
+controls that minimize a functional $J(u)$.
 In the simplest case the functional is
@@
 to the final knot point and the target state
-$J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u(t))} \rvert}^{2}$.
-In general $J$ is a linear combinaion of cost functions on the state
+$J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u)} \rvert}^{2}$.
+In general, $J$ is a linear combinaion of cost functions on the state
 as well as cost functions on the controls \cite{leung2017speedup}.
@@
 which can easily be used to implement first-order optimization methods
-\cite{clarkson2010coresets, hauswirth2016projected}.
+\cite{machnes2015gradient, khaneja2005optimal, leung2017speedup, goerz2019krotov}.
 
@@
 \begin{mini}[2]
-    {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell(x_k,u_k) }{}{}
+    {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell(x_k,u_k) }{}{}
     \addConstraint{x_{k+1} = f(x_k,u_k)}
@@
 where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^n$ and
-$u_k \in \R^m$ are the state and input control variables, $f(x_k,u_k)$ is the discrete
-dynamics function, and $g_k(x_k,u_k)$ and $h_k(x_k,u_k)$ are the inequality
+$u_k \in \R^m$ are the augmented state and input augmented control variables,
+$f(x_k,u_k)$ is the discrete dynamics function, and $g_k(x_k,u_k)$ and
+$h_k(x_k,u_k)$ are the inequality
 and equality constraints, potentially including initial and final conditions,
@@
 
-The ALTRO method has fast tail-convergence and
-allows us to impose arbitrary constraints without restricting
-the optimization to the constraint manifold.
-This contrasts typical first-order QOC methods that employ the same
-update procedure for the duration of optimization
-and employ projected gradient methods
-which restrict the optimization to the constraint manifold.
-To see how to use the ALTRO method for QOC, we will present
-a concrete example in the following section.
+As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
+for imposing arbitrary, simultaneous constraints
+on both the state trajectory and the controls and does not restrict
+optimization to the constraint manifold. First-order solvers like GRAPE
+rely on unconstrained optimization to satisfy desiderata on the state trajectory,
+such as achieving a desired gate fidelity, forbidding occupations of particular quantum states,
+and smoothing the controls below a threshold. The unconstrained approach relies
+on a careful choice of competing penalty weights, which is cumbersome and often intractable.
+To impose constraints on the controls, such as amplitude restrictions, first-order solvers
+employ projected gradient descent methods which restrict optimization to the constraint manifold and
+may hinder convergence to the optimal solution \cite{clarkson2010coresets,
+  hauswirth2016projected}.
+The ALTRO solver delays projecting onto the constraint manifold until the ALM
+solve is already at a coarse tolerance, achieving machine-precision satisfaction
+for constraints on the controls and the state trajectory. These advantage are critical
+for this work, where multiple constraints must be satisfied in addition
+to minimizing auxiliary metrics.

### 2020-11-28 paper/3fluxonium.tex
solid progress through sec4

@@
 The fluxonium qubit is a promising building block for superconducting
-circuits, and the experimental constraints we encode in the optimization reflect
-those of a realistic device. Furthermore, the accurate
-two-level approximation of the system Hamiltonian makes it
-efficient to perform quantum optimal control on
-a classical computer. In the two-level
-approximation the system Hamiltonian takes the form:
+circuits, and the accurate
+two-level approximation of its Hamiltonian makes
+quantum optimal control on a classical computer inexpensive.
+In the two-level
+approximation the Hamiltonian takes the form:
 \begin{align}
@@
 where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. The flux amplitude $a$ is experimentally
-realized by modulating the flux 
-threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
+$a(t)$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
+are Pauli matrices. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
 gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-systematic errors. We compare the gates we obtain with numerical
+Hamiltonian parameter deviations. We compare the gates we obtain with numerical
 methods to the analytically constructed gates reported in
@@
 The initial and final conditions are determined by the
-desired unitary transformation, which requires two quantum states
+desired gate, which requires two quantum states
 to specify $\psi^{(1)}$, $\psi^{(2)}$.
@@
   \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
+  \label{eq:isomorphism}
 \end{equation}
-The discrete dynamics function $f(x_{k}, u_{k})$ integrates the TDSE dynamics for the quantum states
-\eqref{eq:tdse}, \eqref{eq:hamiltonian}
+The discrete dynamics function \eqref{eq:dyn_con} integrates the TDSE dynamics for the quantum states
+using \eqref{eq:tdse} and \eqref{eq:hamiltonian}
 and integrates the moments of the flux drive amplitude. Exposing lower order moments of the flux
@@
 and mitigating AWG ringing due to high frequency transitions.
-The matrices $Q_{k}$ and $R_{k}$ define the penalty metric. We choose $Q$
-to be diagonal because it is computationally efficient. This corresponds
+The matrices $Q_{k}$ and $R_{k}$ define the penalty metric.
+Including objectives at each knot point smoothens the optimization landscape, though
+the importance of the final state objective is encoded in the relative
+weight $Q_{N} \sim N \cdot Q_{k}$.
+We choose $Q$ to be diagonal because it is computationally efficient. This corresponds
 to penalizing phase differences between the quantum states, although
 other phase-insensitive metrics such as infidelity may be employed.
-The final state transfer is the most important objective which is
-encoded in the realtive metric weight $Q_{N} \sim N \cdot Q_{k}$.
 
@@
 and $0$ otherwise.
-In addition the norm of the quantum states
+In addition, the norm of the quantum states
 are constrained to $1$ to ensure discretization error is not
-exploited $c_{k} = x_{k}^{T}x_{k} - 1$.
-
-%% In addition to obeying experimental constraints
-%% and achieving low simulated gate errors, we desire
-%% to make our gates perform well in the presence of decoherence.
-%% Decoherence of the quantum state due to external noise
-%% is modeled by two phenomena: longitudinal relaxation and pure dephasing.
-%% They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively.
-%% Dissipation to the thermal bath via longitudinal
-%% relaxation is an irreversible process
-%% that results in information loss.
-%% Converesely, pure dephasing is a reversible process.
-%% We choose to maximize the longitudinal relaxation time directly
-%% by tuning the flux amplitude, as in Section \ref{sec:longitude}, and
-%% employ robust control techniques to mitigate
-%% pure dephasing in Section \ref{sec:stochastic}.
-
-%% The main contributions to longitudinal relaxation in our
-%% device are dielectric loss in the capacitor, resistive loss in the inductor,
-%% and Purcell loss. The main contributions to pure dephasing in our
-%% device are $1/f$ flux noise and decay via charge and flux coupling
-%% to the control lines.
-%% Our device achieves its best pure dephasing
-%% protection at the flux frustration point
-%% $T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
-%% where the qubit is first-order insensitive to changes in flux.
-%% It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
-%% frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
-%% at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
-%% and increases away from the flux frustration point
-%% $T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
-%% of the decay processes and the tradeoff,
+exploited $c_{k} = x_{k}^{T}x_{k} - 1$. The selection
+of subsets from $x_{k}$ is implied in
+the constraint equations.

### 2020-11-28 paper/4longitude.tex
solid progress through sec4

@@
     gate error is computed after each gate application.
-    The numerically optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
-    in the figure.
+    The numerically optimized $Z/2$ and $Y/2$ gate errors are indistinguishable.
   }
@@
 \section{Longitudinal Relaxation Awareness \label{sec:longitude}}
-Dissipation to the thermal bath via longitudinal relaxation
-is an irreversible process that results in information loss.
-This process is modeled using its $1/e$ decay time $T_{1}$.
-The longitudinal relaxation time $T_{1}$ varies with
-control parameters in a range of superconducting circuit platforms,
-including the fluxonium qubit.
-It is advantageous to tune the controls to extend the longitudinal
-relaxation time.
+The strength of longitudinal relaxation type decoherence varies
+with the control parameters in a range of superconducting circuit platforms,
+including the fluxonium qubit. Because dissipation to the thermal bath
+via longitudinal relaxation is an irreversible process that results in information loss,
+it is advantageous to tune the controls to extend the longitudinal
+relaxation time $T_{1}$--which represents the $1/e$ decay time
+of the quantum state.
 Computing the gate error due to longitudinal relaxation
@@
 We avoid this increase in computational complexity by
-penalizing the integrated rate of longitudinal relaxation,
-i.e. the probability of longitudinal relaxation.
+penalizing the probability (integrated rate) of longitudinal relaxation.
 Using this probability as proxy for the gate error incurred
-is reasonable because losses due to longitudinal relaxation
-increase monotonically in time.
+is reasonable because losses due to longitudinal relaxation increase monotonically
+in time.
 This technique can be extended to
@@
 see Figure \ref{fig:longitudeb}.
-It is also possible to use a spline fit to theoretically obtained data.
+It is also possible to fit a spline to theoretically obtained data.
 However, $T_{1}$ values are known to fluctuate greatly
@@
 The numerically optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitude}.
+waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}.
 They extend their gate times
@@
 generated initial states. The numerical $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The numerical $X/2$ gate
-achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
+the concatenated gate application comparision, suppressing cumulative gate errors to $8 \cdot 10^{-3}$
+over $2000$ gate applications $\sim 40 \mu\textrm{s}$, see Figure \ref{fig:longitudec}.
+The numerical $X/2$ gate achieves a cumulative gate error of
+$1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
+These low cumulative gate errors for high gate counts are critical for
+noisy, intermediate-scale quantum (NISQ) applications.
 Both the analytic and numerical gates attain single gate errors sufficient for
-quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
-The low gate errors achieved by the numerical gates \todo{typo / awkward sentence} 
-for extend computations
-are critical for noisy, intermediate-scale quantum (NISQ) applications.
+quantum error correction ($< 10^{-4}$), which are required for fault-tolerant quantum computing
+\cite{aharonov2008fault, fowler2009high, gottesman1997stabilizer}.
 These improvements are significant for the realistic constraints we have imposed

### 2020-11-28 paper/5static.tex
solid progress through sec4

@@
 making the state evolution insensitive
-to Hamiltonian parameter deviations. As an example
+to Hamiltonian parameter deviations. As an example,
 we mitigate errors arising from the drift and finite measurement
-precision of the qubit frequency $\tilde{f_{q}} = f_{q} \pm \sigma_{f_{q}}$ \todo{maybe 
-reference \eqref{eq:hamiltonian} here?}.
-We consider three robust control techniques.
-The first is the sampling method, which has
-been proposed previously in the context of QOC
-\cite{rembold2020introduction, reinhold2019controlling, carvalho2020error}. In the
-sampling method, multiple states are evolved under distinct deviant dynamics
-to capture the effect of parameter deviations. We also
-study the unscented sampling method, which uses the unscented
-transformation to accurately propagate a distribution
-representing the uncertainty in an evolving state
-due to a parameter deviation.
-The unscented transformation was designed for nonlinear Kalman
-filtering and is frequently utilized in robust control
-\cite{howell2020direct, julier2004unscented, lee2013sigma, manchester2016derivative}.
-Finally, we propose the derivative method. Derivative
-information encoding the sensitivity of the
-state trajectory with respect to the deviant parameter
-is used to penalize state trajectory deviations.
+precision of the qubit frequency in \eqref{eq:hamiltonian}.
+We consider three robust control techniques:
+the sampling method, the unscented sampling method,
+and the derivative method.
 
-\todo{I think you can make this more concrete and simpler to understand by just defining
-the new augmented state vector and explaining you stack multiple copies of the state vector
-together, each evolving with a different static parameter. In this case there's only one 
-parameter being changed, so I'd just state it explicitly.}
-In the sampling method, sample states evolve under a
-Hamiltonian where a parameter is replaced by
-a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
-deviant values $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
-The gate error of each sample state is 
-penalized at each knot point
-$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
-Each initial state is used to initialize two
-sample states $\ket{\psi^{\pm}}$.
-The initial states are chosen so
-that their outer products span the operators on the
-Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-For this method the number of states in the augmented state vector
-scales as $O(dn^{2})$ \todo{is this right?}
-because each of $n^{2}$ initial
-states is represented by $2d$ samples. $n$ is both the dimension of the Hilbert space
-and the size of the state vector. $d$ is the number of deviant parameters.
+The sampling method optimizes over copies of a quantum state,
+called samples.
+Each sample is associated with a distinct deviant parameter value.
+Two samples $\psi_{\pm}$
+are appended to the augmented state vector \eqref{eq:astatecontrols}
+for each of four initial states
+$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
+(\ket{0} - \ket{1}) / \sqrt{2}\}$ which are chosen
+so that their outer products span the operators on the
+Hilbert space  \cite{chow2009randomized}.
+The discrete dynamics \eqref{eq:dyn_con} is modified so that the
+samples evolve under the Hamiltonian given in \eqref{eq:hamiltonian}
+with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$. The final state
+for each sample is the image of the initial state
+under the desired gate, and deviations from the final state are
+penalized in the objective \eqref{eq:costfun}.
 
@@
 
-The strategy of the unscented sampling method is to
-propagate an ensemble of sample states (sigma points) which represent a distribution
-over every element of the initial state. The distribution
-models the uncertainty in each state element arising from the
-parameter deviation. The ensemble consists of $4n + 2d$ states.
-Each state in the
-ensemble is propagated to the next
-knot point using separate deviant dynamics. Then, the mean and covariance of
-the propagated ensemble is calculated. New states are sampled from the distribution
-given by the calculated statistics for propagation to the next knot point.
-This resampling procedure, the unscented transformation, accurately propagates
-first and second moments of the distribution and ensures the points
-lie on the $\sqrt{2n}$\textsuperscript{th}
-covariance contour at each knot point. A detailed update procedure is given
-in Appendix \ref{appendix:unscented}. Each initial state from the operator basis
-is represented with a distribution of $4n + 2d$ samples. Hence,
-the number of states in the augmented state vector scales as $O(n^{3} + dn^{2})$.
-The gate error of each sample state is penalized as in the sampling method.
-\todo{I think this section needs some more detailed explanation.}
+For the unscented sampling method, $2(2n + d)$ samples
+are appended to the augmented state vector \eqref{eq:astatecontrols}
+for each initial state from the operator basis. Here $2n = 4$ is the
+dimension of the Hilbert space under the isomorphism \eqref{eq:isomorphism}
+and $d = 1$ is the number of deviant parameters. The sample
+states are used to represent a Gaussian distribution over all $2n$
+elements of the initial state, modeling
+the uncertainty in the state as a result of the uncertainty in
+the deviant parameter. The unscented transformation
+is applied to the sample ensemble at each knot point in the discrete dynamics
+\eqref{eq:dyn_con} to ensure the first and second moments of the distribution
+are accurately propagated. Each sample evolves under the Hamiltonian given in
+\eqref{eq:hamiltonian}
+where the qubit frequency is replaced by a deviant value determined by the
+statistics of the sample ensemble. As in the sampling method,
+deviations from the final state are penalized according to \eqref{eq:costfun}.
+A detailed update procedure for this method is given
+in Appendix \ref{appendix:unscented}.
 
 The derivative method draws on the intuition that
-the sensitivity of the state evolution to a parameter
-$\lambda$ is encoded in the $l$\textsuperscript{th}-order
+the sensitivity of the state evolution to the deviant parameter
+is encoded in the ($l$\textsuperscript{th}-order)
 derivative of the state with respect to that parameter
-$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
+$\partial_{f_{q}}^{l} \psi$.
+
+The $m$\textsuperscript{th}-order
 derivative method minimizes the norm of the first $m$
@@
 see Appendix \ref{appendix:derivative}.
-For this method $m$ state derivatives are associated with each
-initial state from the operator basis.
-So, the number of states in the augmented state vector scales as
-$O(dmn^{2})$. Additionally, we note that the number of initial
-states and samples we have employed for each of the three
-methods follows standard perscriptions. However, knowledge
-of the problem can typically be used to reduced the number
-of required samples. \todo{I think this also needs some more explanation. It's not 
+\todo{I think this also needs some more explanation. It's not 
 clear exactly what you're doing. You're penalizing the derivatives of the dynamics with
@@
 error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses \cite{merill2014progress},
+to arbitrary order with sufficiently many pulses \cite{merrill2014progress},
 which mirrors the negative correlation between

### 2020-11-28 paper/a4time.tex
solid progress through sec4

\section{Computational Performance and Scaling}
\todo{timing table}

\todo{make complexity analysis a coherent paragraph}
(sampling) For the sampling method the number of sample states in the augmented state vector
scales as $O(dn^{2})$
because each of $n^{2}$ initial
states is represented by $2d$ samples. $n$ is both the dimension of the Hilbert space
and the size of a state vector. $d$ is the number of deviant parameters.

(unscented) Each initial state from the operator basis
is represented with a distribution of $4n + 2d$ samples. Hence,
the number of states in the augmented state vector scales as $O(n^{3} + dn^{2})$.
The gate error of each sample state is penalized as in the sampling method.

(derivative) For this method $m$ state derivatives are associated with each
initial state from the operator basis.
So, the number of states in the augmented state vector scales as
$O(dmn^{2})$. Additionally, we note that the number of initial
states and samples we have employed for each of the three
methods follows standard perscriptions. However, knowledge
of the problem can typically be used to reduced the number
of required samples.

### 2020-11-28 paper/1introduction.tex
prog

@@
 these shortcomings. Analytic techniques to mitigate parameter deviation errors
-include dynamic and geometric phase considerations
-\cite{han2020experimental, xu2020nonadiabatic} and
-composite pulses \cite{cummins2000use, cummins2003tackling,
-  kupce1995stretched, merrill2014progress}.
+include composite pulses \cite{cummins2000use, cummins2003tackling,
+  kupce1995stretched, merrill2014progress},
+dynamic and geometric phase considerations
+\cite{han2020experimental, xu2020nonadiabatic},
+and the DRAG scheme \todo{DRAG citations}.
 To mitigate decoherence, Floquet techniques have been employed

### 2020-11-28 paper/4longitude.tex
prog

@@
 it is advantageous to tune the controls to extend the longitudinal
-relaxation time $T_{1}$--which represents the $1/e$ decay time
+relaxation time $T_{1}$--which represents the $1/e$ relaxation time
 of the quantum state.
@@
     \includegraphics[width=\linewidth]{assets/f2a.png}
-    \label{fig:statica}
+    \caption{\label{fig:statica}}
   \end{subfigure}\hfill
@@
     \includegraphics[width=\linewidth]{assets/f2c.png}
-    \label{fig:staticb}
+    \caption{\label{fig:staticb}}
   \end{subfigure}\hfill
@@
     \includegraphics[width=\linewidth]{assets/f2b.png}
-    \label{fig:staticc}
+    \caption{\label{fig:staticc}}
   \end{subfigure}

### 2020-11-28 paper/5static.tex
prog

@@
 
-The sampling method optimizes over copies of a quantum state,
-called samples.
-Each sample is associated with a distinct deviant parameter value.
+The sampling method optimizes over multiple copies of a state,
+each of which evolves with a distinct deviant parameter value.
 Two samples $\psi_{\pm}$
-are appended to the augmented state vector \eqref{eq:astatecontrols}
+are appended to the augmented state vector \eqref{eq:astatecontrols}.
 for each of four initial states
 $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ which are chosen
+(\ket{0} - \ket{1}) / \sqrt{2}\}$. The initial states are chosen
 so that their outer products span the operators on the
@@
 with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$. The final state
-for each sample is the image of the initial state
-under the desired gate, and deviations from the final state are
+for each sample is the image of its initial state
+under the desired gate. Deviations from the final state are
 penalized in the objective \eqref{eq:costfun}.
+
+For the unscented sampling method, $2(2n + d)$ samples
+are appended to the augmented state vector \eqref{eq:astatecontrols}
+for each initial state from the operator basis. Here $2n = 4$ is twice the
+dimension of the Hilbert space, resulting from the isomorphism \eqref{eq:isomorphism},
+and $d = 1$ is the number of deviant parameters. The samples
+are used to represent a Gaussian distribution over all $2n$
+elements of the initial state, modeling
+the uncertainty in the state as a result of the uncertainty in
+the deviant parameter. Each sample evolves under the Hamiltonian given in
+\eqref{eq:hamiltonian}
+where the qubit frequency is replaced by a deviant value determined by the
+statistics of the ensemble of samples.
+The discrete dynamics \eqref{eq:dyn_con} is modified to 
+evolve the samples under the TDSE dynamics \eqref{eq:tdse}
+and then apply the unscented transformation to the ensemble,
+accurately preserving the first and second moments
+of the distribution.
+As in the sampling method,
+deviations from the final state are penalized according to \eqref{eq:costfun}.
+A detailed procedure for the unscented transformation is given
+in Appendix \ref{appendix:unscented}.
+
+The derivative method draws on the intuition that
+the sensitivity of the state evolution to the deviant parameter
+is encoded in the ($l$\textsuperscript{th}-order)
+derivative of the state with respect to that parameter
+$\partial_{f_{q}}^{l} \psi$. In the $m$\textsuperscript{th}-order
+derivative method, all state derivatives of order $1, \dots, m$
+are appended to the augmented state vector \eqref{eq:astatecontrols}
+for each initial state in the operator basis. The norms of the state derivatives are penalized
+in \eqref{eq:costfun} by setting the corresponding final states to zero.
+The state derivatives could be obtained in the discrete dynamics \eqref{eq:dyn_con}
+with backwards mode differentiation.
+Naively applying automatic differentiation
+would result in differentiating all dynamics at knot points
+$1, \dots, k - 1$ to obtain the state derivative at knot point $k$, requiring
+$O(N^{2})$ matrix multiplications. Instead, we 
+employ forwards mode differentiation on the TDSE to obtain coupled dynamics
+which require $O(N)$ matrix multiplications to integrate.
+For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are,
+\begin{align}
+  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
+  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
+  H \ket{\partial_{f_{q}} \psi} +
+  (\partial_{f_{q}} H) \ket{\psi}
+  \label{eq:d1dyn}
+\end{align}
+Exponential integrators that account for the non-linear
+term in \eqref{eq:d1dyn} may be used to efficiently integrate the coupled dynamics,
+see Appendix \ref{appendix:derivative}. For timing results
+and an analysis of how the three robust methods scale to larger Hilbert spaces
+and more deviant parameters,
+consult Appendix \ref{appendix:time}.
+
+To demonstrate the applicability of these techniques
+to mitigate parameter deviations,
+we consider the task of achieving a single $Z/2$
+gate subject to a static qubit frequency detuning
+$f_{q} \gets f_{q} + \delta f_{q}$.
+We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
+the sampling methods accordingly. For each method we compute the gate error for
+one simulated gate application subject to the deviant dynamics given by the
+stated qubit frequency detuning $\delta f_{q}$.
+
+We compare the numerical methods
+to an analytically derived $Z/2$ gate, see Figure \ref{fig:statica}. 
+This gate corresponds to
+idling at the flux frustration point $a = 0$. The analytic gate
+is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
+is simple to derive. Its erroneous rotation angle $2 \pi t_{Z/2} \delta f_{q}$ is linearly sensitive to
+the qubit frequency detuning, resulting in a gate error that is quadratically sensitive
+to the qubit frequency detutning.
+At a one-percent
+qubit frequency detuning the analytic gate achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
+which is sufficient for quantum error correction, see Figure \ref{fig:staticb}.
+Although the analytic $Z/2$ gate performs well, it works
+only at the gate time $t_{Z/2}$. The ability to perform $Z$ rotations in arbitrary times is critical
+for operating multi-qubit experiments in the lab frame.
+Each numerical method is able to find solutions at
+all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
+These numerical methods offer an effective scheme for synchronizing
+qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
+
+The sampling and unscented sampling methods
+converge on qualitatively similar solutions which combine idling periods
+with fast ramps to the maximum amplitude. The gate error at a one-percent
+detuning from the nominal qubit frequency achieved
+by the sampling method does not improve substantially over the
+range of gate durations. The unscented sampling method
+achieves linear decreases in its gate error with longer gate durations
+until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
+gate error $\sim 3.5 \cdot 10^{-5}$.
+
+The derivative methods converge on qualitatively similar solutions that
+use fast traingle pulses at the boundaries and balance time
+on either side of the flux-frustration point symmetrically at low amplitudes.
+Both methods achieve a super-linear scaling in their gate error as
+a function of the gate duration. The gate error for the 1\textsuperscript{st}-order
+derivative method approaches zero at the larmor period $1 / f_{q}$, see Figure \ref{fig:staticc}.
+We believe the 1\textsuperscript{st}-order method outperforms the 2\textsuperscript{nd}-order
+method due to the low contribution of second-order
+terms to the gate error in this deviation regime, see Appendix \ref{appendix:derivative}.
+\todo{@Dave should we still make a comment on leaving out the composite
+pulse comparison?}
+
+%% Additionally, there are multiple analytic methods 
+%% to mitigate errors due to parameter deviations
+%% including composite pulse sequences \cite{merrill2014progress},
+%% the DRAG scheme \cite{krantz2019quantum}, and
+%% geometric phase considerations
+%% \cite{xu2020nonadiabatic, han2020experimental}.
+%% Composite pulse sequences are derived by computing
+%% the erroneous rotation arising from the parameter deviation
+%% with a finite order magnus expansion.
+%% Pulses are then suitably composed to eliminate the
+%% error. In principle the error may be eliminated
+%% to arbitrary order with sufficiently many pulses \cite{merrill2014progress},
+%% which mirrors the negative correlation between
+%% gate error and gate duration we observe with the derivative method.
+%% It is difficult to choose an appropriate composite pulse
+%% for the problem studied here,
+%% so we propose comparisons between composite pulses and numerical techniques
+%% for future work.
 
@@
     \includegraphics[width=\linewidth]{assets/f3a.png}
+    \caption{}
+    \label{fig:stochastica}
   \end{subfigure}\hspace{0.025\textwidth}
@@
     \includegraphics[width=\linewidth]{assets/f3b.png}
+        \caption{}
+    \label{fig:stochasticb}
   \end{subfigure}
@@
 \end{figure*}
-
-For the unscented sampling method, $2(2n + d)$ samples
-are appended to the augmented state vector \eqref{eq:astatecontrols}
-for each initial state from the operator basis. Here $2n = 4$ is the
-dimension of the Hilbert space under the isomorphism \eqref{eq:isomorphism}
-and $d = 1$ is the number of deviant parameters. The sample
-states are used to represent a Gaussian distribution over all $2n$
-elements of the initial state, modeling
-the uncertainty in the state as a result of the uncertainty in
-the deviant parameter. The unscented transformation
-is applied to the sample ensemble at each knot point in the discrete dynamics
-\eqref{eq:dyn_con} to ensure the first and second moments of the distribution
-are accurately propagated. Each sample evolves under the Hamiltonian given in
-\eqref{eq:hamiltonian}
-where the qubit frequency is replaced by a deviant value determined by the
-statistics of the sample ensemble. As in the sampling method,
-deviations from the final state are penalized according to \eqref{eq:costfun}.
-A detailed update procedure for this method is given
-in Appendix \ref{appendix:unscented}.
-
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to the deviant parameter
-is encoded in the ($l$\textsuperscript{th}-order)
-derivative of the state with respect to that parameter
-$\partial_{f_{q}}^{l} \psi$.
-
-The $m$\textsuperscript{th}-order
-derivative method minimizes the norm of the first $m$
-state derivatives with a quadratic cost at each knot point
-${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
-  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
-The state derivatives could be obtained with backwards mode differentiation.
-Naive automatic differentiation would compute
-the state derivative at all $1, \dots, k - 1$ knot points
-to obtain the state derivative at knot point $k$.
-For a single state derivative and $N$ knot points this
-requires $O(N^2)$ matrix multiplications.
-Instead, we forward propagate the state derivatives in the
-augmented state vector under coupled dynamics, resulting in
-$O(N)$ matrix multiplications. For example, the dynamics
-for the 1\textsuperscript{st}-order derivative method are
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
-  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
-  H \ket{\partial_{\lambda} \psi} +
-  (\partial_{\lambda} H) \ket{\psi}
-\end{align}
-\todo{inconsistent use of $\hbar$ vs $h$.}
-Exponential integrators that account for the non-linear
-term may be used to efficiently integrate the coupled dynamics,
-see Appendix \ref{appendix:derivative}.
-\todo{I think this also needs some more explanation. It's not 
-clear exactly what you're doing. You're penalizing the derivatives of the dynamics with
-respect to the uncertain parameter in the objective, so why do you need to augment the 
-state vector? Or are you adding the derivatives to the state vector so that you have 
-standard quadratic penalty terms in the objective? Again, I think this confusion would 
-be easily cleared up by explicitly writing down your new dynamics, objective function, 
-and any additional constraints you impose on the system. Since this is really the meat of 
-the paper, it's worth spending the time to make sure this is clear and easy to understand,
-especially for those who aren't very familiar with trajectory optimization or optimization 
-in general. Also, I think you should explicitly state the size of the problem you're
-actually solving. Might be worth putting in a table, along with some timing results, to 
-give an idea of how well these methods scale.}
-
-To demonstrate the applicability of these techniques
-to mitigate parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a constant qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$.
-
-We compare the numerical methods
-to an analytically derived $Z/2$ gate, see Figure \ref{fig:static}a 
-\todo{{\LaTeX} suggetion: don't hard-code the sublabels. You can directly reference them.}. 
-This gate corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. Its erroneous rotation angle $2 \pi t_{Z/2} \delta f_{q}$ is linearly sensitive to
-the qubit frequency detuning, resulting in a gate error that is quadratically sensitive
-to the qubit frequency detutning.
-At a one-percent
-qubit frequency detuning the analytic gate achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-which is sufficient for quantum error correction, see Figure \ref{fig:static}b.
-Although the analytic $Z/2$ gate performs well, it works
-only at the gate time $t_{Z/2}$. The ability to perform $Z$ rotations in arbitrary times is critical
-for operating multi-qubit experiments in the lab frame.
-Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
-These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
-
-%% TOOD: intro sentence
-%% TODO: comment on unscented when it is redone
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
-The derivative methods converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-Both methods achieve a super-linear scaling in their gate error as
-a function of the gate duration. The gate error for the 1\textsuperscript{st}-order
-derivative method approaches zero at the larmor period $1 / f_{q}$, see Figure \ref{fig:static}c.
-We believe the 1\textsuperscript{st}-order method outperforms the 2\textsuperscript{nd}-order
-method due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Appendix \ref{appendix:derivative}.
-
-Additionally, there are multiple analytic methods 
-to mitigate errors due to parameter deviations
-including composite pulse sequences \cite{merrill2014progress},
-the DRAG scheme \cite{krantz2019quantum}, and
-geometric phase considerations
-\cite{xu2020nonadiabatic, han2020experimental}.
-Composite pulse sequences are derived by computing
-the erroneous rotation arising from the parameter deviation
-with a finite order magnus expansion.
-Pulses are then suitably composed to eliminate the
-error. In principle the error may be eliminated
-to arbitrary order with sufficiently many pulses \cite{merrill2014progress},
-which mirrors the negative correlation between
-gate error and gate duration we observe with the derivative method.
-It is difficult to choose an appropriate composite pulse
-for the problem studied here,
-so we propose comparisons between composite pulses and numerical techniques
-for future work.

### 2020-11-28 paper/6stochastic.tex
prog

@@
-\section{Robustness to Stochastic Parameter Deviations \label{sec:stochastic}}
-An additional source of experimental error arises from stochastic, time-varying
+\section{Robustness to Time-Dependent Parameter Deviations \label{sec:stochastic}}
+An additional source of experimental error arises from time-dependent
 Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
@@
 source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
-It is well studied that the spectral density of $\delta a$ follows a
+modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a(t)$.
+It is well studied that the spectral density of $\delta a(t)$ follows a
 1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise (citations needed). Analytic methods to combat flux noise
+noise \todo{1/$f$ citations}. Analytic methods to combat flux noise
 take advantage of the low frequency characteristic and
@@
 filtering white noise sampled from a standard normal distribution with a finite
-impulse response filter \cite{saspweb2011}
-\footnote{\url{https://ccrma.stanford.edu/~jos/sasp/Example_Synthesis_1_F_Noise.html}}.
+impulse response filter \cite{saspweb2011}.
 It is then scaled by the 
 flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
+\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$. This filtered noise
+is then used to evolve the state under the Hamiltonian in \eqref{eq:hamiltonian}
+with the flux amplitude replaced $a(t) \gets a(t) + \delta a(t)$.
 The unscented sampling method is modified so that its sampled deviations
@@
 similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case. \todo{can you provide a little more 
+methods require no modification from the static case.
+\todo{can you provide a little more 
 detail on how this noise factors into the dynamics? Are you just adding some white 

### 2020-11-28 paper/a3derivative.tex
solid progress through sec4
prog

@@
-\section{Derivative Method}
-\label{appendix:derivative}
+\section{Derivative Method \label{appendix:derivative}}
 We comment on the optimization metrics of the derivative methods and then
@@
 
-\begin{table}[h]
+\begin{table}[ht]
   \begin{tabular}{c | c | c}
@@
 
-The dynamics for the derivative methods can be integrated efficiently using exponential integrators.
-General exponential integrators break the dynamics into a linear term and a non-linear term. For example, consider
+The dynamics for the derivative methods can be integrated efficiently using exponential integrators
+\cite{auer2018magnus, berland2005solving, einkemmer2017performance}.
+General exponential integrators break the dynamics into a linear term and a non-linear term.
+For example, consider
 integrating the dynamics of the first state derivative $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =

### 2020-11-28 paper/3fluxonium.tex
unlimited powah

@@
 and integrates the moments of the flux drive amplitude. Exposing lower order moments of the flux
-drive amplitude allows us to penalize their norms, smoothing the flux drive amplitude
+drive amplitude allows us to penalize their norms in \eqref{eq:costfun}, smoothing the flux drive amplitude
 and mitigating AWG ringing due to high frequency transitions.

### 2020-11-28 paper/5static.tex
unlimited powah

@@
 \section{Robustness to Static Parameter Deviations \label{sec:static}}
-We have formulated the quantum optimal control
+\todo{standardize gate time vs gate duration} We have formulated the quantum optimal control
 problem as an open loop optimization problem, i.e.
@@
 we mitigate errors arising from the drift and finite measurement
-precision of the qubit frequency in \eqref{eq:hamiltonian}.
+precision of the qubit frequency which modifies \eqref{eq:hamiltonian}
+by $f_{q} \gets f_{q} + \delta f_{q}$.
 We consider three robust control techniques:
@@
 
-The sampling method optimizes over multiple copies of a state,
+The sampling method simultaneously optimizes over copies of a state,
 each of which evolves with a distinct deviant parameter value.
-Two samples $\psi_{\pm}$
-are appended to the augmented state vector \eqref{eq:astatecontrols}.
-for each of four initial states
+A pair of two sample states $\psi_{\pm}$
+is appended to the augmented state vector \eqref{eq:astatecontrols}
+. The samples are initialized in the same state.
+Four pairs
+are appended in total corresponding to the initial states
 $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$. The initial states are chosen
+(\ket{0} - \ket{1}) / \sqrt{2}\}$.
+The initial states are chosen
 so that their outer products span the operators on the
@@
 samples evolve under the Hamiltonian given in \eqref{eq:hamiltonian}
-with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$. The final state
+with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$ for a fixed $\sigma_{f_{q}}$.
+The final state
 for each sample is the image of its initial state
@@
 the uncertainty in the state as a result of the uncertainty in
-the deviant parameter. Each sample evolves under the Hamiltonian given in
-\eqref{eq:hamiltonian}
-where the qubit frequency is replaced by a deviant value determined by the
-statistics of the ensemble of samples.
+the deviant parameter. Each sample evolves under the Hamiltonian
+given in \eqref{eq:hamiltonian} with $f_{q} \gets f_{q} + \delta f_{q}$.
+The qubit frequency detuning $\delta f_{q}$ is determined by the statistics
+of the ensemble of samples at each knot point.
 The discrete dynamics \eqref{eq:dyn_con} is modified to 
-evolve the samples under the TDSE dynamics \eqref{eq:tdse}
+evolve the samples under the deviant Hamiltonian
 and then apply the unscented transformation to the ensemble,
@@
 
-The derivative method draws on the intuition that
-the sensitivity of the state evolution to the deviant parameter
-is encoded in the ($l$\textsuperscript{th}-order)
-derivative of the state with respect to that parameter
-$\partial_{f_{q}}^{l} \psi$. In the $m$\textsuperscript{th}-order
+The derivative method penalizes the sensitivty of the state
+to the deviant parameter, which is encoded in the ($l$\textsuperscript{th}-order)
+derivative $\partial_{f_{q}}^{l} \psi$. In the $m$\textsuperscript{th}-order
 derivative method, all state derivatives of order $1, \dots, m$
 are appended to the augmented state vector \eqref{eq:astatecontrols}
-for each initial state in the operator basis. The norms of the state derivatives are penalized
+for each initial state in the operator basis.
+The norms of the state derivatives are penalized
 in \eqref{eq:costfun} by setting the corresponding final states to zero.
@@
 with backwards mode differentiation.
-Naively applying automatic differentiation
-would result in differentiating all dynamics at knot points
+Naive automatic differentiation
+would differentiate all dynamics at knot points
 $1, \dots, k - 1$ to obtain the state derivative at knot point $k$, requiring
@@
 
-To demonstrate the applicability of these techniques
-to mitigate parameter deviations,
-we consider the task of achieving a single $Z/2$
-gate subject to a static qubit frequency detuning
-$f_{q} \gets f_{q} + \delta f_{q}$.
+We task these methods with achieving
+a $Z/2$ gate subject to a static qubit frequency detuning.
 We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. For each method we compute the gate error for
-one simulated gate application subject to the deviant dynamics given by the
-stated qubit frequency detuning $\delta f_{q}$.
+the sampling methods accordingly. We compute gate errors for each method
+by evolving the optimized trajectory under the Hamiltonian in \eqref{eq:hamiltonain}
+with $f_{q} \gets f_{q} + \delta f_{q}$ at the stated qubit frequency detuning.
 
@@
 to an analytically derived $Z/2$ gate, see Figure \ref{fig:statica}. 
-This gate corresponds to
-idling at the flux frustration point $a = 0$. The analytic gate
+The analytic gate corresponds to
+idling at the flux frustration point $a = 0$. It
 is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
@@
 At a one-percent
-qubit frequency detuning the analytic gate achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
-which is sufficient for quantum error correction, see Figure \ref{fig:staticb}.
-Although the analytic $Z/2$ gate performs well, it works
-only at the gate time $t_{Z/2}$. The ability to perform $Z$ rotations in arbitrary times is critical
-for operating multi-qubit experiments in the lab frame.
+qubit frequency detuning ($\delta f_{q} / f_{q} = 1\%$)
+the analytic gate achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
+which is sufficient for quantum error correction.
+Although the analytic $Z/2$ gate performs well, it
+only works at the gate time $t_{Z/2}$. The ability to perform $Z$ rotations in arbitrary times is critical
+for multi-qubit experiments, where the qubits operate at different frequencies $f_{q, i} \neq f_{q, j}$.
 Each numerical method is able to find solutions at
-all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
+all gate times above $t_{Z/2}$, see Figure \ref{fig:staticb}.
 These numerical methods offer an effective scheme for synchronizing
-qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
+multi-qubit experiments.
 
-The sampling and unscented sampling methods
-converge on qualitatively similar solutions which combine idling periods
-with fast ramps to the maximum amplitude. The gate error at a one-percent
-detuning from the nominal qubit frequency achieved
-by the sampling method does not improve substantially over the
+The sampling and unscented sampling methods combine idling periods
+with time-anti-symmetric ramps.
+The gate error at $\delta f_{q} / f_{q} = 1\%$ for the sampling method
+does not improve substantially over the
 range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error with longer gate durations
+achieves linear decreases in its gate error at $\delta f_{q} / f_{q} = 1\%$ with longer gate durations
 until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
@@
 terms to the gate error in this deviation regime, see Appendix \ref{appendix:derivative}.
-\todo{@Dave should we still make a comment on leaving out the composite
-pulse comparison?}
-
-%% Additionally, there are multiple analytic methods 
-%% to mitigate errors due to parameter deviations
-%% including composite pulse sequences \cite{merrill2014progress},
-%% the DRAG scheme \cite{krantz2019quantum}, and
-%% geometric phase considerations
-%% \cite{xu2020nonadiabatic, han2020experimental}.
-%% Composite pulse sequences are derived by computing
-%% the erroneous rotation arising from the parameter deviation
-%% with a finite order magnus expansion.
-%% Pulses are then suitably composed to eliminate the
-%% error. In principle the error may be eliminated
-%% to arbitrary order with sufficiently many pulses \cite{merrill2014progress},
-%% which mirrors the negative correlation between
-%% gate error and gate duration we observe with the derivative method.
-%% It is difficult to choose an appropriate composite pulse
-%% for the problem studied here,
-%% so we propose comparisons between composite pulses and numerical techniques
-%% for future work.
+The vanishing gate error at longer gate durations exhibited by the derivative methods
+is reminiscent of the
+ability of composite pulses to mitigate parameter deviation errors to arbitrary
+order with sufficiently many pulses \cite{merrill2014progress}.
+It is difficult to choose an appropriate composite pulse
+for the problem studied here due to our Hamiltonian and experimental constraints.
+We propose comparisons between composite pulses and numerical techniques
+for future work.
 
@@
   \caption{
+    \todo{fix the size of the images and where are x axis labels?}
     (a) $X/2$ gates robust to flux offsets constructed with the analytic,

### 2020-11-28 paper/6stochastic.tex
unlimited powah

@@
 source of coherent errors. Magnetic flux noise
-modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a(t)$.
+modifies the fluxonium Hamiltonian \eqref{eq:hamiltonian}
+by $a(t) \gets a(t) + \delta a(t)$.
 It is well studied that the spectral density of $\delta a(t)$ follows a
@@
 
-We compare the analytic gate and those produced by
-the numerical methods discussed in the previous section
-on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
-The flux noise is generated by
+We task the analytic gate and those produced by
+the robust control methods of the previous section
+to realize an $X/2$ gate subject to 1/$f$ flux noise.
+We compute gate errors for each method
+by evolving the optimized trajectory under the Hamiltonian in \eqref{eq:hamiltonain}
+with $a(t) \gets a(t) + \delta a(t)$.
+The flux noise $\delta a(t)$ is generated by
 filtering white noise sampled from a standard normal distribution with a finite
@@
 flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$. This filtered noise
-is then used to evolve the state under the Hamiltonian in \eqref{eq:hamiltonian}
-with the flux amplitude replaced $a(t) \gets a(t) + \delta a(t)$.
-The unscented sampling method is modified so that its sampled deviations
-follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
-in the augmented state vector. In principle the basic sampling method could be modified
-similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
-methods require no modification from the static case.
-\todo{can you provide a little more 
-detail on how this noise factors into the dynamics? Are you just adding some white 
-noise to $a(t)$? Are you also considering uncertainty in $f_q$ at the same time?}
+\delta a (t) \sim \sigma_{a} = 2.5 \cdot 10^{-5} \textrm{GHz}$.
+The unscented sampling method is modified so that its samples
+are subject to 1/$f$ flux noise by carrying the state of a finite impulse response filter
+in the augmented state vector \eqref{eq:astatecontrols}.
+In principle the sampling method could be modified
+similarly but we choose to sample statically at $\sigma_{a}$ for comparision.
+The derivative methods require no algorithmic modification from the static case,
+but the TDSE is now differentiated with respect to $a(t)$ instead of $f_{q}$ as
+in \eqref{d1dyn}.
 
@@
 and numerical methods achieve single gate errors
-sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated gate application comparision. They achieve a two
+sufficient for quantum error correction.
+Despite converging on qualitatively different solutions, the
+numerical methods perform similarly in the concatenated
+gate application comparision. They achieve a two
 order of magnitude cumulative gate error reduction over the analytic method after $200$

### 2020-11-28 paper/8ack.tex
full build
unlimited powah

@@
-%% ACK
 \begin{acknowledgments}
   The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss for useful discussions.
+  and Daniel Weiss, Taylor Howell, and Tanay Roy for useful discussions. \todo{cite software packages}.
+  \todo{funding}.
 \end{acknowledgments}

### 2020-11-30 src/spin/spin.jl
updates for benchmarking

@@ -1030,26 +1030,28 @@ const DT_EN = Dict(
 )
 
 
-@inline gate_error_iso2a(s1::Array{T,1}, s2::Array{T,1};
-                         s1o::Int64=0, s2o::Int64=0) where {T} = (
-    s1[1 + s1o] * s2[1 + s20] + s1[2 + s1o] * s2[2 + s2o]
+@inline gate_error_iso2a(s1::AbstractVector, s2::AbstractVector;
+                         s1o::Int64=0, s2o::Int64=0) = (
+    s1[1 + s1o] * s2[1 + s2o] + s1[2 + s1o] * s2[2 + s2o]
     + s1[3 + s1o] * s2[3 + s2o] + s1[4 + s1o] * s2[4 + s2o]
 )
 
 
-@inline gate_error_iso2b(s1::Array{T,1}, s2::Array{T,1};
-                         s1o::Int64=0, s2o::Int64=0) where {T} = (
+@inline gate_error_iso2b(s1::AbstractVector, s2::AbstractVector;
+                         s1o::Int64=0, s2o::Int64=0) = (
     -s1[3 + s1o] * s2[1 + s2o] - s1[4 + s1o] * s2[2 + s2o]
     + s1[1 + s1o] * s2[3 + s2o] + s1[2 + s1o] * s2[4 + s2o]
 )
 
 
-@inline gate_error_iso2(s1::Array{T,1}, s2::Array{T,1}; s1o::Int64=0, s2o::Int64=0) where {T} = (
+@inline gate_error_iso2(s1::AbstractVector, s2::AbstractVector;
+                        s1o::Int64=0, s2o::Int64=0) = (
     1 - gate_error_iso2a(s1, s2; s1o=s1o, s2o=s2o)^2 - gate_error_iso2b(s1, s2; s1o=s1o, s2o=s2o)^2
 )
 
 
-function jacobian_gate_error_iso2(s1::Array{T,1}, s2::Array{T,1}; s1o::Int64=0, s2o::Int64=0) where {T}
+function jacobian_gate_error_iso2(s1::AbstractVector, s2::AbstractVector;
+                                  s1o::Int64=0, s2o::Int64=0)
     a = 2 * gate_error_iso2a(s1, s2; s1o=s1o, s2o=s2o)
     b = 2 * gate_error_iso2b(s1, s2; s1o=s1o, s2o=s2o)
     jac = [
[…]

### 2020-11-30 src/spin/spin11.jl

@@ -17,17 +17,13 @@ const TO = TrajectoryOptimization
 # paths
 const EXPERIMENT_META = "spin"
 const EXPERIMENT_NAME = "spin11"
-const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
+const SAVE_PATH = abspath(joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME))
 
 # problem
 const CONTROL_COUNT = 1
 const STATE_COUNT = 2
 const ASTATE_SIZE_BASE = STATE_COUNT * HDIM_ISO + 3 * CONTROL_COUNT
 const ACONTROL_SIZE = CONTROL_COUNT
-const INITIAL_STATE1 = [1., 0, 0, 0]
-const INITIAL_STATE2 = [0., 1, 0, 0]
-const INITIAL_STATE3 = [1., 0, 0, 1] ./ sqrt(2)
-const INITIAL_STATE4 = [1., -1, 0, 0] ./ sqrt(2)
 const SAMPLE_COUNT = 4
 # state indices
 const STATE1_IDX = 1:HDIM_ISO
@@ -63,16 +59,16 @@ function RD.discrete_dynamics(::Type{RK3}, model::Model{DO}, astate::SVector,
                               acontrol::SVector, time::Real, dt::Real) where {DO}
     negi_h = (
         FQ_NEGI_H0_ISO
-        + astate[CONTROLS_IDX][1] * NEGI_H1_ISO
+        + astate[CONTROLS_IDX[1]] * NEGI_H1_ISO
     )
     h_prop = exp(negi_h * dt)
     state1_ = astate[STATE1_IDX]
     state2_ = astate[STATE2_IDX]
     state1 =  h_prop * state1_
     state2 = h_prop * state2_
-    intcontrols = astate[INTCONTROLS_IDX] + astate[CONTROLS_IDX] * dt
-    controls = astate[CONTROLS_IDX] + astate[DCONTROLS_IDX] * dt
-    dcontrols = astate[DCONTROLS_IDX] + acontrol[D2CONTROLS_IDX] * dt
+    intcontrols = astate[INTCONTROLS_IDX[1]] + astate[CONTROLS_IDX[1]] * dt
+    controls = astate[CONTROLS_IDX[1]] + astate[DCONTROLS_IDX[1]] * dt
[…]

### 2020-11-30 src/spin/spin12.jl

@@ -1,5 +1,8 @@
 """
-spin12.jl - sampling robustness
+spin12.jl - sampling robustness for the δf problem
+
+This optimization uses the infidelity metric rather than
+the standard diagonal LQR metric.
 """
 
 WDIR = joinpath(@__DIR__, "../../")
@@ -26,10 +29,6 @@ const SAMPLE_COUNT = 8
 const ASTATE_SIZE_BASE = STATE_COUNT * HDIM_ISO + 3 * CONTROL_COUNT
 const ASTATE_SIZE = ASTATE_SIZE_BASE + SAMPLE_COUNT * HDIM_ISO
 const ACONTROL_SIZE = CONTROL_COUNT
-const INITIAL_STATE1 = [1., 0, 0, 0]
-const INITIAL_STATE2 = [0., 1, 0, 0]
-const INITIAL_STATE3 = [1., 0, 0, 1] ./ sqrt(2)
-const INITIAL_STATE4 = [1., -1, 0, 0] ./ sqrt(2)
 # state indices
 const STATE1_IDX = 1:HDIM_ISO
 const STATE2_IDX = STATE1_IDX[end] + 1:STATE1_IDX[end] + HDIM_ISO
@@ -107,31 +106,33 @@ end
 
 @inline TO.stage_cost(cost::Cost{N,M,T}, astate::SVector{N}) where {N,M,T} = (
     0.5 * astate' * cost.Q * astate + cost.q'astate + cost.c
-    + cost.q_ss1 * gate_error_iso2(astate, cost.target_states[1], S1_IDX[1] - 1)
-    + cost.q_ss2 * gate_error_iso2(astate, cost.target_states[2], S2_IDX[1] - 1)
-    + cost.q_ss3 * gate_error_iso2(astate, cost.target_states[3], S3_IDX[1] - 1)
-    + cost.q_ss4 * gate_error_iso2(astate, cost.target_states[4], S4_IDX[1] - 1)
-    + cost.q_ss1 * gate_error_iso2(astate, cost.target_states[1], S5_IDX[1] - 1)
-    + cost.q_ss2 * gate_error_iso2(astate, cost.target_states[2], S6_IDX[1] - 1)
-    + cost.q_ss3 * gate_error_iso2(astate, cost.target_states[3], S7_IDX[1] - 1)
-    + cost.q_ss4 * gate_error_iso2(astate, cost.target_states[4], S8_IDX[1] - 1)
[…]

### 2020-11-30 src/spin/spin15.jl

@@ -11,19 +11,18 @@ using LinearAlgebra
 using RobotDynamics
 using StaticArrays
 using TrajectoryOptimization
+const RD = RobotDynamics
 const TO = TrajectoryOptimization
 
 # paths
 const EXPERIMENT_META = "spin"
 const EXPERIMENT_NAME = "spin15"
-const SAVE_PATH = joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME)
+const SAVE_PATH = abspath(joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME))
 
 # problem
 const CONTROL_COUNT = 1
 const STATE_COUNT = 2
 const ASTATE_SIZE_BASE = STATE_COUNT * HDIM_ISO + 3 * CONTROL_COUNT
-const INITIAL_STATE1 = [1., 0, 0, 0]
-const INITIAL_STATE2 = [0., 1, 0, 0]
 # state indices
 const STATE1_IDX = 1:HDIM_ISO
 const STATE2_IDX = STATE1_IDX[end] + 1:STATE1_IDX[end] + HDIM_ISO
@@ -41,34 +40,32 @@ const DT_IDX = D2CONTROLS_IDX[end] + 1:D2CONTROLS_IDX[end] + 1
 struct Model{DA, TO} <: AbstractModel
     Model(DA::Bool=true, TO::Bool=true) = new{DA, TO}()
 end
-RobotDynamics.state_dim(::Model{false, TO}) where TO = ASTATE_SIZE_BASE
-RobotDynamics.state_dim(::Model{true, TO}) where TO = ASTATE_SIZE_BASE + 1
-RobotDynamics.control_dim(::Model{DA, false}) where DA = CONTROL_COUNT
-RobotDynamics.control_dim(::Model{DA, true}) where DA = CONTROL_COUNT + 1
+@inline RD.state_dim(::Model{false, TO}) where TO = ASTATE_SIZE_BASE
+@inline RD.state_dim(::Model{true, TO}) where TO = ASTATE_SIZE_BASE + 1
+@inline RD.control_dim(::Model{DA, false}) where DA = CONTROL_COUNT
+@inline RD.control_dim(::Model{DA, true}) where DA = CONTROL_COUNT + 1
 
 
 # dynamics
-function RobotDynamics.discrete_dynamics(::Type{RK3}, model::Model{DA, TO}, astate::StaticVector,
[…]

### 2020-11-30 paper/1introduction.tex
updated figures, lots of proofreading

@@
 \section{Introduction}
+\todo{Ask Dave to review paragraph}
 Quantum optimal control (QOC) techniques are a class of optimization
@@
 We adapt robust control techniques from the robotics community to mitigate
-realistic model uncertainties for
+parameter deviation errors for
 a superconducting fluxonium qubit.
@@
 \cite{han2020experimental, xu2020nonadiabatic},
-and the DRAG scheme \todo{DRAG citations}.
+and the DRAG scheme \cite{motzoi2009simple}.
 To mitigate decoherence, Floquet techniques have been employed
@@
   \item A derivative-penalization method, which uses efficient mixed-mode differentiation
-    to compute derivative information of parameter deviations with respect to the 
-    quantum state trajectory.
+    to compute the sensitivity of the quantum state trajectory
+    to deviant parameters.
 \end{enumerate}

### 2020-11-30 paper/2background.tex
updated figures, lots of proofreading

@@
 In this section we
-review the quantum optimal control problem statement
+review the QOC problem statement
 and introduce the ALTRO solver.
-Quantum optimal control concerns the evolution of
-a quantum state $\ket{\psi(t)}$ governed by the time-dependent
+QOC concerns the evolution of
+a (quantum) state $\ket{\psi(t)}$ governed by the time-dependent
 Schr{\"o}dinger equation (TDSE),
@@
 The Hamiltonian has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so called because they are the means the experimentalist has to
+The controls are so-called because they are the means the experimentalist has to
 act on the system. To make the problem numerically tractable,
@@
 controls that minimize a functional $J(u)$.
-In the simplest case the functional is
-the infidelity between the inital state evolved
-to the final knot point and the target state
+In the simplest case, the functional is
+the infidelity between the initial state evolved
+to the final knot point and the desired final state
 $J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u)} \rvert}^{2}$.
-In general, $J$ is a linear combinaion of cost functions on the state
+In general, $J$ is a linear combination of cost functions on the state
 as well as cost functions on the controls \cite{leung2017speedup}.
-Typical quantum optimal control
-algorithms employ automatic differentiation
+Standard QOC
+solvers employ automatic differentiation
 to compute gradients of the functional $\nabla_{u} J(u)$,
@@
 \begin{mini}[2]
-    {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell(x_k,u_k) }{}{}
+    {x_{1:N},u_{1:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell(x_k,u_k) }{}{}
     \addConstraint{x_{k+1} = f(x_k,u_k)}
@@
 \end{mini}
-where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^n$ and
+where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^l$ and
 $u_k \in \R^m$ are the augmented state and input augmented control variables,
@@
 \end{mini}
-where $f(z)$ is the objective function, $c(z) : \R^n \mapsto \R^p$ is the constraint 
+where $f(z)$ is the objective function, $c(z) : \R^{l + m} \mapsto \R^p$ is the constraint 
 function, $\lambda \in \R^p$ is a Lagrange multiplier estimate, and $I_\mu$ is a diagonal matrix
@@
 
+\todo{Ask Dave to review paragraph}
 As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
 for imposing arbitrary, simultaneous constraints
-on both the state trajectory and the controls and does not restrict
+on both the state and the controls and does not restrict
 optimization to the constraint manifold. First-order solvers like GRAPE
-rely on unconstrained optimization to satisfy desiderata on the state trajectory,
+rely on unconstrained optimization to satisfy desiderata on the state,
 such as achieving a desired gate fidelity, forbidding occupations of particular quantum states,
@@
 solve is already at a coarse tolerance, achieving machine-precision satisfaction
-for constraints on the controls and the state trajectory. These advantage are critical
+for constraints on the controls and the state. These advantages are critical
 for this work, where multiple constraints must be satisfied in addition

### 2020-11-30 paper/3fluxonium.tex
updated figures, lots of proofreading

--- 2020-11-28 paper/3fluxonium.tex
@@
 \section{QOC on the Fluxonium \label{sec:fluxonium}}
-In the following we study
-the quantum optimal control problem on the fluxonium qubit.
+In the following, we study
+the QOC problem on the fluxonium qubit.
 The fluxonium qubit is a promising building block for superconducting
@@
 two-level approximation of its Hamiltonian makes
-quantum optimal control on a classical computer inexpensive.
+QOC on a classical computer inexpensive.
 In the two-level
@@
 \begin{mini!}[2]
-  {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
+  {x_{1:N}, u_{1:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
     + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{} \label{eq:costfun}
     \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
-    \addConstraint{(x_1, x_N \; \textrm{given})}
+    \addConstraint{(x_1 \; \textrm{given})}
     \addConstraint{g_k(x_k,u_k)}{\leq 0}
@@
 \end{mini!}
-where the augmented state and controls are given by:
+The augmented state and augmented controls are:
 \begin{equation}
-  x = \begin{bmatrix} \psi^{(1)} \\ \psi^{(2)} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
+  x = \begin{bmatrix} \psi^{1} \\ \psi^{2} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
   u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
@@
 \end{equation}
-The initial and final conditions are determined by the
-desired gate, which requires two quantum states
-to specify $\psi^{(1)}$, $\psi^{(2)}$.
+The initial states $\psi^{1}_{1}$, $\psi^{2}_{1}$
+are basis vectors of the Hilbert space $\ket{0}, \ket{1}$.
+The final states $\psi^{1}_{f}$, $\psi^{2}_{f}$ are the image
+of the initial states under the desired gate.
 The ALTRO implementation we use does not currently support complex numbers so
-we compute in the isomorphism $\mathcal{H}(\mathbb{R}^{2n}) \cong \mathcal{H}(\mathbb{C}^{n})$
-given in \cite{leung2017speedup}:
+we compute in the isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
+given in \cite{leung2017speedup},
 \begin{equation}
@@
 \end{equation}
-The discrete dynamics function \eqref{eq:dyn_con} integrates the TDSE dynamics for the quantum states
+The discrete dynamics function \eqref{eq:dyn_con} integrates the states
 using \eqref{eq:tdse} and \eqref{eq:hamiltonian}
-and integrates the moments of the flux drive amplitude. Exposing lower order moments of the flux
-drive amplitude allows us to penalize their norms, smoothing the flux drive amplitude
+and integrates the moments of the flux amplitude.
+Exposing lower order moments of the flux amplitude allows us to penalize their norms
+in \eqref{eq:costfun}, smoothing the flux amplitude
 and mitigating AWG ringing due to high frequency transitions.
@@
 Including objectives at each knot point smoothens the optimization landscape, though
-the importance of the final state objective is encoded in the relative
+the importance of the final augmented state objective is encoded in the relative
 weight $Q_{N} \sim N \cdot Q_{k}$.
-We choose $Q$ to be diagonal because it is computationally efficient. This corresponds
-to penalizing phase differences between the quantum states, although
-other phase-insensitive metrics such as infidelity may be employed.
+We choose $Q$ to be diagonal because it is computationally efficient. This
+metric penalizes phase differences between the states, although
+phase-insensitive metrics such as infidelity may be employed.
 
-In addition to imposing optimization constraints that
-reflect physical limitations of the apparatus, we impose
+We impose constraints that
+reflect the physical limitations of the apparatus and 
 constraints that improve the experimental realization of the control pulse.
@@
 inducing AWG ringing due to high-frequency transitions,
-we require $a(t = 0) = a(t = t_{N}) = 0$.
-Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
+we require $a_{1} = a_{N} = 0$.
+Furthermore, we require $\int a_{N} \ dt = 0$. This
 constraint ensures the pulse has zero net flux, mitigating
@@
 \cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
-We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
+We require $-0.5 \textrm{GHz} \le a_{k} \le 0.5 \textrm{GHz}$
 to ensure the two-level approximation \eqref{eq:hamiltonian}
-remains valid. Additionally, we require that each gate achieves
-the desired state transition $\psi_{N} = \psi_{f}$.
-Both the zero net flux and target quantum state constraint
-are then handled by ensuring the target augmented state is
+remains valid. Additionally, we require
+$\psi_{N} = \psi_{f}$.
+The zero net flux and final state constraints
+are satisfied if the final augmented state is
 reached $x_{N} = x_{f}$, the constraint function
@@
 $c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
-The equality and inequality constraints on $a$ are handled
+The inequality constraints on $a$ are handled
 with a bound constraint which takes the form
@@
 and $0$ otherwise.
-In addition, the norm of the quantum states
+In addition, the norm of the states
 are constrained to $1$ to ensure discretization error is not

### 2020-11-30 paper/4longitude.tex
updated figures, lots of proofreading

@@
+%% F1
 \begin{figure*}[ht]
@@
   \caption{
-    (a) Numerically optimized gates (red) and analytically optimized gates (blue).
+    \todo{Ask Dave to okay figure color scheme}
+    (a) Numerically optimized gates (dark) and analytically optimized gates (light).
     (b) $T_{1}$ interpolation function used in optimization. Markers
@@
 of the quantum state.
-Computing the gate error due to longitudinal relaxation
-requires propagating density matrices of size $n \times n$ under master equation
-dynamics, rather than state vectors of size $n$ under the TDSE dynamics \eqref{eq:tdse}.
+Typical approaches completely model the gate error due to longitudinal
+relaxation by performing optimal control on $n^{2}$
+density matrices of size $n \times n$ that evolve under master
+equation dynamics, rather than $n$ states vectors of size $n$
+under the TDSE \eqref{eq:tdse}.
 We avoid this increase in computational complexity by
 penalizing the probability (integrated rate) of longitudinal relaxation.
-Using this probability as proxy for the gate error incurred
+Using this probability as a proxy for the gate error incurred
 is reasonable because losses due to longitudinal relaxation increase monotonically
@@
 error channels which share the monotonically increasing property.
-Additionally, for a constant $T_{1}$ time, a shorter gate duration
+Additionally, for a constant $T_{1}$ time, a shorter gate time
 would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate duration in order to minimize the
+the optimizer to tune the gate time to minimize the
 longitudinal relaxation probability. Our scheme for time-optimal
-control is applicable to any time-optimal problem, not only
+control is suitable for any time-optimal problem, not only
 the one we study here.
 
+\todo{Ask Brian to review this paragraph and next}
 The longitudinal relaxation probability is given by,
@@
 
-We allow the optimizer to tune the gate duration by
+We allow the optimizer to tune the gate time by
 making the time step between each knot point $\Delta t_{k}$
@@
   \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
+    \includegraphics[width=\linewidth]{assets/f2b.png}
     \caption{\label{fig:staticb}}
@@
   \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
+    \includegraphics[width=\linewidth]{assets/f2c.png}
     \caption{\label{fig:staticc}}
@@
     (b) Single gate error as
-    a function of the gate duration at a one-percent
-    detuning from the nominal qubit frequency for all methods. Missing
+    a function of the gate time at a one-percent
+    qubit frequency detuning for all methods. Missing
     data points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the detuning from the nominal
-    qubit frequency. The solutions for the analytic and
+    (c) Single gate error as a function of the qubit frequency detuning.
+    The solutions for the analytic and
     1\textsuperscript{st}-order derivative methods are shown at multiples
-    of the analytic gate time. The performance of the two methods is
+    of the analytic gate time. The performance for the two methods is
     indistinguishable at the analytic gate time $18$ns.
@@
 their single gate errors by a factor of $5$ over
-their analytic counterparts which is commensurate to their
+their analytic counterparts, commensurate to their
 probability of longitudinal relaxation reductions, see Appendix \ref{appendix:longitude}.
 The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
-generated initial states. The numerical $Z/2$ and $Y/2$ gates perform simlarly in
-the concatenated gate application comparision, suppressing cumulative gate errors to $8 \cdot 10^{-3}$
+of the evolved state and the final state averaged over 1000 pseudo-randomly
+generated initial states. The numerical $Z/2$ and $Y/2$ gates perform similarly in
+the concatenated gate application comparison, suppressing cumulative gate errors to $8 \cdot 10^{-3}$
 over $2000$ gate applications $\sim 40 \mu\textrm{s}$, see Figure \ref{fig:longitudec}.

### 2020-11-30 paper/5static.tex
updated figures, lots of proofreading

@@
 \section{Robustness to Static Parameter Deviations \label{sec:static}}
-\todo{standardize gate time vs gate duration} We have formulated the quantum optimal control
-problem as an open loop optimization problem, i.e.
-feedback from the experiment is not incorporated in optimization.
+We have formulated the QOC
+problem as an open-loop optimization problem, i.e.
+we do not incorporate feedback from the experiment in optimization.
 However, the device typically deviates from the Hamiltonian we use in optimization,
@@
 
+\todo{Ask Brian to review this paragraph and next two}
 The sampling method simultaneously optimizes over copies of a state,
 each of which evolves with a distinct deviant parameter value.
-A pair of two sample states $\psi_{\pm}$
-is appended to the augmented state vector \eqref{eq:astatecontrols}
-. The samples are initialized in the same state.
-Four pairs
-are appended in total corresponding to the initial states
+We append four sets of two sample states $\psi_{\pm}$
+to the augmented state vector \eqref{eq:astatecontrols}.
+We initialize each pair of samples in the same state.
+The four initial states are chosen so that their outerproducts
+span the operators on the Hilbert space
 $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$.
-The initial states are chosen
-so that their outer products span the operators on the
-Hilbert space  \cite{chow2009randomized}.
+(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
 The discrete dynamics \eqref{eq:dyn_con} is modified so that the
@@
 for each sample is the image of its initial state
-under the desired gate. Deviations from the final state are
-penalized in the objective \eqref{eq:costfun}.
+under the desired gate. We penalize deviations
+from the final state in the objective \eqref{eq:costfun}.
 
-For the unscented sampling method, $2(2n + d)$ samples
-are appended to the augmented state vector \eqref{eq:astatecontrols}
-for each initial state from the operator basis. Here $2n = 4$ is twice the
+For the unscented sampling method, we append $2(2n + d)$ samples
+to the augmented state vector \eqref{eq:astatecontrols}
+for each initial state in the operator basis. Here $2n = 4$ is twice the
 dimension of the Hilbert space, resulting from the isomorphism \eqref{eq:isomorphism},
 and $d = 1$ is the number of deviant parameters. The samples
-are used to represent a Gaussian distribution over all $2n$
+represent a Gaussian distribution over the $2n$
 elements of the initial state, modeling
@@
 given in \eqref{eq:hamiltonian} with $f_{q} \gets f_{q} + \delta f_{q}$.
-The qubit frequency detuning $\delta f_{q}$ is determined by the statistics
-of the ensemble of samples at each knot point.
-The discrete dynamics \eqref{eq:dyn_con} is modified to 
+We determine the qubit frequency detuning $\delta f_{q}$
+at each knot point using the statistics of the samples.
+We modify the discrete dynamics \eqref{eq:dyn_con} to
 evolve the samples under the deviant Hamiltonian
-and then apply the unscented transformation to the ensemble,
+and then apply the unscented transformation to the ensemble
+of samples,
 accurately preserving the first and second moments
@@
 As in the sampling method,
-deviations from the final state are penalized according to \eqref{eq:costfun}.
+we penalize deviations from the final state according to \eqref{eq:costfun}.
 A detailed procedure for the unscented transformation is given
@@
 derivative $\partial_{f_{q}}^{l} \psi$. In the $m$\textsuperscript{th}-order
-derivative method, all state derivatives of order $1, \dots, m$
-are appended to the augmented state vector \eqref{eq:astatecontrols}
+derivative method, we append all state derivatives of order $1, \dots, m$
+to the augmented state vector \eqref{eq:astatecontrols}
 for each initial state in the operator basis.
-The norms of the state derivatives are penalized
+We penalize the norms of the state derivatives
 in \eqref{eq:costfun} by setting the corresponding final states to zero.
-The state derivatives could be obtained in the discrete dynamics \eqref{eq:dyn_con}
+We could obtain the state derivatives in the discrete dynamics \eqref{eq:dyn_con}
 with backwards mode differentiation.
@@
 which require $O(N)$ matrix multiplications to integrate.
-For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are,
+For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
 \begin{align}
@@
 Exponential integrators that account for the non-linear
-term in \eqref{eq:d1dyn} may be used to efficiently integrate the coupled dynamics,
-see Appendix \ref{appendix:derivative}. For timing results
-and an analysis of how the three robust methods scale to larger Hilbert spaces
-and more deviant parameters,
+term in \eqref{eq:d1dyn} are used to efficiently integrate the coupled dynamics,
+see Appendix \ref{appendix:derivative}. For timing
+results and complexity analyses for the three methods,
 consult Appendix \ref{appendix:time}.
 
-We task these methods with achieving
+We challenge these methods to achieve
 a $Z/2$ gate subject to a static qubit frequency detuning.
@@
 the sampling methods accordingly. We compute gate errors for each method
-by evolving the optimized trajectory under the Hamiltonian in \eqref{eq:hamiltonain}
-with $f_{q} \gets f_{q} + \delta f_{q}$ at the stated qubit frequency detuning.
+by evolving the system under the Hamiltonian in \eqref{eq:hamiltonian}
+with the optimized flux amplitude and $f_{q} \gets f_{q} + \delta f_{q}$
+at the stated qubit frequency detuning.
 
@@
 is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. Its erroneous rotation angle $2 \pi t_{Z/2} \delta f_{q}$ is linearly sensitive to
+is simple to derive. Its erroneous rotation angle
+$2 \pi t_{Z/2} \delta f_{q}$ is linearly sensitive to
 the qubit frequency detuning, resulting in a gate error that is quadratically sensitive
-to the qubit frequency detutning.
+to the qubit frequency detuning.
 At a one-percent
@@
 Although the analytic $Z/2$ gate performs well, it
-only works at the gate time $t_{Z/2}$. The ability to perform $Z$ rotations in arbitrary times is critical
-for multi-qubit experiments, where the qubits operate at different frequencies $f_{q, i} \neq f_{q, j}$.
-Each numerical method is able to find solutions at
+only works at the gate time $t_{Z/2}$. The ability to perform $Z$
+rotations in arbitrary times is critical
+for multi-qubit experiments, where the qubits operate at different
+frequencies $f_{q, i} \neq f_{q, j}$.
+Each numerical method can find solutions at
 all gate times above $t_{Z/2}$, see Figure \ref{fig:staticb}.
@@
 does not improve substantially over the
-range of gate durations. The unscented sampling method
-achieves linear decreases in its gate error at $\delta f_{q} / f_{q} = 1\%$ with longer gate durations
-until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
+range of gate times. The unscented sampling method
+achieves linear decreases in its gate error at
+$\delta f_{q} / f_{q} = 1\%$ with longer gate times
+until half the Larmor period $1 / 2 f_{q}$ after which it achieves a consistent
 gate error $\sim 3.5 \cdot 10^{-5}$.
@@
 The derivative methods converge on qualitatively similar solutions that
-use fast traingle pulses at the boundaries and balance time
+use fast triangle pulses at the boundaries and balance time
 on either side of the flux-frustration point symmetrically at low amplitudes.
 Both methods achieve a super-linear scaling in their gate error as
-a function of the gate duration. The gate error for the 1\textsuperscript{st}-order
-derivative method approaches zero at the larmor period $1 / f_{q}$, see Figure \ref{fig:staticc}.
+a function of their gate time.
 We believe the 1\textsuperscript{st}-order method outperforms the 2\textsuperscript{nd}-order
@@
 terms to the gate error in this deviation regime, see Appendix \ref{appendix:derivative}.
-The vanishing gate error at longer gate durations exhibited by the derivative methods
-is reminiscent of the
+The gate error for the 1\textsuperscript{st}-order
+derivative method approaches zero at the Larmor period $1 / f_{q}$, see Figure \ref{fig:staticc}.
+This result mimics the
 ability of composite pulses to mitigate parameter deviation errors to arbitrary
@@
     \label{fig:stochastica}
-  \end{subfigure}\hspace{0.025\textwidth}
+  \end{subfigure}\hspace{0.05\textwidth}
   \begin{subfigure}{.4\textwidth}
@@
   \caption{
-    \todo{fix the size of the images and where are x axis labels?}
-    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
+    (a) $X/2$ gates robust to flux amplitude offsets constructed with the analytic,
     sampling, unscented sampling, and the 1\textsuperscript{st}-

### 2020-11-30 paper/6stochastic.tex
updated figures, lots of proofreading

@@
 by $a(t) \gets a(t) + \delta a(t)$.
-It is well studied that the spectral density of $\delta a(t)$ follows a
+The spectral density of $\delta a(t)$ follows a
 1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise \todo{1/$f$ citations}. Analytic methods to combat flux noise
-take advantage of the low frequency characteristic and
+noise \cite{bialczak20071f, koch2007model,
+  yoshihara2006decoherence, yoshihara2010correlated}.
+Analytic methods to combat flux noise
+take advantage of the low-frequency characteristic and
 treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. This is the strategy employed by the analytic gate
-considered here.
+to compensate for erroneous drift. The analytic gate considered here
+follows this strategy.
 
@@
 We compute gate errors for each method
-by evolving the optimized trajectory under the Hamiltonian in \eqref{eq:hamiltonain}
-with $a(t) \gets a(t) + \delta a(t)$.
+by evolving system under the Hamiltonian in \eqref{eq:hamiltonian}
+where the optimized flux amplitude is modified $a(t) \gets a(t) + \delta a(t)$.
 The flux noise $\delta a(t)$ is generated by
@@
 in the augmented state vector \eqref{eq:astatecontrols}.
-In principle the sampling method could be modified
-similarly but we choose to sample statically at $\sigma_{a}$ for comparision.
+In principle, we could modify the sampling method
+similarly but we choose to sample statically at $\sigma_{a}$ for comparison.
 The derivative methods require no algorithmic modification from the static case,
 but the TDSE is now differentiated with respect to $a(t)$ instead of $f_{q}$ as
-in \eqref{d1dyn}.
+in \eqref{eq:d1dyn}.
 
@@
 numerical methods perform similarly in the concatenated
-gate application comparision. They achieve a two
+gate application comparison. They achieve a two
 order of magnitude cumulative gate error reduction over the analytic method after $200$
 gate applications $\sim 11 \mu\textrm{s}$.
-1/$f$ flux noise is a signficiant source of coherent errors in NISQ applications and
+1/$f$ flux noise is a significiant source of coherent errors in NISQ applications and
 these numerical techniques offer effective avenues to mitigate them.

### 2020-11-30 paper/7conclusion.tex
updated figures, lots of proofreading

--- 2020-11-03 paper/7conclusion.tex
@@
-%% S7
 \section{Conclusion}
-In conclusion, we have demonstrated techniques for achieving robustness to systematic
-errors and mitigating decoherence on a quantum system using state-of-the-art trajectory
-optimization. We have proposed a scheme for mitigating longitudinal relaxation with time-optimal
-control and an efficient optimization metric that comes at a constant computational cost as
-opposed to integrating a master equation which scales quadratically with
+In conclusion, we have applied state-of-the-art trajectory
+optimization techniques to achieve robustness to parameter deviations
+and mitigate decoherence on a quantum system.
+We have proposed a scheme for mitigating
+longitudinal relaxation with time-optimal
+control and the probability of longitudinal relaxation metric,
+avoiding the optimal control problem for density matrices which scales quartically with
 the dimension of the Hilbert space.
 We have proposed the derivative method for robust control which achieves
-super-linear gate error reductions in the gate duration for the problem we studied here.
-We have shown that robust control techniques can be used to mitigate decoherence due
-to 1/$f$ flux noise, a dominant source of coherent errors for flux controlled qubits.
-The numerical techniques we have studied here can be used to perform phase gates in arbitrary times,
-which will be critical for synchronizing multi-qubit systems. Our
-techniques significantly improve the state-of-the-art for controlling quantum systems subject
-to systematic errors and decoherence, and will be used in efforts to lower gate errors
-for fault-tolerant quantum computing and NISQ applications.
+super-linear gate error reductions in its gate time for the static parameter
+deviation problem we studied.
+We have shown that the derivative, sampling, and unscented sampling methods
+can mitigate 1/$f$ flux noise errors
+which dominate coherent errors for flux controlled qubits.
+The derivative and sampling methods scale cubically with
+the dimension of the Hilbert space, allowing for applications
+to larger quantum systems.
+The derivative, sampling, and unscented sampling
+methods will benefit from interleaving optimization with
+experimental characterization, while the sampling and unscented
+sampling methods can also be applied to error channels without models.
+These techniques will be used to achieve the low gate errors
+required for fault-tolerant quantum computing applications. Our
+implementation of the techniques described in this work is available
+at \url{https://github.com/SchusterLab/rbqoc}.

### 2020-11-30 paper/a1longitude.tex
updated figures, lots of proofreading

@@
-\section{Longitudinal Relaxation}
-\label{appendix:longitude}
+\section{Longitudinal Relaxation \label{appendix:longitude}}
 We comment on the longitudinal relaxation metrics and then give
 our procedure for integrating the Lindblad master equation.
-The longitudal relaxation probability and the gate error due to longitudinal
-relaxation metrics are compared in Table \ref{tab:longitude} for the experiment shown in Figure \ref{fig:longitude}c.
+The longitudinal relaxation probability and the gate error due to longitudinal
+relaxation metrics are compared in Table \ref{tab:longitude} for the
+numerical experiment described in Section \ref{sec:longitude}.
 The relative performance of the analytic and numerical techniques
@@
   \caption{
-    Single gate longitudinal relaxation probability ratios
-    and single gate error due to longitudinal relaxation ratios.
-    Values are reported for the analytic (A) and numerical (N) techniques
-    presented in the main text.
+    Single gate longitudinal relaxation probability (P) ratios
+    and single gate error due to longitudinal relaxation (GE) ratios.
+    Values are reported for the analytic (A) and numerical (N) techniques.
   }
@@
 
-To compute the gate error due to longitudinal relaxation,
-we require the final state of the quantum system subject
-to longitudinal relaxation. To obtain the final state we employ
-the Lindblad master equation. This equation takes the form
+To compute the gate error due to longitudinal relaxation
+we employ the Lindblad master equation. This equation takes the form:
 \begin{equation}
@@
 \end{equation}
-Here $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $n = \textrm{dim}(\mathcal{H})$,
-and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
+where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $n$ is the
+dimension of the Hilbert space,
+$[\cdot, \cdot]$ is the algebraic commutator, and $\{\cdot, \cdot \}$ is the algebraic
+anti-commutator.
 For longitudinal relaxation $\gamma_{\uparrow} = T_{1, \uparrow}^{-1}$,
@@
 $L_{\downarrow} = \sigma^{-}/2$ where $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$.
-Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained from the spline shown in Figure \ref{fig:longitude}b.
-The $T_{1}$ values in this spline are obtained experimentally by driving the qubit at the desired flux amplitude
-and monitoring the resultant decay. For more details consult \cite{zhang2020universal}.
+Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained from the spline
+shown in Figure \ref{fig:longitudeb}.
+We obtain the $T_{1}$ values in this spline
+by driving the qubit at the desired flux amplitude
+and monitoring the resultant decay.
+For more details consult \cite{zhang2020universal}.
 
 Exponential integrators can be employed to integrate the Lindblad master equation
-using the Vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018}
+using the Vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018},
 \begin{equation}
@@
 \end{equation}
-Here $\rho = \sum_{i, j} \alpha_{i, j} \ket{i}\bra{j}$
+where $\rho = \sum_{i, j} \alpha_{i, j} \ket{i}\bra{j}$
 and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{i, j} \ket{i} \otimes \ket{j}$.
-We use zero-order hold on the controls so the integration is exact
+We use zero-order hold on the controls so the exact solution is
 $\textrm{vec}(\rho_{k + 1}) = \exp(\Delta t_{k} \hat{\mathcal{L}}_{k}) \textrm{vec}(\rho_{k})$.
@@
 matrix-matrix multiplications to $(n^{4} \times n^{4}) \times n^{4}$ matrix-vector
-multiplications. For small $n$ and zero-order held controls, we find that it is
+multiplications. For small $n$ and zero-order hold on the controls, we find that it is
 faster to use an exponential integrator on the vectorized equation than to perform

### 2020-11-30 paper/a2unscented.tex
updated figures, lots of proofreading

@@
 \label{appendix:unscented}
-In this section we outline the full unscented sampling procedure.
-The unscented sampling method concerns a quantum state $\psi \in \mathbb{R}^{2n}$ with
-deviant parameters $\lambda \in \mathbb{R}^{d}$ and dynamics $\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
-The nominal initial quantum state is given by $\psi_{0}$ with an associated
-positive-definite covariance matrix $P_{0} \in \mathbb{R}_{++}^{n \times n}$
+In this section, we outline the full unscented sampling procedure.
+We consider a state $\psi \in \mathbb{R}^{2n}$ with
+a deviant parameter $\lambda \in \mathbb{R}^{d}$ and dynamics
+$\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
+The nominal initial state is given by $\psi_{0}$ with an associated
+positive-definite covariance matrix $P_{0} \in \mathbb{R}_{++}^{2n \times 2n}$
 which describes the uncertainty in the initial state.
-$P_{0}$ is typically chosen to be non-zero even if the state preparation error is negligible.
+$P_{0}$ is typically non-zero even if the state preparation error is negligible.
 The deviant parameter
-is assumed to have zero mean and its distribution is given by the covariance matrix
+has zero-mean and its distribution is given by the covariance matrix
 $L_{k} \in \mathbb{R}_{++}^{d \times d}$ at knot point $k$. The zero mean assumption
 is convenient for deriving the update procedure. A non-zero mean can be encoded
-in the dynamics $f$.
+in the dynamics.
 
 The initial $4n + 2d$ sigma points and initial $4n + 2d$ deviant parameters are sampled
-from the initial distributions
+from the initial distributions,
 \begin{equation}\label{eq:uupdate}
@@
 column (mod $2n + d$) of the lower triangular Cholesky factor.
-Then, the sigma points are normalized
+Then, the sigma points are normalized,
 \begin{equation}\label{eq:unormalize}
@@
 \end{equation}
-The sigma points are propagated to the next knot point
+The sigma points are propagated to the next knot point,
 \begin{equation}\label{eq:upropagate}
@@
 \end{equation}
-The mean and covariance of the sigma points are computed
+The mean and covariance of the sigma points are computed,
 \begin{align}
@@
 choice of sigma points follows the prescription in equation 11 of \cite{julier2004unscented}.
-Other prescriptions may be chosen to reduce the number of required sigma points \cite{julier2002reduced}.
+Prescriptions that require fewer sigma points exist \cite{julier2002reduced}.

### 2020-11-30 paper/a3derivative.tex
updated figures, lots of proofreading

@@
 to outperform the 2\textsuperscript{nd}-order
-derivative method in the small, static detuning regime, see Figure \ref{fig:static}b.
+derivative method in the small, static detuning regime, see Figure \ref{fig:staticb}.
 The norms of the state derivatives for the 1\textsuperscript{st}- and 2\textsuperscript{nd}-order
-methods are provided in Table \ref{tab:dnorm} for the solutions at $t_{N} = 60$ns.
+methods are provided in Table \ref{tab:dnorm} for the solutions to
+the numerical experiment detailed in Section \ref{sec:static} at $t_{N} = 60$ns.
 The 2\textsuperscript{nd}-order method is able to decrease the 2\textsuperscript{nd}-order
@@
 contribution to the gate error than the 1\textsuperscript{st}-order derivative norm. A
-careful analysis could be completed for future problems to predict the efficacy of the
-derivative method at each order.
+careful analysis of contributions at each order could be used to predict the efficacy of the
+derivative method for future problems.
 
@@
   \begin{tabular}{c | c | c}
-    Method & ${\lvert \braket{\partial_{f_{q}} \psi_{N} | \partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
-    & ${\lvert \braket{\partial^{2}_{f_{q}} \psi_{N} | \partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{6}$)\\
+    Method & $\norm{\partial_{f_{q}} \psi_{N}}_{2}$ ($10^{3}$)
+    & $\norm*{\partial^{2}_{f_{q}} \psi_{N}}_{2}$ ($10^{6}$)\\
     \hline
-    D-1 & 0.436 & 57.817\\
-    D-2 & 1.702 & 9.030\\
+    D1 & 0.436 & 57.817\\
+    D2 & 1.702 & 9.030\\
   \end{tabular}
@@
     for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate's duration $t_{N} = 60$ns and are averaged over the four state derivatives.}
+    at the end of the gate $t_{N} = 60$ns
+    and are averaged over the four state derivatives.}
   \label{tab:dnorm}
@@
 The linear term is $L = H$ and the non-linear term is $N = (\partial_{\lambda} H) \ket{\psi}$.
-With zero-order hold on the controls the exact propagation is
+With zero-order hold on the controls the exact solution is:
 \begin{equation}
@@
   \begin{aligned}
-    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(\Delta t_{k} L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
-    &+ \int_{0}^{\Delta t_{k}} \exp((\Delta t_{k} - t^{'})L_{k}) N(t_{k} + t^{'}) dt^{'}\\
+    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(\Delta t_{k} L_{k})
+    \ket{\partial_{\lambda} \psi_{k}}\\
+    &+ \int_{0}^{\Delta t_{k}} \exp((\Delta t_{k} - t^{'})L_{k})
+    N(t_{k} + t^{'}) dt^{'}\\
   \end{aligned}
@@
 for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler
-method \cite{berland2005solving}
+method \cite{berland2005solving},
 \begin{equation}

### 2020-11-30 paper/a4time.tex
prog
updated figures, lots of proofreading

@@
-\section{Computational Performance and Scaling}
+\section{Computational Performance and Scaling \label{appendix:time}}
+We provide runtimes for our optimizations and comment on the scaling of the
+robustness methods. The runtimes for the base optimization in Section \ref{sec:fluxonium},
+longitudinally aware optimization in Section \ref{sec:longitude},
+and robust optimizations in Section \ref{sec:static} are presented
+in Table \ref{tab:time} for a $Z/2$ gate at multiples
+of the analytic gate time.
 \todo{timing table}
 
-\todo{make complexity analysis a coherent paragraph}
-(sampling) For the sampling method the number of sample states in the augmented state vector
-scales as $O(dn^{2})$
-because each of $n^{2}$ initial
-states is represented by $2d$ samples. $n$ is both the dimension of the Hilbert space
-and the size of a state vector. $d$ is the number of deviant parameters.
+\begin{table}[ht]
+  \begin{tabular} {c | c | c | c | c | c | c}
+    & \multicolumn{6}{c}{Wall Time ($\pm 0.001$ s)}\\
+    \hline
+    Gate Time (ns) & Base & Long. & S & SU & D1 & D2\\
+    \hline
+    $18$ & 0.155 & 1.688 & 1.773 & & 16.713 & 48.398\\ % 0.155049
+    $36$ & 7.014 & - & 48.213 & & 67.838 & 81.030\\
+    $72$ & 15.906 & - & & & &\\
+  \end{tabular}
+  \caption{Programs were executed on a single thread of an AMD Ryzen Threadripper 3970X 32-Core Processor.}
+  \label{tab:time}
+\end{table}
 
-(unscented) Each initial state from the operator basis
-is represented with a distribution of $4n + 2d$ samples. Hence,
-the number of states in the augmented state vector scales as $O(n^{3} + dn^{2})$.
-The gate error of each sample state is penalized as in the sampling method.
-
-(derivative) For this method $m$ state derivatives are associated with each
-initial state from the operator basis.
-So, the number of states in the augmented state vector scales as
-$O(dmn^{2})$. Additionally, we note that the number of initial
-states and samples we have employed for each of the three
-methods follows standard perscriptions. However, knowledge
-of the problem can typically be used to reduced the number
-of required samples.
+\todo{Ask Dave if we should keep complexity analysis in main text.}
+Now we present the problem size complexities for the robustness methods.
+For the sampling method, the size of the augmented state vector
+is $O(dn^{3})$. $d$ is the number of deviant parameters.
+$n$ is the dimension of the Hilbert space. There are $n^{2}$ initial states
+in the operator basis, $2d$ sample states per initial state,
+and each state has $2n$ elements.
+For the unscented sampling method, the size of the augmented state vector
+is $O(dn^{3} + n^{4})$.
+There are $n^{2}$ initial states in the operator basis,
+$2(2n + d)$ sample states per initial state,
+and each state has $2n$ elements.
+For the $m$\textsuperscript{th}-order derivative method, the size of the augmented state vector
+is $O(dmn^{3})$. There are $n^{2}$
+initial states in the operator basis, $dm$ state derivatives per initial state,
+and each state has $2n$ elements.

### 2020-11-30 paper/a4time.tex
updates

@@
-\section{Computational Performance and Scaling \label{appendix:time}}
+\section{Computational Performance \label{appendix:time}}
 We provide runtimes for our optimizations and comment on the scaling of the
-robustness methods. The runtimes for the base optimization in Section \ref{sec:fluxonium},
-longitudinally aware optimization in Section \ref{sec:longitude},
-and robust optimizations in Section \ref{sec:static} are presented
-in Table \ref{tab:time} for a $Z/2$ gate at multiples
-of the analytic gate time.
-\todo{timing table}
+robustness methods. The runtimes for
+the base optimization in Section \ref{sec:fluxonium},
+the longitudinally aware optimization in Section \ref{sec:longitude},
+and the robust optimizations in Section \ref{sec:static}
+are presented in Table \ref{tab:time}
+for a $Z/2$ gate at multiples of the analytic $Z/2$ gate time.
+Compiler optimizations for statically-sized arrays were utilized
+for all methods except for the unscented sampling method.
+The unscented sampling method's augmented state vector size
+was too large to take advantage of these optimizations,
+adversely affecting its run time.
+For the robust methods, the runtime scales super-linearly with the
+gate time. We expect the runtime to scale linearly with the gate time
+for large gate times because
+the number of knot points scales linearly with the gate time.
+We observe this behavior for the base method, which has a smaller
+augmented state vector size than the robust methods.
+We performed optimizations on a single CPU thread in this work as a proof
+of concept. Future work will parallelize the robustness
+methods across the initial states in the operator basis and utilize GPUs
+\cite{leung2017speedup},
+allowing for fast optimization on large Hilbert spaces.
 
-\begin{table}[ht]
+\begin{table}[ht!]
   \begin{tabular} {c | c | c | c | c | c | c}
@@
     \hline
-    Gate Time (ns) & Base & Long. & S & SU & D1 & D2\\
+    $t_{N}$ (ns) & Base & Long. & S & SU & D1 & D2\\
     \hline
@@
     $36$ & 7.014 & - & 48.213 & & 67.838 & 81.030\\
-    $72$ & 15.906 & - & & & &\\
+    $72$ & 15.906 & - & 281.372 & & 266.997 & 332.182\\
   \end{tabular}
-  \caption{Programs were executed on a single thread of an AMD Ryzen Threadripper 3970X 32-Core Processor.}
+  \caption{
+    \todo{finish timing table}
+    Runtimes for $Z/2$ optimizations.
+    Programs were executed on a single thread of an
+    AMD Ryzen Threadripper 3970X 32-Core Processor.
+  }
   \label{tab:time}

### 2020-11-30 src/spin/spin23.jl
update to spin23.jl

@@ -49,7 +49,7 @@ const S8_IDX = SVector{HDIM_ISO}(HDIM_ISO * 7 + 1:HDIM_ISO * 8)
 const S9_IDX = SVector{HDIM_ISO}(HDIM_ISO * 8 + 1:HDIM_ISO * 9)
 const S10_IDX = SVector{HDIM_ISO}(HDIM_ISO * 9 + 1:HDIM_ISO * 10)
 # misc
-const SAMPLE_STATE_COUNT = 1
+const SAMPLE_STATE_COUNT = 4
 const SAMPLES_PER_STATE = 10
 const SAMPLES_PER_STATE_INV = 1//10
 const SAMPLE_COUNT = SAMPLE_STATE_COUNT * SAMPLES_PER_STATE
@@ -77,7 +77,8 @@ Model = Data.Model
         + sample_index * HDIM_ISO
     )
 )
-const SAMPLE_INDICES = [astate_sample_inds(i, j) for i = 1:SAMPLE_STATE_COUNT for j = 1:SAMPLES_PER_STATE]
+const SAMPLE_INDICES = [astate_sample_inds(i, j)
+                        for i = 1:SAMPLE_STATE_COUNT for j = 1:SAMPLES_PER_STATE]
 
 function unscented_transform(model::Model, astate::AbstractVector,
                              negi_hc::AbstractMatrix, dt::Real, i::Int)
@@ -92,20 +93,11 @@ function unscented_transform(model::Model, astate::AbstractVector,
     s8 = astate[offset + S8_IDX]
     s9 = astate[offset + S9_IDX]
     s10 = astate[offset + S10_IDX]
-    # println("s1d")
-    # show_nice(s1)
-    # println("\ns10d")
-    # show_nice(s10)
-    # println("")
-    # @bp
     # compute state mean
     sm = SAMPLES_PER_STATE_INV .* (
         s1 + s2 + s3 + s4 + s5
         + s6 + s7 + s8 + s9 + s10
     )
-    # println("sm:")
-    # show_nice(sm)
-    # println("")
     # compute state covariance
     d1 = s1 - sm
     d2 = s2 - sm
@@ -122,17 +114,11 @@ function unscented_transform(model::Model, astate::AbstractVector,
         + d5 * d5' + d6 * d6' + d7 * d7' + d8 * d8'
         + d9 * d9' + d10 * d10'
     )
-    # println("s_cov:")
-    # show_nice(s_cov)
-    # println("")
[…]

### 2020-12-04 paper/5static.tex
updates

@@
 each of which evolves with a distinct deviant parameter value.
-We append four sets of two sample states $\psi_{\pm}$
+We append four sets of two sample states $\psi^{\pm}$
 to the augmented state vector \eqref{eq:astatecontrols}.
@@
 for each sample is the image of its initial state
-under the desired gate. We penalize deviations
-from the final state in the objective \eqref{eq:costfun}.
+under the desired gate. The infidelity between
+the samples and their final states are penalized
+by adding a term $\sum_{k=1}^{N} q_{k} (1 - \braket{\psi^{\pm}_{N}}{\psi_{f}})$
+to the objective \eqref{eq:costfun}.
 
@@
 of the distribution.
-As in the sampling method,
-we penalize deviations from the final state according to \eqref{eq:costfun}.
+As for the sampling method,
+we penalize the infidelity between the samples and their final states
+by adding an infidelity contribution to \eqref{eq:costfun}.
 A detailed procedure for the unscented transformation is given
@@
 term in \eqref{eq:d1dyn} are used to efficiently integrate the coupled dynamics,
-see Appendix \ref{appendix:derivative}. For timing
-results and complexity analyses for the three methods,
+see Appendix \ref{appendix:derivative}. For runtimes
+and complexity analyses for the three methods,
 consult Appendix \ref{appendix:time}.

### 2020-12-04 paper/a4time.tex
updates

@@
     \hline
-    $18$ & 0.155 & 1.688 & 1.773 & & 16.713 & 48.398\\ % 0.155049
-    $36$ & 7.014 & - & 48.213 & & 67.838 & 81.030\\
-    $72$ & 15.906 & - & 281.372 & & 266.997 & 332.182\\
+    $18$ & 0.155 & 1.688 & 1.773 & 210.573 & 16.713 & 48.398\\ % 0.155049
+    $36$ & 7.014 & - & 48.213 & 4566.236 & 67.838 & 81.030\\
+    $72$ & 15.906 & - & 281.372 & 16575.308 & 266.997 & 332.182\\
   \end{tabular}
   \caption{
-    \todo{finish timing table}
     Runtimes for $Z/2$ optimizations.
@@
 in the operator basis, $2d$ sample states per initial state,
-and each state has $2n$ elements.
+and each state has $2n$ real numbers.
 For the unscented sampling method, the size of the augmented state vector
@@
 $2(2n + d)$ sample states per initial state,
-and each state has $2n$ elements.
+and each state has $2n$ real numbers.
 For the $m$\textsuperscript{th}-order derivative method, the size of the augmented state vector
@@
 initial states in the operator basis, $dm$ state derivatives per initial state,
-and each state has $2n$ elements.
+and each state has $2n$ real numbers.

### 2020-12-04 paper/1introduction.tex
dave checks

@@
 \section{Introduction}
-\todo{Ask Dave to review paragraph}
 Quantum optimal control (QOC) techniques are a class of optimization

### 2020-12-04 paper/2background.tex
dave checks

@@
 
-\todo{Ask Dave to review paragraph}
+\todo{Ask Brian to review paragraph. Are there other advantages to using ALTRO
+  for quantum control that you would like to list in this paragraph? These are the reasons
+  why we think it is better than what we were using previously.}
 As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
@@
 on both the state and the controls and does not restrict
-optimization to the constraint manifold. First-order solvers like GRAPE
-rely on unconstrained optimization to satisfy desiderata on the state,
+optimization to the constraint manifold. First-order solvers such as GRAPE
+use unconstrained optimization to satisfy desiderata on the state,
 such as achieving a desired gate fidelity, forbidding occupations of particular quantum states,

### 2020-12-04 paper/4longitude.tex
dave checks

@@
   \caption{
-    \todo{Ask Dave to okay figure color scheme}
     (a) Numerically optimized gates (dark) and analytically optimized gates (light).
@@
 relaxation by performing optimal control on $n^{2}$
-density matrices of size $n \times n$ that evolve under master
-equation dynamics, rather than $n$ states vectors of size $n$
+density matrices of size $n \times n$ that evolve under a master
+equation, rather than $n$ state vectors of size $n$
 under the TDSE \eqref{eq:tdse}.
@@
 
-\todo{Ask Brian to review this paragraph and next}
+\todo{Ask Brian to review this paragraph and next. Is the
+optimization setup clear?}
 The longitudinal relaxation probability is given by,

### 2020-12-04 paper/5static.tex
dave checks

@@
 
-\todo{Ask Brian to review this paragraph and next two}
+\todo{Ask Brian to review this paragraph and next two.
+Are the optimization setups clear?}
 The sampling method simultaneously optimizes over copies of a state,
@@
 the samples and their final states are penalized
-by adding a term $\sum_{k=1}^{N} q_{k} (1 - \braket{\psi^{\pm}_{N}}{\psi_{f}})$
+by adding a term
+$\sum_{k=1}^{N} q_{k} (1 - {\lvert \braket{\psi^{\pm}_{k}}{\psi_{f}} \rvert}^{2})$
 to the objective \eqref{eq:costfun}.
@@
 of the distribution.
-As for the sampling method,
-we penalize the infidelity between the samples and their final states
-by adding an infidelity contribution to \eqref{eq:costfun}.
+We penalize the infidelity between the samples and their final states
+by adding an infidelity contribution to \eqref{eq:costfun}
+as in the sampling method.
 A detailed procedure for the unscented transformation is given
@@
 see Appendix \ref{appendix:derivative}. For runtimes
-and complexity analyses for the three methods,
+and complexity analyses for the three robust control methods,
 consult Appendix \ref{appendix:time}.

### 2020-12-04 paper/6stochastic.tex
dave checks

@@
 
-We task the analytic gate and those produced by
-the robust control methods of the previous section
+We challenge the analytic method and
+the robust control
+methods of the previous section
 to realize an $X/2$ gate subject to 1/$f$ flux noise.
 We compute gate errors for each method
-by evolving system under the Hamiltonian in \eqref{eq:hamiltonian}
+by evolving the system under the Hamiltonian in \eqref{eq:hamiltonian}
 where the optimized flux amplitude is modified $a(t) \gets a(t) + \delta a(t)$.
@@
 In principle, we could modify the sampling method
-similarly but we choose to sample statically at $\sigma_{a}$ for comparison.
+similarly; however, we choose to sample statically at $\sigma_{a}$ for comparison.
 The derivative methods require no algorithmic modification from the static case,
@@
 and compute the cumulative gate error
-after each application, see Figure \ref{fig:stochastic}. Both the analytic
+after each application, see Figure \ref{fig:stochastic}.
+Both the analytic
 and numerical methods achieve single gate errors
@@
 1/$f$ flux noise is a significiant source of coherent errors in NISQ applications and
-these numerical techniques offer effective avenues to mitigate them.
+these numerical techniques offer effective avenues to mitigate it.

### 2020-12-08 paper/2background.tex
Brian edits

@@
 where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^l$ and
-$u_k \in \R^m$ are the augmented state and input augmented control variables,
+$u_k \in \R^m$ are the augmented state and control variables,
 $f(x_k,u_k)$ is the discrete dynamics function, and $g_k(x_k,u_k)$ and
@@
 for imposing arbitrary, simultaneous constraints
-on both the state and the controls and does not restrict
-optimization to the constraint manifold. First-order solvers such as GRAPE
+on both the state and the controls that converge superlinearly to tight tolerances. 
+While the dynamics remain anytime dynamically feasible and can be integrated with high 
+accuracy, the arbitrary constraints on states and controls are not restricted to the 
+constraint manifold during the solve.
+First-order solvers such as GRAPE
 use unconstrained optimization to satisfy desiderata on the state,

### 2020-12-08 paper/3fluxonium.tex
Brian edits

@@
 \cite{battistel2019fast, krantz2019quantum, zhang2020universal}. 
+Since these terms are all included in the state \eqref{eq:astatecontrols},
+they are enforced via simple initial and final conditions.
 We require $-0.5 \textrm{GHz} \le a_{k} \le 0.5 \textrm{GHz}$
@@
 takes the form
-$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
+$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$ \todo{I'm not sure what you mean by this, 
+this looks like the cost term? Contraints should be of the form $c(x) = 0$ or $c(x) \leq 0$}.
 The inequality constraints on $a$ are handled
@@
 $c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
-and $0$ otherwise.
+and $0$ otherwise \todo{This isn't a bound constraint, this is a piece-wise quadratic 
+constraint? Do you mean $c(x) = [a \leq b \l a \geq -b]$?}.
 In addition, the norm of the states
 are constrained to $1$ to ensure discretization error is not
-exploited $c_{k} = x_{k}^{T}x_{k} - 1$. The selection
-of subsets from $x_{k}$ is implied in
-the constraint equations.
+exploited $c_{k} = x_{k}^{T}x_{k} - 1$ \todo{Is this right? Why are you
+normalizing the entire state vector? Or are you just normalizing the qubit
+states?}. 
+The selection of subsets from $x_{k}$ is implied in
+the constraint equations. \todo{I think it would be cleaner if you just put the actual 
+constraints into \ref{eq:costfun} and then provide a brief explanation here.}

### 2020-12-08 paper/4longitude.tex
Brian edits

@@
 than the number of knot points $N$, preserves the
-Markovian decision structure of the trajectory
+Markovianity of the trajectory
 optimization problem. The square root of the time step $\sqrt{\Delta t_{k}}$
@@
 integration accuracy is maintained we add a bound
-constraint on the time step at each knot point.
+constraint on the time step at each knot point. \todo{(Brian) the setup seems clear to me.}
 

### 2020-12-08 paper/a4time.tex
dave checks
Brian edits

@@
 allowing for fast optimization on large Hilbert spaces.
+\vspace*{\intextsep}
 
-\begin{table}[ht!]
+\begin{table}[H]
   \begin{tabular} {c | c | c | c | c | c | c}
@@
 
-\todo{Ask Dave if we should keep complexity analysis in main text.}
 Now we present the problem size complexities for the robustness methods.
 For the sampling method, the size of the augmented state vector
-is $O(dn^{3})$. $d$ is the number of deviant parameters.
+is $O(dn^{3})$, where $d$ is the number of deviant parameters and
 $n$ is the dimension of the Hilbert space. There are $n^{2}$ initial states

### 2020-12-10 paper/1introduction.tex
added constraints to opt statement

@@
 We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
-We also show that QOC can solve important problems for fluxonium-based qubits,
-in particular, taking advantage of the $T_{1}$-dependence of the controls
-to mitigate longitudinal relaxation type decoherence and
-performing phase gates in arbitrary times.
+We also show that QOC can solve important problems for fluxonium-based qubits: mitigating
+longitudinal relaxation type decoherence
+by taking advantage of the $T_{1}$-dependence of the controls
+and performing phase gates in arbitrary times.
 To mitigate errors due to longitudinal

### 2020-12-10 paper/2background.tex
added constraints to opt statement

--- 2020-11-30 paper/2background.tex
@@
 Trajectory optimization problems are typically of the following form: 
-\begin{mini}[2]
-    {x_{1:N},u_{1:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell(x_k,u_k) }{}{}
-    \addConstraint{x_{k+1} = f(x_k,u_k)}
-    \addConstraint{g_k(x_k,u_k)}{\leq 0}
-    \addConstraint{h_k(x_k,u_k)}{=0}
-    \label{opt:discrete_trajopt},
-\end{mini}
+\begin{mini!}[2]
+    {x_{1:N},u_{1:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell(x_k,u_k)}{}{} \label{eq:gcostfun}
+    \addConstraint{x_{k+1} = f(x_k,u_k)} \label{eq:gdyncon}
+    \addConstraint{g_k(x_k,u_k)}{\leq 0} \label{eq:ineqcon}
+    \addConstraint{h_k(x_k,u_k)}{=0} \label{eq:eqcon}
+\end{mini!}
 where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^l$ and
-$u_k \in \R^m$ are the augmented state and input augmented control variables,
+$u_k \in \R^m$ are the state and control variables,
 $f(x_k,u_k)$ is the discrete dynamics function, and $g_k(x_k,u_k)$ and
@@
 
-Many techniques have been proposed for solving \eqref{opt:discrete_trajopt}. Standard 
+Many techniques have been proposed for solving \eqref{eq:gcostfun}--\eqref{eq:eqcon}. Standard 
 methods include direction collocation \cite{Hargraves87} and differential-dynamic programming
@@
 function, $\lambda \in \R^p$ is a Lagrange multiplier estimate, and $I_\mu$ is a diagonal matrix
-of penalty weights, $\mu$, whose magnitudes depend on whether the corresponding constraint is active or inactive.
-For ALTRO, $f(z)$ is the objective function of \eqref{opt:discrete_trajopt}, $c(z)$ is the 
+of penalty weights, $\mu$, whose magnitudes depend on whether
+the corresponding constraint is active or inactive.
+For ALTRO, $f(z)$ is \eqref{eq:gcostfun}, $c(z)$ is the 
 concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
@@
 \begin{equation}
-	\lambda \gets \lambda + \mu c(z) ,
+	\lambda \gets \lambda + \mu c(z)
 \end{equation}
@@
 
-\todo{Ask Dave to review paragraph}
 As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
 for imposing arbitrary, simultaneous constraints
-on both the state and the controls and does not restrict
-optimization to the constraint manifold. First-order solvers like GRAPE
-rely on unconstrained optimization to satisfy desiderata on the state,
+on both the states and the controls that converge superlinearly to tight tolerances. 
+First-order solvers such as GRAPE
+use unconstrained optimization to satisfy desiderata on the states,
 such as achieving a desired gate fidelity, forbidding occupations of particular quantum states,
@@
 To impose constraints on the controls, such as amplitude restrictions, first-order solvers
-employ projected gradient descent methods which restrict optimization to the constraint manifold and
+employ projected gradient descent methods which restrict
+optimization to the constraint manifold and
 may hinder convergence to the optimal solution \cite{clarkson2010coresets,
   hauswirth2016projected}.
-The ALTRO solver delays projecting onto the constraint manifold until the ALM
+Conversely,
+the ALTRO solver delays projecting onto the constraint manifold defined by
+\eqref{eq:ineqcon} and \eqref{eq:eqcon} until the ALM
 solve is already at a coarse tolerance, achieving machine-precision satisfaction
-for constraints on the controls and the state. These advantages are critical
+for constraints on the controls and the states. These advantages are critical
 for this work, where multiple constraints must be satisfied in addition

### 2020-12-10 paper/3fluxonium.tex
added constraints to opt statement

--- 2020-11-30 paper/3fluxonium.tex
@@
   \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
-    \addConstraint{(x_1 \; \textrm{given})}
-    \addConstraint{g_k(x_k,u_k)}{\leq 0}
-    \addConstraint{h_k(x_k,u_k)}{=0}
+  \addConstraint{\ket{\psi^{0}_{1}} = \ket{0}, \ket{\psi^{1}_{1}} = \ket{1}} \label{eq:istate_con}
+  \addConstraint{\ket{\psi^{i}_{N}} = \ket{\psi^{i}_{f}}} \label{eq:tstate_con}
+  \addConstraint{{\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}}
+      \rvert}^{2} = 1} \label{eq:statenorm_con}
+  \addConstraint{{\textstyle \int^{t_{1}}_{0} a \ \mathrm{d}t} = \pdv*{a_{1}}{t} = 0}
+  \label{eq:icintd_con}
+  \addConstraint{{\textstyle \int^{t_{N}}_{0}} a \ \mathrm{d}t= 0} \label{eq:znf_con}
+  \addConstraint{a_{1} = a_{N} = 0} \label{eq:concat_con}
+  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz}} \label{eq:amp_con}
 \end{mini!}
@@
 \begin{equation}
-  x = \begin{bmatrix} \psi^{1} \\ \psi^{2} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
+  x = \begin{bmatrix} \psi^{0} \\ \psi^{1} \\ \int^{t}_{0} a \ \mathrm{d}t\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
   u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
@@
 \end{equation}
-The initial states $\psi^{1}_{1}$, $\psi^{2}_{1}$
-are basis vectors of the Hilbert space $\ket{0}, \ket{1}$.
-The final states $\psi^{1}_{f}$, $\psi^{2}_{f}$ are the image
-of the initial states under the desired gate.
-The ALTRO implementation we use does not currently support complex numbers so
-we compute in the isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-given in \cite{leung2017speedup},
-\begin{equation}
-  H \psi \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
-  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
-  \label{eq:isomorphism}
-\end{equation}
-The discrete dynamics function \eqref{eq:dyn_con} integrates the states
-using \eqref{eq:tdse} and \eqref{eq:hamiltonian}
-and integrates the moments of the flux amplitude.
-Exposing lower order moments of the flux amplitude allows us to penalize their norms
-in \eqref{eq:costfun}, smoothing the flux amplitude
-and mitigating AWG ringing due to high frequency transitions.
 The matrices $Q_{k}$ and $R_{k}$ define the penalty metric.
 Including objectives at each knot point smoothens the optimization landscape, though
-the importance of the final augmented state objective is encoded in the relative
+the importance of the final objective is encoded in the relative
 weight $Q_{N} \sim N \cdot Q_{k}$.
@@
 reflect the physical limitations of the apparatus and 
-constraints that improve the experimental realization of the control pulse.
-To ensure gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions,
-we require $a_{1} = a_{N} = 0$.
-Furthermore, we require $\int a_{N} \ dt = 0$. This
-constraint ensures the pulse has zero net flux, mitigating
-the hysteresis ubiquitous in flux bias lines,
+improve the experimental realization of the control pulse.
+The discrete dynamics \eqref{eq:dyn_con} integrates the states
+according to the TDSE \eqref{eq:tdse} and the
+fluxonium Hamiltonian \eqref{eq:hamiltonian}
+and integrates the moments of the flux amplitude.
+Exposing lower order moments of the flux amplitude allows us to penalize their norms
+in \eqref{eq:costfun}, smoothing the flux amplitude
+and mitigating AWG ringing due to high frequency transitions.
+We ensure the desired gate is achieved by imposing the constraints
+\eqref{eq:istate_con} and \eqref{eq:tstate_con}.
+The initial states in \eqref{eq:istate_con}
+are chosen to span the Hilbert space.
+The final states in \eqref{eq:tstate_con} are the image
+of the initial states under the desired gate.
+The ALTRO implementation we use does not currently support complex numbers so
+we compute in the isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
+given in \cite{leung2017speedup},
+\begin{equation}
+  H \ket{\psi} \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
+    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
+  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
+  \label{eq:isomorphism}
+\end{equation}
+We ensure the solver does not take advantage of discretization
+error by imposing the normalization constraint \eqref{eq:statenorm_con}.
+The constraint \eqref{eq:icintd_con} is an initial condition.
+We ensure the pulse has zero net flux by imposing \eqref{eq:znf_con},
+which mitigates the hysteresis ubiquitous in flux bias lines
 \cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
-We require $-0.5 \textrm{GHz} \le a_{k} \le 0.5 \textrm{GHz}$
-to ensure the two-level approximation \eqref{eq:hamiltonian}
-remains valid. Additionally, we require
-$\psi_{N} = \psi_{f}$.
-The zero net flux and final state constraints
-are satisfied if the final augmented state is
-reached $x_{N} = x_{f}$, the constraint function
-takes the form
-$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
-The inequality constraints on $a$ are handled
-with a bound constraint which takes the form
-$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
-and $0$ otherwise.
-In addition, the norm of the states
-are constrained to $1$ to ensure discretization error is not
-exploited $c_{k} = x_{k}^{T}x_{k} - 1$. The selection
-of subsets from $x_{k}$ is implied in
-the constraint equations.
+We ensure gates may be concatenated arbitrarily without
+inducing AWG ringing due to high-frequency transitions by imposing
+the boundary condition \eqref{eq:concat_con}.
+Finally, we ensure the two-level
+approximation \eqref{eq:hamiltonian} remains valid
+by imposing the amplitude constraint \eqref{eq:amp_con}.
+All of the gates presented in this work achieve a maximum
+individual constraint violation of $10^{-8}$
+for \eqref{eq:dyn_con}--\eqref{eq:amp_con}.

### 2020-12-10 paper/5static.tex
Brian edits
added constraints to opt statement

@@
 
-\todo{Ask Brian to review this paragraph and next two.
-Are the optimization setups clear?}
 The sampling method simultaneously optimizes over copies of a state,

### 2020-12-10 paper/a3derivative.tex
added constraints to opt statement

@@
 General exponential integrators break the dynamics into a linear term and a non-linear term.
-For example, consider
-integrating the dynamics of the first state derivative $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
-H \ket{\partial_{\lambda} \psi} + (\partial_{\lambda} H) \ket{\psi}$ in units of $i\hbar = 1$.
+For example, the dynamics of the first state derivative in
+units of $i\hbar = 1$ are $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
+H \ket{\partial_{\lambda} \psi} + (\partial_{\lambda} H) \ket{\psi}$.
 The linear term is $L = H$ and the non-linear term is $N = (\partial_{\lambda} H) \ket{\psi}$.

### 2020-12-10 paper/a4time.tex
added constraints to opt statement

@@
 was too large to take advantage of these optimizations,
-adversely affecting its run time.
-For the robust methods, the runtime scales super-linearly with the
-gate time. We expect the runtime to scale linearly with the gate time
-for large gate times because
-the number of knot points scales linearly with the gate time.
-We observe this behavior for the base method, which has a smaller
+which adversely affected its run time.
+The runtimes of the robust methods scale
+super-linearly with the gate time.
+For large gate times, we expect the runtimes
+to scale linearly with the gate time
+because the number of knot points scales linearly with the gate time.
+We observe this trend for the base method, which has a smaller
 augmented state vector size than the robust methods.
@@
 \cite{leung2017speedup},
-allowing for fast optimization on large Hilbert spaces.
+which will enable fast optimizations on large Hilbert spaces.
 \vspace*{\intextsep}
-
 \begin{table}[H]

### 2020-12-10 paper/3fluxonium.tex
added todo in constraint section

@@
 
+\todo{Ask Brian to review this paragraph and the constraints \eqref{eq:dyn_con}-\eqref{eq:amp_con}.
+  Is it okay if I have multiple constraints on the same line? The constraints that are on the same
+  line are imposed for similar reasons, which makes the list shorter and it easier to reference them here.}
 We impose constraints that

### 2021-01-12 paper/1introduction.tex
majority of jens and colm comments adressed

@@
 \section{Introduction}
-Quantum optimal control (QOC) techniques are a class of optimization
+Quantum optimal control (QOC) is a class of optimization
 algorithms for accurately and efficiently manipulating quantum systems.
@@
   nobauer2015smooth, poggiali2018optimal, rembold2020introduction, tian2019optimal},
-and Bose-Einstien condensates \cite{amri2019optimal, sorensen2018quantum}.
+and Bose-Einstein condensates \cite{amri2019optimal, sorensen2018quantum}.
 For quantum computation,
-optimal control techniques are employed to achieve high fidelity gates
+optimal control is employed to achieve high-fidelity gates
 while adhering to experimental constraints.
@@
 to poor experimental performance.
-Robust control techniques improve upon
-standard optimal control techniques by encoding
-model uncertainties
+Robust control improves upon
+standard optimal control by encoding
+model parameter uncertainties
 in optimization objectives, yielding performance
-guarantees over a range of parameters \cite{Zhou97,Morimoto00,Manchester18}.
+guarantees over a range of parameter values \cite{Zhou97,Morimoto00,Manchester18}.
 We adapt robust control techniques from the robotics community to mitigate
-parameter deviation errors for
+parameter uncertainty errors for
 a superconducting fluxonium qubit.
 
-QOC techniques have had tremendous success in
-engineering high fidelity quantum gates \cite{
+QOC has had tremendous success in
+engineering high-fidelity quantum gates \cite{
   chou2015optimal, dolde2014high, egger2013optimized, egger2014adaptive,
@@
   spiteri2018quantum, sporl2007optimal}. 
-While standard QOC techniques can predict system behavior
-with high accuracy, they are sensitive to experimental
-errors such as parameter drift, noise, finite control 
-resolution, and decoherence. Multiple techniques have been developed to address
-these shortcomings. Analytic techniques to mitigate parameter deviation errors
-include composite pulses \cite{cummins2000use, cummins2003tackling,
+However, the system often
+deviates from the model used in optimization
+due to experimental errors such as parameter drift, noise, finite control 
+resolution, and decoherence. Multiple techniques have been
+developed to address this shortcoming.
+Analytically-derived control pulses that
+mitigate parameter uncertainty errors include
+composite pulses \cite{cummins2000use, cummins2003tackling,
   kupce1995stretched, merrill2014progress},
-dynamic and geometric phase considerations
-\cite{han2020experimental, xu2020nonadiabatic},
-and the DRAG scheme \cite{motzoi2009simple}.
-To mitigate decoherence, Floquet techniques have been employed
+pulses designed by considering dynamic and geometric phases
+\cite{han2020experimental, xu2020nonadiabatic}, and
+pulses obtained with the DRAG scheme \cite{motzoi2009simple}.
+To suppress decoherence, Floquet techniques have been employed
 \cite{huang2020engineering, mundada2020floquet}.
-Numerical techniques to mitigate parameter deviation errors
-include closed-loop methods \cite{egger2014adaptive, feng2018gradient, huang2020engineering,
+Numerical schemes to mitigate parameter uncertainty errors
+include closed-loop methods \cite{egger2014adaptive, feng2018gradient,
   wittler2020integrated} and open-loop methods \cite{
@@
   rembold2020introduction, kosut2013robust}.
-Numerical techniques to mitigate decoherence include
-modeling master equations \cite{rembold2020introduction} and employing
-Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
+To model decoherence, numerical techniques typically employ
+master equations \cite{rembold2020introduction, schulteherbruggen2011optimal} or
+Monte Carlo-style quantum trajectories \cite{abdelhafez2019gradient}.
 
 In this work, we study three robust control techniques that
-make the system's state trajectory less sensitive
-to static and time-dependent parameter deviations:
+make the system's quantum state trajectory less sensitive
+to static and time-dependent parameter uncertainty:
 \begin{enumerate}
   \item A sampling method, similar to the work of \cite{allen2019robust,
-    carvalho2020error, reinhold2019controlling, rembold2020introduction}.
-  \item An unscented sampling method adapted from the unscented transform used in the 
-    state estimation community \cite{howell2020direct, julier2004unscented,
+    carvalho2020error, khaneja2005optimal, reinhold2019controlling, rembold2020introduction}.
+  \item An unscented sampling method adapted from the unscented transform used in
+    state estimation \cite{howell2020direct, julier2004unscented,
       lee2013sigma, manchester2016derivative}.
-  \item A derivative-penalization method, which uses efficient mixed-mode differentiation
-    to compute the sensitivity of the quantum state trajectory
-    to deviant parameters.
+  \item A derivative method, which penalizes the sensitivity of the quantum state trajectory
+    to uncertain parameters.
 \end{enumerate}
 We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
-We also show that QOC can solve important problems for fluxonium-based qubits: mitigating
-longitudinal relaxation type decoherence
-by taking advantage of the $T_{1}$-dependence of the controls
-and performing phase gates in arbitrary times.
-To mitigate errors due to longitudinal
-relaxation, we perform time-optimal control and
-utilize an efficient optimization objective that does
-not pay the increased computational cost of integrating a master equation.
+We also show that QOC can solve important problems associated with
+fluxonium-based qubits: mitigating
+depolarization by taking advantage of the $T_{1}$-dependence of the controls
+and synchronizing qubits with distinct frequencies
+by performing phase gates in arbitrary times.
+To mitigate depolarization,
+we perform time-optimal control and
+employ an efficient depolarization model that 
+is significantly less expensive to compute than integrating a master equation.
 Leveraging recent advances in trajectory optimization within the field of robotics, we
-solve these optimization problems using the ALTRO solver, which uses iterative LQR 
-(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
-methods such as GOAT \cite{machnes2015gradient}, GRAPE
-\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
-an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
-each time step \cite{howell2019altro}.
+solve these optimization problems using ALTRO (Augmented Lagrangian TRajectory Optimizer)
+\cite{howell2019altro}, which uses the iterative linear quadratic regulator (iLQR) method
+\cite{Li2004a} within an augmented Lagrangian framework \cite{lantoine2012hybrid,
+  plancher2017constrained}
+to handle nonlinear equality and inequality constraints at 
+each time step.
 
 This paper is organized as follows.
-First, we introduce the ALTRO method in the context of QOC
+First, we reintroduce ALTRO in the context of QOC
 in Section \ref{sec:background}.
-We describe realistic constraints for the fluxonium and
-map them to the ALTRO method in Section \ref{sec:fluxonium}. Then, we
-outline a method for making the optimization aware of longitudinal
-relaxation in Section \ref{sec:longitude}. Next, we outline three methods for achieving
-robustness to static parameter deviations in Section \ref{sec:static}. Finally,
-we employ the robust control techniques to mitigate 1/$f$ flux noise
+We describe realistic constraints for operating the fluxonium and
+define the associated QOC problem in Section \ref{sec:fluxonium}.
+Then, we outline a method for making the optimization aware of depolarization
+in Section \ref{sec:longitude}. Next, we outline three techniques for achieving
+robustness to static parameter uncertainties in Section \ref{sec:static}. We
+adapt the same techniques to mitigate 1/$f$ flux noise
 in Section \ref{sec:stochastic}.

### 2021-01-12 paper/3fluxonium.tex
small constraint updates
more minor constraint updates
majority of jens and colm comments adressed

--- 2020-12-10 paper/3fluxonium.tex
@@
-\section{QOC on the Fluxonium \label{sec:fluxonium}}
-In the following, we study
-the QOC problem on the fluxonium qubit.
-The fluxonium qubit is a promising building block for superconducting
-circuits, and the accurate
+\section{QOC for the Fluxonium \label{sec:fluxonium}}
+In the following, we optimize quantum gates
+(unitary transformations) for the fluxonium qubit.
+The fluxonium is a promising building block
+for quantum computers, and the accurate
 two-level approximation of its Hamiltonian makes
@@
 In the two-level
-approximation the Hamiltonian takes the form:
+approximation, the Hamiltonian takes the form:
 \begin{align}
@@
 \end{align}
-where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
-$a(t)$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
-are Pauli matrices. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
-gates for the fluxonium qubit subject to experimental constraints, decoherence, and
-Hamiltonian parameter deviations. We compare the gates we obtain with numerical
-methods to the analytically constructed gates reported in
-\cite{zhang2020universal} for the same device.
+where $f_{q}$ is the qubit frequency at the flux frustration point,
+$a(t)$ is the flux-drive amplitude, $h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
+are Pauli matrices. We optimize $X/2$, $Y/2$, and
+$Z/2$ gates for the fluxonium presented in \cite{zhang2020universal},
+and compare them to the analytically constructed gates for
+that device.
 
-The optimization problem takes the form:
+First, we outline the constraints for the fluxonium gate problem.
+We formulate this problem as a multi-state-transfer problem.
+The initial conditions on
+the states are $\ket{\psi^{0}_{1}} = \ket{0}$, $\ket{\psi^{1}_{1}} = \ket{1}$
+\eqref{eq:istate_con}
+where the superscript is an index $i \in \{0, 1\}$,
+and the subscript indicates the first knot point $k = 1$.
+The states at the final knot point are constrained to be
+the image of the initial states under the desired gate $U$,
+$\ket{\psi^{i}_{N}} = U \ket{\psi^{i}_{1}} \ \forall \ i$
+\eqref{eq:tstate_con}.
+%% This constraint is sensitive to global phase
+%% differences between the states.
+%% , but it is computationally efficient
+%% because the Jacobian and Hessian
+%% of the corresponding constraint function are diagonal.
+%% We could have analogously chosen a metric which is insensitive
+%% to global phases, such as the fidelity of the states.
+Furthermore, we impose the normalization constraint
+${\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}} \rvert}^{2} = 1 \ \forall \ i,k$
+\eqref{eq:statenorm_con}
+to ensure the solver does not take advantge of discretization errors in numerical integration.
+To refer to the discrete moments the flux amplitude, we introduce the notation
+$\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
+$a(t) \lvert_{t = t_{k}} \equiv a_{k}$,
+$\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$.
+We impose the zero net flux constraint $\int_{t} a_{N} = 0$
+\eqref{eq:znf_con}
+which mitigates the hysteresis ubiquitous in flux-bias lines
+\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
+The flux amplitude is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
+\eqref{eq:amp_con}
+to ensure the two-level approximation \eqref{eq:hamiltonian} remains valid.
+We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con}
+so the gates may be concatenated arbitrarily. Additionally,
+we have the initial condition $\int_{t} a_{1} = \partial_{t} a_{1} = 0$
+\eqref{eq:ic_con}.
+
+%% F1
+\begin{figure*}[ht]
+
+  \begin{subfigure}{.315\textwidth}
+    \includegraphics[width=\linewidth]{assets/f1a.png}
+    \caption{\label{fig:longitudea}}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.23\textwidth}
+    \includegraphics[width=\linewidth]{assets/f1b.png}
+    \caption{\label{fig:longitudeb}}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.4\textwidth}
+    \includegraphics[width=\linewidth]{assets/f1c.png}
+    \caption{\label{fig:longitudec}}
+  \end{subfigure}
+  \caption{
+    (a) Numerically optimized gates (dark blue)
+    and analytically optimized gates (light pink).
+    (b) $T_{1}$ interpolation function used in optimization. Markers
+    denote the time-averaged, absolute amplitude of each gate.
+    (c) Cumulative gate errors due to depolarization as a function of the
+    number of gates applied.
+    Gate errors for the numerically optimized $Z/2$ and $Y/2$ gates
+    are indistinguishable. Inset shows the cumulative gate errors
+    plotted on a log scale for small gate counts.
+  }
+  \label{fig:longitude}
+\end{figure*}
+
+Next, we outline the cost functions for the fluxonium gate problem.
+The augmented state and augmented control are:
+\begin{equation}
+  \begingroup
+  \renewcommand*{\arraystretch}{1.3}
+  x_{k} = \begin{bmatrix} \ket{\psi^{0}_{k}} \\ \ket{\psi^{1}_{k}}
+    \\ \int_{t} a_{k} \\ a_{k} \\ \partial_{t} a_{k} \end{bmatrix} \quad
+  u_{k} = \begin{bmatrix} \partial^{2}_{t} a_{k} \end{bmatrix}
+  \endgroup
+  \label{eq:astatecontrols}
+\end{equation}
+The cost function at each knot point is
+$\ell_{k}(x_{k}, u_{k}) = (x_{k} - x_{f})^{T} Q_{k} (x_{k} - x_{f}) + u^{T}_{k} R_{k} u_{k}$
+where $Q_{k}$ and $R_{k}$ are matrices we supply, and
+$x_{f}$ is given by the constraints we have imposed on
+$\ket{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$ in addition to
+$\partial_{t} a_{f} = 0$. We choose $Q_{k}$ to be diagonal
+so the cost function incentivizes the achievement of the constraints on
+$\ket{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$,
+and penalizes the norm of $\partial_{t} a_{k}$ and $\partial^{2}_{t} a_{k}$,
+mitigating AWG ringing due to high-frequency transitions.
+In the discrete dynamics function for this problem \eqref{eq:dyn_con},
+we integrate the states according to the TDSE \eqref{eq:tdse} and the
+fluxonium Hamiltonian \eqref{eq:hamiltonian} and integrate
+the second-derivative of the flux amplitude to obtain
+the first-derivative, proportional, and integral moments.
+Stated succinctly, the optimization problem takes the form:
 \begin{mini!}[2] 
-  {x_{1:N}, u_{1:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
-    + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{} \label{eq:costfun}
+  {x_{1:N}, u_{1:N\text{-}1}}{\sum_{k=1}^N {(x_k\text{-}x_f)}^{T} Q_k (x_k\text{-}x_{f})
+    + \sum_{k=1}^{N-1} {u_k}^{T} R_k u_{k}}{}{} \label{eq:costfun}
   \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
   \addConstraint{\ket{\psi^{0}_{1}} = \ket{0}, \ket{\psi^{1}_{1}} = \ket{1}} \label{eq:istate_con}
-  \addConstraint{\ket{\psi^{i}_{N}} = \ket{\psi^{i}_{f}}} \label{eq:tstate_con}
+  \addConstraint{\ket{\psi^{i}_{N}} = U \ket{\psi^{i}_{1}}
+    \ \forall \  i} \label{eq:tstate_con}
   \addConstraint{{\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}}
-      \rvert}^{2} = 1} \label{eq:statenorm_con}
-  \addConstraint{{\textstyle \int^{t_{1}}_{0} a \ \mathrm{d}t} = \pdv*{a_{1}}{t} = 0}
-  \label{eq:icintd_con}
-  \addConstraint{{\textstyle \int^{t_{N}}_{0}} a \ \mathrm{d}t= 0} \label{eq:znf_con}
-  \addConstraint{a_{1} = a_{N} = 0} \label{eq:concat_con}
-  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz}} \label{eq:amp_con}
+      \rvert}^{2} = 1 \ \forall \ i, k} \label{eq:statenorm_con}
+  \addConstraint{{\textstyle \int_{t}} a_{N} = 0} \label{eq:znf_con}
+  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz} \ \forall \ k} \label{eq:amp_con}
+  \addConstraint{a_{1} = a_{N} = 0} \label{eq:bound_con}
+  \addConstraint{{\textstyle \int_{t}} a_{1} = \partial_{t} a_{1} = 0} \label{eq:ic_con}
 \end{mini!}
-The augmented state and augmented controls are:
-\begin{equation}
-  x = \begin{bmatrix} \psi^{0} \\ \psi^{1} \\ \int^{t}_{0} a \ \mathrm{d}t\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
-  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
-  \label{eq:astatecontrols}
-\end{equation}
-The matrices $Q_{k}$ and $R_{k}$ define the penalty metric.
-Including objectives at each knot point smoothens the optimization landscape, though
-the importance of the final objective is encoded in the relative
-weight $Q_{N} \sim N \cdot Q_{k}$.
-We choose $Q$ to be diagonal because it is computationally efficient. This
-metric penalizes phase differences between the states, although
-phase-insensitive metrics such as infidelity may be employed.
-
-We impose constraints that
-reflect the physical limitations of the apparatus and 
-improve the experimental realization of the control pulse.
-The discrete dynamics \eqref{eq:dyn_con} integrates the states
-according to the TDSE \eqref{eq:tdse} and the
-fluxonium Hamiltonian \eqref{eq:hamiltonian}
-and integrates the moments of the flux amplitude.
-Exposing lower order moments of the flux amplitude allows us to penalize their norms
-in \eqref{eq:costfun}, smoothing the flux amplitude
-and mitigating AWG ringing due to high frequency transitions.
-We ensure the desired gate is achieved by imposing the constraints
-\eqref{eq:istate_con} and \eqref{eq:tstate_con}.
-The initial states in \eqref{eq:istate_con}
-are chosen to span the Hilbert space.
-The final states in \eqref{eq:tstate_con} are the image
-of the initial states under the desired gate.
-The ALTRO implementation we use does not currently support complex numbers so
-we compute in the isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-given in \cite{leung2017speedup},
-\begin{equation}
-  H \ket{\psi} \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
-    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
-  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
-  \label{eq:isomorphism}
-\end{equation}
-We ensure the solver does not take advantage of discretization
-error by imposing the normalization constraint \eqref{eq:statenorm_con}.
-The constraint \eqref{eq:icintd_con} is an initial condition.
-We ensure the pulse has zero net flux by imposing \eqref{eq:znf_con},
-which mitigates the hysteresis ubiquitous in flux bias lines
-\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
-We ensure gates may be concatenated arbitrarily without
-inducing AWG ringing due to high-frequency transitions by imposing
-the boundary condition \eqref{eq:concat_con}.
-Finally, we ensure the two-level
-approximation \eqref{eq:hamiltonian} remains valid
-by imposing the amplitude constraint \eqref{eq:amp_con}.
-All of the gates presented in this work achieve a maximum
-individual constraint violation of $10^{-8}$
-for \eqref{eq:dyn_con}--\eqref{eq:amp_con}.

### 2021-01-12 paper/6stochastic.tex
added constraints to opt statement
majority of jens and colm comments adressed

@@
-\section{Robustness to Time-Dependent Parameter Deviations \label{sec:stochastic}}
+\section{Robustness to Time-Dependent Parameter Uncertainty \label{sec:stochastic}}
 An additional source of experimental error arises from time-dependent
-Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
+Hamiltonian parameter uncertainty. For many flux-biased and inductively-coupled
 superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Magnetic flux noise
+source of coherent errors \todo{coherent errors?}. Flux noise
 modifies the fluxonium Hamiltonian \eqref{eq:hamiltonian}
-by $a(t) \gets a(t) + \delta a(t)$.
-The spectral density of $\delta a(t)$ follows a
-1/$f$ distribution for a range of devices, consisting primarily of low frequency
-noise \cite{bialczak20071f, koch2007model,
-  yoshihara2006decoherence, yoshihara2010correlated}.
-Analytic methods to combat flux noise
-take advantage of the low-frequency characteristic and
-treat the noise as quasi-static, performing generalizations of the spin-echo technique
-to compensate for erroneous drift. The analytic gate considered here
-follows this strategy.
+by $a(t) \rightarrow a(t) + \delta a(t)$.
+The spectral density of flux noise is known to
+follow a 1/$f$ distribution
+\cite{bialczak20071f, koch2007model,
+  yoshihara2006decoherence, yoshihara2010correlated},
+so the noise is dominated by low-frequency components.
+The analytic gate considered here
+takes advantage of the low-frequency characteristic and
+treat the noise as quasi-static, performing a generalization of the spin-echo
+technique to compensate for erroneous drift.
 
-We challenge the analytic method and
-the robust control
-methods of the previous section
-to realize an $X/2$ gate subject to 1/$f$ flux noise.
-We compute gate errors for each method
-by evolving the system under the Hamiltonian in \eqref{eq:hamiltonian}
-where the optimized flux amplitude is modified $a(t) \gets a(t) + \delta a(t)$.
-The flux noise $\delta a(t)$ is generated by
-filtering white noise sampled from a standard normal distribution with a finite
-impulse response filter \cite{saspweb2011}.
-It is then scaled by the 
+Additionally, we modify the robust control techniques presented
+in the previous section to combat 1/$f$ flux noise.
+The unscented sampling method is modified so that the sample states
+are subject to 1/$f$ flux noise. The noise
+is generated by filtering white noise sampled from a standard
+normal distribution with a finite impulse response filter \cite{saspweb2011}.
+The noise is then scaled by the 
 flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
-\delta a (t) \sim \sigma_{a} = 2.5 \cdot 10^{-5} \textrm{GHz}$.
-The unscented sampling method is modified so that its samples
-are subject to 1/$f$ flux noise by carrying the state of a finite impulse response filter
-in the augmented state vector \eqref{eq:astatecontrols}.
+\sigma_{a} = 2.5 \cdot 10^{-5} \textrm{GHz}$.
 In principle, we could modify the sampling method
-similarly; however, we choose to sample statically at $\sigma_{a}$ for comparison.
-The derivative methods require no algorithmic modification from the static case,
-but the TDSE is now differentiated with respect to $a(t)$ instead of $f_{q}$ as
-in \eqref{eq:d1dyn}.
+similarly; however, we choose to sample statically $a(t) \rightarrow a(t) + \sigma_{a}$
+for comparison. The derivative methods require no algorithmic modification
+from the static case, but the TDSE is now differentiated with respect
+to $a(t)$ instead of $f_{q}$ as in \eqref{eq:d1dyn}.
 
+We compare the gate errors due to 1/$f$ flux noise for $X/2$
+gates constructed with the robust control techniques
+and the analytic method. To compute the gate error,
+an initial state is evolved
+under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
+where the optimized flux amplitude is modified $a(t) \rightarrow a(t) + \delta a(t)$.
+The flux noise $\delta a(t)$ is generated as
+we described for the unscented sampling method.
+Each of the $1000$ averages we perform uses a different random seed for
+generating the flux noise.
 We simulate successive applications of the gate constructed by each method
@@
 Both the analytic
-and numerical methods achieve single gate errors
+and numerical gates yield single gate errors
 sufficient for quantum error correction.
 Despite converging on qualitatively different solutions, the
-numerical methods perform similarly in the concatenated
-gate application comparison. They achieve a two
-order of magnitude cumulative gate error reduction over the analytic method after $200$
-gate applications $\sim 11 \mu\textrm{s}$.
+numerical gates perform similarly in the concatenated
+gate application comparison. Their gate errors
+after $200$ gate applications $\sim 11 \mu\textrm{s}$ are
+two orders of magnitude less than the gate error produced by the analytic gate.
+\todo{motional narrowing}.
 1/$f$ flux noise is a significiant source of coherent errors in NISQ applications and

### 2021-01-12 paper/7conclusion.tex
majority of jens and colm comments adressed

@@
 In conclusion, we have applied state-of-the-art trajectory
-optimization techniques to achieve robustness to parameter deviations
-and mitigate decoherence on a quantum system.
-We have proposed a scheme for mitigating
-longitudinal relaxation with time-optimal
-control and the probability of longitudinal relaxation metric,
-avoiding the optimal control problem for density matrices which scales quartically with
-the dimension of the Hilbert space.
-We have proposed the derivative method for robust control which achieves
-super-linear gate error reductions in its gate time for the static parameter
-deviation problem we studied.
+optimization techniques to mitigate decoherence and
+achieve robustness to parameter uncertainty
+errors on a quantum system.
+We have proposed a scheme for supressing
+depolarization with time-optimal
+control and the depolarization probability model.
+The computational cost of this depolarization model is
+independent of the dimension of the Hilbert space, enabling
+inexpensive optimization on high-dimensional quantum systems.
+We have also proposed the derivative method for robust control which achieves
+super-linear gate error reductions in the gate time for the static parameter
+uncertainty problem we studied.
 We have shown that the derivative, sampling, and unscented sampling methods
@@
 which dominate coherent errors for flux controlled qubits.
-The derivative and sampling methods scale cubically with
-the dimension of the Hilbert space, allowing for applications
-to larger quantum systems.
-The derivative, sampling, and unscented sampling
-methods will benefit from interleaving optimization with
-experimental characterization, while the sampling and unscented
-sampling methods can also be applied to error channels without models.
-These techniques will be used to achieve the low gate errors
+The robust control techniques can be applied
+to any Hamiltonian,
+allowing experimentalists in all disciplines to engineer robust
+operations on their quantum systems.
+These methods will be used to achieve the low gate errors
 required for fault-tolerant quantum computing applications. Our

### 2021-01-12 paper/2background.tex
diffed main

@@
-\section{QOC + ALTRO \label{sec:background}}
-In this section we
+\section{Background \label{sec:background}}
+In this section, we
 review the QOC problem statement
-and introduce the ALTRO solver.
-QOC concerns the evolution of
-a (quantum) state $\ket{\psi(t)}$ governed by the time-dependent
+and reintroduce the ALTRO solver \cite{howell2019altro}.
+QOC concerns a vector of time-dependent control parameters $u(t)$
+that steer the evolution of a (quantum) state $\ket{\psi(t)}$.
+The evolution of the state is governed by the time-dependent
 Schr{\"o}dinger equation (TDSE),
@@
 \end{equation}
-The Hamiltonian has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
-The controls are so-called because they are the means the experimentalist has to
-act on the system. To make the problem numerically tractable,
-the controls and quantum state are discretized into $N$ knot points (time steps).
-The optimization problem is to find the
-controls that minimize a functional $J(u)$.
-In the simplest case, the functional is
-the infidelity between the initial state evolved
-to the final knot point and the desired final state
-$J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u)} \rvert}^{2}$.
-In general, $J$ is a linear combination of cost functions on the state
+The Hamiltonian $H(u(t), t)$ is determined by the quantum system.
+The QOC problem is to find the
+controls that minimize a functional $J(u(t))$.
+To make the problem numerically tractable,
+the controls and state are discretized into $N$ knot points (time steps).
+In the case of a single state transfer problem, the functional is
+the infidelity of the desired final state and the initial state evolved
+to the final knot point,
+$J(u) = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u)} \rvert}^{2}$.
+In general, $J(u)$ is a linear combination of cost functions on the state
 as well as cost functions on the controls \cite{leung2017speedup}.
-Standard QOC
-solvers employ automatic differentiation
-to compute gradients of the functional $\nabla_{u} J(u)$,
-which can easily be used to implement first-order optimization methods
-\cite{machnes2015gradient, khaneja2005optimal, leung2017speedup, goerz2019krotov}.
+Standard QOC solvers compute derivatives of the functional $\nabla_{u} J(u)$,
+which are used to implement first-order optimization methods
+\cite{machnes2015tunable, khaneja2005optimal, leung2017speedup, goerz2019krotov}.
 
-Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
+\myrem{Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
 and solved using a variety of specialized solvers developed by the robotics community
@@
 and equality constraints, potentially including initial and final conditions,
-at time step $k$.
+at time step $k$.}
 
-Many techniques have been proposed for solving \eqref{eq:gcostfun}--\eqref{eq:eqcon}. Standard 
+\myadd{Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
+and solved using a variety of specialized solvers developed by the robotics community
+\cite{Schulman13, Tedrake16, Hereid2017FROST, howell2019altro}.
+The functional $J(u)$ is divided into its constituent
+cost functions at each knot point $\ell_{k}(x_{k}, u_{k})$,
+where $k \in \{1, ..., N\}$ denotes the knot point,
+$x_{k}$ is the augmented state vector,
+and $u_{k}$ is the augmented control vector.
+The augmented state includes all variables of interest that are not decision variables.
+In the case of QOC, the state is a subset of the
+augmented state $\ket{\psi_{k}} \subseteq x_{k}$.
+The solver we use does not currently support complex numbers, so we employ the isomorphism
+$\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
+given in \cite{leung2017speedup},
+\begin{equation}
+  H \ket{\psi} \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
+    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
+  \begin{bmatrix} \ket{\psi}_{\textrm{re}} \\ \ket{\psi}_{\textrm{im}}\end{bmatrix}
+  \label{eq:isomorphism}
+\end{equation}
+The augmented state trajectory obeys the physics of the system if
+the dynamics constraint is satisfied
+$x_{k + 1} = f(x_{k}, u_{k})$. For QOC, the discrete
+dynamics function $f(x_{k}, u_{k})$ propagates the state by
+integrating the TDSE \ref{eq:tdse} with Runge-Kutta methods \cite{jorgensen2011numerical}
+or exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}.}
+
+\myadd{Additional constraints on the augmented states and augmented controls
+are encoded in constraint functions.
+The constraint functions are put into a form such that, when the constraint is satisfied,
+inequality constraint functions obey $g_{k}(x_{k}, u_{k}) \leq 0$ and
+equality constraint functions obey $h_{k}(x_{k}, u_{k}) = 0$. The constraint
+functions may be vector-valued to encode multiple constraints of each type.
+If a constraint function takes a positive value, the
+value is called the constraint's violation and
+the constraint is said to be active.
+The trajectory optimization problem can be stated concisely,
+\begin{mini!}[2]
+    {x_{1:N},u_{1:N-1}}{\ell_{N}(x_N) + \sum_{k=1}^{N-1} \ell_{k}(x_k,u_k)}{}{} \label{eq:gcostfun}
+    \addConstraint{x_{k+1} = f(x_k,u_k)} \label{eq:gdyncon}
+    \addConstraint{g_k(x_k,u_k)}{\leq 0} \label{eq:ineqcon}
+    \addConstraint{h_k(x_k,u_k)}{=0} \label{eq:eqcon}
+\end{mini!}}
+
+\myrem{Many techniques have been proposed for solving
+\eqref{eq:gcostfun}--\eqref{eq:eqcon}. Standard 
 methods include direction collocation \cite{Hargraves87} and differential-dynamic programming
 (DDP) \cite{Mayne1966a}. Recent state-of-the-art solvers, such as ALTRO \cite{howell2019altro},
-have combined principles from both of these approaches.
+have combined principles from both of these approaches.}
 
-ALTRO uses iterative LQR (iLQR) \cite{Li2004a} as the internal solver of an augmented 
+\myadd{Many techniques have been proposed for solving
+\eqref{eq:gcostfun}-\eqref{eq:eqcon}. Standard 
+methods include direction collocation \cite{Hargraves87} and differential-dynamic programming
+(DDP) \cite{Mayne1966a}. Recent state-of-the-art solvers, such as ALTRO,
+have combined principles from both of these approaches. ALTRO uses iLQR
+\cite{Li2004a} as the internal solver of an augmented Lagrangian method (ALM)
+\cite{lantoine2012hybrid, plancher2017constrained}
+and employs a projected Newton method \cite{bertsekas1982projected, rao1998application}
+in its final solving stage.}
+
+\myrem{ALTRO uses iterative LQR (iLQR) \cite{Li2004a} as the internal solver of an augmented 
 Lagrangian method (ALM). iLQR solves an unconstrained trajectory optimization problem 
@@
 \end{equation}
-the penalty terms are updated, and the process repeats until convergence.
+the penalty terms are updated, and the process repeats until convergence.}
 
-ALM converges superlinearly but tends to exhibit slow constraint convergence near the optimal 
+\myadd{iLQR simulates the augmented state forward
+  using the discrete dynamics function 
+  and computes derivatives of the cost functions
+  to update the augmented controls. This procedure is known as
+  indirect shooting, and iLQR shares this approach with GOAT \cite{machnes2015tunable},
+  GRAPE \cite{khaneja2005optimal, leung2017speedup}, and Krotov's method \cite{goerz2019krotov}.
+  iLQR differs from these solvers
+  in its update procedure. iLQR computes the locally optimal update for the augmented
+  controls using a dynamic programming algorithm, and a line search is performed
+  in the direction of this update to ensure a decrease in the sum of the cost functions.}
+
+\myadd{Indirect shooting solvers such as iLQR are popular
+for solving unconstrained trajectory optimization problems \eqref{eq:gcostfun}-\eqref{eq:gdyncon},
+but standard implementations have no ability to handle nonlinear equality and inequality
+constraints \eqref{eq:ineqcon}-\eqref{eq:eqcon}.
+One way to handle constraints is to employ a projected gradient
+method \cite{clarkson2010coresets, hauswirth2016projected}.
+However, within the indirect shooting framework,
+projected gradient methods can
+only handle constraints on the augmented controls,
+not the augmented states.
+Another technique is to promote the constraint functions to cost functions \cite{leung2017speedup}.
+A simple implementation of this approach does not garauntee that the constraints
+are satisfied, as the solver trades
+minimization of the cost functions and constraint functions against each other.
+ALM promotes the constraint functions to cost functions, and adaptively
+adjusts a Lagrange multiplier estimate for each constraint function to ensure
+the constraints are satisfied. ALM considers a decision variable $z \in \R^{s}$ with $s$ given,
+a cost function $f(z) : \R^{s} \mapsto \R$, and a constraint function
+$c(z) : \R^{s} \mapsto \R^{p}$ encoding $p$ constraints.
+For ALTRO, $z$ is the vector concatenation of the augmented states and augmented controls across
+all knot points, $f(z)$ is the sum of the trajectory optimization
+cost functions \eqref{eq:gcostfun},
+and $c(z)$ is the vector concatenation of the trajectory
+optimization constraint functions \eqref{eq:ineqcon}-\eqref{eq:eqcon}.
+ALM successively solves unconstrained minimization 
+problems of the form:
+\begin{mini}[2]
+    {z} {f(z) + \lambda^T c(z) + \half c(z)^T I_\mu c(z)}{}{}
+    \label{opt:alm}
+\end{mini}
+$\lambda \in \R^p$ is a Lagrange multiplier estimate, and $I_\mu$ is a diagonal matrix
+of penalty weights, $\mu \in \R^{p}$, whose magnitudes depend on whether
+the corresponding constraint is active or inactive.
+After minimizing \eqref{opt:alm} using iLQR,
+the Lagrange multiplier estimates are updated according to,
+\begin{equation}
+  \lambda \gets \lambda + I_\mu c(z)
+  \label{eq:almupdate}
+\end{equation}
+the penalty weight is updated, and the process repeats until convergence.}
+
+\myrem{ALM converges superlinearly but tends to exhibit
+  slow constraint convergence near the optimal 
   solution due to poor numerical conditioning. To address this shortcoming, ALTRO provides a 
@@
   machine-precision constraint satisfaction. For more information on the details of the ALTRO
-solver, see \cite{howell2019altro, Jackson2020altroc}.
+  solver, see \cite{howell2019altro, Jackson2020altroc}.}
 
-As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
+\myadd{ALM converges superlinearly, but poor numerical conditioning may lead
+to small decreases in the constraint violations near the locally optimal solution
+\cite{bertsekas1996constrained}. To address this shortcoming, ALTRO
+projects the solution from the ALM solving stage onto the constraint manifold using
+a projected Newton method, achieving ultra-low
+constraint violations $\sim 1 \cdot 10^{-8}$.
+For more information on the details of the ALTRO
+solver, see \cite{howell2019altro, Jackson2020altroc}.}
+
+\myrem{As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
 for imposing arbitrary, simultaneous constraints
@@
 for this work, where multiple constraints must be satisfied in addition
-to minimizing auxiliary metrics.
+to minimizing auxiliary metrics.}
+
+\myadd{As opposed to standard QOC solvers, ALTRO
+is able to satisfy constraints
+on both the augmented states and the augmented controls to tight tolerances.
+This advantage is crucial for this work, where multiple medium-priority cost functions
+are minimized subject to many high-priority constraints.}

### 2021-01-12 paper/3fluxonium.tex
diffed main

@@
 
-First, we outline the constraints for the fluxonium gate problem.
+\myrem{The optimization problem takes the form:
+\begin{mini!}[2] 
+  {x_{1:N}, u_{1:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
+    + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{} \label{eq:costfun}
+  \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
+  \addConstraint{\ket{\psi^{0}_{1}} = \ket{0}, \ket{\psi^{1}_{1}} = \ket{1}} \label{eq:istate_con}
+  \addConstraint{\ket{\psi^{i}_{N}} = \ket{\psi^{i}_{f}}
+    \ \forall \  i \in \{0, 1\}} \label{eq:tstate_con}
+  \addConstraint{{\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}}
+      \rvert}^{2} = 1 \ \forall \ i \in \{0, 1\}} \label{eq:statenorm_con}
+  \addConstraint{{\textstyle \int^{t_{1}}_{0} a \ \mathrm{d}t} = \pdv*{a_{1}}{t} = 0}
+  \label{eq:icintd_con}
+  \addConstraint{{\textstyle \int^{t_{N}}_{0}} a \ \mathrm{d}t= 0} \label{eq:znf_con}
+  \addConstraint{a_{1} = a_{N} = 0} \label{eq:concat_con}
+  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz}} \label{eq:amp_con}
+\end{mini!}
+The augmented state and augmented controls are:
+\begin{equation}
+  x = \begin{bmatrix} \psi^{0} \\ \psi^{1} \\ \int^{t}_{0} a \ \mathrm{d}t\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
+  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
+  \label{eq:astatecontrols}
+\end{equation}
+The matrices $Q_{k}$ and $R_{k}$ define the penalty metric.
+Including objectives at each knot point smoothens the optimization landscape, though
+the importance of the final objective is encoded in the relative
+weight $Q_{N} \sim N \cdot Q_{k}$.
+We choose $Q$ to be diagonal because it is computationally efficient. This
+metric penalizes phase differences between the states, although
+phase-insensitive metrics such as infidelity may be employed.}
+
+\myrem{We impose constraints that
+reflect the physical limitations of the apparatus and 
+improve the experimental realization of the control pulse.
+The discrete dynamics \eqref{eq:dyn_con} integrates the states
+according to the TDSE \eqref{eq:tdse} and the
+fluxonium Hamiltonian \eqref{eq:hamiltonian}
+and integrates the moments of the flux amplitude.
+Exposing lower order moments of the flux amplitude allows us to penalize their norms
+in \eqref{eq:costfun}, smoothing the flux amplitude
+and mitigating AWG ringing due to high frequency transitions.
+We ensure the desired gate is achieved by imposing the constraints
+\eqref{eq:istate_con} and \eqref{eq:tstate_con}.
+The initial states in \eqref{eq:istate_con}
+are chosen to span the Hilbert space.
+The final states in \eqref{eq:tstate_con} are the image
+of the initial states under the desired gate.
+The ALTRO implementation we use does not currently support complex numbers so
+we compute in the isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
+given in \cite{leung2017speedup},
+\begin{equation}
+  H \ket{\psi} \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
+    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
+  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
+  \label{eq:isomorphism}
+\end{equation}
+We ensure the solver does not take advantage of discretization
+error by imposing the normalization constraint \eqref{eq:statenorm_con}.
+The constraint \eqref{eq:icintd_con} is an initial condition.
+We ensure the pulse has zero net flux by imposing \eqref{eq:znf_con},
+which mitigates the hysteresis ubiquitous in flux bias lines
+\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
+We ensure gates may be concatenated arbitrarily without
+inducing AWG ringing due to high-frequency transitions by imposing
+the boundary condition \eqref{eq:concat_con}.
+Finally, we ensure the two-level
+approximation \eqref{eq:hamiltonian} remains valid
+by imposing the amplitude constraint \eqref{eq:amp_con}.
+All of the gates presented in this work achieve a maximum
+individual constraint violation of $10^{-8}$
+for \eqref{eq:dyn_con}--\eqref{eq:amp_con}.}
+
+\myadd{First, we outline the constraints for the fluxonium gate problem.
 We formulate this problem as a multi-state-transfer problem.
@@
 we have the initial condition $\int_{t} a_{1} = \partial_{t} a_{1} = 0$
-\eqref{eq:ic_con}.
+\eqref{eq:ic_con}.}
 
@@
 
-Next, we outline the cost functions for the fluxonium gate problem.
+\myadd{Next, we outline the cost functions for the fluxonium gate problem.
 The augmented state and augmented control are:
@@
   \addConstraint{{\textstyle \int_{t}} a_{1} = \partial_{t} a_{1} = 0} \label{eq:ic_con}
-\end{mini!}
+\end{mini!}}

### 2021-01-12 paper/7conclusion.tex
diffed main

@@
 which dominate coherent errors for flux controlled qubits.
-The robust control techniques can be applied
+\myrem{The derivative and sampling methods scale cubically with
+the dimension of the Hilbert space, allowing for applications
+to larger quantum systems.
+The derivative, sampling, and unscented sampling
+methods will benefit from interleaving optimization with
+experimental characterization, while the sampling and unscented
+sampling methods can also be applied to error channels without models.}
+\myadd{The robust control techniques can be applied
 to any Hamiltonian,
-allowing experimentalists in all disciplines to engineer robust
-operations on their quantum systems.
+allowing experimentalists in all domains to engineer robust
+operations on their quantum systems.}
 These methods will be used to achieve the low gate errors

### 2021-01-14 src/spin/spin.jl
updates for rho2_traces

@@ -1234,6 +1234,29 @@ end
 
 
 """
+Given an array of states, compute their
+"""
+function compute_rho2_traces(states)
+    state_type = length(size(states)) == 2 ? st_state : st_density
+    if state_type == st_state
+        (state_count, _) = size(states)
+        rho2_traces = zeros(state_count)
+        for i = 1:state_count
+            state = states[i, :]
+            state_uniso = get_vec_uniso(state) 
+            density = state_uniso * state_uniso'
+            rho2_trace = real(tr(density * density))
+            rho2_traces[i] = rho2_trace
+        end
+    else
+        (state_count, _, _) = size(states)
+        rho2_traces = zeros(state_count)
+    end
+    return rho2_traces
+end
+
+
+"""
 run_sim_deqjl - Apply a gate multiple times and measure the fidelity
 after each application. Save the output.
 
@@ -1557,6 +1580,8 @@ function run_sim_prop(
     integrator(gate_count, states, state, params)
     # compute fidelities
     fidelities = compute_fidelities(gate_count, gate_type, states)
+    # compute traces
+    rho2_traces = compute_rho2_traces(states)
 
     # report
     result = Dict(
@@ -1570,7 +1595,8 @@ function run_sim_prop(
         "namp" => namp,
         "ndist" => string(ndist),
         "noise_dt_inv" => noise_dt_inv,
-        "save_file_path" => save_file_path
+        "rho2_traces" => rho2_traces,
+        "save_file_path" => save_file_path,
     )
 
     return result

### 2021-01-14 src/spin/spin27.jl

@@ -1,0 +1,500 @@
+"""
+spin27.jl - unscented sampling robustness via nominal penalization for the δfq problem
+"""
+
+WDIR = joinpath(@__DIR__, "../../")
+include(joinpath(WDIR, "src", "spin", "spin.jl"))
+
+using Altro
+using Debugger
+using Distributions
+using HDF5
+using Hyperopt
+using ForwardDiff
+using LinearAlgebra
+using Random
+using RobotDynamics
+using StaticArrays
+using Zygote
+using TrajectoryOptimization
+const RD = RobotDynamics
+const TO = TrajectoryOptimization
+
+# paths
+const EXPERIMENT_META = "spin"
+const EXPERIMENT_NAME = "spin27"
+const SAVE_PATH = abspath(joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME))
+
+# problem
+const CONTROL_COUNT = 1
+const STATE_COUNT = 4
+const ASTATE_SIZE_BASE = STATE_COUNT * HDIM_ISO + 3 * CONTROL_COUNT
+# state indices
+const STATE1_IDX = 1:HDIM_ISO
+const STATE2_IDX = STATE1_IDX[end] + 1:STATE1_IDX[end] + HDIM_ISO
+const STATE3_IDX = STATE2_IDX[end] + 1:STATE2_IDX[end] + HDIM_ISO
+const STATE4_IDX = STATE3_IDX[end] + 1:STATE3_IDX[end] + HDIM_ISO
+const INTCONTROLS_IDX = STATE4_IDX[end] + 1:STATE4_IDX[end] + CONTROL_COUNT
+const CONTROLS_IDX = INTCONTROLS_IDX[end] + 1:INTCONTROLS_IDX[end] + CONTROL_COUNT
+const DCONTROLS_IDX = CONTROLS_IDX[end] + 1:CONTROLS_IDX[end] + CONTROL_COUNT
+# control indices
+const D2CONTROLS_IDX = 1:CONTROL_COUNT
+# sample indices
+const S1_IDX = SVector{HDIM_ISO}(HDIM_ISO * 0 + 1:HDIM_ISO * 1)
+const S2_IDX = SVector{HDIM_ISO}(HDIM_ISO * 1 + 1:HDIM_ISO * 2)
[…]

### 2021-01-14 paper/0abstract.tex
updated figures, lots of proofreading

@@
 %% TITLE
-\title{Robust Control of a Fluxonium Qubit}
+\title{Robust Quantum Optimal Control}
 
@@
   The ability to engineer high-fidelity gates on quantum processors in the presence of
-  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
-  Quantum optimal control techniques have proven effective in experimentally
+  systematic errors remains the primary challenge requisite to achieving quantum advantage.
+  Quantum optimal control methods have proven effective in experimentally
   realizing high-fidelity gates, but they require exquisite calibration to be performant.
   We apply robust trajectory optimization techniques to suppress gate errors arising from system
-  parameter deviations and noise.
-  We propose a method that takes advantage of deviant parameter derivative information while maintaining
-  computational efficiency by utilizing mixed-mode differentiation.
-  Additionally, completely modeling decoherence effects due to longitudinal relaxation requires
-  integrating the Lindblad master equation, which is computationally expensive.
-  We propose a computationally efficient metric
-  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
-  We demonstrate these techniques numerically on a fluxonium qubit with realistic
-  experimental parameters and constraints,
-  achieving orders of magnitude gate error reductions from our baseline gate set.
+  parameter uncertainty.
+  We propose a method that takes advantage of uncertain parameter
+  derivative information while maintaining
+  computational efficiency by transforming high-order differential equations to coupled,
+  first-order ODEs.
+  Additionally, the effect of depolarization on a gate is most accurately modeled by
+  integrating the Lindblad master equation,
+  which is computationally expensive.
+  We propose a computationally efficient model
+  and utilize time-optimal control to achieve high fidelity gates in the presence of depolarization.
+  We apply these techniques to a fluxonium qubit with realistic
+  parameters and constraints,
+  achieving orders of magnitude gate-error reductions from our baseline gate set in simulations.
 \end{abstract}
+
+\maketitle

### 2021-01-14 paper/2background.tex

@@
 
-\myrem{Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
-and solved using a variety of specialized solvers developed by the robotics community
-\cite{Schulman13,Tedrake16,Hereid2017FROST,howell2019altro}.
-Trajectory optimization problems are typically of the following form: 
-\begin{mini!}[2]
-    {x_{1:N},u_{1:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell(x_k,u_k)}{}{} \label{eq:gcostfun}
-    \addConstraint{x_{k+1} = f(x_k,u_k)} \label{eq:gdyncon}
-    \addConstraint{g_k(x_k,u_k)}{\leq 0} \label{eq:ineqcon}
-    \addConstraint{h_k(x_k,u_k)}{=0} \label{eq:eqcon}
-\end{mini!}
-where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^l$ and
-$u_k \in \R^m$ are the state and control variables,
-$f(x_k,u_k)$ is the discrete dynamics function, and $g_k(x_k,u_k)$ and
-$h_k(x_k,u_k)$ are the inequality
-and equality constraints, potentially including initial and final conditions,
-at time step $k$.}
-
-\myadd{Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
+Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
 and solved using a variety of specialized solvers developed by the robotics community
@@
 integrating the TDSE \ref{eq:tdse} with Runge-Kutta methods \cite{jorgensen2011numerical}
-or exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}.}
+or exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}.
 
-\myadd{Additional constraints on the augmented states and augmented controls
+Additional constraints on the augmented states and augmented controls
 are encoded in constraint functions.
@@
     \addConstraint{h_k(x_k,u_k)}{=0} \label{eq:eqcon}
-\end{mini!}}
+\end{mini!}
 
-\myrem{Many techniques have been proposed for solving
-\eqref{eq:gcostfun}--\eqref{eq:eqcon}. Standard 
-methods include direction collocation \cite{Hargraves87} and differential-dynamic programming
-(DDP) \cite{Mayne1966a}. Recent state-of-the-art solvers, such as ALTRO \cite{howell2019altro},
-have combined principles from both of these approaches.}
-
-\myadd{Many techniques have been proposed for solving
-\eqref{eq:gcostfun}-\eqref{eq:eqcon}. Standard 
+Many techniques have been proposed for solving \eqref{eq:gcostfun}-\eqref{eq:eqcon}. Standard 
 methods include direction collocation \cite{Hargraves87} and differential-dynamic programming
@@
 and employs a projected Newton method \cite{bertsekas1982projected, rao1998application}
-in its final solving stage.}
+in its final solving stage.
 
-\myrem{ALTRO uses iterative LQR (iLQR) \cite{Li2004a} as the internal solver of an augmented 
-Lagrangian method (ALM). iLQR solves an unconstrained trajectory optimization problem 
-using a backward Riccati recursion to derive a closed-loop linear feedback law about the current 
-trajectory. By simulating the system forward with the feedback law, the trajectory is 
-brought closer to the (locally) optimal trajectory. DDP-based solvers such as iLQR
-are popular since they are very computationally efficient, are always dynamically feasible,
-and provide a locally stabilizing closed-loop control policy about the optimal trajectory. However,
-standard implementations have no ability to deal with nonlinear equality and inequality 
-constraints. ALM handles constraints by successively solving unconstrained minimization 
-problems of the form:
-\begin{mini}[2]
-    {z}{f(z) + \lambda^T c(z) + \half c(z)^T I_\mu c(z)}{}{}
-    \label{opt:alm}
-\end{mini}
-where $f(z)$ is the objective function, $c(z) : \R^{l + m} \mapsto \R^p$ is the constraint 
-function, $\lambda \in \R^p$ is a Lagrange multiplier estimate, and $I_\mu$ is a diagonal matrix
-of penalty weights, $\mu$, whose magnitudes depend on whether
-the corresponding constraint is active or inactive.
-For ALTRO, $f(z)$ is \eqref{eq:gcostfun}, $c(z)$ is the 
-concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
-across all time steps. After minimizing \eqref{opt:alm} using iLQR,
-the Lagrange multiplier estimates are updated according to,
-\begin{equation}
-	\lambda \gets \lambda + \mu c(z)
-\end{equation}
-the penalty terms are updated, and the process repeats until convergence.}
-
-\myadd{iLQR simulates the augmented state forward
+iLQR simulates the augmented state forward
 using the discrete dynamics function 
@@
 controls using a dynamic programming algorithm, and a line search is performed
-  in the direction of this update to ensure a decrease in the sum of the cost functions.}
+in the direction of this update to ensure a decrease in the sum of the cost functions.
 
-\myadd{Indirect shooting solvers such as iLQR are popular
+Indirect shooting solvers such as iLQR are popular
 for solving unconstrained trajectory optimization problems \eqref{eq:gcostfun}-\eqref{eq:gdyncon},
@@
 \end{equation}
-the penalty weight is updated, and the process repeats until convergence.}
+the penalty weight is updated, and the process repeats until convergence.
 
-\myrem{ALM converges superlinearly but tends to exhibit
-  slow constraint convergence near the optimal 
-  solution due to poor numerical conditioning. To address this shortcoming, ALTRO provides a 
-  solution-polishing phase that takes 1-2 Newton steps on the active constraint set to provide 
-  machine-precision constraint satisfaction. For more information on the details of the ALTRO
-  solver, see \cite{howell2019altro, Jackson2020altroc}.}
-
-\myadd{ALM converges superlinearly, but poor numerical conditioning may lead
+ALM converges superlinearly, but poor numerical conditioning may lead
 to small decreases in the constraint violations near the locally optimal solution
@@
 For more information on the details of the ALTRO
-solver, see \cite{howell2019altro, Jackson2020altroc}.}
+solver, see \cite{howell2019altro, Jackson2020altroc}.
 
-\myrem{As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
-for imposing arbitrary, simultaneous constraints
-on both the states and the controls that converge superlinearly to tight tolerances. 
-First-order solvers such as GRAPE
-use unconstrained optimization to satisfy desiderata on the states,
-such as achieving a desired gate fidelity, forbidding occupations of particular quantum states,
-and smoothing the controls below a threshold. The unconstrained approach relies
-on a careful choice of competing penalty weights, which is cumbersome and often intractable.
-To impose constraints on the controls, such as amplitude restrictions, first-order solvers
-employ projected gradient descent methods which restrict
-optimization to the constraint manifold and
-may hinder convergence to the optimal solution \cite{clarkson2010coresets,
-  hauswirth2016projected}.
-Conversely,
-the ALTRO solver delays projecting onto the constraint manifold defined by
-\eqref{eq:ineqcon} and \eqref{eq:eqcon} until the ALM
-solve is already at a coarse tolerance, achieving machine-precision satisfaction
-for constraints on the controls and the states. These advantages are critical
-for this work, where multiple constraints must be satisfied in addition
-to minimizing auxiliary metrics.}
-
-\myadd{As opposed to standard QOC solvers, ALTRO
+As opposed to standard QOC solvers, ALTRO
 is able to satisfy constraints
@@
 This advantage is crucial for this work, where multiple medium-priority cost functions
-are minimized subject to many high-priority constraints.}
+are minimized subject to many high-priority constraints.

### 2021-01-14 paper/5static.tex

@@
-\section{Robustness to Static Parameter Deviations \label{sec:static}}
+%% F2
+\begin{figure*}[ht]
+  \begin{subfigure}{.315\textwidth}
+    \includegraphics[width=\linewidth]{assets/f2a.png}
+    \caption{\label{fig:statica}}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.4\textwidth}
+    \includegraphics[width=\linewidth]{assets/f2b.png}
+    \caption{\label{fig:staticb}}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.23\textwidth}
+    \includegraphics[width=\linewidth]{assets/f2c.png}
+    \caption{\label{fig:staticc}}
+  \end{subfigure}
+  \caption{
+    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
+    analytic (A), sampling (S), unscented sampling (SU), and the 1\textsuperscript{st}-
+    and 2\textsuperscript{nd}-order derivative methods (D1, D2). The numerical gates shown
+    are optimized for twice the gate time of the analytic gate.
+    (b) Single gate error at a one-percent qubit frequency detuning as
+    a function of the gate time. Missing
+    data points represent solutions with a gate error greater than $5 \cdot 10^{-5}$.
+    (c) Single gate error as a function of the qubit frequency detuning.
+    The gate errors for the analytic and 1\textsuperscript{st}-order derivative
+    methods are shown for gate times which are multiples of $1 / 4 f_{q} \sim 18 \textrm{ns}$.
+    The performance of the two methods is
+    indistinguishable at the gate time $18 \textrm{ns}$.
+  }
+  \label{fig:static}
+\end{figure*}
+
+\section{Robustness to Static Parameter Uncertainty \label{sec:static}}
 We have formulated the QOC
-problem as an open-loop optimization problem, i.e.
+problem as an open-loop optimization problem, equivalently,
 we do not incorporate feedback from the experiment in optimization.
@@
 making the state evolution insensitive
-to Hamiltonian parameter deviations. As an example,
+to Hamiltonian parameter uncertainty. As an example,
 we mitigate errors arising from the drift and finite measurement
-precision of the qubit frequency which modifies \eqref{eq:hamiltonian}
-by $f_{q} \gets f_{q} + \delta f_{q}$.
-We consider three robust control techniques:
+precision of the qubit frequency which modifies the fluxonium Hamiltonian
+\eqref{eq:hamiltonian} by $f_{q} \gets f_{q} + \delta f_{q}$.
+We consider three robust control techniques to accomplish this task:
 the sampling method, the unscented sampling method,
@@
 
-The sampling method simultaneously optimizes over copies of a state,
-each of which evolves with a distinct deviant parameter value.
-We append four sets of two sample states $\psi^{\pm}$
-to the augmented state vector \eqref{eq:astatecontrols}.
-We initialize each pair of samples in the same state.
-The four initial states are chosen so that their outerproducts
+The sampling method incentivizes the optimizer
+to ensure multiple copies of a state, each of which evolves
+with a distinct value of the uncertain parameter, achieve
+the same state transfer. For each initial state,
+we add two sample states $\ket{\psi^{\pm}}$
+to the augmented state \eqref{eq:astatecontrols}, and the discrete dynamics
+function \eqref{eq:dyn_con} is modified
+so the sample states evolve under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
+with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$ for a fixed hyperparameter
+$\sigma_{f_{q}}$ which is the standard deviation of the qubit frequency.
+For this method, the standard orthonormal basis states are an insufficient choice
+for the initial states. As an example, a $Z/2$ gate achieved by idling
+at the flux frustration point $a_{k} = 0 \ \forall \ k$
+will be robust to qubit frequency detunings for the initial states $\ket{0}$
+or $\ket{1}$ because the fidelity metric is insensitive to global phases,
+but this gate will not be robust for any other initial states.
+Thus, we choose four initial states
+so that their outer products
 span the operators on the Hilbert space
@@
 (\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-The discrete dynamics \eqref{eq:dyn_con} is modified so that the
-samples evolve under the Hamiltonian given in \eqref{eq:hamiltonian}
-with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$ for a fixed $\sigma_{f_{q}}$.
-The final state
-for each sample is the image of its initial state
-under the desired gate. The infidelity between
-the samples and their final states are penalized
-by adding a term
-$\sum_{k=1}^{N} q_{k} (1 - {\lvert \braket{\psi^{\pm}_{k}}{\psi_{f}} \rvert}^{2})$
-to the objective \eqref{eq:costfun}.
+The infidelities between the sample states and the images of their
+initial states under the desired gate are penalized
+by adding a cost function
+$\sum_{k, \pm} q_{k} (1 - {\lvert \braket{\psi^{\pm}_{k}}{U \psi^{\pm}_{1}} \rvert}^{2})$
+to the objective \eqref{eq:costfun} for a constant $q_{k}$ we supply.
 
-For the unscented sampling method, we append $2(2n + d)$ samples
-to the augmented state vector \eqref{eq:astatecontrols}
-for each initial state in the operator basis. Here $2n = 4$ is twice the
+For the unscented sampling method, we append $2(2n + d)$ sample
+states to the augmented state vector \eqref{eq:astatecontrols}
+for each initial state used in the sampling method. $2n = 4$ is twice the
 dimension of the Hilbert space, resulting from the isomorphism \eqref{eq:isomorphism},
-and $d = 1$ is the number of deviant parameters. The samples
-represent a Gaussian distribution over the $2n$
-elements of the initial state, modeling
-the uncertainty in the state as a result of the uncertainty in
-the deviant parameter. Each sample evolves under the Hamiltonian
-given in \eqref{eq:hamiltonian} with $f_{q} \gets f_{q} + \delta f_{q}$.
-We determine the qubit frequency detuning $\delta f_{q}$
-at each knot point using the statistics of the samples.
-We modify the discrete dynamics \eqref{eq:dyn_con} to
-evolve the samples under the deviant Hamiltonian
-and then apply the unscented transformation to the ensemble
-of samples,
-accurately preserving the first and second moments
-of the distribution.
-We penalize the infidelity between the samples and their final states
-by adding an infidelity contribution to \eqref{eq:costfun}
+and $d = 1$ is the number of uncertain parameters. The sample states
+encode a unimodal distribution over the $2n$
+elements of the evolving initial state, modeling
+the uncertainty in the state due to the uncertainty in the parameter.
+The discrete dynamics function \eqref{eq:dyn_con} is modified
+so each sample state evolves under the fluxonium Hamiltonian
+\eqref{eq:hamiltonian} where $f_{q} \gets f_{q} + \delta f_{q}$
+and $\delta f_{q}$ is sampled at each knot point from a distribution
+with fixed standard deviation $\sigma_{f_{q}}$.
+After propagating the sample states to each knot point, we apply the unscented
+transformation to the states, which
+preserves the first and second moments of the distribution the states encode.
+This procedure makes the unscented sampling method more resistant to sample
+states collapsing towards the mean than in the sampling method.
+We add a contribution to the objective \eqref{eq:costfun} for the infidelity
+of the sample states and their target states
 as in the sampling method.
@@
 The derivative method penalizes the sensitivty of the state
-to the deviant parameter, which is encoded in the ($l$\textsuperscript{th}-order)
-derivative $\partial_{f_{q}}^{l} \psi$. In the $m$\textsuperscript{th}-order
+to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
+state derivative $\partial_{f_{q}}^{l} \ket{\psi}$. In the $m$\textsuperscript{th}-order
 derivative method, we append all state derivatives of order $1, \dots, m$
 to the augmented state vector \eqref{eq:astatecontrols}
-for each initial state in the operator basis.
+for each initial state used in the sampling method.
 We penalize the norms of the state derivatives
-in \eqref{eq:costfun} by setting the corresponding final states to zero.
-We could obtain the state derivatives in the discrete dynamics \eqref{eq:dyn_con}
-with backwards mode differentiation.
-Naive automatic differentiation 
-would differentiate all dynamics at knot points
-$1, \dots, k - 1$ to obtain the state derivative at knot point $k$, requiring
+in the objective \eqref{eq:costfun} by setting the corresponding elements
+of the final augmented state $x_{f}$ to $0$.
+We could obtain the state derivatives at each knot point
+with backward-mode differentiation.
+In a naive automatic differentiation scheme,
+the discrete dynamics function at knot points
+$1, \dots, k - 1$ would be differentiated to obtain the state
+derivative at knot point $k$, requiring
 $O(N^{2})$ matrix multiplications. Instead, we 
-employ forwards mode differentiation on the TDSE to obtain coupled dynamics
+employ forward-mode differentiation on the TDSE \eqref{eq:tdse}
+to obtain coupled, first-order ODEs
 which require $O(N)$ matrix multiplications to integrate.
@@
 \end{align}
-Exponential integrators that account for the non-linear
-term in \eqref{eq:d1dyn} are used to efficiently integrate the coupled dynamics,
-see Appendix \ref{appendix:derivative}. For runtimes
-and complexity analyses for the three robust control methods,
+We integrate the coupled ODEs with exponential
+integrators, see Appendix \ref{appendix:derivative},
+in the discrete dynamics function \eqref{eq:dyn_con}. For runtimes
+and complexity analyses for the three robust control techniques,
 consult Appendix \ref{appendix:time}.
 
-We challenge these methods to achieve
-a $Z/2$ gate subject to a static qubit frequency detuning.
-We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
-the sampling methods accordingly. We compute gate errors for each method
-by evolving the system under the Hamiltonian in \eqref{eq:hamiltonian}
-with the optimized flux amplitude and $f_{q} \gets f_{q} + \delta f_{q}$
+We compare the gate errors due to a static qubit frequency
+detuning for $Z/2$ gates obtained with the robust control techniques
+and the analytic method.
+To compute the gate error,
+an initial state is evolved
+under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
+two separate times with the transformations
+$f_{q} \rightarrow f_{q} \pm \delta f_{q}$
 at the stated qubit frequency detuning.
+The reported gate error is the infidelity of
+the evolved state and the target state averaged over
+the two transformations and $1000$ initial states.
+We set $\sigma_{f_{q}}/f_{q} = 1\%$
+for the sampling and unscented sampling
+methods.
 
-We compare the numerical methods
-to an analytically derived $Z/2$ gate, see Figure \ref{fig:statica}. 
 The analytic gate corresponds to
-idling at the flux frustration point $a = 0$. It
-is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
-is simple to derive. Its erroneous rotation angle
-$2 \pi t_{Z/2} \delta f_{q}$ is linearly sensitive to
-the qubit frequency detuning, resulting in a gate error that is quadratically sensitive
-to the qubit frequency detuning.
+idling at the flux frustration point $a_{k} = 0 \ \forall \ k$, see Figure
+\ref{fig:statica}. Its gate time $1 / 4 f_{q} \sim 18\textrm{ns}$
+is the shortest possible for a $Z/2$ gate on the device.
+The gate's erroneous rotation angle
+$\pi \delta f_{q} / 2 f_{q}$ is linear in the
+qubit frequency detuning, resulting in a gate error that is quadratic
+in the detuning.
 At a one-percent
-qubit frequency detuning ($\delta f_{q} / f_{q} = 1\%$)
-the analytic gate achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
+qubit frequency detuning ($\lvert \delta f_{q} \rvert / f_{q} = 1\%$),
+the gate error is $\sim 4.5 \cdot 10^{-5}$,
 which is sufficient for quantum error correction.
-Although the analytic $Z/2$ gate performs well, it
-only works at the gate time $t_{Z/2}$. The ability to perform $Z$
-rotations in arbitrary times is critical
+Although the gate performs well, the method used to derive
+it can not be employed to achieve a $Z/2$ gate at gate times other than
+$1 / 4 f_{q}$. The ability to perform
+phase gates in any given time is critical
 for multi-qubit experiments, where the qubits operate at different
 frequencies $f_{q, i} \neq f_{q, j}$.
-Each numerical method can find solutions at
-all gate times above $t_{Z/2}$, see Figure \ref{fig:staticb}.
+We can find solutions using the numerical methods at
+all gate times above $18$ns, see Figure \ref{fig:staticb}.
 These numerical methods offer an effective scheme for synchronizing
@@
 
-The sampling and unscented sampling methods combine idling periods
-with time-anti-symmetric ramps.
-The gate error at $\delta f_{q} / f_{q} = 1\%$ for the sampling method
+The solution for the unscented sampling method combines idling periods
+with fast ramps to the maximum amplitude, whereas that for the sampling method
+does not reach high amplitudes.
+The gate error at a one-percent qubit frequency detuning for the sampling method
 does not improve substantially over the
-range of gate times. The unscented sampling method
-achieves linear decreases in its gate error at
-$\delta f_{q} / f_{q} = 1\%$ with longer gate times
-until half the Larmor period $1 / 2 f_{q}$ after which it achieves a consistent
+range of gate times. Conversely, the gate error at a one-percent detuning
+for the unscented sampling method decreases linearly in the gate time
+until half the Larmor period $1 / 2 f_{q}$ after which it has a consistent
 gate error $\sim 3.5 \cdot 10^{-5}$.
@@
 The derivative methods converge on qualitatively similar solutions that
-use fast triangle pulses at the boundaries and balance time
-on either side of the flux-frustration point symmetrically at low amplitudes.
-Both methods achieve a super-linear scaling in their gate error as
-a function of their gate time.
-We believe the 1\textsuperscript{st}-order method outperforms the 2\textsuperscript{nd}-order
-method due to the low contribution of second-order
-terms to the gate error in this deviation regime, see Appendix \ref{appendix:derivative}.
-The gate error for the 1\textsuperscript{st}-order
+use fast triangle pulses at the boundaries and idle at low amplitudes.
+The gate errors at a one-percent qubit frequency detuning
+for both methods decrease super-linearly in their gate times.
+The 1\textsuperscript{st}-order derivative method
+consistently achieves lower gate errors than the
+2\textsuperscript{nd}-order derivative method,
+which we believe is related to the greater importance
+of minimizing the 1\textsuperscript{st}-order state
+derivative norm than the 2\textsuperscript{nd}-order
+state derivative norm for this problem, see Appendix \ref{appendix:derivative}.
+The gate error at a one-percent detuning for the 1\textsuperscript{st}-order
 derivative method approaches zero at the Larmor period $1 / f_{q}$, see Figure \ref{fig:staticc}.
 This result mimics the
-ability of composite pulses to mitigate parameter deviation errors to arbitrary
+ability of composite pulses to mitigate parameter uncertainty errors to arbitrary
 order with sufficiently many pulses \cite{merrill2014progress}.
@@
   \caption{
-    (a) $X/2$ gates robust to flux amplitude offsets constructed with the analytic,
-    sampling, unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates shown
-    are the solutions at the analytic gate time.
-    (b) Simulation of stochastic 1/$f$ flux noise for
-    successive gate applications. The cumulative
-    gate error is computed after each gate application.
+    (a) $X/2$ gates robust to flux amplitude offsets constructed with the analytic (A),
+    sampling (S), unscented sampling (SU), and the 1\textsuperscript{st}-
+    and 2\textsuperscript{nd}-order derivative methods (D1, D2).
+    (b) Cumulative gate error due to 1/$f$ flux noise for
+    successive gate applications.
   }

### 2021-01-14 paper/a1longitude.tex
added constraints to opt statement

@@
-\section{Longitudinal Relaxation \label{appendix:longitude}}
-We comment on the longitudinal relaxation metrics and then give
+\section{Depolarization \label{appendix:longitude}}
+We comment on the depolarization metrics and then give
 our procedure for integrating the Lindblad master equation.
-The longitudinal relaxation probability and the gate error due to longitudinal
-relaxation metrics are compared in Table \ref{tab:longitude} for the
+The depolarization probability and the gate error due to
+depolarization metrics are compared in Table \ref{tab:longitude} for the
 numerical experiment described in Section \ref{sec:longitude}.
@@
 
-\begin{table}[h!]
+\begin{table}[ht]
   \begin{tabular}{c | c | c | c | c | c | c}
@@
   \caption{
-    Single gate longitudinal relaxation probability (P) ratios
-    and single gate error due to longitudinal relaxation (GE) ratios.
+    Single gate depolarization probability (P) ratios
+    and single gate error due to depolarization (GE) ratios.
     Values are reported for the analytic (A) and numerical (N) techniques.
@@
 
-To compute the gate error due to longitudinal relaxation
+To compute the gate error due to depolarization
 we employ the Lindblad master equation. This equation takes the form:
@@
 anti-commutator.
-For longitudinal relaxation $\gamma_{\uparrow} = T_{1, \uparrow}^{-1}$,
+For depolarization $\gamma_{\uparrow} = T_{1, \uparrow}^{-1}$,
 $\gamma_{\downarrow} = T_{1, \downarrow}^{-1}$,

### 2021-01-14 src/spin/figures.py
updated f1c alpha

@@ -397,7 +397,7 @@ F1C_GT_LS = {
     GateType.xpiby2: "dashdot",
 }
 F1C_LB = 1.1
-F1C_ALPHA = np.array([[1, 0.4], [1, 0.4], [1, 1]])
+F1C_ALPHA = np.array([[1, 1], [1, 1], [1, 1]])
 def make_figure1c():
     """
     Refs:

### 2021-01-14 paper/4longitude.tex
added constraints to opt statement

@@
-%% F1
-\begin{figure*}[ht]
+\section{Depolarization Awareness \label{sec:longitude}}
+The strength of depolarization varies
+with the control parameters in a range of superconducting circuit platforms,
+including the fluxonium. In this section, we outline a method
+for tuning the flux amplitude to mitigate depolarization.
+Previous work has modeled the gate error due to depolarization
+by evolving density matrices under a master
+equation \cite{rembold2020introduction, schulteherbruggen2011optimal}
+or evolving a large number of states in a quantum trajectory approach
+\cite{abdelhafez2019gradient}.
+We avoid the increase in computational complexity required for these
+techniques by
+penalizing the integrated rate (probability) of depolarization in optimization.
+Using this probability as a proxy for the gate error incurred
+is reasonable because losses due to depolarization increase monotonically
+in time. Additionally, for a constant $T_{1}$ time--the $1/e$ depolarization
+time for the qubit--a shorter gate duration
+would favor a lower depolarization probability. We allow
+the optimizer to tune the gate time to minimize the
+depolarization probability.
 
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-    \caption{\label{fig:longitudea}}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-    \caption{\label{fig:longitudeb}}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-    \caption{\label{fig:longitudec}}
-  \end{subfigure}
-  \caption{
-    (a) Numerically optimized gates (dark) and analytically optimized gates (light).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
-    (c) Lindblad master equation simulation with $T_{1}$ dissipation
-    for successive gate applications. The cumulative
-    gate error is computed after each gate application.
-    The numerically optimized $Z/2$ and $Y/2$ gate errors are indistinguishable.
-  }
-  \label{fig:longitude}
-\end{figure*}
-
-\section{Longitudinal Relaxation Awareness \label{sec:longitude}}
-The strength of longitudinal relaxation type decoherence varies
-with the control parameters in a range of superconducting circuit platforms,
-including the fluxonium qubit. Because dissipation to the thermal bath
-via longitudinal relaxation is an irreversible process that results in information loss,
-it is advantageous to tune the controls to extend the longitudinal
-relaxation time $T_{1}$--which represents the $1/e$ relaxation time
-of the quantum state.
-Typical approaches completely model the gate error due to longitudinal
-relaxation by performing optimal control on $n^{2}$
-density matrices of size $n \times n$ that evolve under a master
-equation, rather than $n$ state vectors of size $n$
-under the TDSE \eqref{eq:tdse}.
-We avoid this increase in computational complexity by
-penalizing the probability (integrated rate) of longitudinal relaxation.
-Using this probability as a proxy for the gate error incurred
-is reasonable because losses due to longitudinal relaxation increase monotonically
-in time.
-This technique can be extended to
-error channels which share the monotonically increasing property.
-Additionally, for a constant $T_{1}$ time, a shorter gate time
-would favor a lower longitudinal relaxation probability. We allow
-the optimizer to tune the gate time to minimize the
-longitudinal relaxation probability. Our scheme for time-optimal
-control is suitable for any time-optimal problem, not only
-the one we study here.
-
-\todo{Ask Brian to review this paragraph and next. Is the
-optimization setup clear?}
-The longitudinal relaxation probability is given by,
+The depolarization probability is given by,
 \begin{equation}
@@
 \end{equation}
-This value is appended to the augmented state vector \eqref{eq:astatecontrols}
-and penalized according to \eqref{eq:costfun}.
+This value is appended to the augmented state \eqref{eq:astatecontrols}
+and its norm is penalized in the objective \eqref{eq:costfun} by setting
+the corresponding element of the final augmented state $x_{f}$ to $0$.
 $T_{1}(a_{k})$ is obtained at each knot point by evaluating
@@
 optimization problem. The square root of the time step $\sqrt{\Delta t_{k}}$
-is appended to the augmented control vector \eqref{eq:astatecontrols}
+is appended to the augmented control \eqref{eq:astatecontrols}
 and the squared root
 of the time step $\lvert \Delta t_{k} \rvert$ is used
-for integration in the discrete dynamics function \eqref{eq:dyn_con}.
+for numerical integration in the discrete dynamics function \eqref{eq:dyn_con}.
 To ensure numerical
-integration accuracy is maintained we add a bound
-constraint on the time step at each knot point. \todo{(Brian) the setup seems clear to me.}
+integration accuracy is maintained, we constrain
+the bounds of the time step at each knot point.
 
-%% F2
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2a.png}
-    \caption{\label{fig:statica}}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2b.png}
-    \caption{\label{fig:staticb}}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f2c.png}
-    \caption{\label{fig:staticc}}
-  \end{subfigure}
-  \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic, sampling, unscented sampling, and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods. The gates
-    shown for the numerical methods are the solutions at twice the analytic
-    gate time.
-    (b) Single gate error as
-    a function of the gate time at a one-percent
-    qubit frequency detuning for all methods. Missing
-    data points represent solutions with a gate error above $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the qubit frequency detuning.
-    The solutions for the analytic and
-    1\textsuperscript{st}-order derivative methods are shown at multiples
-    of the analytic gate time. The performance for the two methods is
-    indistinguishable at the analytic gate time $18$ns.
-  }
-  \label{fig:static}
-\end{figure*}
-
-We compare the numerical method we have developed to the analytic gates
-on the task of achieving low gate errors in the presence of longitudinal relaxation
-for the $Z/2$, $Y/2$, and $X/2$ gates.
-The numerically optimized gates converge on similar solutions, a periodic
-waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}.
-They extend their gate times
-beyond their analytic counterparts, trading longer gate times for access
-to higher amplitudes and therefore higher $T_{1}$ times. All numerical gates reduce
-their single gate errors by a factor of $5$ over
-their analytic counterparts, commensurate to their
-probability of longitudinal relaxation reductions, see Appendix \ref{appendix:longitude}.
+We compare the performance of the $X/2$, $Y/2$, and $Z/2$ gates obtained with
+our numerical method to the analytic gates. We
+use the Lindblad master equation to simulate $T_{1}$ dissipation for successive
+gate applications, and compute the cumulative gate error
+after each application, see Appendix \ref{appendix:longitude}.
+The flux amplitudes for the numerical gates are similar; the
+waveforms are approximately periodic
+with amplitudes $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}. \todo{Floquet}.
+Their gate times are greater
+than their analytic counterparts, but they
+reach higher amplitudes and therefore higher $T_{1}$ times.
+Although the analytic and numerical gates attain single gate errors sufficient for
+quantum error correction ($< 10^{-4}$), which are required for fault-tolerant quantum computing
+\cite{aharonov2008fault, fowler2009high, gottesman1997stabilizer},
+the single gate errors for the numerical
+gates are a factor of $5$ less
+than those for the analytic gates, commensurate to the
+depolarization probability reductions, see Appendix \ref{appendix:longitude}.
 The gate error reported in this text is the infidelity
-of the evolved state and the final state averaged over 1000 pseudo-randomly
-generated initial states. The numerical $Z/2$ and $Y/2$ gates perform similarly in
-the concatenated gate application comparison, suppressing cumulative gate errors to $8 \cdot 10^{-3}$
+of the evolved state and the target state averaged over 1000 pseudo-randomly
+generated initial states.
+The numerical $Z/2$ and $Y/2$ gates produce similar
+results in the concatenated gate application comparison,
+yielding cumulative gate errors below $8 \cdot 10^{-3}$
 over $2000$ gate applications $\sim 40 \mu\textrm{s}$, see Figure \ref{fig:longitudec}.
-The numerical $X/2$ gate achieves a cumulative gate error of
+The cumulative gate error for the numerical $X/2$ gate is
 $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-These low cumulative gate errors for high gate counts are critical for
+The low cumulative gate errors for high gate counts produced
+by the numerical gates are critical for
 noisy, intermediate-scale quantum (NISQ) applications.
-Both the analytic and numerical gates attain single gate errors sufficient for
-quantum error correction ($< 10^{-4}$), which are required for fault-tolerant quantum computing
-\cite{aharonov2008fault, fowler2009high, gottesman1997stabilizer}.
 These improvements are significant for the realistic constraints we have imposed

### 2021-01-14 paper/8ack.tex

@@
   The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss, Taylor Howell, and Tanay Roy for useful discussions. \todo{cite software packages}.
+  and Taylor Howell, Jens Koch, Tanay Roy, Colm Ryan, and Daniel Weiss for useful discussions.
+  \todo{cite software packages}.
   \todo{funding}.

### 2021-01-14 paper/a2unscented.tex

@@
 We consider a state $\psi \in \mathbb{R}^{2n}$ with
-a deviant parameter $\lambda \in \mathbb{R}^{d}$ and dynamics
+a deviant parameter $\lambda \in \mathbb{R}^{d}$ and discrete dynamics
 $\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
@@
 is convenient for deriving the update procedure. A non-zero mean can be encoded
-in the dynamics.
+in the discrete dynamics.
 

### 2021-01-26 paper/0abstract.tex
revised based on dave comments

@@
 %% TITLE
-\title{Robust Quantum Optimal Control}
+\title{Robust Trajectory Optimization on a Quantum System}
 
@@
   and utilize time-optimal control to achieve high fidelity gates in the presence of depolarization.
-  We apply these techniques to a fluxonium qubit with realistic
-  parameters and constraints,
-  achieving orders of magnitude gate-error reductions from our baseline gate set in simulations.
+  We apply these techniques to a fluxonium qubit and suppress
+  simulated gate errors due to parameter uncertainty below $10^{-7}$ for
+  static parameter deviations on the order of $1\%$.
 \end{abstract}

### 2021-01-26 paper/1introduction.tex
revised based on dave comments

@@
   chakram2020multimode, fisher2010optimal, gokhale2019partial,
-  huang2020engineering, leng2019robust, leung2017speedup, li2020fast,
-  xu2020nonadiabatic},
+  leng2019robust, leung2017speedup, li2020fast},
 neutral atoms and ions \cite{brouzos2015quantum,
@@
 and Bose-Einstein condensates \cite{amri2019optimal, sorensen2018quantum}.
-For quantum computation,
+In the context of quantum computation,
 optimal control is employed to achieve high-fidelity gates
 while adhering to experimental constraints.
-Experimental errors may cause the system to deviate
+Experimental errors such as parameter drift, noise, and
+finite control resolution cause the system to deviate
 from the model used in optimization, leading
@@
 
-QOC has had tremendous success in
-engineering high-fidelity quantum gates \cite{
-  chou2015optimal, dolde2014high, egger2013optimized, egger2014adaptive,
-  grace2007optimal, heeres2017implementing, huang2014optimal,
-  kelly2014optimal, leng2019robust, liebermann2016optimal,
-  nebendahl2009optimal, rebentrost2009optimal, rebentrost2009optimal2,
-  spiteri2018quantum, sporl2007optimal}. 
-However, the system often
-deviates from the model used in optimization
-due to experimental errors such as parameter drift, noise, finite control 
-resolution, and decoherence. Multiple techniques have been
-developed to address this shortcoming.
-Analytically-derived control pulses that
-mitigate parameter uncertainty errors include
-composite pulses \cite{cummins2000use, cummins2003tackling,
+Analytically-dervied control pulses that mitigate parameter uncertainty
+errors include composite pulses \cite{cummins2000use, cummins2003tackling,
   kupce1995stretched, merrill2014progress},
@@
 pulses obtained with the DRAG scheme \cite{motzoi2009simple}.
-To suppress decoherence, Floquet techniques have been employed
-\cite{huang2020engineering, mundada2020floquet}.
-Numerical schemes to mitigate parameter uncertainty errors
-include closed-loop methods \cite{egger2014adaptive, feng2018gradient,
+These techniques are not amenable to arbitrary experimental constraints,
+so recent work has sought to achieve robustness in quantum optimal
+control frameworks using closed-loop methods \cite{egger2014adaptive, feng2018gradient,
   wittler2020integrated} and open-loop methods \cite{
@@
   rembold2020introduction, kosut2013robust}.
-To model decoherence, numerical techniques typically employ
-master equations \cite{rembold2020introduction, schulteherbruggen2011optimal} or
-Monte Carlo-style quantum trajectories \cite{abdelhafez2019gradient}.
 
@@
 \item A sampling method, similar to the work of \cite{allen2019robust,
-    carvalho2020error, khaneja2005optimal, reinhold2019controlling, rembold2020introduction}.
+  ball2020software, carvalho2020error, khaneja2005optimal,
+  reinhold2019controlling, rembold2020introduction}.
   \item An unscented sampling method adapted from the unscented transform used in
     state estimation \cite{howell2020direct, julier2004unscented,
-      lee2013sigma, manchester2016derivative}.
+      lee2013sigma, thangavel2020robust}.
   \item A derivative method, which penalizes the sensitivity of the quantum state trajectory
@@
 we perform time-optimal control and
-employ an efficient depolarization model that 
-is significantly less expensive to compute than integrating a master equation.
+employ an efficient depolarization model
+for which the computational cost is independent of the
+Hilbert space dimension.
 Leveraging recent advances in trajectory optimization within the field of robotics, we
 solve these optimization problems using ALTRO (Augmented Lagrangian TRajectory Optimizer)
-\cite{howell2019altro}, which uses the iterative linear quadratic regulator (iLQR) method
-\cite{Li2004a} within an augmented Lagrangian framework \cite{lantoine2012hybrid,
-  plancher2017constrained}
-to handle nonlinear equality and inequality constraints at 
-each time step.
+\cite{howell2019altro}, which can enforce constraints on the quantum state trajectory
+and control parameters.
 
 This paper is organized as follows.
-First, we reintroduce ALTRO in the context of QOC
+First, we describe ALTRO in the context of QOC
 in Section \ref{sec:background}.
-We describe realistic constraints for operating the fluxonium and
+We outline realistic constraints for operating the fluxonium and
 define the associated QOC problem in Section \ref{sec:fluxonium}.
-Then, we outline a method for making the optimization aware of depolarization
-in Section \ref{sec:longitude}. Next, we outline three techniques for achieving
+Then, we formulate a method for supressing depolarization
+in Section \ref{sec:longitude}. Next, we describe three techniques for achieving
 robustness to static parameter uncertainties in Section \ref{sec:static}. We

### 2021-01-26 paper/2background.tex
revised based on dave comments

@@
 review the QOC problem statement
-and reintroduce the ALTRO solver \cite{howell2019altro}.
+and describe the ALTRO solver \cite{howell2019altro}.
 QOC concerns a vector of time-dependent control parameters $u(t)$
@@
 \begin{equation}
-  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(u(t), t) \ket{\psi(t)}
+  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(u(t), t) \ket{\psi(t)}.
   \label{eq:tdse}
@@
 In general, $J(u)$ is a linear combination of cost functions on the state
-as well as cost functions on the controls \cite{leung2017speedup}.
+as well as cost functions on the controls.
 Standard QOC solvers compute derivatives of the functional $\nabla_{u} J(u)$,
-which are used to implement first-order optimization methods
+which can easily be used to implement first-order optimization methods
 \cite{machnes2015tunable, khaneja2005optimal, leung2017speedup, goerz2019krotov}.
@@
 and $u_{k}$ is the augmented control vector.
-The augmented state includes all variables of interest that are not decision variables.
-In the case of QOC, the state is a subset of the
-augmented state $\ket{\psi_{k}} \subseteq x_{k}$.
-The solver we use does not currently support complex numbers, so we employ the isomorphism
-$\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
-given in \cite{leung2017speedup},
-\begin{equation}
-  H \ket{\psi} \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
-    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
-  \begin{bmatrix} \ket{\psi}_{\textrm{re}} \\ \ket{\psi}_{\textrm{im}}\end{bmatrix}
-  \label{eq:isomorphism}
-\end{equation}
+The augmented state contains all relevant variables
+that are dependent upon the controls, for example the state
+$\ket{\psi_{k}} \subseteq x_{k}$.
 The augmented state trajectory obeys the physics of the system if
@@
 integrating the TDSE \ref{eq:tdse} with Runge-Kutta methods \cite{jorgensen2011numerical}
-or exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}.
+or exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance,
+shillito2020fast}.
 
-Additional constraints on the augmented states and augmented controls
+Additional constraints on the augmented controls and augmented states
 are encoded in constraint functions.
@@
 If a constraint function takes a positive value, the
-value is called the constraint's violation and
-the constraint is said to be active.
+value is called the constraint's violation.
 The trajectory optimization problem can be stated concisely,
@@
     \addConstraint{g_k(x_k,u_k)}{\leq 0} \label{eq:ineqcon}
-    \addConstraint{h_k(x_k,u_k)}{=0} \label{eq:eqcon}
+    \addConstraint{h_k(x_k,u_k)}{=0} \label{eq:eqcon}.
 \end{mini!}
 
-Many techniques have been proposed for solving \eqref{eq:gcostfun}-\eqref{eq:eqcon}. Standard 
-methods include direction collocation \cite{Hargraves87} and differential-dynamic programming
-(DDP) \cite{Mayne1966a}. Recent state-of-the-art solvers, such as ALTRO,
-have combined principles from both of these approaches. ALTRO uses iLQR
+Standard techniques for solving \eqref{eq:gcostfun}-\eqref{eq:eqcon} typically
+fall into two categories: direct methods \cite{Hargraves87, kelly2017introduction}
+and indirect methods \cite{betts1998survey}. For indirect methods,
+the augmented controls are the decision variables--the
+variables the optimizer adjusts to solve the problem.
+The augmented states are obtained from the augmented controls using the discrete dynamics function,
+and are then used to evaluate derivatives of the cost functions \eqref{eq:gcostfun}.
+Then, the derivative information is employed to update the augmented controls.
+This is the approach taken by standard QOC solvers such as GOAT \cite{machnes2015tunable},
+GRAPE \cite{khaneja2005optimal, leung2017speedup}, and Krotov's method \cite{goerz2019krotov}.
+Conversely, direct methods treat both the augmented controls and the augmented states as decision
+variables. In addition to minimizing the cost functions, the optimizer uses derivative information
+for the discrete dynamics function
+to satisfy the dynamics constraint \eqref{eq:gdyncon} to a specified tolerance.
+In this sense, the TDSE \eqref{eq:tdse} is a constraint which may be violated
+for intermediate steps of the optimization, where the states need not be physical.
+The direct approach lends itself to a nonlinear program formulation, for which
+a variety of general purpose solvers exist \cite{gill2005snopt, wachter2006implementation}.
+Recent state-of-the-art solvers, such as ALTRO,
+have combined principles from both of these approaches. ALTRO uses an
+iterative linear-quadratic regulator (iLQR) algorithm
 \cite{Li2004a} as the internal solver of an augmented Lagrangian method (ALM)
@@
 
-iLQR simulates the augmented state forward
-using the discrete dynamics function 
-and computes derivatives of the cost functions
-to update the augmented controls. This procedure is known as
-indirect shooting, and iLQR shares this approach with GOAT \cite{machnes2015tunable},
-GRAPE \cite{khaneja2005optimal, leung2017speedup}, and Krotov's method \cite{goerz2019krotov}.
-iLQR differs from these solvers
-in its update procedure. iLQR computes the locally optimal update for the augmented
-controls using a dynamic programming algorithm, and a line search is performed
-in the direction of this update to ensure a decrease in the sum of the cost functions.
+iLQR is an indirect method
+for solving the dynamically constrained
+trajectory optimization problem \eqref{eq:gcostfun}-\eqref{eq:gdyncon},
+and its update procedure is based on the differential-dynamic-programming approach
+\cite{mayne1966a}.
+First, iLQR uses an initial guess for the augmented controls to obtain the augmented
+states with the discrete dynamics function.
+iLQR then constructs quadratic models for each cost function using
+their zeroth-, first- and second-order derivatvies in a Taylor expansion
+about the current augmented controls and augmented states.
+These models are used to derive a recurrence relation between knot points
+which gives the locally optimal update for the augmented controls.
+Finally, a line search \cite{zhang2006global}
+is performed in the direction of this update to ensure a
+decrease in the sum of the cost functions. This procedure
+is repeated until convergence.
 
-Indirect shooting solvers such as iLQR are popular
-for solving unconstrained trajectory optimization problems \eqref{eq:gcostfun}-\eqref{eq:gdyncon},
-but standard implementations have no ability to handle nonlinear equality and inequality
+Indirect solvers such as iLQR are popular
+because they are very computationally efficient and maintain
+high accuracy for the discrete dynamics throughout optimization.
+However, standard implementations have no ability to handle
+nonlinear equality and inequality
 constraints \eqref{eq:ineqcon}-\eqref{eq:eqcon}.
-One way to handle constraints is to employ a projected gradient
-method \cite{clarkson2010coresets, hauswirth2016projected}.
-However, within the indirect shooting framework,
-projected gradient methods can
-only handle constraints on the augmented controls,
+Projected gradient methods are a typical approach
+to handle constraints
+\cite{clarkson2010coresets, hauswirth2016projected, morzhin2019minimal,
+  nikolskii2007convergence}. Unfortunately, within the indirect framework,
+they can only be used for constraints on the augmented controls,
 not the augmented states.
-Another technique is to promote the constraint functions to cost functions \cite{leung2017speedup}.
-A simple implementation of this approach does not garauntee that the constraints
-are satisfied, as the solver trades
+Another technique, which is popular for QOC \cite{leung2017speedup},
+is to add the constraint functions to the objective
+\eqref{eq:gcostfun}. 
+This strategy does not guarantee that the constraints
+are satisfied as the solver trades
 minimization of the cost functions and constraint functions against each other.
-ALM promotes the constraint functions to cost functions, and adaptively
-adjusts a Lagrange multiplier estimate for each constraint function to ensure
-the constraints are satisfied. ALM considers a decision variable $z \in \R^{s}$ with $s$ given,
-a cost function $f(z) : \R^{s} \mapsto \R$, and a constraint function
-$c(z) : \R^{s} \mapsto \R^{p}$ encoding $p$ constraints.
-For ALTRO, $z$ is the vector concatenation of the augmented states and augmented controls across
-all knot points, $f(z)$ is the sum of the trajectory optimization
-cost functions \eqref{eq:gcostfun},
-and $c(z)$ is the vector concatenation of the trajectory
-optimization constraint functions \eqref{eq:ineqcon}-\eqref{eq:eqcon}.
-ALM successively solves unconstrained minimization 
-problems of the form:
-\begin{mini}[2]
-    {z} {f(z) + \lambda^T c(z) + \half c(z)^T I_\mu c(z)}{}{}
-    \label{opt:alm}
-\end{mini}
-$\lambda \in \R^p$ is a Lagrange multiplier estimate, and $I_\mu$ is a diagonal matrix
-of penalty weights, $\mu \in \R^{p}$, whose magnitudes depend on whether
-the corresponding constraint is active or inactive.
-After minimizing \eqref{opt:alm} using iLQR,
-the Lagrange multiplier estimates are updated according to,
-\begin{equation}
-  \lambda \gets \lambda + I_\mu c(z)
-  \label{eq:almupdate}
-\end{equation}
-the penalty weight is updated, and the process repeats until convergence.
+ALM remedies this issue by adaptively adjusting a Lagrange multiplier estimate
+for each constraint function to ensure the constraints are satisfied.
+ALM adds terms that are linear and quadratic in the constraint functions
+to the objective. The new objective is then minimized with
+iLQR. If the solution produced by iLQR does not satisfy the constraints,
+the prefactors for the constraint terms in the objective are increased
+and the procedure is repeated.
 
@@
 to small decreases in the constraint violations near the locally optimal solution
-\cite{bertsekas1996constrained}. To address this shortcoming, ALTRO
-projects the solution from the ALM solving stage onto the constraint manifold using
+\cite{bertsekas1996constrained}.
+To address this shortcoming, ALTRO
+projects the solution from the ALM stage onto the constraint manifold using
 a projected Newton method, achieving ultra-low
-constraint violations $\sim 1 \cdot 10^{-8}$.
+constraint violations $\sim 10^{-8}$.
 For more information on the details of the ALTRO
@@
 As opposed to standard QOC solvers, ALTRO
-is able to satisfy constraints
-on both the augmented states and the augmented controls to tight tolerances.
+can satisfy constraints
+on both the augmented controls and the augmented states to tight tolerances.
 This advantage is crucial for this work, where multiple medium-priority cost functions

### 2021-01-26 paper/3fluxonium.tex
revised based on dave comments

--- 2021-01-12 paper/3fluxonium.tex
@@
 QOC on a classical computer inexpensive.
-In the two-level
-approximation, the Hamiltonian takes the form:
+To high accuracy, we approximate the Hamiltonian near the flux frustration
+as a spin-$1/2$ system:
 \begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
+  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2},
   \label{eq:hamiltonian}
@@
 where $f_{q}$ is the qubit frequency at the flux frustration point,
-$a(t)$ is the flux-drive amplitude, $h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
-are Pauli matrices. We optimize $X/2$, $Y/2$, and
-$Z/2$ gates for the fluxonium presented in \cite{zhang2020universal},
+$a(t)$ is the flux bias, $h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
+are Pauli matrices. Although we use an approximation for the Hamiltonian,
+our noise model is realistic and considers the full Hilbert space. We optimize $X/2$,
+$Y/2$, and $Z/2$ gates for the fluxonium presented in \cite{zhang2020universal},
 and compare them to the analytically constructed gates for
@@
 \eqref{eq:tstate_con}.
-%% This constraint is sensitive to global phase
-%% differences between the states.
-%% , but it is computationally efficient
-%% because the Jacobian and Hessian
-%% of the corresponding constraint function are diagonal.
-%% We could have analogously chosen a metric which is insensitive
-%% to global phases, such as the fidelity of the states.
 Furthermore, we impose the normalization constraint
@@
 \eqref{eq:statenorm_con}
-to ensure the solver does not take advantge of discretization errors in numerical integration.
-To refer to the discrete moments the flux amplitude, we introduce the notation
+to ensure the solver does not take advantage of discretization errors in numerical integration.
+To refer to the discrete moments the flux, we introduce the notation
 $\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
-$a(t) \lvert_{t = t_{k}} \equiv a_{k}$,
+$a(t = t_{k}) \equiv a_{k}$,
 $\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$.
@@
 \eqref{eq:znf_con}
-which mitigates the hysteresis ubiquitous in flux-bias lines
+which mitigates the inductive drift ubiquitous in flux-bias lines
 \cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
-The flux amplitude is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
+The flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
 \eqref{eq:amp_con}
@@
   \caption{
-    (a) Numerically optimized gates (dark blue)
-    and analytically optimized gates (light pink).
-    (b) $T_{1}$ interpolation function used in optimization. Markers
-    denote the time-averaged, absolute amplitude of each gate.
+    (a) Flux pulses for the numerical gates (dark blue)
+    and the analytic gates (light pink).
+    (b) $T_{1}$ interpolation function used in optimization. Circle markers
+    indicate measured $T_{1}$ times. Non-circle markers
+    indicate the time-averaged, absolute amplitude of each flux pulse.
     (c) Cumulative gate errors due to depolarization as a function of the
     number of gates applied.
-    Gate errors for the numerically optimized $Z/2$ and $Y/2$ gates
-    are indistinguishable. Inset shows the cumulative gate errors
-    plotted on a log scale for small gate counts.
+    Cumulative gate errors for the numerical $Z/2$ and $Y/2$ gates
+    are indistinguishable. Inset shows log-scaled cumulative gate errors
+    for small gate counts.
   }
@@
 
-Next, we outline the cost functions for the fluxonium gate problem.
-The augmented state and augmented control are:
+Next, we introduce the augmented control and augmented state:
 \begin{equation}
@@
   \renewcommand*{\arraystretch}{1.3}
+  u_{k} = \begin{bmatrix} \partial^{2}_{t} a_{k} \end{bmatrix}, \quad
   x_{k} = \begin{bmatrix} \ket{\psi^{0}_{k}} \\ \ket{\psi^{1}_{k}}
-    \\ \int_{t} a_{k} \\ a_{k} \\ \partial_{t} a_{k} \end{bmatrix} \quad
-  u_{k} = \begin{bmatrix} \partial^{2}_{t} a_{k} \end{bmatrix}
+    \\ \int_{t} a_{k} \\ a_{k} \\ \partial_{t} a_{k} \end{bmatrix}.
   \endgroup
@@
 \end{equation}
-The cost function at each knot point is
-$\ell_{k}(x_{k}, u_{k}) = (x_{k} - x_{f})^{T} Q_{k} (x_{k} - x_{f}) + u^{T}_{k} R_{k} u_{k}$
-where $Q_{k}$ and $R_{k}$ are matrices we supply, and
-$x_{f}$ is given by the constraints we have imposed on
-$\ket{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$ in addition to
-$\partial_{t} a_{f} = 0$. We choose $Q_{k}$ to be diagonal
-so the cost function incentivizes the achievement of the constraints on
-$\ket{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$,
-and penalizes the norm of $\partial_{t} a_{k}$ and $\partial^{2}_{t} a_{k}$,
-mitigating AWG ringing due to high-frequency transitions.
-In the discrete dynamics function for this problem \eqref{eq:dyn_con},
-we integrate the states according to the TDSE \eqref{eq:tdse} and the
+The variables in the augmented state are derived from
+the decision variable in the augmented
+control via coupled, first-order differential equations in the 
+the discrete dynamics function \eqref{eq:dyn_con}.
+We integrate the states according to the TDSE \eqref{eq:tdse} and the
 fluxonium Hamiltonian \eqref{eq:hamiltonian} and integrate
@@
 the first-derivative, proportional, and integral moments.
+The ALTRO implementation we use does not currently
+support complex numbers, so we repesent the states
+in the isomorphism $\mathcal{H}(\mathbb{C}^{n})
+\cong \mathcal{H}(\mathbb{R}^{2n})$ given in \cite{leung2017speedup},
+\begin{equation}
+  H \ket{\psi} \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
+    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
+  \begin{bmatrix} \ket{\psi}_{\textrm{re}} \\ \ket{\psi}_{\textrm{im}}\end{bmatrix}.
+  \label{eq:isomorphism}
+\end{equation}
+The cost function at each knot point is
+$\ell_{k}(x_{k}, u_{k}) = (x_{k} - x_{f})^{T} Q_{k} (x_{k} - x_{f}) + u^{T}_{k} R_{k} u_{k}$
+where $Q_{k}$ and $R_{k}$ are diagonal matrices we supply. The $Q_{k}$ term
+penalizes deviations from the final augmented state $x_{f}$,
+which is given by the constraints we have imposed on
+$\ket*{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$ in addition to
+$\partial_{t} a_{f} = 0$. The $R_{k}$ term penalizes the norm of $\partial^{2}_{t} a_{k}$,
+mitigating AWG ringing due to high-frequency transitions.
 Stated succinctly, the optimization problem takes the form:
@@
   \addConstraint{a_{1} = a_{N} = 0} \label{eq:bound_con}
-  \addConstraint{{\textstyle \int_{t}} a_{1} = \partial_{t} a_{1} = 0} \label{eq:ic_con}
+  \addConstraint{{\textstyle \int_{t}} a_{1} = \partial_{t} a_{1} = 0}. \label{eq:ic_con}
 \end{mini!}

### 2021-01-26 paper/4longitude.tex
revised based on dave comments

@@
-\section{Depolarization Awareness \label{sec:longitude}}
-The strength of depolarization varies
-with the control parameters in a range of superconducting circuit platforms,
-including the fluxonium. In this section, we outline a method
-for tuning the flux amplitude to mitigate depolarization.
-Previous work has modeled the gate error due to depolarization
+\section{Depolarization Mitigation\label{sec:longitude}}
+In this section, we outline a method
+for optimizing the flux to mitigate depolarization.
+For many superconducting circuits, the 1/$e$ depolarization time ($T_{1}$) is independent of the control parameters,
+so the fastest possible gate incurs the least depolarization error
+\cite{schulteherbruggen2011optimal}.
+For the fluxonium, however, $T_{1}$ is strongly dependent on the flux.
+We enable the optimizer to trade longer gate times
+for lower depolarizaiton times or vice-versa by making the gate time a decision variable.
+Additionally, previous work has modeled the gate error due to depolarization
 by evolving density matrices under a master
@@
 We avoid the increase in computational complexity required for these
-techniques by
-penalizing the integrated rate (probability) of depolarization in optimization.
+techniques by penalizing the integrated rate (probability) of depolarization in optimization.
 Using this probability as a proxy for the gate error incurred
-is reasonable because losses due to depolarization increase monotonically
-in time. Additionally, for a constant $T_{1}$ time--the $1/e$ depolarization
-time for the qubit--a shorter gate duration
-would favor a lower depolarization probability. We allow
-the optimizer to tune the gate time to minimize the
-depolarization probability.
+is reasonable because depolarization errors are incoherent and therefore
+increase monotonically in time without interference.
 
@@
 \begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
+  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}.
 \end{equation}
@@
 and its norm is penalized in the objective \eqref{eq:costfun} by setting
-the corresponding element of the final augmented state $x_{f}$ to $0$.
+the corresponding element of the final augmented state to zero.
 $T_{1}(a_{k})$ is obtained at each knot point by evaluating
@@
 It is also possible to fit a spline to theoretically obtained data.
-However, $T_{1}$ values are known to fluctuate greatly
-with laboratory temperatures \cite{klimov2018fluctuations}.
-Interpolating $T_{1}$ from experimental data
-increases the fridge truth of the simulation.
+However, experimental data often deviates from theoretical predictions
+because it is difficult to account for all loss channels \cite{zhang2020universal},
+and $T_{1}$ values often fluctuate on hour
+time scales \cite{klimov2018fluctuations}.
 
@@
 
-We compare the performance of the $X/2$, $Y/2$, and $Z/2$ gates obtained with
-our numerical method to the analytic gates. We
+We analyze the effect of depolarization on
+the $X/2$, $Y/2$, and $Z/2$ gates obtained with
+our numerical method and the analytic gates. We
 use the Lindblad master equation to simulate $T_{1}$ dissipation for successive
@@
 after each application, see Appendix \ref{appendix:longitude}.
-The flux amplitudes for the numerical gates are similar; the
-waveforms are approximately periodic
-with amplitudes $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}. \todo{Floquet}.
-Their gate times are greater
-than their analytic counterparts, but they
-reach higher amplitudes and therefore higher $T_{1}$ times.
-Although the analytic and numerical gates attain single gate errors sufficient for
-quantum error correction ($< 10^{-4}$), which are required for fault-tolerant quantum computing
-\cite{aharonov2008fault, fowler2009high, gottesman1997stabilizer},
-the single gate errors for the numerical
-gates are a factor of $5$ less
-than those for the analytic gates, commensurate to the
-depolarization probability reductions, see Appendix \ref{appendix:longitude}.
+\todo{Jens, Danny, Ziwen, any idea what the correct way to simulate depolarization is
+  when $T_{1}$ is time dependent?}.
 The gate error reported in this text is the infidelity
@@
 generated initial states.
-The numerical $Z/2$ and $Y/2$ gates produce similar
-results in the concatenated gate application comparison,
-yielding cumulative gate errors below $8 \cdot 10^{-3}$
-over $2000$ gate applications $\sim 40 \mu\textrm{s}$, see Figure \ref{fig:longitudec}.
-The cumulative gate error for the numerical $X/2$ gate is
-$1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
-The low cumulative gate errors for high gate counts produced
-by the numerical gates are critical for
-noisy, intermediate-scale quantum (NISQ) applications.
-These improvements are significant for the realistic constraints we have imposed
-on the gates, and do not represent a fundamental limit to the optimization methods we have
-employed.
+The flux pulses for the numerical gates
+are approximately periodic
+with amplitudes $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}.
+They are reminiscent of the analytically determined Floquet operations
+for a fluxonium described in \cite{huang2020engineering}
+and realized in \cite{mundada2020floquet}.
+The numerical gate times are greater
+than the analytic gate times, but the
+numerical flux pulses
+spend more time at higher amplitudes, achieving higher $T_{1}$ times.
+The single gate errors for both the analytic and numerical gates are
+less than $10^{-4}$, which makes them sufficient for quantum error correction--a
+prerequisite for fault-tolerant quantum
+computing \cite{aharonov2008fault, knill2005quantum, gottesman1997stabilizer}.
+However, the numerical gates achieve single gate errors
+$\sim 5$ times less than those for the analytic gates,
+which tracks closely with their relative improvement
+on the depolarization probability metric, see Appendix \ref{appendix:longitude}.
+This single gate error advantage corresponds to a significant reduction in error correction resources.
+Furthermore, for successive gate applications, the gate error due to depolarization is approximately linear
+in the gate count, which we expect for $t \ll T_{1}$, see Figure \ref{fig:longitudec}.
+The gate error reduction for large gate counts is important for noisy, intermediate-scale quantum (NISQ)
+applications. These improvements are significant for the constraints we have imposed on the gates,
+and do not represent a fundamental limit to the optimization methods we have employed.

### 2021-01-26 paper/5static.tex
revised based on dave comments

@@
   \caption{
-    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
-    analytic (A), sampling (S), unscented sampling (SU), and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods (D1, D2). The numerical gates shown
-    are optimized for twice the gate time of the analytic gate.
+    (a) Flux pulses for $Z/2$ gates robust to qubit frequency detunings constructed with the
+    analytic (A), sampling (S), unscented sampling (U), and the 1\textsuperscript{st}-
+    and 2\textsuperscript{nd}-order derivative methods (D1, D2). The flux pulses shown
+    for the sampling, unscented sampling, and derivative methods are optimized
+    for twice the gate time of the analytic gate.
     (b) Single gate error at a one-percent qubit frequency detuning as
     a function of the gate time. Missing
-    data points represent solutions with a gate error greater than $5 \cdot 10^{-5}$.
+    data points represent gates with a gate error greater than $5 \cdot 10^{-5}$.
     (c) Single gate error as a function of the qubit frequency detuning.
@@
     methods are shown for gate times which are multiples of $1 / 4 f_{q} \sim 18 \textrm{ns}$.
-    The performance of the two methods is
+    The gate errors for the two methods are
     indistinguishable at the gate time $18 \textrm{ns}$.
@@
 precision of the qubit frequency which modifies the fluxonium Hamiltonian
-\eqref{eq:hamiltonian} by $f_{q} \gets f_{q} + \delta f_{q}$.
+\eqref{eq:hamiltonian} by $f_{q} \rightarrow f_{q} + \delta f_{q}$.
 We consider three robust control techniques to accomplish this task:
-the sampling method, the unscented sampling method,
-and the derivative method.
+a sampling method, an unscented sampling method,
+and a derivative method.
 
 The sampling method incentivizes the optimizer
-to ensure multiple copies of a state, each of which evolves
+to ensure multiple instances of a state, each of which evolves
 with a distinct value of the uncertain parameter, achieve
-the same state transfer. For each initial state,
+the same target state. Variants of this technique have been proposed
+in the context of QOC
+\cite{allen2019robust, ball2020software, khaneja2005optimal,
+  reinhold2019controlling, rembold2020introduction} and applied
+experimentally \cite{carvalho2020error}.
+For each initial state,
 we add two sample states $\ket{\psi^{\pm}}$
-to the augmented state \eqref{eq:astatecontrols}, and the discrete dynamics
+to the augmented state \eqref{eq:astatecontrols}. The discrete dynamics
 function \eqref{eq:dyn_con} is modified
 so the sample states evolve under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
-with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$ for a fixed hyperparameter
-$\sigma_{f_{q}}$ which is the standard deviation of the qubit frequency.
+with $f_{q} \rightarrow f_{q} \pm \sigma_{f_{q}}$ for a fixed
+hyperparameter $\sigma_{f_{q}}$ which is the standard deviation of the qubit frequency.
+We penalize the infidelities of the sample states and their target state
+by adding a cost function to the objective \eqref{eq:costfun} of the form
+$\sum_{k, \pm} q_{k} (1 - {\lvert \braket{\psi^{\pm}_{k}}{U \psi_{1}} \rvert}^{2})$
+where $q_{k}$ is a constant we supply.
 For this method, the standard orthonormal basis states are an insufficient choice
@@
 will be robust to qubit frequency detunings for the initial states $\ket{0}$
-or $\ket{1}$ because the fidelity metric is insensitive to global phases,
+or $\ket{1}$ because the infidelity metric is insensitive to global phases,
 but this gate will not be robust for any other initial states.
-Thus, we choose four initial states
+Therefore, we choose four initial states
 so that their outer products
@@
 $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
-The infidelities between the sample states and the images of their
-initial states under the desired gate are penalized
-by adding a cost function
-$\sum_{k, \pm} q_{k} (1 - {\lvert \braket{\psi^{\pm}_{k}}{U \psi^{\pm}_{1}} \rvert}^{2})$
-to the objective \eqref{eq:costfun} for a constant $q_{k}$ we supply.
+(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized},
+which we refer to as the operator basis.
 
-For the unscented sampling method, we append $2(2n + d)$ sample
-states to the augmented state vector \eqref{eq:astatecontrols}
-for each initial state used in the sampling method. $2n = 4$ is twice the
-dimension of the Hilbert space, resulting from the isomorphism \eqref{eq:isomorphism},
-and $d = 1$ is the number of uncertain parameters. The sample states
-encode a unimodal distribution over the $2n$
-elements of the evolving initial state, modeling
-the uncertainty in the state due to the uncertainty in the parameter.
-The discrete dynamics function \eqref{eq:dyn_con} is modified
-so each sample state evolves under the fluxonium Hamiltonian
-\eqref{eq:hamiltonian} where $f_{q} \gets f_{q} + \delta f_{q}$
-and $\delta f_{q}$ is sampled at each knot point from a distribution
-with fixed standard deviation $\sigma_{f_{q}}$.
-After propagating the sample states to each knot point, we apply the unscented
-transformation to the states, which
-preserves the first and second moments of the distribution the states encode.
-This procedure makes the unscented sampling method more resistant to sample
-states collapsing towards the mean than in the sampling method.
-We add a contribution to the objective \eqref{eq:costfun} for the infidelity
-of the sample states and their target states
-as in the sampling method.
+Whereas the sampling method penalizes the deviations of the sample states
+from the target state, the unscented sampling method
+penalizes the deviations of the sample states from the nominal state
+\cite{howell2020direct, lee2013sigma,
+  thangavel2020robust}. Accordingly, the cost function we add
+to the objective \eqref{eq:costfun} takes the form
+$\sum_{k, j} q_{k} (\psi^{j}_{k} - \psi_{k})^{T}
+(\psi^{j}_{k} - \psi_{k})$, where $q_{k}$ is a
+constant we supply, $\psi_{k}$ is
+the evolved initial state (nominal state), and $\psi^{j}_{k}$ is a sample state
+that evolves under a modified Hamiltonian similar to that in the sampling method.
+We omit bra-ket notation here to emphasize that
+the states are real vectors, and are given by the right-hand-side of the
+isomorphism \eqref{eq:isomorphism}. Additionally,
+the unscented sampling method employs the unscented transform
+\cite{julier2004unscented, uhlmann1995dynamic}
+to prevent the sample states
+from drifting too far or close to the nominal state during evolution,
+which would result in exceedingly large or small deviations in the cost function.
+In particular, the sample states are chosen to encode a unimodal distribution over
+the $2n$ elements of the nominal state, modeling the uncertainty in the state
+as a result of the uncertainty in the parameter. The unscented transform
+accurately propagates the mean and covariance of this distribution between
+knot points, or equivalently, through the transformation of the TDSE.
+For this method, we sample on one initial state $(\ket{0} + i\ket{1}) / \sqrt{2}$,
+and do not observe a performance increase
+using more initial states, for example those from the operator basis.
 A detailed procedure for the unscented transformation is given
@@
 
-The derivative method penalizes the sensitivty of the state
+The derivative method penalizes the sensitivity of the state
 to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
@@
 to the augmented state vector \eqref{eq:astatecontrols}
-for each initial state used in the sampling method.
+for each initial state. We use the initial state $\ket{0}$ for this method,
+and observe no advantage to using more initial states.
 We penalize the norms of the state derivatives
 in the objective \eqref{eq:costfun} by setting the corresponding elements
-of the final augmented state $x_{f}$ to $0$.
+of the final augmented state to zero.
 We could obtain the state derivatives at each knot point
@@
 \begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
+  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi},\\
   i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
   H \ket{\partial_{f_{q}} \psi} +
-  (\partial_{f_{q}} H) \ket{\psi}
+  (\partial_{f_{q}} H) \ket{\psi}.
   \label{eq:d1dyn}
@@
 
-We compare the gate errors due to a static qubit frequency
-detuning for $Z/2$ gates obtained with the robust control techniques
-and the analytic method.
+We examine the gate errors due to a static qubit frequency
+detuning for the $Z/2$ gates obtained with the robust control techniques
+and the analytic $Z/2$ gate.
 To compute the gate error,
@@
 $f_{q} \rightarrow f_{q} \pm \delta f_{q}$
-at the stated qubit frequency detuning.
+at the stated qubit frequency detuning $\delta f_{q}$.
 The reported gate error is the infidelity of
 the evolved state and the target state averaged over
-the two transformations and $1000$ initial states.
+the two transformations for each of $1000$ pseudorandomly
+generated initial states.
 We set $\sigma_{f_{q}}/f_{q} = 1\%$
@@
 The gate's erroneous rotation angle
-$\pi \delta f_{q} / 2 f_{q}$ is linear in the
+$2 \pi \delta f_{q} / 4 f_{q}$ is linear in the
 qubit frequency detuning, resulting in a gate error that is quadratic
@@
 At a one-percent
-qubit frequency detuning ($\lvert \delta f_{q} \rvert / f_{q} = 1\%$),
+detuning ($\lvert \delta f_{q} \rvert / f_{q} = 1\%$),
 the gate error is $\sim 4.5 \cdot 10^{-5}$,
 which is sufficient for quantum error correction.
-Although the gate performs well, the method used to derive
-it can not be employed to achieve a $Z/2$ gate at gate times other than
+Although the gate performs well, it cannot be extended
+to gate times other than
 $1 / 4 f_{q}$. The ability to perform
@@
 
-The solution for the unscented sampling method combines idling periods
-with fast ramps to the maximum amplitude, whereas that for the sampling method
-does not reach high amplitudes.
+The flux pulse produced by the sampling method
+transitions  smoothly away from idles at the flux frustration point,
+whereas that for the unscented sampling method employs
+three distinct transitions.
 The gate error at a one-percent qubit frequency detuning for the sampling method
@@
 range of gate times. Conversely, the gate error at a one-percent detuning
-for the unscented sampling method decreases linearly in the gate time
-until half the Larmor period $1 / 2 f_{q}$ after which it has a consistent
-gate error $\sim 3.5 \cdot 10^{-5}$.
+for the unscented sampling method reaches a minimum of $\sim 3.9 \cdot 10^{-5}$
+near fractions of the Larmor period $2/4f_{q} \sim 36\textrm{ns},
+3/4f_{q} \sim 54\textrm{ns}, 4/4f_{q} \sim 72\textrm{ns}$.
 
-The derivative methods converge on qualitatively similar solutions that
-use fast triangle pulses at the boundaries and idle at low amplitudes.
+The two derivative methods converge on qualitatively similar flux pulses that
+use fast triangle pulses at the boundaries and idle near the flux frustration point,
+similar to the flux pulse produced by the unscented sampling method.
 The gate errors at a one-percent qubit frequency detuning
-for both methods decrease super-linearly in their gate times.
-The 1\textsuperscript{st}-order derivative method
-consistently achieves lower gate errors than the
-2\textsuperscript{nd}-order derivative method,
-which we believe is related to the greater importance
-of minimizing the 1\textsuperscript{st}-order state
-derivative norm than the 2\textsuperscript{nd}-order
-state derivative norm for this problem, see Appendix \ref{appendix:derivative}.
+for both derivative methods decrease super-linearly in their gate times.
+The 2\textsuperscript{nd}-order method does not offer a substantial improvement
+over the 1\textsuperscript{st}-order method for most gate times.
 The gate error at a one-percent detuning for the 1\textsuperscript{st}-order
-derivative method approaches zero at the Larmor period $1 / f_{q}$, see Figure \ref{fig:staticc}.
+method reaches $10^{-7}$ at the Larmor period $1 / f_{q} \sim 72$ns,
+see Figure \ref{fig:staticc}.
 This result mimics the
@@
   \caption{
-    (a) $X/2$ gates robust to flux amplitude offsets constructed with the analytic (A),
-    sampling (S), unscented sampling (SU), and the 1\textsuperscript{st}-
+    (a) Flux pulses for $X/2$ gates robust to flux offsets
+    constructed with the analytic (A),
+    sampling (S), unscented sampling (U), and the 1\textsuperscript{st}-
     and 2\textsuperscript{nd}-order derivative methods (D1, D2).
     (b) Cumulative gate error due to 1/$f$ flux noise for
-    successive gate applications.
+    successive gate applications. The cumulative gate errors for the
+    sampling, unscented sampling, and the derivative methods are indistinguishable.
   }

### 2021-01-26 paper/6stochastic.tex
revised based on dave comments

@@
 superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors \todo{coherent errors?}. Flux noise
+source of coherent errors. Flux noise
 modifies the fluxonium Hamiltonian \eqref{eq:hamiltonian}
 by $a(t) \rightarrow a(t) + \delta a(t)$.
-The spectral density of flux noise is known to
+The spectral density of the flux noise $\delta a(t)$ is observed to
 follow a 1/$f$ distribution
@@
 takes advantage of the low-frequency characteristic and
-treat the noise as quasi-static, performing a generalization of the spin-echo
-technique to compensate for erroneous drift.
+treats the noise as quasi-static, performing a generalization of the spin-echo
+technique to compensate for erroneous drift \cite{hahn1952spin, meiboom1958modified}.
 
@@
 In principle, we could modify the sampling method
-similarly; however, we choose to sample statically $a(t) \rightarrow a(t) + \sigma_{a}$
+similarly; however, we choose to subject the sample states
+to static noise
+$a(t) \rightarrow a(t) \pm \sigma_{a}$
 for comparison. The derivative methods require no algorithmic modification
@@
 
-We compare the gate errors due to 1/$f$ flux noise for $X/2$
-gates constructed with the robust control techniques
-and the analytic method. To compute the gate error,
+We analyze the gate errors due to 1/$f$ flux noise for
+the $X/2$ gates constructed with the robust control techniques
+and the analytic $X/2$ gate. To compute the gate error,
 an initial state is evolved
 under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
-where the optimized flux amplitude is modified $a(t) \rightarrow a(t) + \delta a(t)$.
-The flux noise $\delta a(t)$ is generated as
+where the optimized flux is modified $a(t) \rightarrow a(t) + \delta a(t)$.
+The flux noise is generated as
 we described for the unscented sampling method.
-Each of the $1000$ averages we perform uses a different random seed for
-generating the flux noise.
-We simulate successive applications of the gate constructed by each method
+The reported gate error is the infidelity
+averaged over $1000$ pseudorandomly generated initial states,
+each of which is subject to a distinct pseudorandomly
+generated flux noise instance.
+To observe the effect of interfering coherent errors,
+we simulate successive applications of the gate constructed by each method
 and compute the cumulative gate error
@@
 two orders of magnitude less than the gate error produced by the analytic gate.
-\todo{motional narrowing}.
-1/$f$ flux noise is a significiant source of coherent errors in NISQ applications and
+1/$f$ flux noise is a significant source of coherent errors in NISQ applications and
 these numerical techniques offer effective avenues to mitigate it.

### 2021-01-26 paper/7conclusion.tex
revised based on dave comments

--- 2021-01-12 paper/7conclusion.tex
@@
 errors on a quantum system.
-We have proposed a scheme for supressing
+We have proposed a scheme for suppressing
 depolarization with time-optimal
 control and the depolarization probability model.
-The computational cost of this depolarization model is
+The computational cost of this model is
 independent of the dimension of the Hilbert space, enabling
@@
 We have shown that the derivative, sampling, and unscented sampling methods
-can mitigate 1/$f$ flux noise errors
-which dominate coherent errors for flux controlled qubits.
-The robust control techniques can be applied
+can mitigate 1/$f$ flux noise errors--which
+dominate coherent errors for flux controlled qubits.
+These robust control techniques can be applied
 to any Hamiltonian,
-allowing experimentalists in all disciplines to engineer robust
+allowing experimentalists in all domains to engineer robust
 operations on their quantum systems.
@@
 required for fault-tolerant quantum computing applications. Our
-implementation of the techniques described in this work is available
-at \url{https://github.com/SchusterLab/rbqoc}.
+implementations of the techniques described in this work are available
+at \url{https://github.com/SchusterLab/rbqoc}. \todo{static code?}

### 2021-01-26 paper/8ack.tex
revised based on dave comments

--- 2020-11-28 paper/8ack.tex
@@
 \begin{acknowledgments}
-  The authors would like to thank Helin Zhang for experimental assistance
-  and Daniel Weiss, Taylor Howell, and Tanay Roy for useful discussions. \todo{cite software packages}.
+  We thank Helin Zhang for experimental assistance
+  and Taylor Howell, Jens Koch, Tanay Roy, Colm Ryan, and Daniel Weiss for useful discussions.
+  This work is funded in part by EPiQC, an NSF Expedition in Computing, under grant CCF-1730449.
+  This work was supported by the Army Research Office under Grant No. W911NF1910016.
   \todo{funding}.
+  \todo{cite software packages}.
 \end{acknowledgments}

### 2021-01-26 paper/a1longitude.tex
revised based on dave comments

@@
   \caption{
-    Single gate depolarization probability (P) ratios
-    and single gate error due to depolarization (GE) ratios.
-    Values are reported for the analytic (A) and numerical (N) techniques.
+    Single gate depolarization probability (P)
+    and single gate error due to depolarization (GE).
+    Values are reported for the analytic (A) and numerical (N) gates.
   }
@@
 
-To compute the gate error due to depolarization
-we employ the Lindblad master equation. This equation takes the form:
+We employ the Lindblad master equation
+to compute the gate error due to depolarization.
+This equation takes the form:
 \begin{equation}
@@
   + \sum_{i = 1}^{n^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger}
-  - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
+  - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\}),
 \end{equation}
@@
 by driving the qubit at the desired flux amplitude
-and monitoring the resultant decay.
-For more details consult \cite{zhang2020universal}.
+and monitoring the resultant decay. For more details
+on these measurements, consult \cite{zhang2020universal}.
 
-Exponential integrators can be employed to integrate the Lindblad master equation
-using the Vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018},
+So that we may use exponential integrators, we employ
+the Vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018},
 \begin{equation}
-  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}} \textrm{vec}({\rho})
+  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}} \textrm{vec}({\rho}),
 \end{equation}
@@
     (L_{i}^{*} \otimes L_{i} - \frac{1}{2} (I \otimes L_{i}^{\dagger}L_{i}
-    - L_{i}^{T}L_{i}^{*} \otimes I))
+    - L_{i}^{T}L_{i}^{*} \otimes I)),
   \end{aligned}
@@
 We use zero-order hold on the controls so the exact solution is
-$\textrm{vec}(\rho_{k + 1}) = \exp(\Delta t_{k} \hat{\mathcal{L}}_{k}) \textrm{vec}(\rho_{k})$.
+$\textrm{vec}(\rho_{k + 1}) = {\exp}{\textstyle(}\Delta t_{k} \hat{\mathcal{L}}_{k}{\textstyle)} \textrm{vec}(\rho_{k})$.
 This isomorphism transforms $(n^{2} \times n^{2}) \times (n^{2} \times n^{2})$

### 2021-01-26 paper/a2unscented.tex
revised based on dave comments

--- 2020-11-30 paper/a2unscented.tex
@@
 In this section, we outline the full unscented sampling procedure.
-We consider a state $\psi \in \mathbb{R}^{2n}$ with
-a deviant parameter $\lambda \in \mathbb{R}^{d}$ and dynamics
+In the unscented sampling method, we consider a state $\psi \in \mathbb{R}^{2n}$, 
+an uncertain parameter $\lambda \in \mathbb{R}^{d}$, and discrete dynamics
 $\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
-The nominal initial state is given by $\psi_{0}$ with an associated
-positive-definite covariance matrix $P_{0} \in \mathbb{R}_{++}^{2n \times 2n}$
+The state is obtained with an isomorphism \eqref{eq:isomorphism} and
+we omit bra-ket notation to emphasize that the state is real.
+The nominal initial state is given by $\bar{\psi}_{1}$ with an associated
+positive-definite covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$
 which describes the uncertainty in the initial state.
-$P_{0}$ is typically non-zero even if the state preparation error is negligible.
-The deviant parameter
-has zero-mean and its distribution is given by the covariance matrix
-$L_{k} \in \mathbb{R}_{++}^{d \times d}$ at knot point $k$. The zero mean assumption
+$P_{1}$ is typically non-zero even if the state preparation error is negligible.
+The uncertain parameter has zero-mean
+and its distribution is given by the covariance matrix
+$L_{k} \in \mathbb{S}_{++}^{d}$ at knot point $k$. The zero-mean assumption
 is convenient for deriving the update procedure. A non-zero mean can be encoded
-in the dynamics.
+in the discrete dynamics function $f(\psi_{k}, \lambda_{k})$.
 
-The initial $4n + 2d$ sigma points and initial $4n + 2d$ deviant parameters are sampled
-from the initial distributions,
+The initial $4n + 2d$ sample states and initial $4n + 2d$
+uncertain parameters are sampled from the initial distributions,
 \begin{equation}\label{eq:uupdate}
-  \begin{bmatrix} \Psi_{0}^{i} \\ \Lambda_{0}^{i} \end{bmatrix} =
-  \begin{bmatrix} \bar{\Psi}_{0} \\ 0\end{bmatrix}
-    \pm \beta \sqrt{\begin{bmatrix} P_{0} & 0\\ 0 & L_{0}\end{bmatrix}}^{\; i}
+  \begin{bmatrix} \psi_{1}^{j} \\ \lambda_{1}^{j} \end{bmatrix} =
+  \begin{bmatrix} \bar{\psi}_{1} \\ 0\end{bmatrix}
+    \pm \beta \sqrt{\begin{bmatrix} P_{1} & 0\\ 0 & L_{1}\end{bmatrix}}^{\; j}.
 \end{equation}
-We have written $\bar{\Psi}_{0} = \psi_{0}$. $\beta$
-is a hyperparameter which controls the spacing of the covariance contour.
-The $(\pm)$ is understood to take $(+)$ for $ i \in \{1, \dots, 2n + d\}$ and $(-)$ for
-$i \in \{2n + d + 1, \dots, 4n + 2d\}$. We use the Cholesky factorization
+$\beta$ is a hyperparameter that controls the spacing of the covariance contour.
+The $(\pm)$ is understood to take $(+)$ for $j \in \{1, \dots, 2n + d\}$ and $(-)$ for
+$j \in \{2n + d + 1, \dots, 4n + 2d\}$. We use the Cholesky factorization
 to compute the square root of the
@@
 column (mod $2n + d$) of the lower triangular Cholesky factor.
-Then, the sigma points are normalized,
+Then, the sample states are normalized,
 \begin{equation}\label{eq:unormalize}
-  \Psi_{0}^{i} \gets \frac{\Psi_{0}^{i}}{\sqrt{\Psi_{0}^{i^{T}} \Psi_{0}^{i}}}
+  \psi_{1}^{j} \rightarrow \frac{\psi_{1}^{j}}{\sqrt{{\psi_{1}^{j}}^{T} \psi_{1}^{j}}}.
 \end{equation}
-The sigma points are propagated to the next knot point,
+The sample states are propagated to the next knot point,
 \begin{equation}\label{eq:upropagate}
-  \Psi_{1} = f(\Psi_{0}, \Lambda_{0})
+  \psi^{j}_{2} = f(\psi^{j}_{1}, \lambda^{j}_{1}).
 \end{equation}
-The mean and covariance of the sigma points are computed,
+The mean and covariance of the sample states are computed,
 \begin{align}
-  \bar{\Psi}_{1} &= \frac{1}{4n + 2d} \sum_{i = 1}^{4n + 2d} \Psi_{1}^{i}\\
-  P_{1} &= \frac{1}{2 \beta^{2}} \sum_{i = 1}^{4n + 2d}
-  (\Psi^{i}_{1} - \bar{\Psi}_{1})^{T}(\Psi^{i}_{1} - \bar{\Psi}_{1})
+  \bar{\psi}_{2} &= \frac{1}{4n + 2d} \sum_{j = 1}^{4n + 2d} \psi_{2}^{j},\\
+  P_{2} &= \frac{1}{2 \beta^{2}} \sum_{j = 1}^{4n + 2d}
+  (\psi^{j}_{2} - \bar{\psi}_{2})(\psi^{j}_{2} - \bar{\psi}_{2})^{T}.
 \end{align}
-The sigma points are then resampled and propagated to the next knot point using
+The sample states are then resampled and propagated to the next knot point using
 \eqref{eq:uupdate}, \eqref{eq:unormalize}, and \eqref{eq:upropagate}. Our
-choice of sigma points follows the prescription in equation 11 of \cite{julier2004unscented}.
+choice of sample states (sigma points) follows equation 11 of
+\cite{julier2004unscented}.
 Prescriptions that require fewer sigma points exist \cite{julier2002reduced}.

### 2021-01-26 paper/a3derivative.tex
majority of jens and colm comments adressed
revised based on dave comments

@@
 \section{Derivative Method \label{appendix:derivative}}
-We comment on the optimization metrics of the derivative methods and then
-outline how to efficiently integrate their dynamics.
-The 1\textsuperscript{st}-order derivative method tends
-to outperform the 2\textsuperscript{nd}-order
-derivative method in the small, static detuning regime, see Figure \ref{fig:staticb}.
-The norms of the state derivatives for the 1\textsuperscript{st}- and 2\textsuperscript{nd}-order
-methods are provided in Table \ref{tab:dnorm} for the solutions to
-the numerical experiment detailed in Section \ref{sec:static} at $t_{N} = 60$ns.
-The 2\textsuperscript{nd}-order method is able to decrease the 2\textsuperscript{nd}-order
-state derivate norm relative to the 1\textsuperscript{st}-order method at the expense of increasing
-its 1\textsuperscript{st}-order derivative norm. In the problem
-we have studied it is likely the 2\textsuperscript{nd}-order derivative norm has a smaller
-contribution to the gate error than the 1\textsuperscript{st}-order derivative norm. A
-careful analysis of contributions at each order could be used to predict the efficacy of the
-derivative method for future problems.
-
-\begin{table}[ht]
-  \begin{tabular}{c | c | c}
-    Method & $\norm{\partial_{f_{q}} \psi_{N}}_{2}$ ($10^{3}$)
-    & $\norm*{\partial^{2}_{f_{q}} \psi_{N}}_{2}$ ($10^{6}$)\\
-    \hline
-    D1 & 0.436 & 57.817\\
-    D2 & 1.702 & 9.030\\
-  \end{tabular}
-  \caption{Norm of state derivates with respect to the qubit frequency
-    for $Z/2$ gates optimized using the derivative methods. The norms are computed
-    at the end of the gate $t_{N} = 60$ns
-    and are averaged over the four state derivatives.}
-  \label{tab:dnorm}
-\end{table}
-
-The dynamics for the derivative methods can be integrated efficiently using exponential integrators
-\cite{auer2018magnus, berland2005solving, einkemmer2017performance}.
+Here, we outline how to efficiently integrate the dynamics
+of the derivative method using exponential integrators.
 General exponential integrators break the dynamics into a linear term and a non-linear term.
@@
     &+ \int_{0}^{\Delta t_{k}} \exp((\Delta t_{k} - t^{'})L_{k})
-    N(t_{k} + t^{'}) dt^{'}\\
+    N(t_{k} + t^{'}) dt^{'}.\\
   \end{aligned}
@@
     \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(\Delta t_{k}L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
-    &+ \exp(\Delta t_{k}L_{k}) N_{k}\\
+    &+ \exp(\Delta t_{k}L_{k}) N_{k} \Delta t_{k}.\\
   \end{aligned}

### 2021-01-26 paper/a4time.tex
majority of jens and colm comments adressed
revised based on dave comments

--- 2020-12-08 paper/a4time.tex
@@
 the base optimization in Section \ref{sec:fluxonium},
-the longitudinally aware optimization in Section \ref{sec:longitude},
+the depolarization optimization in Section \ref{sec:longitude},
 and the robust optimizations in Section \ref{sec:static}
 are presented in Table \ref{tab:time}
-for a $Z/2$ gate at multiples of the analytic $Z/2$ gate time.
-Compiler optimizations for statically-sized arrays were utilized
-for all methods except for the unscented sampling method.
-The unscented sampling method's augmented state vector size
-was too large to take advantage of these optimizations,
-adversely affecting its run time.
-For the robust methods, the runtime scales super-linearly with the
-gate time. We expect the runtime to scale linearly with the gate time
-for large gate times because
-the number of knot points scales linearly with the gate time.
-We observe this behavior for the base method, which has a smaller
-augmented state vector size than the robust methods.
-We performed optimizations on a single CPU thread in this work as a proof
-of concept. Future work will parallelize the robustness
-methods across the initial states in the operator basis and utilize GPUs
+for a $Z/2$ gate at gate times which are multiples of $1/4f_{q} \sim 18$ns.
+We performed optimizations on a single core of
+an AMD Ryzen Threadripper 3970X 32-Core Processor.
+Future work will parallelize the robustness methods using GPUs
 \cite{leung2017speedup},
-allowing for fast optimization on large Hilbert spaces.
-\vspace*{\intextsep}
+which will enable fast optimizations on high-dimensional Hilbert spaces.
 
 \begin{table}[H]
-  \begin{tabular} {c | c | c | c | c | c | c}
-    & \multicolumn{6}{c}{Wall Time ($\pm 0.001$ s)}\\
+  \centering
+  \begin{tabular} {c | c | c | c }
+    & \multicolumn{3}{c}{Average Runtime (s)}\\
     \hline
-    $t_{N}$ (ns) & Base & Long. & S & SU & D1 & D2\\
+    $t_{N}$ (ns) & $18$ & $36$ & $72$\\
     \hline
-    $18$ & 0.155 & 1.688 & 1.773 & 210.573 & 16.713 & 48.398\\ % 0.155049
-    $36$ & 7.014 & - & 48.213 & 4566.236 & 67.838 & 81.030\\
-    $72$ & 15.906 & - & 281.372 & 16575.308 & 266.997 & 332.182\\
+    Base & $0.155 \pm 0.008$ & $7.0 \pm 0.4$ & $15.9 \pm 0.8$\\
+    Depol. & $1.69 \pm 0.08$ & - & -\\
+    S & $1.77 \pm 0.09$ & $48 \pm 2$ & $280 \pm 10$\\
+    U & $75 \pm 4$ & $340 \pm 20$ & $400 \pm 20$\\
+    D1 & $6.1 \pm 0.3$ & $27 \pm 1$ & $65 \pm 3$\\
+    D2 & $15.7 \pm 0.8$ & $17.3 \pm 0.9$ & $54 \pm 3$\\
   \end{tabular}
   \caption{
-    Runtimes for $Z/2$ optimizations.
-    Programs were executed on a single thread of an
-    AMD Ryzen Threadripper 3970X 32-Core Processor.
+    Average runtimes for $Z/2$ optimizations
+    using the base, depolarization, sampling (S),
+    unscented Sampling (U), and the 1\textsuperscript{st}-
+    and 2\textsuperscript{nd}-order derivative methods (D1, D2).
   }
@@
 For the sampling method, the size of the augmented state vector
-is $O(dn^{3})$, where $d$ is the number of deviant parameters and
+is $O(dn^{3})$, where $d$ is the number of uncertain parameters and
 $n$ is the dimension of the Hilbert space. There are $n^{2}$ initial states

### 2021-01-26 paper/2background.tex
updates

@@
 the prefactors for the constraint terms in the objective are increased
-and the procedure is repeated.
+intelligently and the procedure is repeated.
 

### 2021-01-29 paper/0abstract.tex
danny paper edits!

@@
   and utilize time-optimal control to achieve high fidelity gates in the presence of depolarization.
-  We apply these techniques to a fluxonium qubit and suppress
+  We apply these techniques to a two level approximation of a fluxonium qubit and suppress
   simulated gate errors due to parameter uncertainty below $10^{-7}$ for

### 2021-01-29 paper/1introduction.tex
danny paper edits!

@@
 from the model used in optimization, leading
-to poor experimental performance.
+to poor experimental performance (could be a good place for a citation here?
+It obviosuly makes sense in principle, but I'm not aware of a paper that tried to implement optimal control and failed for these reasons.
+*on reading further* Maybe the Zhou book from below).
 Robust control improves upon

### 2021-01-29 paper/2background.tex
danny paper edits!

@@
 iLQR then constructs quadratic models for each cost function using
-their zeroth-, first- and second-order derivatvies in a Taylor expansion
+their zeroth-, first- and second-order derivatives in a Taylor expansion
 about the current augmented controls and augmented states.

### 2021-01-29 paper/3fluxonium.tex
danny paper edits!

@@
 to ensure the solver does not take advantage of discretization errors in numerical integration.
-To refer to the discrete moments the flux, we introduce the notation
+To refer to the discrete moments of the flux, we introduce the notation
 $\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
@@
 We impose the zero net flux constraint $\int_{t} a_{N} = 0$
+(seems kind of funny to be referencing the equations so far below - might be nicer to introduce the equations, referencing Eqs. (2)a-(2)d, then talk about what they mean.)
 \eqref{eq:znf_con}
@@
     indicate measured $T_{1}$ times. Non-circle markers
-    indicate the time-averaged, absolute amplitude of each flux pulse.
+    indicate the time-averaged, absolute amplitude of each flux pulse (seems a bit odd to plot these on this figure - it looks like you're assigning a T1 value to the entire pulse, which I'm not sure how to interpret).
     (c) Cumulative gate errors due to depolarization as a function of the
@@
 $\ell_{k}(x_{k}, u_{k}) = (x_{k} - x_{f})^{T} Q_{k} (x_{k} - x_{f}) + u^{T}_{k} R_{k} u_{k}$
-where $Q_{k}$ and $R_{k}$ are diagonal matrices we supply. The $Q_{k}$ term
+where $Q_{k}$ and $R_{k}$ are diagonal matrices we supply (didn't we say above that iLQR constructs quadratic models for each cost function?
+Is that what's going on here?
+How do we reconcile then that "we supply" Q and R?).
+The $Q_{k}$ term
 penalizes deviations from the final augmented state $x_{f}$,

### 2021-01-29 paper/4longitude.tex
danny paper edits!

@@
 see Figure \ref{fig:longitudeb}.
+
 It is also possible to fit a spline to theoretically obtained data.
@@
 and $T_{1}$ values often fluctuate on hour
-time scales \cite{klimov2018fluctuations}.
+time scales \cite{klimov2018fluctuations}. (You could cut this - no need to justify using experimental data if you've got it! Maybe just cite where the data is coming from, if published.)
 
@@
 is appended to the augmented control \eqref{eq:astatecontrols}
-and the squared root
+and the squared root (this looks like a typo)
 of the time step $\lvert \Delta t_{k} \rvert$ is used
@@
 \todo{Jens, Danny, Ziwen, any idea what the correct way to simulate depolarization is
-  when $T_{1}$ is time dependent?}.
+  when $T_{1}$ is time dependent? Jens is going to have the best opinion on this - personally I think what you've done seems correct}.
 The gate error reported in this text is the infidelity
-of the evolved state and the target state averaged over 1000 pseudo-randomly
+of the evolved state and the target state averaged over 1000 pseudo-randomly (what is this a cryptography paper?! "pseudo" lol)
 generated initial states.
@@
 on the depolarization probability metric, see Appendix \ref{appendix:longitude}.
-This single gate error advantage corresponds to a significant reduction in error correction resources.
+This single gate error advantage corresponds to a significant reduction in error correction resources (this seems like a strong statement - citation?).
 Furthermore, for successive gate applications, the gate error due to depolarization is approximately linear

### 2021-01-29 paper/5static.tex
danny paper edits!

@@
     a function of the gate time. Missing
-    data points represent gates with a gate error greater than $5 \cdot 10^{-5}$.
+    data points represent gates with a gate error greater than $5 \cdot 10^{-5}$(...what's going on here? Should I be concerned? This seems like a relatively big thing that goes unexplained).
     (c) Single gate error as a function of the qubit frequency detuning.
@@
 We have formulated the QOC
-problem as an open-loop optimization problem, equivalently,
+problem as an open-loop optimization problem; equivalently,
 we do not incorporate feedback from the experiment in optimization.
@@
 $\sum_{k, \pm} q_{k} (1 - {\lvert \braket{\psi^{\pm}_{k}}{U \psi_{1}} \rvert}^{2})$
+(I thought of this question here but it applies also to when we first write down the objective 6a.
+Doesn't the sum over k imply that we are penalizing the time it takes to complete the gate?
+Why do we need to append $\Delta t_{k}$ to the augmented state vector?)
 where $q_{k}$ is a constant we supply.
@@
 In particular, the sample states are chosen to encode a unimodal distribution over
-the $2n$ elements of the nominal state, modeling the uncertainty in the state
+the $2n$ (at this point the reader has forgotten what lower-case n refers to) elements of the nominal state, modeling the uncertainty in the state
 as a result of the uncertainty in the parameter. The unscented transform
@@
 in Appendix \ref{appendix:unscented}.
+(this seems far more complicated than the sampling method - if it truly is necessary, perhaps earlier we could fully flesh out what the issues are?
+Does this drifting too close to the nominal state actually occur in practice?
+And if so, it isn't clear to me why that is a bad thing, as for example if it drifts close to the nominal state, then great, isn't the optimizer doing its job?)
 
@@
 to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
-state derivative $\partial_{f_{q}}^{l} \ket{\psi}$. In the $m$\textsuperscript{th}-order
+state derivative $\partial_{f_{q}}^{l} \ket{\psi}$ (seems funny to change from $l$ to $m$ immediately after introduction). In the $m$\textsuperscript{th}-order
 derivative method, we append all state derivatives of order $1, \dots, m$
@@
 for each initial state. We use the initial state $\ket{0}$ for this method,
-and observe no advantage to using more initial states.
+and observe no advantage to using more initial states (comment on why you've chosen different states for this vs. unscented?).
 We penalize the norms of the state derivatives
 in the objective \eqref{eq:costfun} by setting the corresponding elements
-of the final augmented state to zero.
+of the final augmented state to zero (this isn't totally obvious to me - why should this be the right way to penalize?
+Also the wording seems a bit off - surely you don't mean that you're penalizing the norm of the state derivative with itself?
+Isn't it normalized?).
 We could obtain the state derivatives at each knot point
@@
 We integrate the coupled ODEs with exponential
-integrators, see Appendix \ref{appendix:derivative},
-in the discrete dynamics function \eqref{eq:dyn_con}. For runtimes
+integrators
+in the discrete dynamics function \eqref{eq:dyn_con}, see Appendix \ref{appendix:derivative}. For runtimes
 and complexity analyses for the three robust control techniques,
@@
 
+(you have obviously made the choice to first present all three methods as theoretical tools, then discuss their results together, which is fine.
+It might make sense to also consider presenting a method and presenting results associated with that method, before moving on to the next one.
+Might make it easier to justify, e.g., the complexity of the unscented sampling method if there are deficiencies to be addressed that are clear from the results of the sampling method.)
 We examine the gate errors due to a static qubit frequency
@@
 methods.
+(right so reading further - this paragraph has nothing to do with the previous two - might make sense to put this with the corresponding explanation of the method.)
 
@@
 for multi-qubit experiments, where the qubits operate at different
-frequencies $f_{q, i} \neq f_{q, j}$.
+frequencies $f_{q, i} \neq f_{q, j}$ (I wouldn't introduce new notation if it isn't being used anywhere else. Can just delete this eqn).
 We can find solutions using the numerical methods at
@@
 multi-qubit experiments.
+(this seems like a funny way to end this paragraph - I was told that static parameter deviations were the issue the whole time, and suddenly it's that analytic pulses aren't flexible enough w.r.t.
+the time it takes to complete them.
+I would stress instead that you can do better than $\sim 4.5 \cdot 10^{-5}$!
+The time issue, though important, feels like it should be more of an aside given the things you've emphasized are the issues you're trying to solve.)
 
@@
 over the 1\textsuperscript{st}-order method for most gate times.
+(I would say the second order method as compared to the first order method feels the same as the sampling method as compared to the unscented method.
+You say here that there is no improvement, but improvement for unscented looks mostly like noise to me, even if you have some idea of where it's coming from (larmor period fractions).
+I would just be careful with how these things are framed - seems opportunistic to claim that unscented gives some sort of serious advantage here.)
 The gate error at a one-percent detuning for the 1\textsuperscript{st}-order

### 2021-01-29 paper/6stochastic.tex
danny paper edits!

@@
 superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Flux noise
+source of coherent errors (good place for a citation). Flux noise
 modifies the fluxonium Hamiltonian \eqref{eq:hamiltonian}

### 2021-01-29 paper/7conclusion.tex
danny paper edits!

@@
 \section{Conclusion}
-In conclusion, we have applied state-of-the-art trajectory
+We have applied state-of-the-art trajectory
 optimization techniques to mitigate decoherence and
@@
 independent of the dimension of the Hilbert space, enabling
-inexpensive optimization on high-dimensional quantum systems.
+inexpensive optimization on high-dimensional quantum systems (hmm, this seems like something that definitely wasn't emphasized in the text - could be worth talking about this a little earlier.
+Also doesn't seem totally true right?
+Re Eq.(5)).
 We have also proposed the derivative method for robust control which achieves
@@
 operations on their quantum systems.
-These methods will be used to achieve the low gate errors
+These methods can be used to achieve the low gate errors
 required for fault-tolerant quantum computing applications. Our

### 2021-01-29 paper/a2unscented.tex
updates
danny paper edits!

@@
 an uncertain parameter $\lambda \in \mathbb{R}^{d}$, and discrete dynamics
-$\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
+$\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$ (maybe choose one between this and $x_{k + 1} = f(x_{k}, u_{k})$).
 The state is obtained with an isomorphism \eqref{eq:isomorphism} and
@@
 The nominal initial state is given by $\bar{\psi}_{1}$ with an associated
-positive-definite covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$
+positive-definite covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$ (I certainly have never seen this notation before - obviously $\mathbb{R}^{2n}$ doesn't need to be explained but maybe $\mathbb{S}_{++}^{2n}$ is not as well known.)
 which describes the uncertainty in the initial state.
@@
 such as the principal square root may be employed.
-The superscript on the matrix square root indicates the $i$\textsuperscript{th}
+The superscript on the matrix square root indicates the $j$\textsuperscript{th}
 column (mod $2n + d$) of the lower triangular Cholesky factor.

### 2021-01-29 paper/a3derivative.tex
danny paper edits!

@@
 For example, the dynamics of the first state derivative in
-units of $i\hbar = 1$ are $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
+units of $i\hbar = 1$ (is setting i=1 something that is done?
+Seems fraught and scary...just leave it in there bro and set hbar to 1 like the rest of us!)are $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
 H \ket{\partial_{\lambda} \psi} + (\partial_{\lambda} H) \ket{\psi}$.

### 2021-02-05 paper/0abstract.tex
jens edits 2 addressed or todo'd

--- 2021-01-26 paper/0abstract.tex
@@
   parameter uncertainty.
-  We propose a method that takes advantage of uncertain parameter
-  derivative information while maintaining
-  computational efficiency by transforming high-order differential equations to coupled,
-  first-order ODEs.
-  Additionally, the effect of depolarization on a gate is most accurately modeled by
+  We propose a derivative-based approach that maintains
+  computational efficiency by using forward-mode differentiation.
+  Additionally, the effect of depolarization on a gate is typically modeled by
   integrating the Lindblad master equation,
   which is computationally expensive.
-  We propose a computationally efficient model
+  We employ a computationally efficient model
   and utilize time-optimal control to achieve high fidelity gates in the presence of depolarization.

### 2021-02-05 paper/1introduction.tex
jens edits 2 addressed or todo'd

--- 2021-01-26 paper/1introduction.tex
@@
 We also show that QOC can solve important problems associated with
-fluxonium-based qubits: mitigating
-depolarization by taking advantage of the $T_{1}$-dependence of the controls
-and synchronizing qubits with distinct frequencies
-by performing phase gates in arbitrary times.
+fluxonium-based qubits: exploiting the $T_{1}$-dependence of the controls
+to mitigate depolarization
+and synchronizing the phase of qubits with distinct frequencies.
 To mitigate depolarization,

### 2021-02-05 paper/2background.tex
jens edits 2 addressed or todo'd

--- 2021-01-26 paper/2background.tex
@@
 the controls and state are discretized into $N$ knot points (time steps).
-In the case of a single state transfer problem, the functional is
+In the case of a single state-transfer problem, the functional is
 the infidelity of the desired final state and the initial state evolved
@@
 projects the solution from the ALM stage onto the constraint manifold using
-a projected Newton method, achieving ultra-low
+a (direct) projected Newton method, achieving ultra-low
 constraint violations $\sim 10^{-8}$.

### 2021-02-05 paper/3fluxonium.tex
jens edits 2 addressed or todo'd

--- 2021-01-26 paper/3fluxonium.tex
@@
 First, we outline the constraints for the fluxonium gate problem.
-We formulate this problem as a multi-state-transfer problem.
+All gates in this work satisfy these constraints to a tolerance
+of $\sim 10^{-8}$.
 The initial conditions on
@@
 where the superscript is an index $i \in \{0, 1\}$,
-and the subscript indicates the first knot point $k = 1$.
+and the subscript indicates the knot point $k = 1$.
 The states at the final knot point are constrained to be
@@
 $\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
-$a(t = t_{k}) \equiv a_{k}$,
+$a(t_{k}) \equiv a_{k}$,
 $\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$.
@@
     indicate measured $T_{1}$ times. Non-circle markers
-    indicate the time-averaged, absolute amplitude of each flux pulse.
+    are plotted at the time-averaged, absolute amplitude of each flux pulse.
     (c) Cumulative gate errors due to depolarization as a function of the

### 2021-02-05 paper/4longitude.tex
jens edits 2 addressed or todo'd

--- 2021-01-26 paper/4longitude.tex
@@
 We avoid the increase in computational complexity required for these
-techniques by penalizing the integrated rate (probability) of depolarization in optimization.
-Using this probability as a proxy for the gate error incurred
+techniques by penalizing the integrated depolarization rate in optimization.
+Using this metric as a proxy for the gate error incurred
 is reasonable because depolarization errors are incoherent and therefore
@@
 
-The depolarization probability is given by,
+The integrated depolarization rate is given by,
 \begin{equation}
-  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}.
+  D_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}.
 \end{equation}
@@
 We allow the optimizer to tune the gate time by
-making the time step between each knot point $\Delta t_{k}$
-a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
-than the number of knot points $N$, preserves the
-Markovianity of the trajectory
-optimization problem. The square root of the time step $\sqrt{\Delta t_{k}}$
+making the time step between each knot point
+a decision variable. 
+The square root of the time step $\sqrt{\Delta t_{k}}$
 is appended to the augmented control \eqref{eq:astatecontrols}
@@
 the bounds of the time step at each knot point.
+We promote the time steps
+to decision variables, rather
+than the number of knot points $N$, to ensure each cost function depends
+only on the augmented state and augmented
+controls at its own knot point $\ell_{k}(x_{k}, u_{k})$, which is necessary for iLQR.
 
@@
 after each application, see Appendix \ref{appendix:longitude}.
-\todo{Jens, Danny, Ziwen, any idea what the correct way to simulate depolarization is
-  when $T_{1}$ is time dependent?}.
 The gate error reported in this text is the infidelity
@@
 which tracks closely with their relative improvement
-on the depolarization probability metric, see Appendix \ref{appendix:longitude}.
+on the integrated depolarization rate metric, see Appendix \ref{appendix:longitude}.
 This single gate error advantage corresponds to a significant reduction in error correction resources.

### 2021-02-05 paper/5static.tex
jens edits 2 addressed or todo'd

--- 2021-01-26 paper/5static.tex
@@
 we do not incorporate feedback from the experiment in optimization.
-However, the device typically deviates from the Hamiltonian we use in optimization,
+However, the device's parameters deviate from the parameters we use in optimization,
 leading to poor experimental performance. We combat errors
@@
 making the state evolution insensitive
-to Hamiltonian parameter uncertainty. As an example,
+to parameter uncertainty. As an example,
 we mitigate errors arising from the drift and finite measurement
@@
 The sampling method incentivizes the optimizer
-to ensure multiple instances of a state, each of which evolves
+to ensure multiple copies of a state, each of which evolves
 with a distinct value of the uncertain parameter, achieve
@@
 by adding a cost function to the objective \eqref{eq:costfun} of the form
-$\sum_{k, \pm} q_{k} (1 - {\lvert \braket{\psi^{\pm}_{k}}{U \psi_{1}} \rvert}^{2})$
-where $q_{k}$ is a constant we supply.
+$\sum_{k, \pm} b_{k} (1 - {\lvert \bra*{\psi^{\pm}_{k}}U\ket*{\psi_{1}} \rvert}^{2})$
+where $b_{k}$ is a constant we supply.
 For this method, the standard orthonormal basis states are an insufficient choice
@@
 to the objective \eqref{eq:costfun} takes the form
-$\sum_{k, j} q_{k} (\psi^{j}_{k} - \psi_{k})^{T}
-(\psi^{j}_{k} - \psi_{k})$, where $q_{k}$ is a
+$\sum_{k, j} c_{k} (\psi^{j}_{k} - \psi_{k})^{T}
+(\psi^{j}_{k} - \psi_{k})$, where $c_{k}$ is a
 constant we supply, $\psi_{k}$ is
@@
 in the discrete dynamics function \eqref{eq:dyn_con}. For runtimes
-and complexity analyses for the three robust control techniques,
+of the three robust control techniques,
 consult Appendix \ref{appendix:time}.

### 2021-02-05 paper/6stochastic.tex
jens edits 2 addressed or todo'd

--- 2021-01-26 paper/6stochastic.tex
@@
 An additional source of experimental error arises from time-dependent
-Hamiltonian parameter uncertainty. For many flux-biased and inductively-coupled
+parameter uncertainty. For many flux-biased and inductively-coupled
 superconducting circuit elements, magnetic flux noise is a significant

### 2021-02-05 paper/a1longitude.tex
jens edits 2 addressed or todo'd

@@
 our procedure for integrating the Lindblad master equation.
-The depolarization probability and the gate error due to
+The integrated depolarization rate and the gate error due to
 depolarization metrics are compared in Table \ref{tab:longitude} for the
@@
   \begin{tabular}{c | c | c | c | c | c | c}
-         & $P_{1\textrm{A}}$ & $P_{1\textrm{N}}$ & & GE$_{\textrm{A}}$ & GE$_{\textrm{N}}$ &\\
-    Gate & $(10^{-5})$ & $(10^{-5})$ & $P_{1\textrm{A}}/ P_{1\textrm{N}}$ & $(10^{-5})$ &
+         & $D_{1\textrm{A}}$ & $D_{1\textrm{N}}$ & & GE$_{\textrm{A}}$ & GE$_{\textrm{N}}$ &\\
+    Gate & $(10^{-5})$ & $(10^{-5})$ & $D_{1\textrm{A}}/ D_{1\textrm{N}}$ & $(10^{-5})$ &
     $(10^{-5})$ & GE$_{\textrm{A}} / \textrm{GE}_{\textrm{N}}$\\
@@
   \caption{
-    Single gate depolarization probability (P)
+    Single gate integrated depolarization rate ($D_{1}$)
     and single gate error due to depolarization (GE).
@@
   \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho]
-  + \sum_{i = 1}^{n^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger}
+  + \sum_{i} \gamma_{i} (L_{i} \rho L_{i}^{\dagger}
   - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\}),
 \end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $n$ is the
-dimension of the Hilbert space,
+where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix,
 $[\cdot, \cdot]$ is the algebraic commutator, and $\{\cdot, \cdot \}$ is the algebraic
@@
 $L_{\downarrow} = \sigma^{-}/2$ where $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$.
-Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained from the spline
-shown in Figure \ref{fig:longitudeb}.
+Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained at each knot point
+from the spline shown in Figure \ref{fig:longitudeb}.
 We obtain the $T_{1}$ values in this spline
@@
     \hat{\mathcal{L}} &= -i(I \otimes H - H^{T} \otimes I)\\
-    &+ \sum_{i = 1}^{n^{2} - 1} \gamma_{i}
+    &+ \sum_{i} \gamma_{i}
     (L_{i}^{*} \otimes L_{i} - \frac{1}{2} (I \otimes L_{i}^{\dagger}L_{i}
@@
 and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{i, j} \ket{i} \otimes \ket{j}$.
-We use zero-order hold on the controls so the exact solution is
+We use zero-order hold on the controls--equivalently, $a(t)$ is piecewise-constant
+on the intervals $[t_{k}, t_{k + 1}]$. Therefore,
+$H(a(t))$, $L(a(t))$, and $\gamma_{i}(a(t))$ are also piecewise-constant.
+So, the exact solution is
 $\textrm{vec}(\rho_{k + 1}) = {\exp}{\textstyle(}\Delta t_{k} \hat{\mathcal{L}}_{k}{\textstyle)} \textrm{vec}(\rho_{k})$.
-This isomorphism transforms $(n^{2} \times n^{2}) \times (n^{2} \times n^{2})$
-matrix-matrix multiplications to $(n^{4} \times n^{4}) \times n^{4}$ matrix-vector
+This isomorphism transforms $(n \times n) \times (n \times n)$
+matrix-matrix multiplications to $(n^{2} \times n^{2}) \times n^{2}$ matrix-vector
 multiplications. For small $n$ and zero-order hold on the controls, we find that it is

### 2021-02-05 paper/a2unscented.tex
jens edits 2 addressed or todo'd

--- 2021-01-26 paper/a2unscented.tex
@@
 $\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
-The state is obtained with an isomorphism \eqref{eq:isomorphism} and
-we omit bra-ket notation to emphasize that the state is real.
+We omit bra-ket notation to emphasize that
+the state is a real vector, and it is given by the right-hand-side of the
+isomorphism \eqref{eq:isomorphism}.
 The nominal initial state is given by $\bar{\psi}_{1}$ with an associated
-positive-definite covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$
+covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$
 which describes the uncertainty in the initial state.
+We use the notation $\mathbb{S}_{++}^{m}$ to denote the
+set of real, symmetric, and positive-definite $m \times m$ matrices.
 $P_{1}$ is typically non-zero even if the state preparation error is negligible.

### 2021-02-05 paper/a4time.tex
danny paper edits!
jens edits 2 addressed or todo'd

@@
 \section{Computational Performance \label{appendix:time}}
-We provide runtimes for our optimizations and comment on the scaling of the
+We provide runtimes for our optimizations and comment on the problem size of the
 robustness methods. The runtimes for
@@
 
-Now we present the problem size complexities for the robustness methods.
+Now we note the size of the augmented state vector for the robustness methods.
 For the sampling method, the size of the augmented state vector
@@
 initial states in the operator basis, $dm$ state derivatives per initial state,
-and each state has $2n$ real numbers.
+and each state has $2n$ real numbers. \todo{maybe add comment about auto-diff}

### 2021-02-05 paper/2background.tex
more updates

@@
 to the objective. The new objective is then minimized with
-iLQR. If the solution produced by iLQR does not satisfy the constraints,
+iLQR. If the solution obtained with iLQR does not satisfy the constraints,
 the prefactors for the constraint terms in the objective are increased

### 2021-02-05 paper/5static.tex
more updates

@@
 We have formulated the QOC
-problem as an open-loop optimization problem, equivalently,
+problem as an open-loop optimization problem--equivalently,
 we do not incorporate feedback from the experiment in optimization.

### 2021-02-05 paper/a3derivative.tex
more updates

--- 2021-01-26 paper/a3derivative.tex
@@
 Here, we outline how to efficiently integrate the dynamics
-of the derivative method using exponential integrators.
+for the derivative method using exponential integrators.
 General exponential integrators break the dynamics into a linear term and a non-linear term.

### 2021-02-05 paper/1introduction.tex

@@
 from the model used in optimization, leading
-to poor experimental performance.
+to poor experimental performance
+\todo{could be a good place for a citation here?
+  It obviosuly makes sense in principle,
+  but I'm not aware of a paper that tried to implement optimal
+  control and failed for these reasons.
+  *on reading further* Maybe the Zhou book from below}.
 Robust control improves upon
@@
 control frameworks using closed-loop methods \cite{egger2014adaptive, feng2018gradient,
-  wittler2020integrated} and open-loop methods \cite{
+  li2017hybrid, wittler2020integrated} and open-loop methods \cite{
   allen2019robust, carvalho2020error, reinhold2019controlling,
-  rembold2020introduction, kosut2013robust}.
+  rembold2020introduction, kosut2013robust, niu2019universal}.
 

### 2021-02-05 paper/3fluxonium.tex

@@
 We impose the zero net flux constraint $\int_{t} a_{N} = 0$
+\todo{seems kind of funny to be referencing the equations so far below - might be nicer to introduce the equations, referencing Eqs. (2)a-(2)d, then talk about what they mean.}
 \eqref{eq:znf_con}
@@
 \begin{figure*}[ht]
-
   \begin{subfigure}{.315\textwidth}
@@
     indicate measured $T_{1}$ times. Non-circle markers
-    are plotted at the time-averaged, absolute amplitude of each flux pulse.
+    are plotted at the time-averaged, absolute amplitude of each flux pulse
+    \todo{seems a bit odd to plot these on this figure - it looks like you're assigning a T1 value to the entire pulse, which I'm not sure how to interpret}.
     (c) Cumulative gate errors due to depolarization as a function of the
@@
 $\ell_{k}(x_{k}, u_{k}) = (x_{k} - x_{f})^{T} Q_{k} (x_{k} - x_{f}) + u^{T}_{k} R_{k} u_{k}$
-where $Q_{k}$ and $R_{k}$ are diagonal matrices we supply. The $Q_{k}$ term
+where $Q_{k}$ and $R_{k}$ are diagonal matrices we supply
+\todo{didn't we say above that iLQR constructs
+quadratic models for each cost function? Is that what's going on here? How do we reconcile then that "we supply" Q and R?}. The $Q_{k}$ term
 penalizes deviations from the final augmented state $x_{f}$,

### 2021-02-05 paper/4longitude.tex

@@
 time scales \cite{klimov2018fluctuations}.
+\todo{You could cut this - starting from "It is also" to end of paragraph -
+  no need to justify using experimental data if you've got it!
+  Maybe just cite where the data is coming from, if published.}
 
@@
 is appended to the augmented control \eqref{eq:astatecontrols}
-and the squared root
-of the time step $\lvert \Delta t_{k} \rvert$ is used
+and its square $\lvert \Delta t_{k} \rvert$ is used
 for numerical integration in the discrete dynamics function \eqref{eq:dyn_con}.
@@
 on the integrated depolarization rate metric, see Appendix \ref{appendix:longitude}.
-This single gate error advantage corresponds to a significant reduction in error correction resources.
+This single gate error advantage corresponds to a significant reduction in error correction resources
+\todo{this seems like a strong statement - citation?}.
 Furthermore, for successive gate applications, the gate error due to depolarization is approximately linear

### 2021-02-05 paper/5static.tex

@@
     a function of the gate time. Missing
-    data points represent gates with a gate error greater than $5 \cdot 10^{-5}$.
+    data points represent gates with a gate error greater than $5 \cdot 10^{-5}$
+    \todo{...what's going on here? Should I be concerned? This seems like a relatively big thing that goes unexplained}.
     (c) Single gate error as a function of the qubit frequency detuning.
@@
 We have formulated the QOC
-problem as an open-loop optimization problem--equivalently,
+problem as an open-loop optimization problem; equivalently,
 we do not incorporate feedback from the experiment in optimization.
@@
 where $b_{k}$ is a constant we supply.
+\todo{I thought of this question here but it applies also to when we first write down the objective 6a.
+Doesn't the sum over k imply that we are penalizing the time it takes to complete the gate?
+Why do we need to append $\Delta t_{k}$ to the augmented state vector?}
 For this method, the standard orthonormal basis states are an insufficient choice
@@
 In particular, the sample states are chosen to encode a unimodal distribution over
-the $2n$ elements of the nominal state, modeling the uncertainty in the state
+the $2n$ \todo{at this point the reader has forgotten what lower-case n refers to} elements of the nominal state, modeling the uncertainty in the state
 as a result of the uncertainty in the parameter. The unscented transform
@@
 in Appendix \ref{appendix:unscented}.
+\todo{this seems far more complicated than the sampling method - if it truly is necessary,
+  perhaps earlier we could fully flesh out what the issues are?
+  Does this drifting too close to the nominal state actually occur in practice?
+  And if so, it isn't clear to me why that is a bad thing, as for example if it drifts close to the nominal state,
+  then great, isn't the optimizer doing its job?}
 
@@
 to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
-state derivative $\partial_{f_{q}}^{l} \ket{\psi}$. In the $m$\textsuperscript{th}-order
+state derivative $\partial_{f_{q}}^{l} \ket{\psi}$ \todo{seems funny to change from $l$ to $m$ immediately after introduction}.
+In the $m$\textsuperscript{th}-order
 derivative method, we append all state derivatives of order $1, \dots, m$
@@
 for each initial state. We use the initial state $\ket{0}$ for this method,
-and observe no advantage to using more initial states.
+and observe no advantage to using more initial states
+\todo{comment on why you've chosen different states for this vs. unscented?}.
 We penalize the norms of the state derivatives
 in the objective \eqref{eq:costfun} by setting the corresponding elements
-of the final augmented state to zero.
+of the final augmented state to zero
+\todo{this isn't totally obvious to me - why should this be the right way to penalize?
+  Also the wording seems a bit off - surely you don't mean that you're penalizing the norm of the state derivative with itself? Isn't it normalized?}.
 We could obtain the state derivatives at each knot point
@@
 We integrate the coupled ODEs with exponential
-integrators, see Appendix \ref{appendix:derivative},
-in the discrete dynamics function \eqref{eq:dyn_con}. For runtimes
+integrators
+in the discrete dynamics function \eqref{eq:dyn_con}, see Appendix \ref{appendix:derivative}.
+For runtimes
 of the three robust control techniques,
@@
 
+\todo{you have obviously made the choice to first present all three methods as theoretical tools,
+  then discuss their results together, which is fine. It might make sense to also consider presenting
+  a method and presenting results associated with that method, before moving on to the next one.
+  Might make it easier to justify, e.g., the complexity of the unscented sampling method
+  if there are deficiencies to be addressed that are clear from the results of the sampling method.}
 We examine the gate errors due to a static qubit frequency
@@
 methods.
+\todo{right so reading further - this paragraph has nothing to do with the previous two
+  - might make sense to put this with the corresponding explanation of the method.}
 
@@
 for multi-qubit experiments, where the qubits operate at different
-frequencies $f_{q, i} \neq f_{q, j}$.
+frequencies.
 We can find solutions using the numerical methods at
@@
 multi-qubit experiments.
+\todo{this seems like a funny way to end this paragraph - I was told that static parameter deviations were the issue the whole time, and suddenly it's that analytic pulses aren't flexible enough w.r.t.
+the time it takes to complete them.
+I would stress instead that you can do better than $\sim 4.5 \cdot 10^{-5}$!
+The time issue, though important, feels like it should be more of an aside given the things you've emphasized are the issues you're trying to solve.}
 
@@
 over the 1\textsuperscript{st}-order method for most gate times.
+\todo{I would say the second order method as compared to the first order method feels the same as the sampling method as compared
+  to the unscented method. You say here that there is no improvement, but improvement for unscented looks mostly like noise to me,
+  even if you have some idea of where it's coming from (larmor period fractions).
+  I would just be careful with how these things are framed - seems opportunistic to claim that unscented gives some sort of serious advantage here.}
 The gate error at a one-percent detuning for the 1\textsuperscript{st}-order

### 2021-02-05 paper/6stochastic.tex

@@
 superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Flux noise
+source of coherent errors \todo{good place for a citation}. Flux noise
 modifies the fluxonium Hamiltonian \eqref{eq:hamiltonian}

### 2021-02-05 paper/a2unscented.tex

@@
 an uncertain parameter $\lambda \in \mathbb{R}^{d}$, and discrete dynamics
-$\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
+$\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$ (maybe choose one between this and $x_{k + 1} = f(x_{k}, u_{k})$)..
 We omit bra-ket notation to emphasize that

### 2021-02-09 paper/1introduction.tex
resolved danny edits

--- 2021-02-05 paper/1introduction.tex
@@
 and applications now include superconducting circuits \cite{abdelhafez2020universal,
-  chakram2020multimode, fisher2010optimal, gokhale2019partial,
-  leng2019robust, leung2017speedup, li2020fast},
+  chakram2020multimode, egger2013optimized, fisher2010optimal, gokhale2019partial,
+  huang2014optimal, heeres2017implementing, kelly2014optimal, leng2019robust,
+  leung2017speedup, li2020fast,
+  liebermann2016optimal, reinhold2019controlling,
+  rebentrost2009optimal, rebentrost2009optimal2, spiteri2018quantum,
+  sporl2007optimal},
 neutral atoms and ions \cite{brouzos2015quantum,
-  de2008optimal, goerz2011quantum, guo2019high, jensen2019time,
-  larrouy2020fast, omran2019generation, rosi2013fast, sorensen2019qengine,
+  de2008optimal, grace2007optimal, goerz2011quantum, guo2019high, jensen2019time,
+  larrouy2020fast, nebendahl2009optimal, omran2019generation,
+  rosi2013fast,
   treutlein2006microwave, van2016optimal},
@@
   nobauer2015smooth, poggiali2018optimal, rembold2020introduction, tian2019optimal},
-and Bose-Einstein condensates \cite{amri2019optimal, sorensen2018quantum}.
+and Bose-Einstein condensates \cite{amri2019optimal, doria2011optimal,
+  sorensen2019qengine, sorensen2018quantum}.
 In the context of quantum computation,
@@
 finite control resolution cause the system to deviate
-from the model used in optimization, leading
-to poor experimental performance.
+from the model used in optimization, hampering
+experimental performance
+\cite{chakram2020multimode, heeres2017implementing, klimov2020snake,
+  omran2019generation, reinhold2019controlling}.
 Robust control improves upon
@@
 pulses obtained with the DRAG scheme \cite{motzoi2009simple}.
-These techniques are not amenable to arbitrary experimental constraints,
-so recent work has sought to achieve robustness in quantum optimal
-control frameworks using closed-loop methods \cite{egger2014adaptive, feng2018gradient,
-  wittler2020integrated} and open-loop methods \cite{
+As compared to analytical techniques, QOC is advantageous for
+designing pulses that consider all experimental constraints and
+performance tradeoffs
+\cite{leung2017speedup},
+and for constructing operations without a known analytic solution
+\cite{chakram2020multimode, heeres2017implementing}.
+Accordingly, recent work has sought to achieve robustness in QOC
+frameworks using closed-loop methods \cite{egger2014adaptive, feng2018gradient,
+  li2017hybrid, wittler2020integrated} and open-loop methods \cite{
   allen2019robust, carvalho2020error, reinhold2019controlling,
-  rembold2020introduction, kosut2013robust}.
+  rembold2020introduction, kosut2013robust, niu2019universal}.
 
@@
   reinhold2019controlling, rembold2020introduction}.
-  \item An unscented sampling method adapted from the unscented transform used in
-    state estimation \cite{howell2020direct, julier2004unscented,
-      lee2013sigma, thangavel2020robust}.
+\item An unscented sampling method \cite{howell2020direct, lee2013sigma, thangavel2020robust}
+  adapted from the unscented transform \cite{julier2004unscented,
+    uhlmann1995dynamic} used in state estimation.
   \item A derivative method, which penalizes the sensitivity of the quantum state trajectory
@@
 solve these optimization problems using ALTRO (Augmented Lagrangian TRajectory Optimizer)
-\cite{howell2019altro}, which can enforce constraints on the quantum state trajectory
-and control parameters.
+\cite{howell2019altro}, which can enforce constraints on
+the control parameters and the quantum state trajectory.
 
@@
 define the associated QOC problem in Section \ref{sec:fluxonium}.
-Then, we formulate a method for supressing depolarization
+Then, we formulate a method for suppressing depolarization
 in Section \ref{sec:longitude}. Next, we describe three techniques for achieving

### 2021-02-09 paper/2background.tex
resolved danny edits

@@
     {x_{1:N},u_{1:N-1}}{\ell_{N}(x_N) + \sum_{k=1}^{N-1} \ell_{k}(x_k,u_k)}{}{} \label{eq:gcostfun}
-    \addConstraint{x_{k+1} = f(x_k,u_k)} \label{eq:gdyncon}
-    \addConstraint{g_k(x_k,u_k)}{\leq 0} \label{eq:ineqcon}
-    \addConstraint{h_k(x_k,u_k)}{=0} \label{eq:eqcon}.
+    \addConstraint{x_{k+1} = f(x_k,u_k) \ \forall \ k} \label{eq:gdyncon}
+    \addConstraint{g_k(x_k,u_k) \leq 0 \ \forall \ k} \label{eq:ineqcon}
+    \addConstraint{h_k(x_k,u_k) = 0 \ \forall \ k} \label{eq:eqcon}.
 \end{mini!}
@@
 Then, the derivative information is employed to update the augmented controls.
-This is the approach taken by standard QOC solvers such as GOAT \cite{machnes2015tunable},
+This approach is taken by standard QOC solvers such as GOAT \cite{machnes2015tunable},
 GRAPE \cite{khaneja2005optimal, leung2017speedup}, and Krotov's method \cite{goerz2019krotov}.
@@
 to satisfy the dynamics constraint \eqref{eq:gdyncon} to a specified tolerance.
-In this sense, the TDSE \eqref{eq:tdse} is a constraint which may be violated
+In this sense, the TDSE \eqref{eq:tdse} is a constraint that may be violated
 for intermediate steps of the optimization, where the states need not be physical.
 The direct approach lends itself to a nonlinear program formulation, for which
-a variety of general purpose solvers exist \cite{gill2005snopt, wachter2006implementation}.
+a variety of general-purpose solvers exist \cite{gill2005snopt, wachter2006implementation}.
 Recent state-of-the-art solvers, such as ALTRO,
@@
 is performed in the direction of this update to ensure a
-decrease in the sum of the cost functions. This procedure
+decrease in the objective \eqref{eq:gcostfun}. This procedure
 is repeated until convergence.
@@
 because they are very computationally efficient and maintain
-high accuracy for the discrete dynamics throughout optimization.
-However, standard implementations have no ability to handle
+high accuracy for the discrete dynamics throughout the optimization.
+However, standard implementations cannot handle
 nonlinear equality and inequality
@@
 not the augmented states.
-Another technique, which is popular for QOC \cite{leung2017speedup},
+Another technique, which is popular for QOC \cite{heeres2017implementing, leung2017speedup,
+reinhold2019controlling},
 is to add the constraint functions to the objective
@@
 ALM adds terms that are linear and quadratic in the constraint functions
-to the objective. The new objective is then minimized with
+to the objective. Then, the new objective is minimized with
 iLQR. If the solution obtained with iLQR does not satisfy the constraints,

### 2021-02-09 paper/3fluxonium.tex
resolved danny edits

--- 2021-02-05 paper/3fluxonium.tex
@@
-\section{QOC for the Fluxonium \label{sec:fluxonium}}
-In the following, we optimize quantum gates
-(unitary transformations) for the fluxonium qubit.
-The fluxonium is a promising building block
-for quantum computers, and the accurate
-two-level approximation of its Hamiltonian makes
-QOC on a classical computer inexpensive.
-To high accuracy, we approximate the Hamiltonian near the flux frustration
-as a spin-$1/2$ system:
-\begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2},
-  \label{eq:hamiltonian}
-\end{align}
-where $f_{q}$ is the qubit frequency at the flux frustration point,
-$a(t)$ is the flux bias, $h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
-are Pauli matrices. Although we use an approximation for the Hamiltonian,
-our noise model is realistic and considers the full Hilbert space. We optimize $X/2$,
-$Y/2$, and $Z/2$ gates for the fluxonium presented in \cite{zhang2020universal},
-and compare them to the analytically constructed gates for
-that device.
-
-First, we outline the constraints for the fluxonium gate problem.
-All gates in this work satisfy these constraints to a tolerance
-of $\sim 10^{-8}$.
-The initial conditions on
-the states are $\ket{\psi^{0}_{1}} = \ket{0}$, $\ket{\psi^{1}_{1}} = \ket{1}$
-\eqref{eq:istate_con}
-where the superscript is an index $i \in \{0, 1\}$,
-and the subscript indicates the knot point $k = 1$.
-The states at the final knot point are constrained to be
-the image of the initial states under the desired gate $U$,
-$\ket{\psi^{i}_{N}} = U \ket{\psi^{i}_{1}} \ \forall \ i$
-\eqref{eq:tstate_con}.
-Furthermore, we impose the normalization constraint
-${\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}} \rvert}^{2} = 1 \ \forall \ i,k$
-\eqref{eq:statenorm_con}
-to ensure the solver does not take advantage of discretization errors in numerical integration.
-To refer to the discrete moments the flux, we introduce the notation
-$\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
-$a(t_{k}) \equiv a_{k}$,
-$\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$.
-We impose the zero net flux constraint $\int_{t} a_{N} = 0$
-\eqref{eq:znf_con}
-which mitigates the inductive drift ubiquitous in flux-bias lines
-\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
-The flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
-\eqref{eq:amp_con}
-to ensure the two-level approximation \eqref{eq:hamiltonian} remains valid.
-We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con}
-so the gates may be concatenated arbitrarily. Additionally,
-we have the initial condition $\int_{t} a_{1} = \partial_{t} a_{1} = 0$
-\eqref{eq:ic_con}.
-
 %% F1
 \begin{figure*}[ht]
-
   \begin{subfigure}{.315\textwidth}
@@
     indicate measured $T_{1}$ times. Non-circle markers
-    are plotted at the time-averaged, absolute amplitude of each flux pulse.
+    are plotted at the time-averaged 
+    absolute flux and the time-averaged $T_{1}$ time for each pulse.
     (c) Cumulative gate errors due to depolarization as a function of the
@@
 \end{figure*}
+
+\section{QOC for the Fluxonium \label{sec:fluxonium}}
+In the following, we optimize quantum gates
+for the superconducting fluxonium qubit--a promising
+building block for quantum computers due to its high
+coherence times
+\cite{earnest2018realization, lin2018demonstration,
+  manucharyan2009fluxonium, nguyen2019high,
+  zhang2020universal}.
+In this section, we outline the base optimization problem,
+which we extend in subsequent sections to account
+for experimental error channels.
+To high accuracy, we approximate the Hamiltonian near the flux-frustration
+point as a two-level system:
+\begin{align}
+  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2},
+  \label{eq:hamiltonian}
+\end{align}
+where $f_{q}$ is the qubit frequency at the flux-frustration point,
+$a(t)$ is the flux offset from the flux-frustration point,
+$h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
+are Pauli matrices. We optimize $X/2$,
+$Y/2$, and $Z/2$ gates (unitary transformations)
+for the fluxonium presented in \cite{zhang2020universal},
+and compare them to the analytically constructed gates for
+that device.
+
+First, we outline the constraints for the fluxonium gate problem.
+All gates presented in this work satisfy these constraints to
+a tolerance of $\sim 10^{-8}$.
+Casting this problem as a multi-state transfer, the initial conditions on
+the states are $\ket*{\psi^{0}_{1}} = \ket*{0}$, $\ket*{\psi^{1}_{1}} = \ket*{1}$
+\eqref{eq:istate_con}
+where the superscript is an index $i \in \{0, 1\}$,
+and the subscript indicates the knot point $k = 1$.
+The states at the final knot point are constrained to be
+the target states $\ket*{\psi^{i}_{N}} = U \ket*{\psi^{i}_{1}} \ \forall \ i$
+\eqref{eq:tstate_con} where $U = X/2, Y/2, Z/2$ is the desired gate.
+Furthermore, we impose the normalization constraint
+${\lvert \braket*{\psi^{i}_{k}}{\psi^{i}_{k}} \rvert}^{2} = 1 \ \forall \ i,k$
+\eqref{eq:statenorm_con}
+to ensure the solver does not take advantage of discretization errors in numerical integration.
+To refer to the discrete moments of the flux, we introduce the notation
+$\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
+$a(t_{k}) \equiv a_{k}$,
+$\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$.
+We impose the zero net flux constraint $\int_{t} a_{N} = 0$
+\eqref{eq:znf_con}
+which mitigates the inductive drift ubiquitous in flux-bias lines
+\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
+The flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
+\eqref{eq:amp_con}.
+Above $0.5$ GHz, we observe excitation out of the first two levels, disallowing the
+approximation \eqref{eq:hamiltonian}.
+We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con}
+so the gates may be concatenated arbitrarily. Additionally,
+we have the initial condition $\int_{t} a_{1} = \partial_{t} a_{1} = 0$
+\eqref{eq:ic_con}.
 
@@
 \end{equation}
-The variables in the augmented state are derived from
-the decision variable in the augmented
-control via coupled, first-order differential equations in the 
-the discrete dynamics function \eqref{eq:dyn_con}.
+The variables in the augmented state are obtained from the decision variable
+in the augmented control
+via coupled, first-order, differential equations, which are integrated in the 
+discrete dynamics function \eqref{eq:dyn_con}.
 We integrate the states according to the TDSE \eqref{eq:tdse} and the
-fluxonium Hamiltonian \eqref{eq:hamiltonian} and integrate
-the second-derivative of the flux amplitude to obtain
-the first-derivative, proportional, and integral moments.
+fluxonium Hamiltonian \eqref{eq:hamiltonian}.
 The ALTRO implementation we use does not currently
@@
 \end{equation}
+
 The cost function at each knot point is
 $\ell_{k}(x_{k}, u_{k}) = (x_{k} - x_{f})^{T} Q_{k} (x_{k} - x_{f}) + u^{T}_{k} R_{k} u_{k}$
-where $Q_{k}$ and $R_{k}$ are diagonal matrices we supply. The $Q_{k}$ term
+where $Q_{k}$ and $R_{k}$ are diagonal matrices we supply.
+The $Q_{k}$ term
 penalizes deviations from the final augmented state $x_{f}$,
@@
 $\ket*{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$ in addition to
-$\partial_{t} a_{f} = 0$. The $R_{k}$ term penalizes the norm of $\partial^{2}_{t} a_{k}$,
-mitigating AWG ringing due to high-frequency transitions.
+$\partial_{t} a_{f} = 0$.
+The $R_{k}$ term penalizes the norm of $\partial^{2}_{t} a_{k}$,
+smoothing the flux to mitigate high-frequency AWG transitions.
 Stated succinctly, the optimization problem takes the form:
 \begin{mini!}[2] 
-  {x_{1:N}, u_{1:N\text{-}1}}{\sum_{k=1}^N {(x_k\text{-}x_f)}^{T} Q_k (x_k\text{-}x_{f})
+  {x_{1:N}, u_{1:N\text{-}1}}{\sum_{k=1}^N {(x_k-x_f)}^{T} Q_k (x_k-x_{f})
     + \sum_{k=1}^{N-1} {u_k}^{T} R_k u_{k}}{}{} \label{eq:costfun}
-  \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
+  \addConstraint{x_{k+1}}{= f(x_k, u_k) \ \forall \ k}  \label{eq:dyn_con}
   \addConstraint{\ket{\psi^{0}_{1}} = \ket{0}, \ket{\psi^{1}_{1}} = \ket{1}} \label{eq:istate_con}
@@
 \end{mini!}
+
+Lastly, we remark on decisions we have made in the problem definition.
+We penalize deviations from the target state at all knot points \eqref{eq:costfun}
+because it benefits the iLQR solving stage \cite{Jackson2020altroc}.
+This choice is unrelated to schemes
+for incentivizing early achievement of the desired gate \cite{leung2017speedup}.
+In practice, the penalty for deviations from the target state
+at the final knot point ($k = N$) becomes large relative to those at
+intermediate knot points ($k = 1, \dots, N - 1$)
+because ALM increases the penalty at the final knot point
+to satisfy the target state constraint \eqref{eq:tstate_con}.
+Furthermore, the target state constraint \eqref{eq:tstate_con} says
+that the final state must match the target state, including its phase,
+up to the constraint tolerance $\sim 10^{-8}$.
+If we did not make target state achievement a constraint, the optimizer would
+be allowed to sacrifice the nominal gate error to achieve more depolarization protection,
+or more robustness to parameter uncertainties, which is undesirable.
+Additionally, we chose the target state constraint to be
+phase-sensitive because the Hessian of the constraint function is diagonal,
+and therefore computationally efficient, and we wish to optimize $Z/2$ gates,
+which require a phase-sensitive metric for the initial states $\ket{0}$ and $\ket{1}$.

### 2021-02-09 paper/4longitude.tex
resolved danny edits

@@
 for optimizing the flux to mitigate depolarization.
-For many superconducting circuits, the 1/$e$ depolarization time ($T_{1}$) is independent of the control parameters,
+For many superconducting circuits, the 1/$e$ depolarization time
+($T_{1}$) is independent of the control parameters,
 so the fastest possible gate incurs the least depolarization error
@@
 We enable the optimizer to trade longer gate times
-for lower depolarizaiton times or vice-versa by making the gate time a decision variable.
+for longer $T_{1}$ times, or shorter $T_{1}$ times for shorter gate times,
+by making the gate time a decision variable.
 Additionally, previous work has modeled the gate error due to depolarization
 by evolving density matrices under a master
-equation \cite{rembold2020introduction, schulteherbruggen2011optimal}
+equation \cite{rembold2020introduction, schulteherbruggen2011optimal},
 or evolving a large number of states in a quantum trajectory approach
@@
 techniques by penalizing the integrated depolarization rate in optimization.
-Using this metric as a proxy for the gate error incurred
-is reasonable because depolarization errors are incoherent and therefore
-increase monotonically in time without interference.
 
@@
 and its norm is penalized in the objective \eqref{eq:costfun} by setting
-the corresponding element of the final augmented state to zero.
+the corresponding element of the final augmented state to zero,
+which has the effect of incentivizing long $T_{1}$ times at each knot point.
 $T_{1}(a_{k})$ is obtained at each knot point by evaluating
@@
 see Figure \ref{fig:longitudeb}.
-It is also possible to fit a spline to theoretically obtained data.
-However, experimental data often deviates from theoretical predictions
-because it is difficult to account for all loss channels \cite{zhang2020universal},
-and $T_{1}$ values often fluctuate on hour
-time scales \cite{klimov2018fluctuations}.
-\todo{You could cut this - starting from "It is also" to end of paragraph -
-  no need to justify using experimental data if you've got it!
-  Maybe just cite where the data is coming from, if published.}
+Using the integrated depolarization rate as a proxy for the gate error incurred
+is reasonable because depolarization errors are incoherent and therefore
+increase monotonically in time without interference.
+Additionally, a master equation approach would require adding
+density matrices of size $n \times n$
+to the augmented state, and a quantum trajectory approach
+would require adding many states of size $n$ to the augmented state,
+where $n$ is the dimension of the Hilbert space.
+The integrated depolarization rate is a single
+real number; thus, the computational complexity of this
+depolarization model does not scale
+with the dimension of the Hilbert space.
 
@@
 making the time step between each knot point
-a decision variable. 
+a decision variable \cite{howell2019altro}. 
 The square root of the time step $\sqrt{\Delta t_{k}}$
@@
 the bounds of the time step at each knot point.
-We promote the time steps
-to decision variables, rather
-than the number of knot points $N$, to ensure each cost function depends
-only on the augmented state and augmented
-controls at its own knot point $\ell_{k}(x_{k}, u_{k})$, which is necessary for iLQR.
 
@@
 the $X/2$, $Y/2$, and $Z/2$ gates obtained with
-our numerical method and the analytic gates. We
+our numerical method and the corresponding analytic gates. We
 use the Lindblad master equation to simulate $T_{1}$ dissipation for successive
@@
 generated initial states.
+
 The flux pulses for the numerical gates
@@
 numerical flux pulses
-spend more time at higher amplitudes, achieving higher $T_{1}$ times.
+spend more time at higher flux values, achieving higher $T_{1}$ times on average,
+see Figure \ref{fig:longitudeb}.
 The single gate errors for both the analytic and numerical gates are
@@
 on the integrated depolarization rate metric, see Appendix \ref{appendix:longitude}.
-This single gate error advantage corresponds to a significant reduction in error correction resources
-\todo{this seems like a strong statement - citation?}.
-Furthermore, for successive gate applications, the gate error due to depolarization is approximately linear
+This single gate error advantage corresponds to a
+significant reduction in error correction resources
+\cite{paetznick2014resource, suchara2013comparing}.
+Furthermore, for successive gate applications,
+the gate error due to depolarization is approximately linear
 in the gate count, which we expect for $t \ll T_{1}$, see Figure \ref{fig:longitudec}.
-The gate error reduction for large gate counts is important for noisy, intermediate-scale quantum (NISQ)
-applications. These improvements are significant for the constraints we have imposed on the gates,
-and do not represent a fundamental limit to the optimization methods we have employed.
+The gate error reduction for large gate counts is
+important for noisy, intermediate-scale quantum (NISQ)
+applications. These improvements are significant for the constraints
+we have imposed on the gates,
+and do not represent a fundamental limit to the optimization
+methods we have employed.

### 2021-02-09 paper/5static.tex
resolved danny edits

--- 2021-02-05 paper/5static.tex
@@
 we mitigate errors arising from the drift and finite measurement
-precision of the qubit frequency which modifies the fluxonium Hamiltonian
+precision of the qubit frequency, which modifies the fluxonium Hamiltonian
 \eqref{eq:hamiltonian} by $f_{q} \rightarrow f_{q} + \delta f_{q}$.
@@
 We omit bra-ket notation here to emphasize that
-the states are real vectors, and are given by the right-hand-side of the
-isomorphism \eqref{eq:isomorphism}. Additionally,
-the unscented sampling method employs the unscented transform
+the states are real vectors and are given by the right-hand-side of the
+isomorphism \eqref{eq:isomorphism}.
+The sample states are chosen to encode a unimodal distribution over
+the $2n$ elements of the nominal state, modeling the uncertainty in the state
+as a result of the uncertainty in the parameter. We use the unscented transform
 \cite{julier2004unscented, uhlmann1995dynamic}
-to prevent the sample states
-from drifting too far or close to the nominal state during evolution,
-which would result in exceedingly large or small deviations in the cost function.
-In particular, the sample states are chosen to encode a unimodal distribution over
-the $2n$ elements of the nominal state, modeling the uncertainty in the state
-as a result of the uncertainty in the parameter. The unscented transform
-accurately propagates the mean and covariance of this distribution between
-knot points, or equivalently, through the transformation of the TDSE.
-For this method, we sample on one initial state $(\ket{0} + i\ket{1}) / \sqrt{2}$,
-and do not observe a performance increase
-using more initial states, for example those from the operator basis.
+to accurately propagate the mean and covariance of this distribution between
+knot points, or equivalently, through the transformation of the TDSE \eqref{eq:tdse}.
+Unlike the sampling method, the cost function for the unscented sampling method
+is sensitive to global phases. Accordingly, we
+do not observe a performance increase when
+using more than one initial state. 
 A detailed procedure for the unscented transformation is given
@@
 The derivative method penalizes the sensitivity of the state
-to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
-state derivative $\partial_{f_{q}}^{l} \ket{\psi}$. In the $m$\textsuperscript{th}-order
+to the uncertain parameter, which is encoded in the $m$\textsuperscript{th}-order
+state derivative $\ket*{\partial_{f_{q}}^{m} \psi} \equiv \partial_{f_{q}}^{m} \ket*{\psi}$.
+In the $m$\textsuperscript{th}-order
 derivative method, we append all state derivatives of order $1, \dots, m$
 to the augmented state vector \eqref{eq:astatecontrols}
-for each initial state. We use the initial state $\ket{0}$ for this method,
-and observe no advantage to using more initial states.
-We penalize the norms of the state derivatives
-in the objective \eqref{eq:costfun} by setting the corresponding elements
-of the final augmented state to zero.
-We could obtain the state derivatives at each knot point
-with backward-mode differentiation.
-In a naive automatic differentiation scheme,
-the discrete dynamics function at knot points
-$1, \dots, k - 1$ would be differentiated to obtain the state
-derivative at knot point $k$, requiring
-$O(N^{2})$ matrix multiplications. Instead, we 
-employ forward-mode differentiation on the TDSE \eqref{eq:tdse}
-to obtain coupled, first-order ODEs
-which require $O(N)$ matrix multiplications to integrate.
+for each initial state.
+We obtain the state derivatives at each knot point by performing forward-mode
+differentiation on the TDSE \eqref{eq:tdse}.
 For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
@@
 We integrate the coupled ODEs with exponential
-integrators, see Appendix \ref{appendix:derivative},
-in the discrete dynamics function \eqref{eq:dyn_con}. For runtimes
-of the three robust control techniques,
+integrators in the discrete dynamics function \eqref{eq:dyn_con},
+see Appendix \ref{appendix:derivative}.
+While the state $\ket*{\psi}$ has unit norm,
+the state derivatives $\ket*{\partial^{m}_{f_{q}} \psi}$ need not, as is evident
+from the non-unitary dynamics \eqref{eq:d1dyn}.
+We penalize the norms of the state derivatives
+in the objective \eqref{eq:costfun} by setting the corresponding elements
+of the final augmented state to zero. Intuitively, this corresponds to penalizing
+the sensitivity of each state element to the uncertain parameter. As was the case for
+the unscented sampling
+method, we do not observe a performance increase when using more than one initial state
+for the derivative method.
+For runtimes of the three robust control techniques,
 consult Appendix \ref{appendix:time}.
@@
 in the detuning.
-At a one-percent
-detuning ($\lvert \delta f_{q} \rvert / f_{q} = 1\%$),
-the gate error is $\sim 4.5 \cdot 10^{-5}$,
+At a one-percent detuning ($\lvert \delta f_{q} \rvert / f_{q} = 1\%$),
+the gate error is $\sim 4.7 \cdot 10^{-5}$,
 which is sufficient for quantum error correction.
-Although the gate performs well, it cannot be extended
-to gate times other than
-$1 / 4 f_{q}$. The ability to perform
-phase gates in any given time is critical
-for multi-qubit experiments, where the qubits operate at different
-frequencies $f_{q, i} \neq f_{q, j}$.
-We can find solutions using the numerical methods at
-all gate times above $18$ns, see Figure \ref{fig:staticb}.
-These numerical methods offer an effective scheme for synchronizing
-multi-qubit experiments.
 
-The flux pulse produced by the sampling method
-transitions  smoothly away from idles at the flux frustration point,
-whereas that for the unscented sampling method employs
-three distinct transitions.
-The gate error at a one-percent qubit frequency detuning for the sampling method
-does not improve substantially over the
-range of gate times. Conversely, the gate error at a one-percent detuning
-for the unscented sampling method reaches a minimum of $\sim 3.9 \cdot 10^{-5}$
-near fractions of the Larmor period $2/4f_{q} \sim 36\textrm{ns},
-3/4f_{q} \sim 54\textrm{ns}, 4/4f_{q} \sim 72\textrm{ns}$.
-
-The two derivative methods converge on qualitatively similar flux pulses that
-use fast triangle pulses at the boundaries and idle near the flux frustration point,
-similar to the flux pulse produced by the unscented sampling method.
-The gate errors at a one-percent qubit frequency detuning
-for both derivative methods decrease super-linearly in their gate times.
-The 2\textsuperscript{nd}-order method does not offer a substantial improvement
-over the 1\textsuperscript{st}-order method for most gate times.
-The gate error at a one-percent detuning for the 1\textsuperscript{st}-order
-method reaches $10^{-7}$ at the Larmor period $1 / f_{q} \sim 72$ns,
-see Figure \ref{fig:staticc}.
-This result mimics the
-ability of composite pulses to mitigate parameter uncertainty errors to arbitrary
-order with sufficiently many pulses \cite{merrill2014progress}.
-It is difficult to choose an appropriate composite pulse
-for the problem studied here due to our Hamiltonian and experimental constraints.
-We propose comparisons between composite pulses and numerical techniques
-for future work.
+For the sampling method, the gate error at a one-percent qubit frequency detuning
+does not decrease substantially over the
+range of gate times, and begins to increase above $5 \cdot 10^{-5}$ for gate times
+greater than $\sim 50$ns, see Figure \ref{fig:staticb}.
+Optimization results for the sampling method reveal that it is typically
+able to achieve a high fidelity for
+one sample $\ket{\psi^{\pm}}$,
+but not the other $\ket{\psi^{\mp}}$, indicating that it is difficult for the optimizer
+to make progress on both objectives.
+For the unscented sampling method,
+the gate error at a one-percent detuning
+does not decrease substantially 
+over the gate times, but it does reach
+a minimum of $\sim 3.9 \cdot 10^{-5}$
+near fractions of the Larmor period $2/4f_{q} \sim 36\textrm{ns}$,
+$3/4f_{q} \sim 54\textrm{ns}$, $4/4f_{q} \sim 72\textrm{ns}$.
 
@@
   \caption{
-    (a) Flux pulses for $X/2$ gates robust to flux offsets
+    (a) Flux pulses for $X/2$ gates robust to flux noise
     constructed with the analytic (A),
@@
 \end{figure*}
+
+The two derivative methods converge on qualitatively similar flux pulses that
+idle near the flux frustration point and use fast triangle movements at the boundaries,
+similar to the flux pulse produced by the unscented sampling method.
+For both methods, the gate error at a one-percent qubit frequency detuning
+decreases super-linearly in the gate time.
+For the 1\textsuperscript{st}-order method, the gate error at a one-percent detuning 
+reaches $10^{-7}$ at the Larmor period $1 / f_{q} \sim 72$ns,
+see Figure \ref{fig:staticc}.
+This result mimics the
+ability of composite pulses to mitigate parameter uncertainty errors to arbitrary
+order with sufficiently many pulses \cite{merrill2014progress}.
+It is difficult to choose an appropriate composite pulse
+for the problem studied here due to our Hamiltonian and experimental constraints.
+We propose comparisons between composite pulses and numerical techniques
+for future work.
+
+Furthermore, the ability to perform
+$Z$-type gates in any given time is critical
+for synchronizing phases in multi-qubit experiments,
+where the qubits have distinct
+frequencies. Notably, the analytic gate studied here cannot be extended
+to gate times other than $1 / 4 f_{q}$. 
+We can find gates using the numerical methods at
+all gate times above $18$ns, see Figure \ref{fig:staticb}.
+These numerical methods offer an effective scheme for synchronizing
+multi-qubit experiments.

### 2021-02-09 paper/6stochastic.tex
resolved danny edits

--- 2021-02-05 paper/6stochastic.tex
@@
 parameter uncertainty. For many flux-biased and inductively-coupled
-superconducting circuit elements, magnetic flux noise is a significant
-source of coherent errors. Flux noise
+superconducting circuit elements, magnetic flux noise is the dominant
+source of coherent errors \cite{bialczak20071f, kakuyanagi2007dephasing,
+  kumar2016origin, yoshihara2006decoherence}. Flux noise
 modifies the fluxonium Hamiltonian \eqref{eq:hamiltonian}
-by $a(t) \rightarrow a(t) + \delta a(t)$.
-The spectral density of the flux noise $\delta a(t)$ is observed to
+by $a(t) \rightarrow a(t) + \delta a(t)$ where $\delta a(t)$ is the flux noise.
+The spectral density of flux noise is observed to
 follow a 1/$f$ distribution
-\cite{bialczak20071f, koch2007model,
-  yoshihara2006decoherence, yoshihara2010correlated},
+\cite{bialczak20071f, koch2007model, kakuyanagi2007dephasing,
+  kumar2016origin, yoshihara2006decoherence, yoshihara2010correlated,
+  zhang2020universal},
 so the noise is dominated by low-frequency components.
@@
 
-Additionally, we modify the robust control techniques presented
+We modify the robust control techniques presented
 in the previous section to combat 1/$f$ flux noise.
@@
 and the analytic $X/2$ gate. To compute the gate error,
-an initial state is evolved
+we evolve an initial state
 under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
 where the optimized flux is modified $a(t) \rightarrow a(t) + \delta a(t)$.
-The flux noise is generated as
+We generate the flux noise as
 we described for the unscented sampling method.
@@
 To observe the effect of interfering coherent errors,
-we simulate successive applications of the gate constructed by each method
-and compute the cumulative gate error
+we simulate successive applications of the gate constructed by each method;
+we compute the cumulative gate error
 after each application, see Figure \ref{fig:stochastic}.
@@
 two orders of magnitude less than the gate error produced by the analytic gate.
-1/$f$ flux noise is a significant source of coherent errors in NISQ applications and
-these numerical techniques offer effective avenues to mitigate it.
+1/$f$ flux noise is a significant source of coherent errors in NISQ applications,
+and these numerical techniques offer effective avenues to mitigate it.

### 2021-02-09 paper/7conclusion.tex
jens edits 2 addressed or todo'd
resolved danny edits

@@
 depolarization with time-optimal
-control and the depolarization probability model.
+control and the integrated depolarization rate model.
 The computational cost of this model is
 independent of the dimension of the Hilbert space, enabling
-inexpensive optimization on high-dimensional quantum systems (hmm, this seems like something that definitely wasn't emphasized in the text - could be worth talking about this a little earlier.
-Also doesn't seem totally true right?
-Re Eq.(5)).
+inexpensive optimization on high-dimensional quantum systems.
 We have also proposed the derivative method for robust control which achieves

### 2021-02-09 paper/a1longitude.tex
resolved danny edits

@@
 anti-commutator.
-For depolarization $\gamma_{\uparrow} = T_{1, \uparrow}^{-1}$,
-$\gamma_{\downarrow} = T_{1, \downarrow}^{-1}$,
-$L_{\uparrow} = \sigma^{+}/2$, and
-$L_{\downarrow} = \sigma^{-}/2$ where $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$.
-Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained at each knot point
+For depolarization $\gamma_{+} = T_{+}^{-1}$,
+$\gamma_{-} = T_{-}^{-1}$,
+$L_{+} = \sigma^{+}/2$, and
+$L_{-} = \sigma^{-}/2$ where $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$.
+\todo{Ask Jens and company to okay the rest of this paragraph}
+$T_{+} = T_{-} = T_{1}$ are obtained at each knot point
 from the spline shown in Figure \ref{fig:longitudeb}.
 We obtain the $T_{1}$ values in this spline
-by driving the qubit at the desired flux amplitude
+by driving the qubit at the desired flux bias
 and monitoring the resultant decay. For more details
 on these measurements, consult \cite{zhang2020universal}.
+Because $T_{1}$ depends on the flux, so do
+the decay rates, $\gamma_{i}$.
+Integrating the master equation with time-dependent decay crates
+provides a heuristic for how the gates we study may perform in
+the experiment, but it may not be strictly correct in the regime we
+are operating in, where the decay rates change much faster
+than the relaxation time, $\Delta t_{k} \ll T_{1}$. Standard
+derivations of the Lindblad master equation do not account for
+time-dependent decay rates \cite{manzano2020a}, and a more thorough
+treatment in future work would unlock new insights for
+quantum computing platforms where decoherence is strongly
+dependent on the control parameters.
 
@@
 \end{equation}
-where $\rho = \sum_{i, j} \alpha_{i, j} \ket{i}\bra{j}$
-and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{i, j} \ket{i} \otimes \ket{j}$.
-We use zero-order hold on the controls--equivalently, $a(t)$ is piecewise-constant
-on the intervals $[t_{k}, t_{k + 1}]$. Therefore,
-$H(a(t))$, $L(a(t))$, and $\gamma_{i}(a(t))$ are also piecewise-constant.
+where $\rho = \sum_{i, j} \alpha_{ij} \ket{i}\bra{j}$
+and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{ij} \ket{i} \otimes \ket{j}$.
+We use zero-order hold on the controls; equivalently, $a(t)$ is constant
+on each integration interval $[t_{k}, t_{k + 1}]$. Therefore,
+$H(a(t))$, and $\gamma_{i}(a(t))$ are also piecewise-constant.
 So, the exact solution is
-$\textrm{vec}(\rho_{k + 1}) = {\exp}{\textstyle(}\Delta t_{k} \hat{\mathcal{L}}_{k}{\textstyle)} \textrm{vec}(\rho_{k})$.
+$\textrm{vec}(\rho_{k + 1}) = {\exp}{\textstyle(}\Delta t_{k}
+\hat{\mathcal{L}}_{k}{\textstyle)} \textrm{vec}(\rho_{k})$.
 This isomorphism transforms $(n \times n) \times (n \times n)$

### 2021-02-09 paper/a3derivative.tex
resolved danny edits

--- 2021-02-05 paper/a3derivative.tex
@@
 General exponential integrators break the dynamics into a linear term and a non-linear term.
-For example, the dynamics of the first state derivative in
-units of $i\hbar = 1$ are $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
-H \ket{\partial_{\lambda} \psi} + (\partial_{\lambda} H) \ket{\psi}$.
-The linear term is $L = H$ and the non-linear term is $N = (\partial_{\lambda} H) \ket{\psi}$.
+For example, the dynamics for the first state derivative
+are $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
+-\frac{i}{\hbar} H  \ket{\partial_{\lambda} \psi} -
+\frac{i}{\hbar} (\partial_{\lambda} H) \ket{\psi}$.
+The linear term is $L = -\frac{i}{\hbar} H$ and the non-linear
+term is $N = -\frac{i}{\hbar} (\partial_{\lambda} H) \ket{\psi}$.
 With zero-order hold on the controls the exact solution is:
@@
   \begin{aligned}
-    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(\Delta t_{k}L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
+    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(\Delta t_{k}L_{k})
+    \ket{\partial_{\lambda} \psi_{k}}\\
     &+ \exp(\Delta t_{k}L_{k}) N_{k} \Delta t_{k}.\\

### 2021-02-23 paper/0abstract.tex
resolved danny edits
updated bib and added brian edits 3

--- 2021-02-05 paper/0abstract.tex
@@
 %% TITLE
-\title{Robust Trajectory Optimization on a Quantum System}
+\title{Robust Quantum Optimal Control with Trajectory Optimization}
 
@@
 }
-\author{Brian Jackson}
+\author{Brian E. Jackson}
 \author{Zachary Manchester}
@@
   We employ a computationally efficient model
-  and utilize time-optimal control to achieve high fidelity gates in the presence of depolarization.
+  and utilize time-optimal control to achieve high-fidelity gates in the presence of depolarization.
   We apply these techniques to a fluxonium qubit and suppress

### 2021-02-23 paper/2background.tex
updated bib and added brian edits 3

@@
 dynamics function $f(x_{k}, u_{k})$ propagates the state by
-integrating the TDSE \ref{eq:tdse} with Runge-Kutta methods \cite{jorgensen2011numerical}
-or exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance,
+integrating the TDSE \eqref{eq:tdse} with Runge-Kutta methods \cite{jorgensen2011numerical}
+or exponential integrators \cite{auer2018magnus, berland2006solving, einkemmer2017performance,
 shillito2020fast}.
@@
 functions may be vector-valued to encode multiple constraints of each type.
-If a constraint function takes a positive value, the
-value is called the constraint's violation.
-The trajectory optimization problem can be stated concisely,
+The constraint function's violation is the magnitude by which it is not satisfied:
+$\textrm{max}(g(\cdot), 0)$ or $\textrm{max}(\abs{h(\cdot)})$
+where $\textrm{max}(\cdot)$ and $\abs{\cdot}$ act element-wise.
+Stated concisely, the trajectory optimization problem is,
 \begin{mini!}[2]
-    {x_{1:N},u_{1:N-1}}{\ell_{N}(x_N) + \sum_{k=1}^{N-1} \ell_{k}(x_k,u_k)}{}{} \label{eq:gcostfun}
+  {x_{1:N},u_{1:N-1}}{\ell_{N}(x_N) + \sum_{k=1}^{N-1}
+    \ell_{k}(x_k,u_k)}{}{} \label{eq:gcostfun}
     \addConstraint{x_{k+1} = f(x_k,u_k) \ \forall \ k} \label{eq:gdyncon}
     \addConstraint{g_k(x_k,u_k) \leq 0 \ \forall \ k} \label{eq:ineqcon}
-    \addConstraint{h_k(x_k,u_k) = 0 \ \forall \ k} \label{eq:eqcon}.
+    \addConstraint{h_k(x_k,u_k) = 0 \ \forall \ k} \label{eq:eqcon}
 \end{mini!}

### 2021-02-23 paper/3fluxonium.tex
updated bib and added brian edits 3

@@
   zhang2020universal}.
-In this section, we outline the base optimization problem,
+In this section, we outline the base optimization problem \eqref{eq:costfun}-\eqref{eq:ic_con},
 which we extend in subsequent sections to account
@@
 All gates presented in this work satisfy these constraints to
-a tolerance of $\sim 10^{-8}$.
+a maximum violation of $\sim 10^{-8}$.
 Casting this problem as a multi-state transfer, the initial conditions on
@@
 \eqref{eq:amp_con}.
-Above $0.5$ GHz, we observe excitation out of the first two levels, disallowing the
+Above $0.5$ GHz, we observe population beyond the first two levels, disallowing the
 approximation \eqref{eq:hamiltonian}.
@@
 $\ell_{k}(x_{k}, u_{k}) = (x_{k} - x_{f})^{T} Q_{k} (x_{k} - x_{f}) + u^{T}_{k} R_{k} u_{k}$
-where $Q_{k}$ and $R_{k}$ are diagonal matrices we supply.
+where $Q_{k}$ and $R_{k}$ are positive-definite diagonal matrices we supply.
 The $Q_{k}$ term
@@
 Furthermore, the target state constraint \eqref{eq:tstate_con} says
-that the final state must match the target state, including its phase,
-up to the constraint tolerance $\sim 10^{-8}$.
+that the final state must match the target state, including its global phase,
+up to the maximum constraint violation $\sim 10^{-8}$.
 If we did not make target state achievement a constraint, the optimizer would
@@
 Additionally, we chose the target state constraint to be
-phase-sensitive because the Hessian of the constraint function is diagonal,
-and therefore computationally efficient, and we wish to optimize $Z/2$ gates,
-which require a phase-sensitive metric for the initial states $\ket{0}$ and $\ket{1}$.
+global phase-sensitive because the Hessian of the constraint function is diagonal--and
+therefore computationally efficient--and we wish to optimize $Z/2$ gates,
+which require a global phase-sensitive metric for the initial states $\ket{0}$ and $\ket{1}$.

### 2021-02-23 paper/4longitude.tex
updated bib and added brian edits 3

@@
 and its norm is penalized in the objective \eqref{eq:costfun} by setting
-the corresponding element of the final augmented state to zero,
-which has the effect of incentivizing long $T_{1}$ times at each knot point.
+the corresponding element of the final augmented state to zero.
 $T_{1}(a_{k})$ is obtained at each knot point by evaluating

### 2021-02-23 paper/5static.tex
updated bib and added brian edits 3

@@
 the states are real vectors and are given by the right-hand-side of the
-isomorphism \eqref{eq:isomorphism}.
+complex-to-real isomorphism \eqref{eq:isomorphism}.
 The sample states are chosen to encode a unimodal distribution over
@@
 The derivative method penalizes the sensitivity of the state
-to the uncertain parameter, which is encoded in the $m$\textsuperscript{th}-order
-state derivative $\ket*{\partial_{f_{q}}^{m} \psi} \equiv \partial_{f_{q}}^{m} \ket*{\psi}$.
+to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
+state derivative $\ket*{\partial_{f_{q}}^{l} \psi} \equiv \partial_{f_{q}}^{l} \ket*{\psi}$.
 In the $m$\textsuperscript{th}-order
@@
 in the detuning.
-At a one-percent detuning ($\lvert \delta f_{q} \rvert / f_{q} = 1\%$),
+At a one-percent detuning ($\lvert \delta f_{q} / f_{q} \rvert = 1\%$),
 the gate error is $\sim 4.7 \cdot 10^{-5}$,

### 2021-02-23 paper/8ack.tex
updated bib and added brian edits 3

@@
   This work was supported by the Army Research Office under Grant No. W911NF1910016.
-  \todo{funding}.
-  \todo{cite software packages}.
+  This work was made possible by many open source software projects,
+  including but not limited to:
+  Altro.jl \cite{howell2019altro},
+  DifferentialEquations.jl \cite{rackauckas2017differentialequations},
+  Distributions.jl \cite{besancon2019distributions},
+  ForwardDiff.jl \cite{revelsLubinPapamarkou2016},
+  Matplotlib \cite{hunter2007matplotlib},
+  NumPy \cite{harris2020array},
+  and Zygote.jl \cite{innes2018don}.
 \end{acknowledgments}

### 2021-02-23 paper/a1longitude.tex
updated bib and added brian edits 3

@@
     \hline
-    Z/2 & 5.745 & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
-    Y/2 & 5.253 & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
-    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
+    Z/2 & 5.745 & 1.149 & 5.000 & 0.888   & 0.185  & 4.791 \\
+    Y/2 & 5.253 & 1.157 & 4.540 & 0.770 & 0.186   & 4.132\\
+    X/2 & 16.251 & 2.660 & 6.109 & 2.674 & 0.432  & 6.200\\
   \end{tabular}
@@
 anti-commutator.
-For depolarization $\gamma_{+} = T_{+}^{-1}$,
-$\gamma_{-} = T_{-}^{-1}$,
-$L_{+} = \sigma^{+}/2$, and
-$L_{-} = \sigma^{-}/2$ where $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$.
-\todo{Ask Jens and company to okay the rest of this paragraph}
-$T_{+} = T_{-} = T_{1}$ are obtained at each knot point
+For depolarization, $\gamma_{\pm} = T_{\pm}^{-1}$,
+$L_{\pm} = \sigma^{\pm} \equiv (\sigma_{x} \pm i \sigma_{y})/2$.
+The depolarization times $T_{+} = T_{-} = 2 T_{1}$ are obtained at each knot point
 from the spline shown in Figure \ref{fig:longitudeb}.
@@
 on these measurements, consult \cite{zhang2020universal}.
-Because $T_{1}$ depends on the flux, so do
-the decay rates, $\gamma_{i}$.
-Integrating the master equation with time-dependent decay crates
+Because $T_{1}$ depends on the flux $a(t)$, so do
+the decay rates $\gamma_{i}$.
+Integrating the master equation with time-dependent decay rates
 provides a heuristic for how the gates we study may perform in
-the experiment, but it may not be strictly correct in the regime we
-are operating in, where the decay rates change much faster
-than the relaxation time, $\Delta t_{k} \ll T_{1}$. Standard
+the experiment. 
+This procedure may not be strictly correct
+when the decay rates change much faster
+than the relaxation time $\Delta t_{k} \ll T_{1}$,
+which is the regime we are operating in. Standard
 derivations of the Lindblad master equation do not account for
-time-dependent decay rates \cite{manzano2020a}, and a more thorough
-treatment in future work would unlock new insights for
+time-dependent decay rates \cite{manzano2020a}. A more thorough
+treatment of this regime in future work would unlock new insights for
 quantum computing platforms where decoherence is strongly
@@
 on each integration interval $[t_{k}, t_{k + 1}]$. Therefore,
-$H(a(t))$, and $\gamma_{i}(a(t))$ are also piecewise-constant.
+$H(a(t))$ and $\gamma_{i}(a(t))$ are also piecewise-constant.
 So, the exact solution is

### 2021-02-23 paper/a4time.tex
resolved danny edits
updated bib and added brian edits 3

@@
     using the base, depolarization, sampling (S),
-    unscented Sampling (U), and the 1\textsuperscript{st}-
+    unscented sampling (U), and the 1\textsuperscript{st}-
     and 2\textsuperscript{nd}-order derivative methods (D1, D2).
@@
 initial states in the operator basis, $dm$ state derivatives per initial state,
-and each state has $2n$ real numbers. \todo{maybe add comment about auto-diff}
+and each state has $2n$ real numbers.

### 2021-03-05 paper/2background.tex
updates to paper

@@
 \cite{Li2004a} as the internal solver of an augmented Lagrangian method (ALM)
-\cite{lantoine2012hybrid, plancher2017constrained}
+\cite{lantoine2012hybrid, plancher2017constrained, nocedal2006numerical}
 and employs a projected Newton method \cite{bertsekas1982projected, rao1998application}

### 2021-03-05 paper/3fluxonium.tex
updates to paper

@@
-%% F1
-\begin{figure*}[ht]
-  \begin{subfigure}{.315\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1a.png}
-    \caption{\label{fig:longitudea}}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.23\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1b.png}
-    \caption{\label{fig:longitudeb}}
-  \end{subfigure}\hfill
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f1c.png}
-    \caption{\label{fig:longitudec}}
-  \end{subfigure}
-  \caption{
-    (a) Flux pulses for the numerical gates (dark blue)
-    and the analytic gates (light pink).
-    (b) $T_{1}$ interpolation function used in optimization. Circle markers
-    indicate measured $T_{1}$ times. Non-circle markers
-    are plotted at the time-averaged 
-    absolute flux and the time-averaged $T_{1}$ time for each pulse.
-    (c) Cumulative gate errors due to depolarization as a function of the
-    number of gates applied.
-    Cumulative gate errors for the numerical $Z/2$ and $Y/2$ gates
-    are indistinguishable. Inset shows log-scaled cumulative gate errors
-    for small gate counts.
-  }
-  \label{fig:longitude}
-\end{figure*}
-
 \section{QOC for the Fluxonium \label{sec:fluxonium}}
@@
   zhang2020universal}.
-In this section, we outline the base optimization problem \eqref{eq:costfun}-\eqref{eq:ic_con},
+In this section, we use the trajectory optimization
+formalism \eqref{eq:gcostfun}-\eqref{eq:ineqcon}
+to define the problem \eqref{eq:costfun}-\eqref{eq:ic_con},
 which we extend in subsequent sections to account
@@
 The states at the final knot point are constrained to be
-the target states $\ket*{\psi^{i}_{N}} = U \ket*{\psi^{i}_{1}} \ \forall \ i$
+the target states $\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{f}} \equiv U
+\ket*{\psi^{i}_{1}} \ \forall \ i$
 \eqref{eq:tstate_con} where $U = X/2, Y/2, Z/2$ is the desired gate.
@@
 
-Next, we introduce the augmented control and augmented state:
+The augmented control and augmented state are:
 \begin{equation}
@@
 \end{equation}
-The variables in the augmented state are obtained from the decision variable
-in the augmented control
-via coupled, first-order, differential equations, which are integrated in the 
-discrete dynamics function \eqref{eq:dyn_con}.
+The time evolution of these variables are encoded
+in coupled, first-order, differential equations
+which are integrated in the discrete dynamics function \eqref{eq:dyn_con}.
 We integrate the states according to the TDSE \eqref{eq:tdse} and the
@@
 The $Q_{k}$ term
-penalizes deviations from the final augmented state $x_{f}$,
+penalizes deviations from the target augmented state $x_{f}$,
 which is given by the constraints we have imposed on
@@
   \addConstraint{\ket{\psi^{0}_{1}} = \ket{0}, \ket{\psi^{1}_{1}} = \ket{1}} \label{eq:istate_con}
-  \addConstraint{\ket{\psi^{i}_{N}} = U \ket{\psi^{i}_{1}}
+  \addConstraint{\ket{\psi^{i}_{N}} = \ket{\psi^{i}_{f}}
     \ \forall \  i} \label{eq:tstate_con}
@@
 
-Lastly, we remark on decisions we have made in the problem definition.
-We penalize deviations from the target state at all knot points \eqref{eq:costfun}
-because it benefits the iLQR solving stage \cite{Jackson2020altroc}.
-This choice is unrelated to schemes
-for incentivizing early achievement of the desired gate \cite{leung2017speedup}.
-In practice, the penalty for deviations from the target state
-at the final knot point ($k = N$) becomes large relative to those at
-intermediate knot points ($k = 1, \dots, N - 1$)
-because ALM increases the penalty at the final knot point
-to satisfy the target state constraint \eqref{eq:tstate_con}.
-Furthermore, the target state constraint \eqref{eq:tstate_con} says
-that the final state must match the target state, including its global phase,
-up to the maximum constraint violation $\sim 10^{-8}$.
-If we did not make target state achievement a constraint, the optimizer would
-be allowed to sacrifice the nominal gate error to achieve more depolarization protection,
-or more robustness to parameter uncertainties, which is undesirable.
-Additionally, we chose the target state constraint to be
-global phase-sensitive because the Hessian of the constraint function is diagonal--and
-therefore computationally efficient--and we wish to optimize $Z/2$ gates,
-which require a global phase-sensitive metric for the initial states $\ket{0}$ and $\ket{1}$.
+Next, we remark on our problem formulation.
+We penalize the squared error of the final state and the target
+state \eqref{eq:costfun}, rather than their infidelity,
+because the Hessian of the squared error cost function is diagonal--which
+makes matrix multiplications fast--and we wish to optimize $Z/2$ gates,
+which requires a metric that is sensitive to global phases for the initial
+states $\ket{0}$ and $\ket{1}$.
+We penalize the squared error at all knot points
+because it benefits the iLQR solving stage \cite{Jackson2020altroc},
+not to incentivize early achievement
+of the desired gate \cite{leung2017speedup}.
+
+In addition to the target state cost function, we also
+impose the target state constraint \eqref{eq:tstate_con}, which
+says that the final state must match the target state, including its global phase,
+up to our chosen maximum constraint violation $\sim 10^{-8}$.
+If we did not impose this constraint, the optimizer would be
+allowed to sacrifice the closed-system gate error to achieve better
+performance on the depolarization or robustness cost functions we will
+introduce in the following sections, which is undesirable.
+To enforce a constraint in standard QOC frameworks,
+the prefactor for the constraint function is manually increased
+between separate optimization instances until the constraint is satisfied
+\cite{heeres2017implementing, leung2017speedup, reinhold2019controlling},
+which becomes infeasible for more than one constraint.
+ALM automates these prefactor updates to find
+a solution that satisfies all of the constraints.
+Hence, ALTRO's ability to handle multiple constraints makes it
+an attractive solver for robust QOC problems.
+
+In extraordinarily difficult cases of
+QOC \cite{abdelhafez2020universal}, it may be impossible
+to obey the physics of the system and achieve the desired gate,
+equivalently, the dynamics constraint \eqref{eq:dyn_con}
+and the target state constraint \eqref{eq:tstate_con} may be mutually unsatisfiable.
+In this case, the prefactors for the constraint function terms
+in the ALM objective will tend to infinity--leading to numerical instability--and the
+optimization will not converge. To maintain a constrained approach in this situation,
+the maximum constraint violation for the target state constraint can be raised
+to a level commensurate with the minimum acceptable gate error.
+
+Finally, for the indirect ALM-iLQR solving stage,
+the state at each knot point is obtained by integrating
+the TDSE, so the dynamics constraint \eqref{eq:dyn_con} is satisfied by construction.
+This is not the case for the direct projected Newton stage, where the states
+are free parameters that are adjusted to satisfy the TDSE. However,
+the solution's deviation from the TDSE is never more than the maximum constraint violation.
+When we report gate errors in this text, we explicity integrate the TDSE with the
+flux produced by the optimization to ensure accuracy. Exploring the benefit
+of direct optimization approaches for QOC is an interesting direction for future work.

### 2021-03-05 paper/4longitude.tex
updates to paper

@@
+%% F1
+\begin{figure*}[ht]
+  \begin{subfigure}{.315\textwidth}
+    \includegraphics[width=\linewidth]{assets/f1a.png}
+    \caption{\label{fig:longitudea}}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.23\textwidth}
+    \includegraphics[width=\linewidth]{assets/f1b.png}
+    \caption{\label{fig:longitudeb}}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.4\textwidth}
+    \includegraphics[width=\linewidth]{assets/f1c.png}
+    \caption{\label{fig:longitudec}}
+  \end{subfigure}
+  \caption{
+    (a) Flux pulses for the numerical gates (dark blue)
+    and the analytic gates (light pink).
+    (b) $T_{1}$ interpolation function used in optimization. Circle markers
+    indicate measured $T_{1}$ times. Non-circle markers
+    are plotted at the time-averaged 
+    absolute flux and the time-averaged $T_{1}$ time for each pulse.
+    (c) Cumulative gate errors due to depolarization as a function of the
+    number of gates applied.
+    Cumulative gate errors for the numerical $Z/2$ and $Y/2$ gates
+    are indistinguishable. Inset shows log-scaled cumulative gate errors
+    for small gate counts.
+  }
+  \label{fig:longitude}
+\end{figure*}
+
 \section{Depolarization Mitigation\label{sec:longitude}}
@@
 and its norm is penalized in the objective \eqref{eq:costfun} by setting
-the corresponding element of the final augmented state to zero.
+the corresponding element of the target augmented state to zero.
 $T_{1}(a_{k})$ is obtained at each knot point by evaluating

### 2021-03-05 paper/5static.tex
updates to paper

@@
 by adding a cost function to the objective \eqref{eq:costfun} of the form
-$\sum_{k, \pm} b_{k} (1 - {\lvert \bra*{\psi^{\pm}_{k}}U\ket*{\psi_{1}} \rvert}^{2})$
+$\sum_{k, \pm} b_{k} (1 - {\lvert \braket*{\psi^{\pm}_{k}}{\psi_{f}} \rvert}^{2})$
 where $b_{k}$ is a constant we supply.
@@
 in the objective \eqref{eq:costfun} by setting the corresponding elements
-of the final augmented state to zero. Intuitively, this corresponds to penalizing
+of the target augmented state to zero. Intuitively, this corresponds to penalizing
 the sensitivity of each state element to the uncertain parameter. As was the case for
@@
 
-For the sampling method, the gate error at a one-percent qubit frequency detuning
-does not decrease substantially over the
-range of gate times, and begins to increase above $5 \cdot 10^{-5}$ for gate times
-greater than $\sim 50$ns, see Figure \ref{fig:staticb}.
-Optimization results for the sampling method reveal that it is typically
-able to achieve a high fidelity for
-one sample $\ket{\psi^{\pm}}$,
-but not the other $\ket{\psi^{\mp}}$, indicating that it is difficult for the optimizer
-to make progress on both objectives.
-For the unscented sampling method,
-the gate error at a one-percent detuning
-does not decrease substantially 
-over the gate times, but it does reach
-a minimum of $\sim 3.9 \cdot 10^{-5}$
-near fractions of the Larmor period $2/4f_{q} \sim 36\textrm{ns}$,
-$3/4f_{q} \sim 54\textrm{ns}$, $4/4f_{q} \sim 72\textrm{ns}$.
-
 %% F3
@@
 \end{figure*}
+
+For the sampling method, the gate error at a one-percent qubit frequency detuning
+does not decrease substantially over the
+range of gate times, and begins to increase above $5 \cdot 10^{-5}$ for gate times
+greater than $\sim 50$ns, see Figure \ref{fig:staticb}.
+Optimization results for the sampling method reveal that it is typically
+able to achieve a high fidelity for
+one sample $\ket{\psi^{\pm}}$,
+but not the other $\ket{\psi^{\mp}}$, indicating that it is difficult for the optimizer
+to make progress on both objectives.
+For the unscented sampling method,
+the gate error at a one-percent detuning
+does not decrease substantially 
+over the gate times, but it does reach
+a minimum of $\sim 3.9 \cdot 10^{-5}$
+near fractions of the Larmor period $2/4f_{q} \sim 36\textrm{ns}$,
+$3/4f_{q} \sim 54\textrm{ns}$, $4/4f_{q} \sim 72\textrm{ns}$.
 

### 2021-03-13 paper/0abstract.tex
dave comments up to sec iv and readme updates

@@
   The ability to engineer high-fidelity gates on quantum processors in the presence of
-  systematic errors remains the primary challenge requisite to achieving quantum advantage.
+  systematic errors remains the primary barrier to achieving quantum advantage.
   Quantum optimal control methods have proven effective in experimentally

### 2021-03-13 paper/2background.tex
dave comments up to sec iv and readme updates

@@
 To make the problem numerically tractable,
-the controls and state are discretized into $N$ knot points (time steps).
+the controls and state are discretized into $N$ time steps.
 In the case of a single state-transfer problem, the functional is
-the infidelity of the desired final state and the initial state evolved
-to the final knot point,
+the infidelity of the target state and the initial state evolved
+to the final time step,
 $J(u) = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u)} \rvert}^{2}$.
@@
 The functional $J(u)$ is divided into its constituent
-cost functions at each knot point $\ell_{k}(x_{k}, u_{k})$,
-where $k \in \{1, ..., N\}$ denotes the knot point,
+cost functions at each time step $\ell_{k}(x_{k}, u_{k})$,
+where $k \in \{1, ..., N\}$ denotes the time step,
 $x_{k}$ is the augmented state vector,
@@
     \addConstraint{g_k(x_k,u_k) \leq 0 \ \forall \ k} \label{eq:ineqcon}
-    \addConstraint{h_k(x_k,u_k) = 0 \ \forall \ k} \label{eq:eqcon}
+    \addConstraint{h_k(x_k,u_k) = 0 \ \forall \ k.} \label{eq:eqcon}
 \end{mini!}
@@
 about the current augmented controls and augmented states.
-These models are used to derive a recurrence relation between knot points
+These models are used to derive a recurrence relation between time steps
 which gives the locally optimal update for the augmented controls.

### 2021-03-13 paper/3fluxonium.tex
dave comments up to sec iv and readme updates

@@
 In the following, we optimize quantum gates
-for the superconducting fluxonium qubit--a promising
+(unitary transformations) for the superconducting fluxonium qubit--a promising
 building block for quantum computers due to its high
@@
 In this section, we use the trajectory optimization
-formalism \eqref{eq:gcostfun}-\eqref{eq:ineqcon}
-to define the problem \eqref{eq:costfun}-\eqref{eq:ic_con},
+formalism \eqref{eq:gcostfun}-\eqref{eq:eqcon}
+to define the optimization problem \eqref{eq:costfun}-\eqref{eq:ic_con},
 which we extend in subsequent sections to account
@@
 $h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
-are Pauli matrices. We optimize $X/2$,
-$Y/2$, and $Z/2$ gates (unitary transformations)
-for the fluxonium presented in \cite{zhang2020universal},
-and compare them to the analytically constructed gates for
-that device.
+are Pauli matrices. Although the coherent dynamics can be modeled as a two-level
+system, our noise model, experimental constraints, and system parameters
+consider the full system, and they are representative of the fluxonium
+presented in \cite{zhang2020universal}.
 
@@
 where the superscript is an index $i \in \{0, 1\}$,
-and the subscript indicates the knot point $k = 1$.
-The states at the final knot point are constrained to be
-the target states $\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{f}} \equiv U
-\ket*{\psi^{i}_{1}} \ \forall \ i$
-\eqref{eq:tstate_con} where $U = X/2, Y/2, Z/2$ is the desired gate.
+and the subscript indicates the time step $k = 1$.
+The states at the final time step are constrained to be
+the target states $\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{f}} \equiv
+U \ket*{\psi^{i}_{1}} \ \forall \ i$
+\eqref{eq:tstate_con} where $U = X/2, Y/2, Z/2$ is the target gate.
 Furthermore, we impose the normalization constraint
@@
 \eqref{eq:amp_con}.
-Above $0.5$ GHz, we observe population beyond the first two levels, disallowing the
-approximation \eqref{eq:hamiltonian}.
+Above $0.5$ GHz, we observe poor experimental agreement with the two-level approximation \eqref{eq:hamiltonian}.
 We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con}
@@
 
-The cost function at each knot point is
+The cost function at each time step is
 $\ell_{k}(x_{k}, u_{k}) = (x_{k} - x_{f})^{T} Q_{k} (x_{k} - x_{f}) + u^{T}_{k} R_{k} u_{k}$
@@
 states $\ket{0}$ and $\ket{1}$.
-We penalize the squared error at all knot points
+We penalize the squared error at all time steps
 because it benefits the iLQR solving stage \cite{Jackson2020altroc},
@@
 
-In addition to the target state cost function, we also
+In addition to the target state cost function, we
 impose the target state constraint \eqref{eq:tstate_con}, which
@@
 allowed to sacrifice the closed-system gate error to achieve better
-performance on the depolarization or robustness cost functions we will
-introduce in the following sections, which is undesirable.
+performance on the other cost functions, which is undesirable.
 To enforce a constraint in standard QOC frameworks,
@@
 ALM automates these prefactor updates to find
-a solution that satisfies all of the constraints.
+a solution that satisfies all of the given constraints.
 Hence, ALTRO's ability to handle multiple constraints makes it
-an attractive solver for robust QOC problems.
+an attractive solver for QOC problems.
 
@@
 QOC \cite{abdelhafez2020universal}, it may be impossible
-to obey the physics of the system and achieve the desired gate,
+to obey the physics of the system and achieve the desired gate;
 equivalently, the dynamics constraint \eqref{eq:dyn_con}
@@
 Finally, for the indirect ALM-iLQR solving stage,
-the state at each knot point is obtained by integrating
-the TDSE, so the dynamics constraint \eqref{eq:dyn_con} is satisfied by construction.
+the state at each time step is obtained by integrating
+the TDSE with the discrete dynamics function,
+so the dynamics constraint \eqref{eq:dyn_con} is satisfied by construction.
 This is not the case for the direct projected Newton stage, where the states
-are free parameters that are adjusted to satisfy the TDSE. However,
-the solution's deviation from the TDSE is never more than the maximum constraint violation.
-When we report gate errors in this text, we explicity integrate the TDSE with the
-flux produced by the optimization to ensure accuracy. Exploring the benefit
-of direct optimization approaches for QOC is an interesting direction for future work.
+are free parameters that are adjusted to satisfy the TDSE.
+Although the final solution's deviation from the TDSE is never more than the maximum constraint violation,
+we explicitly integrate the TDSE when reporting gate errors to ensure accuracy.
+Exploring the benefit of direct optimization approaches for QOC is an interesting direction for future work.

### 2021-03-13 paper/4longitude.tex
dave comments up to sec iv and readme updates

@@
 for optimizing the flux to mitigate depolarization.
-For many superconducting circuits, the 1/$e$ depolarization time
-($T_{1}$) is independent of the control parameters,
+For many superconducting circuits, the depolarization time
+$T_{1}$ is independent of the control parameters,
 so the fastest possible gate incurs the least depolarization error
@@
 \end{equation}
-This value is appended to the augmented state \eqref{eq:astatecontrols}
+For the gates we consider here, where the gate time is small compared to $T_{1}$,
+the integrated depolarization rate is proportional to the probability of a depolarization event.
+Additionally, the integrated depolarization rate is a reasonable proxy for the gate error incurred
+because depolarization errors are incoherent--they increase monotonically in time without interference.
+The integrated depolarization rate is appended to the augmented state \eqref{eq:astatecontrols}
 and its norm is penalized in the objective \eqref{eq:costfun} by setting
 the corresponding element of the target augmented state to zero.
-$T_{1}(a_{k})$ is obtained at each knot point by evaluating
+$T_{1}(a_{k})$ is obtained at each time step by evaluating
 a spline fit to experimental data of the form $\{(a, T_{1})\}$,
 see Figure \ref{fig:longitudeb}.
-Using the integrated depolarization rate as a proxy for the gate error incurred
-is reasonable because depolarization errors are incoherent and therefore
-increase monotonically in time without interference.
+
 Additionally, a master equation approach would require adding
@@
 We allow the optimizer to tune the gate time by
-making the time step between each knot point
+making the duration between time steps
 a decision variable \cite{howell2019altro}. 
-The square root of the time step $\sqrt{\Delta t_{k}}$
+The square root of the duration $\sqrt{\Delta t_{k}}$
 is appended to the augmented control \eqref{eq:astatecontrols}
 and its square $\lvert \Delta t_{k} \rvert$ is used
-for numerical integration in the discrete dynamics function \eqref{eq:dyn_con}.
-To ensure numerical
-integration accuracy is maintained, we constrain
-the bounds of the time step at each knot point.
+for integration in the discrete dynamics function \eqref{eq:dyn_con}.
+To ensure numerical stability, we constrain
+the bounds of the duration.
 
@@
 the $X/2$, $Y/2$, and $Z/2$ gates obtained with
-our numerical method and the corresponding analytic gates. We
-use the Lindblad master equation to simulate $T_{1}$ dissipation for successive
+our numerical method and the corresponding analytic gates presented in \cite{zhang2020universal}.
+We use the Lindblad master equation to simulate $T_{1}$ dissipation for successive
 gate applications, and compute the cumulative gate error

### 2021-03-13 paper/5static.tex
dave comments up to sec iv and readme updates

@@
 to accurately propagate the mean and covariance of this distribution between
-knot points, or equivalently, through the transformation of the TDSE \eqref{eq:tdse}.
+time steps, or equivalently, through the transformation of the TDSE \eqref{eq:tdse}.
 Unlike the sampling method, the cost function for the unscented sampling method
@@
 for each initial state.
-We obtain the state derivatives at each knot point by performing forward-mode
+We obtain the state derivatives at each time step by performing forward-mode
 differentiation on the TDSE \eqref{eq:tdse}.
@@
 While the state $\ket*{\psi}$ has unit norm,
-the state derivatives $\ket*{\partial^{m}_{f_{q}} \psi}$ need not, as is evident
+the state derivatives $\ket*{\partial^{l}_{f_{q}} \psi}$ need not, as is evident
 from the non-unitary dynamics \eqref{eq:d1dyn}.
@@
 
-The analytic gate corresponds to
-idling at the flux frustration point $a_{k} = 0 \ \forall \ k$, see Figure
-\ref{fig:statica}. Its gate time $1 / 4 f_{q} \sim 18\textrm{ns}$
-is the shortest possible for a $Z/2$ gate on the device.
-The gate's erroneous rotation angle
-$2 \pi \delta f_{q} / 4 f_{q}$ is linear in the
-qubit frequency detuning, resulting in a gate error that is quadratic
-in the detuning.
-At a one-percent detuning ($\lvert \delta f_{q} / f_{q} \rvert = 1\%$),
-the gate error is $\sim 4.7 \cdot 10^{-5}$,
-which is sufficient for quantum error correction.
-
 %% F3
@@
 
+The analytic gate corresponds to
+idling at the flux frustration point $a_{k} = 0 \ \forall \ k$, see Figure
+\ref{fig:statica}. Its gate time $1 / 4 f_{q} \sim 18\textrm{ns}$
+is the shortest possible for a $Z/2$ gate on the device.
+The gate's erroneous rotation angle
+$2 \pi \delta f_{q} / 4 f_{q}$ is linear in the
+qubit frequency detuning, resulting in a gate error that is quadratic
+in the detuning.
+At a one-percent detuning $\lvert \delta f_{q} / f_{q} \rvert = 1\%$,
+the gate error is $\sim 4.7 \cdot 10^{-5}$,
+which is sufficient for quantum error correction.
+
 For the sampling method, the gate error at a one-percent qubit frequency detuning
@@
 similar to the flux pulse produced by the unscented sampling method.
-For both methods, the gate error at a one-percent qubit frequency detuning
+For both derivative methods, the gate error at a one-percent qubit frequency detuning
 decreases super-linearly in the gate time.

### 2021-03-13 paper/a1longitude.tex
dave comments up to sec iv and readme updates

@@
 $L_{\pm} = \sigma^{\pm} \equiv (\sigma_{x} \pm i \sigma_{y})/2$.
-The depolarization times $T_{+} = T_{-} = 2 T_{1}$ are obtained at each knot point
+The depolarization times $T_{+} = T_{-} = 2 T_{1}$ are obtained at each time step
 from the spline shown in Figure \ref{fig:longitudeb}.
@@
 Because $T_{1}$ depends on the flux $a(t)$, so do
-the decay rates $\gamma_{i}$.
+the decay rates $\gamma_{\pm}$.
 Integrating the master equation with time-dependent decay rates
@@
 and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{ij} \ket{i} \otimes \ket{j}$.
-We use zero-order hold on the controls; equivalently, $a(t)$ is constant
-on each integration interval $[t_{k}, t_{k + 1}]$. Therefore,
-$H(a(t))$ and $\gamma_{i}(a(t))$ are also piecewise-constant.
-So, the exact solution is
+Because $a(t)$ is constant between time steps $[t_{k}, t_{k + 1})$,
+so are $H(a(t))$ and $\gamma_{\pm}(a(t))$.
+Therefore, the exact solution is
 $\textrm{vec}(\rho_{k + 1}) = {\exp}{\textstyle(}\Delta t_{k}
 \hat{\mathcal{L}}_{k}{\textstyle)} \textrm{vec}(\rho_{k})$.
-This isomorphism transforms $(n \times n) \times (n \times n)$
+The vector isomorphism transforms $(n \times n) \times (n \times n)$
 matrix-matrix multiplications to $(n^{2} \times n^{2}) \times n^{2}$ matrix-vector
@@
 faster to use an exponential integrator on the vectorized equation than to perform
-Runge-Kutta on the unvectorized equation. The latter requires decreasing the integration
-time step to maintain accuracy, resulting in more knot points.
+Runge-Kutta on the unvectorized equation. The latter requires decreasing the interval $\Delta t_{k}$
+to maintain accuracy, resulting in more time steps.

### 2021-03-13 paper/a2unscented.tex
resolved danny edits
dave comments up to sec iv and readme updates

--- 2021-02-05 paper/a2unscented.tex
@@
-\section{Unscented Transformation}
+\section{Unscented Sampling Method}
 \label{appendix:unscented}
 In this section, we outline the full unscented sampling procedure.
-In the unscented sampling method, we consider a state $\psi \in \mathbb{R}^{2n}$, 
+We consider a state $\psi \in \mathbb{R}^{2n}$, 
 an uncertain parameter $\lambda \in \mathbb{R}^{d}$, and discrete dynamics
@@
 the state is a real vector, and it is given by the right-hand-side of the
-isomorphism \eqref{eq:isomorphism}.
+complex-to-real isomorphism \eqref{eq:isomorphism}.
 The nominal initial state is given by $\bar{\psi}_{1}$ with an associated
@@
 set of real, symmetric, and positive-definite $m \times m$ matrices.
-$P_{1}$ is typically non-zero even if the state preparation error is negligible.
+By the positive-definite requirement, $P_{1}$ must be non-zero even if the state-preparation error
+is negligible.
 The uncertain parameter has zero-mean
 and its distribution is given by the covariance matrix
-$L_{k} \in \mathbb{S}_{++}^{d}$ at knot point $k$. The zero-mean assumption
+$L_{k} \in \mathbb{S}_{++}^{d}$ at time step $k$. The zero-mean assumption
 is convenient for deriving the update procedure. A non-zero mean can be encoded
@@
 \end{equation}
-The sample states are propagated to the next knot point,
+The sample states are propagated to the next time step,
 \begin{equation}\label{eq:upropagate}
@@
 \end{align}
-The sample states are then resampled and propagated to the next knot point using
+The sample states are then resampled and propagated to the next time step using
 \eqref{eq:uupdate}, \eqref{eq:unormalize}, and \eqref{eq:upropagate}. Our

### 2021-03-16 paper/2background.tex
updates based on dave comments

@@
 These models are used to derive a recurrence relation between time steps
-which gives the locally optimal update for the augmented controls.
+that gives the locally optimal update for the augmented controls.
 Finally, a line search \cite{zhang2006global}
@@
 reinhold2019controlling},
-is to add the constraint functions to the objective
-\eqref{eq:gcostfun}. 
+is to add the constraint functions to the objective.
 This strategy does not guarantee that the constraints

### 2021-03-16 paper/4longitude.tex
updates based on dave comments

@@
 and its square $\lvert \Delta t_{k} \rvert$ is used
-for integration in the discrete dynamics function \eqref{eq:dyn_con}.
-To ensure numerical stability, we constrain
-the bounds of the duration.
+for integration in the discrete dynamics function \eqref{eq:dyn_con},
+which ensures positivity in the event that
+the optimizer assigns a negative value to $\sqrt{\Delta t_{k}}$.
+Further, we constraint the minimum and maximum value of the
+duration to ensure integration accuracy.
 

### 2021-03-16 paper/5static.tex
updates based on dave comments

@@
 for the problem studied here due to our Hamiltonian and experimental constraints.
-We propose comparisons between composite pulses and numerical techniques
-for future work.
+A comparison between composite pulses and numerical techniques could be an
+interesting topic for future study
 
@@
 We can find gates using the numerical methods at
-all gate times above $18$ns, see Figure \ref{fig:staticb}.
+all gate times at and above $1 / 4 f_{q}$, see Figure \ref{fig:staticb}.
 These numerical methods offer an effective scheme for synchronizing

### 2021-03-16 paper/7conclusion.tex
updated bib and added brian edits 3
updates based on dave comments

@@
 \section{Conclusion}
-We have applied state-of-the-art trajectory
-optimization techniques to mitigate decoherence and
+We have introduced state-of-the-art trajectory optimization
+techniques in the context of quantum optimal control, enabling
+us to achieve tight tolerances for multiple constraints on the
+controls and quantum states. We used these capabilities
+to mitigate decoherence and
 achieve robustness to parameter uncertainty
@@
 implementations of the techniques described in this work are available
-at \url{https://github.com/SchusterLab/rbqoc}. \todo{static code?}
+at \url{https://github.com/SchusterLab/rbqoc}.

### 2021-03-16 paper/a1longitude.tex
updates based on dave comments

@@
 when the decay rates change much faster
-than the relaxation time $\Delta t_{k} \ll T_{1}$,
+than the relaxation time.
 which is the regime we are operating in. Standard
@@
 faster to use an exponential integrator on the vectorized equation than to perform
-Runge-Kutta on the unvectorized equation. The latter requires decreasing the interval $\Delta t_{k}$
+Runge-Kutta integration on the unvectorized equation. The latter requires decreasing the interval $\Delta t_{k}$
 to maintain accuracy, resulting in more time steps.

### 2021-03-16 paper/a3derivative.tex
updated bib and added brian edits 3
updates based on dave comments

@@
 for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler
-method \cite{berland2005solving},
+method \cite{berland2006solving},
 \begin{equation}
@@
 This method provides a good tradeoff between accuracy and efficiency, requiring one unique matrix
-exponential computation per stage. Integration accuracy is not of the utmost importance because the
-state derivatives guide the optimization, and do not correspond to experimental parameters
-which must be realized with high accuracy.
+exponential computation per stage. Integration accuracy for the state derivatives is not of the
+utmost importance because they are used in the robustness cost function--as opposed
+to the states themselves which are experimental parameters that must be realized
+with high accuracy.

### 2021-03-16 paper/1introduction.tex
jens initial final edits

@@
 We adapt robust control techniques from the robotics community to mitigate
-parameter uncertainty errors for
+parameter-uncertainty errors for
 a superconducting fluxonium qubit.
 
-Analytically-dervied control pulses that mitigate parameter uncertainty
+Analytically-derived control pulses that mitigate parameter-uncertainty
 errors include composite pulses \cite{cummins2000use, cummins2003tackling,
@@
 In this work, we study three robust control techniques that
-make the system's quantum state trajectory less sensitive
-to static and time-dependent parameter uncertainty:
+make the system's quantum state trajectory less sensitive to uncertainties of static and time-dependent parameters:
 \begin{enumerate}
-\item A sampling method, similar to the work of \cite{allen2019robust,
+\item A sampling method, similar to the work in Refs.~\cite{allen2019robust,
   ball2020software, carvalho2020error, khaneja2005optimal,
@@
 We also show that QOC can solve important problems associated with
-fluxonium-based qubits: exploiting the $T_{1}$-dependence of the controls
+fluxonium-based qubits: exploiting the $T_{1}$-dependence of the controls \textcolor{blue}{[as in the microwave generator output is $T_1$ dependent?]} 
 to mitigate depolarization
 and synchronizing the phase of qubits with distinct frequencies.
-To mitigate depolarization,
+To mitigate \textcolor{blue}{[close-range repetition]} depolarization,
 we perform time-optimal control and

### 2021-03-16 paper/2background.tex
jens initial final edits

--- 2021-03-13 paper/2background.tex
@@
 and describe the ALTRO solver \cite{howell2019altro}.
-QOC concerns a vector of time-dependent control parameters $u(t)$
+QOC concerns a vector $\mathbf{u}(t)$ of time-dependent control parameters 
 that steer the evolution of a (quantum) state $\ket{\psi(t)}$.
@@
 \begin{equation}
-  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(u(t), t) \ket{\psi(t)}.
+  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(\mathbf{u}(t), t) \ket{\psi(t)}.
   \label{eq:tdse}
 \end{equation}
-The Hamiltonian $H(u(t), t)$ is determined by the quantum system.
+The Hamiltonian $H(\mathbf{u}(t), t)$ is determined by the quantum system \textcolor{blue}{and the external control fields}.
 The QOC problem is to find the

### 2021-03-22 paper/1introduction.tex
jens all edits plus some thomas revisions

@@
 In this work, we study three robust control techniques that
-make the system's quantum state trajectory less sensitive to uncertainties of static and time-dependent parameters:
+make the system's quantum state trajectory less sensitive to the uncertainties of static and time-dependent parameters:
 \begin{enumerate}
@@
 We also show that QOC can solve important problems associated with
-fluxonium-based qubits: exploiting the $T_{1}$-dependence of the controls \textcolor{blue}{[as in the microwave generator output is $T_1$ dependent?]} 
+fluxonium-based qubits: exploiting the dependence of $T_{1}$ on the controls
 to mitigate depolarization
 and synchronizing the phase of qubits with distinct frequencies.
-To mitigate \textcolor{blue}{[close-range repetition]} depolarization,
+To ameliorate depolarization,
 we perform time-optimal control and

### 2021-03-22 paper/2background.tex
jens all edits plus some thomas revisions

--- 2021-03-13 paper/2background.tex
@@
 and describe the ALTRO solver \cite{howell2019altro}.
-QOC concerns a vector of time-dependent control parameters $u(t)$
-that steer the evolution of a (quantum) state $\ket{\psi(t)}$.
+QOC concerns a vector $\mathbf{a}(t)$ of time-dependent control fields
+that steer the evolution of a quantum state $\ket{\psi(t)}$.
 The evolution of the state is governed by the time-dependent
@@
 \begin{equation}
-  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(u(t), t) \ket{\psi(t)}.
+  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(\mathbf{a}(t), t) \ket{\psi(t)}.
   \label{eq:tdse}
 \end{equation}
-The Hamiltonian $H(u(t), t)$ is determined by the quantum system.
+The Hamiltonian $H(\mathbf{a}(t), t)$ is determined by the quantum system
+and the external control fields.
 The QOC problem is to find the
-controls that minimize a functional $J(u(t))$.
+controls that minimize a functional $J[\mathbf{a}(t)]$,
+which we call the objective.
 To make the problem numerically tractable,
-the controls and state are discretized into $N$ time steps.
-In the case of a single state-transfer problem, the functional is
-the infidelity of the target state and the initial state evolved
-to the final time step,
-$J(u) = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u)} \rvert}^{2}$.
-In general, $J(u)$ is a linear combination of cost functions on the state
-as well as cost functions on the controls.
-Standard QOC solvers compute derivatives of the functional $\nabla_{u} J(u)$,
+the quantum state and controls are discretized into $N$ time steps,
+$\ket{\psi(t_{k})} \to \ket{\psi_{k}}$,
+and $\mathbf{a}(t_{k})\to \mathbf{a}_{k}$, where $t_{k + 1} = t_{k} + \Delta t$,
+and $k \in \{1, ..., N\}$.
+In the case of a single state-transfer problem, the objective is
+the infidelity of the time-evolved final state $\ket{\psi_{N}}$ and
+the intended target state $\ket{\psi_{T}}$,
+$J(\mathbf{a}) = 1 - {\lvert \braket{\psi_{T}}{\psi_{N}(\mathbf{a})} \rvert}^{2}$.
+Standard QOC solvers compute derivatives of the objective $\nabla J(\mathbf{a})$,
 which can easily be used to implement first-order optimization methods
@@
 \cite{Schulman13, Tedrake16, Hereid2017FROST, howell2019altro}.
-The functional $J(u)$ is divided into its constituent
-cost functions at each time step $\ell_{k}(x_{k}, u_{k})$,
-where $k \in \{1, ..., N\}$ denotes the time step,
-$x_{k}$ is the augmented state vector,
-and $u_{k}$ is the augmented control vector.
-The augmented state contains all relevant variables
-that are dependent upon the controls, for example the state
-$\ket{\psi_{k}} \subseteq x_{k}$.
-The augmented state trajectory obeys the physics of the system if
-the dynamics constraint is satisfied
-$x_{k + 1} = f(x_{k}, u_{k})$. For QOC, the discrete
-dynamics function $f(x_{k}, u_{k})$ propagates the state by
-integrating the TDSE \eqref{eq:tdse} with Runge-Kutta methods \cite{jorgensen2011numerical}
+The objective $J(\mathbf{a})=\sum_k \ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k})$ is expressed in terms of the cost function
+at each time step $\ell_k$, where
+$\mathbf{x}_{k}$ is the state vector,
+and $\mathbf{u}_{k}$ is the control vector.
+The control vector contains all variables that the experimentalist
+may manipulate, and the state vector contains all variables that
+depend upon those in the control vector.
+For example, the control field 
+$\mathbf{a}_{k}$ is a component of the control vector $\mathbf{u}_{k}$,
+and the quantum state $\ket{\psi_{k}}$ is a component of
+the state vector $\mathbf{x}_{k}$.
+
+The dependence
+of the state on the controls is encoded in the discrete relationship
+$\mathbf{x}_{k + 1} = \mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$.
+For QOC, the discrete dynamics function $\mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$
+propagates the quantum state by integrating the TDSE with Runge-Kutta methods \cite{jorgensen2011numerical}
 or exponential integrators \cite{auer2018magnus, berland2006solving, einkemmer2017performance,
-shillito2020fast}.
+shillito2020fast}. \todo{Jens wanted TDSE related before ddf}
 
-Additional constraints on the augmented controls and augmented states
-are encoded in constraint functions.
-The constraint functions are put into a form such that, when the constraint is satisfied,
-inequality constraint functions obey $g_{k}(x_{k}, u_{k}) \leq 0$ and
-equality constraint functions obey $h_{k}(x_{k}, u_{k}) = 0$. The constraint
-functions may be vector-valued to encode multiple constraints of each type.
-The constraint function's violation is the magnitude by which it is not satisfied:
-$\textrm{max}(g(\cdot), 0)$ or $\textrm{max}(\abs{h(\cdot)})$
-where $\textrm{max}(\cdot)$ and $\abs{\cdot}$ act element-wise.
-Stated concisely, the trajectory optimization problem is,
+Additional constraints imposed on the augmented controls and augmented states
+are formulated as inequalities  $\mathbf{g}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) \leq \mathbf{0}$ or equalities $\mathbf{h}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = \mathbf{0}$.
+\textcolor{blue}{[How to understand the $k$?
+labels here?]}
+The constraint functions $\mathbf{g}_k$ and $\mathbf{h}_k$ may be vector-valued to encode multiple constraints, and equalities and inequalities are understood component-wise.
+Each constraint function's \emph{violation} is defined as the magnitude of the maximum deviation:
+$\textrm{max}(\mathbf{g}(\cdot), 0)$ or $\textrm{max}(\abs{\mathbf{h}(\cdot)})$
+where $\textrm{max}(\cdot)$ and $\abs{\cdot}$ are understood to act element-wise. \textcolor{blue}{[Is this the right place for this definition in the text. Seems unmotivated/out of place right now?]}
+Stated concisely, the trajectory optimization problem is
 \begin{mini!}[2]
-  {x_{1:N},u_{1:N-1}}{\ell_{N}(x_N) + \sum_{k=1}^{N-1}
-    \ell_{k}(x_k,u_k)}{}{} \label{eq:gcostfun}
-    \addConstraint{x_{k+1} = f(x_k,u_k) \ \forall \ k} \label{eq:gdyncon}
-    \addConstraint{g_k(x_k,u_k) \leq 0 \ \forall \ k} \label{eq:ineqcon}
-    \addConstraint{h_k(x_k,u_k) = 0 \ \forall \ k.} \label{eq:eqcon}
+  {\substack{\mathbf{x}_1,\ldots,\mathbf{x}_N \\ \mathbf{u}_1,\ldots,\mathbf{u}_{N-1}}}{\ell_{N}(\mathbf{x}_N) + \sum_{k=1}^{N-1}
+    \ell_{k}(\mathbf{x}_k,\mathbf{u}_k)}{}{} \label{eq:gcostfun}
+    \addConstraint{\mathbf{x}_{k+1} = \mathbf{f}(\mathbf{x}_k,\mathbf{u}_k) \quad \forall \ k} \label{eq:gdyncon}
+    \addConstraint{\mathbf{g}_k(\mathbf{x}_k,\mathbf{u}_k) \leq \mathbf{0} \quad \forall \ k} \label{eq:ineqcon}
+    \addConstraint{\mathbf{h}_k(\mathbf{x}_k,\mathbf{u}_k) = \mathbf{0} \quad \forall \ k.} \label{eq:eqcon}
 \end{mini!}
+\textcolor{blue}{[Can't we just define $\mathbf{u}_N=0$ to simplify the expression in (2a)?]}
 
@@
 and indirect methods \cite{betts1998survey}. For indirect methods,
-the augmented controls are the decision variables--the
+the augmented controls are the \emph{decision variables}, i.e., the
 variables the optimizer adjusts to solve the problem.
@@
 a variety of general-purpose solvers exist \cite{gill2005snopt, wachter2006implementation}.
+
 Recent state-of-the-art solvers, such as ALTRO,
-have combined principles from both of these approaches. ALTRO uses an
+ combine the indirect and direct method in a two-stage approach.  Specifically, ALTRO first employs the indirect method as implemented by an
 iterative linear-quadratic regulator (iLQR) algorithm
 \cite{Li2004a} as the internal solver of an augmented Lagrangian method (ALM)
-\cite{lantoine2012hybrid, plancher2017constrained, nocedal2006numerical}
-and employs a projected Newton method \cite{bertsekas1982projected, rao1998application}
-in its final solving stage.
+\cite{lantoine2012hybrid, plancher2017constrained, nocedal2006numerical}.
+In the second and final stage to the solution, the direct-method approach is utilized in the form of a projected Newton method \cite{bertsekas1982projected, rao1998application}.
+We next provide a more detailed summary of these two stages.
 
@@
 and its update procedure is based on the differential-dynamic-programming approach
-\cite{mayne1966a}.
+\cite{mayne1966a}. \textcolor{blue}{[Start with the familiar or the understandable explanation. The imposing sounding names can be given afterwards. (Purely psychology.)]}
 First, iLQR uses an initial guess for the augmented controls to obtain the augmented
@@
 about the current augmented controls and augmented states.
-These models are used to derive a recurrence relation between time steps
+These models are used to derive a recurrence relation between time steps \textcolor{blue}{[for some quantity (which?) at consecutive time steps]}
 which gives the locally optimal update for the augmented controls.
@@
 is performed in the direction of this update to ensure a
-decrease in the objective \eqref{eq:gcostfun}. This procedure
-is repeated until convergence.
+decrease in the objective \eqref{eq:gcostfun}.
+\textcolor{blue}{[again, I suspect that (2a) is not the ``objective"; rather the objective is what I know as cost function, which would exclude the minimization that is part of (2a).]} This procedure
+is repeated until convergence is reached.
 
-Indirect solvers such as iLQR are popular
-because they are very computationally efficient and maintain
-high accuracy for the discrete dynamics throughout the optimization.
-However, standard implementations cannot handle
+While indirect solvers like iLQR are computationally efficient and maintain high accuracy for the discrete dynamics throughout the optimization, they typically cannot handle
 nonlinear equality and inequality
 constraints \eqref{eq:ineqcon}-\eqref{eq:eqcon}.
-Projected gradient methods are a typical approach
-to handle constraints
+Projected gradient methods are a preferred approach
+to handle such constraints
 \cite{clarkson2010coresets, hauswirth2016projected, morzhin2019minimal,
@@
 not the augmented states.
+\textcolor{blue}{[Seems like my edit attempts and understanding of the two-stage process are still not appropriate.
+On the previous page, it seemed like the projected Newton method was the final stage and a direct method at that.
+However, here it sounds now like it is not good enough and we need yet another method?
+I'm confused.]}
 Another technique, which is popular for QOC \cite{heeres2017implementing, leung2017speedup,
@@
 is to add the constraint functions to the objective
-\eqref{eq:gcostfun}. 
+\eqref{eq:gcostfun}. \textcolor{blue}{[(2a) is of the form $\text{min}_x L(x)$ do you really mean to add to the minimum? Or only to the cost function $L(x)$?]}
 This strategy does not guarantee that the constraints

### 2021-03-22 paper/3fluxonium.tex
jens all edits plus some thomas revisions

@@
 In the following, we optimize quantum gates
-(unitary transformations) for the superconducting fluxonium qubit--a promising
+(unitary transformations) \textcolor{blue}{[for a physics journal, you will want as much handholding and explanation for the robotics community language as possible; bu contrast, clarifying that a quantum gate is a unitary tranformation seems out of place here.]} for the superconducting fluxonium qubit -- a promising
 building block for quantum computers due to its high
@@
 for experimental error channels.
-To high accuracy, we approximate the Hamiltonian near the flux-frustration
-point as a two-level system:
+To high accuracy, we can approximate the fluxonium Hamiltonian near the flux-frustration
+point by a two-level system:
 \begin{align}
-  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2},
+  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}.
   \label{eq:hamiltonian}
 \end{align}
-where $f_{q}$ is the qubit frequency at the flux-frustration point,
-$a(t)$ is the flux offset from the flux-frustration point,
+Here, $f_{q}$ is the qubit frequency at the flux-frustration point,
+$a(t)$ is the control governing the flux offset from the flux-frustration point,
 $h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
-are Pauli matrices. Although the coherent dynamics can be modeled as a two-level
-system, our noise model, experimental constraints, and system parameters
+are Pauli matrices. Although the coherent dynamics can be described based on this two-level
+system model, our noise model, experimental constraints, and system parameters
 consider the full system, and they are representative of the fluxonium
@@
 First, we outline the constraints for the fluxonium gate problem.
-All gates presented in this work satisfy these constraints to
-a maximum violation of $\sim 10^{-8}$.
-Casting this problem as a multi-state transfer, the initial conditions on
-the states are $\ket*{\psi^{0}_{1}} = \ket*{0}$, $\ket*{\psi^{1}_{1}} = \ket*{1}$
+Casting this problem in terms of a multi-state transfer problem, we fix as the initial states
+ $\ket*{\psi^{0}_{1}} = \ket*{0}$, $\ket*{\psi^{1}_{1}} = \ket*{1}$
 \eqref{eq:istate_con}
-where the superscript is an index $i \in \{0, 1\}$,
-and the subscript indicates the time step $k = 1$.
+where the superscript $i \in \{0, 1\}$ is the state index,
+and the subscript labels the time step -- here, $k = 1$.
 The states at the final time step are constrained to be
@@
 U \ket*{\psi^{i}_{1}} \ \forall \ i$
-\eqref{eq:tstate_con} where $U = X/2, Y/2, Z/2$ is the target gate.
+\eqref{eq:tstate_con} with $U = X/2, Y/2, Z/2$ denoting the target gate.
 Furthermore, we impose the normalization constraint
@@
 $a(t_{k}) \equiv a_{k}$,
-$\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$.
-We impose the zero net flux constraint $\int_{t} a_{N} = 0$
+$\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$. \textcolor{blue}{[Why is this a partial rather than a single-variable derivative?]}
+We impose the zero net-flux constraint $\int_{t} a_{N} = 0$
 \eqref{eq:znf_con}
@@
 Above $0.5$ GHz, we observe poor experimental agreement with the two-level approximation \eqref{eq:hamiltonian}.
-We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con}
+We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con} \textcolor{blue}{[Thomas, this would actually be consistent with my suggestion to set $\mathbf{u}_N$ to zero?]}
 so the gates may be concatenated arbitrarily. Additionally,
 we have the initial condition $\int_{t} a_{1} = \partial_{t} a_{1} = 0$
-\eqref{eq:ic_con}.
+\eqref{eq:ic_con}. All gates presented in this work satisfy these constraints to
+a maximum violation of $\sim 10^{-8}$.
 
@@
   \renewcommand*{\arraystretch}{1.3}
-  u_{k} = \begin{bmatrix} \partial^{2}_{t} a_{k} \end{bmatrix}, \quad
-  x_{k} = \begin{bmatrix} \ket{\psi^{0}_{k}} \\ \ket{\psi^{1}_{k}}
+  \mathbf{u}_{k} = \partial^{2}_{t} {a}_{k} , \quad
+  \mathbf{x}_{k} = \begin{bmatrix} \ket{\psi^{0}_{k}} \\ \ket{\psi^{1}_{k}}
     \\ \int_{t} a_{k} \\ a_{k} \\ \partial_{t} a_{k} \end{bmatrix}.
@@
 \end{equation}
-The time evolution of these variables are encoded
-in coupled, first-order, differential equations
-which are integrated in the discrete dynamics function \eqref{eq:dyn_con}.
-We integrate the states according to the TDSE \eqref{eq:tdse} and the
+(We maintain vector notation for $\mathbf{u}_k$ in anticipation of augmentation with additional components in Section \ref{sec:longitude}.)
+The time evolution of these variables is governed by 
+ coupled first-order differential equations
+which are encoded in the discrete dynamics function \eqref{eq:dyn_con}.
+We integrate the states according to the TDSE \eqref{eq:tdse} based on the
 fluxonium Hamiltonian \eqref{eq:hamiltonian}.
-The ALTRO implementation we use does not currently
-support complex numbers, so we repesent the states
-in the isomorphism $\mathcal{H}(\mathbb{C}^{n})
+Since the ALTRO implementation we use does not currently
+support complex numbers, we follow \cite{leung2017speedup} in  representing the states
+via the isomorphism $\mathcal{H}(\mathbb{C}^{n})
 \cong \mathcal{H}(\mathbb{R}^{2n})$ given in \cite{leung2017speedup},
@@
 The cost function at each time step is
-$\ell_{k}(x_{k}, u_{k}) = (x_{k} - x_{f})^{T} Q_{k} (x_{k} - x_{f}) + u^{T}_{k} R_{k} u_{k}$
-where $Q_{k}$ and $R_{k}$ are positive-definite diagonal matrices we supply.
+$\ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = (\mathbf{x}_{k} - \mathbf{x}_{f})^{T} Q_{k} (\mathbf{x}_{k} - \mathbf{x}_{f}) + \mathbf{u}^{T}_{k} R_{k} \mathbf{u}_{k}$
+where $Q_{k}$ and $R_{k}$ are positive-definite diagonal matrices we supply. \textcolor{blue}{[``We supply" -- sure, but not very helpful. How are they defined/constructed specifically?]}
 The $Q_{k}$ term
-penalizes deviations from the target augmented state $x_{f}$,
+penalizes deviations from the target augmented state $\mathbf{x}_{f}$,
 which is given by the constraints we have imposed on
 $\ket*{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$ in addition to
-$\partial_{t} a_{f} = 0$.
+$\partial_{t} a_{f} = 0$. \textcolor{blue}{[Should this have been $a_N$ rather than $a_f$?]}
 The $R_{k}$ term penalizes the norm of $\partial^{2}_{t} a_{k}$,
@@
 \begin{mini!}[2] 
-  {x_{1:N}, u_{1:N\text{-}1}}{\sum_{k=1}^N {(x_k-x_f)}^{T} Q_k (x_k-x_{f})
-    + \sum_{k=1}^{N-1} {u_k}^{T} R_k u_{k}}{}{} \label{eq:costfun}
-  \addConstraint{x_{k+1}}{= f(x_k, u_k) \ \forall \ k}  \label{eq:dyn_con}
+  {\substack{\mathbf{x}_1,\ldots,\mathbf{x}_N \\ \mathbf{u}_1,\ldots,\mathbf{u}_{N-1}}}{\sum_{k=1}^N {(\mathbf{x}_k-\mathbf{x}_f)}^{T} Q_k (\mathbf{x}_k-\mathbf{x}_{f})
+    + \sum_{k=1}^{N-1} {\mathbf{u}_k}^{T} R_k \mathbf{u}_{k}}{}{} \label{eq:costfun}
+  \addConstraint{\mathbf{x}_{k+1}}{= \mathbf{f}(\mathbf{x}_k, u_k) \quad \forall \ k}  \label{eq:dyn_con}
   \addConstraint{\ket{\psi^{0}_{1}} = \ket{0}, \ket{\psi^{1}_{1}} = \ket{1}} \label{eq:istate_con}
   \addConstraint{\ket{\psi^{i}_{N}} = \ket{\psi^{i}_{f}}
-    \ \forall \  i} \label{eq:tstate_con}
+    \quad \forall \  i} \label{eq:tstate_con}
   \addConstraint{{\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}}
-      \rvert}^{2} = 1 \ \forall \ i, k} \label{eq:statenorm_con}
+      \rvert}^{2} = 1 \quad \forall \ i, k} \label{eq:statenorm_con}
   \addConstraint{{\textstyle \int_{t}} a_{N} = 0} \label{eq:znf_con}
-  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz} \ \forall \ k} \label{eq:amp_con}
+  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz} \quad \forall \ k} \label{eq:amp_con}
   \addConstraint{a_{1} = a_{N} = 0} \label{eq:bound_con}
@@
 We penalize the squared error of the final state and the target
-state \eqref{eq:costfun}, rather than their infidelity,
-because the Hessian of the squared error cost function is diagonal--which
-makes matrix multiplications fast--and we wish to optimize $Z/2$ gates,
+state \eqref{eq:costfun} \textcolor{blue}{[referencing (6a) is puzzling here unless you specify $R_k$; how is ``squared error" different from infidelity?]}, rather than their infidelity,
+because the Hessian of the squared-error cost function is diagonal -- which
+makes matrix multiplications fast -- and we wish to optimize $Z/2$ gates,
 which requires a metric that is sensitive to global phases for the initial
@@
 because it benefits the iLQR solving stage \cite{Jackson2020altroc},
-not to incentivize early achievement
+not to incentivize early achievement \textcolor{blue}{[NOT to incentivize or to incentivize??]}
 of the desired gate \cite{leung2017speedup}.
 
-In addition to the target state cost function, we
-impose the target state constraint \eqref{eq:tstate_con}, which
-says that the final state must match the target state, including its global phase,
+In addition to the target-state cost function, we
+impose a target-state constraint \eqref{eq:tstate_con} which  requires the final state to match the target state, including its global phase,
 up to our chosen maximum constraint violation $\sim 10^{-8}$.
@@
 equivalently, the dynamics constraint \eqref{eq:dyn_con}
-and the target state constraint \eqref{eq:tstate_con} may be mutually unsatisfiable.
+and the target-state constraint \eqref{eq:tstate_con} may be mutually unsatisfiable.
 In this case, the prefactors for the constraint function terms
-in the ALM objective will tend to infinity--leading to numerical instability--and the
+in the ALM objective will tend to infinity -- leading to numerical instability -- and the
 optimization will not converge. To maintain a constrained approach in this situation,
-the maximum constraint violation for the target state constraint can be raised
+the maximum constraint violation for the target-state constraint can be raised
 to a level commensurate with the minimum acceptable gate error.

### 2021-03-22 paper/4longitude.tex
jens all edits plus some thomas revisions

--- 2021-03-13 paper/4longitude.tex
@@
 \begin{equation}
-  D_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}.
+  D_{1}(t) = \int_{0}^{t} T_{1}^{-1}[a(t^{\prime})] dt^{\prime}.
 \end{equation}
@@
 Additionally, the integrated depolarization rate is a reasonable proxy for the gate error incurred
-because depolarization errors are incoherent--they increase monotonically in time without interference.
+because depolarization errors are incoherent -- they increase monotonically in time without interference.
 The integrated depolarization rate is appended to the augmented state \eqref{eq:astatecontrols}
@@
 $T_{1}(a_{k})$ is obtained at each time step by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$,
+a spline fit to experimental data of the form $\{(a, T_{1})\}$ \textcolor{blue}{[What's the meaning of the nested parentheses here?
+Is the outer supposed to be a set?
+Then elements would need to be indexed by $k$.]},
 see Figure \ref{fig:longitudeb}.
+\textcolor{blue}{Phys. Rev. will want this to read 1(b). Not sure how to change this the way you are doing automatic referencing.}
 
-Additionally, a master equation approach would require adding
+An alternative master equation approach would require adding
 density matrices of size $n \times n$
@@
 where $n$ is the dimension of the Hilbert space.
-The integrated depolarization rate is a single
+By contrast, the integrated depolarization rate is a single
 real number; thus, the computational complexity of this
@@
 
+\textcolor{blue}{[abrupt start of a new paragraph; provide transition/motivation for what's next; the entire following paragraph feels a bit out of place/inserted after the fact.]}
 We allow the optimizer to tune the gate time by
@@
 see Figure \ref{fig:longitudeb}.
-The single gate errors for both the analytic and numerical gates are
+The single-gate errors for both the analytic and numerical gates are
 less than $10^{-4}$, which makes them sufficient for quantum error correction--a
@@
 computing \cite{aharonov2008fault, knill2005quantum, gottesman1997stabilizer}.
-However, the numerical gates achieve single gate errors
+However, the numerical gates achieve single-gate errors
 $\sim 5$ times less than those for the analytic gates,
@@
 on the integrated depolarization rate metric, see Appendix \ref{appendix:longitude}.
-This single gate error advantage corresponds to a
+This advantage in single-gate errors corresponds to a
 significant reduction in error correction resources
@@
 important for noisy, intermediate-scale quantum (NISQ)
-applications. These improvements are significant for the constraints
+applications. These improvements are significant for the constraints \textcolor{blue}{[improvements are significant FOR THE CONSTRAINTS?]}
 we have imposed on the gates,

### 2021-03-22 paper/5static.tex
jens all edits plus some thomas revisions

--- 2021-03-13 paper/5static.tex
@@
 We have formulated the QOC
-problem as an open-loop optimization problem; equivalently,
-we do not incorporate feedback from the experiment in optimization.
-However, the device's parameters deviate from the parameters we use in optimization,
+problem as an \emph{open-loop} optimization problem, i.e.,
+we do not incorporate feedback from the experiment into the optimization.
+However, the precise device parameters will realistically deviate from the parameters we use in optimization,
 leading to poor experimental performance. We combat errors
@@
 
+\textcolor{blue}{[Consider structuring the following into subsections to clearly separate the three methods.]}
 The sampling method incentivizes the optimizer
-to ensure multiple copies of a state, each of which evolves
+to ensure that multiple copies of a state, each evolving
 with a distinct value of the uncertain parameter, achieve
@@
 with $f_{q} \rightarrow f_{q} \pm \sigma_{f_{q}}$ for a fixed
-hyperparameter $\sigma_{f_{q}}$ which is the standard deviation of the qubit frequency.
-We penalize the infidelities of the sample states and their target state
+ standard deviation $\sigma_{f_{q}}$ of the qubit frequency, acting as a hyperparameter.
+We penalize the infidelities of the sample states with respect to the target state
 by adding a cost function to the objective \eqref{eq:costfun} of the form
@@
 for the initial states. As an example, a $Z/2$ gate achieved by idling
-at the flux frustration point $a_{k} = 0 \ \forall \ k$
+at the flux frustration point ($a_{k} = 0 \ \forall \ k$)
 will be robust to qubit frequency detunings for the initial states $\ket{0}$
@@
 but this gate will not be robust for any other initial states.
-Therefore, we choose four initial states
-so that their outer products
-span the operators on the Hilbert space
-$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
+Therefore, we choose the four initial states $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
 (\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized},
-which we refer to as the operator basis.
+whose outer products span the operators on the Hilbert space, and  we refer to them as the operator basis.
+\textcolor{blue}{[I cannot find the term ``operator basis" in the rest of the paper.
+If not used, then don't introduce terminology.]}
 
@@
 We integrate the coupled ODEs with exponential
-integrators in the discrete dynamics function \eqref{eq:dyn_con},
+integrators in the discrete dynamics function \eqref{eq:dyn_con} \textcolor{blue}{[integrators IN A function? sounds like mix-up between the function vs. how it is implemented numerically]},
 see Appendix \ref{appendix:derivative}.
@@
 for the derivative method.
-For runtimes of the three robust control techniques,
+For runtimes of the three robust control techniques \textcolor{blue}{[a ``technique" is abstract/general; a technique does not have a concrete runtime; do you mean computational complexity?]},
 consult Appendix \ref{appendix:time}.

### 2021-03-22 paper/7conclusion.tex
updated conclusiong
jens all edits plus some thomas revisions

--- 2021-02-09 paper/7conclusion.tex
@@
 control and the integrated depolarization rate model.
-The computational cost of this model is
+The computational cost of this model \textcolor{blue}{[I would argue that a \emph{model} has no cost... Computing something does.]} is
 independent of the dimension of the Hilbert space, enabling

### 2021-03-22 paper/1introduction.tex
jens final comments

@@
 In this work, we study three robust control techniques that
-make the system's quantum state trajectory less sensitive to the uncertainties of static and time-dependent parameters:
+make the system's quantum state trajectory less sensitive to uncertainties of static and time-dependent parameters:
 \begin{enumerate}

### 2021-03-22 paper/2background.tex
jens final comments

@@
 and describe the ALTRO solver \cite{howell2019altro}.
-QOC concerns a vector $\mathbf{a}(t)$ of time-dependent control fields
-that steer the evolution of a quantum state $\ket{\psi(t)}$.
+QOC concerns a vector $\mathbf{a}(t)$ of time-dependent control parameters 
+that steer the evolution of a (quantum) state $\ket{\psi(t)}$.
 The evolution of the state is governed by the time-dependent
@@
 \end{equation}
-The Hamiltonian $H(\mathbf{a}(t), t)$ is determined by the quantum system
-and the external control fields.
+The Hamiltonian $H(\mathbf{a}(t), t)$ is determined by the quantum system and the external control fields.
 The QOC problem is to find the
-controls that minimize a functional $J[\mathbf{a}(t)]$,
-which we call the objective.
+controls that minimize a functional $J[\mathbf{a}(t)]$.
 To make the problem numerically tractable,
-the quantum state and controls are discretized into $N$ time steps,
-$\ket{\psi(t_{k})} \to \ket{\psi_{k}}$,
-and $\mathbf{a}(t_{k})\to \mathbf{a}_{k}$, where $t_{k + 1} = t_{k} + \Delta t$,
-and $k \in \{1, ..., N\}$.
-In the case of a single state-transfer problem, the objective is
-the infidelity of the time-evolved final state $\ket{\psi_{N}}$ and
-the intended target state $\ket{\psi_{T}}$,
-$J(\mathbf{a}) = 1 - {\lvert \braket{\psi_{T}}{\psi_{N}(\mathbf{a})} \rvert}^{2}$.
-Standard QOC solvers compute derivatives of the objective $\nabla J(\mathbf{a})$,
+the quantum state and controls are discretized into $N$ time steps, $\ket{\psi(t)}\to\ket{\psi_k}$ and $\mathbf{a}(t)\to \mathbf{a}_k$ where $k \in \{1, ..., N\}$.
+In the case of a single state-transfer problem, the functional is
+the infidelity of the time-evolved final state $\ket{\psi_N}$ with respect to the intended target state $\ket{\psi_f}$ \textcolor{blue}{[to me, ``f" would mean ``final" rather than target...]},
+$J(\mathbf{a}) = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(\mathbf{a})} \rvert}^{2}$.
+In general, $J(\mathbf{u})$ is a linear combination of cost functions depending on the state
+ or the controls.
+Standard QOC solvers compute derivatives of the functional $\nabla J(\mathbf{a})$,
 which can easily be used to implement first-order optimization methods
@@
 \cite{Schulman13, Tedrake16, Hereid2017FROST, howell2019altro}.
-The objective $J(\mathbf{a})=\sum_k \ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k})$ is expressed in terms of the cost function
-at each time step $\ell_k$, where
-$\mathbf{x}_{k}$ is the state vector,
-and $\mathbf{u}_{k}$ is the control vector.
-The control vector contains all variables that the experimentalist
-may manipulate, and the state vector contains all variables that
-depend upon those in the control vector.
-For example, the control field 
-$\mathbf{a}_{k}$ is a component of the control vector $\mathbf{u}_{k}$,
-and the quantum state $\ket{\psi_{k}}$ is a component of
-the state vector $\mathbf{x}_{k}$.
-
-The dependence
-of the state on the controls is encoded in the discrete relationship
-$\mathbf{x}_{k + 1} = \mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$.
-For QOC, the discrete dynamics function $\mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$
-propagates the quantum state by integrating the TDSE with Runge-Kutta methods \cite{jorgensen2011numerical}
+The functional $J(\mathbf{u})=\sum_k \ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k})$ is expressed in terms of its constituent
+cost functions $\ell_k$, where
+$\mathbf{x}_{k}$ is the augmented state vector,
+and $\mathbf{u}_{k}$ the augmented control vector \textcolor{blue}{[augmented by what?]}.
+The augmented state contains all relevant variables
+that are dependent upon the controls, for example the state
+$\ket{\psi_{k}}$ is a component of the augmented state vector $\mathbf{x}_{k}$.
+The augmented state trajectory obeys the physics of the system if
+the dynamics constraint 
+$\mathbf{x}_{k + 1} = \mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$ is satisfied.
+\textcolor{blue}{["discrete dynamics function" is not physics language.
+Relate $\mathbf{f}$ to Schr\"odinger equation first, then give it the unfamiliar name.
+Separate point: if $\mathbf{x}$ contains components other than the quantum state, then what does "obeying the physics of the system" mean?]}.
+For QOC, the discrete
+dynamics function $\mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$ propagates the state by \textcolor{blue}{[definition of $\mathbf{f}$ should be separated from its numerical implementation?]}
+integrating the TDSE \eqref{eq:tdse} with Runge-Kutta methods \cite{jorgensen2011numerical}
 or exponential integrators \cite{auer2018magnus, berland2006solving, einkemmer2017performance,
-shillito2020fast}. \todo{Jens wanted TDSE related before ddf}
+shillito2020fast}.
 

### 2021-03-22 paper/8ack.tex
jens final comments

@@
   We thank Helin Zhang for experimental assistance
-  and Taylor Howell, Jens Koch, Tanay Roy, Colm Ryan, and Daniel Weiss for useful discussions.
+  and Taylor Howell, Tanay Roy, Colm Ryan, and Daniel Weiss for useful discussions.
   This work is funded in part by EPiQC, an NSF Expedition in Computing, under grant CCF-1730449.

### 2021-03-22 paper/a1longitude.tex
jens final comments

@@
 The integrated depolarization rate and the gate error due to
-depolarization metrics are compared in Table \ref{tab:longitude} for the
+depolarization metrics \textcolor{blue}{[The error is due to depolarization METRICS?]} are compared in Table \ref{tab:longitude} for the
 numerical experiment described in Section \ref{sec:longitude}.
-The relative performance of the analytic and numerical techniques
+The relative performance \textcolor{blue}{[what is relative performance?]} of the analytic and numerical techniques
 is similar across the two metrics.
@@
   \caption{
-    Single gate integrated depolarization rate ($D_{1}$)
-    and single gate error due to depolarization (GE).
+    Single-gate integrated depolarization rate ($D_{1}$)
+    and single-gate error due to depolarization (GE).
     Values are reported for the analytic (A) and numerical (N) gates.
@@
 \begin{equation}
-  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho]
+  \frac{d}{dt} \rho = -\frac{i}{\hbar} [H, \rho]
   + \sum_{i} \gamma_{i} (L_{i} \rho L_{i}^{\dagger}
   - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\}),
-\end{equation}
-where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix,
+\end{equation}\textcolor{blue}{[I suggest to remove the rest of the sentence completely; should not state any of this for physics journal]}
+where $\rho = \ket{\psi}\bra{\psi}$ \textcolor{blue}{[This is not correct (unless your state were to stay pure for all times)]} is the density matrix,
 $[\cdot, \cdot]$ is the algebraic commutator, and $\{\cdot, \cdot \}$ is the algebraic
@@
 Integrating the master equation with time-dependent decay rates
-provides a heuristic for how the gates we study may perform in
+provides a heuristic for how gates might  perform in
 the experiment. 
 This procedure may not be strictly correct
-when the decay rates change much faster
-than the relaxation time.
+when decay rates change significantly on the time scale of the relaxation time $\Delta t_{k} \ll T_{1}$,
 which is the regime we are operating in. Standard
@@
 
-So that we may use exponential integrators, we employ
-the Vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018},
+In order to use exponential integrators, we employ
+the vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018},
 \begin{equation}
-  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}} \textrm{vec}({\rho}),
+  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}}\, \textrm{vec}({\rho}),
 \end{equation}
@@
   \begin{aligned}
-    \hat{\mathcal{L}} &= -i(I \otimes H - H^{T} \otimes I)\\
+    \hat{\mathcal{L}} &= -i(\openone \otimes H - H^{T} \otimes \openone)\\
     &+ \sum_{i} \gamma_{i}
-    (L_{i}^{*} \otimes L_{i} - \frac{1}{2} (I \otimes L_{i}^{\dagger}L_{i}
-    - L_{i}^{T}L_{i}^{*} \otimes I)),
+    (L_{i}^{*} \otimes L_{i} - \frac{1}{2} (\openone \otimes L_{i}^{\dagger}L_{i}
+    - L_{i}^{T}L_{i}^{*} \otimes \openone)),
   \end{aligned}
@@
 and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{ij} \ket{i} \otimes \ket{j}$.
-Because $a(t)$ is constant between time steps $[t_{k}, t_{k + 1})$,
+Because $a(t)$ is constant between time steps $[t_{k}, t_{k + 1})$ \textcolor{blue}{[for real, because of the AWG?
+even that would get filtered, right?
+or as an approximation because of discretization?]},
 so are $H(a(t))$ and $\gamma_{\pm}(a(t))$.
@@
 matrix-matrix multiplications to $(n^{2} \times n^{2}) \times n^{2}$ matrix-vector
-multiplications. For small $n$ and zero-order hold on the controls, we find that it is
+multiplications. For small $n$ and zero-order hold \textcolor{blue}{[what is zero-order hold?]} on the controls, we find that it is
 faster to use an exponential integrator on the vectorized equation than to perform
-Runge-Kutta integration on the unvectorized equation. The latter requires decreasing the interval $\Delta t_{k}$
+Runge-Kutta on the unvectorized equation. The latter requires decreasing the interval $\Delta t_{k}$
 to maintain accuracy, resulting in more time steps.

### 2021-03-22 paper/a2unscented.tex
jens final comments

@@
 We consider a state $\psi \in \mathbb{R}^{2n}$, 
-an uncertain parameter $\lambda \in \mathbb{R}^{d}$, and discrete dynamics
+an uncertain set of parameters $\lambda\in \mathbb{R}^{d}$, and discrete dynamics
 $\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
-We omit bra-ket notation to emphasize that
+We abandon bra-ket notation to emphasize that
 the state is a real vector, and it is given by the right-hand-side of the
@@
 The nominal initial state is given by $\bar{\psi}_{1}$ with an associated
-covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$
+covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$ \textcolor{blue}{[why do we need a double plus?]}
 which describes the uncertainty in the initial state.

### 2021-03-24 paper/2background.tex
revised jens edits

--- 2021-03-22 paper/2background.tex
@@
 \end{equation}
-The Hamiltonian $H(\mathbf{a}(t), t)$ is determined by the quantum system
-and the external control fields.
+The Hamiltonian $H(\mathbf{a}(t), t)$ is determined by the quantum system and the external control fields.
 The QOC problem is to find the
@@
 the quantum state and controls are discretized into $N$ time steps,
-$\ket{\psi(t_{k})} \to \ket{\psi_{k}}$,
-and $\mathbf{a}(t_{k})\to \mathbf{a}_{k}$, where $t_{k + 1} = t_{k} + \Delta t$,
-and $k \in \{1, ..., N\}$.
+$\ket{\psi(t_{k})} \to \ket{\psi_k}$ and $\mathbf{a}(t_{k}) \to \mathbf{a}_k$ where
+$t_{k + 1} = t_{k} + \Delta t$ and $k \in \{1, ..., N\}$.
 In the case of a single state-transfer problem, the objective is
-the infidelity of the time-evolved final state $\ket{\psi_{N}}$ and
-the intended target state $\ket{\psi_{T}}$,
+the infidelity of the time-evolved final state $\ket{\psi_N}$ and
+the intended target state $\ket{\psi_T}$
+\textcolor{blue}{[to me, ``f" would mean ``final" rather than target...]}
+\textcolor{red}{[changed $f$ to $T$]},
 $J(\mathbf{a}) = 1 - {\lvert \braket{\psi_{T}}{\psi_{N}(\mathbf{a})} \rvert}^{2}$.
@@
 Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
-and solved using a variety of specialized solvers developed by the robotics community
+and solved using specialized solvers developed by the robotics community
 \cite{Schulman13, Tedrake16, Hereid2017FROST, howell2019altro}.
-The objective $J(\mathbf{a})=\sum_k \ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k})$ is expressed in terms of the cost function
-at each time step $\ell_k$, where
-$\mathbf{x}_{k}$ is the state vector,
-and $\mathbf{u}_{k}$ is the control vector.
-The control vector contains all variables that the experimentalist
-may manipulate, and the state vector contains all variables that
-depend upon those in the control vector.
-For example, the control field 
-$\mathbf{a}_{k}$ is a component of the control vector $\mathbf{u}_{k}$,
-and the quantum state $\ket{\psi_{k}}$ is a component of
-the state vector $\mathbf{x}_{k}$.
-
-The dependence
-of the state on the controls is encoded in the discrete relationship
+The objective $J(\mathbf{a})=\sum_k \ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k})$
+is expressed in terms of the
+cost function at each time step $\ell_k$, where
+$\mathbf{x}_{k}$ is the augmented state vector
+and $\mathbf{u}_{k}$ is the augmented control vector.
+We use the term \emph{augmented} because these
+vectors contain all of the relevant variables in the optimization problem,
+not just the quantum state and the control fields,
+for an example see Section \ref{sec:fluxonium}.
+The augmented control contains all variables that the experimentalist
+may manipulate, and the augmented state contains all variables that depend
+on those in the augmented control.
+The variables in the augmented states depend on those in the augmented
+controls as defined by the differential equations governing the physical system, which are
+encoded in the discrete relation
 $\mathbf{x}_{k + 1} = \mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$.
 For QOC, the discrete dynamics function $\mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$
-propagates the quantum state by integrating the TDSE with Runge-Kutta methods \cite{jorgensen2011numerical}
+propagates the quantum state by
+integrating the TDSE \eqref{eq:tdse} with the given control fields
+using Runge-Kutta methods \cite{jorgensen2011numerical}
 or exponential integrators \cite{auer2018magnus, berland2006solving, einkemmer2017performance,
-shillito2020fast}. \todo{Jens wanted TDSE related before ddf}
+  shillito2020fast}.
+\textcolor{blue}{[”discrete dynamics function” is not physics language.
+    Relate f to Schrodinger equation first, then give it the un- familiar name.]}
+\textcolor{red}{[I think the sentence flows better with "The discrete dynamics function
+    $f(., .)$ propagates" rather than "$f(., .)$ propagates... By the way, $f(., .)$ is
+    called the ddf". Furthermore, I cannot produce a sentence which relates
+    the TDSE dependence of $\ket{\psi_{k}}$ and $\mathbf{a}_{k}$ to each other
+    and then introduces $\mathbf{x}_{k + 1} = \mathbf{f}(\mathbf{u}_{k}, \mathbf{x}_{k})$
+    because then I feel it needs to be explained again that $\ket{\psi_{k}}$
+    and $\mathbf{a}_{k}$ are contained in $\mathbf{x}_{k}$ and $\mathbf{u}_{k}$.
+    I think introducing the specific case of QOC later in "For QOC"
+    is less verbose and more words will detract from understanding in this instance.]}
+\textcolor{blue}{[Separate point: if x contains components other than the quantum state,
+    then what does "obeying the physics of the system” mean?]}
+\textcolor{red}{[everything time evolves under a differential equation even if it's
+    really simple $\mathrm{d}a/\mathrm{d}t = \dot{a}$]}
+\textcolor{blue}{[definition of $\mathbf{f}$ should be separated from its numerical
+    implementation?]}
+\textcolor{red}{[I think that the lead in "For QOC," makes it evident that this is
+    a specific case,
+    similar to how above it says "For example," when talking about specific components
+    of the state and control vectors]}.
 
-Additional constraints imposed on the augmented controls and augmented states
-are formulated as inequalities  $\mathbf{g}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) \leq \mathbf{0}$ or equalities $\mathbf{h}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = \mathbf{0}$.
-\textcolor{blue}{[How to understand the $k$?
-labels here?]}
-The constraint functions $\mathbf{g}_k$ and $\mathbf{h}_k$ may be vector-valued to encode multiple constraints, and equalities and inequalities are understood component-wise.
-Each constraint function's \emph{violation} is defined as the magnitude of the maximum deviation:
-$\textrm{max}(\mathbf{g}(\cdot), 0)$ or $\textrm{max}(\abs{\mathbf{h}(\cdot)})$
-where $\textrm{max}(\cdot)$ and $\abs{\cdot}$ are understood to act element-wise. \textcolor{blue}{[Is this the right place for this definition in the text. Seems unmotivated/out of place right now?]}
+Additional constraints on the augmented controls and states
+are formulated as inequalities  $\mathbf{g}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) \leq \mathbf{0}$
+or equalities $\mathbf{h}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = \mathbf{0}$.
+\textcolor{blue}{[How to understand the $k$? labels here?]}
+\textcolor{red}{[they denote the time step, which i think is consistent with everything else]}
+The constraint functions $\mathbf{g}_k$ and $\mathbf{h}_k$ may be vector-valued to encode multiple constraints,
+and equalities and inequalities are understood component-wise. Each constraint's \emph{violation}
+is defined as the magnitude of its deviation:
+$\textrm{max}(g(\cdot), 0)$ or $\abs{h(\cdot)}$, where $g$ and $h$
+are individual components of constraint functions $\mathbf{g}_{k}$ and $\mathbf{h}_{k}$,
+respectively.
+\textcolor{blue}{[Is this the right place for this definition in the text.
+    Seems unmotivated/out of place right now?]}
+\textcolor{red}{[Yes, first place it comes up in ALM paragraph is a worse place to introduce it]}
 Stated concisely, the trajectory optimization problem is
 \begin{mini!}[2]
-  {\substack{\mathbf{x}_1,\ldots,\mathbf{x}_N \\ \mathbf{u}_1,\ldots,\mathbf{u}_{N-1}}}{\ell_{N}(\mathbf{x}_N) + \sum_{k=1}^{N-1}
-    \ell_{k}(\mathbf{x}_k,\mathbf{u}_k)}{}{} \label{eq:gcostfun}
-    \addConstraint{\mathbf{x}_{k+1} = \mathbf{f}(\mathbf{x}_k,\mathbf{u}_k) \quad \forall \ k} \label{eq:gdyncon}
-    \addConstraint{\mathbf{g}_k(\mathbf{x}_k,\mathbf{u}_k) \leq \mathbf{0} \quad \forall \ k} \label{eq:ineqcon}
-    \addConstraint{\mathbf{h}_k(\mathbf{x}_k,\mathbf{u}_k) = \mathbf{0} \quad \forall \ k.} \label{eq:eqcon}
+  {\substack{\mathbf{x}_1, \ldots, \mathbf{x}_N \\\mathbf{u}_1, \ldots, \mathbf{u}_{N - 1}}}{
+    \ell_{N}(\mathbf{x}_{N}) + \sum^{N - 1}_{k = 1}\ell_{k}
+    (\mathbf{x}_k,\mathbf{u}_k)}{}{} \label{eq:gcostfun}
+  \addConstraint{\mathbf{x}_{k+1} = \mathbf{f}(\mathbf{x}_k,\mathbf{u}_k) \quad \forall \ k}
+  \label{eq:gdyncon}
+  \addConstraint{\mathbf{g}_k(\mathbf{x}_k,\mathbf{u}_k) \leq \mathbf{0} \quad \forall \ k}
+  \label{eq:ineqcon}
+  \addConstraint{\mathbf{h}_k(\mathbf{x}_k,\mathbf{u}_k) = \mathbf{0} \quad \forall \ k.}
+  \label{eq:eqcon}
 \end{mini!}
-\textcolor{blue}{[Can't we just define $\mathbf{u}_N=0$ to simplify the expression in (2a)?]}
+\textcolor{blue}{[Can't we just define $\mathbf{u}_N=\mathbf{0}$
+    to simplify the expression in (2a)?]}
+\textcolor{red}{[I'm not certain that is the same thing as
+    saying that $\mathbf{u}_{N}$ is undefined.
+    What I think you are saying is $\ell_{N}(\mathbf{x}_{N},
+    \mathbf{u}_{N} = \mathbf{0})$, but what is really going
+    on is that there is no dependence at all $\ell_{N}(\mathbf{x}_{N})$]}
 
@@
 The augmented states are obtained from the augmented controls using the discrete dynamics function,
-and are then used to evaluate derivatives of the cost functions \eqref{eq:gcostfun}.
+and they are used to evaluate derivatives of the cost functions.
 Then, the derivative information is employed to update the augmented controls.
@@
 GRAPE \cite{khaneja2005optimal, leung2017speedup}, and Krotov's method \cite{goerz2019krotov}.
-Conversely, direct methods treat both the augmented controls and the augmented states as decision
+Conversely, direct methods treat both the augmented controls and states as decision
 variables. In addition to minimizing the cost functions, the optimizer uses derivative information
-for the discrete dynamics function
-to satisfy the dynamics constraint \eqref{eq:gdyncon} to a specified tolerance.
+for the discrete dynamics function to satisfy the dynamics constraint
+\eqref{eq:gdyncon} to a specified tolerance.
 In this sense, the TDSE \eqref{eq:tdse} is a constraint that may be violated
@@
 a variety of general-purpose solvers exist \cite{gill2005snopt, wachter2006implementation}.
-
 Recent state-of-the-art solvers, such as ALTRO,
- combine the indirect and direct method in a two-stage approach.  Specifically, ALTRO first employs the indirect method as implemented by an
-iterative linear-quadratic regulator (iLQR) algorithm
+combine the indirect and direct methods in a two-stage approach.
+First, ALTRO employs an indirect solving stage using
+the iterative linear-quadratic regulator (iLQR) algorithm
 \cite{Li2004a} as the internal solver of an augmented Lagrangian method (ALM)
 \cite{lantoine2012hybrid, plancher2017constrained, nocedal2006numerical}.
-In the second and final stage to the solution, the direct-method approach is utilized in the form of a projected Newton method \cite{bertsekas1982projected, rao1998application}.
-We next provide a more detailed summary of these two stages.
+In the second direct stage, ALTRO uses a projected Newton method
+\cite{bertsekas1982projected, rao1998application}.
+Next, we provide a more detailed summary of these two stages.
 
 iLQR is an indirect method
-for solving the dynamically constrained
-trajectory optimization problem \eqref{eq:gcostfun}-\eqref{eq:gdyncon},
-and its update procedure is based on the differential-dynamic-programming approach
-\cite{mayne1966a}. \textcolor{blue}{[Start with the familiar or the understandable explanation. The imposing sounding names can be given afterwards. (Purely psychology.)]}
-First, iLQR uses an initial guess for the augmented controls to obtain the augmented
-states with the discrete dynamics function.
+for minimizing the objective subject to the dynamics constraint,
+i.e., solving \eqref{eq:gcostfun}-\eqref{eq:gdyncon}.
+\textcolor{blue}{[Start with the familiar or the understandable explanation.
+    The imposing sounding names can be given afterwards. (Purely psychology.)]}
+\textcolor{red}{[resolved]}
+First, iLQR uses an initial guess for the augmented controls to obtain the 
+augmented states with the discrete dynamics function.
 iLQR then constructs quadratic models for each cost function using
 their zeroth-, first- and second-order derivatives in a Taylor expansion
-about the current augmented controls and augmented states.
-These models are used to derive a recurrence relation between time steps \textcolor{blue}{[for some quantity (which?) at consecutive time steps]}
-which gives the locally optimal update for the augmented controls.
+about the current augmented controls and states.
+These models are used in a recurrence relation between time steps
+to derive the locally optimal update for the augmented controls.
+\textcolor{blue}{[for some quantity (which?) at consecutive time steps]}
+\textcolor{red}{[resolved]}
 Finally, a line search \cite{zhang2006global}
 is performed in the direction of this update to ensure a
-decrease in the objective \eqref{eq:gcostfun}.
-\textcolor{blue}{[again, I suspect that (2a) is not the ``objective"; rather the objective is what I know as cost function, which would exclude the minimization that is part of (2a).]} This procedure
+decrease in the objective.
+\textcolor{blue}{[again, I suspect that (2a) is not the ``objective"; rather the objective
+    is what I know as cost function, which would exclude the minimization that is part of (2a).]}
+\textcolor{red}{[resolved, i introduce objective earlier now]}
+This procedure
 is repeated until convergence is reached.
 
-While indirect solvers like iLQR are computationally efficient and maintain high accuracy for the discrete dynamics throughout the optimization, they typically cannot handle
+While indirect solvers like iLQR are computationally
+efficient and maintain high accuracy for the discrete dynamics
+throughout the optimization, they cannot handle
 nonlinear equality and inequality
 constraints \eqref{eq:ineqcon}-\eqref{eq:eqcon}.
-Projected gradient methods are a preferred approach
-to handle such constraints
-\cite{clarkson2010coresets, hauswirth2016projected, morzhin2019minimal,
-  nikolskii2007convergence}. Unfortunately, within the indirect framework,
-they can only be used for constraints on the augmented controls,
-not the augmented states.
-\textcolor{blue}{[Seems like my edit attempts and understanding of the two-stage process are still not appropriate.
-On the previous page, it seemed like the projected Newton method was the final stage and a direct method at that.
-However, here it sounds now like it is not good enough and we need yet another method?
-I'm confused.]}
-Another technique, which is popular for QOC \cite{heeres2017implementing, leung2017speedup,
-reinhold2019controlling},
+\textcolor{blue}{[Seems like my edit attempts and understanding
+    of the two-stage process are still not appropriate.
+    On the previous page, it seemed like the projected Newton
+    method was the final stage and a direct method at that.
+    However, here it sounds now like it is not good enough
+    and we need yet another method? I'm confused.]}
+\textcolor{red}{[removed discussion of projected gradient methods]}
+For QOC, a popular approach to handle such constraints
 is to add the constraint functions to the objective
-\eqref{eq:gcostfun}. \textcolor{blue}{[(2a) is of the form $\text{min}_x L(x)$ do you really mean to add to the minimum? Or only to the cost function $L(x)$?]}
-This strategy does not guarantee that the constraints
+\cite{heeres2017implementing, leung2017speedup, reinhold2019controlling}.
+\textcolor{blue}{[(2a) is of the form $\text{min}_x L(x)$
+    do you really mean to add to the minimum? Or only to the cost function $L(x)$?]}
+\textcolor{red}{[resolved, i introduce objective earlier now]}
+However, this strategy does not guarantee that the constraints
 are satisfied as the solver trades
@@
 can satisfy constraints
-on both the augmented controls and the augmented states to tight tolerances.
+on both the augmented controls and states to tight tolerances.
 This advantage is crucial for this work, where multiple medium-priority cost functions

### 2021-03-24 paper/3fluxonium.tex
revised jens edits

@@
 In the following, we optimize quantum gates
-(unitary transformations) \textcolor{blue}{[for a physics journal, you will want as much handholding and explanation for the robotics community language as possible; bu contrast, clarifying that a quantum gate is a unitary tranformation seems out of place here.]} for the superconducting fluxonium qubit -- a promising
+for the superconducting fluxonium qubit -- a promising
 building block for quantum computers due to its high
@@
   manucharyan2009fluxonium, nguyen2019high,
-  zhang2020universal}.
+  zhang2020universal}
+\textcolor{blue}{[for a physics journal, you will want as much handholding and explanation
+    for the robotics community language as possible; by contrast, clarifying that
+    a quantum gate is a unitary tranformation seems out of place here.]}
+\textcolor{red}{[resolved, removed unitary comment]}.
 In this section, we use the trajectory optimization
@@
 for experimental error channels.
-To high accuracy, we can approximate the fluxonium Hamiltonian near the flux-frustration
-point by a two-level system:
+To high accuracy, we approximate the fluxonium Hamiltonian near the flux-frustration
+point as a two-level system:
 \begin{align}
@@
 $h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
-are Pauli matrices. Although the coherent dynamics can be described based on this two-level
+are Pauli matrices. Although the coherent dynamics can be described with this two-level
 system model, our noise model, experimental constraints, and system parameters
@@
 
-First, we outline the constraints for the fluxonium gate problem.
-Casting this problem in terms of a multi-state transfer problem, we fix as the initial states
- $\ket*{\psi^{0}_{1}} = \ket*{0}$, $\ket*{\psi^{1}_{1}} = \ket*{1}$
-\eqref{eq:istate_con}
-where the superscript $i \in \{0, 1\}$ is the state index,
-and the subscript labels the time step -- here, $k = 1$.
-The states at the final time step are constrained to be
-the target states $\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{f}} \equiv
-U \ket*{\psi^{i}_{1}} \ \forall \ i$
-\eqref{eq:tstate_con} with $U = X/2, Y/2, Z/2$ denoting the target gate.
-Furthermore, we impose the normalization constraint
-${\lvert \braket*{\psi^{i}_{k}}{\psi^{i}_{k}} \rvert}^{2} = 1 \ \forall \ i,k$
-\eqref{eq:statenorm_con}
-to ensure the solver does not take advantage of discretization errors in numerical integration.
+First, we introduce the augmented control and state for this problem.
 To refer to the discrete moments of the flux, we introduce the notation
@@
 $a(t_{k}) \equiv a_{k}$,
-$\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$. \textcolor{blue}{[Why is this a partial rather than a single-variable derivative?]}
-We impose the zero net-flux constraint $\int_{t} a_{N} = 0$
-\eqref{eq:znf_con}
-which mitigates the inductive drift ubiquitous in flux-bias lines
-\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
-The flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
-\eqref{eq:amp_con}.
-Above $0.5$ GHz, we observe poor experimental agreement with the two-level approximation \eqref{eq:hamiltonian}.
-We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con} \textcolor{blue}{[Thomas, this would actually be consistent with my suggestion to set $\mathbf{u}_N$ to zero?]}
-so the gates may be concatenated arbitrarily. Additionally,
-we have the initial condition $\int_{t} a_{1} = \partial_{t} a_{1} = 0$
-\eqref{eq:ic_con}. All gates presented in this work satisfy these constraints to
-a maximum violation of $\sim 10^{-8}$.
-
-The augmented control and augmented state are:
+$\mathrm{d}^{n}a(t)/{\mathrm{d}t}^{n} \lvert_{t = t_{k}} \equiv \mathrm{d}^{n}_{t} a_{k}$.
+The augmented control and state are:
 \begin{equation}
@@
   \renewcommand*{\arraystretch}{1.3}
-  \mathbf{u}_{k} = \partial^{2}_{t} {a}_{k} , \quad
+  \mathbf{u}_{k} = \begin{bmatrix} \mathrm{d}^{2}_{t} a_{k} \end{bmatrix}, \quad
   \mathbf{x}_{k} = \begin{bmatrix} \ket{\psi^{0}_{k}} \\ \ket{\psi^{1}_{k}}
-    \\ \int_{t} a_{k} \\ a_{k} \\ \partial_{t} a_{k} \end{bmatrix}.
+    \\ \int_{t} a_{k} \\ a_{k} \\ \mathrm{d}_{t} a_{k} \end{bmatrix}.
   \endgroup
@@
 \end{equation}
-(We maintain vector notation for $\mathbf{u}_k$ in anticipation of augmentation with additional components in Section \ref{sec:longitude}.)
-The time evolution of these variables is governed by 
- coupled first-order differential equations
-which are encoded in the discrete dynamics function \eqref{eq:dyn_con}.
-We integrate the states according to the TDSE \eqref{eq:tdse} based on the
-fluxonium Hamiltonian \eqref{eq:hamiltonian}.
+Here, the superscript on the quantum states $i \in \{0, 1\}$ acts as a label.
+\textcolor{blue}{(We maintain vector notation for $\mathbf{u}_k$ in anticipation
+  of augmentation with additional components in Section \ref{sec:longitude}.)]}
+\textcolor{red}{[I think this sentence is redundant given
+    elaboration of augmented now in Section \ref{sec:background}]}
 Since the ALTRO implementation we use does not currently
-support complex numbers, we follow \cite{leung2017speedup} in  representing the states
-via the isomorphism $\mathcal{H}(\mathbb{C}^{n})
+support complex numbers, we represent the states
+in the isomorphism $\mathcal{H}(\mathbb{C}^{n})
 \cong \mathcal{H}(\mathbb{R}^{2n})$ given in \cite{leung2017speedup},
@@
 \end{equation}
+In standard QOC frameworks, the derivatives of the control fields
+are obtained with finite difference methods, e.g.
+$\mathrm{d}_{t} a_{k} \approx (a_{k + 1} - a_{k}) / \Delta t$ \cite{leung2017speedup}.
+Because iLQR requires that cost functions depend only on the augmented control and state at a given
+time step, i.e. are of the form $\ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k})$,
+we make $\mathrm{d}^{2}_{t} a_{k}$ the decision variable and
+numerically integrate coupled ODEs to obtain $\mathrm{d}_{t} a_{k}$, $a_{k}$, and $\int_{t} a_{k}$
+so that we may penalize them in cost functions.
+Similarly, the quantum states are obtained by numerically integrating
+the TDSE \eqref{eq:tdse} with the fluxonium Hamiltonian \eqref{eq:hamiltonian}
+and the given flux $a_{k}$. These numerical integration rules are implemented
+in the discrete dynamics function for the problem, and give rise to the
+dynamics constraint \eqref{eq:dyn_con}.
+
+Next, we outline the constraints for the fluxonium gate problem.
+Casting this problem in terms of a multi-state transfer problem, we fix as the initial states
+ $\ket*{\psi^{0}_{1}} = \ket*{0}$, $\ket*{\psi^{1}_{1}} = \ket*{1}$
+\eqref{eq:istate_con}.
+The states at the final time step are constrained to be
+the target states $\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{T}} \equiv
+U \ket*{\psi^{i}_{1}} \ \forall \ i$
+\eqref{eq:tstate_con} where $U = X/2, Y/2, Z/2$ denotes the target gate.
+Furthermore, we impose the normalization constraint
+${\lvert \braket*{\psi^{i}_{k}}{\psi^{i}_{k}} \rvert}^{2} = 1 \ \forall \ i,k$
+\eqref{eq:statenorm_con}
+to ensure the solver does not take advantage of discretization errors in numerical integration.
+We impose the zero net-flux constraint $\int_{t} a_{N} = 0$
+\eqref{eq:znf_con}
+which mitigates the inductive drift ubiquitous in flux-bias lines
+\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
+The flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
+\eqref{eq:amp_con} to ensure the two-level approximation
+\eqref{eq:hamiltonian} remains valid.
+We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con}
+\textcolor{blue}{[Thomas, this would actually be consistent with
+    my suggestion to set $\mathbf{u}_N$ to zero?]}
+\textcolor{red}{[Not quite, because $a_{N}$ is a component of $\mathbf{x}_{N}$,
+    see \eqref{eq:astatecontrols}]}
+so the gates may be concatenated arbitrarily. Additionally,
+we have the initial condition $\int_{t} a_{1} = \mathrm{d}_{t} a_{1} = 0$
+\eqref{eq:ic_con}. All gates presented in this work satisfy these constraints to
+a maximum violation of $\sim 10^{-8}$.
 
 The cost function at each time step is
-$\ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = (\mathbf{x}_{k} - \mathbf{x}_{f})^{T} Q_{k} (\mathbf{x}_{k} - \mathbf{x}_{f}) + \mathbf{u}^{T}_{k} R_{k} \mathbf{u}_{k}$
-where $Q_{k}$ and $R_{k}$ are positive-definite diagonal matrices we supply. \textcolor{blue}{[``We supply" -- sure, but not very helpful. How are they defined/constructed specifically?]}
+$\ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = (\mathbf{x}_{k}
+- \mathbf{x}_{T})^{T} Q_{k} (\mathbf{x}_{k} - \mathbf{x}_{T})
++ \mathbf{u}^{T}_{k} R_{k} \mathbf{u}_{k}$
+where $Q_{k}$ and $R_{k}$ are diagonal matrices we supply
+that assign weights to the components in the cost function.
+\textcolor{blue}{[``We supply" -- sure, but not very helpful.
+    How are they defined/constructed specifically?]}
+\textcolor{red}{[in QOC speak these are cost function weights, they are hyperparameters
+and set based on heuristics]}
 The $Q_{k}$ term
-penalizes deviations from the target augmented state $\mathbf{x}_{f}$,
-which is given by the constraints we have imposed on
-$\ket*{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$ in addition to
-$\partial_{t} a_{f} = 0$. \textcolor{blue}{[Should this have been $a_N$ rather than $a_f$?]}
-The $R_{k}$ term penalizes the norm of $\partial^{2}_{t} a_{k}$,
-smoothing the flux to mitigate high-frequency AWG transitions.
+penalizes deviations from the target augmented state
+$\mathbf{x}_{T} = \textrm{col}(\ket*{\psi^{0}_{T}}, \ket*{\psi^{1}_{T}}, 0, 0, 0)$,
+which is consistent with the constraints we have imposed on $\ket*{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$.
+Accordingly, this term penalizes the squared
+difference of $\ket*{\psi^{i}_{k}}$ and $\ket*{\psi^{i}_{T}}$
+and penalizes the norm of $\int_{t} a_{k}$, $a_{k}$, and $\mathrm{d}_{t} a_{k}$.
+We penalize the squared difference of the final and target
+quantum states, rather than their infidelities,
+because the Hessian of the squared-difference cost function is diagonal -- which
+makes matrix multiplications fast -- and we wish to optimize $Z/2$ gates,
+which requires a metric that is sensitive to global phases for the initial
+states $\ket*{0}$ and $\ket*{1}$.
+Additionally, the $R_{k}$ term penalizes the norm of $\mathrm{d}^{2}_{t} a_{k}$.
+Penalizing the norm of $\mathrm{d}^{2}_{t} a_{k}$ and $\mathrm{d}_{t} a_{k}$
+makes $a_{k}$ smooth, which mitigates high-frequency AWG transitions.
 Stated succinctly, the optimization problem takes the form:
 \begin{mini!}[2] 
-  {\substack{\mathbf{x}_1,\ldots,\mathbf{x}_N \\ \mathbf{u}_1,\ldots,\mathbf{u}_{N-1}}}{\sum_{k=1}^N {(\mathbf{x}_k-\mathbf{x}_f)}^{T} Q_k (\mathbf{x}_k-\mathbf{x}_{f})
+  {\substack{\mathbf{x}_1,\ldots,\mathbf{x}_N \\ \mathbf{u}_1,\ldots,\mathbf{u}_{N-1}}}{\sum_{k=1}^N {(\mathbf{x}_k-\mathbf{x}_{T})}^{T} Q_k (\mathbf{x}_k-\mathbf{x}_{T})
     + \sum_{k=1}^{N-1} {\mathbf{u}_k}^{T} R_k \mathbf{u}_{k}}{}{} \label{eq:costfun}
   \addConstraint{\mathbf{x}_{k+1}}{= \mathbf{f}(\mathbf{x}_k, u_k) \quad \forall \ k}  \label{eq:dyn_con}
-  \addConstraint{\ket{\psi^{0}_{1}} = \ket{0}, \ket{\psi^{1}_{1}} = \ket{1}} \label{eq:istate_con}
-  \addConstraint{\ket{\psi^{i}_{N}} = \ket{\psi^{i}_{f}}
+  \addConstraint{\ket*{\psi^{0}_{1}} = \ket*{0}, \ket*{\psi^{1}_{1}} = \ket*{1}} \label{eq:istate_con}
+  \addConstraint{\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{T}}
     \quad \forall \  i} \label{eq:tstate_con}
@@
   \addConstraint{a_{1} = a_{N} = 0} \label{eq:bound_con}
-  \addConstraint{{\textstyle \int_{t}} a_{1} = \partial_{t} a_{1} = 0}. \label{eq:ic_con}
+  \addConstraint{{\textstyle \int_{t}} a_{1} = \mathrm{d}_{t} a_{1} = 0}. \label{eq:ic_con}
 \end{mini!}
@@
 Next, we remark on our problem formulation.
-We penalize the squared error of the final state and the target
-state \eqref{eq:costfun} \textcolor{blue}{[referencing (6a) is puzzling here unless you specify $R_k$; how is ``squared error" different from infidelity?]}, rather than their infidelity,
-because the Hessian of the squared-error cost function is diagonal -- which
-makes matrix multiplications fast -- and we wish to optimize $Z/2$ gates,
-which requires a metric that is sensitive to global phases for the initial
-states $\ket{0}$ and $\ket{1}$.
-We penalize the squared error at all time steps
-because it benefits the iLQR solving stage \cite{Jackson2020altroc},
-not to incentivize early achievement \textcolor{blue}{[NOT to incentivize or to incentivize??]}
-of the desired gate \cite{leung2017speedup}.
-
+We put a cost function at all time steps \eqref{eq:costfun}
+because it benefits the iLQR solving stage \cite{Jackson2020altroc};
+putting a cost function at all time steps does not incentivize early achievement
+of the desired gate, as in \cite{leung2017speedup}, due to the target state constraint.
+\textcolor{blue}{[NOT to incentivize or to incentivize??]}
+\textcolor{red}{[NOT to incentivize]}
 In addition to the target-state cost function, we
-impose a target-state constraint \eqref{eq:tstate_con} which  requires the final state to match the target state, including its global phase,
+impose a target-state constraint \eqref{eq:tstate_con} which
+requires the final state to match the target state, including its global phase,
 up to our chosen maximum constraint violation $\sim 10^{-8}$.
@@
 QOC \cite{abdelhafez2020universal}, it may be impossible
-to obey the physics of the system and achieve the desired gate;
-equivalently, the dynamics constraint \eqref{eq:dyn_con}
+to obey the physics of the system and achieve the desired gate, i.e.
+the dynamics constraint \eqref{eq:dyn_con}
 and the target-state constraint \eqref{eq:tstate_con} may be mutually unsatisfiable.
@@
 Finally, for the indirect ALM-iLQR solving stage,
-the state at each time step is obtained by integrating
+the states at each time step are obtained by integrating
 the TDSE with the discrete dynamics function,

### 2021-03-24 paper/4longitude.tex
jens final comments
revised jens edits

@@
     are plotted at the time-averaged 
-    absolute flux and the time-averaged $T_{1}$ time for each pulse.
+    absolute flux and $T_{1}$ time for each pulse.
     (c) Cumulative gate errors due to depolarization as a function of the
@@
 Additionally, the integrated depolarization rate is a reasonable proxy for the gate error incurred
-because depolarization errors are incoherent -- they increase monotonically in time without interference.
+because depolarization errors are incoherent -- they increase
+monotonically in time without interference.
 The integrated depolarization rate is appended to the augmented state \eqref{eq:astatecontrols}
-and its norm is penalized in the objective \eqref{eq:costfun} by setting
+and its norm is penalized in the objective by setting
 the corresponding element of the target augmented state to zero.
-$T_{1}(a_{k})$ is obtained at each time step by evaluating
-a spline fit to experimental data of the form $\{(a, T_{1})\}$ \textcolor{blue}{[What's the meaning of the nested parentheses here?
-Is the outer supposed to be a set?
-Then elements would need to be indexed by $k$.]},
-see Figure \ref{fig:longitudeb}.
-\textcolor{blue}{Phys. Rev. will want this to read 1(b). Not sure how to change this the way you are doing automatic referencing.}
+$T_{1}$ as a function of the flux is obtained by evaluating
+a spline fit to experimental data, see Figure \ref{fig:longitudeb}.
+\textcolor{blue}{[Phys. Rev. will want this to read 1(b).
+    Not sure how to change this the way you are doing automatic referencing.]}
+\textcolor{red}{[resolved]}
 
-An alternative master equation approach would require adding
+Alternatively, modeling the depolarization with a master equation approach would require adding
 density matrices of size $n \times n$
@@
 By contrast, the integrated depolarization rate is a single
-real number; thus, the computational complexity of this
+real number; thus, the computational complexity of evaluating this
 depolarization model does not scale
@@
 
-\textcolor{blue}{[abrupt start of a new paragraph; provide transition/motivation for what's next; the entire following paragraph feels a bit out of place/inserted after the fact.]}
-We allow the optimizer to tune the gate time by
-making the duration between time steps
+\textcolor{blue}{[abrupt start of a new paragraph; provide
+    transition/motivation for what's next; the entire following
+    paragraph feels a bit out of place/inserted after the fact.]}
+\textcolor{red}{[better?]}
+To perform time-optimal control, we make
+the duration between time steps
 a decision variable \cite{howell2019altro}. 
@@
 and its square $\lvert \Delta t_{k} \rvert$ is used
-for integration in the discrete dynamics function \eqref{eq:dyn_con}.
-To ensure numerical stability, we constrain
-the bounds of the duration.
+for integration in the discrete dynamics function.
+Although we constrain the bounds of the duration between reasonable positive values to maintain
+numerical stability, the optimizer may assign negative values to the
+duration for intermediate optimization iterations, so this squaring
+approach maintains positivity.
 
@@
 important for noisy, intermediate-scale quantum (NISQ)
-applications. These improvements are significant for the constraints \textcolor{blue}{[improvements are significant FOR THE CONSTRAINTS?]}
+applications. These improvements are significant given the constraints
+\textcolor{blue}{[improvements are significant FOR THE CONSTRAINTS?]}
+\textcolor{red}{[changed "for" to "given"]}
 we have imposed on the gates,

### 2021-03-24 paper/5static.tex
revised jens edits

@@
     for twice the gate time of the analytic gate.
-    (b) Single gate error at a one-percent qubit frequency detuning as
+    (b) Single-gate error at a one-percent qubit frequency detuning as
     a function of the gate time. Missing
     data points represent gates with a gate error greater than $5 \cdot 10^{-5}$.
-    (c) Single gate error as a function of the qubit frequency detuning.
+    (c) Single-gate error as a function of the qubit frequency detuning.
     The gate errors for the analytic and 1\textsuperscript{st}-order derivative
@@
 
+\subsection{Sampling Method}
 \textcolor{blue}{[Consider structuring the following into subsections to clearly separate the three methods.]}
+\textcolor{red}{[better?]}
 The sampling method incentivizes the optimizer
@@
 to the augmented state \eqref{eq:astatecontrols}. The discrete dynamics
-function \eqref{eq:dyn_con} is modified
+function is modified
 so the sample states evolve under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
@@
 We penalize the infidelities of the sample states with respect to the target state
-by adding a cost function to the objective \eqref{eq:costfun} of the form
-$\sum_{k, \pm} b_{k} (1 - {\lvert \braket*{\psi^{\pm}_{k}}{\psi_{f}} \rvert}^{2})$
+by adding a cost function to the objective of the form
+$\sum_{k, \pm} b_{k} (1 - {\lvert \braket*{\psi^{\pm}_{k}}{\psi_{T}} \rvert}^{2})$
 where $b_{k}$ is a constant we supply.
@@
 (\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized},
-whose outer products span the operators on the Hilbert space, and  we refer to them as the operator basis.
+whose outer products span the operators on the Hilbert space,
+and  we refer to them as the operator basis.
 \textcolor{blue}{[I cannot find the term ``operator basis" in the rest of the paper.
     If not used, then don't introduce terminology.]}
+\textcolor{red}{[Operator basis is used in the appendix \ref{appendix:time}]}
 
+\subsection{Unscented Sampling Method}
 Whereas the sampling method penalizes the deviations of the sample states
@@
   thangavel2020robust}. Accordingly, the cost function we add
-to the objective \eqref{eq:costfun} takes the form
+to the objective takes the form
 $\sum_{k, j} c_{k} (\psi^{j}_{k} - \psi_{k})^{T}
@@
 
+\subsection{Derivative Method}
 The derivative method penalizes the sensitivity of the state
@@
 We integrate the coupled ODEs with exponential
-integrators in the discrete dynamics function \eqref{eq:dyn_con} \textcolor{blue}{[integrators IN A function? sounds like mix-up between the function vs. how it is implemented numerically]},
-see Appendix \ref{appendix:derivative}.
+integrators, see Appendix \ref{appendix:derivative}.
+\textcolor{blue}{[integrators IN A function? sounds like mix-up between the function vs. how it is implemented numerically]}
+\textcolor{red}{[removed ref. to ddf]}
 While the state $\ket*{\psi}$ has unit norm,
@@
 We penalize the norms of the state derivatives
-in the objective \eqref{eq:costfun} by setting the corresponding elements
+in the objective by setting the corresponding elements
 of the target augmented state to zero. Intuitively, this corresponds to penalizing
@@
 for the derivative method.
-For runtimes of the three robust control techniques \textcolor{blue}{[a ``technique" is abstract/general; a technique does not have a concrete runtime; do you mean computational complexity?]},
-consult Appendix \ref{appendix:time}.
+We present the runtimes of our implementations of the three robust control methods
+in Appendix \ref{appendix:time}.
+\textcolor{blue}{[a ``technique" is abstract/general; a technique does not have a concrete runtime; do you mean computational complexity?]},
+\textcolor{red}{[better?]}
 
+\subsection{Comparison}
 We examine the gate errors due to a static qubit frequency

### 2021-03-24 paper/7conclusion.tex
revised jens edits

--- 2021-02-09 paper/7conclusion.tex
@@
 control and the integrated depolarization rate model.
-The computational cost of this model is
+The computational complexity of evaluating this model is
 independent of the dimension of the Hilbert space, enabling
 inexpensive optimization on high-dimensional quantum systems.
+\textcolor{blue}{[I would argue that a \emph{model} has no cost... Computing something does.]}
+\textcolor{red}{[better?]}
 We have also proposed the derivative method for robust control which achieves

### 2021-03-24 paper/a1longitude.tex
revised jens edits

@@
 The integrated depolarization rate and the gate error due to
-depolarization metrics \textcolor{blue}{[The error is due to depolarization METRICS?]} are compared in Table \ref{tab:longitude} for the
+depolarization
+\textcolor{blue}{[The error is due to depolarization METRICS?]}
+\textcolor{red}{[resolved, removed the word metrics]}
+are compared in Table \ref{tab:longitude} for the
 numerical experiment described in Section \ref{sec:longitude}.
-The relative performance \textcolor{blue}{[what is relative performance?]} of the analytic and numerical techniques
+The ratio of the value obtained on the metric with the analytic technique
+to the value obtained with the numerical technique
 is similar across the two metrics.
+\textcolor{blue}{what is relative performance?}
+\textcolor{red}{replaced relative performance with more precise phrasing}
 
@@
   - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\}),
-\end{equation}\textcolor{blue}{[I suggest to remove the rest of the sentence completely; should not state any of this for physics journal]}
-where $\rho = \ket{\psi}\bra{\psi}$ \textcolor{blue}{[This is not correct (unless your state were to stay pure for all times)]} is the density matrix,
-$[\cdot, \cdot]$ is the algebraic commutator, and $\{\cdot, \cdot \}$ is the algebraic
-anti-commutator.
+\end{equation}
+\textcolor{blue}{[I suggest to remove the rest of the sentence completely; should not state any of this for physics journal]}
+\textcolor{red}{[sentence on commutator definitions removed]}
 For depolarization, $\gamma_{\pm} = T_{\pm}^{-1}$,
@@
 This procedure may not be strictly correct
-when decay rates change significantly on the time scale of the relaxation time $\Delta t_{k} \ll T_{1}$,
+when decay rates change significantly on the time scale of the relaxation time,
 which is the regime we are operating in. Standard
@@
 and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{ij} \ket{i} \otimes \ket{j}$.
-Because $a(t)$ is constant between time steps $[t_{k}, t_{k + 1})$ \textcolor{blue}{[for real, because of the AWG?
-even that would get filtered, right?
-or as an approximation because of discretization?]},
+Because $a(t)$ is constant between time steps $[t_{k}, t_{k + 1})$,
+  due to our numerical discretization,
+  \textcolor{blue}{[for real, because of the AWG? even that would get filtered,
+      right? or as an approximation because of discretization?]}
+  \textcolor{red}{[added comment bout discretization]},
 so are $H(a(t))$ and $\gamma_{\pm}(a(t))$.
@@
 matrix-matrix multiplications to $(n^{2} \times n^{2}) \times n^{2}$ matrix-vector
-multiplications. For small $n$ and zero-order hold \textcolor{blue}{[what is zero-order hold?]} on the controls, we find that it is
+multiplications. For small $n$, we find that it is
 faster to use an exponential integrator on the vectorized equation than to perform
-Runge-Kutta on the unvectorized equation. The latter requires decreasing the interval $\Delta t_{k}$
+Runge-Kutta on the unvectorized equation.
+\textcolor{blue}{[what is zero-order hold?]}
+\textcolor{red}{[removed ZOH comment]}
+The latter requires decreasing the interval $\Delta t_{k}$
 to maintain accuracy, resulting in more time steps.

### 2021-03-24 paper/a2unscented.tex
revised jens edits

@@
 The nominal initial state is given by $\bar{\psi}_{1}$ with an associated
-covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$ \textcolor{blue}{[why do we need a double plus?]}
+covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$
+\textcolor{blue}{[why do we need a double plus?]}
+\textcolor{red}{[the convention is $+ =$ positive-semidefinite, $++ =$ positive-definite]}
 which describes the uncertainty in the initial state.

### 2021-03-26 paper/1introduction.tex
final thomas edits

@@
 Accordingly, recent work has sought to achieve robustness in QOC
-frameworks using closed-loop methods \cite{egger2014adaptive, feng2018gradient,
-  li2017hybrid, wittler2020integrated} and open-loop methods \cite{
-  allen2019robust, carvalho2020error, reinhold2019controlling,
-  rembold2020introduction, kosut2013robust, niu2019universal}.
+frameworks using closed-loop methods
+\cite{egger2014adaptive, feng2018gradient,
+  li2017hybrid, wittler2020integrated}
+and open-loop methods
+\cite{ball2020software, carvalho2020error, allen2019robust, reinhold2019controlling,
+  rembold2020introduction, kosut2013robust, niu2019universal, khaneja2005optimal}.
 
-In this work, we study three robust control techniques that
-make the system's quantum state trajectory less sensitive to uncertainties of static and time-dependent parameters:
+In this work, we study three open-loop robust control techniques that
+make the quantum state trajectory less sensitive to the uncertainties
+of static and time-dependent parameters:
 \begin{enumerate}
-\item A sampling method, similar to the work in Refs.~\cite{allen2019robust,
-  ball2020software, carvalho2020error, khaneja2005optimal,
-  reinhold2019controlling, rembold2020introduction}.
+\item A sampling method, similar to the work in Refs.~
+  \cite{allen2019robust, khaneja2005optimal, reinhold2019controlling, rembold2020introduction}.
 \item An unscented sampling method \cite{howell2020direct, lee2013sigma, thangavel2020robust}
@@
 \cite{howell2019altro}, which can enforce constraints on
-the control parameters and the quantum state trajectory.
+the control fields and the quantum state trajectory.
 
@@
 First, we describe ALTRO in the context of QOC
-in Section \ref{sec:background}.
+in Sec. \ref{sec:background}.
 We outline realistic constraints for operating the fluxonium and
-define the associated QOC problem in Section \ref{sec:fluxonium}.
+define the associated QOC problem in Sec. \ref{sec:fluxonium}.
 Then, we formulate a method for suppressing depolarization
-in Section \ref{sec:longitude}. Next, we describe three techniques for achieving
-robustness to static parameter uncertainties in Section \ref{sec:static}. We
+in Sec. \ref{sec:longitude}. Next, we describe three techniques for achieving
+robustness to static parameter uncertainties in Sec. \ref{sec:static}. We
 adapt the same techniques to mitigate 1/$f$ flux noise
-in Section \ref{sec:stochastic}.
+in Sec. \ref{sec:stochastic}.

### 2021-03-26 paper/2background.tex
final thomas edits

@@
 $\ket{\psi(t_{k})} \to \ket{\psi_k}$ and $\mathbf{a}(t_{k}) \to \mathbf{a}_k$ where
-$t_{k + 1} = t_{k} + \Delta t$ and $k \in \{1, ..., N\}$.
+$t_{k} = t_{k - 1} + \Delta t$ and $k \in \{1, ..., N\}$.
 In the case of a single state-transfer problem, the objective is
 the infidelity of the time-evolved final state $\ket{\psi_N}$ and
-the intended target state $\ket{\psi_T}$
-\textcolor{blue}{[to me, ``f" would mean ``final" rather than target...]}
-\textcolor{red}{[changed $f$ to $T$]},
+the intended target state $\ket{\psi_T}$,
 $J(\mathbf{a}) = 1 - {\lvert \braket{\psi_{T}}{\psi_{N}(\mathbf{a})} \rvert}^{2}$.
@@
 not just the quantum state and the control fields,
-for an example see Section \ref{sec:fluxonium}.
+for an example see Sec. \ref{sec:fluxonium}.
 The augmented control contains all variables that the experimentalist
@@
 $\mathbf{x}_{k + 1} = \mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$.
-For QOC, the discrete dynamics function $\mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$
-propagates the quantum state by
-integrating the TDSE \eqref{eq:tdse} with the given control fields
-using Runge-Kutta methods \cite{jorgensen2011numerical}
-or exponential integrators \cite{auer2018magnus, berland2006solving, einkemmer2017performance,
+For QOC, $\mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$ -- which we call
+the discrete dynamics function -- propagates the quantum state by
+integrating the TDSE \eqref{eq:tdse}
+using a Runge-Kutta method \cite{jorgensen2011numerical}
+or an exponential integrator \cite{auer2018magnus, berland2006solving, einkemmer2017performance,
   shillito2020fast}.
-\textcolor{blue}{[”discrete dynamics function” is not physics language.
-    Relate f to Schrodinger equation first, then give it the un- familiar name.]}
-\textcolor{red}{[I think the sentence flows better with "The discrete dynamics function
-    $f(., .)$ propagates" rather than "$f(., .)$ propagates... By the way, $f(., .)$ is
-    called the ddf". Furthermore, I cannot produce a sentence which relates
-    the TDSE dependence of $\ket{\psi_{k}}$ and $\mathbf{a}_{k}$ to each other
-    and then introduces $\mathbf{x}_{k + 1} = \mathbf{f}(\mathbf{u}_{k}, \mathbf{x}_{k})$
-    because then I feel it needs to be explained again that $\ket{\psi_{k}}$
-    and $\mathbf{a}_{k}$ are contained in $\mathbf{x}_{k}$ and $\mathbf{u}_{k}$.
-    I think introducing the specific case of QOC later in "For QOC"
-    is less verbose and more words will detract from understanding in this instance.]}
-\textcolor{blue}{[Separate point: if x contains components other than the quantum state,
-    then what does "obeying the physics of the system” mean?]}
-\textcolor{red}{[everything time evolves under a differential equation even if it's
-    really simple $\mathrm{d}a/\mathrm{d}t = \dot{a}$]}
-\textcolor{blue}{[definition of $\mathbf{f}$ should be separated from its numerical
-    implementation?]}
-\textcolor{red}{[I think that the lead in "For QOC," makes it evident that this is
-    a specific case,
-    similar to how above it says "For example," when talking about specific components
-    of the state and control vectors]}.
 
-Additional constraints on the augmented controls and states
-are formulated as inequalities  $\mathbf{g}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) \leq \mathbf{0}$
+We incorporate constraints on the augmented controls and states
+by formulating them as inequalities  $\mathbf{g}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) \leq \mathbf{0}$
 or equalities $\mathbf{h}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = \mathbf{0}$.
-\textcolor{blue}{[How to understand the $k$? labels here?]}
-\textcolor{red}{[they denote the time step, which i think is consistent with everything else]}
 The constraint functions $\mathbf{g}_k$ and $\mathbf{h}_k$ may be vector-valued to encode multiple constraints,
-and equalities and inequalities are understood component-wise. Each constraint's \emph{violation}
-is defined as the magnitude of its deviation:
+and equalities and inequalities are understood component-wise.
+To quantify constraint satisfaction, we define each constraint's \emph{violation}
+as the magnitude of its deviation:
 $\textrm{max}(g(\cdot), 0)$ or $\abs{h(\cdot)}$, where $g$ and $h$
-are individual components of constraint functions $\mathbf{g}_{k}$ and $\mathbf{h}_{k}$,
+are components of constraint functions $\mathbf{g}_{k}$ and $\mathbf{h}_{k}$,
 respectively.
-\textcolor{blue}{[Is this the right place for this definition in the text.
-    Seems unmotivated/out of place right now?]}
-\textcolor{red}{[Yes, first place it comes up in ALM paragraph is a worse place to introduce it]}
-Stated concisely, the trajectory optimization problem is
+Stated concisely, the trajectory optimization problem is:
 \begin{mini!}[2]
   {\substack{\mathbf{x}_1, \ldots, \mathbf{x}_N \\\mathbf{u}_1, \ldots, \mathbf{u}_{N - 1}}}{
-    \ell_{N}(\mathbf{x}_{N}) + \sum^{N - 1}_{k = 1}\ell_{k}
-    (\mathbf{x}_k,\mathbf{u}_k)}{}{} \label{eq:gcostfun}
-  \addConstraint{\mathbf{x}_{k+1} = \mathbf{f}(\mathbf{x}_k,\mathbf{u}_k) \quad \forall \ k}
+    \sum^{N}_{k = 1}\ell_{k}
+    (\mathbf{x}_k,\mathbf{u}_k),}{}{} \label{eq:gcostfun}
+  \addConstraint{\mathbf{x}_{k+1} = \mathbf{f}(\mathbf{x}_k,\mathbf{u}_k) \ \forall \ k}
   \label{eq:gdyncon}
-  \addConstraint{\mathbf{g}_k(\mathbf{x}_k,\mathbf{u}_k) \leq \mathbf{0} \quad \forall \ k}
+  \addConstraint{\mathbf{g}_k(\mathbf{x}_k,\mathbf{u}_k) \leq \mathbf{0} \ \forall \ k}
   \label{eq:ineqcon}
-  \addConstraint{\mathbf{h}_k(\mathbf{x}_k,\mathbf{u}_k) = \mathbf{0} \quad \forall \ k.}
+  \addConstraint{\mathbf{h}_k(\mathbf{x}_k,\mathbf{u}_k) = \mathbf{0} \ \forall \ k.}
   \label{eq:eqcon}
 \end{mini!}
-\textcolor{blue}{[Can't we just define $\mathbf{u}_N=\mathbf{0}$
-    to simplify the expression in (2a)?]}
-\textcolor{red}{[I'm not certain that is the same thing as
-    saying that $\mathbf{u}_{N}$ is undefined.
-    What I think you are saying is $\ell_{N}(\mathbf{x}_{N},
-    \mathbf{u}_{N} = \mathbf{0})$, but what is really going
-    on is that there is no dependence at all $\ell_{N}(\mathbf{x}_{N})$]}
+We have formulated the problem such that the cost and constraint functions at
+time step $k$ may only depend on the augmented control and state at time step $k$.
+Although this structure may appear limiting, the problem can
+typically be reformulated to accomodate any cost or constraint function,
+for an example see Sec. \ref{sec:fluxonium}, and the ALTRO solver,
+which we introduce in the following discussion, exploits this
+structure to efficiently solve the problem.
 
@@
 In this sense, the TDSE \eqref{eq:tdse} is a constraint that may be violated
-for intermediate steps of the optimization, where the states need not be physical.
+for intermediate steps of the optimization, where the quantum states need not be physical.
 The direct approach lends itself to a nonlinear program formulation, for which
 a variety of general-purpose solvers exist \cite{gill2005snopt, wachter2006implementation}.
+
 Recent state-of-the-art solvers, such as ALTRO,
@@
 i.e., solving \eqref{eq:gcostfun}-\eqref{eq:gdyncon}.
-\textcolor{blue}{[Start with the familiar or the understandable explanation.
-    The imposing sounding names can be given afterwards. (Purely psychology.)]}
-\textcolor{red}{[resolved]}
 First, iLQR uses an initial guess for the augmented controls to obtain the 
@@
 about the current augmented controls and states.
-These models are used in a recurrence relation between time steps
-to derive the locally optimal update for the augmented controls.
-\textcolor{blue}{[for some quantity (which?) at consecutive time steps]}
-\textcolor{red}{[resolved]}
+These models are used with a recurrence relation between time steps
+to obtain the locally optimal update for the augmented controls.
+This recurrence relation is possible to derive in closed
+form because cost function contributions come only from the augmented
+control and state at a single time step \cite{mayne1966a}.
 Finally, a line search \cite{zhang2006global}
-is performed in the direction of this update to ensure a
+is performed in the direction of the locally optimal update to ensure a
 decrease in the objective.
-\textcolor{blue}{[again, I suspect that (2a) is not the ``objective"; rather the objective
-    is what I know as cost function, which would exclude the minimization that is part of (2a).]}
-\textcolor{red}{[resolved, i introduce objective earlier now]}
-This procedure
-is repeated until convergence is reached.
+This procedure is repeated until convergence is reached.
 
@@
 constraints \eqref{eq:ineqcon}-\eqref{eq:eqcon}.
-\textcolor{blue}{[Seems like my edit attempts and understanding
-    of the two-stage process are still not appropriate.
-    On the previous page, it seemed like the projected Newton
-    method was the final stage and a direct method at that.
-    However, here it sounds now like it is not good enough
-    and we need yet another method? I'm confused.]}
-\textcolor{red}{[removed discussion of projected gradient methods]}
 For QOC, a popular approach to handle such constraints
 is to add the constraint functions to the objective
-\cite{heeres2017implementing, leung2017speedup, reinhold2019controlling}.
-\textcolor{blue}{[(2a) is of the form $\text{min}_x L(x)$
-    do you really mean to add to the minimum? Or only to the cost function $L(x)$?]}
-\textcolor{red}{[resolved, i introduce objective earlier now]}
+\cite{heeres2017implementing, leung2017speedup, reinhold2019controlling, niu2019universal}.
 However, this strategy does not guarantee that the constraints
@@
 to small decreases in the constraint violations near the locally optimal solution
-\cite{bertsekas1996constrained}.
+\cite{bertsekas2014constrained}.
 To address this shortcoming, ALTRO
@@
 For more information on the details of the ALTRO
-solver, see \cite{howell2019altro, Jackson2020altroc}.
+solver, see Refs. \cite{howell2019altro, Jackson2020altroc}.
 
 As opposed to standard QOC solvers, ALTRO
-can satisfy constraints
-on both the augmented controls and states to tight tolerances.
+can satisfy constraints on both the control fields and quantum states to tight tolerances.
 This advantage is crucial for this work, where multiple medium-priority cost functions

### 2021-03-26 paper/3fluxonium.tex
final thomas edits

@@
 \cite{earnest2018realization, lin2018demonstration,
-  manucharyan2009fluxonium, nguyen2019high,
-  zhang2020universal}
-\textcolor{blue}{[for a physics journal, you will want as much handholding and explanation
-    for the robotics community language as possible; by contrast, clarifying that
-    a quantum gate is a unitary tranformation seems out of place here.]}
-\textcolor{red}{[resolved, removed unitary comment]}.
+  manucharyan2009fluxonium, somoroff2021millisecond, nguyen2019high,
+  zhang2020universal}.
 In this section, we use the trajectory optimization
@@
 
-First, we introduce the augmented control and state for this problem.
+First, we introduce the augmented control and state for the fluxonium gate problem.
+Since the ALTRO implementation we use does not currently
+support complex numbers, we represent the quantum states
+in the isomorphism $\mathcal{H}(\mathbb{C}^{n})
+\cong \mathcal{H}(\mathbb{R}^{2n})$ given in \cite{leung2017speedup},
+\begin{equation}
+  H \ket{\psi} \cong \begin{pmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
+    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{pmatrix}
+  \begin{pmatrix} \ket{\psi}_{\textrm{re}} \\ \ket{\psi}_{\textrm{im}}\end{pmatrix}.
+  \label{eq:isomorphism}
+\end{equation}
+We use $\psi$ -- abandoning bra-ket notation -- to denote the real representation of a state
+given by the right-hand-side of \eqref{eq:isomorphism}.
 To refer to the discrete moments of the flux, we introduce the notation
-$\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
-$a(t_{k}) \equiv a_{k}$,
-$\mathrm{d}^{n}a(t)/{\mathrm{d}t}^{n} \lvert_{t = t_{k}} \equiv \mathrm{d}^{n}_{t} a_{k}$.
+$\int_{t} a_{k} \equiv \int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t$,
+$a_{k} \equiv a(t_{k})$,
+$\mathrm{d}^{n}_{t} a_{k} \equiv \mathrm{d}^{n}a(t)/{\mathrm{d}t}^{n} \lvert_{t = t_{k}}$.
 The augmented control and state are:
@@
   \renewcommand*{\arraystretch}{1.3}
-  \mathbf{u}_{k} = \begin{bmatrix} \mathrm{d}^{2}_{t} a_{k} \end{bmatrix}, \quad
-  \mathbf{x}_{k} = \begin{bmatrix} \ket{\psi^{0}_{k}} \\ \ket{\psi^{1}_{k}}
-    \\ \int_{t} a_{k} \\ a_{k} \\ \mathrm{d}_{t} a_{k} \end{bmatrix}.
+  \mathbf{u}_{k} = \begin{pmatrix} \mathrm{d}^{2}_{t} a_{k} \end{pmatrix}, \quad
+  \mathbf{x}_{k} = \begin{pmatrix} \psi^{0}_{k} \\ \psi^{1}_{k}
+    \\ \int_{t} a_{k} \\ a_{k} \\ \mathrm{d}_{t} a_{k} \end{pmatrix}.
   \endgroup
@@
 Here, the superscript on the quantum states $i \in \{0, 1\}$ acts as a label.
-\textcolor{blue}{(We maintain vector notation for $\mathbf{u}_k$ in anticipation
-  of augmentation with additional components in Section \ref{sec:longitude}.)]}
-\textcolor{red}{[I think this sentence is redundant given
-    elaboration of augmented now in Section \ref{sec:background}]}
-Since the ALTRO implementation we use does not currently
-support complex numbers, we represent the states
-in the isomorphism $\mathcal{H}(\mathbb{C}^{n})
-\cong \mathcal{H}(\mathbb{R}^{2n})$ given in \cite{leung2017speedup},
-\begin{equation}
-  H \ket{\psi} \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
-    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
-  \begin{bmatrix} \ket{\psi}_{\textrm{re}} \\ \ket{\psi}_{\textrm{im}}\end{bmatrix}.
-  \label{eq:isomorphism}
-\end{equation}
 In standard QOC frameworks, the derivatives of the control fields
-are obtained with finite difference methods, e.g.
+are obtained with finite difference methods, e.g.,
 $\mathrm{d}_{t} a_{k} \approx (a_{k + 1} - a_{k}) / \Delta t$ \cite{leung2017speedup}.
-Because iLQR requires that cost functions depend only on the augmented control and state at a given
-time step, i.e. are of the form $\ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k})$,
-we make $\mathrm{d}^{2}_{t} a_{k}$ the decision variable and
+Because ALTRO requires that cost functions do not use information from multiple time steps,
+we make $\mathrm{d}^{2}_{t} a_{k}$ a decision variable and
 numerically integrate coupled ODEs to obtain $\mathrm{d}_{t} a_{k}$, $a_{k}$, and $\int_{t} a_{k}$
@@
 and the given flux $a_{k}$. These numerical integration rules are implemented
-in the discrete dynamics function for the problem, and give rise to the
+in the discrete dynamics function for the problem, and they give rise to the
 dynamics constraint \eqref{eq:dyn_con}.
 
-Next, we outline the constraints for the fluxonium gate problem.
+Next, we outline the constraints for this problem.
 Casting this problem in terms of a multi-state transfer problem, we fix as the initial states
@@
 to ensure the solver does not take advantage of discretization errors in numerical integration.
+For the flux moments,
+we have the initial condition $\int_{t} a_{1} = \mathrm{d}_{t} a_{1} = 0$
+\eqref{eq:ic_con}.
+We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:ic_con}, \eqref{eq:znf_con}
+so the gates may be concatenated arbitrarily. 
 We impose the zero net-flux constraint $\int_{t} a_{N} = 0$
@@
 \cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
-The flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
-\eqref{eq:amp_con} to ensure the two-level approximation
-\eqref{eq:hamiltonian} remains valid.
-We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con}
-\textcolor{blue}{[Thomas, this would actually be consistent with
-    my suggestion to set $\mathbf{u}_N$ to zero?]}
-\textcolor{red}{[Not quite, because $a_{N}$ is a component of $\mathbf{x}_{N}$,
-    see \eqref{eq:astatecontrols}]}
-so the gates may be concatenated arbitrarily. Additionally,
-we have the initial condition $\int_{t} a_{1} = \mathrm{d}_{t} a_{1} = 0$
-\eqref{eq:ic_con}. All gates presented in this work satisfy these constraints to
+Additionally, the flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz}
+\ \forall \ k$ \eqref{eq:amp_con} to ensure the two-level approximation
+\eqref{eq:hamiltonian} remains valid. Above $0.5$ GHz,
+we experimentally observe that the relationship
+between the energy levels and the flux becomes strongly non-linear.
+All gates presented in this work satisfy these constraints to
 a maximum violation of $\sim 10^{-8}$.
@@
 + \mathbf{u}^{T}_{k} R_{k} \mathbf{u}_{k}$
-where $Q_{k}$ and $R_{k}$ are diagonal matrices we supply
-that assign weights to the components in the cost function.
-\textcolor{blue}{[``We supply" -- sure, but not very helpful.
-    How are they defined/constructed specifically?]}
-\textcolor{red}{[in QOC speak these are cost function weights, they are hyperparameters
-and set based on heuristics]}
+where $Q_{k}$ and $R_{k}$ are diagonal matrices
+of hyperparameters that assign weights to cost function contributions.
 The $Q_{k}$ term
 penalizes deviations from the target augmented state
-$\mathbf{x}_{T} = \textrm{col}(\ket*{\psi^{0}_{T}}, \ket*{\psi^{1}_{T}}, 0, 0, 0)$,
-which is consistent with the constraints we have imposed on $\ket*{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$.
+$\mathbf{x}_{T} = (\psi^{0}_{T}, \psi^{1}_{T}, 0, 0, 0)^{T}$,
+which is consistent with the constraints we have
+imposed on $\ket*{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$.
 Accordingly, this term penalizes the squared
-difference of $\ket*{\psi^{i}_{k}}$ and $\ket*{\psi^{i}_{T}}$
+difference of $\psi^{i}_{k}$ and $\psi^{i}_{T}$
 and penalizes the norm of $\int_{t} a_{k}$, $a_{k}$, and $\mathrm{d}_{t} a_{k}$.
@@
 \begin{mini!}[2] 
-  {\substack{\mathbf{x}_1,\ldots,\mathbf{x}_N \\ \mathbf{u}_1,\ldots,\mathbf{u}_{N-1}}}{\sum_{k=1}^N {(\mathbf{x}_k-\mathbf{x}_{T})}^{T} Q_k (\mathbf{x}_k-\mathbf{x}_{T})
+  {\substack{\mathbf{x}_1,\ldots,\mathbf{x}_N \\ \mathbf{u}_1,\ldots,\mathbf{u}_{N-1}}}
+  {\sum_{k=1}^N {(\mathbf{x}_k-\mathbf{x}_{T})}^{T} Q_k (\mathbf{x}_k-\mathbf{x}_{T})
     + \sum_{k=1}^{N-1} {\mathbf{u}_k}^{T} R_k \mathbf{u}_{k}}{}{} \label{eq:costfun}
-  \addConstraint{\mathbf{x}_{k+1}}{= \mathbf{f}(\mathbf{x}_k, u_k) \quad \forall \ k}  \label{eq:dyn_con}
-  \addConstraint{\ket*{\psi^{0}_{1}} = \ket*{0}, \ket*{\psi^{1}_{1}} = \ket*{1}} \label{eq:istate_con}
+  \addConstraint{\mathbf{x}_{k+1}}{= \mathbf{f}(\mathbf{x}_k, \mathbf{u}_k) \ \forall \ k}
+  \label{eq:dyn_con}
+  \addConstraint{\ket*{\psi^{0}_{1}} = \ket*{0}, \ket*{\psi^{1}_{1}} = \ket*{1}}
+  \label{eq:istate_con}
+  \addConstraint{{\textstyle \int_{t}} a_{1} = a_{1} = \mathrm{d}_{t} a_{1} = 0}
+  \label{eq:ic_con}
   \addConstraint{\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{T}}
-    \quad \forall \  i} \label{eq:tstate_con}
+    \ \forall \  i} \label{eq:tstate_con}
+  \addConstraint{{\textstyle \int_{t}} a_{N} = a_{N} = 0} \label{eq:znf_con}
   \addConstraint{{\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}}
-      \rvert}^{2} = 1 \quad \forall \ i, k} \label{eq:statenorm_con}
-  \addConstraint{{\textstyle \int_{t}} a_{N} = 0} \label{eq:znf_con}
-  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz} \quad \forall \ k} \label{eq:amp_con}
-  \addConstraint{a_{1} = a_{N} = 0} \label{eq:bound_con}
-  \addConstraint{{\textstyle \int_{t}} a_{1} = \mathrm{d}_{t} a_{1} = 0}. \label{eq:ic_con}
+      \rvert}^{2} = 1 \ \forall \ i, k} \label{eq:statenorm_con}
+  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz} \ \forall \ k.} \label{eq:amp_con}
 \end{mini!}
@@
 Next, we remark on our problem formulation.
-We put a cost function at all time steps \eqref{eq:costfun}
+We put a cost function at all time steps
 because it benefits the iLQR solving stage \cite{Jackson2020altroc};
-putting a cost function at all time steps does not incentivize early achievement
-of the desired gate, as in \cite{leung2017speedup}, due to the target state constraint.
-\textcolor{blue}{[NOT to incentivize or to incentivize??]}
-\textcolor{red}{[NOT to incentivize]}
-In addition to the target-state cost function, we
-impose a target-state constraint \eqref{eq:tstate_con} which
+although this may incentivize early achievement of the desired gate,
+as in Ref. \cite{leung2017speedup}, we are primarily concerned
+with achieving the gate at the final time step, which the target-state
+constraint \eqref{eq:tstate_con} ensures.
+Additionally, the target-state constraint
 requires the final state to match the target state, including its global phase,
@@
 In extraordinarily difficult cases of
-QOC \cite{abdelhafez2020universal}, it may be impossible
-to obey the physics of the system and achieve the desired gate, i.e.
-the dynamics constraint \eqref{eq:dyn_con}
+QOC, it may be impossible
+to obey the physics of the system and achieve the desired gate \cite{abdelhafez2020universal},
+i.e., the dynamics constraint \eqref{eq:dyn_con}
 and the target-state constraint \eqref{eq:tstate_con} may be mutually unsatisfiable.
@@
 
-Finally, for the indirect ALM-iLQR solving stage,
-the states at each time step are obtained by integrating
-the TDSE with the discrete dynamics function,
-so the dynamics constraint \eqref{eq:dyn_con} is satisfied by construction.
-This is not the case for the direct projected Newton stage, where the states
-are free parameters that are adjusted to satisfy the TDSE.
-Although the final solution's deviation from the TDSE is never more than the maximum constraint violation,
+Finally, for ALTRO's first indirect stage,
+the augmented states are obtained explicitly with the discrete
+dynamics function, so the dynamics constraint and initial conditions
+\eqref{eq:dyn_con}-\eqref{eq:ic_con} are satisfied by construction.
+In this stage, the rest of the constraint functions \eqref{eq:tstate_con}-\eqref{eq:amp_con}
+are added to the objective in their isomorphism-equivalent form \eqref{eq:isomorphism}.
+Conversely, for the second direct stage, all of the constraints
+\eqref{eq:dyn_con}-\eqref{eq:amp_con} are used to define
+the projection onto the constraint manifold, and the objective is unmodified.
+Hence, the quantum states become
+free parameters that are adjusted to satisfy the TDSE.
+Although the final solution's deviation from the TDSE is never
+more than the maximum constraint violation,
 we explicitly integrate the TDSE when reporting gate errors to ensure accuracy.
-Exploring the benefit of direct optimization approaches for QOC is an interesting direction for future work.
+Exploring the benefit of direct optimization approaches
+for QOC is an interesting direction for future work.

### 2021-03-26 paper/4longitude.tex
final thomas edits

@@
     are plotted at the time-averaged 
-    absolute flux and $T_{1}$ time for each pulse.
+    absolute flux and the time-averaged $T_{1}$ time for each pulse.
     (c) Cumulative gate errors due to depolarization as a function of the
@@
 The integrated depolarization rate is appended to the augmented state \eqref{eq:astatecontrols}
-and its norm is penalized in the objective by setting
-the corresponding element of the target augmented state to zero.
+and its norm is penalized in the $Q_{k}$ term of the objective by setting
+the corresponding element of the target augmented state to zero, see \eqref{eq:costfun}.
 $T_{1}$ as a function of the flux is obtained by evaluating
-a spline fit to experimental data, see Figure \ref{fig:longitudeb}.
-\textcolor{blue}{[Phys. Rev. will want this to read 1(b).
-    Not sure how to change this the way you are doing automatic referencing.]}
-\textcolor{red}{[resolved]}
+a spline fit to experimental data, see Fig. \ref{fig:longitudeb}.
 
@@
 
-\textcolor{blue}{[abrupt start of a new paragraph; provide
-    transition/motivation for what's next; the entire following
-    paragraph feels a bit out of place/inserted after the fact.]}
-\textcolor{red}{[better?]}
 To perform time-optimal control, we make
@@
 are approximately periodic
-with amplitudes $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}.
+with amplitudes $\sim 0.2 \textrm{GHz}$, see Fig. \ref{fig:longitudea}.
 They are reminiscent of the analytically determined Floquet operations
-for a fluxonium described in \cite{huang2020engineering}
+for a fluxonium described in \cite{huang2021engineering}
 and realized in \cite{mundada2020floquet}.
@@
 numerical flux pulses
-spend more time at higher flux values, achieving higher $T_{1}$ times on average,
-see Figure \ref{fig:longitudeb}.
+spend more time at larger flux values, achieving higher $T_{1}$ times on average,
+see Fig. \ref{fig:longitudeb}.
 The single-gate errors for both the analytic and numerical gates are
-less than $10^{-4}$, which makes them sufficient for quantum error correction--a
+less than $10^{-4}$, which makes them sufficient for quantum error correction -- a
 prerequisite for fault-tolerant quantum
@@
 the gate error due to depolarization is approximately linear
-in the gate count, which we expect for $t \ll T_{1}$, see Figure \ref{fig:longitudec}.
+in the gate count, which we expect for $t \ll T_{1}$, see Fig. \ref{fig:longitudec}.
 The gate error reduction for large gate counts is
@@
 applications. These improvements are significant given the constraints
-\textcolor{blue}{[improvements are significant FOR THE CONSTRAINTS?]}
-\textcolor{red}{[changed "for" to "given"]}
 we have imposed on the gates,

### 2021-03-26 paper/5static.tex
final thomas edits

@@
+\section{Robustness to Static Parameter Uncertainty \label{sec:static}}
+We have formulated the QOC
+problem as an \emph{open-loop} optimization problem, i.e.,
+we do not incorporate feedback from the experiment into the optimization.
+However, the precise device parameters will deviate from the parameters we use in optimization,
+leading to poor experimental performance. We combat errors
+of this form using robust control techniques,
+making the state evolution insensitive
+to parameter uncertainty. As an example,
+we mitigate errors arising from the drift and finite measurement
+precision of the qubit frequency, which modifies the fluxonium Hamiltonian
+\eqref{eq:hamiltonian} by $f_{q} \rightarrow f_{q} + \delta f_{q}$.
+We consider three robust control techniques to accomplish this task:
+a sampling method, an unscented sampling method,
+and a derivative method.
+
+\subsection{Sampling Method}
+The sampling method incentivizes the optimizer
+to ensure that multiple copies of a state, each evolving
+with a distinct value of the uncertain parameter, achieve
+the same target state. Variants of this technique have been proposed
+in the context of QOC
+\cite{allen2019robust, khaneja2005optimal,
+  reinhold2019controlling, rembold2020introduction}.
+For each initial state,
+we add two sample states $\ket*{\psi^{\pm}}$
+to the augmented state \eqref{eq:astatecontrols}. The discrete dynamics
+function is modified
+so the sample states evolve under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
+with $f_{q} \rightarrow f_{q} \pm \sigma_{f_{q}}$ for a fixed
+ standard deviation $\sigma_{f_{q}}$ of the qubit frequency, acting as a hyperparameter.
+We penalize the infidelities of the sample states with respect to the target state
+by adding a cost function to the objective of the form
+$\sum_{k, \pm} b_{k} (1 - {\lvert \braket*{\psi_{T}}{\psi^{\pm}_{k}} \rvert}^{2})$
+where $b_{k}$ is a constant we supply.
+For this method, the standard orthonormal basis states are an insufficient choice
+for the initial states. As an example, a $Z/2$ gate achieved by idling
+at the flux frustration point ($a_{k} = 0 \ \forall \ k$)
+will be robust to qubit frequency detunings for the initial states $\ket{0}$
+or $\ket{1}$ because the infidelity metric is insensitive to global phases,
+but this gate will not be robust for any other initial states.
+Therefore, we choose the four initial states $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
+(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized},
+whose outer products span the operators on the Hilbert space,
+and  we refer to them as the operator basis.
+
+\subsection{Unscented Sampling Method}
+Whereas the sampling method penalizes the deviations of the sample states
+from the target state, the unscented sampling method
+penalizes the deviations of the sample states from the nominal state
+\cite{howell2020direct, lee2013sigma,
+  thangavel2020robust}. Accordingly, the cost function we add
+to the objective takes the form
+$\sum_{k, j} c_{k} (\psi^{j}_{k} - \psi_{k})^{T}
+(\psi^{j}_{k} - \psi_{k})$, where $c_{k}$ is a
+constant we supply, $\psi_{k}$ is
+the evolved initial state (nominal state), and $\psi^{j}_{k}$ is a sample state
+that evolves under a modified Hamiltonian similar to that in the sampling method.
+The sample states are chosen to encode a unimodal distribution over
+the $2n$ elements of the nominal state, modeling the uncertainty in the state
+as a result of the uncertainty in the parameter. We use the unscented transform
+\cite{julier2004unscented, uhlmann1995dynamic}
+to accurately propagate the mean and covariance of this distribution between
+time steps, or equivalently, through the transformation of the TDSE \eqref{eq:tdse}.
+Unlike the sampling method, the cost function for the unscented sampling method
+is sensitive to global phases. Accordingly, we
+do not observe a performance increase when
+using more than one initial state. 
+A detailed procedure for the unscented transformation is given
+in Appendix \ref{appendix:unscented}.
+
+\subsection{Derivative Method}
+The derivative method penalizes the sensitivity of the state
+to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
+state derivative $\ket*{\partial_{f_{q}}^{l} \psi} \equiv \partial_{f_{q}}^{l} \ket*{\psi}$.
+In the $m$\textsuperscript{th}-order
+derivative method, we append all state derivatives of order $1, \dots, m$
+to the augmented state \eqref{eq:astatecontrols}
+for each initial state.
+We obtain the state derivatives at each time step by performing forward-mode
+differentiation on the TDSE \eqref{eq:tdse}.
+For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
+\begin{align}
+  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi},\\
+  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
+  H \ket{\partial_{f_{q}} \psi} +
+  (\partial_{f_{q}} H) \ket{\psi}.
+  \label{eq:d1dyn}
+\end{align}
+We integrate the coupled ODEs with exponential
+integrators, see Appendix \ref{appendix:derivative}.
+While the state $\ket*{\psi}$ has unit norm,
+the state derivatives $\ket*{\partial^{l}_{f_{q}} \psi}$ need not, as is evident
+from the non-unitary dynamics \eqref{eq:d1dyn}.
+We penalize the norms of the isomorphism-equivalent state derivatives
+in the $Q_{k}$ term of the objective by setting the corresponding elements
+of the target augmented state to zero, see \eqref{eq:costfun}. Intuitively, this corresponds to penalizing
+the sensitivity of each state element to the uncertain parameter. As was the case for
+the unscented sampling
+method, we do not observe a performance increase when using more than one initial state
+for the derivative method.
+We present the runtimes of our implementations of the three robust control methods
+in Appendix \ref{appendix:time}.
+
 %% F2
@@
 
-\section{Robustness to Static Parameter Uncertainty \label{sec:static}}
-We have formulated the QOC
-problem as an \emph{open-loop} optimization problem, i.e.,
-we do not incorporate feedback from the experiment into the optimization.
-However, the precise device parameters will realistically deviate from the parameters we use in optimization,
-leading to poor experimental performance. We combat errors
-of this form using robust control techniques,
-making the state evolution insensitive
-to parameter uncertainty. As an example,
-we mitigate errors arising from the drift and finite measurement
-precision of the qubit frequency, which modifies the fluxonium Hamiltonian
-\eqref{eq:hamiltonian} by $f_{q} \rightarrow f_{q} + \delta f_{q}$.
-We consider three robust control techniques to accomplish this task:
-a sampling method, an unscented sampling method,
-and a derivative method.
-
-\subsection{Sampling Method}
-\textcolor{blue}{[Consider structuring the following into subsections to clearly separate the three methods.]}
-\textcolor{red}{[better?]}
-The sampling method incentivizes the optimizer
-to ensure that multiple copies of a state, each evolving
-with a distinct value of the uncertain parameter, achieve
-the same target state. Variants of this technique have been proposed
-in the context of QOC
-\cite{allen2019robust, ball2020software, khaneja2005optimal,
-  reinhold2019controlling, rembold2020introduction} and applied
-experimentally \cite{carvalho2020error}.
-For each initial state,
-we add two sample states $\ket{\psi^{\pm}}$
-to the augmented state \eqref{eq:astatecontrols}. The discrete dynamics
-function is modified
-so the sample states evolve under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
-with $f_{q} \rightarrow f_{q} \pm \sigma_{f_{q}}$ for a fixed
- standard deviation $\sigma_{f_{q}}$ of the qubit frequency, acting as a hyperparameter.
-We penalize the infidelities of the sample states with respect to the target state
-by adding a cost function to the objective of the form
-$\sum_{k, \pm} b_{k} (1 - {\lvert \braket*{\psi^{\pm}_{k}}{\psi_{T}} \rvert}^{2})$
-where $b_{k}$ is a constant we supply.
-For this method, the standard orthonormal basis states are an insufficient choice
-for the initial states. As an example, a $Z/2$ gate achieved by idling
-at the flux frustration point ($a_{k} = 0 \ \forall \ k$)
-will be robust to qubit frequency detunings for the initial states $\ket{0}$
-or $\ket{1}$ because the infidelity metric is insensitive to global phases,
-but this gate will not be robust for any other initial states.
-Therefore, we choose the four initial states $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
-(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized},
-whose outer products span the operators on the Hilbert space,
-and  we refer to them as the operator basis.
-\textcolor{blue}{[I cannot find the term ``operator basis" in the rest of the paper.
-    If not used, then don't introduce terminology.]}
-\textcolor{red}{[Operator basis is used in the appendix \ref{appendix:time}]}
-
-\subsection{Unscented Sampling Method}
-Whereas the sampling method penalizes the deviations of the sample states
-from the target state, the unscented sampling method
-penalizes the deviations of the sample states from the nominal state
-\cite{howell2020direct, lee2013sigma,
-  thangavel2020robust}. Accordingly, the cost function we add
-to the objective takes the form
-$\sum_{k, j} c_{k} (\psi^{j}_{k} - \psi_{k})^{T}
-(\psi^{j}_{k} - \psi_{k})$, where $c_{k}$ is a
-constant we supply, $\psi_{k}$ is
-the evolved initial state (nominal state), and $\psi^{j}_{k}$ is a sample state
-that evolves under a modified Hamiltonian similar to that in the sampling method.
-We omit bra-ket notation here to emphasize that
-the states are real vectors and are given by the right-hand-side of the
-complex-to-real isomorphism \eqref{eq:isomorphism}.
-The sample states are chosen to encode a unimodal distribution over
-the $2n$ elements of the nominal state, modeling the uncertainty in the state
-as a result of the uncertainty in the parameter. We use the unscented transform
-\cite{julier2004unscented, uhlmann1995dynamic}
-to accurately propagate the mean and covariance of this distribution between
-time steps, or equivalently, through the transformation of the TDSE \eqref{eq:tdse}.
-Unlike the sampling method, the cost function for the unscented sampling method
-is sensitive to global phases. Accordingly, we
-do not observe a performance increase when
-using more than one initial state. 
-A detailed procedure for the unscented transformation is given
-in Appendix \ref{appendix:unscented}.
-
-\subsection{Derivative Method}
-The derivative method penalizes the sensitivity of the state
-to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
-state derivative $\ket*{\partial_{f_{q}}^{l} \psi} \equiv \partial_{f_{q}}^{l} \ket*{\psi}$.
-In the $m$\textsuperscript{th}-order
-derivative method, we append all state derivatives of order $1, \dots, m$
-to the augmented state vector \eqref{eq:astatecontrols}
-for each initial state.
-We obtain the state derivatives at each time step by performing forward-mode
-differentiation on the TDSE \eqref{eq:tdse}.
-For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
-\begin{align}
-  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi},\\
-  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
-  H \ket{\partial_{f_{q}} \psi} +
-  (\partial_{f_{q}} H) \ket{\psi}.
-  \label{eq:d1dyn}
-\end{align}
-We integrate the coupled ODEs with exponential
-integrators, see Appendix \ref{appendix:derivative}.
-\textcolor{blue}{[integrators IN A function? sounds like mix-up between the function vs. how it is implemented numerically]}
-\textcolor{red}{[removed ref. to ddf]}
-While the state $\ket*{\psi}$ has unit norm,
-the state derivatives $\ket*{\partial^{l}_{f_{q}} \psi}$ need not, as is evident
-from the non-unitary dynamics \eqref{eq:d1dyn}.
-We penalize the norms of the state derivatives
-in the objective by setting the corresponding elements
-of the target augmented state to zero. Intuitively, this corresponds to penalizing
-the sensitivity of each state element to the uncertain parameter. As was the case for
-the unscented sampling
-method, we do not observe a performance increase when using more than one initial state
-for the derivative method.
-We present the runtimes of our implementations of the three robust control methods
-in Appendix \ref{appendix:time}.
-\textcolor{blue}{[a ``technique" is abstract/general; a technique does not have a concrete runtime; do you mean computational complexity?]},
-\textcolor{red}{[better?]}
-
 \subsection{Comparison}
@@
 
-%% F3
-\begin{figure*}[ht]
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3a.png}
-    \caption{}
-    \label{fig:stochastica}
-  \end{subfigure}\hspace{0.05\textwidth}
-  \begin{subfigure}{.4\textwidth}
-    \includegraphics[width=\linewidth]{assets/f3b.png}
-        \caption{}
-    \label{fig:stochasticb}
-  \end{subfigure}
-  \caption{
-    (a) Flux pulses for $X/2$ gates robust to flux noise
-    constructed with the analytic (A),
-    sampling (S), unscented sampling (U), and the 1\textsuperscript{st}-
-    and 2\textsuperscript{nd}-order derivative methods (D1, D2).
-    (b) Cumulative gate error due to 1/$f$ flux noise for
-    successive gate applications. The cumulative gate errors for the
-    sampling, unscented sampling, and the derivative methods are indistinguishable.
-  }
-  \label{fig:stochastic}
-\end{figure*}
-
 The analytic gate corresponds to
-idling at the flux frustration point $a_{k} = 0 \ \forall \ k$, see Figure
+idling at the flux frustration point $a_{k} = 0 \ \forall \ k$, see Fig.
 \ref{fig:statica}. Its gate time $1 / 4 f_{q} \sim 18\textrm{ns}$
@@
 range of gate times, and begins to increase above $5 \cdot 10^{-5}$ for gate times
-greater than $\sim 50$ns, see Figure \ref{fig:staticb}.
+greater than $\sim 50$ns, see Fig. \ref{fig:staticb}.
 Optimization results for the sampling method reveal that it is typically
@@
 a minimum of $\sim 3.9 \cdot 10^{-5}$
-near fractions of the Larmor period $2/4f_{q} \sim 36\textrm{ns}$,
-$3/4f_{q} \sim 54\textrm{ns}$, $4/4f_{q} \sim 72\textrm{ns}$.
+near fractions of the Larmor period: $2/4f_{q} \sim 36\textrm{ns}$,
+$3/4f_{q} \sim 54\textrm{ns}$, and $4/4f_{q} \sim 72\textrm{ns}$.
 
@@
 reaches $10^{-7}$ at the Larmor period $1 / f_{q} \sim 72$ns,
-see Figure \ref{fig:staticc}.
+see Fig. \ref{fig:staticc}.
 This result mimics the
@@
 for the problem studied here due to our Hamiltonian and experimental constraints.
-We propose comparisons between composite pulses and numerical techniques
-for future work.
+A comparison between composite pulses and numerical techniques could
+be an interesting topic for future study.
 
@@
 frequencies. Notably, the analytic gate studied here cannot be extended
-to gate times other than $1 / 4 f_{q}$. 
+to gate times other than $1/4f_{q}$. 
 We can find gates using the numerical methods at
-all gate times above $18$ns, see Figure \ref{fig:staticb}.
+all gate times at and above $1/4f_{q}$, see Fig. \ref{fig:staticb}.
 These numerical methods offer an effective scheme for synchronizing

### 2021-03-26 paper/6stochastic.tex
revised jens edits
final thomas edits

@@
+%% F3
+\begin{figure*}[ht]
+  \begin{subfigure}{.4\textwidth}
+    \includegraphics[width=\linewidth]{assets/f3a.png}
+    \caption{}
+    \label{fig:stochastica}
+  \end{subfigure}\hspace{0.05\textwidth}
+  \begin{subfigure}{.4\textwidth}
+    \includegraphics[width=\linewidth]{assets/f3b.png}
+        \caption{}
+    \label{fig:stochasticb}
+  \end{subfigure}
+  \caption{
+    (a) Flux pulses for $X/2$ gates robust to flux noise
+    constructed with the analytic (A),
+    sampling (S), unscented sampling (U), and the 1\textsuperscript{st}-
+    and 2\textsuperscript{nd}-order derivative methods (D1, D2).
+    (b) Cumulative gate error due to 1/$f$ flux noise for
+    successive gate applications. The cumulative gate errors for the
+    sampling, unscented sampling, and the derivative methods are indistinguishable.
+  }
+  \label{fig:stochastic}
+\end{figure*}
+
 \section{Robustness to Time-Dependent Parameter Uncertainty \label{sec:stochastic}}
@@
 we compute the cumulative gate error
-after each application, see Figure \ref{fig:stochastic}.
+after each application, see Fig. \ref{fig:stochastic}.
 Both the analytic
-and numerical gates yield single gate errors
+and numerical gates yield single-gate errors
 sufficient for quantum error correction.

### 2021-03-26 paper/7conclusion.tex
final thomas edits

@@
 \section{Conclusion}
-We have applied state-of-the-art trajectory
-optimization techniques to mitigate decoherence and
-achieve robustness to parameter uncertainty
-errors on a quantum system.
+We have introduced state-of-the-art trajectory optimization
+techniques in the context of quantum optimal control, enabling
+us to achieve tight tolerances for multiple constraints on the
+control fields and quantum states. Using these capabilities,
+we have mitigated decoherence and
+achieved robustness to parameter uncertainty
+errors on a superconducting fluxonium qubit.
 We have proposed a scheme for suppressing
@@
 inexpensive optimization on high-dimensional quantum systems.
-\textcolor{blue}{[I would argue that a \emph{model} has no cost... Computing something does.]}
-\textcolor{red}{[better?]}
 We have also proposed the derivative method for robust control which achieves
-super-linear gate error reductions in the gate time for the static parameter
+superlinear gate error reductions in the gate time for the static parameter
 uncertainty problem we studied.
 We have shown that the derivative, sampling, and unscented sampling methods
-can mitigate 1/$f$ flux noise errors--which
+can mitigate 1/$f$ flux noise errors -- which
 dominate coherent errors for flux controlled qubits.
@@
 operations on their quantum systems.
-These methods can be used to achieve the low gate errors
+Furthermore, they can be used to achieve the low gate errors
 required for fault-tolerant quantum computing applications. Our

### 2021-03-26 paper/8ack.tex
jens edits
final thomas edits

--- 2021-01-26 paper/8ack.tex
@@
   We thank Helin Zhang for experimental assistance
-  and Taylor Howell, Jens Koch, Tanay Roy, Colm Ryan, and Daniel Weiss for useful discussions.
+  and Taylor Howell, Tanay Roy, Colm Ryan, and Daniel Weiss for useful discussions.
+  This work was made possible by many open source software projects,
+  including but not limited to:
+  DifferentialEquations.jl \cite{rackauckas2017differentialequations},
+  Distributions.jl \cite{besancon2019distributions},
+  ForwardDiff.jl \cite{revelsLubinPapamarkou2016},
+  Matplotlib \cite{hunter2007matplotlib},
+  NumPy \cite{harris2020array},
+  TrajectoryOptimization.jl \cite{howell2019altro},
+  and Zygote.jl \cite{innes2018don}.
   This work is funded in part by EPiQC, an NSF Expedition in Computing, under grant CCF-1730449.
   This work was supported by the Army Research Office under Grant No. W911NF1910016.
-  \todo{funding}.
-  \todo{cite software packages}.
 \end{acknowledgments}

### 2021-03-26 paper/a1longitude.tex
addressed jens comments
final thomas edits

@@
 depolarization
-\textcolor{blue}{[The error is due to depolarization METRICS?]}
-\textcolor{red}{[resolved, removed the word metrics]}
 are compared in Table \ref{tab:longitude} for the
-numerical experiment described in Section \ref{sec:longitude}.
+numerical experiment described in Sec. \ref{sec:longitude}.
 The ratio of the value obtained on the metric with the analytic technique
@@
 is similar across the two metrics.
-\textcolor{blue}{what is relative performance?}
-\textcolor{red}{replaced relative performance with more precise phrasing}
 
 \begin{table}[ht]
-  \begin{tabular}{c | c | c | c | c | c | c}
-         & $D_{1\textrm{A}}$ & $D_{1\textrm{N}}$ & & GE$_{\textrm{A}}$ & GE$_{\textrm{N}}$ &\\
-    Gate & $(10^{-5})$ & $(10^{-5})$ & $D_{1\textrm{A}}/ D_{1\textrm{N}}$ & $(10^{-5})$ &
-    $(10^{-5})$ & GE$_{\textrm{A}} / \textrm{GE}_{\textrm{N}}$\\
+  \begin{tabular}{| c | c | c | c | c | c | c | c | c |}
     \hline
-    Z/2 & 5.745 & 1.149 & 5.000 & 0.888   & 0.185  & 4.791 \\
+    \multirow{2}{*}{Gate} &
+    $D_{1\textrm{A}}$ &
+    $D_{1\textrm{N}}$ &
+    \multirow{2}{*}{$\dfrac{ D_{1\textrm{A}} }{ D_{1\textrm{N}} }$} &
+    $\textrm{GE}_{\textrm{A}}$ &
+    $\textrm{GE}_{\textrm{N}}$ &
+    \multirow{2}{*}{$\dfrac{ \textrm{GE}_{\textrm{A}} }{ \textrm{GE}_{1\textrm{N}} }$}\\
+    &
+    $(10^{-5})$ &
+    $(10^{-5})$ &
+    &
+    $(10^{-5})$ &
+    $(10^{-5})$ &\\
+    \hline\hline
+    Z/2 & 5.745 & 1.149 & 5.000 & 0.888   & 0.185  & 4.791\\
+    \hline
     Y/2 & 5.253 & 1.157 & 4.540 & 0.770 & 0.186   & 4.132\\
+    \hline
     X/2 & 16.251 & 2.660 & 6.109 & 2.674 & 0.432  & 6.200\\
+    \hline
   \end{tabular}
@@
 \end{equation}
-\textcolor{blue}{[I suggest to remove the rest of the sentence completely; should not state any of this for physics journal]}
-\textcolor{red}{[sentence on commutator definitions removed]}
 For depolarization, $\gamma_{\pm} = T_{\pm}^{-1}$,
 $L_{\pm} = \sigma^{\pm} \equiv (\sigma_{x} \pm i \sigma_{y})/2$.
-The depolarization times $T_{+} = T_{-} = 2 T_{1}$ are obtained at each time step
-from the spline shown in Figure \ref{fig:longitudeb}.
+We make the approximation $hf \ll k_{\textrm{B}}T$ such that
+$T_{+} = T_{-} = 2 T_{1}$, where $T_{1}$ is obtained at each time step
+from the spline shown in Fig. \ref{fig:longitudeb}.
 We obtain the $T_{1}$ values in this spline
@@
 on these measurements, consult \cite{zhang2020universal}.
-Because $T_{1}$ depends on the flux $a(t)$, so do
+Because $T_{1}$ depends on the flux, so do
 the decay rates $\gamma_{\pm}$.
@@
 In order to use exponential integrators, we employ
-the vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018},
+the vector (Choi-Jamiolkowski) isomorphism \cite{Landi2018},
 \begin{equation}
-  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}}\, \textrm{vec}({\rho}),
+  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}}\, \textrm{vec}({\rho})
+  \label{eq:veclindblad},
 \end{equation}
@@
 and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{ij} \ket{i} \otimes \ket{j}$.
-Because $a(t)$ is constant between time steps $[t_{k}, t_{k + 1})$,
+Because the flux is constant between time steps
 due to our numerical discretization,
-  \textcolor{blue}{[for real, because of the AWG? even that would get filtered,
-      right? or as an approximation because of discretization?]}
-  \textcolor{red}{[added comment bout discretization]},
-so are $H(a(t))$ and $\gamma_{\pm}(a(t))$.
-Therefore, the exact solution is
-$\textrm{vec}(\rho_{k + 1}) = {\exp}{\textstyle(}\Delta t_{k}
-\hat{\mathcal{L}}_{k}{\textstyle)} \textrm{vec}(\rho_{k})$.
+the Hamiltonian and decay rates are also constant
+between time steps.
+Therefore, the exact solution to \eqref{eq:veclindblad} is,
+\begin{equation}
+  \textrm{vec}(\rho_{k + 1}) = {\exp}{\textstyle(}\Delta t_{k}
+  \hat{\mathcal{L}}_{k}{\textstyle)} \textrm{vec}(\rho_{k}).
+\end{equation}
 The vector isomorphism transforms $(n \times n) \times (n \times n)$
@@
 Runge-Kutta on the unvectorized equation.
-\textcolor{blue}{[what is zero-order hold?]}
-\textcolor{red}{[removed ZOH comment]}
 The latter requires decreasing the interval $\Delta t_{k}$

### 2021-03-26 paper/a2unscented.tex
final thomas edits

--- 2021-03-13 paper/a2unscented.tex
@@
 We consider a state $\psi \in \mathbb{R}^{2n}$, 
-an uncertain parameter $\lambda \in \mathbb{R}^{d}$, and discrete dynamics
+an uncertain set of parameters $\lambda\in \mathbb{R}^{d}$, and discrete dynamics
 $\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
-We omit bra-ket notation to emphasize that
-the state is a real vector, and it is given by the right-hand-side of the
-complex-to-real isomorphism \eqref{eq:isomorphism}.
 The nominal initial state is given by $\bar{\psi}_{1}$ with an associated
@@
 \begin{equation}\label{eq:uupdate}
-  \begin{bmatrix} \psi_{1}^{j} \\ \lambda_{1}^{j} \end{bmatrix} =
-  \begin{bmatrix} \bar{\psi}_{1} \\ 0\end{bmatrix}
-    \pm \beta \sqrt{\begin{bmatrix} P_{1} & 0\\ 0 & L_{1}\end{bmatrix}}^{\; j}.
+  \begin{pmatrix} \psi_{1}^{j} \\ \lambda_{1}^{j} \end{pmatrix} =
+  \begin{pmatrix} \bar{\psi}_{1} \\ 0\end{pmatrix}
+    \pm \beta \sqrt{\begin{pmatrix} P_{1} & 0\\ 0 & L_{1}\end{pmatrix}}^{\; j}.
 \end{equation}
-$\beta$ is a hyperparameter that controls the spacing of the covariance contour.
+Here, $\beta$ is a hyperparameter that controls the spacing of the covariance contour.
 The $(\pm)$ is understood to take $(+)$ for $j \in \{1, \dots, 2n + d\}$ and $(-)$ for

### 2021-03-26 paper/a3derivative.tex
jens final comments
final thomas edits

@@
 For example, the dynamics for the first state derivative
-are $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
+are,
+\begin{equation}
+\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
 -\frac{i}{\hbar} H  \ket{\partial_{\lambda} \psi} -
-\frac{i}{\hbar} (\partial_{\lambda} H) \ket{\psi}$.
+\frac{i}{\hbar} (\partial_{\lambda} H) \ket{\psi}. \label{eq:gd1dyn}
+\end{equation}
 The linear term is $L = -\frac{i}{\hbar} H$ and the non-linear
 term is $N = -\frac{i}{\hbar} (\partial_{\lambda} H) \ket{\psi}$.
-With zero-order hold on the controls the exact solution is:
+With piecewise-constant controls, the exact solution to \eqref{eq:gd1dyn} is,
 \begin{equation}
@@
 This method provides a good tradeoff between accuracy and efficiency, requiring one unique matrix
-exponential computation per stage. Integration accuracy for the state derivatives is not of the
-utmost importance because they are used in the robustness cost function--as opposed
-to the states themselves which are experimental parameters that must be realized
+exponential computation per stage. Integration accuracy for the state derivatives
+is not of the utmost importance because they are used in the robustness cost function -- as
+opposed to the states themselves which are experimental parameters that must be realized
 with high accuracy.

### 2021-03-26 paper/a4time.tex
final thomas edits

--- 2021-01-26 paper/a4time.tex
@@
 \section{Computational Performance \label{appendix:time}}
-We provide runtimes for our optimizations and comment on the scaling of the
-robustness methods. The runtimes for
-the base optimization in Section \ref{sec:fluxonium},
-the depolarization optimization in Section \ref{sec:longitude},
-and the robust optimizations in Section \ref{sec:static}
+In this section we provide runtimes for our optimizations. The runtimes for
+the base optimization in Sec. \ref{sec:fluxonium},
+the depolarization optimization in Sec. \ref{sec:longitude},
+and the robust optimizations in Sec. \ref{sec:static}
 are presented in Table \ref{tab:time}
@@
   \centering
-  \begin{tabular} {c | c | c | c }
-    & \multicolumn{3}{c}{Average Runtime (s)}\\
+  \begin{tabular} {| c | c | c | c | c |}
     \hline
-    $t_{N}$ (ns) & $18$ & $36$ & $72$\\
+    \backslashbox{Method}{$t_{N} (\textrm{ns})$} & $18$ & $36$ & $72$\\
     \hline
     Base & $0.155 \pm 0.008$ & $7.0 \pm 0.4$ & $15.9 \pm 0.8$\\
+    \hline
     Depol. & $1.69 \pm 0.08$ & - & -\\
+    \hline
     S & $1.77 \pm 0.09$ & $48 \pm 2$ & $280 \pm 10$\\
+    \hline
     U & $75 \pm 4$ & $340 \pm 20$ & $400 \pm 20$\\
+    \hline
     D1 & $6.1 \pm 0.3$ & $27 \pm 1$ & $65 \pm 3$\\
+    \hline
     D2 & $15.7 \pm 0.8$ & $17.3 \pm 0.9$ & $54 \pm 3$\\
+    \hline
   \end{tabular}
   \caption{
-    Average runtimes for $Z/2$ optimizations
+    Average runtimes in seconds for $Z/2$ optimizations
     using the base, depolarization, sampling (S),
-    unscented Sampling (U), and the 1\textsuperscript{st}-
+    unscented sampling (U), and the 1\textsuperscript{st}-
     and 2\textsuperscript{nd}-order derivative methods (D1, D2).
@@
 \end{table}
-
-Now we present the problem size complexities for the robustness methods.
-For the sampling method, the size of the augmented state vector
-is $O(dn^{3})$, where $d$ is the number of uncertain parameters and
-$n$ is the dimension of the Hilbert space. There are $n^{2}$ initial states
-in the operator basis, $2d$ sample states per initial state,
-and each state has $2n$ real numbers.
-For the unscented sampling method, the size of the augmented state vector
-is $O(dn^{3} + n^{4})$.
-There are $n^{2}$ initial states in the operator basis,
-$2(2n + d)$ sample states per initial state,
-and each state has $2n$ real numbers.
-For the $m$\textsuperscript{th}-order derivative method, the size of the augmented state vector
-is $O(dmn^{3})$. There are $n^{2}$
-initial states in the operator basis, $dm$ state derivatives per initial state,
-and each state has $2n$ real numbers.

### 2021-03-29 paper/3fluxonium.tex
final edits

@@
 formalism \eqref{eq:gcostfun}-\eqref{eq:eqcon}
-to define the optimization problem \eqref{eq:costfun}-\eqref{eq:ic_con},
+to define the optimization problem \eqref{eq:costfun}-\eqref{eq:amp_con},
 which we extend in subsequent sections to account
@@
 to ensure the solver does not take advantage of discretization errors in numerical integration.
-For the flux moments,
+For the flux,
 we have the initial condition $\int_{t} a_{1} = \mathrm{d}_{t} a_{1} = 0$
@@
 Additionally, the flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz}
-\ \forall \ k$ \eqref{eq:amp_con} to ensure the two-level approximation
-\eqref{eq:hamiltonian} remains valid. Above $0.5$ GHz,
-we experimentally observe that the relationship
-between the energy levels and the flux becomes strongly non-linear.
+\ \forall \ k$ \eqref{eq:amp_con} to ensure the two-level
+approximation remains valid \eqref{eq:hamiltonian}. Above $0.5$ GHz,
+the relationship between the energy levels and the flux becomes strongly non-linear.
 All gates presented in this work satisfy these constraints to

### 2021-03-29 paper/a1longitude.tex
final edits

@@
     $\textrm{GE}_{\textrm{N}}$ &
-    \multirow{2}{*}{$\dfrac{ \textrm{GE}_{\textrm{A}} }{ \textrm{GE}_{1\textrm{N}} }$}\\
+    \multirow{2}{*}{$\dfrac{ \textrm{GE}_{\textrm{A}} }{ \textrm{GE}_{\textrm{N}} }$}\\
     &
@@
     $(10^{-5})$ &\\
-    \hline\hline
+    \hline
     Z/2 & 5.745 & 1.149 & 5.000 & 0.888   & 0.185  & 4.791\\
@@
 $L_{\pm} = \sigma^{\pm} \equiv (\sigma_{x} \pm i \sigma_{y})/2$.
-We make the approximation $hf \ll k_{\textrm{B}}T$ such that
+Our device operates in the regime where $hf \ll k_{\textrm{B}}T$ such that
 $T_{+} = T_{-} = 2 T_{1}$, where $T_{1}$ is obtained at each time step
@@
 and monitoring the resultant decay. For more details
-on these measurements, consult \cite{zhang2020universal}.
+on these measurements, consult Ref. \cite{zhang2020universal}.
 Because $T_{1}$ depends on the flux, so do

### 2021-05-17 paper/main.tex
full build
prog
updated figures, lots of proofreading
final thomas edits
updated paper

--- 2020-11-03 paper/main.tex
@@
   amsmath,
-  tbtags,
   amssymb,
-  aps,
-  nobibnotes,
-  %% prl,
+  pra,
   twocolumn,
@@
 
-\input{preamble}
+% imports
+\usepackage{appendix}
+\usepackage[english]{babel}
+\usepackage[justification=centering]{caption}
+\usepackage{float}
+\usepackage{graphicx}
+\usepackage{hyperref}
+\usepackage[utf8]{inputenc}
+\usepackage{layouts}
+\usepackage{multirow}
+\usepackage{optidef}
+\usepackage{physics}
+\usepackage{setspace}
+\usepackage{diagbox}
+\usepackage[labelformat=simple]{subcaption}
+\usepackage{xcolor}
+
+% configure imports
+\DeclareMathOperator*{\argmin}{arg\,min}
+\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
+\definecolor{darkgreen}{RGB}{46, 184, 46}
+\newcommand{\half}{\frac{1}{2}}
+\newcommand{\R}{\mathbb{R}}
+\captionsetup[subfigure]{labelformat=empty, skip=0pt}
+\captionsetup{labelsep=period}
+\restylefloat{table}
+\renewcommand\thesubfigure{(\alph{subfigure})}
+\addto\captionsenglish{%
+  \renewcommand{\figurename}{FIG.}%
+  \renewcommand{\tablename}{TABLE}%
+  \renewcommand{\appendixname}{APPENDIX}%
+}
 
@@
 %% TITLE
-\title{Robust Control of a Fluxonium Qubit}
+\title{Robust Quantum Optimal Control with Trajectory Optimization}
 
@@
 }
-\author{Brian Jackson}
-\author{Zac Manchester}
+\author{Brian E. Jackson}
 \affiliation{
-  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
+  Robotics Institute, Carnegie Mellon University, Pittsburgh, Pennsylvania 15213, USA
+}
+
+\author{Jens Koch}
+\affiliation{Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
+\affiliation{Northwestern–Fermilab Center for Applied Physics and Superconducting Technologies, Northwestern University, Evanston, Illinois 60208, USA}
+
+\author{Zachary Manchester}
+\affiliation{
+  Robotics Institute, Carnegie Mellon University, Pittsburgh, Pennsylvania 15213, USA
 }
@@
 
-\input{0abstract.tex}
+%% ABSTRACT
+\begin{abstract}
+  The ability to engineer high-fidelity gates on quantum processors in the presence of
+  systematic errors remains the primary barrier to achieving quantum advantage.
+  Quantum optimal control methods have proven effective in experimentally
+  realizing high-fidelity gates, but they require exquisite calibration to be performant.
+  We apply robust trajectory optimization techniques to suppress gate errors arising from system
+  parameter uncertainty.
+  We propose a derivative-based approach that maintains
+  computational efficiency by using forward-mode differentiation.
+  Additionally, the effect of depolarization on a gate is typically modeled by
+  integrating the Lindblad master equation,
+  which is computationally expensive.
+  We employ a computationally efficient model
+  and utilize time-optimal control to achieve high-fidelity gates in the presence of depolarization.
+  We apply these techniques to a fluxonium qubit and suppress
+  simulated gate errors due to parameter uncertainty below $10^{-7}$ for
+  static parameter deviations on the order of $1\%$.
+\end{abstract}
 
@@
 
-\input{1introduction.tex}
-\input{2background.tex}
-\input{3fluxonium.tex}
-\input{4longitude.tex}
-\input{5static.tex}
-\input{6stochastic.tex}
-\input{7conclusion.tex}
-\input{8ack.tex}
+%% SECTION 1
+\section{Introduction}
+Quantum optimal control (QOC) is a class of optimization
+algorithms for accurately and efficiently manipulating quantum systems.
+Early techniques were proposed for nuclear magnetic resonance experiments
+\cite{vandersypen2005nmr, kehlet2004improving, khaneja2005optimal,
+  maximov2008optimal, nielsen2010optimal, skinner2003application, tosner2009optimal},
+and applications now include superconducting circuits \cite{abdelhafez2020universal,
+  chakram2020multimode, egger2013optimized, fisher2010optimal, gokhale2019partial,
+  huang2014optimal, heeres2017implementing, kelly2014optimal, leng2019robust,
+  leung2017speedup, li2020fast,
+  liebermann2016optimal, reinhold2019controlling,
+  rebentrost2009optimal, rebentrost2009optimal2, spiteri2018quantum,
+  sporl2007optimal},
+neutral atoms and ions \cite{brouzos2015quantum,
+  de2008optimal, grace2007optimal, goerz2011quantum, guo2019high, jensen2019time,
+  larrouy2020fast, nebendahl2009optimal, omran2019generation,
+  rosi2013fast,
+  treutlein2006microwave, van2016optimal},
+nitrogen-vacancy centers in diamond \cite{chou2015optimal,
+  dolde2014high, geng2016experimental,
+  nobauer2015smooth, poggiali2018optimal, rembold2020introduction, tian2019optimal},
+and Bose-Einstein condensates \cite{amri2019optimal, doria2011optimal,
+  sorensen2019qengine, sorensen2018quantum}.
+In the context of quantum computation,
+optimal control is employed to achieve high-fidelity gates
+while adhering to experimental constraints.
+Experimental errors such as parameter drift, noise, and
+finite control resolution cause the system to deviate
+from the model used in optimization, hampering
+experimental performance
+\cite{chakram2020multimode, heeres2017implementing, klimov2020snake,
+  omran2019generation, reinhold2019controlling}.
+Robust control improves upon
+standard optimal control by encoding
+model parameter uncertainties
+in optimization objectives, yielding performance
+guarantees over a range of parameter values \cite{Zhou97,Morimoto00,Manchester18}.
+We adapt robust control techniques from the robotics community to mitigate
+parameter-uncertainty errors for
+a superconducting fluxonium qubit.
+
+Analytically-derived control pulses that mitigate parameter-uncertainty
+errors include composite pulses \cite{cummins2000use, cummins2003tackling,
+  kupce1995stretched, merrill2014progress},
+pulses designed by considering dynamic and geometric phases
+\cite{han2020experimental, xu2020nonadiabatic, dridi2020optimal}, and
+pulses obtained with the DRAG scheme \cite{motzoi2009simple}.
+As compared to analytical techniques, QOC is advantageous for
+designing pulses that consider all experimental constraints and
+performance tradeoffs
+\cite{leung2017speedup},
+and for constructing operations without a known analytic solution
+\cite{chakram2020multimode, heeres2017implementing}.
+Accordingly, recent work has sought to achieve robustness in QOC
+frameworks using closed-loop methods
+\cite{egger2014adaptive, feng2018gradient,
+  li2017hybrid, wittler2020integrated, pechen2012trap}
+and open-loop methods
+\cite{ball2020software, carvalho2020error, allen2019robust, reinhold2019controlling,
+  rembold2020introduction, kosut2013robust, niu2019universal, khaneja2005optimal}.
+
+In this work, we study three open-loop robust control techniques that
+make the quantum state trajectory less sensitive to the uncertainties
+of static and time-dependent parameters:
+\begin{enumerate}
+\item A sampling method, similar to the work in Refs.~
+  \cite{allen2019robust, khaneja2005optimal, reinhold2019controlling, rembold2020introduction}.
+\item An unscented sampling method \cite{howell2020direct, lee2013sigma, thangavel2020robust}
+  adapted from the unscented transform \cite{julier2004unscented,
+    uhlmann1995dynamic} used in state estimation.
+  \item A derivative method, which penalizes the sensitivity of the quantum state trajectory
+    to uncertain parameters.
+\end{enumerate}
+We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
+We also show that QOC can solve important problems associated with
+fluxonium-based qubits: exploiting the dependence of $T_{1}$ on the controls
+to mitigate depolarization
+and synchronizing the phase of qubits with distinct frequencies.
+To ameliorate depolarization,
+we perform time-optimal control and
+employ an efficient depolarization model
+for which the computational cost is independent of the
+Hilbert space dimension.
+Leveraging recent advances in trajectory optimization within the field of robotics, we
+solve these optimization problems using ALTRO (Augmented Lagrangian TRajectory Optimizer)
+\cite{howell2019altro}, which can enforce constraints on
+the control fields and the quantum state trajectory.
+
+This paper is organized as follows.
+First, we describe ALTRO in the context of QOC
+in Sec. \ref{sec:background}.
+We outline realistic constraints for operating the fluxonium and
+define the associated QOC problem in Sec. \ref{sec:fluxonium}.
+Then, we formulate a method for suppressing depolarization
+in Sec. \ref{sec:longitude}. Next, we describe three techniques for achieving
+robustness to static parameter uncertainties in Sec. \ref{sec:static}. We
+adapt the same techniques to mitigate 1/$f$ flux noise
+in Sec. \ref{sec:stochastic}.
+
+\section{Background \label{sec:background}}
+In this section, we
+review the QOC problem statement
+and describe the ALTRO solver \cite{howell2019altro}.
+QOC concerns a vector $\mathbf{a}(t)$ of time-dependent control fields
+that steer the evolution of a quantum state $\ket{\psi(t)}$.
+The evolution of the state is governed by the time-dependent
+Schr{\"o}dinger equation (TDSE),
+\begin{equation}
+  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(\mathbf{a}(t), t) \ket{\psi(t)}.
+  \label{eq:tdse}
+\end{equation}
+The Hamiltonian $H(\mathbf{a}(t), t)$ is determined by the quantum system and the external control fields.
+The QOC problem is to find the
+controls that minimize a functional $J[\mathbf{a}(t)]$,
+which we call the objective.
+To make the problem numerically tractable,
+the quantum state and controls are discretized into $N$ time steps,
+$\ket{\psi(t_{k})} \to \ket{\psi_k}$ and $\mathbf{a}(t_{k}) \to \mathbf{a}_k$ where
+$t_{k} = t_{k - 1} + \Delta t$ and $k \in \{1, ..., N\}$.
+In the case of a single state-transfer problem, the objective is
+the infidelity of the time-evolved final state $\ket{\psi_N}$ and
+the intended target state $\ket{\psi_T}$,
+$J(\mathbf{a}) = 1 - {\lvert \braket{\psi_{T}}{\psi_{N}(\mathbf{a})} \rvert}^{2}$.
+Standard QOC solvers compute derivatives of the objective $\nabla J(\mathbf{a})$,
+which can easily be used to implement first-order optimization methods
+\cite{machnes2015tunable, khaneja2005optimal, leung2017speedup, goerz2019krotov}.
+
+Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
+and solved using specialized solvers developed by the robotics community
+\cite{Schulman13, Tedrake16, Hereid2017FROST, howell2019altro}.
+The objective $J(\mathbf{a})=\sum_k \ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k})$
+is expressed in terms of the
+cost function at each time step $\ell_k$, where
+$\mathbf{x}_{k}$ is the augmented state vector
+and $\mathbf{u}_{k}$ is the augmented control vector.
+We use the term \emph{augmented} because these
+vectors contain all of the relevant variables in the optimization problem,
+not just the quantum state and the control fields,
+for an example see Sec. \ref{sec:fluxonium}.
+The augmented control contains all variables that the experimentalist
+may manipulate, and the augmented state contains all variables that depend
+on those in the augmented control.
+The variables in the augmented states depend on those in the augmented
+controls as defined by the differential equations governing the physical system, which are
+encoded in the discrete relation
+$\mathbf{x}_{k + 1} = \mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$.
+For QOC, $\mathbf{f}(\mathbf{x}_{k}, \mathbf{u}_{k})$ -- which we call
+the discrete dynamics function -- propagates the quantum state by
+integrating the TDSE \eqref{eq:tdse}
+using a Runge-Kutta method \cite{jorgensen2011numerical}
+or an exponential integrator \cite{auer2018magnus, berland2006solving, einkemmer2017performance,
+  shillito2020fast}.
+
+We incorporate constraints on the augmented controls and states
+by formulating them as inequalities  $\mathbf{g}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) \leq \mathbf{0}$
+or equalities $\mathbf{h}_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = \mathbf{0}$.
+The constraint functions $\mathbf{g}_k$ and $\mathbf{h}_k$ may be vector-valued to encode multiple constraints,
+and equalities and inequalities are understood component-wise.
+To quantify constraint satisfaction, we define each constraint's \emph{violation}
+as the magnitude of its deviation:
+$\textrm{max}(g(\cdot), 0)$ or $\abs{h(\cdot)}$, where $g$ and $h$
+are components of constraint functions $\mathbf{g}_{k}$ and $\mathbf{h}_{k}$,
+respectively.
+Stated concisely, the trajectory optimization problem is:
+\begin{mini!}[2]
+  {\substack{\mathbf{x}_1, \ldots, \mathbf{x}_N \\\mathbf{u}_1, \ldots, \mathbf{u}_{N - 1}}}{
+    \sum^{N}_{k = 1}\ell_{k}
+    (\mathbf{x}_k,\mathbf{u}_k),}{}{} \label{eq:gcostfun}
+  \addConstraint{\mathbf{x}_{k+1} = \mathbf{f}(\mathbf{x}_k,\mathbf{u}_k) \ \forall \ k}
+  \label{eq:gdyncon}
+  \addConstraint{\mathbf{g}_k(\mathbf{x}_k,\mathbf{u}_k) \leq \mathbf{0} \ \forall \ k}
+  \label{eq:ineqcon}
+  \addConstraint{\mathbf{h}_k(\mathbf{x}_k,\mathbf{u}_k) = \mathbf{0} \ \forall \ k.}
+  \label{eq:eqcon}
+\end{mini!}
+We have formulated the problem such that the cost and constraint functions at
+time step $k$ may only depend on the augmented control and state at time step $k$.
+Although this structure may appear limiting, the problem can
+typically be reformulated to accomodate any cost or constraint function,
+for an example see Sec. \ref{sec:fluxonium}, and the ALTRO solver,
+which we introduce in the following discussion, exploits this
+structure to efficiently solve the problem.
+
+Standard techniques for solving \eqref{eq:gcostfun}-\eqref{eq:eqcon} typically
+fall into two categories: direct methods \cite{Hargraves87, kelly2017introduction}
+and indirect methods \cite{betts1998survey}. For indirect methods,
+the augmented controls are the \emph{decision variables}, i.e., the
+variables the optimizer adjusts to solve the problem.
+The augmented states are obtained from the augmented controls using the discrete dynamics function,
+and they are used to evaluate derivatives of the cost functions.
+Then, the derivative information is employed to update the augmented controls.
+This approach is taken by standard QOC solvers such as GOAT \cite{machnes2015tunable},
+GRAPE \cite{khaneja2005optimal, leung2017speedup}, and Krotov's method \cite{goerz2019krotov}.
+Conversely, direct methods treat both the augmented controls and states as decision
+variables. In addition to minimizing the cost functions, the optimizer uses derivative information
+for the discrete dynamics function to satisfy the dynamics constraint
+\eqref{eq:gdyncon} to a specified tolerance.
+In this sense, the TDSE \eqref{eq:tdse} is a constraint that may be violated
+for intermediate steps of the optimization, where the quantum states need not be physical.
+The direct approach lends itself to a nonlinear program formulation, for which
+a variety of general-purpose solvers exist \cite{gill2005snopt, wachter2006implementation}.
+
+Recent state-of-the-art solvers, such as ALTRO,
+combine the indirect and direct methods in a two-stage approach.
+First, ALTRO employs an indirect solving stage using
+the iterative linear-quadratic regulator (iLQR) algorithm
+\cite{Li2004a} as the internal solver of an augmented Lagrangian method (ALM)
+\cite{lantoine2012hybrid, plancher2017constrained, nocedal2006numerical}.
+In the second direct stage, ALTRO uses a projected Newton method
+\cite{bertsekas1982projected, rao1998application}.
+Next, we provide a more detailed summary of these two stages.
+
+iLQR is an indirect method
+for minimizing the objective subject to the dynamics constraint,
+i.e., solving \eqref{eq:gcostfun}-\eqref{eq:gdyncon}.
+First, iLQR uses an initial guess for the augmented controls to obtain the 
+augmented states with the discrete dynamics function.
+iLQR then constructs quadratic models for each cost function using
+their zeroth-, first- and second-order derivatives in a Taylor expansion
+about the current augmented controls and states.
+These models are used with a recurrence relation between time steps
+to obtain the locally optimal update for the augmented controls.
+This recurrence relation is possible to derive in closed
+form because cost function contributions come only from the augmented
+control and state at a single time step \cite{mayne1966a}.
+Finally, a line search \cite{zhang2006global}
+is performed in the direction of the locally optimal update to ensure a
+decrease in the objective.
+This procedure is repeated until convergence is reached.
+
+While indirect solvers like iLQR are computationally
+efficient and maintain high accuracy for the discrete dynamics
+throughout the optimization, they cannot handle
+nonlinear equality and inequality
+constraints \eqref{eq:ineqcon}-\eqref{eq:eqcon}.
+For QOC, a popular approach to handle such constraints
+is to add the constraint functions to the objective
+\cite{heeres2017implementing, leung2017speedup, reinhold2019controlling, niu2019universal}.
+However, this strategy does not guarantee that the constraints
+are satisfied as the solver trades
+minimization of the cost functions and constraint functions against each other.
+ALM remedies this issue by adaptively adjusting a Lagrange multiplier estimate
+for each constraint function to ensure the constraints are satisfied.
+ALM adds terms that are linear and quadratic in the constraint functions
+to the objective. Then, the new objective is minimized with
+iLQR. If the solution obtained with iLQR does not satisfy the constraints,
+the prefactors for the constraint terms in the objective are increased
+intelligently and the procedure is repeated.
+
+ALM converges superlinearly, but poor numerical conditioning may lead
+to small decreases in the constraint violations near the locally optimal solution
+\cite{bertsekas2014constrained}.
+To address this shortcoming, ALTRO
+projects the solution from the ALM stage onto the constraint manifold using
+a (direct) projected Newton method, achieving ultra-low
+constraint violations $\sim 10^{-8}$.
+For more information on the details of the ALTRO
+solver, see Refs. \cite{howell2019altro, Jackson2020altroc}.
+
+As opposed to standard QOC solvers, ALTRO
+can satisfy constraints on both the control fields and quantum states to tight tolerances.
+This advantage is crucial for this work, where multiple medium-priority cost functions
+are minimized subject to many high-priority constraints.
+
+\section{QOC for the Fluxonium \label{sec:fluxonium}}
+In the following, we optimize quantum gates
+for the superconducting fluxonium qubit -- a promising
+building block for quantum computers due to its high
+coherence times
+\cite{earnest2018realization, lin2018demonstration,
+  manucharyan2009fluxonium, somoroff2021millisecond, nguyen2019high,
+  zhang2020universal}.
+In this section, we use the trajectory optimization
+formalism \eqref{eq:gcostfun}-\eqref{eq:eqcon}
+to define the optimization problem \eqref{eq:costfun}-\eqref{eq:amp_con},
+which we extend in subsequent sections to account
+for experimental error channels.
+To high accuracy, we approximate the fluxonium Hamiltonian near the flux-frustration
+point as a two-level system:
+\begin{align}
+  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}.
+  \label{eq:hamiltonian}
+\end{align}
+Here, $f_{q}$ is the qubit frequency at the flux-frustration point,
+$a(t)$ is the control governing the flux offset from the flux-frustration point,
+$h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
+are Pauli matrices. Although the coherent dynamics can be described with this two-level
+system model, our noise model, experimental constraints, and system parameters
+consider the full system, and they are representative of the fluxonium
+presented in \cite{zhang2020universal}.
+
+First, we introduce the augmented control and state for the fluxonium gate problem.
+Since the ALTRO implementation we use does not currently
+support complex numbers, we represent the quantum states
+in the isomorphism $\mathcal{H}(\mathbb{C}^{n})
+\cong \mathcal{H}(\mathbb{R}^{2n})$ given in \cite{leung2017speedup},
+\begin{equation}
+  H \ket{\psi} \cong \begin{pmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
+    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{pmatrix}
+  \begin{pmatrix} \ket{\psi}_{\textrm{re}} \\ \ket{\psi}_{\textrm{im}}\end{pmatrix}.
+  \label{eq:isomorphism}
+\end{equation}
+We use $\psi$ -- abandoning bra-ket notation -- to denote the real representation of a state
+given by the right-hand-side of \eqref{eq:isomorphism}.
+To refer to the discrete moments of the flux, we introduce the notation
+$\int_{t} a_{k} \equiv \int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t$,
+$a_{k} \equiv a(t_{k})$,
+$\mathrm{d}^{n}_{t} a_{k} \equiv \mathrm{d}^{n}a(t)/{\mathrm{d}t}^{n} \lvert_{t = t_{k}}$.
+The augmented control and state are:
+\begin{equation}
+  \begingroup
+  \renewcommand*{\arraystretch}{1.3}
+  \mathbf{u}_{k} = \begin{pmatrix} \mathrm{d}^{2}_{t} a_{k} \end{pmatrix}, \quad
+  \mathbf{x}_{k} = \begin{pmatrix} \psi^{0}_{k} \\ \psi^{1}_{k}
+    \\ \int_{t} a_{k} \\ a_{k} \\ \mathrm{d}_{t} a_{k} \end{pmatrix}.
+  \endgroup
+  \label{eq:astatecontrols}
+\end{equation}
+Here, the superscript on the quantum states $i \in \{0, 1\}$ acts as a label.
+In standard QOC frameworks, the derivatives of the control fields
+are obtained with finite difference methods, e.g.,
+$\mathrm{d}_{t} a_{k} \approx (a_{k + 1} - a_{k}) / \Delta t$ \cite{leung2017speedup}.
+Because ALTRO requires that cost functions do not use information from multiple time steps,
+we make $\mathrm{d}^{2}_{t} a_{k}$ a decision variable and
+numerically integrate coupled ODEs to obtain $\mathrm{d}_{t} a_{k}$, $a_{k}$, and $\int_{t} a_{k}$
+so that we may penalize them in cost functions.
+Similarly, the quantum states are obtained by numerically integrating
+the TDSE \eqref{eq:tdse} with the fluxonium Hamiltonian \eqref{eq:hamiltonian}
+and the given flux $a_{k}$. These numerical integration rules are implemented
+in the discrete dynamics function for the problem, and they give rise to the
+dynamics constraint \eqref{eq:dyn_con}.
+
+Next, we outline the constraints for this problem.
+Casting this problem in terms of a multi-state transfer problem, we fix as the initial states
+ $\ket*{\psi^{0}_{1}} = \ket*{0}$, $\ket*{\psi^{1}_{1}} = \ket*{1}$
+\eqref{eq:istate_con}.
+The states at the final time step are constrained to be
+the target states $\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{T}} \equiv
+U \ket*{\psi^{i}_{1}} \ \forall \ i$
+\eqref{eq:tstate_con} where $U = X/2, Y/2, Z/2$ denotes the target gate.
+Furthermore, we impose the normalization constraint
+${\lvert \braket*{\psi^{i}_{k}}{\psi^{i}_{k}} \rvert}^{2} = 1 \ \forall \ i,k$
+\eqref{eq:statenorm_con}
+to ensure the solver does not take advantage of discretization errors in numerical integration.
+For the flux,
+we have the initial condition $\int_{t} a_{1} = \mathrm{d}_{t} a_{1} = 0$
+\eqref{eq:ic_con}.
+We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:ic_con}, \eqref{eq:znf_con}
+so the gates may be concatenated arbitrarily. 
+We impose the zero net-flux constraint $\int_{t} a_{N} = 0$
+\eqref{eq:znf_con}
+which mitigates the inductive drift ubiquitous in flux-bias lines
+\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
+Additionally, the flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz}
+\ \forall \ k$ \eqref{eq:amp_con} to ensure the two-level
+approximation remains valid \eqref{eq:hamiltonian}. Above $0.5$ GHz,
+the relationship between the energy levels and the flux becomes strongly non-linear.
+All gates presented in this work satisfy these constraints to
+a maximum violation of $\sim 10^{-8}$.
+
+The cost function at each time step is
+$\ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = (\mathbf{x}_{k}
+- \mathbf{x}_{T})^{T} Q_{k} (\mathbf{x}_{k} - \mathbf{x}_{T})
++ \mathbf{u}^{T}_{k} R_{k} \mathbf{u}_{k}$
+where $Q_{k}$ and $R_{k}$ are diagonal matrices
+of hyperparameters that assign weights to cost function contributions.
+The $Q_{k}$ term
+penalizes deviations from the target augmented state
+$\mathbf{x}_{T} = (\psi^{0}_{T}, \psi^{1}_{T}, 0, 0, 0)^{T}$,
+which is consistent with the constraints we have
+imposed on $\ket*{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$.
+Accordingly, this term penalizes the squared
+difference of $\psi^{i}_{k}$ and $\psi^{i}_{T}$
+and penalizes the norm of $\int_{t} a_{k}$, $a_{k}$, and $\mathrm{d}_{t} a_{k}$.
+We penalize the squared difference of the final and target
+quantum states, rather than their infidelities,
+because the Hessian of the squared-difference cost function is diagonal -- which
+makes matrix multiplications fast -- and we wish to optimize $Z/2$ gates,
+which requires a metric that is sensitive to global phases for the initial
+states $\ket*{0}$ and $\ket*{1}$.
+Additionally, the $R_{k}$ term penalizes the norm of $\mathrm{d}^{2}_{t} a_{k}$.
+Penalizing the norm of $\mathrm{d}^{2}_{t} a_{k}$ and $\mathrm{d}_{t} a_{k}$
+makes $a_{k}$ smooth, which mitigates high-frequency AWG transitions.
+Stated succinctly, the optimization problem takes the form:
+\begin{mini!}[2] 
+  {\substack{\mathbf{x}_1,\ldots,\mathbf{x}_N \\ \mathbf{u}_1,\ldots,\mathbf{u}_{N-1}}}
+  {\sum_{k=1}^N {(\mathbf{x}_k-\mathbf{x}_{T})}^{T} Q_k (\mathbf{x}_k-\mathbf{x}_{T})
+    + \sum_{k=1}^{N-1} {\mathbf{u}_k}^{T} R_k \mathbf{u}_{k}}{}{} \label{eq:costfun}
+  \addConstraint{\mathbf{x}_{k+1}}{= \mathbf{f}(\mathbf{x}_k, \mathbf{u}_k) \ \forall \ k}
+  \label{eq:dyn_con}
+  \addConstraint{\ket*{\psi^{0}_{1}} = \ket*{0}, \ket*{\psi^{1}_{1}} = \ket*{1}}
+  \label{eq:istate_con}
+  \addConstraint{{\textstyle \int_{t}} a_{1} = a_{1} = \mathrm{d}_{t} a_{1} = 0}
+  \label{eq:ic_con}
+  \addConstraint{\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{T}}
+    \ \forall \  i} \label{eq:tstate_con}
+  \addConstraint{{\textstyle \int_{t}} a_{N} = a_{N} = 0} \label{eq:znf_con}
+  \addConstraint{{\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}}
+      \rvert}^{2} = 1 \ \forall \ i, k} \label{eq:statenorm_con}
+  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz} \ \forall \ k.} \label{eq:amp_con}
+\end{mini!}
+
+Next, we remark on our problem formulation.
+We put a cost function at all time steps
+because it benefits the iLQR solving stage \cite{Jackson2020altroc};
+although this may incentivize early achievement of the desired gate,
+as in Ref. \cite{leung2017speedup}, we are primarily concerned
+with achieving the gate at the final time step, which the target-state
+constraint \eqref{eq:tstate_con} ensures.
+Additionally, the target-state constraint
+requires the final state to match the target state, including its global phase,
+up to our chosen maximum constraint violation $\sim 10^{-8}$.
+If we did not impose this constraint, the optimizer would be
+allowed to sacrifice the closed-system gate error to achieve better
+performance on the other cost functions, which is undesirable.
+To enforce a constraint in standard QOC frameworks,
+the prefactor for the constraint function is manually increased
+between separate optimization instances until the constraint is satisfied
+\cite{heeres2017implementing, leung2017speedup, reinhold2019controlling},
+which becomes infeasible for more than one constraint.
+ALM automates these prefactor updates to find
+a solution that satisfies all of the given constraints.
+Hence, ALTRO's ability to handle multiple constraints makes it
+an attractive solver for QOC problems.
+
+In extraordinarily difficult cases of
+QOC, it may be impossible
+to obey the physics of the system and achieve the desired gate \cite{abdelhafez2020universal},
+i.e., the dynamics constraint \eqref{eq:dyn_con}
+and the target-state constraint \eqref{eq:tstate_con} may be mutually unsatisfiable.
+In this case, the prefactors for the constraint function terms
+in the ALM objective will tend to infinity -- leading to numerical instability -- and the
+optimization will not converge. To maintain a constrained approach in this situation,
+the maximum constraint violation for the target-state constraint can be raised
+to a level commensurate with the minimum acceptable gate error.
+
+Finally, for ALTRO's first indirect stage,
+the augmented states are obtained explicitly with the discrete
+dynamics function, so the dynamics constraint and initial conditions
+\eqref{eq:dyn_con}-\eqref{eq:ic_con} are satisfied by construction.
+In this stage, the rest of the constraint functions \eqref{eq:tstate_con}-\eqref{eq:amp_con}
+are added to the objective in their isomorphism-equivalent form \eqref{eq:isomorphism}.
+Conversely, for the second direct stage, all of the constraints
+\eqref{eq:dyn_con}-\eqref{eq:amp_con} are used to define
+the projection onto the constraint manifold, and the objective is unmodified.
+Hence, the quantum states become
+free parameters that are adjusted to satisfy the TDSE.
+Although the final solution's deviation from the TDSE is never
+more than the maximum constraint violation,
+we explicitly integrate the TDSE when reporting gate errors to ensure accuracy.
+Exploring the benefit of direct optimization approaches
+for QOC is an interesting direction for future work.
+
+%% F1
+\begin{figure*}[ht]
+  \begin{subfigure}{.315\textwidth}
+    \includegraphics[width=\linewidth]{f1a.png}
+    \caption{\label{fig:longitudea}}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.23\textwidth}
+    \includegraphics[width=\linewidth]{f1b.png}
+    \caption{\label{fig:longitudeb}}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.4\textwidth}
+    \includegraphics[width=\linewidth]{f1c.png}
+    \caption{\label{fig:longitudec}}
+  \end{subfigure}
+  \caption{
+    (a) Flux pulses for the numerical gates (dark blue)
+    and the analytic gates (light pink).
+    (b) $T_{1}$ interpolation function used in optimization. Circle markers
+    indicate measured $T_{1}$ times. Non-circle markers
+    are plotted at the time-averaged 
+    absolute flux and the time-averaged $T_{1}$ time for each pulse.
+    (c) Cumulative gate errors due to depolarization as a function of the
+    number of gates applied.
+    Cumulative gate errors for the numerical $Z/2$ and $Y/2$ gates
+    are indistinguishable. Inset shows log-scaled cumulative gate errors
+    for small gate counts.
+  }
+  \label{fig:longitude}
+\end{figure*}
+
+\section{Depolarization Mitigation\label{sec:longitude}}
+In this section, we outline a method
+for optimizing the flux to mitigate depolarization.
+For many superconducting circuits, the depolarization time
+$T_{1}$ is independent of the control parameters,
+so the fastest possible gate incurs the least depolarization error
+\cite{schulteherbruggen2011optimal}.
+For the fluxonium, however, $T_{1}$ is strongly dependent on the flux.
+We enable the optimizer to trade longer gate times
+for longer $T_{1}$ times, or shorter $T_{1}$ times for shorter gate times,
+by making the gate time a decision variable.
+Additionally, previous work has modeled the gate error due to depolarization
+by evolving density matrices under a master
+equation \cite{rembold2020introduction, schulteherbruggen2011optimal},
+or evolving a large number of states in a quantum trajectory approach
+\cite{abdelhafez2019gradient}.
+We avoid the increase in computational complexity required for these
+techniques by penalizing the integrated depolarization rate in optimization.
+
+The integrated depolarization rate is given by,
+\begin{equation}
+  D_{1}(t) = \int_{0}^{t} T_{1}^{-1}[a(t^{\prime})] dt^{\prime}.
+\end{equation}
+For the gates we consider here, where the gate time is small compared to $T_{1}$,
+the integrated depolarization rate is proportional to the probability of a depolarization event.
+Additionally, the integrated depolarization rate is a reasonable proxy for the gate error incurred
+because depolarization errors are incoherent -- they increase
+monotonically in time without interference.
+The integrated depolarization rate is appended to the augmented state \eqref{eq:astatecontrols}
+and its norm is penalized in the $Q_{k}$ term of the objective by setting
+the corresponding element of the target augmented state to zero, see \eqref{eq:costfun}.
+$T_{1}$ as a function of the flux is obtained by evaluating
+a spline fit to experimental data, see Fig. \ref{fig:longitudeb}.
+
+Alternatively, modeling the depolarization with a master equation approach would require adding
+density matrices of size $n \times n$
+to the augmented state, and a quantum trajectory approach
+would require adding many states of size $n$ to the augmented state,
+where $n$ is the dimension of the Hilbert space.
+By contrast, the integrated depolarization rate is a single
+real number; thus, the computational complexity of evaluating this
+depolarization model does not scale
+with the dimension of the Hilbert space.
+
+To perform time-optimal control, we make
+the duration between time steps
+a decision variable \cite{howell2019altro}. 
+The square root of the duration $\sqrt{\Delta t_{k}}$
+is appended to the augmented control \eqref{eq:astatecontrols}
+and its square $\lvert \Delta t_{k} \rvert$ is used
+for integration in the discrete dynamics function.
+Although we constrain the bounds of the duration between reasonable positive values to maintain
+numerical stability, the optimizer may assign negative values to the
+duration for intermediate optimization iterations, so this squaring
+approach maintains positivity.
+
+We analyze the effect of depolarization on
+the $X/2$, $Y/2$, and $Z/2$ gates obtained with
+our numerical method and the corresponding analytic gates presented in \cite{zhang2020universal}.
+We use the Lindblad master equation to simulate $T_{1}$ dissipation for successive
+gate applications, and compute the cumulative gate error
+after each application, see Appendix \ref{appendix:longitude}.
+The gate error reported in this text is the infidelity
+of the evolved state and the target state averaged over 1000 pseudo-randomly
+generated initial states.
+
+The flux pulses for the numerical gates
+are approximately periodic
+with amplitudes $\sim 0.2 \textrm{GHz}$, see Fig. \ref{fig:longitudea}.
+They are reminiscent of the analytically determined Floquet operations
+for a fluxonium described in \cite{huang2021engineering}
+and realized in \cite{mundada2020floquet}.
+The numerical gate times are greater
+than the analytic gate times, but the
+numerical flux pulses
+spend more time at larger flux values, achieving higher $T_{1}$ times on average,
+see Fig. \ref{fig:longitudeb}.
+The single-gate errors for both the analytic and numerical gates are
+less than $10^{-4}$, which makes them sufficient for quantum error correction -- a
+prerequisite for fault-tolerant quantum
+computing \cite{aharonov2008fault, knill2005quantum, gottesman1997stabilizer}.
+However, the numerical gates achieve single-gate errors
+$\sim 5$ times less than those for the analytic gates,
+which tracks closely with their relative improvement
+on the integrated depolarization rate metric, see Appendix \ref{appendix:longitude}.
+This advantage in single-gate errors corresponds to a
+significant reduction in error correction resources
+\cite{paetznick2014resource, suchara2013comparing}.
+Furthermore, for successive gate applications,
+the gate error due to depolarization is approximately linear
+in the gate count, which we expect for $t \ll T_{1}$, see Fig. \ref{fig:longitudec}.
+The gate error reduction for large gate counts is
+important for noisy, intermediate-scale quantum (NISQ)
+applications. These improvements are significant given the constraints
+we have imposed on the gates,
+and do not represent a fundamental limit to the optimization
+methods we have employed.
+
+\section{Robustness to Static Parameter Uncertainty \label{sec:static}}
+We have formulated the QOC
+problem as an \emph{open-loop} optimization problem, i.e.,
+we do not incorporate feedback from the experiment into the optimization.
+However, the precise device parameters will deviate from the parameters we use in optimization,
+leading to poor experimental performance. We combat errors
+of this form using robust control techniques,
+making the state evolution insensitive
+to parameter uncertainty. As an example,
+we mitigate errors arising from the drift and finite measurement
+precision of the qubit frequency, which modifies the fluxonium Hamiltonian
+\eqref{eq:hamiltonian} by $f_{q} \rightarrow f_{q} + \delta f_{q}$.
+We consider three robust control techniques to accomplish this task:
+a sampling method, an unscented sampling method,
+and a derivative method.
+
+\subsection{Sampling Method}
+The sampling method incentivizes the optimizer
+to ensure that multiple copies of a state, each evolving
+with a distinct value of the uncertain parameter, achieve
+the same target state. Variants of this technique have been proposed
+in the context of QOC
+\cite{allen2019robust, khaneja2005optimal,
+  reinhold2019controlling, rembold2020introduction}.
+For each initial state,
+we add two sample states $\ket*{\psi^{\pm}}$
+to the augmented state \eqref{eq:astatecontrols}. The discrete dynamics
+function is modified
+so the sample states evolve under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
+with $f_{q} \rightarrow f_{q} \pm \sigma_{f_{q}}$ for a fixed
+ standard deviation $\sigma_{f_{q}}$ of the qubit frequency, acting as a hyperparameter.
+We penalize the infidelities of the sample states with respect to the target state
+by adding a cost function to the objective of the form
+$\sum_{k, \pm} b_{k} (1 - {\lvert \braket*{\psi_{T}}{\psi^{\pm}_{k}} \rvert}^{2})$
+where $b_{k}$ is a constant we supply.
+For this method, the standard orthonormal basis states are an insufficient choice
+for the initial states. As an example, a $Z/2$ gate achieved by idling
+at the flux frustration point ($a_{k} = 0 \ \forall \ k$)
+will be robust to qubit frequency detunings for the initial states $\ket{0}$
+or $\ket{1}$ because the infidelity metric is insensitive to global phases,
+but this gate will not be robust for any other initial states.
+Therefore, we choose the four initial states $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
+(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized},
+whose outer products span the operators on the Hilbert space,
+and  we refer to them as the operator basis.
+
+\subsection{Unscented Sampling Method}
+Whereas the sampling method penalizes the deviations of the sample states
+from the target state, the unscented sampling method
+penalizes the deviations of the sample states from the nominal state
+\cite{howell2020direct, lee2013sigma,
+  thangavel2020robust}. Accordingly, the cost function we add
+to the objective takes the form
+$\sum_{k, j} c_{k} (\psi^{j}_{k} - \psi_{k})^{T}
+(\psi^{j}_{k} - \psi_{k})$, where $c_{k}$ is a
+constant we supply, $\psi_{k}$ is
+the evolved initial state (nominal state), and $\psi^{j}_{k}$ is a sample state
+that evolves under a modified Hamiltonian similar to that in the sampling method.
+The sample states are chosen to encode a unimodal distribution over
+the $2n$ elements of the nominal state, modeling the uncertainty in the state
+as a result of the uncertainty in the parameter. We use the unscented transform
+\cite{julier2004unscented, uhlmann1995dynamic}
+to accurately propagate the mean and covariance of this distribution between
+time steps, or equivalently, through the transformation of the TDSE \eqref{eq:tdse}.
+Unlike the sampling method, the cost function for the unscented sampling method
+is sensitive to global phases. Accordingly, we
+do not observe a performance increase when
+using more than one initial state. 
+A detailed procedure for the unscented transformation is given
+in Appendix \ref{appendix:unscented}.
+
+\subsection{Derivative Method}
+The derivative method penalizes the sensitivity of the state
+to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
+state derivative $\ket*{\partial_{f_{q}}^{l} \psi} \equiv \partial_{f_{q}}^{l} \ket*{\psi}$.
+In the $m$\textsuperscript{th}-order
+derivative method, we append all state derivatives of order $1, \dots, m$
+to the augmented state \eqref{eq:astatecontrols}
+for each initial state.
+We obtain the state derivatives at each time step by performing forward-mode
+differentiation on the TDSE \eqref{eq:tdse}.
+For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
+\begin{align}
+  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi},\\
+  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
+  H \ket{\partial_{f_{q}} \psi} +
+  (\partial_{f_{q}} H) \ket{\psi}.
+  \label{eq:d1dyn}
+\end{align}
+We integrate the coupled ODEs with exponential
+integrators, see Appendix \ref{appendix:derivative}.
+While the state $\ket*{\psi}$ has unit norm,
+the state derivatives $\ket*{\partial^{l}_{f_{q}} \psi}$ need not, as is evident
+from the non-unitary dynamics \eqref{eq:d1dyn}.
+We penalize the norms of the isomorphism-equivalent state derivatives
+in the $Q_{k}$ term of the objective by setting the corresponding elements
+of the target augmented state to zero, see \eqref{eq:costfun}. Intuitively, this corresponds to penalizing
+the sensitivity of each state element to the uncertain parameter. As was the case for
+the unscented sampling
+method, we do not observe a performance increase when using more than one initial state
+for the derivative method.
+We present the runtimes of our implementations of the three robust control methods
+in Appendix \ref{appendix:time}.
+
+%% F2
+\begin{figure*}[ht]
+  \begin{subfigure}{.315\textwidth}
+    \includegraphics[width=\linewidth]{f2a.png}
+    \caption{\label{fig:statica}}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.4\textwidth}
+    \includegraphics[width=\linewidth]{f2b.png}
+    \caption{\label{fig:staticb}}
+  \end{subfigure}\hfill
+  \begin{subfigure}{.23\textwidth}
+    \includegraphics[width=\linewidth]{f2c.png}
+    \caption{\label{fig:staticc}}
+  \end{subfigure}
+  \caption{
+    (a) Flux pulses for $Z/2$ gates robust to qubit frequency detunings constructed with the
+    analytic (A), sampling (S), unscented sampling (U), and the 1\textsuperscript{st}-
+    and 2\textsuperscript{nd}-order derivative methods (D1, D2). The flux pulses shown
+    for the sampling, unscented sampling, and derivative methods are optimized
+    for twice the gate time of the analytic gate.
+    (b) Single-gate error at a one-percent qubit frequency detuning as
+    a function of the gate time. Missing
+    data points represent gates with a gate error greater than $5 \cdot 10^{-5}$.
+    (c) Single-gate error as a function of the qubit frequency detuning.
+    The gate errors for the analytic and 1\textsuperscript{st}-order derivative
+    methods are shown for gate times which are multiples of $1 / 4 f_{q} \sim 18 \textrm{ns}$.
+    The gate errors for the two methods are
+    indistinguishable at the gate time $18 \textrm{ns}$.
+  }
+  \label{fig:static}
+\end{figure*}
+
+\subsection{Comparison}
+We examine the gate errors due to a static qubit frequency
+detuning for the $Z/2$ gates obtained with the robust control techniques
+and the analytic $Z/2$ gate.
+To compute the gate error,
+an initial state is evolved
+under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
+two separate times with the transformations
+$f_{q} \rightarrow f_{q} \pm \delta f_{q}$
+at the stated qubit frequency detuning $\delta f_{q}$.
+The reported gate error is the infidelity of
+the evolved state and the target state averaged over
+the two transformations for each of $1000$ pseudorandomly
+generated initial states.
+We set $\sigma_{f_{q}}/f_{q} = 1\%$
+for the sampling and unscented sampling
+methods.
+
+The analytic gate corresponds to
+idling at the flux frustration point $a_{k} = 0 \ \forall \ k$, see Fig.
+\ref{fig:statica}. Its gate time $1 / 4 f_{q} \sim 18\textrm{ns}$
+is the shortest possible for a $Z/2$ gate on the device.
+The gate's erroneous rotation angle
+$2 \pi \delta f_{q} / 4 f_{q}$ is linear in the
+qubit frequency detuning, resulting in a gate error that is quadratic
+in the detuning.
+At a one-percent detuning $\lvert \delta f_{q} / f_{q} \rvert = 1\%$,
+the gate error is $\sim 4.7 \cdot 10^{-5}$,
+which is sufficient for quantum error correction.
+
+For the sampling method, the gate error at a one-percent qubit frequency detuning
+does not decrease substantially over the
+range of gate times, and begins to increase above $5 \cdot 10^{-5}$ for gate times
+greater than $\sim 50$ns, see Fig. \ref{fig:staticb}.
+Optimization results for the sampling method reveal that it is typically
+able to achieve a high fidelity for
+one sample $\ket{\psi^{\pm}}$,
+but not the other $\ket{\psi^{\mp}}$, indicating that it is difficult for the optimizer
+to make progress on both objectives.
+For the unscented sampling method,
+the gate error at a one-percent detuning
+does not decrease substantially 
+over the gate times, but it does reach
+a minimum of $\sim 3.9 \cdot 10^{-5}$
+near fractions of the Larmor period: $2/4f_{q} \sim 36\textrm{ns}$,
+$3/4f_{q} \sim 54\textrm{ns}$, and $4/4f_{q} \sim 72\textrm{ns}$.
+
+The two derivative methods converge on qualitatively similar flux pulses that
+idle near the flux frustration point and use fast triangle movements at the boundaries,
+similar to the flux pulse produced by the unscented sampling method.
+For both derivative methods, the gate error at a one-percent qubit frequency detuning
+decreases super-linearly in the gate time.
+For the 1\textsuperscript{st}-order method, the gate error at a one-percent detuning 
+reaches $10^{-7}$ at the Larmor period $1 / f_{q} \sim 72$ns,
+see Fig. \ref{fig:staticc}.
+This result mimics the
+ability of composite pulses to mitigate parameter uncertainty errors to arbitrary
+order with sufficiently many pulses \cite{merrill2014progress}.
+It is difficult to choose an appropriate composite pulse
+for the problem studied here due to our Hamiltonian and experimental constraints.
+A comparison between composite pulses and numerical techniques could
+be an interesting topic for future study.
+
+Furthermore, the ability to perform
+$Z$-type gates in any given time is critical
+for synchronizing phases in multi-qubit experiments,
+where the qubits have distinct
+frequencies. Notably, the analytic gate studied here cannot be extended
+to gate times other than $1/4f_{q}$. 
+We can find gates using the numerical methods at
+all gate times at and above $1/4f_{q}$, see Fig. \ref{fig:staticb}.
+These numerical methods offer an effective scheme for synchronizing
+multi-qubit experiments.
+
+%% F3
+\begin{figure*}[ht]
+  \begin{subfigure}{.4\textwidth}
+    \includegraphics[width=\linewidth]{f3a.png}
+    \caption{}
+    \label{fig:stochastica}
+  \end{subfigure}\hspace{0.05\textwidth}
+  \begin{subfigure}{.4\textwidth}
+    \includegraphics[width=\linewidth]{f3b.png}
+        \caption{}
+    \label{fig:stochasticb}
+  \end{subfigure}
+  \caption{
+    (a) Flux pulses for $X/2$ gates robust to flux noise
+    constructed with the analytic (A),
+    sampling (S), unscented sampling (U), and the 1\textsuperscript{st}-
+    and 2\textsuperscript{nd}-order derivative methods (D1, D2).
+    (b) Cumulative gate error due to 1/$f$ flux noise for
+    successive gate applications. The cumulative gate errors for the
+    sampling, unscented sampling, and the derivative methods are indistinguishable.
+  }
+  \label{fig:stochastic}
+\end{figure*}
+
+\section{Robustness to Time-Dependent Parameter Uncertainty \label{sec:stochastic}}
+An additional source of experimental error arises from time-dependent
+parameter uncertainty. For many flux-biased and inductively-coupled
+superconducting circuit elements, magnetic flux noise is the dominant
+source of coherent errors \cite{bialczak20071f, kakuyanagi2007dephasing,
+  kumar2016origin, yoshihara2006decoherence}. Flux noise
+modifies the fluxonium Hamiltonian \eqref{eq:hamiltonian}
+by $a(t) \rightarrow a(t) + \delta a(t)$ where $\delta a(t)$ is the flux noise.
+The spectral density of flux noise is observed to
+follow a 1/$f$ distribution
+\cite{bialczak20071f, koch2007model, kakuyanagi2007dephasing,
+  kumar2016origin, yoshihara2006decoherence, yoshihara2010correlated,
+  zhang2020universal},
+so the noise is dominated by low-frequency components.
+The analytic gate considered here
+takes advantage of the low-frequency characteristic and
+treats the noise as quasi-static, performing a generalization of the spin-echo
+technique to compensate for erroneous drift \cite{hahn1952spin, meiboom1958modified}.
+
+We modify the robust control techniques presented
+in the previous section to combat 1/$f$ flux noise.
+The unscented sampling method is modified so that the sample states
+are subject to 1/$f$ flux noise. The noise
+is generated by filtering white noise sampled from a standard
+normal distribution with a finite impulse response filter \cite{saspweb2011}.
+The noise is then scaled by the 
+flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
+\sigma_{a} = 2.5 \cdot 10^{-5} \textrm{GHz}$.
+In principle, we could modify the sampling method
+similarly; however, we choose to subject the sample states
+to static noise
+$a(t) \rightarrow a(t) \pm \sigma_{a}$
+for comparison. The derivative methods require no algorithmic modification
+from the static case, but the TDSE is now differentiated with respect
+to $a(t)$ instead of $f_{q}$ as in \eqref{eq:d1dyn}.
+
+We analyze the gate errors due to 1/$f$ flux noise for
+the $X/2$ gates constructed with the robust control techniques
+and the analytic $X/2$ gate. To compute the gate error,
+we evolve an initial state
+under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
+where the optimized flux is modified $a(t) \rightarrow a(t) + \delta a(t)$.
+We generate the flux noise as
+we described for the unscented sampling method.
+The reported gate error is the infidelity
+averaged over $1000$ pseudorandomly generated initial states,
+each of which is subject to a distinct pseudorandomly
+generated flux noise instance.
+To observe the effect of interfering coherent errors,
+we simulate successive applications of the gate constructed by each method;
+we compute the cumulative gate error
+after each application, see Fig. \ref{fig:stochastic}.
+Both the analytic
+and numerical gates yield single-gate errors
+sufficient for quantum error correction.
+Despite converging on qualitatively different solutions, the
+numerical gates perform similarly in the concatenated
+gate application comparison. Their gate errors
+after $200$ gate applications $\sim 11 \mu\textrm{s}$ are
+two orders of magnitude less than the gate error produced by the analytic gate.
+1/$f$ flux noise is a significant source of coherent errors in NISQ applications,
+and these numerical techniques offer effective avenues to mitigate it.
+
+\section{Conclusion}
+We have introduced state-of-the-art trajectory optimization
+techniques in the context of quantum optimal control, enabling
+us to achieve tight tolerances for multiple constraints on the
+control fields and quantum states. Using these capabilities,
+we have mitigated decoherence and
+achieved robustness to parameter uncertainty
+errors on a superconducting fluxonium qubit.
+We have proposed a scheme for suppressing
+depolarization with time-optimal
+control and the integrated depolarization rate model.
+The computational complexity of evaluating this model is
+independent of the dimension of the Hilbert space, enabling
+inexpensive optimization on high-dimensional quantum systems.
+We have also proposed the derivative method for robust control which achieves
+superlinear gate error reductions in the gate time for the static parameter
+uncertainty problem we studied.
+We have shown that the derivative, sampling, and unscented sampling methods
+can mitigate 1/$f$ flux noise errors -- which
+dominate coherent errors for flux controlled qubits.
+These robust control techniques can be applied
+to any Hamiltonian,
+allowing experimentalists in all domains to engineer robust
+operations on their quantum systems.
+Furthermore, they can be used to achieve the low gate errors
+required for fault-tolerant quantum computing applications. Our
+implementations of the techniques described in this work are available
+at \url{https://github.com/SchusterLab/rbqoc}.
+
+\begin{acknowledgments}
+  We thank Helin Zhang for experimental assistance
+  and Taylor Howell, Tanay Roy, Colm Ryan, and Daniel Weiss for useful discussions.
+  This work was made possible by many open source software projects,
+  including but not limited to:
+  DifferentialEquations.jl \cite{rackauckas2017differentialequations},
+  Distributions.jl \cite{besancon2019distributions},
+  ForwardDiff.jl \cite{revelsLubinPapamarkou2016},
+  Matplotlib \cite{hunter2007matplotlib},
+  NumPy \cite{harris2020array},
+  TrajectoryOptimization.jl \cite{howell2019altro},
+  and Zygote.jl \cite{innes2018don}.
+  This work is funded in part by EPiQC, an NSF Expedition in Computing, under grant CCF-1730449.
+  This work was supported by the Army Research Office under Grant No. W911NF1910016.
+\end{acknowledgments}
 
 \appendix
-\input{a1experiment.tex}
-\input{a2dissipation.tex}
-\input{a3derivative.tex}
-\input{a4ctensor.tex}
+\section{DEPOLARIZATION \label{appendix:longitude}}
+We comment on the depolarization metrics and then give
+our procedure for integrating the Lindblad master equation.
+The integrated depolarization rate and the gate error due to
+depolarization
+are compared in Table \ref{tab:longitude} for the
+numerical experiment described in Sec. \ref{sec:longitude}.
+The ratio of the value obtained on the metric with the analytic technique
+to the value obtained with the numerical technique
+is similar across the two metrics.
+
+\begin{table}[ht]
+  \begin{tabular}{| c | c | c | c | c | c | c | c | c |}
+    \hline
+    \multirow{2}{*}{Gate} &
+    $D_{1\textrm{A}}$ &
+    $D_{1\textrm{N}}$ &
+    \multirow{2}{*}{$\dfrac{ D_{1\textrm{A}} }{ D_{1\textrm{N}} }$} &
+    $\textrm{GE}_{\textrm{A}}$ &
+    $\textrm{GE}_{\textrm{N}}$ &
+    \multirow{2}{*}{$\dfrac{ \textrm{GE}_{\textrm{A}} }{ \textrm{GE}_{\textrm{N}} }$}\\
+    &
+    $(10^{-5})$ &
+    $(10^{-5})$ &
+    &
+    $(10^{-5})$ &
+    $(10^{-5})$ &\\
+    \hline
+    Z/2 & 5.745 & 1.149 & 5.000 & 0.888   & 0.185  & 4.791\\
+    \hline
+    Y/2 & 5.253 & 1.157 & 4.540 & 0.770 & 0.186   & 4.132\\
+    \hline
+    X/2 & 16.251 & 2.660 & 6.109 & 2.674 & 0.432  & 6.200\\
+    \hline
+  \end{tabular}
+  \caption{
+    Single-gate integrated depolarization rate ($D_{1}$)
+    and single-gate error due to depolarization (GE).
+    Values are reported for the analytic (A) and numerical (N) gates.
+  }
+  \label{tab:longitude}
+\end{table}
+
+We employ the Lindblad master equation
+to compute the gate error due to depolarization.
+This equation takes the form:
+\begin{equation}
+  \frac{d}{dt} \rho = -\frac{i}{\hbar} [H, \rho]
+  + \sum_{i} \gamma_{i} (L_{i} \rho L_{i}^{\dagger}
+  - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\}),
+\end{equation}
+For depolarization, $\gamma_{\pm} = T_{\pm}^{-1}$,
+$L_{\pm} = \sigma^{\pm} \equiv (\sigma_{x} \pm i \sigma_{y})/2$.
+Our device operates in the regime where $hf \ll k_{\textrm{B}}T$ such that
+$T_{+} = T_{-} = 2 T_{1}$, where $T_{1}$ is obtained at each time step
+from the spline shown in Fig. \ref{fig:longitudeb}.
+We obtain the $T_{1}$ values in this spline
+by driving the qubit at the desired flux bias
+and monitoring the resultant decay. For more details
+on these measurements, consult Ref. \cite{zhang2020universal}.
+Because $T_{1}$ depends on the flux, so do
+the decay rates $\gamma_{\pm}$.
+Integrating the master equation with time-dependent decay rates
+provides a heuristic for how gates might  perform in
+the experiment. 
+This procedure may not be strictly correct
+when decay rates change significantly on the time scale of the relaxation time,
+which is the regime we are operating in. Standard
+derivations of the Lindblad master equation do not account for
+time-dependent decay rates \cite{manzano2020a}. A more thorough
+treatment of this regime in future work would unlock new insights for
+quantum computing platforms where decoherence is strongly
+dependent on the control parameters.
+
+In order to use exponential integrators, we employ
+the vector (Choi-Jamiolkowski) isomorphism \cite{Landi2018},
+\begin{equation}
+  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}}\, \textrm{vec}({\rho})
+  \label{eq:veclindblad},
+\end{equation}
+\begin{equation}
+  \begin{aligned}
+    \hat{\mathcal{L}} &= -i(\openone \otimes H - H^{T} \otimes \openone)\\
+    &+ \sum_{i} \gamma_{i}
+    (L_{i}^{*} \otimes L_{i} - \frac{1}{2} (\openone \otimes L_{i}^{\dagger}L_{i}
+    - L_{i}^{T}L_{i}^{*} \otimes \openone)),
+  \end{aligned}
+\end{equation}
+where $\rho = \sum_{i, j} \alpha_{ij} \ket{i}\bra{j}$
+and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{ij} \ket{i} \otimes \ket{j}$.
+Because the flux is constant between time steps
+due to our numerical discretization,
+the Hamiltonian and decay rates are also constant
+between time steps.
+Therefore, the exact solution to \eqref{eq:veclindblad} is,
+\begin{equation}
+  \textrm{vec}(\rho_{k + 1}) = {\exp}{\textstyle(}\Delta t_{k}
+  \hat{\mathcal{L}}_{k}{\textstyle)} \textrm{vec}(\rho_{k}).
+\end{equation}
+The vector isomorphism transforms $(n \times n) \times (n \times n)$
+matrix-matrix multiplications to $(n^{2} \times n^{2}) \times n^{2}$ matrix-vector
+multiplications. For small $n$, we find that it is
+faster to use an exponential integrator on the vectorized equation than to perform
+Runge-Kutta on the unvectorized equation.
+The latter requires decreasing the interval $\Delta t_{k}$
+to maintain accuracy, resulting in more time steps.
+
+\section{UNSCENTED SAMPLING METHOD}
+\label{appendix:unscented}
+In this section, we outline the full unscented sampling procedure.
+We consider a state $\psi \in \mathbb{R}^{2n}$, 
+an uncertain set of parameters $\lambda\in \mathbb{R}^{d}$, and discrete dynamics
+$\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
+The nominal initial state is given by $\bar{\psi}_{1}$ with an associated
+covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$
+which describes the uncertainty in the initial state.
+We use the notation $\mathbb{S}_{++}^{m}$ to denote the
+set of real, symmetric, and positive-definite $m \times m$ matrices.
+By the positive-definite requirement, $P_{1}$ must be non-zero even if the state-preparation error
+is negligible.
+The uncertain parameter has zero-mean
+and its distribution is given by the covariance matrix
+$L_{k} \in \mathbb{S}_{++}^{d}$ at time step $k$. The zero-mean assumption
+is convenient for deriving the update procedure. A non-zero mean can be encoded
+in the discrete dynamics function $f(\psi_{k}, \lambda_{k})$.
+
+The initial $4n + 2d$ sample states and initial $4n + 2d$
+uncertain parameters are sampled from the initial distributions,
+\begin{equation}\label{eq:uupdate}
+  \begin{pmatrix} \psi_{1}^{j} \\ \lambda_{1}^{j} \end{pmatrix} =
+  \begin{pmatrix} \bar{\psi}_{1} \\ 0\end{pmatrix}
+    \pm \beta \sqrt{\begin{pmatrix} P_{1} & 0\\ 0 & L_{1}\end{pmatrix}}^{\; j}.
+\end{equation}
+Here, $\beta$ is a hyperparameter that controls the spacing of the covariance contour.
+The $(\pm)$ is understood to take $(+)$ for $j \in \{1, \dots, 2n + d\}$ and $(-)$ for
+$j \in \{2n + d + 1, \dots, 4n + 2d\}$. We use the Cholesky factorization
+to compute the square root of the
+joint covariance matrix, though other methods
+such as the principal square root may be employed.
+The superscript on the matrix square root indicates the $j$\textsuperscript{th}
+column (mod $2n + d$) of the lower triangular Cholesky factor.
+Then, the sample states are normalized,
+\begin{equation}\label{eq:unormalize}
+  \psi_{1}^{j} \rightarrow \frac{\psi_{1}^{j}}{\sqrt{{\psi_{1}^{j}}^{T} \psi_{1}^{j}}}.
+\end{equation}
+The sample states are propagated to the next time step,
+\begin{equation}\label{eq:upropagate}
+  \psi^{j}_{2} = f(\psi^{j}_{1}, \lambda^{j}_{1}).
+\end{equation}
+The mean and covariance of the sample states are computed,
+\begin{align}
+  \bar{\psi}_{2} &= \frac{1}{4n + 2d} \sum_{j = 1}^{4n + 2d} \psi_{2}^{j},\\
+  P_{2} &= \frac{1}{2 \beta^{2}} \sum_{j = 1}^{4n + 2d}
+  (\psi^{j}_{2} - \bar{\psi}_{2})(\psi^{j}_{2} - \bar{\psi}_{2})^{T}.
+\end{align}
+The sample states are then resampled and propagated to the next time step using
+\eqref{eq:uupdate}, \eqref{eq:unormalize}, and \eqref{eq:upropagate}. Our
+choice of sample states (sigma points) follows equation 11 of Ref.
+\cite{julier2004unscented}.
+Prescriptions that require fewer sigma points exist \cite{julier2002reduced}.
+
+\section{DERIVATIVE METHOD \label{appendix:derivative}}
+Here, we outline how to efficiently integrate the dynamics
+for the derivative method using exponential integrators.
+General exponential integrators break the dynamics into a linear term and a non-linear term.
+For example, the dynamics for the first state derivative
+are,
+\begin{equation}
+\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
+-\frac{i}{\hbar} H  \ket{\partial_{\lambda} \psi} -
+\frac{i}{\hbar} (\partial_{\lambda} H) \ket{\psi}. \label{eq:gd1dyn}
+\end{equation}
+The linear term is $L = -\frac{i}{\hbar} H$ and the non-linear
+term is $N = -\frac{i}{\hbar} (\partial_{\lambda} H) \ket{\psi}$.
+With piecewise-constant controls, the exact solution to \eqref{eq:gd1dyn} is,
+\begin{equation}
+  \label{eq:dgeneralexp}
+  \begin{aligned}
+    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(\Delta t_{k} L_{k})
+    \ket{\partial_{\lambda} \psi_{k}}\\
+    &+ \int_{0}^{\Delta t_{k}} \exp((\Delta t_{k} - t^{'})L_{k})
+    N(t_{k} + t^{'}) dt^{'}.\\
+  \end{aligned}
+\end{equation}
+General exponential integrators proceed by breaking the integral in \eqref{eq:dgeneralexp}
+into a discrete sum, similar to the procedure
+for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler
+method \cite{berland2006solving},
+\begin{equation}
+  \begin{aligned}
+    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(\Delta t_{k}L_{k})
+    \ket{\partial_{\lambda} \psi_{k}}\\
+    &+ \exp(\Delta t_{k}L_{k}) N_{k} \Delta t_{k}.\\
+  \end{aligned}
+\end{equation}
+This method provides a good tradeoff between accuracy and efficiency, requiring one unique matrix
+exponential computation per stage. Integration accuracy for the state derivatives
+is not of the utmost importance because they are used in the robustness cost function -- as
+opposed to the states themselves which are experimental parameters that must be realized
+with high accuracy.
+
+\section{COMPUTATIONAL PERFORMANCE \label{appendix:time}}
+In this section we provide runtimes for our optimizations. The runtimes for
+the base optimization in Sec. \ref{sec:fluxonium},
+the depolarization optimization in Sec. \ref{sec:longitude},
+and the robust optimizations in Sec. \ref{sec:static}
+are presented in Table \ref{tab:time}
+for a $Z/2$ gate at gate times which are multiples of $1/4f_{q} \sim 18$ns.
+We performed optimizations on a single core of
+an AMD Ryzen Threadripper 3970X 32-Core Processor @ 3.7 GHz.
+Future work will parallelize the robustness methods using GPUs
+\cite{leung2017speedup},
+which will enable fast optimizations on high-dimensional Hilbert spaces.
+
+\begin{table}[H]
+  \centering
+  \begin{tabular} {| c | c | c | c | c |}
+    \hline
+    \backslashbox{Method}{$t_{N} (\textrm{ns})$} & $18$ & $36$ & $72$\\
+    \hline
+    Base & $0.155 \pm 0.008$ & $7.0 \pm 0.4$ & $15.9 \pm 0.8$\\
+    \hline
+    Depol. & $1.69 \pm 0.08$ & - & -\\
+    \hline
+    S & $1.77 \pm 0.09$ & $48 \pm 2$ & $280 \pm 10$\\
+    \hline
+    U & $75 \pm 4$ & $340 \pm 20$ & $400 \pm 20$\\
+    \hline
+    D1 & $6.1 \pm 0.3$ & $27 \pm 1$ & $65 \pm 3$\\
+    \hline
+    D2 & $15.7 \pm 0.8$ & $17.3 \pm 0.9$ & $54 \pm 3$\\
+    \hline
+  \end{tabular}
+  \caption{
+    Average runtimes in seconds for $Z/2$ optimizations
+    using the base, depolarization, sampling (S),
+    unscented sampling (U), and the 1\textsuperscript{st}-
+    and 2\textsuperscript{nd}-order derivative methods (D1, D2).
+  }
+  \label{tab:time}
+\end{table}
 

### 2021-05-17 paper/main.tex
added blais reset reference

@@
 and applications now include superconducting circuits \cite{abdelhafez2020universal,
-  chakram2020multimode, egger2013optimized, fisher2010optimal, gokhale2019partial,
+  boutin2017resonator, chakram2020multimode, egger2013optimized,
+  fisher2010optimal, gokhale2019partial,
   huang2014optimal, heeres2017implementing, kelly2014optimal, leng2019robust,
@@
   treutlein2006microwave, van2016optimal},
-nitrogen-vacancy centers in diamond \cite{chou2015optimal,
+vacancy centers in diamond \cite{chou2015optimal,
   dolde2014high, geng2016experimental,

### 2021-05-17 src/twospin/twospin.jl
two spin updates

@@ -4,12 +4,46 @@ twospin.jl - common definitions for the twospin directory derived from spin.jl
 
 WDIR = joinpath(@__DIR__, "../../")
 include(joinpath(WDIR, "src", "spin", "spin.jl"))
+using CSV
+using DataFrames
 
+
+chi_minus_df = CSV.read(joinpath(WDIR, "src/twospin/chi_minus_values.csv"), DataFrame, header=false)
+gs_minus_expect_df = CSV.read(joinpath(WDIR, "src/twospin/gs_minus_expect_values.csv"), DataFrame, header=false)
+J_eff_vals_df = CSV.read(joinpath(WDIR, "src/twospin/J_eff_values.csv"), DataFrame, header=false)
+x_vals_df = CSV.read(joinpath(WDIR, "src/twospin/flux_vals.csv"), DataFrame, header=false)
+chi_minus_vals = convert(Matrix{Float64}, chi_minus_df)[:, 1]
+gs_minus_expect_vals = convert(Matrix{Float64}, gs_minus_expect_df)[:, 1]
+J_eff_vals = convert(Matrix{Float64}, J_eff_vals_df)[:, 1]
+x_vals = convert(Matrix{Float64}, x_vals_df)[:, 1]
+
+J_eff_spline_itp = extrapolate(interpolate((x_vals,), J_eff_vals,
+                                Gridded(Linear())), Periodic())
+J_eff_spline_dkx = Spline1D(x_vals[:, 1], J_eff_vals[:, 1], periodic=true)
+chi_minus_spline_itp = extrapolate(interpolate((x_vals,), chi_minus_vals,
+                                   Gridded(Linear())), Periodic())
+gs_minus_spline_itp = extrapolate(interpolate((x_vals,), gs_minus_expect_vals,
+                                  Gridded(Linear())), Periodic())
+
+
+function J_eff_func(x)
[…]

### 2021-05-17 src/twospin/twospin12.jl

@@ -1,5 +1,5 @@
 """
-twospin11.jl - vanilla two spin based on spin13.jl
+twospin12.jl - vanilla two spin based on spin13.jl
 """
 
 WDIR = get(ENV, "ROBUST_QOC_PATH", "../../")
@@ -16,11 +16,11 @@ const TO = TrajectoryOptimization
 
 # paths
 const EXPERIMENT_META = "twospin"
-const EXPERIMENT_NAME = "twospin11"
+const EXPERIMENT_NAME = "twospin12"
 const SAVE_PATH = abspath(joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME))
 
 # problem
-const CONTROL_COUNT = 3
+const CONTROL_COUNT = 1
 const STATE_COUNT = 4
 const ASTATE_SIZE = STATE_COUNT * HDIM_TWOSPIN_ISO + 3 * CONTROL_COUNT
 const ACONTROL_SIZE = CONTROL_COUNT
@@ -46,13 +46,7 @@ end
 function RD.discrete_dynamics(::Type{RK3}, model::Model, astate::SVector,
                               acontrol::SVector, time::Real, dt::Real)
     h_prop = exp(dt * (NEGI_H0_TWOSPIN_ISO
-                       + XI_coeff(astate[CONTROLS_IDX[1]],
-                                  astate[CONTROLS_IDX[2]],
-                                  astate[CONTROLS_IDX[3]]) * NEGI_H1_TWOSPIN_ISO_1
-                       + IX_coeff(astate[CONTROLS_IDX[1]],
-                                  astate[CONTROLS_IDX[2]],
-                                  astate[CONTROLS_IDX[3]]) * NEGI_H1_TWOSPIN_ISO_2
-                       + XX_coeff * J_eff(astate[CONTROLS_IDX[3]]) * NEGI_H1_TWOSPIN_ISO_3))
+     + J_eff(astate[CONTROLS_IDX[1]]) * NEGI_H1_TWOSPIN_ISO_3))
     state1 = h_prop * astate[STATE1_IDX]
     state2 = h_prop * astate[STATE2_IDX]
     state3 = h_prop * astate[STATE3_IDX]
@@ -69,7 +63,7 @@ function RD.discrete_dynamics(::Type{RK3}, model::Model, astate::SVector,
 end
 
 # main
[…]

### 2021-05-17 src/twospin/twospin13.jl

@@ -1,5 +1,6 @@
 """
-twospin11.jl - vanilla two spin based on spin13.jl
+twospin12.jl - vanilla two spin based on spin13.jl
+want to directly control the magnitude of the XX term
 """
 
 WDIR = get(ENV, "ROBUST_QOC_PATH", "../../")
@@ -16,11 +17,11 @@ const TO = TrajectoryOptimization
 
 # paths
 const EXPERIMENT_META = "twospin"
-const EXPERIMENT_NAME = "twospin11"
+const EXPERIMENT_NAME = "twospin13"
 const SAVE_PATH = abspath(joinpath(WDIR, "out", EXPERIMENT_META, EXPERIMENT_NAME))
 
 # problem
-const CONTROL_COUNT = 3
+const CONTROL_COUNT = 1
 const STATE_COUNT = 4
 const ASTATE_SIZE = STATE_COUNT * HDIM_TWOSPIN_ISO + 3 * CONTROL_COUNT
 const ACONTROL_SIZE = CONTROL_COUNT
@@ -36,6 +37,9 @@ const DCONTROLS_IDX = CONTROLS_IDX[end] + 1:CONTROLS_IDX[end] + CONTROL_COUNT
 # control indices
 const D2CONTROLS_IDX = 1:CONTROL_COUNT
 
+const MIN_J_NORM = -0.03
+const MAX_J_NORM = 0.04
+
 # model
 struct Model <: AbstractModel
 end
@@ -46,13 +50,7 @@ end
 function RD.discrete_dynamics(::Type{RK3}, model::Model, astate::SVector,
                               acontrol::SVector, time::Real, dt::Real)
     h_prop = exp(dt * (NEGI_H0_TWOSPIN_ISO
-                       + XI_coeff(astate[CONTROLS_IDX[1]],
-                                  astate[CONTROLS_IDX[2]],
-                                  astate[CONTROLS_IDX[3]]) * NEGI_H1_TWOSPIN_ISO_1
-                       + IX_coeff(astate[CONTROLS_IDX[1]],
-                                  astate[CONTROLS_IDX[2]],
-                                  astate[CONTROLS_IDX[3]]) * NEGI_H1_TWOSPIN_ISO_2
-                       + XX_coeff * J_eff(astate[CONTROLS_IDX[3]]) * NEGI_H1_TWOSPIN_ISO_3))
[…]

### 2021-08-25 paper/main.tex
paper edits

@@
 
-%% SECTION 1
+%%S1
 \section{Introduction}
@@
 
+%% S2
 \section{Background \label{sec:background}}
@@
 
+%% S3
 \section{QOC for the Fluxonium \label{sec:fluxonium}}
@@
 
+%% S4
 \section{Depolarization Mitigation\label{sec:longitude}}
@@
 
+%% S5
 \section{Robustness to Static Parameter Uncertainty \label{sec:static}}
@@
 
+%% S6
 \section{Robustness to Time-Dependent Parameter Uncertainty \label{sec:stochastic}}
@@
 superconducting circuit elements, magnetic flux noise is the dominant
-source of coherent errors \cite{bialczak20071f, kakuyanagi2007dephasing,
+source of coherent errors, and therefore pure dephasing \cite{bialczak20071f, kakuyanagi2007dephasing,
   kumar2016origin, yoshihara2006decoherence}. Flux noise
@@
 
+For qubit platforms where pure dephasing is caused by high-frequency noise,
+i.e., noise that has a non-negligible spectral weight at frequencies which are greater
+than the sampling rate of the control pulses, it is unclear that pulse-shaping alone
+can ameliorate pure dephasing. In such cases, applying the time-optimal control
+scheme we have developed in Section \ref{sec:longitude} to decrease the gate duration may be an effective
+strategy to mitigate pure dephasing.
+
+%% S7
 \section{Conclusion}

### 2021-11-02 paper/main.tex
added qsde ref

@@
 \cite{ball2020software, carvalho2020error, allen2019robust, reinhold2019controlling,
-  rembold2020introduction, kosut2013robust, niu2019universal, khaneja2005optimal}.
+  rembold2020introduction, kosut2013robust, niu2019universal, khaneja2005optimal,
+petersen2013robustness}.
 
